### 2020-07-16 cases_and_policies/rmd/counterfactuals.Rmd
Version submitted in early July 2020.
Compared to submitted version the following code changes have been made:

- Removed safegraph and unacast data (they can't be publicly shared)
- Minor update to policy data loading due to data format change
- Added explicit subsetting of data to June 2, 2020 and earlier (to match what we used in submitted paper).

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Counterfactuals"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->

 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->

```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

# Data Preparation
```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))

datelims <- c(as.Date("2020-03-07"), max(df$date))
```

# Counterfactuals

```{r cffuncs, cache=FALSE}
#' Given an estimated felm model with lags and differences, returns a
#' function that simulates the model.
#'
#' @param m estimated felm model
#' @param df Data
#' @param dvar Name of variable repressenting differences in the
#' model. dvar should be a difference of length difflag.
#' @param var Name of varaible representing levels in the model.
#' @param difflag difference length for dvar
#' @param lhsdiff if true the outcome is also dvar with the same
#' difflag, if not the outcome is var
#'
#' @return A function with arguments idvar==value of identifier for
#'   which to simulate and shockgen=function(n) that generates values
#'   for residuals. The default value of shockgen is to resample model
#'   residuals with replacement. The function returns a simulated path
#'   var beginning from the initial values for id==idvar and using the
#'   exogenous covariates of id==idvar.
#'
#' Example:
#' df <- df[order(df$id,$df$t),]
#' df$dy <- paneldiff(df$y, df$id, df$t, lag=2)
#' fmla <- dy ~ lag(y,3) + lag(dy,1) + x
#' # hack to get get id
#'
#' sim <- createforwardsim(fmla, df, "dy", "y", 2, TRUE)
#' # simulate with errors set to 0's
#' yhat <- sim(df$id[1], function(n) fill(0, n))
#' #' # simulate sampling residuals with replacement
#' yhat <- sim(df$id[1])
createforwardsim <- function(m, df, id="state",
                             dvar="dlogdc", var="logdc",
                             difflag=7, lhsdiff=TRUE) {
  m$coef[is.nan(m$coef)] <- 0

  # get index of observations used in mdl
  df$idxcfs <- 1:nrow(df)
  mid <- felm(update.felm.formula(m$formula, idxcfs ~ .), data=df, keepX=TRUE)
  idx <- mid$response

  # This  regexp will match dvar if preceded by start of string or (
  # and followed by end of string or ,
  di <- grepl(sprintf("(^|\\()%s(,|$)",dvar), rownames(m$coef))
  li <- grepl(sprintf("(^|\\()%s(,|$)",var), rownames(m$coef))
  dc <- m$coef[di]
  names(dc) <- rownames(m$coef)[di]
  lc <- m$coef[li]
  names(lc) <- rownames(m$coef)[li]
  # exogenous part of model
  if (!is.null(getfe(m))) {
    stop("createforwardsim not implemented with fixed effects. Include factors in the formula instead")
  }

  xt <- mid$X %*% (m$coef * as.numeric(!(di | li)))
  #xtold <- m$fitted.values - m$X %*% (m$coef * as.numeric(di | li))

  ## convert into AR model
  llags <- as.numeric(gsub("lag\\(.+, (\\d+)\\)","\\1", names(lc)))
  dlags <- as.numeric(gsub("lag\\(.+, (\\d+)\\)","\\1", names(dc)))
  maxlags <- max(c(llags, dlags+difflag, ifelse(lhsdiff, difflag, 0)))
  ar <- rep(0, maxlags)
  if (lhsdiff) ar[difflag]  <- ar[difflag] + 1
  ar[llags]  <- ar[llags] + lc
  ar[dlags] <- ar[dlags] +dc
  ar[dlags+difflag] <- ar[dlags+difflag]-dc

  if (!is(df, "pdata.frame")) {
    warning("df is not a pdata.frame. This function assumes df is sorted by id, time, and within each id, time is consecutive. Results will be incorrect if these assumptions are false.")
  } else {
    stopifnot(all(is.pconsecutive(df)))
  }

  sim <- function(idvar, shockgen=function(n=1) { sample(m$residuals, size=n,
                                                         replace=TRUE)},
                  samplecoefs=FALSE, nsim=100) {
    ids <- df[idx,id]
    if (samplecoefs) {
      V <- vcov(m)
      nac <- is.na(m$coefficients)
      e <- rep(0,nrow(V))
      Vnona <- V[!nac,!nac]
      e[!nac] <- t(rmvnorm(1, mean=rep(0, nrow(Vnona)), sigma=Vnona))

      dc <- m$coef[di] + e[di]
      lc <- m$coef[li] + e[li]
      ## convert into AR model
      ar <- rep(0, maxlags)
      if (lhsdiff) ar[difflag]  <- ar[difflag] + 1
      ar[llags]  <- ar[llags] + lc
      ar[dlags] <- ar[dlags] +dc
      ar[dlags+difflag] <- ar[dlags+difflag]-dc

      x <- xt[ids==idvar] + (mid$X[ids==idvar,] %*% (e * as.numeric(!(di | li))))
    } else {
      x <- xt[ids==idvar]
    }

    date <- df$date[idx[df[idx,id]==idvar]]
    names(date) <- NULL
    it0 <- which(df$date==date[1] & df[,id]==idvar)
    Y0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,var], k)[it0]))
    dY0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,dvar], k)[it0]))
    lagy0 <- Y0-dY0
    check <- lagy0[1:(length(Y0)-difflag)]- Y0[(difflag+1):length(Y0)]
    stopifnot(all(abs(check)<1e-8, na.rm=TRUE))
    na0 <- is.na(Y0[(difflag+1):length(Y0)])
    Y0[(difflag+1):length(Y0)][na0] <- lagy0[1:(length(Y0)-difflag)][na0]

    yt <- matrix(0,nrow=nsim,ncol=length(x))
    e <- matrix(shockgen(length(x)*nsim),nrow=nsim)
    Y0 <- t(matrix(rep(Y0,nsim),ncol=nsim))
    Y <- Y0
    for (t in 1:length(x)) {
      yt[,t] <- x[t] + Y %*% ar + e[,t]
      #Y <- c(yt[t], Y[-length(Y)])
      Y[,2:ncol(Y)] <- Y[,1:(ncol(Y)-1)]
      Y[,1] <- yt[,t]
    }
    return(list(y=cbind(Y0[, ncol(Y0):1],yt),
                date=c(date[1]-((maxlags):1),date)))
  }
}

meanandci <- function(yi, cfyi, date, p=0.9, difflag=7) {

  y <- colMeans(yi)
  cfy <- colMeans(cfyi)
  dy <- matrix(NA,nrow=nrow(y),ncol=ncol(y))
  dy[(difflag+1):nrow(y),]  <- diff(y,difflag)

  dcfy <- matrix(NA,nrow=nrow(cfy),ncol=ncol(cfy))
  dcfy[(difflag+1):nrow(cfy),]  <- diff(cfy,difflag)
  statframe <- function(y,suffix) {
    d <- data.frame(m=rowMeans(y, na.rm=TRUE),
                    cl=apply(y, 1, function(x) quantile(x, (1-p)/2,na.rm=TRUE)),
                    ch=apply(y, 1, function(x) quantile(x,1-(1-p)/2,na.rm=TRUE)))
    names(d) <- paste(names(d), suffix, sep="")
    d
  }

  cbind(data.frame(date=date),
        statframe(colMeans(exp(yi)),"y"), statframe(dy,"dy"), statframe(colMeans(exp(cfyi)),"cf"),
        statframe(dcfy,"dcf"), statframe(colMeans(exp(cfyi)-exp(yi)),"p"),
        statframe(dcfy-dy,"dp"))

}

nopstatesim <- function(st, nsim=200, simobs, simcf, data, var="logdc", dvar="dlogdc") {
  seed <- .Random.seed
  set.seed(seed)
  yi <- replicate(nsim, simobs(st, samplecoefs=TRUE, nsim=nsim/2)$y)
  set.seed(seed)
  cfyi <- replicate(nsim, simcf(st, samplecoefs=TRUE, nsim=nsim/2)$y)
  date <- simobs(st, samplecoefs=TRUE, nsim=2)$date
  est <- meanandci(yi,cfyi, date)
  est <- merge(est, data.frame(state=st,
                        obs=exp(data[data$state==st, var]),
                        dobs=data[data$state==st, dvar],
                        date=data[data$state==st,]$date),
               by="date", all=TRUE)
  return(list(edf=est, y=yi, cfy=cfyi))
}

cfplots <- function(st, cfdata=NULL, simobs=NULL, simcf=NULL,
                    yvar="cases", data) {
  xlims <- datelims
  pds <-
    c("Mandate.face.mask.use.by.employees.in.public.facing.businesses",
      "Date.closed.K.12.schools",
      "Stay.at.home..shelter.in.place",
      "Closed.movie.theaters",
      "Closed.restaurants.except.take.out",
      "Closed.non.essential.businesses")
  if (is.null(cfdata)) {
    est <- nopstatesim(st, simobs=simobs, simcf=simcf, data=data)
  } else {
    est <- subset(cfdata, cfdata$state==st)
  }

  dt  <- c()
  labels <- c()
  for (lbl in pds) {
    d <- unique(data[data$state==st,lbl])
    i <- which(dt==d)
    if (length(i)==0) {
      if (length(dt)==0)
        dt <- d
      else
        dt <- c(dt,d)
      labels <- c(labels,lbl)
    } else {
      labels[i]  <- paste(labels[i],lbl,sep=", ")
    }
  }
  labels <- labels[!is.na(dt)]
  dt <- dt[!is.na(dt)]

  figl <-
    ggplot(data=est, aes(x=date, y=obs)) +
    geom_line(aes(color="Observed", fill="Observed"), size=1.5) +
    geom_line(aes(y=my,color="Estimated"), size=1.2) +
    #geom_ribbon(aes(ymin=cly, ymax=chy, fill="Estimated"),
    #            alpha=0.3, show.legend=FALSE) +
    geom_line(aes(y=mcf, color="Counterfactual"), size=1.2) +
    #geom_ribbon(aes(ymin=clcf, ymax=chcf, fill="Without Policies"),
    #            alpha=0.3, show.legend=FALSE) +
    figtheme + colors() + colors_fill() +
    ylab(sprintf("%s in past week",yvar)) + labs(color="")  +
    theme(legend.position=c(0.2, 0.8)) +
    ggtitle(TeX(sprintf("%s in past week",yvar))) + xlim(xlims)

  figp <- ggplot(data=est, aes(x=date, y=mp)) +
    geom_line(size=1.2) +
    geom_ribbon(aes(ymin=clp, ymax=chp), alpha=0.3) +
    ylab(sprintf("Change in %s in past week", yvar)) + figtheme + colors() + colors_fill() +
    ggtitle(TeX(sprintf("Change in %s",yvar))) +
    xlim(xlims) +
    annotate("text",x=dt, y=rep(0.01,length(dt)),
             label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)

  figd <- ggplot(data=est, aes(x=date, y=dobs)) +
    geom_line(aes(y=mdcf, color="Counterfactual"), size=1.5) +
    geom_line(aes(color="Observed", fill="Observed"), size=1.5) +
    geom_line(aes(y=mdy,color="Estimated"), size=1.5) +
    #geom_ribbon(aes(ymin=cldy, ymax=chdy, fill="Estimated"),
    #            alpha=0.3, show.legend=FALSE) +
    #geom_ribbon(aes(ymin=cldcf, ymax=chdcf, fill="Without Policies"),
    #            alpha=0.3, show.legend=FALSE) +
    figtheme + colors() + colors_fill() +
    ylab(TeX(sprintf("$\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    labs(color="")  +
    theme(legend.position=c(0.8, 0.8)) +
    ggtitle(TeX(sprintf("$\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    xlim(xlims)

  figdp <- ggplot(data=est, aes(x=date, y=mdp)) +
    geom_line(size=1.2) +
    geom_ribbon(aes(ymin=cldp, ymax=chdp), alpha=0.3) +
    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    figtheme + colors() + colors_fill() +
    ggtitle(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
    xlim(xlims) +
    annotate("text",x=dt, y=rep(0.01,length(dt)),
             label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)

  return(list(figd=figd, figdp=figdp, figl=figl, figp=figp))
}

# Extract E[log(dc) |state, t, parameters] with parameters ~ N(m$coef, vcov(m))
getmeans <- function(sl) {
  date <- sl$edf$date[!is.na(sl$edf$my)]
  state <- sl$edf$state[!is.na(sl$edf$state)]
  stopifnot(length(unique(state))==1)
  y0 <- apply(exp(sl$y), c(2,3), mean)
  long <- melt(y0)
  names(long)[names(long)=="value"] <- "meany0"

  log0 <- apply(sl$y, c(2,3), mean)
  long1 <- melt(log0)
  names(long1)[names(long1)=="value"] <- "meanlog0"
  long <- merge(long, long1, by=c("Var1","Var2"))

  y1 <- apply(exp(sl$cf), c(2,3), mean)
  long1 <- melt(y1)
  names(long1)[names(long1)=="value"] <- "meany1"
  long <- merge(long,long1, by=c("Var1","Var2"))

  log1 <- apply(sl$cf, c(2,3), mean)
  long1 <- melt(log1)
  names(long1)[names(long1)=="value"] <- "meanlog1"
  long <- merge(long, long1, by=c("Var1","Var2"))

  long$date <- date[long$Var1]
  long$state <- unique(state)
  long$sidx <- long$Var2
  long$Var1 <- NULL
  long$Var2 <- NULL
  return(long)
}

```

```{r statesim, cache=FALSE}
infovars <- list(c("dlogdc", "logdc")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
m <- regs$piy[[1]]$reg
cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

subcoefs <- function(di, coef) {
  for (r in 1:(nrow(di)-1)) {
    i <- grep(rownames(di)[r] , rownames(coef))
    coef[i,1] <- di[r,"Average"]
  }
  return(coef)
}

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))

V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf)

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf)

rerunsim=FALSE
if (!file.exists("cfnop.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(sdf$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop.Rda")
}
load("cfnop.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies

We now examine the impact changing from the observed policies to
none. For illustrative purpose, we being by focusing on
Washington. Figure \ref{fig:WA-nop-dlogdc} shows the observed,
estimated average, and counterfactual without policies average of
$\Delta \log \Delta C$ in Washington. To compute the estimated and
counterfactual paths we use the estimates from column 2 of table
\ref{tab:policy-cases}. We set initial $\Delta \log \Delta C$ and
$\log \Delta C$ to their values first observed in the state we are
simulating. We hold all other regressors at their observed
values. Error terms are drawn with replacement from the residuals. We
do this many times and report the average over draws of the
residuals. To calculate confidence intervals, we repeat the above with
coefficients drawn randomly from their asymptotic distribution. The
shaded region is a pointwise 90\% confidence interval.

```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

```{r WA-nop-cases, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
if (!knitr::is_latex_output()) {
  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
    f <- cfplots(st, cfnop, data=df)
    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
  }
}
```

```{r nop-state-figures, cache=TRUE, warning=FALSE}
states <- c("Washington","Illinois","Massachusetts")
for(st in states) {
  f <- cfplots(st, cfnop, data=df)

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figd + ylim(c(-1,5)))
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figl)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}
```

## National

These figures show the average effect on $\Delta \log \Delta C$ across
states, and the total national effect on cases in past week.

```{r nop-all, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

nationalplots <- function(alldf,cfdf, cfname, yvar="cases") {
  alldf[order(state, sidx, date), dlog1:=c(rep(NA,7),diff(meanlog1,7)), by=.(state, sidx)]
  alldf[order(state, sidx, date),dlog0:=c(rep(NA,7),diff(meanlog0,7)), by=.(state, sidx)]
  national <- alldf[ , .(diff=sum(meany1-meany0), y0=sum(meany0), dlog=mean(dlog1-dlog0)), by=.(date, sidx)]

  adf <- national[, .(mp=mean(diff),
                      clo=quantile(diff, 0.05),
                      chi=quantile(diff, 0.95),
                      rel=mean(diff/y0),
                      rlo=quantile(diff/y0, 0.05),
                      rhi=quantile(diff/y0, 0.95),
                      mlog=mean(dlog, na.rm=TRUE),
                      llo=quantile(dlog, 0.025, na.rm=TRUE),
                      lhi=quantile(dlog, 0.975, na.rm=TRUE)
                      ),
                  keyby=.(date)]

  figdlog <- ggplot(cfdf, aes(x=date)) +
    #geom_text(aes(y=mdp, label=ST),alpha=0.2) +
    geom_point(aes(y=mdp),alpha=0.2, size=0.5) +
    xlim(datelims) +
    geom_line(data=adf,  aes(x=date, y=mlog, color=""), size=2) + figtheme +
    geom_ribbon(data=adf, aes(ymin=llo, ymax=lhi, fill=""),alpha=0.3) +
    colors() + theme(legend.position="none") +
    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",
                     toupper(substring(yvar,1,1))
                     ))) +
    colors_fill() +
     ggtitle(sprintf("Effect of %s on %s growth",cfname, substring(yvar,1,nchar(yvar)-1)))

  figc <- ggplot(cfdf, aes(x=date)) +
    #geom_text(aes(y=mp, label=ST),alpha=0.5) +
    geom_point(aes(y=mp),alpha=0.2, size=0.5) +
    xlim(datelims) +
    geom_line(data=adf, aes(x=date, y=mp, color=""), size=2) +
    figtheme +
    geom_ribbon(data=adf, aes(ymin=clo, ymax=chi, fill=""),alpha=0.3) +
    colors() + theme(legend.position="none") +
    ylab(sprintf("%s in the past week", yvar)) +
    colors_fill() +
    ggtitle(sprintf("Effect of %s on %s",cfname, yvar))

  figr <- ggplot(adf, aes(x=date, y=rel)) +
    xlim(datelims) +
    geom_line(aes(color=""), size=2) + figtheme +
    geom_ribbon(data=adf, aes(ymin=rlo, ymax=rhi, fill=""),alpha=0.3) +
    colors() + theme(legend.position="none") +
    ylab(sprintf("relative increase in %s",yvar)) +
    colors_fill() +
    ggtitle(sprintf("Relative effect of %s",cfname))
  return(list(g=figdlog, c=figc, r=figr))
}

figs <- nationalplots(allmeans,cfnop, "removing policies")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non-Essential Business Closures

```{r business_sim, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf)

if (!file.exists("cfnb.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb.Rda")
}
load("cfnb.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf)

if (!file.exists("cfmask.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask.Rda")
}
load("cfmask.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf)

if (!file.exists("cfshelter.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter.Rda")
}
load("cfshelter.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df)
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

# Deaths

```{r statesim-deaths, cache=FALSE}
infovars <- list(c("dlogdd", "logdd")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- NULL
xlist <-list("")
interactions <-list("month", statevars)
ilist <- list(interactions) #, interactions)
iv <- list("0") #,"(testratedc ~ lag(testratedc,14))")

L <- 21
sdf <- subset(df, df$date>=datelims[1]) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)

m <- regs$piy[[1]]$reg

cnames <- c(pols, infovars[[1]])
di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                  regs$pbiy[[1]]$reg$coef[,1],
                  regs$piy[[1]]$reg$coef[,1],
                  policies=cnames)

pib <- lapply(regs$pib[[1]][1:4], function(x) x$reg)
pbiy <- regs$pbiy[[1]]$reg
piy <- regs$piy[[1]]$reg

S <- 999
bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
bdi <- lapply(1:S, function(i)
  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
              bs[[5]][,i], bs[[6]][,i],
              policies=cnames)
  )

bcoefs <- t(do.call(cbind, lapply(1:S, function(i) {
  coef <- matrix(bs[[6]][,i],ncol=1)
  rownames(coef) <- rownames(bs[[6]])
  return(subcoefs(bdi[[i]],coef ))
})))
V <- cov(bcoefs)
m$coefficients <- subcoefs(di, m$coefficients)
m$vcv <- V
m$clustervcv <- V
m$robustvcv <- V

simmodel <- createforwardsim(m, sdf, dvar="dlogdd", var="logdd")

# remove policies
cf <- sdf
cf[,pols] <- 0
nopmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

rerunsim=FALSE
if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
  cfnoplist <- lapply(unique(df$state), function(s)
    nopstatesim(s,simobs=simmodel,simcf=nopmodel, data=sdf, var="logdd", dvar="dlogdd"))
  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
  save(cfnoplist,allmeans, file="cfnop_deaths.Rda")
}
load("cfnop_deaths.Rda")
cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
```

### Removing Policies

```{r WA-nop-dlogdd, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
st <- "Washington"
f <- cfplots(st, cfnop, yvar="deaths", data=df)
grid.arrange(f$figd, f$figdp, nrow=2, top =
                                        textGrob(st,gp=gpar(fontsize=20,font=2)))

```

```{r WA-nop-deaths, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
grid.arrange(f$figl, f$figp, nrow=2) #,
#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
```

```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
if (!knitr::is_latex_output()) {
  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
    f <- cfplots(st, cfnop, yvar="deaths", data=df)
    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
  }
}
```

```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
states <- c("Washington","Illinois","Massachusetts")
for(st in states) {
  f <- cfplots(st, cfnop, yvar="deaths", data=df)

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figd + ylim(c(-1,5)))
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figl)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}
```

## National

```{r nop-all-deaths, cache=TRUE}
abbrv <- unique(df[,c("state","ST")])
cfnop <- merge(cfnop, abbrv, by="state")

figs <- nationalplots(allmeans,cfnop, "removing policies", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Non Essential Business Closures

```{r business_sim-deaths, cache=FALSE}
# remove non essential business closure
cf <- sdf
cf[,"pnonessential"] <- 0
nobmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
  nball <- rbindlist(lapply(cfnblist, getmeans))
  save(cfnblist,nball, file="cfnb_deaths.Rda")
}
load("cfnb_deaths.Rda")
cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
```

```{r, cache=TRUE}
cfnb <- merge(cfnb, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfnb, yvar="deaths", data=df)
  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
  #textGrob(st,gp=gpar(fontsize=20,font=2)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(nball, cfnb,
                      "never shutting down non-essential\nbusinesses", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Masks

We now look at what happens when all states impose a mask mandate on
April 1st.

```{r mask_sim-deaths, cache=FALSE}
cf <- df
cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
cf[,"pmaskbus"] <-
  smoothedpolicy(cf,"Mandate.face.mask.use.by.employees.in.public.facing.businesses",type="ma",bw=6)
cf <- subset(cf, cf$date>=datelims[1]) #+L)
maskmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
  maskall <- rbindlist(lapply(cfmasklist, getmeans))
  save(cfmasklist, maskall, file="cfmask_deaths.Rda")
}
load("cfmask_deaths.Rda")
cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
```

```{r, cache=TRUE}
cfmask <- merge(cfmask, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfmask, yvar="deaths", data=df)
  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

## Stay at home

We now look at what happens if stay at home orders were never issued.

```{r home-deaths, cache=FALSE}
cf <- df
cf[,"pshelter"] <- 0
cf <- subset(cf, cf$date>=datelims[1]) #+L)
noshelter <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")

if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
  save(cfshelterlist, shelterall, file="cfshelter_deaths.Rda")
}
load("cfshelter_deaths.Rda")
cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
```

```{r, cache=TRUE}
cfshelter <- merge(cfshelter, abbrv, by="state")
for (st in c("Washington","Illinois","Massachusetts")) {
  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figdp)
  dev.off()

  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
  print(f$figp)
  dev.off()
}

figs <- nationalplots(shelterall, cfshelter, "no stay at home orders", yvar="deaths")

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$g
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$c
dev.off()

figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
figs$r
dev.off()

for (f in figs) print(f)
```

### 2020-07-16 cases_and_policies/rmd/counterfactuals_national.Rmd
Version submitted in early July 2020.
Compared to submitted version the following code changes have been made:

- Removed safegraph and unacast data (they can't be publicly shared)
- Minor update to policy data loading due to data format change
- Added explicit subsetting of data to June 2, 2020 and earlier (to match what we used in submitted paper).

--- 2020-07-16 cases_and_policies/rmd/counterfactuals.Rmd
@@
 ```{r statesim, cache=FALSE}
-infovars <- list(c("dlogdc", "logdc")) #,
+infovars <- list(c("dlogdc", "logdc", "dlogdc.national", "logdc.national")) #,
 #c("dlogdc", "logdc"))
@@
 rerunsim=FALSE
-if (!file.exists("cfnop.Rda") || rerunsim) {
+if (!file.exists("cfnop_NI.Rda") || rerunsim) {
   cfnoplist <- lapply(unique(sdf$state), function(s)
@@
   allmeans <- rbindlist(lapply(cfnoplist, getmeans))
-  save(cfnoplist,allmeans, file="cfnop.Rda")
-}
-load("cfnop.Rda")
+  save(cfnoplist,allmeans, file="cfnop_NI.Rda")
+}
+load("cfnop_NI.Rda")
 cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth.pdf",rootdir,st), width=5,height=3.75)
   print(f$figd + ylim(c(-1,5)))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases.pdf",rootdir,st), width=5,height=3.75)
   print(f$figl)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 ```{r business_sim, cache=FALSE}
-# remove non essential business closure
+# remove non-essential business closure
 cf <- sdf
@@
 
-if (!file.exists("cfnb.Rda") || rerunsim) {
+if (!file.exists("cfnb_NI.Rda") || rerunsim) {
   cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf))
   nball <- rbindlist(lapply(cfnblist, getmeans))
-  save(cfnblist,nball, file="cfnb.Rda")
-}
-load("cfnb.Rda")
+  save(cfnblist,nball, file="cfnb_NI.Rda")
+}
+load("cfnb_NI.Rda")
 cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfmask.Rda") || rerunsim) {
+if (!file.exists("cfmask_NI.Rda") || rerunsim) {
   cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
-  save(cfmasklist, maskall, file="cfmask.Rda")
-}
-load("cfmask.Rda")
+  save(cfmasklist, maskall, file="cfmask_NI.Rda")
+}
+load("cfmask_NI.Rda")
 cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-growth.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figd + ylim(c(-1,5)))
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfshelter.Rda") || rerunsim) {
+if (!file.exists("cfshelter_NI.Rda") || rerunsim) {
   cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfshelter.Rda")
-}
-load("cfshelter.Rda")
+  save(cfshelterlist, shelterall, file="cfshelter_NI.Rda")
+}
+load("cfshelter_NI.Rda")
 cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 ```{r statesim-deaths, cache=FALSE}
-infovars <- list(c("dlogdd", "logdd")) #,
+infovars <- list(c("dlogdd", "logdd", "dlogdd.national", "logdd.national")) #,
 #c("dlogdc", "logdc"))
@@
 rerunsim=FALSE
-if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
+if (!file.exists("cfnop_deaths_NI.Rda") || rerunsim) {
   cfnoplist <- lapply(unique(df$state), function(s)
@@
   allmeans <- rbindlist(lapply(cfnoplist, getmeans))
-  save(cfnoplist,allmeans, file="cfnop_deaths.Rda")
-}
-load("cfnop_deaths.Rda")
+  save(cfnoplist,allmeans, file="cfnop_deaths_NI.Rda")
+}
+load("cfnop_deaths_NI.Rda")
 cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figd + ylim(c(-1,5)))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figl)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-## Non Essential Business Closures
+## Non-Essential Business Closures
 
 ```{r business_sim-deaths, cache=FALSE}
-# remove non essential business closure
+# remove non-essential business closure
 cf <- sdf
@@
 
-if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
+if (!file.exists("cfnb_deaths_NI.Rda") || rerunsim) {
   cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
   nball <- rbindlist(lapply(cfnblist, getmeans))
-  save(cfnblist,nball, file="cfnb_deaths.Rda")
-}
-load("cfnb_deaths.Rda")
+  save(cfnblist,nball, file="cfnb_deaths_NI.Rda")
+}
+load("cfnb_deaths_NI.Rda")
 cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
+if (!file.exists("cfmask_deaths_NI.Rda") || rerunsim) {
   cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
-  save(cfmasklist, maskall, file="cfmask_deaths.Rda")
-}
-load("cfmask_deaths.Rda")
+  save(cfmasklist, maskall, file="cfmask_deaths_NI.Rda")
+}
+load("cfmask_deaths_NI.Rda")
 cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-growth_deaths.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figd + ylim(c(-1,5)))
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
+if (!file.exists("cfshelter_deaths_NI.Rda") || rerunsim) {
   cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfshelter_deaths.Rda")
-}
-load("cfshelter_deaths.Rda")
+  save(cfshelterlist, shelterall, file="cfshelter_deaths_NI.Rda")
+}
+load("cfshelter_deaths_NI.Rda")
 cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths.pdf",sep="/"), width=5,height=3.75)
 figs$r

### 2020-07-16 cases_and_policies/rmd/datasummary.Rmd
Version submitted in early July 2020.
Compared to submitted version the following code changes have been made:

- Removed safegraph and unacast data (they can't be publicly shared)
- Minor update to policy data loading due to data format change
- Added explicit subsetting of data to June 2, 2020 and earlier (to match what we used in submitted paper).

---
title       : ""
subtitle    : "Data description and summary statistics"
author      : Paul Schrimpf, Hiro Kasahara, Victor Chernozhukov
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        code_folding: hide
        fig_width: 8
        fig_height: 6
## To create html files from this, in R enter `library(rmarkdown);
## rmarkdown::render("regressionsWithPlots.Rmd")`
##

## to create the gh-pages hosted website in R enter
## rmarkdown::render_site()
##
## Then in a shell
## $ ghp-import -p -n -m "$(date)" _site
##

---

```{r setup}
library(ggthemes)
library(ggplot2)
library(gridExtra)
library(knitr)
library(kableExtra)

rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
datafile <- paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/")
cat("Loading data last updated on ", as.character(file.mtime(datafile)), "\n")
cat("Run ", paste(rootdir,"cases_and_policies/R/cases_and_policies.R", sep="/"), " to update data.\n")
load(datafile)
source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))

plottheme <- figtheme
colors <-  scale_color_solarized

kable <- function(tbl, ...) kable_styling(knitr::kable(tbl, ...),
                                          bootstrap_options=c("striped","condensed","responsive"),
                                          full_width=TRUE)

cs <- covidstates
cs <- cs[cs$fips<60, ] # drop territories (Puerto Rico, Guam, etc)
cs <- cs[order(cs$state, cs$date),]
```

# Comparing Datasets on Cases and Deaths

```{r, cache=TRUE}
vars <- c("cases.nyt","cases","positive")
tbl <- cor(cs[,vars],use="complete")
rownames(tbl) <- colnames(tbl) <- c("NYT","JHU","CTP")
kable(tbl, digits=5, caption="Correlation of cumulative cases \\label{tab:casecor}")

tbl2 <- matrix(NA,3,3)
for(i in 1:3) {
  for(j in 1:3) {
    tbl2[i,j] <- mean(cs[,vars[i]]==cs[,vars[j]], na.rm=TRUE)
  }
}
rownames(tbl2) <- rownames(tbl)
colnames(tbl2) <- colnames(tbl)
kable(tbl2, digits=5,
      caption="Portion of days that cumulative cases are equal between data sets \\label{tab:eqcases}")

```

## Daily Cases

```{r newcases, cache=FALSE}
startday <- as.Date("2020-02-15")
cs$newcases.nyt <- paneldiff(cs$cases.nyt, cs$state, cs$date, lag=1)
cs$newcases.jhu <- paneldiff(cs$cases, cs$state, cs$date, lag=1)
cs$newcases.ctp <- paneldiff(cs$positive, cs$state, cs$date, lag=1)

fctp <- ggplot(cs, aes(x=date, y=newcases.ctp, colour=state)) + geom_line(alpha=1) +
  xlim(as.Date(startday),max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("CTP") + ylab("New cases") +
  theme(legend.position = "none")
fnyt <- ggplot(cs, aes(x=date, y=newcases.nyt, colour=state)) + geom_line(alpha=1) +
  xlim(as.Date(startday),max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("NYT") + ylab("New cases") +
  theme(legend.position = "none")
fjhu <- ggplot(cs, aes(x=date, y=newcases.jhu, colour=state)) + geom_line(alpha=1) +
  xlim(as.Date(startday),max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("JHU") + ylab("New cases") +
  theme(legend.position = "none")

grid.arrange(fnyt,fjhu,fctp)

X  <- cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]
X[X<0] <- NA
X[X==0] <- 0.1
tbl  <- colSums(cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]<0, na.rm=TRUE)
tbl <- cor(log(X), use="complete")
tbl <- rbind(tbl, diag(cov(log(X), use="complete")))
colnames(tbl) <- c("NYT","JHU","CTP")
rownames(tbl) <- c("NYT","JHU","CTP","Variance")

kable(tbl, digits=4, caption="Correlation and variance of log daily new cases\\label{tab:newcasecor}")
```

## Weekly Cases

```{r}
cs$newcases.nyt <- paneldiff(cs$cases.nyt, cs$state, cs$date, lag=7)
cs$newcases.jhu <- paneldiff(cs$cases, cs$state, cs$date, lag=7)
cs$newcases.ctp <- paneldiff(cs$positive, cs$state, cs$date, lag=7)
X  <- cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]
X[X<0] <- NA
X[X==0] <- 0.1
tbl  <- colSums(cs[,c("newcases.nyt","newcases.jhu","newcases.ctp")]<0, na.rm=TRUE)
tbl <- cor(log(X), use="complete")
tbl <- rbind(tbl, diag(cov(log(X), use="complete")))
colnames(tbl) <- c("NYT","JHU","CTP")
rownames(tbl) <- c("NYT","JHU","CTP","Variance")
kable(tbl, digits=4, caption="Correlation and variance of log weekly new cases\\label{tab:weekcasecor}")

fctp <- ggplot(cs, aes(x=date, y=newcases.ctp, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("CTP") + ylab("New cases") +
  theme(legend.position = "none")
fnyt <- ggplot(cs, aes(x=date, y=newcases.nyt, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("NYT") + ylab("New cases") +
  theme(legend.position = "none")
fjhu <- ggplot(cs, aes(x=date, y=newcases.jhu, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("JHU") + ylab("New cases") +
  theme(legend.position = "none")

grid.arrange(fnyt,fjhu,fctp)
```

## Deaths

```{r deathplots, cache=TRUE}
cs$newdeaths.nyt <- paneldiff(cs$deaths.nyt, cs$state, cs$date, lag=7)
cs$newdeaths.jhu <- paneldiff(cs$deaths, cs$state, cs$date, lag=7)
cs$newdeaths.ctp <- paneldiff(cs$deaths.ctp, cs$state, cs$date, lag=7)
X  <- cs[,c("newdeaths.nyt","newdeaths.jhu","newdeaths.ctp")]
tbl  <- colSums(cs[,c("newdeaths.nyt","newdeaths.jhu","newdeaths.ctp")]<0, na.rm=TRUE)
tbl <- cor(X, use="complete")
tbl <- rbind(tbl, diag(cov(X, use="complete")))
colnames(tbl) <- c("NYT","JHU","CTP")
rownames(tbl) <- c("NYT","JHU","CTP","Variance")
kable(tbl, digits=4, caption="Correlation and variance of weekly deaths\\label{tab:weekdeathcor}")

fctp <- ggplot(cs, aes(x=date, y=newdeaths.ctp, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("CTP") + ylab("Deaths") +
  theme(legend.position = "none")
fnyt <- ggplot(cs, aes(x=date, y=newdeaths.nyt, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("NYT") + ylab("Deaths") +
  theme(legend.position = "none")
fjhu <- ggplot(cs, aes(x=date, y=newdeaths.jhu, colour=state)) + geom_line(alpha=1) +
  xlim(startday,max(cs$date)) + scale_y_log10() + plottheme +
  ggtitle("JHU") + ylab("Deaths") +
  theme(legend.position = "none")

grid.arrange(fnyt,fjhu,fctp)
```

# Testing data

Our data on tests is from the [Covid Tracking
Project](https://covidtracking.com/). This dataset contains the following variables.

```{r , cache=TRUE}
kable(dataguide[dataguide$source=="Covid-Tracking Project",c("name","source")])
```

We have been using the `totalTestResults` variable.

```{r testplots, cache=TRUE}
fign <- ggplot(aggregate(totalTestResults ~ date, FUN=function(x) sum(!is.na(x)), data=cs)) +
  geom_line(aes(x=date, y=totalTestResults, colour="Tests"), size=1.5) +
  geom_line(data=aggregate(cases ~ date, FUN=function(x) sum(!is.na(x)), data=cs),
            aes(x=date, y=cases, colour="Cases"), size=1.3) +
  geom_line(data=aggregate(deaths ~ date, FUN=function(x) sum(!is.na(x)), data=cs),
            aes(x=date, y=deaths, colour="Deaths"), size=1.3) +
            ylab("Number non-missing") +
            xlab("") + plottheme + ggtitle("Number of non-missing observations") + colors()

p <- c(0.05, 0.25, 0.5, 0.75, 0.95)
adf <- aggregate(totalTestResults ~ date,
                 FUN=function(x) quantile(x, probs=p,
                                          na.rm=TRUE), data=cs)

## The following code sucks and the fact that the equivalent loop
## doesn't work reflects all that is wrong with ggplot2 and the
## tidyverse
p <- as.factor(p)
s <- 1.5
figm <- ggplot(adf) +
  geom_line(aes(x=date, y=adf$totalTestResults[,1], color=p[1]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,2], color=p[2]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,3], color=p[3]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,4], color=p[4]), size=s) +
  geom_line(aes(x=date, y=adf$totalTestResults[,5], color=p[5]), size=s) +
  plottheme + scale_y_log10() + ylab("Cumulative tests") +
  colors() + ggtitle("Quantiles of cumulative tests") +
  geom_line(data=cs, aes(x=date, y=totalTestResults, group=state), alpha=0.2) +
  theme(legend.position="bottom")

grid.arrange(fign, figm, layout_matrix=matrix(1:2, nrow=2), heights=c(0.5,1))
```

The top panel shows that states began reporting test numbers in early
March at earliest, and test numbers are included for all states by mid
March. A few states began reporting cases earlier than March, but
mostly case reporting and test reporting coincide.

In the bottom panel, the thick colored lines are quantiles of total
cumulative tests, and the gray lines are tests in each state.

# Policies

```{r, cache=TRUE}
rvars <- dataguide[dataguide$source=="Raifman et al", "name"]
pols  <- unique(cs[,c("state","fips",rvars)])
di <- sapply(cs[,rvars], function(x) inherits(x,"Date"))
pvars <- rvars[di]

tbl <- cbind(sapply(pols[,pvars], function(x) sum(!is.na(x)), USE.NAMES=FALSE),
             sapply(pols[,pvars], function(x) min(x, na.rm=TRUE), USE.NAMES=FALSE),
             sapply(pols[,pvars], function(x) median(x, na.rm=TRUE), USE.NAMES=FALSE),
             sapply(pols[,pvars], function(x) max(x, na.rm=TRUE), USE.NAMES=FALSE))
tbl[,2:4] <- as.character(as.Date(as.numeric(tbl[,2:4]), origin=as.Date("1970-01-01", format="%Y-%m-%d")))
rownames(tbl) <- gsub("\\."," ",pvars)
colnames(tbl) <- c("N","Min","Median","Max")
kable(tbl, digits=0, caption="State Policies \\label{tab:policies}", )
```

```{r policyplot, cache=TRUE, warning=FALSE}
rvars <- dataguide[dataguide$source=="Raifman et al", "name"]
pols  <- unique(cs[,c("state","fips",rvars)])
di <- sapply(cs[,rvars], function(x) inherits(x,"Date"))
pols <- rvars[di]

plotstatepolicies  <- function(st) {
  sdf  <- subset(cs, state==st)
  x  <- c()
  labels <- c()
  for (lbl in pols) {
    d <- unique(sdf[,lbl])
    i <- which(x==d)
    if (length(i)==0) {
      if (length(x)==0)
        x <- d
      else
        x <- c(x,d)
      labels <- c(labels,lbl)
    } else {
      labels[i]  <- paste(labels[i],lbl,sep=", ")
    }
  }
  labels <- labels[!is.na(x)]
  x <- x[!is.na(x)]
  y <- rep(1, length(x))
  fig <- ggplot(sdf,aes(x=date))  +
    geom_line(aes(y=cases)) +
    geom_line(aes(y=deaths), colour="red") +
    xlim(startday,max(sdf$date)) +
    ylab("count") +
    scale_y_continuous(trans="log10") + plottheme + colors() +
    ggtitle(st)

  fig + annotate("text", x=x,y=y,label=gsub("\\."," ",labels),
                 size=2, angle=90, hjust=0)

}

i <- 0
for (st in unique(cs$state)) {
  if (i == 0) fig <- list()
  fig <- c(fig, list(plotstatepolicies(st)))
  i <- i+1
  if (i %% 3 == 0) {
    grid.arrange(grobs=fig, nrow=3, ncol=1)
    i = 0
  }
}
if (i != 0) {
  fig <- c(fig, rep(list(ggplot() + plottheme), 3-i))
  grid.arrange(grobs=fig, nrow=3, ncol=1)
}
```

```{r npolicies, warning=FALSE, cache=TRUE}
df <- cs[order(cs$state,cs$date),]
pdates <- c("Date.closed.K.12.schools",
            "Stay.at.home..shelter.in.place",
            "Closed.movie.theaters",
            "Closed.restaurants.except.take.out",
            "Closed.non.essential.businesses",
            "Mandate.face.mask.use.by.all.individuals.in.public.spaces",
            "Mandate.face.mask.use.by.employees.in.public.facing.businesses",
            "State.of.emergency")
enddates <- c(NA,
              "End.relax.stay.at.home.shelter.in.place",
              "Reopened.movie.theaters",
              "Reopen.restaurants",
              "Began.to.reopen.businesses.statewide",
              NA,
              NA,
              NA)
stopifnot(length(enddates)==length(pdates))

pvars <- c("pk12", "pshelter", "pmovie", "prestaurant", "pnonessential",
           "pmaskall","pmaskbus", "psoe")
for(i in 1:length(pdates)) df[,pvars[i]] <-
                             smoothedpolicy(df,pdates[i],enddatevar=enddates[i],type="ma",bw=0)

pols <- c("psoe","pmaskbus","pk12", "pshelter", "pmovie", "prestaurant", "pnonessential")

ppplot <- function(pol, df) {
  df$y <- df[,pol]
  adf <- aggregate(y ~ date, data=df, function(x) mean(x, na.rm=TRUE))
  fig <- ggplot(adf, aes(x=date, y=y)) + geom_line(size=1.2) +
    plottheme + ylab("Portion of States") +
    colors() + ggtitle(varlabels[varlabels$Variable==pol,"Label"]) +
    theme(legend.position="none") +
    xlim(startday, max(df$date)) + ylim(0,1)
  return(fig)
}
figs <- lapply(pols,function(p) ppplot(p, df))

for (i in 1:length(pols)) {
  pdf(sprintf("%s/tex/tables_and_figures/%s_p.pdf",rootdir, pols[i]), width=4, height=2)
  print(figs[[i]])
  dev.off()
}
grid.arrange(grobs=figs[2:7], nrow=3, ncol=2)
```

# Social Distancing Measures

```{r, warning=FALSE}
gv <- dataguide$name[dataguide$source=="Google Mobility Reports"]
hv <- dataguide$name[dataguide$source=="Homebase"]
uv <- c("daily_distance_diff", "daily_visitation_diff",
        "encounters_rate", "log_encounters_rate")
cs$log_encounters_rate <- log(1+cs$encounters_rate)
```

## Google Mobility Reports

```{r gplots, cache=TRUE, warning=FALSE}
## fign <-
##   ggplot(aggregate(retail_and_recreation_percent_change_from_baseline
##                    ~ date,
##                    FUN=function(x) sum(!is.na(x)), data=cs)) +
##   geom_line(aes(x=date, y=retail_and_recreation_percent_change_from_baseline), size=1) +
##   ylab("Number non-missing") +
##   theme(legend.position = "none") +
##   xlab("") + plottheme + ggtitle("Number of non-missing observations of Google Mobility Reports") +
##   colors() + xlim(startday, max(cs$date))
## fign

quantileplot <- function(v, p=c(0.05, 0.25, 0.5, 0.75, 0.95), data=cs) {
  data$y <- data[,v]
  adf <- aggregate(y ~ date, data=data,
                   FUN=function(x) quantile(x, probs=p, na.rm=TRUE))
  p <- as.factor(p)
  s <- 1.2
  txt <- data.frame(date=max(adf$date), y=adf$y[nrow(adf$y),], p=p)
  names(adf)[1] <- "date"
  figm <- ggplot(adf) +
    geom_line(aes(x=date, y=adf$y[,1], color=p[1]), size=s) +
    geom_line(aes(x=date, y=adf$y[,2], color=p[2]), size=s) +
    geom_line(aes(x=date, y=adf$y[,3], color=p[3]), size=s) +
    geom_line(aes(x=date, y=adf$y[,4], color=p[4]), size=s) +
    geom_line(aes(x=date, y=adf$y[,5], color=p[5]), size=s) +
    geom_text(data=txt,aes(x=date,y=y,color=p, label=p,
                           hjust="left",vjust="center", size=2)) +
    plottheme + ylab("Percent Change") +
    colors() + ggtitle(TeX(paste("Evolution of",
                             varlabels[varlabels$Variable==v,"Label"]))) +
    geom_line(data=data, aes(x=date, y=y, group=state), alpha=0.2) +
    theme(legend.position="none") +
    #labs(color='Quantile')  +
    xlim(startday, max(data$date))
}

figm <- lapply(gv, quantileplot)

for (i in seq(1,length(figm),by=2)) {
  grid.arrange(grobs=figm[i:(i+1)], nrow=2)
}
for (i in 1:length(figm)) {
  vl <- strsplit(gv[i],"_")[[1]][1]
  pdf(sprintf("%s/tex/tables_and_figures/%s.pdf",rootdir,vl), width=7, height=3.5)
  print(figm[[i]])
  dev.off()
}

```

The following figure shows the residuals from projecting each of the
above variables on the social distancing metrics we use in our main
analysis,
```{r, cache=TRUE}
sdused <- c(#"residential_percent_change_from_baseline",
    "retail_and_recreation_percent_change_from_baseline",
    "transit_stations_percent_change_from_baseline",
    "workplaces_percent_change_from_baseline",
    "grocery_and_pharmacy_percent_change_from_baseline")
kable(t(varlabels[sapply(sdused, function(x) which(x==varlabels$Variable)),"Label"]))
```
no smoothing is used in the figure.

The table shows the $R^2$ from regressing each of the social distance
measures on the ones used in our main regressions. We report this both
with and without smoothing.

```{r, cache=TRUE}
residanalysis <- function(yvars, xvars=sdused, df=cs) {

  madf <- df[,c("state","date",yvars,xvars)]
  for(v in c(yvars, xvars)) {
    madf[,v] <- panelma(madf[,v], madf$state, madf$date, 7)
  }

  r <- lapply(yvars, function(v) {
    if (v %in% xvars) {
      list(r.squared=1, r.squared.ma=1, residuals=rep(0,nrow(df)))
    } else {
      m <- lm(as.formula(paste(v,"~",
                               paste(xvars, collapse=" + "), sep=" ")),
              data=df,na.action=na.exclude)
      ma <- lm(as.formula(paste(v,"~",
                               paste(xvars, collapse=" + "), sep=" ")),
               data=madf,na.action=na.exclude)
      list(r.squared=summary(m)$r.squared,
           r.squared.ma=summary(ma)$r.squared,
           residuals=residuals(m))
    }
  })

  tmp <- data.frame()
  for (i in 1:length(yvars)) {
    tmp <- rbind(tmp,
                 data.frame(var=varlabels[varlabels$Variable==yvars[i],"Label"],
                            date=df$date, state=df$state,
                            residuals=r[[i]]$residuals))
  }
  tbl <- t(matrix(c(sapply(r, function(x) x$r.squared),
                    sapply(r, function(x) x$r.squared.ma)),
                 nrow=length(yvars)))
  colnames(tbl) <- varlabels[sapply(yvars, function(x) which(x==varlabels$Variable)),"Label"]
  rownames(tbl) <- c("No smoothing", "7 day moving average")

  fig <- ggplot(tmp, aes(x=date, y=residuals,color=var)) +
    geom_point(alpha=0.7) + colors() + plottheme + ylab("Residuals") +
    labs(color="") + theme(legend.position=c(0.05, 0.95)) +
    xlim(as.Date("2020-03-01"),max(df$date))
  return(list(fig, tbl))
}

out <- residanalysis(gv, sdused)

kable(out[[2]],
      caption="Variance explained by Google Mobility measures used in regressions")

out[[1]]
```

## Homebase

```{r hplots, cache=TRUE, warning=FALSE}
## fign <-
##   ggplot(aggregate(percentchangehours ~ date,
##                    FUN=function(x) sum(!is.na(x)), data=cs)) +
##   geom_line(aes(x=date, y=percentchangehours), size=1) +
##   ylab("Number non-missing") +
##   theme(legend.position = "none") +
##   xlab("") + plottheme + ggtitle("Number of non-missing observations of HomeBase Variables") +
##   colors() + xlim(startday, max(cs$date))
## fign

figm <- lapply(hv, quantileplot)
grid.arrange(grobs=figm, nrow=2)

out <- residanalysis(hv, sdused)

kable(out[[2]],
      caption="Variance explained by Google Mobility measures used in regressions")

out[[1]]
```

<!-- ## Unacast -->

<!-- ```{r uplots, cache=TRUE, warning=FALSE} -->
<!-- ## fign <- -->
<!-- ##   ggplot(aggregate(encounters_rate ~ date, -->
<!-- ##                    FUN=function(x) sum(!is.na(x)), data=cs)) + -->
<!-- ##   geom_line(aes(x=date, y=encounters_rate), size=1) + -->
<!-- ##   ylab("Number non-missing") + -->
<!-- ##   theme(legend.position = "none") + -->
<!-- ##   xlab("") + plottheme + ggtitle("Number of non-missing observations of Unacast Variables") + -->
<!-- ##   colors() + xlim(startday, max(cs$date)) -->
<!-- ## fign -->

<!-- figm <- lapply(uv, quantileplot) -->
<!-- figm <- lapply(figm, function(x) x + ylab("")) -->
<!-- for (i in seq(1,length(figm), by=2)) { -->
<!--   grid.arrange(grobs=figm[i:(i+1)], nrow=2) -->
<!-- } -->

<!-- out <- residanalysis(uv, sdused) -->

<!-- kable(out[[2]], -->
<!--       caption="Variance explained by Google Mobility measures used in regressions") -->

<!-- out[[1]] -->
<!-- ``` -->

<!-- ## SafeGraph -->

<!-- ```{r splots, cache=TRUE, warning=FALSE} -->
<!-- sv <- c("distance_traveled_from_home", "completely_home_device_count", -->
<!--         "median_home_dwell_time", "part_time_work_behavior_devices", -->
<!--         "full_time_work_behavior_devices","portion_completely_home", -->
<!--         "portion_full_time_work", "portion_part_time_work") -->
<!-- cs$portion_completely_home <- -->
<!--   cs$completely_home_device_count/cs$device_count -->
<!-- cs$portion_full_time_work <- -->
<!--   cs$full_time_work_behavior_devices/cs$device_count -->
<!-- cs$portion_part_time_work <- -->
<!--   cs$part_time_work_behavior_devices/cs$device_count -->

<!-- ## fign <- -->
<!-- ##   ggplot(aggregate(distance_traveled_from_home ~ date, -->
<!-- ##                    FUN=function(x) sum(!is.na(x)), data=cs)) + -->
<!-- ##   geom_line(aes(x=date, y=distance_traveled_from_home), size=1) + -->
<!-- ##   ylab("Number non-missing") + -->
<!-- ##   theme(legend.position = "none") + -->
<!-- ##   xlab("") + plottheme + -->
<!-- ##   ggtitle("Number of non-missing observations of Safegraph Variables") + -->
<!-- ##   colors() + xlim(startday, max(cs$date)) -->
<!-- ## fign -->

<!-- figm <- lapply(sv, quantileplot) -->
<!-- figm <- lapply(figm, function(x) x + ylab("")) -->
<!-- for (i in seq(1,length(figm), by=2)) { -->
<!--   grid.arrange(grobs=figm[i:(i+1)], nrow=2) -->
<!-- } -->

<!-- out <- residanalysis(sv, sdused) -->

<!-- kable(out[[2]], -->
<!--       caption="Variance explained by Google Mobility measures used in regressions") -->

<!-- out[[1]] -->
<!-- ``` -->

# Variables

```{r varlabels, cache=TRUE}
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
kable(varlabels)
```

# Case and death growth

```{r cdgrowth, cache=FALSE}
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
fig <- quantileplot("dlogdc", data=df) + ylab("") + xlim(c(as.Date("2020-03-01"),max(df$date+3)))
pdf(sprintf("%s/tex/tables_and_figures/casequantiles.pdf",rootdir), width=7, height=3.5)
print(fig)
dev.off()

fig <- quantileplot("dlogdd", data=df) + ylab("") + xlim(c(as.Date("2020-03-01"),max(df$date+3)))
pdf(sprintf("%s/tex/tables_and_figures/deathquantiles.pdf",rootdir), width=7, height=3.5)
print(fig)
dev.off()

```

### 2020-07-16 cases_and_policies/rmd/regressionsWithPlots.Rmd
Version submitted in early July 2020.
Compared to submitted version the following code changes have been made:

- Removed safegraph and unacast data (they can't be publicly shared)
- Minor update to policy data loading due to data format change
- Added explicit subsetting of data to June 2, 2020 and earlier (to match what we used in submitted paper).

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Tables and Figures"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
job         :
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_cononontained : true
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->

 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->

```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

This file creates the figures and tables in Chernozhukov, Kasahara and
Schrimpf (2020) "Causal Impact of Masks, Policies, Behavior on Early
Covid-19 Pandemic in the U.S."

# Data Preparation
```{r setup, cache=FALSE, warning=FALSE, include=FALSE}

library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
```

# Empirical Results

## Correlations

```{r corr, cache=TRUE}
cmat <-  cor(df[,c(bvars,pols)], use="pairwise.complete.obs")
scmat <-
  matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)))

rownames(scmat) <- getlabel(rownames(cmat))
colnames(scmat) <- getlabel(colnames(cmat))
scmat[upper.tri(cmat)] <- ""

ltbl <- knitr::kable(scmat, format="latex", booktabs=TRUE, escape=FALSE,
              align=(rep("c",ncol(scmat)))) %>%
  row_spec(0, angle = 90)
cat(ltbl, file=paste(rootdir,"tex/tables_and_figures/corr.tex",sep="/"))

knitr::kable(scmat)
```

## Test Smoothing

We rearrange and linearly interpolate constant test counts, but do not
apply any additional smoothing.

```{r testsmooth, cache=FALSE}
rearrange <- function(x, id) {
  for (i in unique(id)) {
    xi <- x[id==i]
    xi[!is.na(xi)] <- sort(xi[!is.na(xi)])
    x[id==i] <- xi
  }
  return(x)
}

dropnotincreasing <- function(x, id, t) {
  const <- (x <= panellag(x, id, t))
  const[is.na(const)] <- FALSE
  x[const & x>0] <- NA
  return(x)
}

lininterp <- function(x, id, t) {
  for (i in unique(id)) {
    xi <- x[id==i]
    ti <- t[id==i]
    xi[is.na(xi)] <- approx(ti[!is.na(xi)], xi[!is.na(xi)], ti[is.na(xi)])$y
    x[id==i] <- xi
  }
  return(x)
}

smoothtests <- function(x, id, t) {
  x[is.na(x)] <- 0
  return(lininterp(dropnotincreasing(rearrange(x,id),id,t),id,t))
}

df$neg.interp <- smoothtests(df$negative, df$state, df$date)
df$pos.interp <- df$cases #smoothtests(df$positive, df$state, df$date)
df$tot.interp <- df$neg.interp + df$pos.interp

L <- 7
df$tests <- paneldiff(df$tot.interp, df$state, df$date, lag=L)
df$testrate  <- df$tests/df$Population.2018*1000
df$testrate[df$testrate<0] <- NA
df$dlogtests <- paneldiff(log(sapply(df$testrate, function(x)
  max(x,exp(-1)))), df$state, df$date, lag=L)
df$testratedc <- df$ddeath/df$dcases*df$testrate
df$testratedc[df$dcases==0] <- 0

df$tests.ma <- paneldiff(panelma(df$neg.interp, df$state, df$date, 7) + df$cases,
                         df$state, df$date, lag=L)
df$testrate.ma  <- df$tests.ma/df$Population.2018*1000
df$testrate.ma[df$testrate.ma<0] <- NA
df$dlogtests.ma <- paneldiff(log(sapply(df$testrate.ma, function(x)
  max(x,exp(-1)))), df$state, df$date, lag=L)
df$testratedc.ma <- df$ddeath/df$dcases*df$testrate.ma
df$testratedc.ma[df$dcases==0] <- 0
```

## Cases

```{r caseregs, cache=TRUE, warning=FALSE, results='hide'}
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
#c("dlogdc", "logdc"))
#infovars2 <- list(c( "logdc"))
tvars <- "dlogtests"
xlist <-list("", "")
interactions <-list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0") #,"(testratedc ~ lag(testratedc,14))")

L <- 14
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
```

```{r casetables, cache=TRUE, warning=FALSE, results='asis'}
showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy)
savetextables(regs$pib[[1]], NULL, NULL, prefix="behavior")
savetextables(c(regs$pib[[1]][1:4], regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-NI")

savetextables(c(regs$pib[[1]][1:4]), NULL, NULL, prefix="behavior-cases-stateinfo")
savetextables(c(regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-cases-nationinfo")
savetextables(NULL, regs$pbiy, regs$piy, prefix="cases")
savetextables(NULL, regs$pbiy[c(1,3)], regs$piy[c(1,3)], prefix="cases-nosplit")
```

```{r casedieff, cache=TRUE, warning=FALSE, results='asis'}
for (use.national in c(TRUE,FALSE)) {
  sb <- ifelse(use.national, 2, 1)
  sy <- ifelse(use.national, 3, 1)
  pib <- lapply(regs$pib[[sb]][1:4], function(x) x$reg)
  pbiy <- regs$pbiy[[sy]]$reg
  piy <- regs$piy[[sy]]$reg

  S <- 999
  bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)

  addse <- function(di, bdi) {
    se <- di*NA
    for (i in 1:length(se)) {
      se[i] <- sd(sapply(bdi, function(b) b[i]))
    }

    tbl <- matrix(NA, nrow=2*nrow(di), ncol=ncol(di))
    for (i in 1:nrow(di)) {
      tbl[2*(i-1) + 1,] <- sapply(1:ncol(di), function(j) printstars(di[i,j], se[i,j]))
      tbl[2*i,] <- sprintf("(%.3f)", se[i,])
    }
    colnames(tbl) <- colnames(di)
    rownames(tbl) <- rep("", nrow(tbl))
    rownames(tbl)[seq(1,nrow(tbl),by=2)] <- relabel(rownames(di))
  return(tbl)
  }

  diall <- dieff_table(lapply(regs$pib[[sb]][1:4], function(x) x$reg$coef[,1]),
                       regs$pbiy[[sy]]$reg$coef[,1],
                       regs$piy[[sy]]$reg$coef[,1],
                       policies=c(pols, infovars[[sb]]), nsum=length(pols))

  bdiall <- lapply(1:S, function(i)
    dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
                bs[[5]][,i], bs[[6]][,i],
                policies=c(pols, infovars[[sb]]), nsum=length(pols))
    )

  tbl <- addse(diall, bdiall)

  #tbl <- tbl[-c(nrow(tbl), nrow(tbl)-1),]

  ltbl <- kable(tbl, digits=2, format="latex", booktabs=TRUE, escape=FALSE,
                linesep="",
                align=(rep("c",ncol(tbl))))  %>%
    column_spec(column=ncol(tbl)+1, border_left=TRUE)
  cat(ltbl, file=paste(rootdir,
                       sprintf("tex/tables_and_figures/dieff-cases-%s.tex",ifelse(use.national, "NI","SI")),
                       sep="/"))

  kable(tbl)
}
```

## Deaths (with cases as information)

```{r deathtables, cache=TRUE, warning=FALSE, results='asis'}
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
df$zero <- 0
L <- 21
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovars, "zero", xlist, ilist, iv, L=L)
showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-casesinfo")
```

## Deaths (with deaths as information)

```{r deathtablesd, cache=TRUE, warning=FALSE, results='asis'}
df$zero <- 0
infovarsd <- list(c("dlogdd", "logdd"),
                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
# infovarsd <- list(c("logdd")
L <- 21
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
regs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)

showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
savetextables(regs$pib[[1]], NULL, NULL, prefix="behavior-deathsinfo")
savetextables(c(regs$pib[[1]][1:4], regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-deathsinfo-NI")
savetextables(c(regs$pib[[1]][1:4]), NULL, NULL, prefix="behavior-deathsinfo-stateinfo")
savetextables(c(regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-deathsinfo-nationinfo")
savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-deathsinfo")
savetextables(NULL, regs$pbiy[c(1,3)], regs$piy[c(1,3)], prefix="deaths-deathsinfo-nosplit")

```

```{r dieff-deaths, cache=TRUE, results='asis', warning=FALSE}
for (use.national in c(TRUE,FALSE)) {
  sb <- ifelse(use.national, 2, 1)
  sy <- ifelse(use.national, 3, 1)
  pib <- lapply(regs$pib[[sb]][1:4], function(x) x$reg)
  pbiy <- regs$pbiy[[sy]]$reg
  piy <- regs$piy[[sy]]$reg

  S <- 999
  bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)

  diall <- dieff_table(lapply(regs$pib[[sb]][1:4], function(x) x$reg$coef[,1]),
                       regs$pbiy[[sy]]$reg$coef[,1],
                       regs$piy[[sy]]$reg$coef[,1],
                       policies=c(pols, infovarsd[[sb]]),
                       nsum=length(pols)
                       )

  bdiall <- lapply(1:S, function(i)
    dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
                bs[[5]][,i], bs[[6]][,i],
                policies=c(pols, infovarsd[[sb]]),
                nsum=length(pols)
                ))

  tbl <- addse(diall, bdiall)

  #tbl <- tbl[-c(nrow(tbl), nrow(tbl)-1),]

  ltbl <- kable(tbl, digits=2, format="latex", booktabs=TRUE, escape=FALSE,
                linesep="",
                align=(rep("c",ncol(tbl))))  %>%
    column_spec(column=ncol(tbl)+1, border_left=TRUE)
  cat(ltbl, file=paste(rootdir,
                       sprintf("tex/tables_and_figures/dieff-deaths-deathsinfo-%s.tex",ifelse(use.national, "NI","SI")),
                       sep="/"))
  kable(tbl)
}
```

<!-- ## $\log(\Delta D/\Delta C)$ -->

<!-- ```{r logdddc, cache=TRUE, warning=FALSE, results='asis'} -->
<!-- df$logdddc <- log(df$ddeath/panellag(df$cases, df$state, df$date, 7)) -->
<!-- df$logdddc[df$ddeath<=0] <- NA -->
<!-- df$logdddc[is.infinite(df$logdddc)] <- NA -->
<!-- df$logtestrate <- log(df$testrate) -->
<!-- df$logtestrate[df$testrate<=0] <- NA -->
<!-- regs <- mainregressions(df, "logdddc", pols, bvars, infovars, "lag(logtestrate,7)", xlist[1], ilist[1], list("0"), L=21) -->

<!-- showhtmltables(NULL, regs$pbiy, regs$piy); -->
<!-- ``` -->

# Figures

## Policy

The following plot illustrates the effect of mandatory masks for
employees.  The lighter line is the average across states that did not
have mask policy 14 days ago of case growth rate. The darker line is
the average across states that did have a mask policy 14 days
ago. (It's actually the fit from regressing dlogdc on date dummies
interacted with the 14 day lagged, 7 day moving average of a mask
policy indicator). Light points are the state observations without a
mask policy. Dark points are states with a mask policy 14 days ago.

```{r fig, cache=TRUE,  fig_width=10, fig_height=6, warning=FALSE}
pfigure <- function(df, pol, stlab=1, outcome="dlogdc", L=14) {
  df$p <- panellag(df[,pol],df$state, df$date, L)
  df$p <- ceiling(df$p)
  df$y <- df[,outcome]
  df$week <- week(df$date)
  m <- lm(y ~ as.factor(date) + as.factor(date):p, data=df)
  adf <- rbind(data.frame(date=m$xlevels$`as.factor(date)`, p=0),
               data.frame(date=m$xlevels$`as.factor(date)`, p=1))
  adf$week <- week(as.Date(adf$date))
  pred <- predict(m, newdata=adf, interval="confidence")
  adf$hat <- pred[,"fit"]
  adf$lo <- pred[,"lwr"]
  adf$hi <- pred[,"upr"]
  adf$date <- as.Date(adf$date)
  if (stlab==1) {
    pdf <- subset(df, df$p<=0.01)
    ldf <- subset(df, df$p>0.01)
  } else if (stlab==0) {
    pdf <- subset(df, df$p>=0.99)
    ldf <- subset(df, df$p<0.99)
  } else {
    pdf <- df
    ldf <- NULL #data.frame(date=NA,y=NA,ST=NA,p=NA)
  }
  clrs <- solarized_pal(2)(2)
  fig <- ggplot(df,aes(color=p)) +
    geom_point(data=pdf,aes(x=date, y=y, color=p),alpha=0.3, size=0.5) +
    xlim(c(as.Date("2020-04-10"), max(df$date))) +
    geom_line(data=adf, aes(x=date, y=hat, group=p), size=2) +
    figtheme + scale_color_gradient(low=clrs[1], high=clrs[2]) + #colors_grad(palette="Green-Gold") +
    theme(legend.position="none") +
    xlab("") +
    ylim(c(-1.2,1.2))
  if (!is.null(ldf))
    fig <- fig +
      geom_text(data=ldf,
                aes(x=date, y=y, label=ST, color=p), alpha=1.0, size=5)
  return(fig)
}
fig <- pfigure(df, "pmaskbus", stlab=-1) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("dlogdc")))

p <- "pmaskbus"
figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
print(fig)
dev.off()
fig
```

We now look at similar figures for other policies.

```{r polplots, cache=TRUE, fig_width=10, fig_height=6}
fig <- pfigure(df, "pk12", stlab=-1) +
  ggtitle(TeX(relabel("dlogdc given pk12"))) +
  xlim(c(as.Date("2020-03-15"), max(df$date))) +
  ylim(c(-1,3)) + ylab(TeX(relabel("dlogdc")))

p <- "pk12"
figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
print(fig)
dev.off()
fig

for (p in pols[3:length(pols)]) {
  fig <- pfigure(df, p, stlab=-1) +
    ggtitle(TeX(relabel(sprintf("dlogdc given %s",p)))) +
    xlim(c(as.Date("2020-04-01"), max(df$date))) +
    ylim(c(-1,1)) + ylab(TeX(relabel("dlogdc")))
  figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
  print(fig)
  dev.off()
}
```

# Death Growth Plots

```{r poldeathsplots, cache=TRUE, fig_width=10, fig_height=6, warning=FALSE}
fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=21) +
  ggtitle(TeX(relabel("dlogdd given pmaskbus"))) +
  ylab(TeX(relabel("dlogdd")))

p <- "pmaskbus"
figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
print(fig)
dev.off()
fig

# fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=14) +
#   ggtitle(TeX(relabel("dlogdd given pmaskbus"))) +
#   ylab(TeX(relabel("dlogdd")))
# p <- "pmaskbus"
# figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
# print(fig)
# dev.off()
# fig

fig <- pfigure(df, "pk12", stlab=-1, outcome="dlogdd", L=21) +
  ggtitle(TeX(relabel("dlogdd given pk12"))) +
  xlim(c(as.Date("2020-03-15"), max(df$date))) +
  ylim(c(-1,3)) + ylab(TeX(relabel("dlogdd")))

p <- "pk12"
figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
print(fig)
dev.off()

for (p in pols[3:length(pols)]) {
  fig <- pfigure(df, p, stlab=-1, outcome="dlogdd", L=21) +
    ggtitle(TeX(relabel(sprintf("dlogdd given %s",p)))) +
    xlim(c(as.Date("2020-04-01"), max(df$date))) +
    ylim(c(-1,1)) + ylab(TeX(relabel("dlogdd")))
  figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p), width=4, height=3)
  print(fig)
  dev.off()
}
```

```{r, results='asis', echo=FALSE}
if (!knitr::is_latex_output()) {
  cat("
[![](https://i.creativecommons.org/l/by-sa/4.0/88x31.png)](http://creativecommons.org/licenses/by-sa/4.0/)

This work is licensed under a [Creative Commons Attribution-ShareAlike
4.0 International License](http://creativecommons.org/licenses/by-sa/4.0/)
") }
```

### 2020-07-16 tex/covidEffectOfNPI-COVIDEconomics.tex
Version submitted in early July 2020.
Compared to submitted version the following code changes have been made:

- Removed safegraph and unacast data (they can't be publicly shared)
- Minor update to policy data loading due to data format change
- Added explicit subsetting of data to June 2, 2020 and earlier (to match what we used in submitted paper).

\documentclass[11pt,reqno,letter]{amsart}
%\pdfoutput=1
\usepackage{amsmath}

%----------------------------------------------------------------------------------------%

\usepackage{amsfonts}
\usepackage{amssymb}
\usepackage{amsthm}

\usepackage{graphicx}
\usepackage{hyperref}
\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
\usepackage{multirow}
\usepackage{verbatim}
\usepackage{float}
\usepackage{rotating}
%\usepackage{pxfonts}
%\usepackage{isomath}
%\usepackage{newpxtext}
%\usepackage{newpxmath}
\usepackage[dvipsnames]{xcolor}
\usepackage[normalem]{ulem}
\usepackage{booktabs}

% packages for rmd output
\usepackage{tabu}
\providecommand{\tightlist}{%
  \setlength{\itemsep}{0pt}\setlength{\parskip}{0pt}}
\usepackage{float}
\usepackage{booktabs}

\usepackage{subcaption}

\newtheorem{theorem}{Theorem}[section]
\newtheorem{conjecture}{Conjecture}[section]
\newtheorem{corollary}{Corollary}[section]
\newtheorem{lemma}{Lemma}[section]
\newtheorem{proposition}{Proposition}[section]
\theoremstyle{definition}
\newtheorem{assumption}{}[section]
\renewcommand{\theassumption}{A\arabic{assumption}}
\newtheorem{definition}{Definition}[section]
\newtheorem{step}{Step}[section]
\newtheorem{remark}{Comment}[section]
\newtheorem{example}{Example}[section]
\newtheorem*{example*}{Example}
\newtheorem{mpart}{Part}
\newtheorem{problem}{Problem}

\usepackage{calc,tikz,nicefrac}

\usepackage[section]{placeins}

\usetikzlibrary{positioning}
\geometry{letterpaper}
%\usepackag\Ep[parfill]{parskip}    % Activate to begin paragraphs
%with an empty line rather than an indent

\setlength{\parskip}{.4cm}
\geometry{centering}
\usepackage{mathtools}

\usepackage{bm}

\usepackage[authoryear]{natbib}
\usepackage{pdflscape}
\usepackage{afterpage}
\usepackage{capt-of}

\newcommand{\argmax}{\operatornamewithlimits{arg\,max}}
\newcommand{\argmin}{\operatornamewithlimits{arg\,min}}
\def\inprobLOW{\rightarrow_p}
\def\inprobHIGH{\,{\buildrel p \over \rightarrow}\,}
\def\inprob{\,{\inprobHIGH}\,}
\def\indist{\,{\buildrel d \over \rightarrow}\,}
\def\F{\mathbb{F}}
\newcommand{\gmatrix}[1]{\begin{pmatrix} {#1}_{11} & \cdots &
    {#1}_{1n} \\ \vdots & \ddots & \vdots \\ {#1}_{m1} & \cdots &
    {#1}_{mn} \end{pmatrix}}
\newcommand{\iprod}[2]{\left\langle {#1} , {#2} \right\rangle}
\newcommand{\norm}[1]{\left\Vert {#1} \right\Vert}
\newcommand{\abs}[1]{\left\vert {#1} \right\vert}
\renewcommand{\det}{\mathrm{det}}
\newcommand{\rank}{\mathrm{rank}}
\newcommand{\spn}{\mathrm{span}}
\newcommand{\row}{\mathrm{Row}}
\newcommand{\col}{\mathrm{Col}}
\renewcommand{\dim}{\mathrm{dim}}
\newcommand{\prefeq}{\succeq}
\newcommand{\pref}{\succ}
\newcommand{\seq}[1]{\{{#1}_n \}_{n=1}^\infty }
\renewcommand{\to}{{\rightarrow}}
\providecommand{\Infected}{{\mathcal{I}}}
\providecommand{\Recovered}{{R}}

\providecommand{\Er}{{\mathrm{E}}}
\providecommand{\Var}{{\mathrm{Var}}}
\providecommand{\set}[1]{\left\{#1\right\}}
\providecommand{\plim}{\operatornamewithlimits{plim}}
\newcommand\indep{\protect\mathpalette{\protect\independenT}{\perp}}
\def\independenT#1#2{\mathrel{\setbox0\hbox{$#1#2$}%
    \copy0\kern-\wd0\mkern4mu\box0}}

\newcommand{\ad}{\overset{\mathrm{a}}{\sim}}
\newcommand{\Z}{\mathbb{Z}}
\newcommand{\R}{\mathbb{R}}
%\renewcommand{\C}{\mathbb{C}}
\newcommand{\N}{\mathbb{N}}
\newcommand{\mS}{\mathcal{S}}
\newcommand{\mF}{\mathcal{F}}
\newcommand{\mE}{\mathcal{E}}
\newcommand{\mG}{\mathcal{G}}
\newcommand{\mL}{\mathcal{L}}
\renewcommand{\qed}{\hfill{\tiny \ensuremath{\blacksquare} }}%

\newcommand{\Ep}{{\mathrm{E}}}
\newcommand{\barEp}{\bar \Ep}
\newcommand{\En}{{\mathbb{E}_n}}
\renewcommand{\Pr}{{\mathrm{P}}}

%\newcommand{\mathbbeta}{\mbox{$\beta$}}
%\newcommand{\btheta}{\mbox{$\theta$}}
%\newcommand{\bdelta}{\mbox{$\delta$}}
\newcommand{\balpha}{\mbox{$\alpha$}}
\newcommand{\stheta}{\mbox{\scriptsize \theta}}

\newcommand{\mC}{\mathcal{C}}
\newcommand{\mU}{\mathcal{U}}
\newcommand{\mX}{\mathcal{X}}
\newcommand{\mJ}{\mathcal{J}}
\newcommand{\df}{\mathrm{df}}
\newcommand{\bP}{\mathbb{P}}
\newcommand{\bE}{\mathbb{E}}
\newcommand{\bEn}{\mathbb{E}_{n}}
\newcommand{\bG}{\mathbb{G}}

\newcommand{\eps}{\varepsilon}

\usepackage{neuralnetwork}

\def\bcolor{\color{ForestGreen}}
\def\pcolor{\color{blue}}
\def\icolor{\color{magenta}}
\def\wcolor{\color{gray}}
\def\ycolor{\color{red}}

\DeclareMathOperator{\dist}{dist} % The distance.

\begin{document}

\title[Causal Impact of Masks, Policies, Behavior]
{Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.}\thanks{We are grateful to Daron Acemoglu, V.V. Chari, Raj Chetty,  Christian Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green,  Ido Rosen, Konstantin Sonin, James Stock, and Ivan Werning for helpful comments. We also thank Chiyoung Ahn,
Joshua Catalano,  Jason Chau, Samuel Gyetvay, Sev Chenyu Hou,
Jordan Hutchings, and Dongxiao Zhang  for excellent research assistance.  All mistakes are our own. }

\author{Victor Chernozhukov}
\address{Department of Economics and Center for Statistics and Data Science, MIT,  MA 02139}
\email{vchern@mit.edu}
%\urladdr{www.mit.edu/$\sim$vcherm} % Delete if not wanted.
\author{Hiroyuki Kasahara}
\address{ Vancouver School of Economics, UBC, 6000 Iona Drive, Vancouver, BC.}
\email{hkasahar@mail.ubc.ca}

\author{Paul Schrimpf}
\address{ Vancouver School of Economics, UBC, 6000 Iona Drive, Vancouver, BC.}
\email{schrimpf@mail.ubc.ca}

\date{\today; \textit{First public} version posted to ArXiv:  May 28, 2020}

\begin{abstract}
This paper evaluates the dynamic impact of various policies adopted by US states on the growth rates of confirmed Covid-19  cases and deaths as well as social distancing behavior measured by Google Mobility Reports, where we take into consideration people's voluntarily behavioral response to new information of transmission risks. Our analysis finds that both policies and information on transmission risks are important determinants of Covid-19  cases and deaths and shows that a change in policies explains a large fraction of observed changes in social distancing behavior. Our counterfactual experiments suggest that nationally mandating face masks for employees on April 1st could have reduced the growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 17 to 55 percent less deaths nationally by the end of May, which roughly translates into 17 to 55 thousand saved lives.   Our estimates imply that removing non-essential business closures  (while maintaining school closures, restrictions on  movie theaters and restaurants)  could have led to -20 to 60 percent more cases and deaths by the end of May. We also find that, without stay-at-home orders, cases would have been larger by  25 to 170 percent, which implies that 0.5 to 3.4 million more Americans could have been infected if stay-at-home orders had not been implemented. Finally,  not having implemented any policies could have led to at least a 7 fold increase with an uninformative upper bound in cases (and deaths) by the end of May  in the US, with considerable uncertainty over the effects of school closures, which had little cross-sectional variation.%  (the potential range of effects could be even larger the effect of school closures on behavior is hard to identify separately from aggregate time effects).
\end{abstract}

\keywords{Covid-19, causal impact, masks, non-essential business, policies, behavior}

% JEL: C33, H75, I10, I18

\maketitle

%\tableofcontents
\thispagestyle{empty}

\pagebreak

\pagestyle{empty}

 \section{Introduction}

Accumulating evidence suggests that various policies in the US have reduced social interactions and slowed down the growth of Covid-19  infections.\footnote{
See
\citet{courtemanche2020}, \citet{hsiang2020}, \citet{pei2020},  \citet{abouk2020}, and \cite{wright2020}.} An important outstanding issue, however, is how much of the observed slow down in the spread is attributable to the effect of policies as opposed to a voluntarily change in people's behavior out of fear of being infected. This question is critical for evaluating the effectiveness of restrictive policies  in the US relative to an alternative policy of just providing recommendations and information such as the one adopted by Sweden.  More generally, understanding people's dynamic  behavioral response to policies and information  is indispensable for properly evaluating the effect of policies on the spread of Covid-19.

This paper  quantitatively assesses the impact of various policies adopted by US states on the spread of Covid-19, such as non-essential business closure and mandatory  face masks, paying particular attention to how people adjust their behavior in response to policies as well as new information on cases and deaths.

We present a conceptual framework that spells out the causal structure on how the Covid-19 spread  is dynamically determined by policies and human behavior. Our approach explicitly recognizes that policies not only directly affect the spread of Covid-19 (e.g., mask requirement) but also indirectly affect its spread by changing people's behavior (e.g., stay-at-home order). It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.  Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007}; see \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book} for  modern developments, especially in computer science and epidemiology), with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}. As such it naturally converses in all languages for causal inference.}
% It also allows us  to explicitly define counterfactual scenarios to properly evaluate the effect of alternative policies on the spread of Covid-19.

Guided by the causal model, our empirical analysis examines how the weekly growth rates of confirmed Covid-19 cases and deaths are determined by (the lags of) policies and behavior using US state-level data. To examine how policies and information affect people's behavior, we also regress social distancing measures  on policy and information variables. Our regression specification for case and death growths is explicitly guided by a SIR model although  our causal approach does not hinge on the validity of a SIR model.

As policy variables, we consider  mandatory face masks for employees in public businesses, stay-at-home orders (or shelter-in-place orders), closure of K-12 schools, closure of restaurants except take out, closure of movie theaters, and closure of non-essential businesses. Our behavior variables are four mobility measures that capture the intensity of visits to ``transit,'' ``grocery,'' ``retail,'' and ``workplaces''  from Google Mobility Reports. We take the lagged growth rate of cases and deaths and the log of lagged cases and deaths at both state-level and national-level as our measures of information on infection risks that affects people's  behavior. We also consider the growth rate of tests, month dummies, and state-level characteristics (e.g., population size and total area) as confounders that have to be controlled for in order to identify the causal relationship between policy/behavior and the growth rate of cases and deaths.

Our key findings from regression analysis are as follows.  We find
that both policies and information on past cases and deaths are important
determinants of people's social distancing behavior, where policy
effects explain more than $50\%$ of the observed decline in the four
behavior variables.\footnote{The behavior accounts for the other half. This is in line with theoretical study by \cite{gitmez2020} that investigates the role of private behavior and
negative external effects for individual decisions over policy compliance as well as information acquisition during pandemics.} Our estimates suggest that there are both large policy effects and large behavioral effects on the growth of cases and deaths. Except for mandatory masks, the effect of policies on cases and deaths is indirectly materialized through their impact on behavior; the effect of mandatory mask policy is direct without affecting behavior.

Using the estimated model, we evaluate the dynamic impact of the following counterfactual policies on Covid-19 cases and deaths: mandating face masks, allowing non-essential businesses to open, not implementing a stay-at-home order, and removing all policies.  The counterfactual experiments show a large impact of  those policies on  the number of cases and deaths.  They also highlight the importance of voluntary behavioral response to infection risks for evaluating the dynamic policy effects.

Figure \ref{fig:US-mask-dgrowth} shows that nationally  implementing mandatory
face masks  for employees in public businesses on April 1st would have reduced the growth rate of cases (top panel)
and that of deaths (bottom panel) by more than 10 percentage points in late April. This leads to reductions of $25$\% and $35$\% in reported cases and deaths, respectively,  by the end of May with a 90 percent confidence interval of $[10,45]$\% and $[17,55]$\%, which  roughly implies that as many as 17 to 55
  thousand lives could have been saved.\footnote{As of May 27, 2020,
   the US Centers for Disease Control and Prevention  reports 99,031
   deaths in the US.}
This finding is significant: given this potentially large  benefit of reducing the spread of Covid-19, mandating masks is an attractive policy instrument especially because it involves relatively little economic disruption. These estimates contribute to the ongoing efforts towards designing approaches to minimize risks from reopening \citep{stock2020b}.

Figure \ref{fig:US-nb-dgrowth} illustrates how allowing non-essential businesses to remain open could
have affected the growth of cases.  We estimate that
non-essential business closures have a small impact on growth rates,
with a 90\% confidence interval that includes both negative and
positive effects. When this effect on growth rates is converted to a
change in levels, the point estimates indicate that keeping
non-essential businesses open (other than movie theaters, gyms, and
keeping restaurants in the ``take-out" mode) could have increased
cases and  deaths by  $15\%$ (with a 90 percent confidence interval of
$-20\%$ to $60\%$).  These estimates contribute to the ongoing efforts of evaluating various reopening approaches.

In Figure \ref{fig:US-shelter-dgrowth}, we find that, without stay-at-home orders, the case
growth rate would have been nearly 10 percentage points higher in late April. No stay-at-home orders could have led to 80\% more cases by the start of June with a 90 precent confidence interval given by  $25\%$ to $170\%$. This implies that 0.5 to 3.4 million more Americans would have been infected without stay-at-home orders, providing suggestive evidence that reopening via removal of stay-at-home orders could lead to a substantial increase in cases and deaths.

In our counterfactual experiment of removing all policies, we find that the results are sensitive to whether  the number of past national cases/deaths is included  in a specification or not. This sensitivity arises because there is little variation across states in the timing of school closures.  This makes the effect of school closures difficult to identify. In Figure \ref{fig:US-nop-SI}, we show that in a specification that excludes past national cases (which allow for greater attribution of effects to school closures), the number of cases by the end of May could have increased 7-fold or more with a very large upper bound. On the other hand, as shown in Figure \ref{fig:US-nop}, under a specification with past national cases, our counterfactual experiment implies a 0 to 10 fold increase in cases by the end of May.  This highlights the uncertainty regarding the impact of all policies versus private behavioral responses to information. Evaluation of re-opening policies needs to be aware of this uncertainty.

%Figure \ref{fig:US-nop-dgrowth} shows how not having implementing any policies would have
%changed the national average growth rate of cases  over
%time. The effect of removing all policies on the growth rate of cases
% peaks in April at 0.1,  and then
%plateaus. The estimate implies a 0 to 10 fold increase in cases by the end of May.
%
%Even this wide confidence interval arguably understates uncertainty regarding the effect of all policies.
%There is little variation across states in the timing of school closures.  This makes the effect of school closures difficult to identify.
%To illustrate this sensitivity of conclusions regarding school closures, in Figure \ref{fig:US-nop-SI} we show that with a less flexible parameterization of time effects (which allow for greater attribution of effects to schooling), the number of cases by the end of May could have been even larger (the confidence band is wide, ranging from at least 7 fold increase to a very large upper bound).  This highlights the uncertainty regarding the impact of all policies versus private behavioral responses to information.
%Evaluation of re-opening policies needs to be aware of this uncertainty.

%, where about 80 \% of states closed schools within a day or two of March 15th.

\begin{figure}[ht]
  \caption{Effect  of nationally mandating masks for employees on April
    1st in the US \label{fig:US-mask-dgrowth}}
  \begin{minipage}{\linewidth}
    \centering
    \medskip
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_v1}
      &
        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_v1}
      &
        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
    \end{tabular}
  \end{minipage}
%    \begin{flushleft}
%      \footnotesize
%      In the left column, the dots are the average change in growth in
%      each state. The blue line is the average across states of the
%      change in growth. The shaded region is a point-wise 90\% confidence
%      interval. The right column shows the change in cases or deaths
%      relative to the baseline of actual policies.
%    \end{flushleft}
\end{figure}

\begin{figure}[ht]
 % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
  \caption{Effect  of   leaving non-essential businesses open on cases in the US  \label{fig:US-nb-dgrowth}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
%      \\
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths}
    \end{tabular}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
 % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
    \end{tabular}
  \end{minipage}
\end{figure}

%\begin{figure}[ht]
% \caption{Average effect  of removing
%   all policies on cases \label{fig:US-nop-dgrowth}}
%%  \caption{Effect of removing policies: national average \label{fig:US-nop}}
%  \begin{minipage}{\linewidth}
%    \centering
%    \begin{tabular}{cc}
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
%%      \\
%%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
%%      &
%%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
%    \end{tabular}
%  \end{minipage}
%\end{figure}

%\subsection{Related literature}
%
%
%... and why/how we are different.

A growing number of other papers have examined the link between non-pharmaceutical interventions and Covid-19  cases.\footnote{We refer the reader to \cite{avery2020policy} for a comprehensive review of  a larger body of work researching Covid-19; here we focus on few quintessential comparisons on our work with other works that we are aware of.} \cite{hsiang2020} estimate the effect of policies on the growth rate of cases using
data from the United States, China, Iran, Italy, France, and South
Korea. In the United States, they find that the combined effect of all policies they consider on the
growth rate is $-0.347$ $(0.061)$. \cite{courtemanche2020} use US county level data to analyze the effect
of interventions on case growth rates. They find that the combination
of policies they study reduced growth rates by 9.1 percentage points
16-20 days after implementation, out of which 5.9 percentage points are attributed to shelter in place orders. Both \cite{hsiang2020} and \cite{courtemanche2020} adopted a reduced-form approach to estimate the total policy effect on case growth without using any social distancing behavior measures. In contrast, our study highlights the role of behavioral response to policies and information.

Existing evidence for the impact of social distancing policies on behavior in the US is mixed.
\cite{abouk2020} employ a difference-in-differences methodology to  find that statewide stay-at-home orders  have  strong causal impacts on reducing social interactions. In contrast, using data from Google Mobility Reports,
\cite{maloney2020}  find that the increase in social distancing is
largely voluntary and driven by information.\footnote{Specifically, they find
that of the 60 percentage point drop in workplace intensity, 40
percentage points can be explained by changes in information as
proxied by case numbers, while roughly 8 percentage points can be
explained by policy changes.} Another study by  \cite{gupta2020} also found little evidence that stay-at-home mandates induced distancing by
using mobility measures from PlaceIQ and SafeGraph. Using data from SafeGraph, \cite{anderson2020} show that there has been substantial voluntary social distancing but also provide evidence that mandatory measures such as stay-at-home orders have  been effective at reducing the frequency of visits  outside of one’s home.
% Overall, these evidence suggests the importance of voluntary social distancing behavior when we evaluate the effect of polices.

 \cite{pei2020} use county-level observations of reported infections and deaths in conjunction with mobility  data from SafeGraph to estimate how effective reproductive numbers in major metropolitan areas change over time. They conduct simulation of implementing all policies 1-2 weeks earlier and found that it would have resulted in reducing the number of cases and deaths by more than half. However, their study does not  explicitly analyze how policies are related to
the effective reproduction numbers. % Our study  analyzes how policies  affects the reproduction numbers through their impact on people's behavior.

Epidemiologists use model simulations to predict how cases and deaths evolve for the purpose of policy recommendation. As reviewed by \cite{avery2020policy}, there exist substantial uncertainty about the values of key epidimiological parameters \citep[see also][]{atkeson2020b,stock2020}. Simulations are often done under strong assumptions about the impact of social distancing policies without connecting to the relevant data  \citep[e.g.,][]{ferguson2020}. Furthermore, simulated models do not take into account that people may limit their contact with other people in response to higher transmission risks.\footnote{See \cite{atkeson2020a} and \cite{stock2020} for the implications of the SIR model for Covid-19 in the US. \cite{NBERw27128} estimate a SIRD model in which time-varying reproduction numbers depend on the daily deaths to capture feedback from daily deaths
to future behavior and infections.}  When such a voluntary behavioral response is ignored, simulations would produce exponential spread of disease and would over-predict  cases and deaths. Our counterfactual experiments  illustrate the importance of this voluntary behavioral change.

Whether wearing  masks in public place should be mandatory or not has been one of the most contested policy issues with health authorities of different countries providing contradiction recommendations. Reviewing evidence, \cite{Greenhalghm2020}  recognize that there is no randomized controlled trial evidence for the effectiveness of face masks,  but they state ``indirect evidence exists to support the argument for the public wearing masks in the Covid-19 pandemic."\footnote{The virus remains viable in the air for several hours, for which surgical masks may be effective. Also, a substantial fraction of individual who are infected become infectious before showing symptom onset.}
\cite{howard2020} also review available medical evidence and conclude that ``mask wearing reduces the transmissibility per contact by reducing transmission of infected droplets in both laboratory and clinical contexts.''  The laboratory findings in \cite{hou2020} suggest that the nasal cavity may be the initial site of infection followed by aspiration to the lung, supporting the argument  ``for the widespread use of masks to prevent aersol, large droplet, and/or mechanical exposure to the nasal passages.''

Given the lack of experimental evidence on the effect of masks, conducting observational studies  is useful and important. To the best of our knowledge, our paper is the first  empirical study that shows the effectiveness of  mask mandates on reducing the spread of Covid-19 by analyzing the US state-level data. This finding corroborates and is complementary to the medical observational evidence in \cite{howard2020}. Analyzing mitigation measures in New York, Wuhan, and Italy, \cite{zhangr2020} conclude that mandatory face coverings substantially reduced infections. \cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms. Our finding
is also corroborated by a completely different causal methodology based on synthetic control 
using German data in \cite{Mitze2020}.\footnote{Our study was first released in ArXiv on May 28, 2020 whereas
 \cite{Mitze2020} was released at SSRN on June 8, 2020.}

Our empirical results contribute to informing the economic-epidemiological models that combine economic models with variants of SIR models to evaluate the efficiency of various economic policies aimed at gradual ``reopening" of various sectors of economy.\footnote{\cite{adda2016} analyzes the effect of policies reducing interpersonal contacts such as school closures or the closure of public transportation networks on the spread of  influenza, gastroenteritis, and chickenpox using high frequency data from France.} For example, the estimated effects of masks, stay-home mandates, and various other policies on behavior, and of behavior on infection can serve as useful inputs and validation checks in the calibrated macro, sectoral, and micro models (see, e.g., \cite{alvarez2020simple,baqaee2020reopening,NBERw27128,acemoglu2020multi,lsmith,nashSIR} and references therein). Furthermore, the causal framework developed in this paper could be applicable, with appropriate extensions, to the impact of policies on economic outcomes replacing health outcomes (see, e.g., \cite{chetty2020real,coibion2020labor}).

\section{The Causal Model for the Effect of Policies, Behavior, and Information on Growth of Infection}\label{sec:causal-model}

\subsection{The Causal Model and Its Structural Equation Form}
We introduce our approach through the Wright-style causal diagram shown in Figure \ref{Wright}.  The diagram describes how policies, behavior, and information interact together:
\begin{itemize}
\item The \textit{forward} health outcome,
$Y_{i,t+\ell}$, is determined last, after all other variables have been determined;
\item The  adopted policies, $P_{it}$,  affect health outcome $Y_{i,t+\ell}$ either directly, or indirectly by altering  human behavior  $B_{it}$;
\item  Information variables, $I_{it}$, such as lagged values of outcomes can affect human behavior and  policies, as well as  outcomes;
\item The confounding factors $W_{it}$, which vary across states and time, affect all other variables.
\end{itemize}
The index $i$ denotes observational unit, the state, and $t$ and $t+\ell$ denotes the time, where  $\ell$ represents the time lag  between infection and case confirmation or death.
\begin{figure}[ht]
\begin{center}
\begin {tikzpicture}[-latex, auto, node distance =1.5cm and 3cm, on grid, thick,
  empty/.style ={circle, top color=white, bottom color = white, draw, white, text=white , minimum width =1.25 cm},
  policy/.style ={circle, top color=white, bottom color = blue, draw, black, text=black , minimum width =1.25 cm},
    behavior/.style ={circle, top color=white, bottom color = ForestGreen, draw, black, text=black , minimum width =1.25 cm},
  observed/.style ={circle, top color=white, bottom color = magenta, draw, black, text=black , minimum width =1.25 cm} ,
   confounder/.style ={circle, top color=white, bottom color =  gray, draw, black, text=black , minimum width =1.25 cm} ,
outcome/.style ={circle, top color=white, bottom color = red, draw, black, text=black , minimum width =1.25 cm} ]

\node[policy]   (P) {\tiny $P_{it}$};
\node[empty] (E) [below=of P] {\tiny $I_{it}$};
\node[outcome]  (Y)[ right =of E] {\tiny $Y_{i,t+\ell}$};
\node[behavior] (B) [below =of E] {\tiny $B_{it}$};
\node[observed] (I) [left =of P] {\tiny $I_{it}$};
\node[confounder] (W) [left=of B] {\tiny  $W_{it}$};
\path[->] (P) edge (Y);
\path[->] (B) edge (Y);
\path[->] (P) edge (B);
\path[->] (I) edge (B);
\path[->] (I) edge (P);
\path[->] (W) edge (Y);
\path[->] (W) edge (B);
\path[->] (W) edge (P);
\path[->] (W) edge (I);
\path[->] (I) edge (Y);

\end{tikzpicture} \end{center}
\caption{S. \& P. Wright type causal path diagram for our model. }\label{Wright}
\end{figure}

Our main outcomes of interest are the growth rates in Covid-19 cases and deaths,  behavioral variables include proportion of time spent in transit or shopping and others, policy variables include stay-at-home orders and school and business closures, and the information variables include lagged values of outcome. We provide a detailed description of these variables and their timing in the next section.

The causal structure  allows for  the effect of the policy to be either direct or indirect -- through-behavior or through dynamics; and all of these effects are not mutually exclusive. The structure also allows for changes in behavior to be brought by change in policies and information. These are all realistic properties that we expect from the contextual knowledge of the problem. Policies such as closures of schools, non-essential business, and restaurants, alter and constrain behavior in strong ways.  In contrast, policies such as mandating employees to wear masks can potentially affect the Covid-19 transmission directly.  The information variables, such as recent growth in the number of cases, can cause people to spend more time at home, regardless of adopted state policies; these changes in
behavior in turn affect the transmission of Covid-19.

The causal ordering induced by this directed acyclical graph is determined by the following
timing sequence: % \textit{within a period}:%\footnote{Mathematically this corresponds to the factorization of the data likelihood, omitting indexes, $$f_{Y,P,B, I, W} = f_{Y\mid P, B,I, W} f_{B \mid P,I,W } f_{P \mid I, W} f_{I,W}.$$ }
\begin{itemize}
\item[(1)]  information and confounders get determined at $t$,
\item[(2)] policies are set in place, given information and confounders at $t$;
\item[(3)] behavior is realized, given policies, information, and confounders at $t$;
\item[(4)] outcomes get realized at $t+\ell$ given policies, behavior, information, and confounders.
\end{itemize}

The model also allows for direct dynamic effects of information variables on the outcome through autoregressive structures that capture persistence in growth patterns. As  highlighted below, realized outcomes may become new information for future periods, inducing dynamics over multiple periods.

Our quantitative model for causal structure in Figure \ref{Wright} is given by the following econometric structural equation model:\begin{equation} \label{eqn:SEM1} \tag{SEM}
  \begin{aligned}
   &  Y_{i,t+\ell} (b,p,\iota) &:=& {\bcolor \alpha ' b}  +  {\pcolor \pi 'p} +
    {\icolor \mu'\iota } + {\wcolor \delta_Y 'W_{it}} + \varepsilon^y_{it}, \\
   &  B_{it} (p,\iota) &:= & {\pcolor \beta'p } + {\icolor \gamma'\iota} +      {\wcolor\delta_B 'W_{it} } + \varepsilon^b_{it},
      \end{aligned}
 \end{equation}
which is a collection of functional relations with stochastic shocks, decomposed into observable part $\delta' W$ and unobservable part $\varepsilon$.
The terms $\varepsilon^y_{it}$ and  $\varepsilon^b_{it} $  are the centered stochastic shocks that obey the orthogonality restrictions posed below.

The policies can be modeled via a linear form as well,
\begin{equation}\label{eq:Policy} \tag{P}
 P_{it}   (\iota) :=  {\icolor\eta'\iota} + {\wcolor \delta_P' W_{it}} +   \varepsilon^p_{it},   \end{equation}
although  linearity is not critical.\footnote{Under some additional independence conditions, this
can be replaced by an arbitrary non-additive function $P_{it}(\iota) = p (\iota, W_{it},  \varepsilon^p_{it})$, such that the unconfoundedness condition stated in the next footnote holds.}

The orthogonality restrictions on the stochastic components are as follows:
\begin{equation}\label{eqn:SEM2} \tag{O}
\begin{aligned}
   & \varepsilon^y_{it} &  \perp &  \quad (\varepsilon^b_{it}, \varepsilon^p_{it}, {\wcolor W_{it}}, {\icolor I_{it}}), \\
&  \varepsilon^b_{it}  & \perp & \quad  (\varepsilon^p_{it}, {\wcolor W_{it}}, {\icolor I_{it}}), \\
&   \varepsilon^p_{it} &  \perp &  \quad ({\wcolor W_{it}}, {\icolor I_{it}}),
\end{aligned}
\end{equation}
where we say that $V \perp U$ if $\Ep VU = 0$. This is a standard way of representing restrictions on errors in structural equation modeling.\footnote{ An alternative useful
starting point is to impose the Rubin-Rosenbaum type unconfoudedness condition:
$$
Y_{i,t+\ell} (\cdot,\cdot,\cdot)  \indep (P_{it}, B_{it}, I_{it})  \mid W_{it}, \
B_{it} (\cdot,\cdot)  \indep (P_{it}, I_{it})  \mid W_{it}, \
 P_{it} (\cdot)  \indep  I_{it}  \mid W_{it},
$$
which imply, with treating stochastic errors as independent additive components, the orthogonal conditions stated above.
 }\footnote{The structural equations of this form are connected to triangular structural equation models, appearing
 in microeconometrics and macroeconometrics (SVARs), going back to the work of  \cite{strotz1960recursive}. .}

 The observed variables are generated by setting $\iota = I_{it}$ and propagating
the system from the last equation to the first:
%\begin{equation} \label{eqn:SEM2}
\[
  \begin{aligned}
& {\ycolor Y_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B_{it} } ,{\pcolor P_{it}}, {\icolor I_{it}}), \\
& {\bcolor B_{it} } & := &   B_{it}({\pcolor P_{it} } ,{\icolor I_{it}}), \\
& {\pcolor P_{it} }& := &  P_{it}({\icolor I_{it}}). \end{aligned}
\]
%\end{equation}

The system above together with  orthogonality restrictions (\ref{eqn:SEM2}) implies the following collection of stochastic equations for realized variables:
\begin{align}
   &  {\ycolor  Y_{i,t+\ell}}
    = {\bcolor\alpha ' B_{it}} + {\pcolor\pi 'P_{it}} + {\icolor\mu'I_{it}} + {\wcolor\delta_Y 'W_{it}}  + \varepsilon^y_{it},
    &  & \varepsilon^y_{it} \perp {\bcolor B_{it}}, {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R1} \tag{BPI$\to$Y} \\
    &  {\bcolor B_{it}}
     =  {\pcolor \beta' P_{it}} + {\icolor \gamma'I_{it}} +  {\wcolor \delta_B' W_{it}} + \varepsilon^b_{it},
   & & \varepsilon^b_{it} \perp {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}}  \label{eq:R2} \tag{PI$\to$B}  \\
    & {\pcolor P_{it}}
    =  {\icolor\eta'I_{it}} + {\wcolor \delta_P' W_{it}} +   \varepsilon^p_{it},   & & \varepsilon^p_{it} \perp   {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R3}  \tag{I$\to$P}
       \end{align}
and
\begin{equation}
   {\ycolor  Y_{i,t+\ell}}
   = ( {\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'} )
    {\pcolor P_{it}} + ( {\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})
    {\icolor I_{it} }+ {\wcolor \bar{\delta} '}{\wcolor W_{it}}  + {\bar \varepsilon}_{it},  \quad   {\bar \varepsilon}_{it} \perp
  {\pcolor P_{it}},  {\icolor I_{it}}, {\wcolor W_{it}}.  \label{eq:R4} \tag{PI$\to$Y}
\end{equation}
These equations form the basis of our empirical analysis.

As discussed below, the information variable includes case growth. Therefore,  an orthogonality restriction  $ \varepsilon^y_{it} \perp   {\pcolor P_{it}}$ holds if  the government does not have knowledge on future case growth beyond what is predicted by today's case growth, policies,  behavior, and  confounders; even when the government has some knowledge on $\varepsilon^y_{it}$, the orthogonality restriction may hold if there is a time lag for the government to implement its policies based on $\varepsilon^y_{it}$.

The orthogonality condition in (\ref{eq:R4}) is weaker than the orthogonality conditions in  (\ref{eq:R1})-(\ref{eq:R2}) in that the former is implied by the latter but not vice versa.
The system over-identifies the regression coefficients because  $({\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'})$ and $({\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})$ in  (\ref{eq:R4}) can be also identified from  ${\bcolor\alpha ' }$,  $ {\pcolor\pi '}$, ${\icolor\mu'}$,   ${\pcolor \beta' }$, and ${\icolor \gamma'}$ in (\ref{eq:R1})-(\ref{eq:R2}).  Comparing the estimates of $({\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'})$ and $({\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})$ from  (\ref{eq:R4})  with  those implied by  the estimates of ${\bcolor\alpha ' }$,  $ {\pcolor\pi '}$, ${\icolor\mu'}$,   ${\pcolor \beta' }$, and ${\icolor \gamma'}$ from   (\ref{eq:R1})-(\ref{eq:R2}) provides a useful specification test.

\subsection*{Identification and Parameter Estimation}

The orthogonality equations imply that these are all projection equations, and the parameters of the SEM are identified by the parameters of these regression equation, provided the latter are identified by sufficient joint variation of these variables across states and time.

The last point can be stated formally as follows. Consider the previous system of equations, after partialling out the confounders:
\begin{equation}\label{eqn:SEM-PO}
 \begin{aligned}
   & {\ycolor { \tilde Y_{i,t+\ell}}}  & =& {\bcolor {\alpha ' \tilde B_{it}}} +{\pcolor {\pi '\tilde P_{it}}} +{\icolor \mu'\tilde I_{it} } + \varepsilon^y_{it},&\varepsilon^y_{it} &\perp {\tilde B_{it}}, {\tilde  P_{it}}, { \tilde I_{it}},  \\
   & {\bcolor { \tilde B_{it}}} & = &{\pcolor { \beta' \tilde P_{it}} } + {\icolor { \gamma' \tilde I_{it}} } + \varepsilon^b_{it},
   &\quad \varepsilon^b_{it} &\perp { \tilde P_{it}}, { \tilde I_{it}},  \\
   & {\pcolor { \tilde P_{it}} } &= &  {\icolor {\eta' \tilde I_{it}} } +   \varepsilon^p_{it}, &\quad \varepsilon^p_{it} &\perp {\tilde I_{it}}  \\
  \end{aligned}
\end{equation}
where $ \tilde V_{it} = V_{it}   -     {\wcolor W_{it}'} \Ep[{\wcolor W_{it}W_{it}'}]^{-} \Ep[{\wcolor W_{it}} V_{it}]$ denotes
the residual after removing the orthogonal projection of $V_{it}$ on ${\wcolor W_{it}}$. The residualization is a linear operator, implying that (\ref{eqn:SEM-PO}) follows immediately from the above. The parameters of (\ref{eqn:SEM-PO})  are identified as projection coefficients in these equations, provided that residualized vectors appearing in each of the equations have non-singular variance, that is
 \begin{equation}
 \Var ( {\pcolor \tilde P_{it}'} ,{\bcolor \tilde B_{it}'},{\icolor \tilde I_{it}'})>0,
 \ \Var ({\pcolor \tilde P_{it}'}, {\icolor \tilde I_{it}'})> 0 , \ \text{ and  }  \Var ( {\icolor \tilde I_{it}'}) >0.
 \end{equation}

Our main estimation method is the standard correlated random effects estimator, where the random effects
are parameterized as functions of observable characteristic, ${\wcolor W_{it}}$, which include both state-level and time random effects.  The state-level random effects
are modeled as a function of state level characteristics, and the time random effects are modeled
by including month dummies and their interactions with state level characteristics.  The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$
are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.

A secondary estimation method is the fixed effects estimator, where ${\wcolor W_{it}}$ includes
latent (unobserved) state level effects ${\wcolor W_i}$ and  and time level effects ${\wcolor W_t}$,
which must be estimated from the data.  This approach is much more demanding of the data and relies on long cross-sectional
and time histories. When histories are relatively short,  large biases emerge and
they need to be removed using debiasing methods. In our context, debiasing
  changes the estimates substantially, often changing the sign of coefficients.\footnote{This is a pre-cautionary
message that may be useful for other researchers using fixed effects estimators in the context
of Covid-19 analysis. We recommend using debiased fixed effects estimators, see e.g., \cite{chen2019mastering}
for expository treatment. }
However, we find the debiased fixed effect estimates are
qualitatively and quantitatively similar to the correlated random effects estimates.
Given this finding, we chose to focus on the latter, as it is a more standard and familiar method, and report the former
estimates in the supplementary materials for this paper.\footnote{The similarity of the debiased fixed
effects and correlated random effects served as a useful specification check. Moreover, using the fixed
effects estimators only yielded minor gains in predictive performances, as judging by the adjusted
$R^2$'s, providing another useful specification check.}

\subsection{Information Structures and Induced Dynamics}  We consider three examples of information
structures: Information variable is  a function of time:
$${\icolor I_{it}} = g(t);$$
 Information variable is lagged value of outcome:
$$
{\icolor I_{it}} =  {\ycolor Y_{it}};
 $$
 and finally:
 \begin{itemize}
 \item[(\textsc{I})] Information variables include time, lagged and integrated values of outcome:
 $$
{\icolor I_{it}} = \left (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{t/\ell}
{\ycolor Y_{i,t - \ell m}} \right) ',
 $$
 with the convention that ${\ycolor Y_{it}} = 0$  for $t \leq 0$.
\end{itemize}
The first information structure captures the  basic idea that, as individuals discover more information about covid over time, they adapt to safer modes of  behavior (stay-at-home, wear masks, wash hands). {Under this structure, information is common across states and exogenously evolves over time, independent of the number of cases.}
The second  structure  arises from considering
autoregressive components {and captures people's behavioral response to information on cases in the state they reside}.  {Specifically,} we model persistence in growth rates, ${\ycolor Y_{i,t+\ell}}$, through an AR(1) model, which leads to ${\icolor I_{it}}= {\ycolor Y_{it}}$. This provides useful \textit{local}, state-specific, information about the forward growth rate and people may adjust their behavior to safer modes when they see a high value. We model this adjustment via the term ${\icolor \gamma'I_{t}}$ in the behavior equation. The {third information} structure
is the combination of the first two structures plus an additional  term representing the log of the total number of new cases in the state. We use this information structure in our empirical specification. In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases.
The last element of the information set can be thought of as a local stochastic trend in cases.

All of these examples fold into a specification of the form:
\begin{equation}\label{eq:I}\tag{I}
 {\icolor I_{it} } :=   I_{it}({\icolor I_{i,t-\ell}}, {\ycolor Y_{it}}, t),  \quad t =1,..., T,
 \end{equation}
 with the initialization ${\icolor I_{i 0}} = 0$ and ${\ycolor Y_{i0}} = 0$.\footnote{This initialization is appropriate
 in our context for analyzing pandemics from the very beginning, but other initializations could be appropriate in other contexts. The  lagged values of behavior variable may be also included in the information set.
 }

With any structure of this form, realized outcomes may become new information for future periods, inducing a dynamical system over multiple periods. We show the resulting dynamical system in a causal diagram of Figure \ref{fig:DynS}.  Specification
of this system is useful for studying delayed effects of policies and behaviors and in considering the counterfactual policy analysis.

\begin{figure}[ht]
\begin{center}
\begin {tikzpicture}[-latex, auto, node distance =3cm and 3cm, on grid, thick,
  empty/.style ={circle, top color=white, bottom color = white, draw, white, text=white , minimum width =1.25 cm},
  policy/.style ={circle, top color=white, bottom color = blue, draw, black, text=black , minimum width =1.25 cm},
    behavior/.style ={circle, top color=white, bottom color = ForestGreen, draw, black, text=black , minimum width =1.25 cm},
  observed/.style ={circle, top color=white, bottom color = magenta, draw, black, text=black , minimum width =1.25 cm} ,
   confounder/.style ={circle, top color=white, bottom color =  gray, draw, black, text=black , minimum width =1.25 cm} ,
outcome/.style ={circle, top color=white, bottom color = red, draw, black, text=black , minimum width =1.25 cm},
myarrow/.style={-Stealth}]

\node[observed] (I)  {\tiny $I_{i,t-\ell}$};
\node[outcome]  (Y)[below=of I ]{\tiny $Y_{it}$};
\node[observed] (In)  [right=of I ]{\tiny $I_{it}$};
\node[outcome]  (Yn)[below=of In ]{\tiny $Y_{i,t+\ell}$};
\node[observed] (Inn)  [right=of In ]{\tiny $I_{i,t+\ell}$};
\node[outcome]  (Ynn)[below=of Inn ]{\tiny $Y_{i,t+2\ell}$};
 \draw [->] (I) -- node[sloped,font=\small,below] {\tiny SEM(t-$\ell$)} (Y);
 \draw [->] (I) --  (In);
 \draw [->] (In) -- node[sloped,font=\small,below] {\tiny SEM(t)} (Yn);
\draw [->] (Y) --  (In);
\draw [->] (In) --  (Inn);
 \draw [->] (Inn) -- node[sloped,font=\small,below] {\tiny SEM(t+$\ell$) } (Ynn);
\draw [->] (Yn) --  (Inn);

\end{tikzpicture} \end{center}
\caption{ Dynamic System Induced by Information Structure and SEM}\label{fig:DynS}
\end{figure}

%
%\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel}
%Letting $C_{it}$ denotes cumulative number of  confirmed cases in state $i$ at time $t$, our outcome
%\begin{equation} \label{eq:y}
%{\ycolor Y_{it}} :=\Delta \log(\Delta C_{it}):= \log( \Delta C_{it} ) -
%\log( \Delta C_{i,t-7})
%\end{equation}
%approximates the weekly growth rate in new cases from $t-7$ to $t$.\footnote{We may show that $ \log( \Delta C_{it} ) -
%\log( \Delta C_{i,t-7})$ approximates the average
%growth rate of cases from $t-7$ to $t$.} Here $\Delta$ denotes the differencing operator over 7 days from $t$ to $t-7$, so that $\Delta C_{it}:=C_{it}-C_{i,t-7}$ is the number of new confirmed cases in the past 7 days.
%
%
%We chose this metric as this is the key metric for policy makers deciding when to relax Covid mitigation policies.  The U.S. government's guidelines for state reopening
%recommend that states display a
%``downward trajectory of documented cases within a 14-day period''
%\citep{whitehouse2020}. A negative value of
%$Y_{it}$ is an indication of meeting this
%criteria for reopening. By focusing on weekly cases  rather than daily cases, we smooth idiosyncratic daily fluctuations as well as periodic fluctuations associated with days of the week.
%
%Our measurement equation for estimating equations (\ref{eq:R1}) and (\ref{eq:R4}) will take the form:
%\begin{align}
%{\ycolor Y_{it}}  =      \theta' X_{it}  + \delta_T \Delta   \log(T)_{it}  +
% \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  + \epsilon_{it},
% \label{eq:M} \tag{M}
%\end{align}
%where $i$ is state, $t$ is day, $C_{it}$ is cumulative confirmed
%cases, $D_{it}$ is deaths, $T_{it}$ is the number of tests over 7 days, $\Delta$ is
%a 7-days differencing operator, $\epsilon_{it}$ is an unobserved error term,
%where $X_{it}$ collects other behavioral, policy, and confounding variables, depending
%on whether we estimate (\ref{eq:R1}) or (\ref{eq:R4}).  Here $$\Delta   \log(T)_{it} \ \text {and } \ \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}$$ are the key confounding variables,
%derived from considering the SIR model below. We describe other confounders in the empirical section.
%
% Our main estimating equation (M) is motivated by a variant of a SIR
%model, where we incorporate a
%delay between infection and death instead of a constant death rate,
%and we add confirmed cases to the model. Let $S$, $I$, $\mathcal{R}$, and $D$ denote the number of susceptible,
%infected, recovered, and dead individuals in a given state or
%province. Each of these variables are a function of time. We model
%them as evolving as
%\begin{align}
%  \dot{S}(t) & = -\frac{S(t)}{N} \beta(t) I(t) \label{eq:s} \\
%  \dot{I}(t) & = \frac{S(t)}{N} \beta(t) I(t) - \frac{S(t-\ell)}{N}
%               \beta(t-\ell) I(t-\ell) \label{eq:i} \\
%  \dot{\mathcal{R}}(t) & = (1-\pi) \frac{S(t-\ell)}{N} \beta(t-\ell)
%                         I(t-\ell) \label{eq:r} \\
%  \dot{D}(t) & = \pi \frac{S(t-\ell)}{N} \beta(t-\ell) I(t-\ell) \label{eq:d}
%\end{align}
%where $N$ is the population, $\beta(t)$ is the rate of infection
%spread, $\ell$ is the duration between infection and
%death, and $\pi$ is the
%probability of death conditional on infection.\footnote{The model can easily be extended to allow a
%  non-deterministic disease duration. This leaves our main estimating
%  equation (\ref{eq:c2}) unchanged, as long as the distribution of
%  duration between infection and death is equal to the distribution of
%  duration between infection and recovery.}
%
%Confirmed cases, $C(t)$, evolve as
%\begin{equation}
%  \dot{C}(t) = \tau(t) I(t), \label{eq:c}
%\end{equation}
%where $\tau(t)$ is the rate that infections are detected.
%
%Our goal is to examine how the rate of infection $\beta(t)$ varies with observed policies
%and measures of social distancing behavior. A key challenge is that we only
%observed $C(t)$ and $D(t)$, but not $I(t)$. The unobserved $I(t)$ can
%be eliminated by differentiating (\ref{eq:c}) and using (\ref{eq:i}) and (\ref{eq:d}) as
%\begin{align}
%%  \ddot{C}(t) & = \dot{\tau}(t) + \dot{I}(t) \notag \\
%%  & = \dot{\tau}(t)I(t) + \frac{S(t)}{N} \beta(t) I(t) -
%%    \frac{S(t-\ell)}{N} \beta(t-\ell) I(t-\ell)  \notag \\
%%  & = \dot{\tau}(t) I(t) + \frac{S(t)}{N} \beta(t) I(t) -
%%    \frac{1}{\pi} \dot{D}(t) \notag \\
%  \frac{\ddot{C}(t)}{\dot{C}(t)}
%              & =
%                \frac{S(t)}{N} \beta(t)  + \frac{\dot{\tau}(t)}{\tau(t)}  -
%                \frac{1}{\pi}
%                \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}. \label{eq:c2}
%\end{align}
%We consider a discrete-time analogue of equation (\ref{eq:c2}) to motivate our empirical
%specification by relating the detection rate $\tau(t)$  to the number of tests $T_{it}$ while specifying $\frac{S(t)}{N}\beta(t)$ as a linear function of variables $X_{it}$.
%This results in
%\begin{align}
%  \underbracket{Y_{it}}_{\frac{\ddot{C}(t)}{\dot{C}(t)}}
%  =
%      \underbracket{X_{it}' \theta + \epsilon_{it}}_{\frac{S(t)}{N}\beta(t)} +
%       & \underbracket{\delta_T \Delta
%      \log(T)_{it}}_{\frac{\dot{\tau}(t)}{\tau(t)} } +
%      \underbracket{\delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  }_{\frac{1}{\pi}
%      \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}}, \nonumber % \label{eq:reg2}
%\end{align}
%which is equation (\ref{eq:M}), where $X_{it}$ captures a vector of variables related to $\beta(t)$.
%
%\begin{quote}
%\textsc{Structural Interpretation}. In the early pandemics, when
%the number of susceptibles is approximately the same as the entire population, i.e. $S_{it}/N_{it} \approx 1$, the component $X_{it}' \theta$
%is the projection of infection rate $ \beta_i(t)$ on   $X_{it}$ (policy, behavioral, information, and confounders other than testing rate), provided
%the stochastic component $\epsilon_{it}$ is orthogonal
%to $X_{it}$ and the testing variables:
%$$
%\beta_i(t)S_{it}/N_{it}  = X_{it}' \theta + \epsilon_{it}, \quad \epsilon_{it} \perp X_{it}.
%$$
%\end{quote}

\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel}
Letting $C_{it}$ denote the cumulative number of  confirmed cases in state $i$ at time $t$, our outcome
\begin{equation} \label{eq:y}
{\ycolor Y_{it}} =\Delta \log(\Delta C_{it}):= \log( \Delta C_{it} ) -
\log( \Delta C_{i,t-7})
\end{equation}
approximates the weekly growth rate in new cases from $t-7$ to $t$.\footnote{We may show that $ \log( \Delta C_{it} ) -
\log( \Delta C_{i,t-7})$ approximates the average
growth rate of cases from $t-7$ to $t$.} Here $\Delta$ denotes the differencing operator over 7 days from $t$ to $t-7$, so that $\Delta C_{it}:=C_{it}-C_{i,t-7}$ is the number of new confirmed cases in the past 7 days.

We chose this metric as this is the key metric for policy makers deciding when to relax Covid mitigation policies.  The U.S. government's guidelines for state reopening
recommend that states display a
``downward trajectory of documented cases within a 14-day period''
\citep{whitehouse2020}. A negative value of
$Y_{it}$ is an indication of meeting this
criteria for reopening. By focusing on weekly cases  rather than daily cases, we smooth idiosyncratic daily fluctuations as well as periodic fluctuations associated with days of the week.

Our measurement equation for estimating equations (\ref{eq:R1}) and (\ref{eq:R4}) will take the form:
\begin{align}
{\ycolor \Delta \log(\Delta C_{it})}  =    X_{i,t-14} '   \theta  -\gamma +  \delta_T \Delta   \log(T_{it})  + \epsilon_{it},
% \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  + \epsilon_{it},
 \label{eq:M} \tag{M-C}
\end{align}
where $i$ is state, $t$ is day, $C_{it}$ is cumulative confirmed
cases, $T_{it}$ is the number of tests over 7 days, $\Delta$ is
a 7-days differencing operator, $\epsilon_{it}$ is an unobserved error term.
 $X_{i,t-14}$  collects other behavioral, policy, and confounding variables, depending
on whether we estimate (\ref{eq:R1}) or (\ref{eq:R4}), where the lag of $14$ days captures the time lag between infection and confirmed case (see the Appendix \ref{sec:timing}). %for available evidence on a delay between infection and confirmation of cases or deaths.
   Here
$$\Delta   \log(T_{it} ):=  \log(T_{it}) - \log(T_{i,t-7})  $$ %\ \text {and } \ \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}$$
is the key confounding variable,
derived from considering the SIR model below. We describe other confounders in the empirical section.

% Our main estimating equation (M) is motivated by a variant of a SIR
%model, where we incorporate a
%delay between infection and death instead of a constant death rate,
%and we add confirmed cases to the model.

 Our main estimating equation (\ref{eq:M}) is motivated by a variant of SIR
model, where we add confirmed cases and infection detection via testing.
Let $S$, $\Infected$, $\Recovered$, and $D$ denote the number of susceptible,
infected, recovered, and dead individuals in a given state. Each of these variables are a function of time. We model
them as evolving as
\begin{align}
  \dot{S}(t) & = -\frac{S(t)}{N} \beta(t) \Infected(t) \label{eq:s} \\
  \dot{\Infected}(t) & = \frac{S(t)}{N} \beta(t) \Infected(t) - \gamma  \Infected(t) \label{eq:i}\\
  % \frac{S(t-\ell)}{N} \beta(t-\ell)  \Infected(t-\ell) \label{eq:i} \\
  \dot{\Recovered}(t) & = (1-\kappa) \gamma  \Infected(t) \label{eq:r}\\ %  \frac{S(t-\ell)}{N} \beta(t-\ell)    \Infected(t-\ell) \label{eq:r} \\
  \dot{D}(t) & = \kappa \gamma \Infected(t) %  \frac{S(t-\ell)}{N} \beta(t-\ell)   \Infected(t-\ell)
  \label{eq:d}
\end{align}
where $N$ is the population, $\beta(t)$ is the rate of infection
spread, $\gamma$ is the rate of recovery or death, and $\kappa$ is the
probability of death conditional on infection.
%\footnote{The model can easily be extended to allow a
%  non-deterministic disease duration. This leaves our main estimating
%  equation (\ref{eq:c2}) unchanged, as long as the distribution of
%  duration between infection and death is equal to the distribution of
%  duration between infection and recovery.}

Confirmed cases, $C(t)$, evolve as
\begin{equation}
  \dot{C}(t) = \tau(t) \Infected(t), \label{eq:c}
\end{equation}
where $\tau(t)$ is the rate that infections are detected.

Our goal is to examine how the rate of infection $\beta(t)$ varies with observed policies
and measures of social distancing behavior. A key challenge is that we only
observed $C(t)$ and $D(t)$, but not $\Infected(t)$. The unobserved $\Infected(t)$ can
be eliminated by differentiating (\ref{eq:c}) and using (\ref{eq:i})  as
\begin{align}
%  \ddot{C}(t) & = \dot{\tau}(t) + \dot{\Infected}(t) \notag \\
%  & = \dot{\tau}(t)\Infected(t) + \frac{S(t)}{N} \beta(t) \Infected(t) -
%    \frac{S(t-\ell)}{N} \beta(t-\ell) \Infected(t-\ell)  \notag \\
%  & = \dot{\tau}(t) \Infected(t) + \frac{S(t)}{N} \beta(t) \Infected(t) -
%    \frac{1}{\kappa} \dot{D}(t) \notag \\
  \frac{\ddot{C}(t)}{\dot{C}(t)}
              & =
                \frac{S(t)}{N} \beta(t) -\gamma  + \frac{\dot{\tau}(t)}{\tau(t)}. \label{eq:c2}
%                 -
%                \frac{1}{\kappa}
%                \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}. \label{eq:c2}
\end{align}
We consider a discrete-time analogue of equation (\ref{eq:c2}) to motivate our empirical
specification by relating the detection rate $\tau(t)$  to the number of tests $T_{it}$ while specifying $\frac{S(t)}{N}\beta(t)$ as a linear function of variables $X_{i,t-14}$.
This results in
\begin{align}
  \underbracket{\Delta \log(\Delta C_{it})}_{\frac{\ddot{C}(t)}{\dot{C}(t)}}
  =
      \underbracket{X_{i,t-14}' \theta + \epsilon_{it}}_{\frac{S(t)}{N}\beta(t) -\gamma}
       +
       & \underbracket{\delta_T \Delta
      \log(T)_{it}}_{\frac{\dot{\tau}(t)}{\tau(t)} } \nonumber
      %+
%      \underbracket{\delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  }_{-\frac{1}{\kappa}
%      \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}}, \nonumber % \label{eq:reg2}
\end{align}
which is equation (\ref{eq:M}), where $X_{i,t-14}$ captures a vector of variables related to $\beta(t)$.

\begin{quote}
\textsc{Structural Interpretation}. Early in the pandemic, when
the number of susceptibles is approximately the same as the entire population, i.e. $S_{it}/N_{it} \approx 1$, the component $X_{i,t-14}' \theta$
is the projection of infection rate $ \beta_i(t)$ on   $X_{i,t-14}$ (policy, behavioral, information, and confounders other than testing rate), provided
the stochastic component $\epsilon_{it}$ is orthogonal
to $X_{i,t-14}$ and the testing variables:
$$
\beta_i(t)S_{it}/N_{it}  - \gamma = X_{i,t-14}' \theta + \epsilon_{it}, \quad \epsilon_{it} \perp X_{i,t-14}.
$$
\end{quote}

\subsection{Growth Rate in Deaths as Outcome}\label{sec:sirmodel-death}
By differentiating (\ref{eq:d}) and (\ref{eq:c}) with respect to $t$ and using (\ref{eq:c2}), we obtain
\begin{align}
\frac{\ddot{D}(t) }{\dot D(t)}& = \frac{\ddot{C}(t) }{\dot C(t)}  - \frac{\dot{\tau}(t) }{ \tau(t)}    =  \frac{S(t)}{N}\beta(t)  -   \gamma.\label{eq:d2}
\end{align}
Our measurement equation for the growth rate of deaths is based on equation (\ref{eq:d2}) but   account for a $21$ day lag between infection and death as
\begin{align}
{\ycolor \Delta \log(\Delta D_{it})}  = X_{i,t-21}' \theta + \epsilon_{it},\label{eq:M-D} \tag{M-D}
\end{align}
where
\begin{equation} \label{eq:y-d}
 \Delta \log(\Delta D_{it}):= \log( \Delta D_{it} ) -
\log( \Delta D_{i,t-7})
\end{equation}
approximates the weekly growth rate in deaths from $t-7$ to $t$ in state $i$. % In our baseline specification, we consider 7 days lag between case confirmation and death by setting   $\ell=7$ in (\ref{eq:M-D}).  %The appendix provides additional results under alternative assumptions including non-deterministic disease durations.

\section{Decomposition and Counterfactual Policy Analysis}

\subsection{Total Change Decomposition}
Given the SEM formulation above, we can carry out the following decomposition analysis, after removing the effects of confounders. For example,
we can decompose the total change $\Ep \tilde Y_{i,t+\ell} - \Ep \tilde Y_{io}$ in the expected outcome, measured at two different time points $t+\ell$ and $o$ into the sum of three components:
\begin{equation} \label{eqn:DC}
  \begin{aligned}
\underbracket{\Ep \tilde Y_{i,t+\ell} - \Ep \tilde Y_{io}}_{\text{Total Change}}& =
{\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left( \Ep \tilde P_{it} - \Ep \tilde P_{io} \right)}}_{\text{Policy Effect via Behavior}}} \\
& +    {\underbracket{\pcolor\pi' \left( \Ep \tilde P_{it} - \Ep \tilde P_{io} \right)}_{\text{Direct Policy Effect}} } \\
& +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( \Ep \tilde I_{it} - \Ep \tilde I_{io} \right)}+ {\icolor{\mu}'  \left( \Ep \tilde I_{it} - \Ep \tilde I_{io} \right)}}_{\text{ Dynamic Effect}}} \\
& =:   \mathrm{PEB}_t + \mathrm{PED}_t + \mathrm{DynE}_t,
  \end{aligned}
\end{equation}
where the first two components capture the immediate effect and the third represents the delayed or dynamic effect.

In the three examples of information structure given earlier, we have the following forms for the dynamic effect: for the trend model,
$$
   \mathrm{DynE}_t   = (\gamma \alpha + \mu) \Delta g_t, \quad  \Delta g_t = (g(t) - g(t-\ell))\\
$$
and for the lag model, $$
   \mathrm{DynE}_t  =    \sum_{m=1}^{t/\ell} \left( \gamma \alpha+ \mu   \right)^m (\mathrm{PEB}_{t-m \ell }  +\mathrm{PED}_{t-m \ell } ),
 $$
  interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
 For the general model we use, the dynamic effect is
 $$
 \begin{array}{lll}
(\mathrm{I})    & \mathrm{DynE}_t  & = \sum_{m=0}^{t/\ell} \left(( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m  \ \ ( (\gamma \alpha)_{1} + \mu_1) \Delta g_t  \\
& & +   \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}_{t-m \ell } +  \mathrm{DPE}_{t- m \ell} \right)
  \\
 &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}_{t-(m +1) \ell } +  \mathrm{DPE}_{t- (m+1) \ell}  \right).
 \end{array}
 $$
 The effects can be decomposed into (a) delayed policy effects via behavior by
summing terms containing $PEB$, (b) delayed policy effects via direct impact
by summing terms containing $DPE$,   (c) pure behavior effects, and (d)  pure dynamic feedback effects.

\subsection{Counterfactuals}

We  also consider simple counterfactual exercises, where we examine the effects of setting
a sequence of counterfactual policies for each state:
$$
\{{\pcolor P^\star_{it}} \}_{t=1}^T, \quad i=1, \ldots N.
$$
We assume that the SEM remains invariant, except of course for the policy equation.\footnote{It is possible to consider counterfactual exercises in which policy  responds to information through the policy equation if we are interested in endogenous policy responses to information. Although this is beyond the scope of the current paper, counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information. } Given the policies,
we propagate the dynamic equations:
\begin{equation} \label{eqn:CF-SEM} \tag{CEF-SEM}
  \begin{aligned}
& {\ycolor Y^\star_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B^\star_{it} } , {\pcolor P^\star_{it}}, {\icolor I^\star_{it}}), \\
& {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
& {\icolor I^\star_{it} }& := &  I_{it}({\icolor I^\star_{i,t-\ell}}, {\ycolor Y^*_{it}}, t),  \end{aligned}
\end{equation}
with the initialization $I^\star_{i0} = 0$, $ Y^\star_{i0} =0$, $B^\star_{i0} = 0$, $P^\star_{i0} = 0$.
In stating this counterfactual system of equations, we make the following invariance assumption
\begin{quote}
\textsc{Invariance Assumption.} The equations of (CF-SEM) remain exactly of the same form
as in the (SEM) and (I). That is, under the policy intervention $\{{\pcolor P^\star_{it}} \}$,  parameters and stochastic shocks in (SEM) and (I) remain the same as under the original policy intervention $\{{\pcolor P_{it}} \}$.
\end{quote}

Let $\mathcal{P} Y^\star_{i,t+\ell}$  and  $\mathcal{P} Y_{i,t+\ell}$ denote the predicted values produced
by working with the counterfactual system (\ref{eqn:CF-SEM}) and the factual system (SEM):
$$
\mathcal{P} Y^\star_{i,t+\ell}  = ( {\bcolor\alpha '} {\pcolor \beta'}  + {\pcolor\pi'})
    {\pcolor P^\star_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
    {\icolor I^\star_{it} }+ {\wcolor \bar \delta 'W_{it}},$$$$\mathcal{P} Y_{i,t+\ell} \ =
    ( {\bcolor\alpha '}  {\pcolor \beta'} + {\pcolor\pi'})
    {\pcolor P_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
    {\icolor I_{it} }+ {\wcolor \bar \delta 'W_{it}}.
$$
In generating these predictions, we make the assumption of invariance stated above.

Then we can write  the difference into the sum of three components:
  \begin{align}
\underbracket{\mathcal{P} Y^\star_{i,t+\ell} - \mathcal{P} Y_{i,t+\ell} }_{\text{Predicted CF Change}} & =
\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}  +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \nonumber\\
&\qquad +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \nonumber \\
 & =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}. \label{eqn:DC2}
  \end{align}
%\begin{equation} \label{eqn:DC2}
%  \begin{aligned}
%\underbracket{\mathcal{P} Y^\star_{it} - \mathcal{P} Y_{it} }_{\text{Predicted CF Change}}& =
%{\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}} \\
%& +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \\
%& +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \\
%& =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}.
%  \end{aligned}
%\end{equation}

Similarly to what we had before,  the counterfactual dynamic effects take the form:
$$
\begin{array}{lll}
(\mathrm{I})  & \mathrm{DynE}^\star_{it} &=
  \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}^\star_{i,t-m \ell } +  \mathrm{DPE}^\star_{i,t- m\ell} \right)
  \\
 &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}^\star_{i,t-(m +1) \ell } +  \mathrm{DPE}^\star_{i,t- (m+1)\ell}  \right),
 \end{array}
 $$
 interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
 The effects can be decomposed into (a) delayed policy effects via behavior by
summing terms containing $PEB$, (b) delayed policy effects via direct impact
by summing terms containing $DPE$,  (c) pure behavior effects, and (d)  pure dynamic feedback effects.

\section{Empirical Analysis}
\subsection{Data}

Our baseline measures for daily Covid-19   cases and deaths are from
\href{https://github.com/nytimes/covid-19-data}{The New York Times (NYT)}. When there are missing values in NYT, we use reported cases and deaths from \href{https://github.com/CSSEGISandData/COVID-19}{JHU CSSE}, and then the
\href{https://github.com/COVID19Tracking/covid-tracking-data}{Covid
Tracking Project}.  The number of tests for each state is from  \href{https://github.com/COVID19Tracking/covid-tracking-data}{Covid
 Tracking Project}. As shown in Figure \ref{fig:test} in the appendix, there was a rapid increase in testing in the second half of March and then the
number of tests increased very slowly in each state in April.

We use the database on US state policies created by
\cite{raifman2020}.
%This database records that each state implemented
%25 Covid-19  related policies.
In our analysis, we focus on 6
policies:  stay-at-home, closed nonessential
businesses, closed K-12 schools, closed restaurants except takeout, closed movie theaters, and mandate face mask by  employees in public facing businesses.
We believe that the first four of these
policies are the most widespread and important.
%, where state of emergency was declared first and the implementation of other policies followed with some time lag.
Closed movie theaters is included because it captures common bans on gatherings of more than
a handful of people. We also include mandatory face mask use by employees because its effectiveness on slowing down Covid-19 spread is a controversial policy issue \citep{howard2020,Greenhalghm2020,zhangr2020}.
Table \ref{tab:policies_inreg} provides summary statistics, where $N$ is the number of states that have ever implemented the policy. We also obtain information on state-level covariates from \citet{raifman2020}, which include population size, total area, unemployment rate, poverty rate, and a percentage of people who are subject to illness.These confounders are motivated by \cite{wheaton2020} who finds that case growth is associated  with residential density and per capita income.

\input{tables_and_figures/policies_inregs}

We obtain social distancing behavior measures from``\href{https://www.google.com/covid19/mobility/}{Google COVID-19 Community Mobility Reports}''
\citep{google2020}.  The dataset provides six measures of ``mobility trends''  that report a percentage change in visits and length of stay at different places relative to a baseline computed by their median values of the same day of the week from January 3 to February 6, 2020.  Our analysis focuses on the following four measures: ``Grocery \& pharmacy," ``Transit stations,'' ``Retail \& recreation,'' and ``Workplaces.''\footnote{The other two measures are ``Residential'' and  ``Parks.''  We drop ``Residential'' because it is highly correlated with both ``Workplaces'' and  ``Retail \& recreation''  at correlation coefficients of -0.98. We also drop ``Parks'' because it does not have clear implication on the spread of Covid-19.}

Figure \ref{fig:transit-workplaces} shows the evolution of  ``Transit stations'' and ``Workplaces,'' where thin lines are the value in each state and date while thicker colored lines are quantiles  conditional on date. The figures illustrate a sharp decline in people's movements starting from mid-March as well as differences in their evolutions across states. They also reveal periodic fluctuations associated with days of the week, which motivates our use of weekly measures.

\begin{figure}\caption{The Evolution of Google Mobility Measures: Transit stations and Workplaces\label{fig:transit-workplaces}}\vspace{0.1cm}
    \begin{tabular}{cc}
      \includegraphics[width=0.5\linewidth]{tables_and_figures/transit}
      \includegraphics[width=0.5\linewidth]{tables_and_figures/workplaces}
       \end{tabular}
        \begin{flushleft}
        \scriptsize{This figure shows the evolution of ``Transit stations'' and  ``Workplaces'' of Google Mobility
      Reports. Thin gray lines are the value in each state and date. Thicker colored lines are
      quantiles of the variables conditional on date.}
      \end{flushleft}
\end{figure}

In our empirical analysis, we use weekly measures for cases, deaths,
and tests by summing up their daily measures from day \(t\) to
\(t-6\).  We focus on weekly cases and deaths because daily new cases and deaths are
affected by the timing of reporting and testing, and are quite
volatile as shown in Figure \ref{fig:dailycases} in the
appendix.  Aggregating to weekly new cases/deaths/tests smooths
out idiosyncratic daily noises as well as periodic fluctuations
associated with days of the week.
We also construct weekly policy and
behavior variables by taking 7 day moving averages from day \(t-14\) to
\(t-21\) for case growth, where the  delay reflects the time lag between infection and case confirmation. The four weekly behavior variables are referred as ``Transit
Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
Intensity.'' Consequently, our empirical analysis uses 7 day moving averages of all variables recorded at daily frequencies. Our sample period is from March 7, 2020 to June 3, 2020.

\begin{table}\caption{Correlations among Policies and Behavior \label{tab:correlation}}\vspace{-0.2cm}
  \begin{minipage}{\linewidth}
    \resizebox{\linewidth}{!}{
      \input{tables_and_figures/corr}
    }\smallskip
       \begin{flushleft}
         \scriptsize
         Each off-diagonal entry reports a correlation coefficient of
         a pair of policy and behavior variables.
       \end{flushleft}
  \end{minipage}
\end{table}

Table \ref{tab:correlation} reports that weekly policy and behavior
variables are highly correlated with each other, except for the``masks
for employees'' policy.  High correlations may cause multicolinearity
problems and could limit our ability to separately identify the
effect of each policy or behavior variable on case growth, but this
does not prevent us from identifying the aggregate effect of all
policies and behavior variables on case or death growth.

Figure \ref{fig:policyportion} shows the portion of states that have
each policy in place at each date. For most policies, there is
considerable variation across states in the time in which the policies
are active. The one exception is K-12 school closures. About 80\% of
states closed schools within a day or two of March 15th, and all
states closed schools by early April. This makes the effect of school
closings difficult to separate from aggregate time series variation.

\begin{figure}[ht]\caption{Portion of states with each
    policy \label{fig:policyportion}}
  \begin{minipage}{\linewidth}
    \begin{tabular}{ccc}
      \includegraphics[width=0.31\textwidth]{tables_and_figures/pmaskbus_p}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/pk12_p}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/pshelter_p}
      \\
      \includegraphics[width=0.31\textwidth]{tables_and_figures/pmovie_p}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/prestaurant_p}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/pnonessential_p}
    \end{tabular}
  \end{minipage}
\end{figure}

\subsection{The Effect of Policies and Information on Behavior\label{policies-and-behavior}}

We first examine how policies and information affect social distancing behaviors by estimating a version of (\ref{eq:R2}):
\begin{align}
  {\bcolor B_{it}^j}
  %& = {\pcolor (\beta^j)' P_{it} }  + {\icolor \gamma_{1}^j \log(t) + \gamma_{2}^j \log (\Delta C_{i,t-14}) +  \gamma_{3}^j Y_{i,t-7}} + {\wcolor (\delta_B^j)' W_{it}} + \varepsilon_{it}^j \nonumber\\
  & = {\pcolor (\beta^j)' P_{it}} + {\icolor (\gamma^j)' I_{it}} +
    {\wcolor (\delta_B^j)' W_{it}} + \varepsilon_{it}^{bj}, \notag
\end{align}
where ${\bcolor B_{it}^j}$ represents behavior variable $j$  in state  $i$ at time $t$.
${\pcolor P_{it}}$ collects the Covid related policies  in state $i$ at $t$.
Confounders, ${\wcolor W_{it}}$, include state-level covariates, month
indicators, and their interactions.
${\icolor I_{it}}$ is a set of information variables that affect
people's behaviors at $t$. As information, we include each state's
growth of cases (in panel \ref{tab:PItoBcases}) or deaths (in panel
\ref{tab:PItoBdeaths}), and log cases or deaths.  Additionally, in
columns (5)-(8) of each panel, we include national growth and log
of cases or deaths.

\afterpage{\clearpage
  %\thispagestyle{empty}
  \begin{landscape}
    \begin{table}[!htbp] \centering
      \caption{The Effect of Policies and Information on Behavior ($P
        I \to B$)}\vspace{-0.3cm}\label{tab:PItoB}

      \begin{subtable}{\linewidth}\caption{Cases as Information\label{tab:PItoBcases}}
      \begin{minipage}{\linewidth}
        \centering
        \resizebox{\textwidth}{!}{
          \footnotesize
          \input{tables_and_figures/behavior-NI-pib}
        }
      \begin{flushleft}
        \scriptsize
        Dependent variables are   ``Transit
        Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
        Intensity" defined as 7 days moving averages of corresponding daily measures obtained from Google Mobility Reports.  All specifications include state-level characteristics (population, area, unemployment rate, poverty rate, and a percentage of people subject to illness) as well as their interactions with the log of days since Jan 15, 2020. The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients. Standard errors are clustered at the state level.
        \end{flushleft}
      \end{minipage}
    \end{subtable}
  \end{table}
  \pagebreak
  \begin{table}[!htbp] \centering
    \ContinuedFloat
    \begin{subtable}{\linewidth}\caption{Deaths as Information \label{tab:PItoBdeaths}}
      \begin{minipage}{\linewidth}
        \centering
        \resizebox{\textwidth}{!}{
          \footnotesize
            \input{tables_and_figures/behavior-deathsinfo-NI-pib}
          }
        \begin{flushleft}
          \scriptsize
          Dependent variables are   ``Transit
          Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
          Intensity" defined as 7 days moving averages of corresponding daily measures obtained from Google Mobility Reports.  All specifications include state-level characteristics (population, area, unemployment rate, poverty rate, and a percentage of people subject to illness) as well as their interactions with the log of days since Jan 15, 2020. The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients. Standard errors are clustered at the state level.
        \end{flushleft}
      \end{minipage}
    \end{subtable}
  \end{table}
\end{landscape}
\clearpage}

Table \ref{tab:PItoB} reports the estimates with standard errors clustered at the state level.  Across different
specifications, our results imply that policies have large effects on
behavior. Comparing columns (1)-(4) with columns (5)-(8),
the magnitude of policy effects are sensitive to whether national
cases or deaths are included as information. The coefficient on school
closures is particularly sensitive to the inclusion of national
information variables.  As shown in Figure
\ref{fig:policyportion}, there is little variation across states in
the timing of school closures. Consequently, it is difficult to
separate the effect of school closures from a behavioral response to the national
trend in cases and deaths.

The other policy coefficients are less sensitive to the inclusion of
national case/death variables. After school closures, stay-at-home
orders and restaurant closures have the next largest effect. Somewhat
surprisingly, closure of nonessential businesses appears to have
a modest effect on behavior. Closing movie theaters has a similar,
small effect on behavior. The effect of masks for employees is also
small. The comparison of effects across policies should be interpreted
with caution.  Differences between policy effects are often
statistically insignificant; except for masks for employees, the
policies are highly correlated and it is difficult to separate their
effects.

The row ``$\sum_j \mathrm{Policy}_j$'' reports the sum of the estimated
effect of all policies, which is substantial and can account for a
large fraction of the observed declines in behavior variables.  For
example,  in Figure \ref{fig:transit-workplaces}, transit intensity for a median state was approximately -50\% at its lowest point
 in early April. The estimated policy coefficients in columns imply
that imposing all six policies would lead to roughly 85\% (in columns
1-4) or roughly 50\% (in columns 5-8) of the observed decline. The large impact of policies on transit intensity suggests that the policies may have reduced the Covid-19 infection by reducing people's use of public transportation.\footnote{Analyzing the New York City's subway ridership,  \cite{NBERw27021} finds  a strong link between public transit and spread of infection.}

%retail intensity was approximately -40\% at its lowest point
%in mid April, and the estimated policy coefficients in columns imply
%that imposing all six policies would lead to either all (in columns
%1-4) or half (in columns 5-8) of the observed decline.

In panel \ref{tab:PItoBdeaths}, estimated coefficients of deaths and
death growth are generally negative. This suggests that the higher
number of deaths reduces social interactions measured by Google
Mobility Reports perhaps because people are increasingly aware of
prevalence of Covid-19 \citep{maloney2020}.  The coefficients on log
cases and case growth in panel \ref{tab:PItoBcases} are more mixed. In
columns (5)-(8) of both panels, we see that national case/death
variables have large, negative coefficients. This suggests that behavior responded to national conditions although it is also likely that national case/death variables capture unobserved aggregate time effects beyond information (e.g., latent policy variables and time-varying confounders that are common across states) which are not fully controlled by month dummies.

%For grocery, the positive estimated coefficients of cases and case growth in columns (3) and (7) might be capturing stockpiling behavior, which could temporarily increase the visits to grocery.

%In columns (5)-(8), the effects of masks for employees are allowed to be different before and after May 1st. The negative effect of mask mandates on mobility measures appears to be slightly larger in May than in April but the difference is not statistically significant.

\begin{figure}[ht]
  \caption{Case and death growth conditional on mask mandates \label{fig:masks}}\vspace{0.2cm}
 \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases}
      &
        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths}
    \end{tabular}
    \begin{flushleft}
      \scriptsize In these figures, red points are the case or death
      growth rate in states without a mask mandate. Blue points are
      states with a mask mandate 14 (21 for deaths) days prior. The
      red line is the average across states without a mask mandate 14
      (21 for deaths) days earlier. The blue line is the average
      across states with a mask mandate 14 (21 for deaths) earlier.
    \end{flushleft}
  \end{minipage}
\end{figure}

\afterpage{
 \clearpage
 %\thispagestyle{empty}
 \begin{landscape}
\begin{table}[!htbp] \centering
 \caption{The Direct Effect of Behavior and Policies on Case and
   Death Growth ($BPI \to Y$)}\vspace{-0.3cm}
 \label{tab:BPItoY}
 \begin{minipage}{\linewidth}
   \centering
   \resizebox{\textwidth}{!}{
   \tiny
   \begin{tabular}{c|c}
   \begin{minipage}{0.49\linewidth}
     \centering
     \input{tables_and_figures/cases-pbiy}
   \end{minipage}
     %\hspace{1pt}
     &
   \begin{minipage}{0.49\linewidth}
     \centering
     \input{tables_and_figures/deaths-deathsinfo-pbiy}
   \end{minipage}
   \end{tabular}
   }
 \begin{flushleft}
     \scriptsize Dependent variable is the weekly growth rate of
     confirmed cases (in the left panel) or deaths (in the right
     panel) as defined in equation (\ref{eq:y}). The covariates
     include lagged policy and behavior variables, which are
     constructed as 7 day moving averages between $t$ to $t-7$ of
     corresponding daily measures.  The row
     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
     coefficients.  The row ``$\sum_k w_k \mathrm{Behavior}_k$''
     reports the sum of four coefficients of behavior variables
     weighted by the average of each behavioral variable from April
     1st-10th.  Standard errors are clustered at the state level.
   \end{flushleft}
 \end{minipage}
\end{table}
\end{landscape}
\clearpage
}

\subsection{The Direct Effect of Policies and Behavior on Case
and Death Growth\label{policy-behavior-and-case-growth}}

We now analyze how behavior and policies together influence case and death
growth rates. We begin with some simple graphical evidence of the
effect of policies on case and death growth. Figure \ref{fig:masks}
shows average case and death growth conditional on date and whether
masks are mandatory for employees.\footnote{We take 14 and 21 day lags of mask policies for case and death growths, respectively, to  identify the states with a mask mandate because policies affect cases and deaths with time lags. See our discussion in the Appendix \ref{sec:timing}.  } The left panel of the figure shows
that states with a mask mandate consistently have 0-0.2 lower case
growth than states without. The right panel  also illustrates that
states with a mask mandate tend to have lower average death growth than states without a mask mandate.

Similar plots are shown for other policies in Figures
\ref{fig:growthpolicies1} and \ref{fig:growthpolicies2} in the
appendix. The  figures for stay-at-home orders and closure
of nonessential businesses are qualitatively similar to that for
masks. States with these two policies appear to have about 0.1 percentage point lower
case growth  than states without. The effects of school closures, movie theater closures, and
restaurant closures are not clearly visible in these figures. These
figures are merely suggestive; the patterns observed in them may be
driven by confounders.

We more formally analyze the effect of policies by estimating
regressions. We first look at the direct effect of policies on case and death growth
conditional on behavior by estimating equation (\ref{eq:R1}):
% We partition $X_{it}$ in (\ref{eq:M}) into behavior variables $B_{it}$,  policy variables $P_{it}$, information variables $I_{it}$, and other confounding variables $W_{X,it}$ as $X_{it}= (B_{it}',P_{it}', I_{it}', W_{X,it}')'$ with $\theta=(\alpha',\pi',\mu',\delta_X')'$ and rewrite  (\ref{eq:M}) as:
\begin{align}
  {\ycolor  Y_{i,t+\ell}}
  & = {\bcolor \alpha ' B_{it}} + {\pcolor\pi 'P_{it}} + {\icolor \mu'I_{it}} +  {\wcolor\delta_Y 'W_{it}}  + \varepsilon^y_{it},
  \label{eq:reg-y}
\end{align}
where the outcome variable, $\ycolor Y_{i,t+\ell}$, is either case growth  or death growth.

For case growth as the outcome, we choose a lag length of $\ell=14$ days for behavior, policy, and information variables to reflect the delay between infection and
confirmation of case.\footnote{As we review in the Appendix \ref{sec:timing},
  a lag length of 14 days between exposure and case reporting, as well as a lag length of 21 days between exposure and deaths, is broadly consistent with currently
  available evidence.  }   $\bcolor B_{it}=(B_{it}^1,...,B_{it}^4)'$ is a vector of four behavior
variables  in state $i$. $\pcolor P_{it}$ includes the Covid-related policies
in state $i$  that directly affect the spread of Covid-19
after controlling for behavior variables (e.g., masks for employees).
We include information variables,  $\icolor I_{it}$, that include the past cases and case growths
because  the past cases may be
correlated with (latent) government  policies or people's behaviors
that are not fully captured by our observed policy and behavior
variables. We also consider a specification that includes the past cases and case growth at the national level as additional information variables. $\wcolor W_{it}$ is a set of confounders that include month dummies, state-level covariates,  and the interaction terms between month dummies and state-level covariates.\footnote{Month dummies also represent the latent information that is not fully captured by the past cases and growths.}
For case growth, $\wcolor W_{it}$ also includes the  test rate growth  $\Delta \log(T)_{it}$  to capture the effect of changing test rates on confirmed cases.
Equation (\ref{eq:reg-y}) corresponds to (\ref{eq:M})  derived from the SIR model.

%We also estimate the effect of policies on death growth by estimating  (\ref{eq:reg-y}) with death growth
%$\Delta \log \Delta D_{it}$  as outcome variable.
For death growth as the outcome, we take a lag length of $\ell=21$ days. The information variables $\icolor I_{it}$ include  past deaths and death growth rates;  $\wcolor W_{it}$ is the same as that of the case growth equation except that the growth rate of test rates is excluded from $\wcolor W_{it}$ as implied by equation (\ref{eq:M-D}).

% Tables \ref{tab:BPItoY-c} and  \ref{tab:BPItoY-d} show
Table \ref{tab:BPItoY} shows
the results of estimating
(\ref{eq:reg-y}) for case and death growth rates. %, respectively.
Column (1) represents our baseline specification while column (2) allows the effect of masks to be different before and after May 1st. Columns (3) and (4) include past cases/deaths and growth rates at national level as additional regressors.

The estimates indicate that mandatory face masks for employees
reduce the growth rate of infections and deaths by 8-15 percent, while holding behavior constant. This suggests that
requiring masks for employees in public-facing businesses may be an effective preventive measure.\footnote{Note that we are \textit{not} evaluating the effect of \textit{universal} mask-wearing for the public but that of mask-wearing for employees. The effect of \textit{universal} mask-wearing for the public could be larger if people comply with such a policy measure. \cite{tian2020calibrated} considered a  model in which mask wearing reduces  the reproduction number by a factor $(1-e \cdot pm)^2$, where $e$ is the efficacy of trapping viral particles inside the mask and $pm$ is the percentage of mask-wearing population. Given an estimate of $R_0=2.4$, \cite{howard2020} argue that  50\% mask usage and a 50\% mask efficacy level would reduce the reproduction number from 2.4 to 1.35, an order of magnitude
impact.} The estimated effect of masks on death growth is larger than the effect on case growth, but this difference between the two estimated effects is not statistically significant.

Except for mask requirements, policies appear to have little direct
effect on case or death growth when behavior is held constant. The one
exception is that closing schools has a large and statistically
significant coefficient in the death growth regressions. As discussed
above, there is little cross-state variation in the timing of school
closures, making estimates of its effect less reliable.

The row ``$\sum_k w_k \mathrm{Behavior}_k$'' reports the sum of estimated coefficients weighted by the average of the behavioral
variables from April 1st-10th. The estimates of $-0.76$ and $-0.87$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 76 and 83 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
When including national cases and deaths in information, as shown in columns (3) and (4), the estimated aggregate impact of behavior is substantially smaller, but remains large and statistically significant.

A useful practical implication of these results are that Google Mobility
Reports and similar data might be useful as a leading indicator of
potential case or death growth. This should be done with caution, however, because other
changes in the environment might alter the relationship between behavior
and infections. Preventative
measures, including mandatory face masks, and changes in habit that are not captured in our data might alter
the future relationship between Google Mobility Reports and case/death growth.

The negative coefficients of past cases or deaths in
% tables \ref{tab:BPItoY-c} and  \ref{tab:BPItoY-d}  are
Table \ref{tab:BPItoY} is
consistent with a hypothesis that higher reported cases and deaths change people's behavior to reduce transmission risks. Such behavioral changes in response to new information   are
 partly captured by Google mobility measures, but the negative estimated coefficient of past cases or deaths imply that other latent behavioral changes that are not fully captured by Google mobility measures (e.g., frequent hand-washing, wearing masks, and keeping 6ft/2m distancing) are also important for reducing future cases and deaths.

If policies are enacted and behavior changes, then future cases/deaths and
information will change, which will induce further behavior changes.  However, since the model includes lags of cases/deaths as well as their growth rates,  computing a long-run effect is not completely
straightforward. We investigate dynamic effects that incorporate
feedback through information in section \ref{counterfactuals}.

% \begin{table}[!htbp] \centering
%   \caption{\label{tab:PtoY-c}
%     The Total Effect of Policies on Case Growth ($PI \to Y$)}\vspace{-0.3cm}
%   \begin{minipage}{\linewidth}
%     \centering
%     \scriptsize
%     \begin{tabular}{c}
%     \begin{minipage}{\linewidth}
%       \centering
%       \input{tables_and_figures/cases-piy}
%     \end{minipage}
%     \end{tabular}
%     \begin{flushleft}
%       \scriptsize Dependent variable is the weekly growth rate of
%       confirmed cases (in the left panel) or deaths (in the right
%       panel) as defined in equation (\ref{eq:y}). The covariates
%       include lagged policy variables, which are
%       constructed as 7 day moving averages between $t$ to $t-7$ of
%       corresponding daily measures.  The row
%       ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
%       coefficients.
%     \end{flushleft}
%   \end{minipage}
% \end{table}

% \begin{table}[!htbp] \centering
%   \caption{\label{tab:PtoY-d}
%     The Total Effect of Policies on Death Growth ($PI \to Y$)}\vspace{-0.3cm}
%   \begin{minipage}{\linewidth}
%     \centering
%     \scriptsize
%     \begin{tabular}{c}
%     \begin{minipage}{\linewidth}
%       \centering
%       \input{tables_and_figures/deaths-deathsinfo-piy}
%     \end{minipage}

%     \end{tabular}
%     \begin{flushleft}
%       \scriptsize Dependent variable is the weekly growth rate of
%       confirmed cases (in the left panel) or deaths (in the right
%       panel) as defined in equation (\ref{eq:y}). The covariates
%       include lagged policy variables, which are
%       constructed as 7 day moving averages between $t$ to $t-7$ of
%       corresponding daily measures.  The row
%       ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
%       coefficients.
%     \end{flushleft}
%   \end{minipage}
% \end{table}

%Here we present the results of counterfactual analyses that don't include the national cases/deaths as
%the the information variables.  The growth in the national cases nearly perfectly predicts school closure
%policies, which have little cross-sectional variation.  Here we remove the national cases from the information
%and check what happens to the counterfactual results.  The counterfactual results of mask policies, shelter-in-place,
%and closing
%non-essential businesses remain robust. The results on school closures effects change substantially,
%attributing a much bigger role to school closures. The resulting counterfactual of  removing
%all policies has a much larger effect on cases, with the lower bound of the confidence band implying at
%least a 7 fold increase in cases (see Figure \ref{fig:US-nop-SI} given in the main text).

  \afterpage{
 \begin{landscape}
\begin{table}[!htbp] \centering
 \caption{\label{tab:PtoY}
   The Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
 \begin{minipage}{\linewidth}
   \resizebox{\textwidth}{!}{
   \centering
   \tiny
   \begin{tabular}{c|c}
   \begin{minipage}{0.49\linewidth}
     \centering
     \input{tables_and_figures/cases-piy}
   \end{minipage}
     &
   \begin{minipage}{0.49\linewidth}
     \centering
     \input{tables_and_figures/deaths-deathsinfo-piy}
   \end{minipage}

   \end{tabular}
   }
   \begin{flushleft}
     \scriptsize Dependent variable is the weekly growth rate of
     confirmed cases (in the left panel) or deaths (in the right
     panel) as defined in equation (\ref{eq:y}). The covariates
     include lagged policy variables, which are
     constructed as 7 day moving averages between $t$ to $t-7$ of
     corresponding daily measures.  The row
     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
     coefficients.  Standard errors are clustered at the state level.
   \end{flushleft}
 \end{minipage}
\end{table}
\end{landscape}
}

\afterpage{
\begin{table}[!b]
  \caption{\label{tab:dieff-si}Direct and Indirect Policy Effects
    without national case/death variables}
\begin{minipage}{\linewidth}
  \centering
    \scriptsize
  \begin{tabular}{c}
    \textbf{Cases}
    \\
    \input{tables_and_figures/dieff-cases-SI}
    \\
    \textbf{Deaths}
    \\
    \input{tables_and_figures/dieff-deaths-deathsinfo-SI}
  \end{tabular}
  \smallskip
  \begin{flushleft}
    {\tiny Direct effects capture the effect of policy on case
      growth holding behavior, information, and confounders
      constant. Direct effects are given by ${\pcolor \pi}$ in
      equation (\ref{eq:R1}). Indirect effects capture how policy
      changes behavior and behavior shift case growth. They are given
      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
      from (\ref{eq:R2}). The total effect is
      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
      \ref{eq:R4}. Columns ``Difference''    are  the differences between
      the estimates from (\ref{eq:R4}) and the combination of
      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
      Standard errors are computed by
      bootstrap and clustered on state.}
    %  \ref{tab:PtoY}.
    \end{flushleft}
  \end{minipage}
\end{table}
}

\afterpage{
\begin{table}
  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
\begin{minipage}{\linewidth}
  \centering
    \scriptsize
  \begin{tabular}{c}
    \textbf{Cases}
    \\
    \input{tables_and_figures/dieff-cases-NI}
    \\ \\
    \textbf{Deaths}
    \\
    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
    \\
  \end{tabular}
  \smallskip
  \begin{flushleft}
      \scriptsize Direct effects capture the effect of policy on case
      growth holding behavior, information, and confounders
      constant. Direct effects are given by ${\pcolor \pi}$ in
      equation (\ref{eq:R1}). Indirect effects capture how policy
      changes behavior and behavior shift case growth. They are given
      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
      from (\ref{eq:R2}). The total effect is
      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
      \ref{eq:R4}. Columns ``Difference''    are  the differences between
      the estimates from (\ref{eq:R4}) and the combination of
      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
      Standard errors are computed by
      bootstrap and clustered on state.
    %  \ref{tab:PtoY}.
    \end{flushleft}
  \end{minipage}
\end{table}
\clearpage}

\subsection{The Total Effect of Policies on Case Growth\label{total-policy-effect}}

In this section, we focus our analysis on policy effects when we hold
information constant. The estimated effect of policy on
behavior in Table \ref{tab:PItoB} and those of policies and
behavior on case/death growth in Table  \ref{tab:BPItoY}  can be
combined to calculate the total effect of policy as well as its
decomposition into  direct and indirect effects.

The first three columns of Table \ref{tab:dieff-si} show the direct
(holding behavior constant) and indirect (through behavior changes)
effects of policy under a specification that excludes national information variables. These are computed from the specification with
national cases or deaths included as information (columns (1)-(4) of
Table \ref{tab:PItoB} and column (1) of Table \ref{tab:BPItoY}). The estimates imply
that all policies combined would reduce the growth rate of cases  and deaths by
0.70 and 0.98, respectively,   out of which about two-third to
three-fourth is attributable to the indirect effect through their impact on behavior.  The estimate also indicates that the
effect of mandatory masks for employees is mostly direct.

We can also examine the total effect of policies and information on
case  or death growth, by estimating (\ref{eq:R4}). The coefficients on policy
in this regression combine both the direct and indirect effects.

Table \ref{tab:PtoY} shows
the full set of coefficient estimates for
(\ref{eq:R4}). The results are broadly consistent with what we found
above.  As in Table \ref{tab:PItoB}, the effect of school closures is
sensitive to the inclusion of national information variables.  Also as
above, mask mandates have a significant negative effect on growth
rates.

In columns (2) and (4) of Table \ref{tab:PtoY}, we find that the estimated effect of mask mandates in April is larger than that in May for both case and death regressions.  This may reflect a wider \textit{voluntary} adoption of masks in May than in April --- if more people wear masks even without mandatory mask policy,  the policy effect of mandating masks for employees becomes weaker.

Table \ref{tab:dieff} presents the estimates for the specification with past national case/death variables. The effects of school closures and  the sum of policies are estimated substantially smaller in Table \ref{tab:dieff} when national case/death variables are included than in Table \ref{tab:dieff-si}. This sensitivity reflects the difficulty in identifying the aggregate time effect---which is largely captured by national cases/deaths---given little cross-sectional variation in the timing of school closures across states. On the other hand, the estimated effects of policies other than school closures are similar between Table \ref{tab:dieff-si} and Table \ref{tab:dieff}; the effect of other policies are well-identified from cross-sectional variations.

Column ``Difference''  in Tables   \ref{tab:dieff-si} and \ref{tab:dieff}  show the
difference between the estimate of (\ref{eq:R4}) in column  ``PI$\to$Y
Coefficient''   and the implied estimate from
(\ref{eq:R1})-(\ref{eq:R2}) in   column ``Total.''   Differences  are generally small and statistically insignificant, broadly supporting the validity of extra orthogonality condition in (\ref{eq:R1}).  The difference for school closures as well as  the sum of all policies in Table   \ref{tab:dieff-si} is significantly different from zero, which may be due to the aforementioned difficulty in identifying the effect of school closures.   There is substantial external epidemiological evidence that suggests that schooling closures may have substantial effects on the spread of the virus: studies like \cite{children:germany} and \cite{children:nature} establish that children carry substantial amounts of viral loads and can contribute to the transmission (due to higher contact rate than other age groups).\footnote{The evidence presented in \cite{children:germany}  has lead German to make the decision to close schools early.} The US data does not allow us to pint down the effect of closing schools reliably due to their approximate collinearity with trends in national cases.

Column ``Average'' of  Tables \ref{tab:dieff-si} and \ref{tab:dieff}   reports the average of
``Total'' and ``PI$\to$Y Coefficient'' columns.  The average is an appealing and simple way
to combine the two estimates of the total effect: one relying on the causal structure and another inferred from a direct estimation of equation (PI $\to$ Y).\footnote{Averaging the two estimates theoretically reduces noise, albeit in our case the reductions are small.
Another approach would be to use precision averaging, which would give similar result. Finally, another approach would be to use generalized method of moments to estimate
all of the equations jointly. We don't pursue this approach since it is likely to be non-robust under local deviations from correct specification; simple model averaging is more appealing in this case.}  We shall be using the average estimate in generating the counterfactuals in the next section. Turning to the results, the estimates of Tables \ref{tab:dieff-si} and \ref{tab:dieff}  imply that all policies combined would reduce
\(\Delta \log \Delta D\) by  0.97 and  0.40, respectively. For
comparison, the median of \(\Delta \log \Delta D_{it}\) reached its
peak in mid-March of about 1.3 (see Figure \ref{fig:growthq} in the
appendix). Since then it has declined to near 0. Therefore,  -0.97 and -0.40 imply
that policy changes can account for roughly one-third to two-third of the observed
decrease in death growth.  The remainder of the decline is likely due
to changes in behavior from information.

\section{Empirical Evaluation of Counterfactual Policies}\label{counterfactuals}

We now turn our focus to dynamic feedback effects. Policy
and behavior changes that reduce case and death growth today can lead to a more
optimistic, riskier behavior in the future, attenuating longer run effects.
We perform the main counterfactual experiments using the average of two estimated coefficients as reported in column ``Average'' of Table \ref{tab:dieff-si} under a specification that excludes the number of past national cases and deaths from information variables. In the appendix, we also report additional counterfactual experiment results with the specification that includes the national information variables, and find that they are very similar. The results on mask policies, business closures, stay-at-home orders are robust with respect to this variation (see Figures \ref{fig:US-mask}-\ref{fig:US-shelter} in the appendix). On the other hand, the results on removing all policies, particularly closure of schools, reported in the next section, are sensitive to the inclusion of national information variables, highlighting the large uncertainty regarding the size of the effect.  In Figures \ref{fig:WA-mask}-\ref{fig:US-shelter} below, the top panel presents the result on cases while the bottom panel presents the result on deaths.

%Additional counterfactual experiments include the
%} %\footnote{\textbf{The results of counterfactual experiments are not sensitive to whether we use the estimate of  table \ref{tab:dieff-si} or that of table \ref{tab:dieff} except for a counterfactual experiment of removing all policies  in section \ref{removing-policies}.}}

\subsection{Business Mask Mandate}

\begin{figure}[ht]
  \caption{Effect of mandating masks on April 1st in Washington State \label{fig:WA-mask}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{ccc}
      \multicolumn{3}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_v1}
      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_v1}
      \\
      \\
      \multicolumn{3}{c}{\textbf{Deaths}}
      \\      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths}
      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_v1}
      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_v1}
    \end{tabular}
    \begin{flushleft}
      \footnotesize To compute the estimated and counterfactual paths
      we use the average of two estimated coefficients as reported in
      column ``Average'' of Table \ref{tab:dieff-si}. We set initial
      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
      observed in the state we are simulating. We hold all other
      regressors at their observed values. Error terms are drawn with
      replacement from the residuals. We do this many times and report
      the average over draws of the residuals. The shaded region is a
      point-wise 90\% confidence interval.
    \end{flushleft}
  \end{minipage}
\end{figure}

We first consider the impact of a nationwide mask mandate for
employees beginning on April 1st. As discussed earlier, we find that
mask mandates reduce case and death growth even when holding behavior
constant. In other words, mask mandates may reduce infections with
relatively little economic disruption. This makes mask mandates a
particularly attractive policy instrument. In this section we examine
what would have happened to the number of cases if all states had imposed a mask
mandate on April 1st.\footnote{We feel this is a very plausible
  counterfactual policy. In a paper made publicly available on
  April 1st, \cite{abaluck2020} argued for mask usage based on
  comparisons between countries with and without pre-existing norms of
  widespread mask usage.}

For illustrative purpose, we begin by focusing on Washington State.
The
left column of  Figure \ref{fig:WA-mask} shows the observed, estimated
average, and counterfactual average of $\Delta \log \Delta C$ (top panel) and
$\Delta \log \Delta D$ (bottom panel). To compute the estimated and counterfactual
paths, we use the   estimate   in
column ``Average'' of Table \ref{tab:dieff-si}. We set initial
$\Delta \log \Delta C$ and $\log \Delta C$ to their values first
observed in the state we are simulating. We hold all other regressors
at their observed values. Error terms are drawn with replacement from
the residuals. We do this many times and report the average over draws
of the residuals. The shaded region is a point-wise 90\% confidence
interval. The left column shows that the fit of the estimated and
observed growth rate is quite good.

The middle column of Figure \ref{fig:WA-mask} shows the change in
growth rate from mandating masks on April 1st. The shaded region is a
90\% pointwise confidence interval. As shown, mandating masks on April
1st lowers the growth of cases or deaths 14 or 21 days later by 0.1 to
0.15. This effect then gradually declines due to information
feedback. Mandatory masks reduce past cases or deaths, which leads to
less cautious behavior, attenuating the impact of the policy. The
reversal of the decrease in growth in late April is due to our
comparison of a mask mandate on April 1st with Washington's actual
mask mandate in early May. By late April, the counterfactual mask
effect has decayed through information feedback, and we are comparing
it the undecayed impact of Washington's actual, later mask mandate.

The right column of Figure \ref{fig:WA-mask} shows how the changes in
case and death growth translate into changes in cases and deaths. The
estimates imply that mandating masks on April 1st would have led to
500 fewer cases and 250 fewer deaths in Washington by the start of
June.

\begin{figure}[ht]
  \caption{Effect of nationally mandating masks for employees on April
    1st in the US\label{fig:US-mask}}
  \begin{minipage}{\linewidth}
    \centering
    \medskip
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_v1}
      &
        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_v1}
      &
        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
    \end{tabular}
  \end{minipage}
    \begin{flushleft}
      \footnotesize In the left column, the dots are the average
      change in growth in each state. The blue line is the average
      across states of the change in growth. The shaded region is a
      point-wise 90\% confidence interval. The right column shows the
      change in cases or deaths relative to the baseline of actual
      policies.
    \end{flushleft}
\end{figure}

The results for other states are similar to those for Washington. In
the appendix, Figures  \ref{fig:MA-mask} and \ref{fig:IL-mask} display
similar results for Massachusetts and Illinois.
Figure \ref{fig:US-mask} shows the average change in cases and
deaths across states, where the top panel shows the effect on cases and the bottom panel shows the effect on deaths.
The point estimates indicate that mandating masks on April 1st
could have led to 25\% fewer cumulative cases and 37\% fewer
cumulative deaths by the end of May with their 90 percent intervals given by $[10,47]$\% and $[18,55]$\%, respectively. The result roughly translates into $18$ to $55$ thousand saved lives.

\subsection{Non-essential Business Closures}

A particularly controversial policy is the closure of non-essential
businesses. We now examine a counterfactual where non-essential
businesses are never closed. Figure \ref{fig:WA-nb} shows the effect
of leaving non-essential businesses open in Washington. The point
estimate implies that the closure of non-essential businesses reduced cases and
deaths by a small amount. However, this estimate is relatively
imprecise; 90\% confidence intervals for the change in cases and deaths from
leaving non-essential businesses open by the end of May are [-250,700] and
[-100,1200], respectively.

\begin{figure}[ht]
  \caption{Effect of leaving non-essential businesses open in Washington \label{fig:WA-nb}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_v1}\\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_deaths_v1}
      & \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_deaths_v1}
    \end{tabular}
  \end{minipage}
\end{figure}

Figure \ref{fig:US-nb} shows the national effect of leaving non
essential businesses open on cases and deaths. For cases, the
estimates imply that with non-essential businesses open, cases would
be about -15 to 60\% higher in late May. The results for deaths are
similar but less precise.

\begin{figure}[ht]
  \caption{Effect of leaving non-essential businesses open in the US\label{fig:US-nb}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths_v1}
    \end{tabular}
  \end{minipage}
\end{figure}

\subsection{Stay-at-home orders}

We next examine a counterfactual where stay-at-home orders had been never
issued.  Figure \ref{fig:US-shelter} shows the average effect of no
stay-at-home orders. On average, without stay-at-home orders, case
growth rate would have been nearly 0.1 higher in late April. This
translates to 80\% [25\%,170\%] more cases by the start of June. The
results for deaths are similar, but slightly less precise, with no
increase included in a 90 percent confidence interval.

\begin{figure}[ht]
  \caption{Effect of having no stay-at-home orders in the US\label{fig:US-shelter}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths_v1}
    \end{tabular}
  \end{minipage}
\end{figure}

\section{Counterfactual Effect of Removing All Policies and its Sensitivity}\label{removing-policies}

%We now consider the impact of changing from the observed policies to
%none.  Figure \ref{fig:WA-nb} shows the effect of removing all
%policies in Washington. The left column shows that without any policy
%intervention, case and death growth would have been higher by
%about 0.1 and 0.2. The confidence intervals are wide relative to the
%mean effect. In the right column, we see that the upper bounds of the
%confidence intervals include large increases in cases, and especially
%deaths.

%\begin{figure}[ht]
%  \caption{Effect of removing all policies in
%    Washington \label{fig:WA-nop}}
%  \begin{minipage}{\linewidth}
%    \centering
%    \begin{tabular}{cc}
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dgrowth}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dcases}
%      \\
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dgrowth_deaths}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dcases_deaths}
%    \end{tabular}
%  \end{minipage}
%\end{figure}

We now consider the impact of changing from the observed policies to none.
Figure \ref{fig:US-nop-SI} shows the average across states of the change
in case growth and relative increase in cases under a specification without past national case variables.
Removing policies leads to an increase of above 0.2 in
case growth throughout April and May. The confidence interval is
fairly wide, and its upper bound includes a very large increase in
cases by the end of May.  The right panel displays the national
increase in aggregate cases without any policy intervention. The
estimates imply at least a 7 fold increase in cases with a large upper bound by the end of May, or
at least 14 million additional cases. The estimated impact on deaths is larger than
cases, and even more imprecise.

The effect of removing all policies includes the effect of school
closures. The visual evidence on growth rates for states with and without school closures,
presented blow, suggest that there may be a potentially large effect, though the history is very short. The main results presented in Section 3 also support the hypothesis that the school closures were important at lowering the growth rates.  This evidence is consistent
with the emerging evidence of prevalence of Covid-19 among children
aged 10-17. \cite{children:nature} find that although children's
transmission and susceptibility rates are half that of ages 20-30,
children's contact rates are much higher.
This type of evidence, as well as, evidence that children carry viral
loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.

\begin{figure}
  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
    \end{tabular}
    \footnotetext{ In these figures, red points are the case or death
      growth rate in states without each policy 14 (or 21 for deaths)
      days earlier. Blue points are states with each policy 14 (or 21
      for deaths) days earlier. The red line is the average across
      states without each policy. The blue line is the average across
      states with each policy.}
  \end{minipage}
\end{figure}

As discussed above, there is little variation across states
in the timing of school closures. Consequently, the effect of school
closures is difficult to identify statistically, because it is hard to separate it from aggregate time effect, and its estimate is sensitive to an inclusion of some aggregate variables such as national cases. To support this point, Figure \ref{fig:US-nop} shows the
effect of removing all policies on cases based on the estimates with national cases
included as information. When national case variables are included
in the specification, the estimated effect of school closures, and hence that of removing all
policies, is much smaller with a 90\% confidence interval of [0,10] fold increases.

Given this sensitivity, we conclude that there still exists a lot of uncertainty as to the effect of removing all policies, especially schooling. The impact of not implementing any policies on cases and deaths can be quite large, but  the effect of school closures, hence that of removing all policies, is not well identified statistically from the US state-level data alone, because of the lack of cross-sectional variations.  Any analyses of re-opening plans need to be aware of this uncertainty.  An important research question is how to resolve this uncertainty using additional data sources.

\begin{figure}[ht]
  \caption{Effect of removing policies on cases in the US under a specification with only state-level cases/deaths as information \label{fig:US-nop-SI}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
%      \\
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths_v1}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths_v1}
    \end{tabular}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
  \caption{Effect of removing policies on cases in the US under a specification with both
  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nop}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel}
%      \\
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
    \end{tabular}
  \end{minipage}
\end{figure}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

\afterpage{\clearpage}

\section{Conclusion}

This paper assesses the effects of policies on the spread of Covid-19 in the US using state-level data on cases, tests, policies, and social distancing behavior measures from Google Mobility Reports. Our findings are summarized as follows.

First, our empirical analysis indicates that mandating face masks has reduced the spread of Covid-19 without affecting people's social distancing behavior measured by Google Mobility Reports. Our counterfactual experiment based on the estimated model suggests that if all states had have adopted mandatory face mask policies on April 1st of 2020, then the number of deaths by the end of May would have been smaller by as much as  $17$ to $55$\%, which roughly translates to $17$ to $55$ thousand saved lives.

Second, we find that keeping non-essential businesses open would have led to -20 to 60\% more cases  while not implementing stay-at-home orders would have increased cases by 25 to 170 \% by the start of June. % Except for mask mandates, the effect of all other policies on case and death growth is realized through their impact on social distancing behavior.

Third, we find considerable uncertainty over  the impact of all policies combined on case or death growth because it is difficult to identify the effect of school closures from the US state-level data due to the lack of variation in the timing of school closures across states.

Fourth, our analysis shows that people voluntarily reduce their visits to workplace, retails, grocery stores, and limit their use of public transit when they receive  information on a higher number of new cases and deaths. This suggests that individuals make decisions to voluntarily limit their contact with others in response to greater transmission risks, leading to an important feedback mechanism that affects future cases and deaths. Model simulations that ignore this voluntary behavioral response to information on transmission risks would over-predict the future number of cases and deaths.

Beyond these findings, our paper presents a useful conceptual framework to investigate the relative roles of policies and information on determining the spread of Covid-19 through their impact on people's behavior. Our causal model allows us  to explicitly define counterfactual scenarios to properly evaluate the effect of alternative policies on the spread of Covid-19. %, and can be used to quantitatively examine various Covid-19 related issues  for future research.
 More broadly, our  causal framework can be useful for quantitatively analyzing not only health outcomes but also various economic outcomes \citep{bartik2020, chetty2020real}.

\FloatBarrier

\begin{footnotesize}

\bibliographystyle{jpe}
\bibliography{covid}

\end{footnotesize}

\newpage

\appendix

\section{Data Construction}

\subsection{Measuring $\Delta C$ and $\Delta\log \Delta C$}

We have three data sets with information on daily cumulative confirmed
cases in each state. As shown in Table \ref{tab:casecor}, these
cumulative case numbers are very highly correlated. However, Table
\ref{tab:casediff} shows that the numbers are different more often
than not.

\input{tables_and_figures/casecor.tex}

\input{tables_and_figures/casediff.tex}

Figure \ref{fig:dailycases} shows the evolution of new cases in each
of these three datasets. In all cases, daily changes in cumulative
cases displays some excessive volatility. This is likely due to delays
and bunching in testing and reporting of results. Table
\ref{tab:newcasecor} shows the variance of log new cases in each data
set, as well as their correlations. As shown, the correlations are
approximately $0.9$. The NYT new case
numbers have the lowest variance.\footnote{This comparison is somewhat
  sensitive to how you handle negative and zero cases when taking
  logs. Here, we replaced $\log(0)$ with $-1$. In our main
  results, we work with weekly new cases, which are very rarely zero.}
In our subsequent results, we will primarily use the case numbers from
The New York Times.

\begin{figure}[!ht]\caption{Daily cases \label{fig:dailycases}}
  \centering
  \begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=0.45\textwidth]{tables_and_figures/newcases}
    \footnotetext{Each line shows daily new cases in a state.}
  \end{minipage}
\end{figure}

\input{tables_and_figures/newcasecor.tex}

For most of our results, we focus on new cases in a week instead of in
a day. We do this for two reasons as discussed in the main text. First, a decline of new cases over
two weeks has become a key metric for decision makers. Secondly, aggregating to weekly new cases smooths out the noise associated with  the timing of
reporting and testing.

Table \ref{tab:weekcasecor} reports the correlation and variance of
weekly log new cases across the three data sets. Figure
\ref{fig:weekcases} shows the evolution of weekly new cases in each
state over time.

\input{tables_and_figures/weekcasecor}

\begin{figure}[!ht]\caption{Weekly Cases \label{fig:weekcases}}
  \centering
  \begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekcases}
    \footnotetext{Each line shows weekly new cases in a state.}
  \end{minipage}
\end{figure}

\subsection{Deaths}

Table \ref{tab:weekdeathcor} reports the correlation and variance of
weekly deaths in the three data sets. Figure \ref{fig:weekdeaths}
shows the evolution of weekly deaths in each state. As with cases, we
use death data from The New York Times in our main results.

\input{tables_and_figures/weekdeathcor}

\begin{figure}[!ht]\caption{Weekly Deaths \label{fig:weekdeaths}}
  \centering
  \begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekdeaths}
    \footnotetext{Each line shows weekly deaths in a state.}
  \end{minipage}
\end{figure}

\begin{figure}[!ht]\caption{Case and death growth \label{fig:growthq}}
  \centering
  \begin{minipage}{\textwidth}
    \centering
    \begin{tabular}{c}
      \includegraphics[width=0.75\textwidth]{tables_and_figures/casequantiles}
      \\
      \includegraphics[width=0.75\textwidth]{tables_and_figures/deathquantiles}
      \footnotetext{Thin gray lines are case or death growth in each
      state and date. Thicker colored lines are quantiles of case or
      death growth conditional on date.}
    \end{tabular}
  \end{minipage}
\end{figure}

\subsection{Tests}

Our test data comes from The Covid Tracking Project. Figure
\ref{fig:test} shows the evolution of tests over time.

\begin{figure}[!ht]\caption{Number of Tests \label{fig:test}}
  \centering
  \begin{minipage}{\textwidth}
    \centering
    \includegraphics[width=0.45\textwidth]{tables_and_figures/test}
    \footnotetext{These figures use the ``total test results''
      reported by The Covid Tracking Project. This is meant to reflect
      the number of people tested
      (\href{https://covidtracking.com/about-data/faq}{as opposed to
        the number of specimens tested}).}
  \end{minipage}
\end{figure}

\FloatBarrier

\subsection{Social Distancing Measures}

In measuring social distancing, we focus on Google Mobility
Reports. This data has international coverage and is publicly
available. Figure \ref{fig:gmr} shows the evolution of the four Google
Mobility Reports variables that we use in our analysis.

%% The following is true but we haven't included any of the
%% results. We can add if needed.
% We also compare Google Mobility Reports to other social
% distancing measures from SafeGraph, Unacast, and PlaceIQ. Each of
% these other data sources are limited to the United States. SafeGraph
% and Unacast data is not publicly available (both companies make the
% data available to researchers at no cost).
\begin{figure}[ht]\caption{Evolution of Google Mobility Reports \label{fig:gmr}}
  \begin{minipage}{\linewidth}
    \begin{tabular}{cc}
      \includegraphics[width=0.5\linewidth]{tables_and_figures/workplaces}
      &
      \includegraphics[width=0.5\linewidth]{tables_and_figures/retail}
      \\
      \includegraphics[width=0.5\linewidth]{tables_and_figures/grocery}
      &
      \includegraphics[width=0.5\linewidth]{tables_and_figures/transit}
    \end{tabular}
    \footnotetext{This figure shows the evolution of Google Mobility
      Reports over time. Thin gray lines are the value of the
      variables in each state and date. Thicker colored lines are
      quantiles of the variables conditional on date.}
  \end{minipage}
\end{figure}

\subsection{Policy Variables}

We use the database on US state policies created by
\cite{raifman2020}. %This database records that each state implemented 25 Covid-19 related policies.
As discussed in the main text, our analysis focuses on seven policies. For stay-at-home orders, closed nonessential
businesses, closed K-12 schools, closed restaurants except takeout,
and closed movie theaters, we double-checked any
state for which \cite{raifman2020} does not record a date. We filled
in a few missing dates. Our modified data is available
\href{"https://docs.google.com/spreadsheets/d/1E6HRkgbdSnZ9ZxrneydU6q4hhOCCt9oTl_5fa3OFVZE/edit?usp=sharing}{here}. Our
modifications fill in 1 value for school closures, 2 for stay-at-home orders, 3 for movie theater closure, and 4 for
non-essential business closures. Table \ref{tab:policies} displays all 25 dated policy variables in
\cite{raifman2020}'s database with our modifications described above.

\afterpage{%
  \clearpage% Flush earlier floats (otherwise order might not be correct)
  \thispagestyle{empty}% empty page style (?)
  \begin{landscape}% Landscape page
    \centering % Center table
    \input{tables_and_figures/policies}
  \end{landscape}
  \clearpage% Flush page
}

\begin{figure}
  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-deaths}
      \\
      \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
      \\
      \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-deaths}
      \\
    \end{tabular}
    \footnotetext{ In these figures, red points are the case or death
      growth rate in states without each policy 14 (or 21 for deaths)
      days earlier. Blue points are states with each policy 14 (or 21
      for deaths) days earlier. The red line is the average across
      states without each policy. The blue line is the average across
      states with each policy.}
  \end{minipage}
\end{figure}

\begin{figure}
  \caption{Case and death growth conditional on policies \label{fig:growthpolicies2}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-deaths}
      \\
      \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-deaths}
      \\
      \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-cases}
      &
        \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-deaths}
    \end{tabular}
    \footnotetext{ In these figures, red points are the case or death
      growth rate in states without each policy 14 (or 21 for deaths)
      days earlier.  Blue points are states with each policy14 (or 21
      for deaths) days earlier.  The red line is the average across
      states without each policy. The blue line is the average across
      states with each policy.}
  \end{minipage}
\end{figure}

\subsection{Timing\label{sec:timing}}

There is a delay between infection and when a person is tested and
appears in our case data. \cite{midas2020} maintain a list of
estimates of the duration of various stages of Covid-19
infections. The incubation period, the time from infection to symptom
onset, is widely believed to be 5 days. For example, using data from
Wuhan, \cite{li2020} estimate a mean incubation period of 5.2 days.
\cite{siorda2020} reviews the literature and concludes the mean
incubation period is 3-9 days.

Estimates of the time between symptom onset and case reporting or death
are less common. Using Italian data, \cite{cereda2020} estimate an
average of 7.3 days between symptom onset and
reporting. \cite{zhang2020} find an average of 7.4 days using Chinese data
from December to early February, but they find this period declined
from 8.9 days in January to 5.4 days in the first week of
February. Both of these papers on time from symptom onset to reporting
have large confidence intervals covering approximately 1 to 20 days.

Studying publicly available data on infected persons diagnosed outside of Wuhan,
\cite{linton2020}  estimate an average of 15 days  from  onset to death. Similarly, using publicly available reports of 140 confirmed Covid-19 cases in China, mostly outside Hubei Province, \cite{sanche2020} estimate the time from onset to death to be  16.1 days.

Based on the above, we expect a delay of roughly two weeks between
changes in behavior or policies, and changes in reported
cases while a corresponding delay of roughly three weeks for deaths.

% Our growth rate variable,
% $$
% {\ycolor Y_{it}} = \Delta \log \Delta C_{t} = \log(C_{t} - C_{t-7}) - \log(C_{t-7}-C_{t-14})
% $$
% reflects an average growth rate of cases from time $t-7$ to $t$.\footnote{This
% can be seen applying the intermediate value theorem to the dynamics
% of the SIR model from section \label{sec:sirmodel}. Specifically,
% \begin{align*}
%   C_{t} - C_{t-7} = & \int_{t-7}^t \dot{C}(s) ds \\
%   = & \int_{t-7}^t \tau(s) I(s) ds \\
%   = & \overline{\tau}(t-7,t) I(t-7) \int_{t-7}^t \exp(\int_{t-7}^s \frac{S(r)}{N}
%       \beta(r) - \gamma(r) dr \right) ds \\
%   = & \overline{\tau}(t-7,t) \overline{\exp(\frac{S}{N} \beta - \gamma)}(t-7,t) I(t-7) \\
%   = & \overline{\tau}(t-7,t) \overline{\exp(\frac{S}{N} \beta -
%       \gamma)}(t-7,t)
%       \overline{\exp(\frac{S}{N} \beta - \gamma)}(t-14,t-7) I(t-14)
% \end{align*}
% where $\gamma(r) = \frac{S(r-\ell)}{N}\beta(r-\ell)$ and expressions
% with a line over them are constants that exist by the intermediate
% value theorem. $\overline{\exp(\frac{S}{N} \beta - \gamma)}(t-7,t)$ is
% a constant growth rate from $t-7$ to $t$ that gives rise to the same
% number of infections at time $t$ as the original, potentially time
% varying, growth rate. Applying similar reasoning to
% $C_{t-7} - C_{t-14}$ and combining gives
% \begin{align*}
%   {\ycolor Y_{it}} = \log(C_{t} - C_{t-7}) - \log(C_{t-7}-C_{t-14})
%   = & \overline{\exp(\frac{S}{N} \beta -  \gamma)}(t-7,t) +
%       \log(\overline{\tau}(t-7,t)) -  \log(\overline{\tau}(t-14,t-7)).
% \end{align*}
% }

\FloatBarrier

\subsection{Counterfactuals for Massachusetts and
  Illinois}
 Figures \ref{fig:MA-mask} and \ref{fig:IL-mask} present the fit of estimated cases as well as the counterfactual effect of mandating masks on April 1st in Massachusetts and Illinois, respectively. Figures \ref{fig:MA-nb} and \ref{fig:IL-nb}   show the counterfactual effect of leaving non-essential business open in Massachusetts and Illinois, respectively.

\FloatBarrier

\begin{figure}[b]
  \caption{Effect of mandating masks on April 1st in Massachusetts \label{fig:MA-mask}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{ccc}
      \multicolumn{3}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_v1}
      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_v1}
\\
      \multicolumn{3}{c}{\textbf{Deaths}} \\
     \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth_deaths}
      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_deaths_v1}
      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_deaths_v1}
    \end{tabular}
    \begin{flushleft}
      \footnotesize To compute the estimated and counterfactual paths
      we use the average of two estimated coefficients as reported in
      column ``Average'' of Table \ref{tab:dieff}. We set initial
      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
      observed in the state we are simulating. We hold all other
      regressors at their observed values. Error terms are drawn with
      replacement from the residuals. We do this many times and report
      the average over draws of the residuals. The shaded region is a
      point-wise 90\% confidence interval.
    \end{flushleft}
  \end{minipage}
\end{figure}

\begin{figure}[h]
  \caption{Effect of mandating masks on April 1st in Illinois \label{fig:IL-mask}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{ccc}
      \multicolumn{3}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth}
      &
        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_v1}
      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_v1}      \\
      \multicolumn{3}{c}{\textbf{Deaths}} \\
    \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth_deaths}
      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_deaths_v1}
      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_deaths_v1}
    \end{tabular}
    \begin{flushleft}
      \footnotesize To compute the estimated and counterfactual paths
      we use the average of two estimated coefficients as reported in
      column ``Average'' of Table \ref{tab:dieff}. We set initial
      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
      observed in the state we are simulating. We hold all other
      regressors at their observed values. Error terms are drawn with
      replacement from the residuals. We do this many times and report
      the average over draws of the residuals. The shaded region is a
      point-wise 90\% confidence interval.
    \end{flushleft}
  \end{minipage}
\end{figure}

\begin{figure}[h]
  \caption{Effect of leaving businesses open in Massachusetts \label{fig:MA-nb}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_v1}
       &
         \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_v1}
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_deaths_v1}
       &      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_deaths_v1}
    \end{tabular}
  \end{minipage}
\end{figure}

\begin{figure}[h]
  \caption{Effect of leaving businesses open in Illinois \label{fig:IL-nb}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_v1}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_v1} \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
           \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_deaths_v1}
      &       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_deaths_v1}    \end{tabular}
  \end{minipage}
\end{figure}

%\begin{figure}[h]
%  \caption{Effect of removing all policies in
%    Illinois \label{fig:IL-nop}}
%  \begin{minipage}{\linewidth}
%    \centering
%    \begin{tabular}{cc}
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dgrowth_v1}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dcases_v1}
%      \\
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dgrowth_deaths_v1}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dcases_deaths_v1}
%    \end{tabular}
%  \end{minipage}
%\end{figure}
%
%\begin{figure}[h]
%  \caption{Effect of removing all policies in
%    Massachusetts \label{fig:MA-nop}}
%  \begin{minipage}{\linewidth}
%    \centering
%    \begin{tabular}{cc}
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dgrowth_v1}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dcases_v1}
%      \\
%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dgrowth_deaths_v1}
%      &
%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dcases_deaths_v1}
%    \end{tabular}
%  \end{minipage}
%\end{figure}

\FloatBarrier

%\subsection{Policy Effects without National Cases as Information Variables}
%
%Table \ref{tab:dieff-si} presents the estimates under a specification that excludes national information variables. Comparing the estimates in table \ref{tab:dieff}, the effects of school closures and  the sum of policies are estimated substantially larger in table \ref{tab:dieff-si} when national information variables are excluded.
%
%%Here we present the results of counterfactual analyses that don't include the national cases/deaths as
%%the the information variables.  The growth in the national cases nearly perfectly predicts school closure
%%policies, which have little cross-sectional variation.  Here we remove the national cases from the information
%%and check what happens to the counterfactual results.  The counterfactual results of mask policies, shelter-in-place,
%%and closing
%%non-essential businesses remain robust. The results on school closures effects change substantially,
%%attributing a much bigger role to school closures. The resulting counterfactual of  removing
%%all policies has a much larger effect on cases, with the lower bound of the confidence band implying at
%%least a 7 fold increase in cases (see Figure \ref{fig:US-nop-SI} given in the main text).
%
%\begin{table}[!b]
%  \caption{\label{tab:dieff-si}Direct and Indirect Policy Effects
%    (without national information variables)}
%\begin{minipage}{\linewidth}
%  \centering
%    \scriptsize
%  \begin{tabular}{c}
%    \textbf{Cases}
%    \\
%    \input{tables_and_figures/dieff-cases-SI}
%    \\
%    \textbf{Deaths}
%    \\
%    \input{tables_and_figures/dieff-deaths-deathsinfo-SI}
%  \end{tabular}
%  \smallskip
%  \begin{flushleft}
%    {\tiny Direct effects capture the effect of policy on case
%      growth holding behavior, information, and confounders
%      constant. Direct effects are given by ${\pcolor \pi}$ in
%      equation (\ref{eq:R1}). Indirect effects capture how policy
%      changes behavior and behavior shift case growth. They are given
%      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
%      from (\ref{eq:R2}). The total effect is
%      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
%      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
%      \ref{eq:R4}. Columns ``Difference''    are  the differences between
%      the estimates from (\ref{eq:R4}) and the combination of
%      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
%      Standard errors are computed by
%      bootstrap and clustered on state.}
%    %  \ref{tab:PtoY}.
%    \end{flushleft}
%  \end{minipage}
%\end{table}

\FloatBarrier

\subsection{Counterfactuals with National Cases as Information
  Variables}

Figures \ref{fig:US-mask-SI}-\ref{fig:US-shelter-SI}  present the results of counterfactual analyses that include the national cases/deaths as the information variables. To create this figure, we repeat the same counterfactual
simulation that we did for Washington with each state. For each state,
we hold national cases constant, but endogenize state specific
information. Thus, these figures should be interpreted as an average
of state specific counterfactuals, and not a national counterfactual.

 The counterfactual results of mask policies, shelter-in-place,
 and closing non-essential businesses remain robust with respect to the inclusion of national case/death variables. This contrasts to the resulting counterfactual of  removing all policies  discussed in section \ref{removing-policies}.

\begin{figure}[!b]
  \caption{Effect of mandating masks for employees on April
    1st
     under a specification with both
  state-level cases/deaths and national-level cases/deaths as information
    \label{fig:US-mask-SI}}
  \begin{minipage}{\linewidth}
    \centering
    \medskip
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth}
      &
        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths}
      &
        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths}
    \end{tabular}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
  \caption{Effect of leaving non-essential businesses open      under a specification with both
  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nb-SI}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths}
    \end{tabular}
  \end{minipage}
\end{figure}

\begin{figure}[ht]
  \caption{Effect of having no stay-at-home orders     under a specification with both
  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-shelter-SI}}
  \begin{minipage}{\linewidth}
    \centering
    \begin{tabular}{cc}
      \multicolumn{2}{c}{\textbf{Cases}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel}
      \\
      \\
      \multicolumn{2}{c}{\textbf{Deaths}} \\
      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths}
      &
        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths}
    \end{tabular}
  \end{minipage}
\end{figure}

\FloatBarrier

$\;$

\end{document}

### 2020-07-16 tex/covidEffectOfNPI.tex
Version submitted in early July 2020.
Compared to submitted version the following code changes have been made:

- Removed safegraph and unacast data (they can't be publicly shared)
- Minor update to policy data loading due to data format change
- Added explicit subsetting of data to June 2, 2020 and earlier (to match what we used in submitted paper).

--- 2020-07-16 tex/covidEffectOfNPI-COVIDEconomics.tex
@@
 We also find that, without stay-at-home orders, cases would have been larger by  25 to 170 percent, which implies that 0.5 to 3.4 million more Americans could have been infected if stay-at-home orders had not been implemented.
-Finally,  not having implemented any policies could have led to at least a 7 fold increase with an uninformative upper bound in cases (and deaths) by the end of May  in the US, with considerable uncertainty over the effects of school closures, which had little cross-sectional variation.%  (the potential range of effects could be even larger the effect of school closures on behavior is hard to identify separately from aggregate time effects).
+Finally,  we find considerable uncertainty over the effects of removing all policies because the timing of school closures had little cross-sectional variation.%  (the potential range of effects could be even larger the effect of school closures on behavior is hard to identify separately from aggregate time effects).
 \end{abstract}
@@
 %\tableofcontents
-\thispagestyle{empty}
-
-\pagebreak
-
-\pagestyle{empty}
 
@@
 It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
-Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007}; see \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book} for  modern developments, especially in computer science and epidemiology), with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-As such it naturally converses in all languages for causal inference.}
+Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{
+The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007} and genetics \citep{wright1923}, with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+See \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book}  for modern developments, especially in computer
+science and epidemiology. The particular causal diagram has several “mediation” components, where variables affect
+outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
+The father and son,
+P. Wright and S. Wright closely collaborated to develop structural equation models and causal path diagrams; P.Wright’s
+key work represented supply-demand system as a directed acyclical graph and established its identification using
+exclusion restrictions on instrumental variables.
+}
+%The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007}.
+See \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book} for  modern developments, especially in computer science and epidemiology), with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+As such it naturally converses in all languages for causal inference.
+
 % It also allows us  to explicitly define counterfactual scenarios to properly evaluate the effect of alternative policies on the spread of Covid-19.
@@
 
-A growing number of other papers have examined the link between non-pharmaceutical interventions and Covid-19  cases.\footnote{We refer the reader to \cite{avery2020policy} for a comprehensive review of  a larger body of work researching Covid-19; here we focus on few quintessential comparisons on our work with other works that we are aware of.} \cite{hsiang2020} estimate the effect of policies on the growth rate of cases using
+A growing number of other papers have examined the link between non-pharmaceutical interventions and Covid-19  cases.\footnote{We refer the reader to \cite{avery2020} for a comprehensive review of  a larger body of work researching Covid-19; here we focus on few quintessential comparisons on our work with other works that we are aware of.} \cite{hsiang2020} estimate the effect of policies on the growth rate of cases using
 data from the United States, China, Iran, Italy, France, and South
@@
 Epidemiologists use model simulations to predict how cases and deaths evolve for the purpose of policy recommendation.
-As reviewed by \cite{avery2020policy}, there exist substantial uncertainty about the values of key epidimiological parameters \citep[see also][]{atkeson2020b,stock2020}.
+As reviewed by \cite{avery2020}, there exist substantial uncertainty about the values of key epidimiological parameters \citep[see also][]{atkeson2020b,stock2020}.
 Simulations are often done under strong assumptions about the impact of social distancing policies without connecting to the relevant data  \citep[e.g.,][]{ferguson2020}.

### 2020-07-16 cases_and_policies/R/cases_and_policies.R

@@ -1,0 +1,286 @@
+library(rvest)
+library(gsheet)
+
+stub_ <- function() {}
+thisPath <- function() {
+  cmdArgs <- commandArgs(trailingOnly = FALSE)
+  if (length(grep("^-f$", cmdArgs)) > 0) {
+    # R console option
+    normalizePath(dirname(cmdArgs[grep("^-f", cmdArgs) + 1]))[1]
+  } else if (length(grep("^--file=", cmdArgs)) > 0) {
+    # Rscript/R console option
+    scriptPath <- normalizePath(dirname(sub("^--file=", "", cmdArgs[grep("^--file=", cmdArgs)])))[1]
+  } else if (Sys.getenv("RSTUDIO") == "1") {
+    # RStudio
+    dirname(rstudioapi::getSourceEditorContext()$path)
+  } else if (is.null(attr(stub_, "srcref")) == FALSE) {
+    # 'source'd via R console
+    dirname(normalizePath(attr(attr(stub_, "srcref"), "srcfile")$filename))
+  } else {
+    "."
+  }
+}
+
+gitrootdir <- function() {
+  rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
+  if (is.na(rootdir)) {
+    thisfilepath <- thisPath()
+    rootdir <- normalizePath(paste(thifilepath,"../..",sep="/"))
+  }
+  validationfile <- normalizePath(paste(rootdir,"cases_and_policies/",sep="/"))
+  if (!file.exists(validationfile)) {
+    stop("Could not determine location of git tree for covid project. Setting R's working directory to anywhere within the git repo should resolve this problem.")
+  }
+  return(rootdir)
+}
+
+rootdir <- gitrootdir()
+
+# Data on policies from Raifman and collaborators
+urls <- c("https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=993060716",
+          "https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=1357478819",
+          "https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=1894978869",
[…]

### 2020-07-16 cases_and_policies/R/dataprep.R

@@ -1,0 +1,62 @@
+## This will break if you didn't check out the git repo.
+## If it breaks, just set rootdir to whatever you need
+rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
+datafile <- paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/")
+cat("Loading data last updated on ", as.character(file.mtime(datafile)), "\n")
+cat("Run ", paste(rootdir,"cases_and_policies/R/cases_and_policies.R", sep="/"), " to update data.\n")
+load(paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/"))
+source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
+
+cs <- covidstates
+cs <- cs[cs$fips<60, ] # drop territories (Puerto Rico, Guam, etc)
+cs <- cs[order(cs$state, cs$date),]
+
+## Outcome variable
+L <- 7
+dlogd <- function(x, id, t, lag=L, minval=0.1) {
+  dx  <- paneldiff(x, id, t, lag=lag)
+  dx[dx<minval] <- minval
+  dlogdx <- paneldiff(log(dx),id,t,lag=lag)
+}
+
+cs$dlogdc <- dlogd(cs$cases.nyt, cs$state, cs$date, lag=L)
+cs$dlogdc.jhu <- dlogd(cs$cases, cs$state, cs$date, lag=L)
+cs$dcases <- paneldiff(cs$cases.nyt,cs$state, cs$date, lag=L)
+cs$dcases.jhu <- paneldiff(cs$cases,cs$state, cs$date, lag=L)
+
+## Social distancing measures
+newnames <- c("residential","transit","workplaces","retail","grocery")
+idx <- sapply(newnames,function(x) grep(paste(x,"_",sep=""), names(cs)))
+for(v in newnames) {
+  cs[,v] <- panelma(cs[,idx[v]], cs$state, cs$date, len=L-1)
+  #cs[,v] <- panellag(cs[,v], cs$state, cs$date, lag=L)
+}
+
+## Policy variables
+bw <- 7
+newnames <- c("movie","shelter","k12","restaurant","nonessential")
+oldnames <- c("Closed.movie.theaters","Stay.at.home..shelter.in.place",
[…]

### 2020-07-16 cases_and_policies/R/exploratory.R

@@ -1,0 +1,189 @@
+library(ggplot2)
+library(ggthemes)
+library(xtable)
+library(gridExtra)
+
+rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
+datafile <- paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/")
+cat("Loading data last updated on ", as.character(file.mtime(datafile)), "\n")
+cat("Run ", paste(rootdir,"cases_and_policies/R/cases_and_policies.R", sep="/"), " to update data.\n")
+load(datafile)
+source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
+outdir <- paste(rootdir,"tex/tables_and_figures", sep="/")
+
+plottheme <- theme_pander() + theme(plot.title=element_text(face="plain"))
+#figtheme
+colors <-  scale_color_solarized
+
+cs <- covidstates
+cs <- cs[cs$fips<60, ] # drop territories (Puerto Rico, Guam, etc)
+cs <- cs[order(cs$state, cs$date),]
+
+################################################################################
+## Case Measures
+vars <- c("cases.nyt","cases","positive")
+tbl <- cor(cs[,vars],use="complete")
+rownames(tbl) <- colnames(tbl) <- c("NYT","JHU","CTP")
+cat(print(xtable(tbl, digits=5, caption="Correlation of cumulative cases \\label{tab:casecor}")),
+    file=paste(outdir, "casecor.tex", sep="/"))
+
+tbl2 <- matrix(NA,3,3)
+for(i in 1:3) {
+  for(j in 1:3) {
+    tbl2[i,j] <- mean(cs[,vars[i]]==cs[,vars[j]], na.rm=TRUE)
+  }
+}
+rownames(tbl2) <- rownames(tbl)
+colnames(tbl2) <- colnames(tbl2)
+cat(print(xtable(tbl2, digits=2, caption="Portion of cumulative cases that are equal between data sets\\label{tab:casediff}")),
+    file=paste(outdir, "casediff.tex", sep="/"))
+
+cs$newcases.nyt <- paneldiff(cs$cases.nyt, cs$state, cs$date, lag=1)
+cs$newcases.jhu <- paneldiff(cs$cases, cs$state, cs$date, lag=1)
[…]

### 2020-07-16 cases_and_policies/R/mainregressions.R

@@ -1,0 +1,113 @@
+library(lfe)
+library(plm)
+
+rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
+source(paste(rootdir,"tex/dataprep.R",sep="/"))
+
+cs$t <- as.numeric(cs$date)
+df <- pdata.frame(cs, index=c("state","t"), stringsAsFactors=FALSE)
+
+behave <- c("residential","transit","workplaces","retail","grocery")
+policy <- c("pmovie_7","pshelter_7","pk12_7","prestaurant_7","pnonessential_7")
+cor(df[,c(behave,policy)], use="pairwise.complete.obs")
+
+df$tau  <- df$tests/df$Population.2018*1000
+df$taudc <- (df$ddeath.jhu/df$dcases.jhu)*df$tau
+df$trend <- log(as.numeric(cs$date - as.Date("2020-01-29")))
+df$Y <- df$dlogdc.jhu
+df$Y[lag(df$cases,7)<10 ] <- NA
+df$taudc <- df$ddeath.jhu/df$dcases.jhu*df$tau
+df$taudc[df$cases<10] <- NA
+# normalize taudc
+df$taudc <- df$taudc*1000
+
+# Table 1
+yvar <- "Y"
+xvars <- c("dlogtests",
+           "lag(tau,14)",
+           "lag(logdc.jhu,14)",
+           "trend",
+           "lag(Y,7)")
+interactions <- list(c("lag(taudc,7)"),
+                     c("log(Population.2018)",
+                       "log(Square.Miles)",
+                       "Percent.Unemployed..2018.",
+                       "Percent.living.under.the.federal.poverty.line..2018.",
+                       "Percent.at.risk.for.serious.illness.due.to.COVID"))
+cluster  <- "state"
+smoothedpolicy <- function(df, pdatevar, type=c("pnorm","ma"), bw=7) {
+  if (type=="pnorm") {
+    p <- as.numeric(pnorm(df$date, df[,pdatevar], bw))
+    p[is.na(p)] <- 0.0
+  } else if(type=="ma") {
+    d <- df$date >= df[,pdatevar]
+    d[is.na(d)] <- FALSE
+    p <- panelma(d, df$state, df$date, len=bw)
+  }
+  return(p)
+}
+
+createfmla <- function(yvar, xvars, interactions, fe = "0", iv="0", cluster="state") {
[…]

### 2020-07-16 cases_and_policies/R/plot_policies.R

@@ -1,0 +1,57 @@
+library(ggplot2)
+library(gridExtra)
+load(file="../data/covidstates.Rda")
+## Some summary plots to check
+df  <- covidstates
+plotstatepolicies  <- function(st) {
+  sdf  <- subset(df, state==st)
+  #sdf$cases  <- pmax(sdf$cases, 1) # to avoid log(0)
+  pols <- c("State.of.emergency",
+            "Date.closed.K.12.schools",
+            "Closed.day.cares",
+            "Date.banned.visitors.to.nursing.homes",
+            "Stay.at.home..shelter.in.place",
+            "Closed.non.essential.businesses",
+            "Closed.restaurants.except.take.out",
+            "Closed.gyms",
+            "Closed.movie.theaters",
+            "Order.freezing.utility.shut.offs",
+            "Froze.mortgage.payments"
+            )
+  x  <- c()
+  labels <- c()
+  for (lbl in pols) {
+    d <- unique(sdf[,lbl])
+    i <- which(x==d)
+    if (length(i)==0) {
+      if (length(x)==0)
+        x <- d
+      else
+        x <- c(x,d)
+      labels <- c(labels,lbl)
+    } else {
+      labels[i]  <- paste(labels[i],lbl,sep=", ")
+    }
+  }
+  labels <- labels[!is.na(x)]
+  x <- x[!is.na(x)]
+  y <- rep(1, length(x))
+  fig <- ggplot(sdf,aes(x=date))  +  theme_minimal() +
+    geom_line(aes(y=cases)) +
+    geom_line(aes(y=deaths), colour="red") +
+    xlim(as.Date("2020-03-01"),max(sdf$date)) +
+    ylab("count") +
+    scale_y_continuous(trans="log10") +
+    ggtitle(sprintf("Cases, deaths, and policies in %s",st))
+
+  fig + annotate("text", x=x,y=y,label=gsub("\\."," ",labels),
+             size=2, angle=90, hjust=0)
+
+}
+
+fig <- list()
+for (st in c("Washington","New York","Florida","Alabama","Wyoming","California"))
+  fig[[st]] <- plotstatepolicies(st)
+pdf("state_policies.pdf", width=14,height=28)
+grid.arrange(grobs=fig, nrow=length(fig))
+dev.off()

### 2020-07-16 cases_and_policies/R/utils.R

@@ -1,0 +1,105 @@
+#' Create lags for panel data.
+#'
+#' This function creates lags (or leads) of panel data variables.
+#' Input data should be sorted by i, t --- e.g.
+#' df <- df[order(i,t),]
+#' @param x Vector or matrix to get lags of.
+#' @param i unit index
+#' @param t time index
+#' @param lag How many time periods to lag. Can be negative if leading
+#' values are desired.
+#' @return Lagged copy of x.
+panellag <- function(x, i, t, lag=1) {
+  if (!identical(order(i,t),1:length(i))) {
+    stop("inputs not sorted.")
+  }
+  if (is.matrix(x)) {
+    return(apply(x,MARGIN=2,FUN=function(c) { panel.lag(c,i,t,lag) }))
+  }
+  if (length(i) != length(x) || length(i) != length(t) ) {
+    stop("Inputs not same length")
+  }
+  if (lag>0) {
+    x.lag <- x[1:(length(x)-lag)]
+    x.lag[i[1:(length(i)-lag)]!=i[(1+lag):length(i)] ] <- NA
+    x.lag[t[1:(length(i)-lag)]+lag!=t[(1+lag):length(i)] ] <- NA
+    val <- (c(rep(NA,lag),x.lag))
+  } else if (lag<0) {
+    lag <- abs(lag)
+    x.lag <- x[(1+lag):length(x)]
+    x.lag[i[1:(length(i)-lag)]!=i[(1+lag):length(i)] ] <- NA
+    x.lag[t[1:(length(i)-lag)]+lag!=t[(1+lag):length(i)] ] <- NA
+    val <- (c(x.lag,rep(NA,lag)))
+  } else { # lag=0
+    return (x)
+  }
+  if (inherits(x,"Date") & class(val)=="numeric") {
+    stopifnot(0==as.numeric(as.Date("1970-01-01")))
+    val <- as.Date(val, origin="1970-01-01")
+  }
+  return(val)
+}
+
+#' Create differences of panel data.
+#'
+#' @param x Vector or matrix to get differences of.
+#' @param id unit index
+#' @param t time index
+#' @param lag How many time periods to lag. Can be negative if leading
+#' values are desired.
+#' @return Differenced x.
+paneldiff <- function(x, id, t, lag=1) {
+  return(x - panellag(x,id,t,lag=lag))
+}
+
+
+#' Create moving average from panel data
+#'
[…]

### 2020-07-16 cases_and_policies/rmd/bootstrap_felm.R

@@ -1,0 +1,58 @@
+#' Gaussian multiplier bootstrap for a collection of felm models
+#'
+#'
+bootstrap_felm <- function(data, models, S=999) {
+
+
+  data[,"rowid"] <- 1:nrow(data)
+  fmlas <- lapply(models, function(x) x$formula)
+
+  # find cluster identifier
+  cvar <- unique(sapply(fmlas, function(x) {
+    trimws(strsplit(as.character(x)[3], "\\|")[[1]][4])
+  }))
+  cvar[is.na(cvar)] <- "rowid" # no clustering
+  stopifnot(length(cvar)==1) # formulas have different clustering
+
+  # make sure keepCX==TRUE
+  models <- lapply(models, function(m) {
+    if (m$keepCX) return(m)
+    m <- felm(m$formula, data=data, keepCX=TRUE)
+  })
+
+
+  # extract index of observations in each model
+  idx <- lapply(models, function(m) {
+    yvar <- as.character(m$formula)[2]
+    data[,"rid"] <- data$rowid
+    data$rid[is.na(data[,yvar])] <- NA
+    f <- update.felm.formula(m$formula, rid ~ .)
+    return(felm(f, data=data)$response)
+  })
+  for (i in 1:length(models)) stopifnot(length(idx[[i]])==length(models[[i]]$cY))
+
+
+  betahat <- lapply(models, function(m) m$coef)
+  betastar <- list()
+
+  # draw multipliers
+  cvals <- unique(data[,cvar])
+  gc <- matrix(rnorm(S*length(cvals)), nrow=S)
+  colnames(gc) <- cvals
+  gi <- gc[,data[,cvar]]
+
+  # bootstrap
+  for (i in 1:length(models)) {
+    X <- models[[i]]$cX[,!is.na(models[[i]]$coef)]
+    score <- matrix(0, nrow=nrow(data), ncol=ncol(X))
+    iH <- solve(t(X) %*% X)
+    score[idx[[i]],] <-X * as.vector(models[[i]]$residuals)
+    betastar[[i]] <- matrix(NA, nrow=length(betahat[[i]]), ncol=S)
+    betastar[[i]][!is.na(betahat[[i]]),] <-
+      as.vector(betahat[[i]][!is.na(betahat[[i]])]) +
+      iH %*% t(score) %*% t(gi)
+    rownames(betastar[[i]]) <- rownames(betahat[[i]])
+  }
+
+  return(betastar)
+}

### 2020-07-16 cases_and_policies/rmd/generatetables.R

@@ -1,0 +1,401 @@
+#' Wrapper for stargazer that saves latex output and displays html table
+#'
+regtable <- function(res, landscape=FALSE, texfile=NULL,
+                     add.lines=NULL, ...) {
+  if (sgtype=="latex" & !is.null(add.lines)) {
+    tbl  <- capture.output(stargazer(res, type=sgtype, style=sgstyle,
+                                     df=FALSE, header=FALSE,
+                                     add.lines=add.lines,
+                                     no.space=TRUE,
+                                     p=c(0.1,0.05,0.01), ...))
+  } else {
+    # add.lines doesn't work for other output types ...
+    tbl <- capture.output(stargazer(res, type=sgtype, style=sgstyle,
+                                    df=FALSE, header=FALSE,
+                                    p=c(0.1,0.05,0.01),
+                                    ...))
+  }
+  if (landscape && knitr::is_latex_output()) {
+    cat("\\afterpage{
+    \\clearpage
+    \\thispagestyle{empty}
+    \\begin{landscape}
+")
+  }
+  tbl <- relabel(tbl)
+  tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
+  cat(paste(relabel(tbl),collapse="\n"))
+  if (landscape && knitr::is_latex_output()) {
+    cat("    \\end{landscape}
+    \\clearpage
+}")
+  }
+  if (!is.null(texfile)) {
+    tbl  <- capture.output(stargazer(res, type="latex", style=sgstyle,
+                                     df=FALSE, header=FALSE,
+                                     add.lines=add.lines,
+                                     no.space=TRUE,
+                                     #p=c(0.1,0.05,0.01),
+                                     ...))
+    tbl <- relabel(tbl)
+    tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
+    tbl <- gsub("No","Yes", tbl)
+    cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
+  }
+  return(tbl)
+};
+
+#' Create formula for use with felm
+#'
+createfmla <- function(yvar, xvars, interactions, fe = "0", iv="0", cluster="state") {
+  rhs <- paste(xvars, collapse=" + ")
[…]

### 2020-07-16 cases_and_policies/rmd/regprep.R

@@ -1,0 +1,124 @@
+cs$t <- as.numeric(cs$date)
+
+warning("Using data only up to 2020-06-02 as in submitted version of paper. Modify lines 4-5 of regprep.R to change")
+df <- pdata.frame(subset(cs, cs$date>=as.Date("2020-01-01") &
+                             cs$date<=as.Date("2020-06-02")),
+                  index=c("state","t"), stringsAsFactors=FALSE)
+
+#df$tests[df$testmissing] <- 0
+df$testrate  <- df$tests/df$Population.2018*1000
+df$testrate[df$testrate<0] <- NA
+
+dlogd <- function(x, id, t, lag=L, minval=0.0, b=0, g=0) {
+  dx  <- paneldiff(x, id, t, lag=lag)
+  dx[dx<minval] <- minval
+  dlogdx <- log(dx + exp(g)*b) - panellag(log(dx+b),id,t,lag=lag)
+  dlogdx[!is.finite(dlogdx)] <- NA
+  return(dlogdx)
+}
+df$cases.jhu <- df$cases
+df$deaths.jhu <- df$deaths
+
+df$cases <- df$cases.nyt
+df$cases[is.na(df$cases)] <- df$cases.jhu[is.na(df$cases)]
+df$cases[is.na(df$cases)] <- df$positive[is.na(df$cases)]
+
+df$deaths <- df$deaths.nyt
+df$deaths[is.na(df$deaths)] <- df$deaths.jhu[is.na(df$deaths)]
+df$deaths[is.na(df$deaths)] <- df$deaths.ctp[is.na(df$deaths)]
+
+
+df$dcases <- paneldiff(df$cases,df$state, df$date, lag=L)
+df$ddeath <- paneldiff(df$deaths,df$state, df$date, lag=L)
+
+df$logdc <- log(sapply(df$dcases, function(x) max(x, exp(-1))))
+
+df$dlogdc <- dlogd(df$cases, df$state, df$date, lag=L,
+                   minval=exp(-1))
+
+df$dlogdd <-  dlogd(df$deaths, df$state, df$date, lag=L,
+                    minval=exp(-1))
+
+df$logdd <- log(sapply(df$ddeath, function(x) max(x, exp(-1))))
+
+df$dlogtests <- paneldiff(log(sapply(df$testrate, function(x)
+                                     max(x,exp(-1)))), df$state, df$date, lag=L)
+df$testratedc <- df$ddeath/df$dcases*df$testrate
[…]

### 2020-07-16 cases_and_policies/rmd/varlabels.R

@@ -1,0 +1,170 @@
+varlabels <-
+  data.frame(
+  t(matrix(c(
+    "dlogdc", "$\\Delta \\log \\Delta C_{it}$",
+    "log(cases[t]-cases[t-7]) - log(cases[t-7] - cases[t-14])", "NYT",TRUE,
+
+    "dlogdd", "$\\Delta \\log \\Delta D_{it}$",
+    "log(deaths[t]-deaths[t-7]) - log(deaths[t-7] - deaths[t-14])", "NYT",TRUE,
+
+    "dlogtests", "$\\Delta \\log T_{it}$",
+    "log(tests from t to t-7) - log(tests from t-7 to t-14)",
+    "Covid Tracking Project",TRUE,
+
+    "testrate", "tests per 1000", "tests in past week per 1000 people",
+    "Covid Tracking Project",TRUE,
+
+    "logdc", "$\\log \\Delta C_{it}$",
+    "log(cases[t]-cases[t-7])", "NYT", TRUE,
+
+    "logdd", "$\\log \\Delta D_{it}$",
+    "log(deaths[t]-deaths[t-7])", "NYT", TRUE,
+
+    "logt", "log(days since 2020-01-15)", "log(days since January 15, 2020)", "", TRUE,
+
+    "testratedc", "$\\frac{\\Delta{D}}{\\Delta{C}} T_{it}$",
+    "tests in past week time deaths in past week divided by cases in past week",
+    "NYT, Covid Tracking Project",TRUE,
+
+    "Population.2018", "population", "Population in 2018",
+    "Raifman et al",TRUE,
+
+    "Square.Miles", "square miles", "State area in square miles",
+    "Raifman et al",TRUE,
+
+    "Percent.Unemployed..2018.", "Unemployment",
+    "State unemployment in 2018", "Raifman et al",TRUE,
+
+    "Percent.living.under.the.federal.poverty.line..2018.", "Poverty",
+    "% under poverty line in 2018", "Raifman et al",TRUE,
+
+    "Percent.at.risk.for.serious.illness.due.to.COVID", "Comorbidity",
+    "% at risk for serious illness due to COVID", "Raifman et al",TRUE,
+
+    "pmovie", "closed movie theaters",
+    "closed movie theaters", "Raifman et al",TRUE,
+
+    "pshelter", "stay at home",
+    "stay at home / shelter in place", "Raifman et al",TRUE,
+
[…]

### 2020-07-21 cases_and_policies/rmd/datasummary.Rmd
new file:   covid.bib 	modified:   datasummary.Rmd

@@
 gv <- dataguide$name[dataguide$source=="Google Mobility Reports"]
+gv <- gv[grep("_change_from_baseline",gv)]
 hv <- dataguide$name[dataguide$source=="Homebase"]
-uv <- c("daily_distance_diff", "daily_visitation_diff",
-        "encounters_rate", "log_encounters_rate")
-cs$log_encounters_rate <- log(1+cs$encounters_rate)
 ```

### 2020-07-23 cases_and_policies/rmd/counterfactuals.Rmd
Reorganized counterfactuals code.
Added longer description of counterfactuals.

Added counterfactual with fixed residuals for masks with cases.

@@
 ```{r cffuncs, cache=FALSE}
-#' Given an estimated felm model with lags and differences, returns a
-#' function that simulates the model.
-#'
-#' @param m estimated felm model
-#' @param df Data
-#' @param dvar Name of variable repressenting differences in the
-#' model. dvar should be a difference of length difflag.
-#' @param var Name of varaible representing levels in the model.
-#' @param difflag difference length for dvar
-#' @param lhsdiff if true the outcome is also dvar with the same
-#' difflag, if not the outcome is var
-#'
-#' @return A function with arguments idvar==value of identifier for
-#'   which to simulate and shockgen=function(n) that generates values
-#'   for residuals. The default value of shockgen is to resample model
-#'   residuals with replacement. The function returns a simulated path
-#'   var beginning from the initial values for id==idvar and using the
-#'   exogenous covariates of id==idvar.
-#'
-#' Example:
-#' df <- df[order(df$id,$df$t),]
-#' df$dy <- paneldiff(df$y, df$id, df$t, lag=2)
-#' fmla <- dy ~ lag(y,3) + lag(dy,1) + x
-#' # hack to get get id
-#'
-#' sim <- createforwardsim(fmla, df, "dy", "y", 2, TRUE)
-#' # simulate with errors set to 0's
-#' yhat <- sim(df$id[1], function(n) fill(0, n))
-#' #' # simulate sampling residuals with replacement
-#' yhat <- sim(df$id[1])
-createforwardsim <- function(m, df, id="state",
-                             dvar="dlogdc", var="logdc",
-                             difflag=7, lhsdiff=TRUE) {
-  m$coef[is.nan(m$coef)] <- 0
-
-  # get index of observations used in mdl
-  df$idxcfs <- 1:nrow(df)
-  mid <- felm(update.felm.formula(m$formula, idxcfs ~ .), data=df, keepX=TRUE)
-  idx <- mid$response
-
-  # This  regexp will match dvar if preceded by start of string or (
-  # and followed by end of string or ,
-  di <- grepl(sprintf("(^|\\()%s(,|$)",dvar), rownames(m$coef))
-  li <- grepl(sprintf("(^|\\()%s(,|$)",var), rownames(m$coef))
-  dc <- m$coef[di]
-  names(dc) <- rownames(m$coef)[di]
-  lc <- m$coef[li]
-  names(lc) <- rownames(m$coef)[li]
-  # exogenous part of model
-  if (!is.null(getfe(m))) {
-    stop("createforwardsim not implemented with fixed effects. Include factors in the formula instead")
-  }
-
-  xt <- mid$X %*% (m$coef * as.numeric(!(di | li)))
-  #xtold <- m$fitted.values - m$X %*% (m$coef * as.numeric(di | li))
-
-  ## convert into AR model
-  llags <- as.numeric(gsub("lag\\(.+, (\\d+)\\)","\\1", names(lc)))
-  dlags <- as.numeric(gsub("lag\\(.+, (\\d+)\\)","\\1", names(dc)))
-  maxlags <- max(c(llags, dlags+difflag, ifelse(lhsdiff, difflag, 0)))
-  ar <- rep(0, maxlags)
-  if (lhsdiff) ar[difflag]  <- ar[difflag] + 1
-  ar[llags]  <- ar[llags] + lc
-  ar[dlags] <- ar[dlags] +dc
-  ar[dlags+difflag] <- ar[dlags+difflag]-dc
-
-  if (!is(df, "pdata.frame")) {
-    warning("df is not a pdata.frame. This function assumes df is sorted by id, time, and within each id, time is consecutive. Results will be incorrect if these assumptions are false.")
-  } else {
-    stopifnot(all(is.pconsecutive(df)))
-  }
-
-  sim <- function(idvar, shockgen=function(n=1) { sample(m$residuals, size=n,
-                                                         replace=TRUE)},
-                  samplecoefs=FALSE, nsim=100) {
-    ids <- df[idx,id]
-    if (samplecoefs) {
-      V <- vcov(m)
-      nac <- is.na(m$coefficients)
-      e <- rep(0,nrow(V))
-      Vnona <- V[!nac,!nac]
-      e[!nac] <- t(rmvnorm(1, mean=rep(0, nrow(Vnona)), sigma=Vnona))
-
-      dc <- m$coef[di] + e[di]
-      lc <- m$coef[li] + e[li]
-      ## convert into AR model
-      ar <- rep(0, maxlags)
-      if (lhsdiff) ar[difflag]  <- ar[difflag] + 1
-      ar[llags]  <- ar[llags] + lc
-      ar[dlags] <- ar[dlags] +dc
-      ar[dlags+difflag] <- ar[dlags+difflag]-dc
-
-      x <- xt[ids==idvar] + (mid$X[ids==idvar,] %*% (e * as.numeric(!(di | li))))
-    } else {
-      x <- xt[ids==idvar]
-    }
-
-    date <- df$date[idx[df[idx,id]==idvar]]
-    names(date) <- NULL
-    it0 <- which(df$date==date[1] & df[,id]==idvar)
-    Y0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,var], k)[it0]))
-    dY0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,dvar], k)[it0]))
-    lagy0 <- Y0-dY0
-    check <- lagy0[1:(length(Y0)-difflag)]- Y0[(difflag+1):length(Y0)]
-    stopifnot(all(abs(check)<1e-8, na.rm=TRUE))
-    na0 <- is.na(Y0[(difflag+1):length(Y0)])
-    Y0[(difflag+1):length(Y0)][na0] <- lagy0[1:(length(Y0)-difflag)][na0]
-
-    yt <- matrix(0,nrow=nsim,ncol=length(x))
-    e <- matrix(shockgen(length(x)*nsim),nrow=nsim)
-    Y0 <- t(matrix(rep(Y0,nsim),ncol=nsim))
-    Y <- Y0
-    for (t in 1:length(x)) {
-      yt[,t] <- x[t] + Y %*% ar + e[,t]
-      #Y <- c(yt[t], Y[-length(Y)])
-      Y[,2:ncol(Y)] <- Y[,1:(ncol(Y)-1)]
-      Y[,1] <- yt[,t]
-    }
-    return(list(y=cbind(Y0[, ncol(Y0):1],yt),
-                date=c(date[1]-((maxlags):1),date)))
-  }
-}
-
-meanandci <- function(yi, cfyi, date, p=0.9, difflag=7) {
-
-  y <- colMeans(yi)
-  cfy <- colMeans(cfyi)
-  dy <- matrix(NA,nrow=nrow(y),ncol=ncol(y))
-  dy[(difflag+1):nrow(y),]  <- diff(y,difflag)
-
-  dcfy <- matrix(NA,nrow=nrow(cfy),ncol=ncol(cfy))
-  dcfy[(difflag+1):nrow(cfy),]  <- diff(cfy,difflag)
-  statframe <- function(y,suffix) {
-    d <- data.frame(m=rowMeans(y, na.rm=TRUE),
-                    cl=apply(y, 1, function(x) quantile(x, (1-p)/2,na.rm=TRUE)),
-                    ch=apply(y, 1, function(x) quantile(x,1-(1-p)/2,na.rm=TRUE)))
-    names(d) <- paste(names(d), suffix, sep="")
-    d
-  }
-
-  cbind(data.frame(date=date),
-        statframe(colMeans(exp(yi)),"y"), statframe(dy,"dy"), statframe(colMeans(exp(cfyi)),"cf"),
-        statframe(dcfy,"dcf"), statframe(colMeans(exp(cfyi)-exp(yi)),"p"),
-        statframe(dcfy-dy,"dp"))
-
-}
-
-nopstatesim <- function(st, nsim=200, simobs, simcf, data, var="logdc", dvar="dlogdc") {
-  seed <- .Random.seed
-  set.seed(seed)
-  yi <- replicate(nsim, simobs(st, samplecoefs=TRUE, nsim=nsim/2)$y)
-  set.seed(seed)
-  cfyi <- replicate(nsim, simcf(st, samplecoefs=TRUE, nsim=nsim/2)$y)
-  date <- simobs(st, samplecoefs=TRUE, nsim=2)$date
-  est <- meanandci(yi,cfyi, date)
-  est <- merge(est, data.frame(state=st,
-                        obs=exp(data[data$state==st, var]),
-                        dobs=data[data$state==st, dvar],
-                        date=data[data$state==st,]$date),
-               by="date", all=TRUE)
-  return(list(edf=est, y=yi, cfy=cfyi))
-}
-
-cfplots <- function(st, cfdata=NULL, simobs=NULL, simcf=NULL,
-                    yvar="cases", data) {
-  xlims <- datelims
-  pds <-
-    c("Mandate.face.mask.use.by.employees.in.public.facing.businesses",
-      "Date.closed.K.12.schools",
-      "Stay.at.home..shelter.in.place",
-      "Closed.movie.theaters",
-      "Closed.restaurants.except.take.out",
-      "Closed.non.essential.businesses")
-  if (is.null(cfdata)) {
-    est <- nopstatesim(st, simobs=simobs, simcf=simcf, data=data)
-  } else {
-    est <- subset(cfdata, cfdata$state==st)
-  }
-
-  dt  <- c()
-  labels <- c()
-  for (lbl in pds) {
-    d <- unique(data[data$state==st,lbl])
-    i <- which(dt==d)
-    if (length(i)==0) {
-      if (length(dt)==0)
-        dt <- d
-      else
-        dt <- c(dt,d)
-      labels <- c(labels,lbl)
-    } else {
-      labels[i]  <- paste(labels[i],lbl,sep=", ")
-    }
-  }
-  labels <- labels[!is.na(dt)]
-  dt <- dt[!is.na(dt)]
-
-  figl <-
-    ggplot(data=est, aes(x=date, y=obs)) +
-    geom_line(aes(color="Observed", fill="Observed"), size=1.5) +
-    geom_line(aes(y=my,color="Estimated"), size=1.2) +
-    #geom_ribbon(aes(ymin=cly, ymax=chy, fill="Estimated"),
-    #            alpha=0.3, show.legend=FALSE) +
-    geom_line(aes(y=mcf, color="Counterfactual"), size=1.2) +
-    #geom_ribbon(aes(ymin=clcf, ymax=chcf, fill="Without Policies"),
-    #            alpha=0.3, show.legend=FALSE) +
-    figtheme + colors() + colors_fill() +
-    ylab(sprintf("%s in past week",yvar)) + labs(color="")  +
-    theme(legend.position=c(0.2, 0.8)) +
-    ggtitle(TeX(sprintf("%s in past week",yvar))) + xlim(xlims)
-
-  figp <- ggplot(data=est, aes(x=date, y=mp)) +
-    geom_line(size=1.2) +
-    geom_ribbon(aes(ymin=clp, ymax=chp), alpha=0.3) +
-    ylab(sprintf("Change in %s in past week", yvar)) + figtheme + colors() + colors_fill() +
-    ggtitle(TeX(sprintf("Change in %s",yvar))) +
-    xlim(xlims) +
-    annotate("text",x=dt, y=rep(0.01,length(dt)),
-             label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)
-
-  figd <- ggplot(data=est, aes(x=date, y=dobs)) +
-    geom_line(aes(y=mdcf, color="Counterfactual"), size=1.5) +
-    geom_line(aes(color="Observed", fill="Observed"), size=1.5) +
-    geom_line(aes(y=mdy,color="Estimated"), size=1.5) +
-    #geom_ribbon(aes(ymin=cldy, ymax=chdy, fill="Estimated"),
-    #            alpha=0.3, show.legend=FALSE) +
-    #geom_ribbon(aes(ymin=cldcf, ymax=chdcf, fill="Without Policies"),
-    #            alpha=0.3, show.legend=FALSE) +
-    figtheme + colors() + colors_fill() +
-    ylab(TeX(sprintf("$\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
-    labs(color="")  +
-    theme(legend.position=c(0.8, 0.8)) +
-    ggtitle(TeX(sprintf("$\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
-    xlim(xlims)
-
-  figdp <- ggplot(data=est, aes(x=date, y=mdp)) +
-    geom_line(size=1.2) +
-    geom_ribbon(aes(ymin=cldp, ymax=chdp), alpha=0.3) +
-    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
-    figtheme + colors() + colors_fill() +
-    ggtitle(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",toupper(substring(yvar,1,1))))) +
-    xlim(xlims) +
-    annotate("text",x=dt, y=rep(0.01,length(dt)),
-             label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)
-
-  return(list(figd=figd, figdp=figdp, figl=figl, figp=figp))
-}
-
-# Extract E[log(dc) |state, t, parameters] with parameters ~ N(m$coef, vcov(m))
-getmeans <- function(sl) {
-  date <- sl$edf$date[!is.na(sl$edf$my)]
-  state <- sl$edf$state[!is.na(sl$edf$state)]
-  stopifnot(length(unique(state))==1)
-  y0 <- apply(exp(sl$y), c(2,3), mean)
-  long <- melt(y0)
-  names(long)[names(long)=="value"] <- "meany0"
-
-  log0 <- apply(sl$y, c(2,3), mean)
-  long1 <- melt(log0)
-  names(long1)[names(long1)=="value"] <- "meanlog0"
-  long <- merge(long, long1, by=c("Var1","Var2"))
-
-  y1 <- apply(exp(sl$cf), c(2,3), mean)
-  long1 <- melt(y1)
-  names(long1)[names(long1)=="value"] <- "meany1"
-  long <- merge(long,long1, by=c("Var1","Var2"))
-
-  log1 <- apply(sl$cf, c(2,3), mean)
-  long1 <- melt(log1)
-  names(long1)[names(long1)=="value"] <- "meanlog1"
-  long <- merge(long, long1, by=c("Var1","Var2"))
-
-  long$date <- date[long$Var1]
-  long$state <- unique(state)
-  long$sidx <- long$Var2
-  long$Var1 <- NULL
-  long$Var2 <- NULL
-  return(long)
-}
-
+source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
 ```
@@
 estimated average, and counterfactual without policies average of
-$\Delta \log \Delta C$ in Washington. To compute the estimated and
-counterfactual paths we use the estimates from column 2 of table
-\ref{tab:policy-cases}. We set initial $\Delta \log \Delta C$ and
-$\log \Delta C$ to their values first observed in the state we are
-simulating. We hold all other regressors at their observed
-values. Error terms are drawn with replacement from the residuals. We
-do this many times and report the average over draws of the
-residuals. To calculate confidence intervals, we repeat the above with
-coefficients drawn randomly from their asymptotic distribution. The
-shaded region is a pointwise 90\% confidence interval.
+$\Delta \log \Delta C$ in Washington.
+
+To compute the estimated and counterfactual paths we use the estimates
+from column 2 of table \ref{tab:policy-cases}. We set initial $\Delta
+\log \Delta C$ and $\log \Delta C$ to their values first observed in
+the state we are simulating. We hold all other regressors at their
+observed values. Error terms are drawn with replacement from the
+residuals. We do this many times and report the average over draws of
+the residuals. To calculate confidence intervals, we repeat the above
+with coefficients drawn randomly from their asymptotic
+distribution. The shaded region is a pointwise 90\% confidence
+interval.
+
+To be more precise, we can rewrite our regression model as:
+
+$$
+  Y_{it} = P_{it}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \epsilon_{it}
+$$
+
+where $Y_{it} = \log \Delta C_{it}$, $P_{it}$ are policies, and
+$X_{it}$ are controls. Each counterfactual involves changing $P_{it}$
+to some other value, $P_{it}^\ast$. We then generate counterfactual paths
+for cases by in state $i$ conditional on
+$Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast \}_{t=1}^T, \theta=(\alpha, \beta, \rho)$ by setting:
+
+$$
+  Y_{it}^{\ast,r} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}^{\ast,r}
+$$
+
+where $\hat{\epsilon}_{it}^{\ast,r}$ are resampled with replacement
+from $\{\hat{\epsilon}_{it}\}_{i=1,t=1}^{N,T}$. We do resample the
+residuals $R$ times, repeating the above each time. We then compute
+
+$$
+\hat{E}[Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
+\frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}
+$$
+
+Finally, we repeatedly draw $\theta$ from its asymptotic distribution
+and repeat all of the above to compute confidence intervals.
+
+For case growth, the same procedure is used, but with $\Delta Y_{it} = \Delta \log \Delta C_{it}$ in place of $Y_{it}$. The top panel of the following figure shows
+
+$$
+E[\Delta Y_{it} | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}\}_{t=1}^T]
+$$
+
+and
+
+$$
+E[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T]
+$$
+
+along with the observed $Y_{it}$, and the bottom panel shows the difference.
 
@@
 ```
+
+To look at cases we instead compute,
+
+$$
+\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
+\frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r})
+$$
 
@@
 
+That is for the first figure, we take:
+
+$$
+\hat{E}[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
+\frac{1}{N} \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \Delta Y_{it}^{\ast,r}
+$$
+
+and plot the difference of the above between two different values of $P_{it}^\ast$
+
+The second figure shows the total change in cases, that is the change in:
+
+$$
+\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
+\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}).
+$$
+
+The final figure shows
+$$
+\frac{\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}) - \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^r)} {\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{r})}
+$$
+that is, the (average over residual draws) change in national cases relative to (average over residual draws) national cases.
+
 ```{r nop-all, cache=TRUE}
@@
 cfnop <- merge(cfnop, abbrv, by="state")
-
-nationalplots <- function(alldf,cfdf, cfname, yvar="cases") {
-  alldf[order(state, sidx, date), dlog1:=c(rep(NA,7),diff(meanlog1,7)), by=.(state, sidx)]
-  alldf[order(state, sidx, date),dlog0:=c(rep(NA,7),diff(meanlog0,7)), by=.(state, sidx)]
-  national <- alldf[ , .(diff=sum(meany1-meany0), y0=sum(meany0), dlog=mean(dlog1-dlog0)), by=.(date, sidx)]
-
-  adf <- national[, .(mp=mean(diff),
-                      clo=quantile(diff, 0.05),
-                      chi=quantile(diff, 0.95),
-                      rel=mean(diff/y0),
-                      rlo=quantile(diff/y0, 0.05),
-                      rhi=quantile(diff/y0, 0.95),
-                      mlog=mean(dlog, na.rm=TRUE),
-                      llo=quantile(dlog, 0.025, na.rm=TRUE),
-                      lhi=quantile(dlog, 0.975, na.rm=TRUE)
-                      ),
-                  keyby=.(date)]
-
-  figdlog <- ggplot(cfdf, aes(x=date)) +
-    #geom_text(aes(y=mdp, label=ST),alpha=0.2) +
-    geom_point(aes(y=mdp),alpha=0.2, size=0.5) +
-    xlim(datelims) +
-    geom_line(data=adf,  aes(x=date, y=mlog, color=""), size=2) + figtheme +
-    geom_ribbon(data=adf, aes(ymin=llo, ymax=lhi, fill=""),alpha=0.3) +
-    colors() + theme(legend.position="none") +
-    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",
-                     toupper(substring(yvar,1,1))
-                     ))) +
-    colors_fill() +
-     ggtitle(sprintf("Effect of %s on %s growth",cfname, substring(yvar,1,nchar(yvar)-1)))
-
-  figc <- ggplot(cfdf, aes(x=date)) +
-    #geom_text(aes(y=mp, label=ST),alpha=0.5) +
-    geom_point(aes(y=mp),alpha=0.2, size=0.5) +
-    xlim(datelims) +
-    geom_line(data=adf, aes(x=date, y=mp, color=""), size=2) +
-    figtheme +
-    geom_ribbon(data=adf, aes(ymin=clo, ymax=chi, fill=""),alpha=0.3) +
-    colors() + theme(legend.position="none") +
-    ylab(sprintf("%s in the past week", yvar)) +
-    colors_fill() +
-    ggtitle(sprintf("Effect of %s on %s",cfname, yvar))
-
-  figr <- ggplot(adf, aes(x=date, y=rel)) +
-    xlim(datelims) +
-    geom_line(aes(color=""), size=2) + figtheme +
-    geom_ribbon(data=adf, aes(ymin=rlo, ymax=rhi, fill=""),alpha=0.3) +
-    colors() + theme(legend.position="none") +
-    ylab(sprintf("relative increase in %s",yvar)) +
-    colors_fill() +
-    ggtitle(sprintf("Relative effect of %s",cfname))
-  return(list(g=figdlog, c=figc, r=figr))
-}
 
@@
 for (f in figs) print(f)
+```
+
+### Holding Residuals fixed
+
+In this subsection, instead of resampling residuals, we condition on them. That is we look at
+
+$$
+  Y_{it}^{\ast} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}
+$$
+
+instead of
+
+$$
+  \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}.
+$$
+
+```{r masks-resids-fixed}
+simmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
+  simmodel(st, samplecoefs=samplecoefs, nsim=nsim,
+           shockgen=NULL)
+maskmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
+  maskmodel(st, samplecoefs=samplecoefs, nsim=nsim,
+            shockgen=NULL)
+
+rerunsim <- FALSE
+if (!file.exists("cfmask1.Rda") || rerunsim) {
+  cfmasklist1 <- lapply(unique(df$state),
+                        function(s) nopstatesim(s,simobs=simmodel1,simcf=maskmodel1, data=sdf, nresid=1))
+  maskall1 <- rbindlist(lapply(cfmasklist1, getmeans))
+  save(cfmasklist1, maskall1, file="cfmask1.Rda")
+}
+load("cfmask1.Rda")
+cfmask1 <- do.call(rbind, lapply(cfmasklist1, function(x) x$edf))
+
+for (st in c("Washington","Illinois","Massachusetts")) {
+  f <- cfplots(st, cfmask1, yvar="deaths", data=df)
+  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))
+
+}
+
+figs1 <- nationalplots(maskall1, cfmask1, "mandating masks on April 1st\n")
+
+for (f in figs1) print(f)
 ```

### 2020-07-23 cases_and_policies/rmd/counterfactuals_national.Rmd
Reorganized counterfactuals code.
Added longer description of counterfactuals.

Added counterfactual with fixed residuals for masks with cases.

@@
 cfnop <- merge(cfnop, abbrv, by="state")
-
-nationalplots <- function(alldf,cfdf, cfname, yvar="cases") {
-  alldf[order(state, sidx, date), dlog1:=c(rep(NA,7),diff(meanlog1,7)), by=.(state, sidx)]
-  alldf[order(state, sidx, date),dlog0:=c(rep(NA,7),diff(meanlog0,7)), by=.(state, sidx)]
-  national <- alldf[ , .(diff=sum(meany1-meany0), y0=sum(meany0), dlog=mean(dlog1-dlog0)), by=.(date, sidx)]
-
-  adf <- national[, .(mp=mean(diff),
-                      clo=quantile(diff, 0.05),
-                      chi=quantile(diff, 0.95),
-                      rel=mean(diff/y0),
-                      rlo=quantile(diff/y0, 0.05),
-                      rhi=quantile(diff/y0, 0.95),
-                      mlog=mean(dlog, na.rm=TRUE),
-                      llo=quantile(dlog, 0.025, na.rm=TRUE),
-                      lhi=quantile(dlog, 0.975, na.rm=TRUE)
-                      ),
-                  keyby=.(date)]
-
-  figdlog <- ggplot(cfdf, aes(x=date)) +
-    #geom_text(aes(y=mdp, label=ST),alpha=0.2) +
-    geom_point(aes(y=mdp),alpha=0.2, size=0.5) +
-    xlim(datelims) +
-    geom_line(data=adf,  aes(x=date, y=mlog, color=""), size=2) + figtheme +
-    geom_ribbon(data=adf, aes(ymin=llo, ymax=lhi, fill=""),alpha=0.3) +
-    colors() + theme(legend.position="none") +
-    ylab(TeX(sprintf("Change in $\\Delta \\log \\Delta %s$",
-                     toupper(substring(yvar,1,1))
-                     ))) +
-    colors_fill() +
-    ggtitle(sprintf("Effect of %s on %s growth",cfname, substring(yvar,1,nchar(yvar)-1)))
-
-  figc <- ggplot(cfdf, aes(x=date)) +
-    #geom_text(aes(y=mp, label=ST),alpha=0.5) +
-    geom_point(aes(y=mp),alpha=0.2, size=0.5) +
-    xlim(datelims) +
-    geom_line(data=adf, aes(x=date, y=mp, color=""), size=2) +
-    figtheme +
-    geom_ribbon(data=adf, aes(ymin=clo, ymax=chi, fill=""),alpha=0.3) +
-    colors() + theme(legend.position="none") +
-    ylab(sprintf("%s in the past week", yvar)) +
-    colors_fill() +
-    ggtitle(sprintf("Effect of %s on %s",cfname, yvar))
-
-  figr <- ggplot(adf, aes(x=date, y=rel)) +
-    xlim(datelims) +
-    geom_line(aes(color=""), size=2) + figtheme +
-    geom_ribbon(data=adf, aes(ymin=rlo, ymax=rhi, fill=""),alpha=0.3) +
-    colors() + theme(legend.position="none") +
-    ylab(sprintf("relative increase in %s",yvar)) +
-    colors_fill() +
-    ggtitle(sprintf("Relative effect of %s",cfname))
-  return(list(g=figdlog, c=figc, r=figr))
-}
 

### 2020-07-25 cases_and_policies/rmd/regressionsWithPlots.Rmd
update with log(vote)

@@
 rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
+rootdir <- "/Users/Hiro/Documents/GitHub/covid-impact" 
 source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
@@
 
-```{r corr, cache=TRUE}
+```{r corr, cache=FALSE}
 cmat <-  cor(df[,c(bvars,pols)], use="pairwise.complete.obs")
@@
 
-```{r caseregs, cache=TRUE, warning=FALSE, results='hide'}
+```{r caseregs, cache=FALSE, warning=FALSE, results='hide'}
 infovars <- list(c("dlogdc", "logdc"),
@@
 
-```{r casetables, cache=TRUE, warning=FALSE, results='asis'}
-showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy)
+```{r casetables, cache=FALSE, warning=FALSE, results='asis'}
+showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy, regs$ip)
 savetextables(regs$pib[[1]], NULL, NULL, prefix="behavior")
@@
 
-```{r casedieff, cache=TRUE, warning=FALSE, results='asis'}
+```{r casedieff, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 for (use.national in c(TRUE,FALSE)) {
@@
 
-```{r deathtables, cache=TRUE, warning=FALSE, results='asis'}
+```{r deathtables, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
 infovars <- list(c("dlogdc", "logdc"),
@@
 
-```{r deathtablesd, cache=TRUE, warning=FALSE, results='asis'}
+```{r deathtablesd, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 df$zero <- 0
@@
 
-```{r dieff-deaths, cache=TRUE, results='asis', warning=FALSE}
+```{r dieff-deaths, eval = TRUE, cache=FALSE, results='asis', warning=FALSE}
 for (use.national in c(TRUE,FALSE)) {
@@
 
-<!-- ```{r logdddc, cache=TRUE, warning=FALSE, results='asis'} -->
+<!-- ```{r logdddc, cache=FALSE, warning=FALSE, results='asis'} -->
 <!-- df$logdddc <- log(df$ddeath/panellag(df$cases, df$state, df$date, 7)) -->
@@
 <!-- ``` -->
+
+## Robustness Check
+```{r regsrobust, cache=FALSE, warning=FALSE, results='asis'}
+infovars <- list(c("dlogdc", "logdc"),
+                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
+#c("dlogdc", "logdc"))
+#infovars2 <- list(c( "logdc"))
+tvars <- "dlogtests"  
+# xlist <-list(c("z.mask","vote"), c("z.mask","vote")) 
+xlist <-list(c("z.mask"), c("z.mask")) 
+#xlist <-list(c("mask_percent","vote"), c("mask_percent","vote")) 
+interactions <-list("month", statevars)
+ilist <- list(interactions, interactions)
+iv <- list("0","0") #,"(testratedc ~ lag(testratedc,14))") 
+L <- 14
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+regs <- mainregressions(sdf,
+                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
+
+showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
+savetextables(NULL, regs$pbiy, regs$piy, prefix="cases-robust")
+
+df$zero <- 0
+infovarsd <- list(c("dlogdd", "logdd"),
+                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
+# infovarsd <- list(c("logdd")
+L <- 21
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+regs <- mainregressions(sdf,
+                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)
+showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
+
+```
 
@@
 
-```{r fig, cache=TRUE,  fig_width=10, fig_height=6, warning=FALSE}
+```{r fig, cache=FALSE,  fig_width=10, fig_height=6, warning=FALSE}
 pfigure <- function(df, pol, stlab=1, outcome="dlogdc", L=14) {
@@
 
-```{r polplots, cache=TRUE, fig_width=10, fig_height=6}
+```{r polplots, cache=FALSE, fig_width=10, fig_height=6}
 fig <- pfigure(df, "pk12", stlab=-1) +
@@
 
-```{r poldeathsplots, cache=TRUE, fig_width=10, fig_height=6, warning=FALSE}
+```{r poldeathsplots, cache=FALSE, fig_width=10, fig_height=6, warning=FALSE}
 fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=21) +

### 2020-07-25 tex/tables_and_figures/policies.tex
update with log(vote)

% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Sat Jul 25 16:38:15 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & N & Min & Median & Max \\ 
  \hline
State of emergency & 50 & 2020-02-29 & 2020-03-11 & 2020-03-16 \\ 
  Date closed K 12 schools & 50 & 2020-03-13 & 2020-03-17 & 2020-04-03 \\ 
  Closed day cares & 14 & 2020-03-16 & 2020-03-23 & 2020-04-06 \\ 
  Date banned visitors to nursing homes & 29 & 2020-03-09 & 2020-03-16 & 2020-04-06 \\ 
  Closed non essential businesses & 42 & 2020-03-19 & 2020-03-25 & 2020-04-06 \\ 
  Closed restaurants except take out & 47 & 2020-03-15 & 2020-03-17 & 2020-04-03 \\ 
  Closed gyms & 45 & 2020-03-16 & 2020-03-20 & 2020-04-03 \\ 
  Closed movie theaters & 48 & 2020-03-16 & 2020-03-20 & 2020-04-06 \\ 
  Stay at home  shelter in place & 41 & 2020-03-19 & 2020-03-28 & 2020-04-07 \\ 
  End relax stay at home shelter in place & 37 & 2020-04-24 & 2020-05-16 & 2020-06-19 \\ 
  Began to reopen businesses statewide & 49 & 2020-04-20 & 2020-05-07 & 2020-06-05 \\ 
  Reopen restaurants & 47 & 2020-04-24 & 2020-05-18 & 2020-06-22 \\ 
  Reopened gyms & 43 & 2020-04-24 & 2020-05-18 & 2020-07-03 \\ 
  Reopened movie theaters & 36 & 2020-04-27 & 2020-05-31 & 2020-07-03 \\ 
  Resumed elective medical procedures & 34 & 2020-04-20 & 2020-04-29 & 2020-05-29 \\ 
  Mandate face mask use by all individuals in public spaces & 31 & 2020-04-08 & 2020-06-18 & 2020-07-27 \\ 
  Mandate face mask use by employees in public facing businesses & 41 & 2020-04-03 & 2020-05-01 & 2020-07-03 \\ 
  Stop Initiation of Evictions overall or due to COVID related issues & 29 & 2020-03-16 & 2020-03-24 & 2020-05-30 \\ 
  Stop enforcement of evictions overall or due to COVID related issues & 28 & 2020-03-15 & 2020-03-24 & 2020-06-08 \\ 
  Renter grace period or use of security deposit to pay rent & 3 & 2020-04-10 & 2020-04-24 & 2020-05-20 \\ 
  Order freezing utility shut offs & 33 & 2020-03-12 & 2020-03-19 & 2020-04-13 \\ 
  Froze mortgage payments & 1 & 2020-03-21 & 2020-03-21 & 2020-03-21 \\ 
  Waived one week waiting period for unemployment insurance & 35 & 2020-03-08 & 2020-03-18 & 2020-04-06 \\ 
   \hline
\end{tabular}
\caption{State Policies \label{tab:policies}} 
\end{table}

### 2020-07-25 cases_and_policies/R/cases_and_policies.R

@@ -36,6 +36,8 @@ gitrootdir <- function() {
 
 rootdir <- gitrootdir()
 
+rootdir <- "/Users/Hiro/Documents/GitHub/covid-impact"
+
 # Data on policies from Raifman and collaborators
 urls <- c("https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=993060716",
           "https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=1357478819",
@@ -266,6 +268,18 @@ stopifnot(length(unique(df$ST))==51)
 stopifnot(length(unique(df$fips))==51)
 stopifnot(nrow(unique(df[,c("state","ST","fips")]))==51)
 stopifnot(nrow(unique(df[,c("fips","date","state","ST")]))==nrow(df))
+
+temp <- read.csv(paste(rootdir,"cases_and_policies/data/Facemask_Mar26_April29.csv",sep="/"))
+temp1 <- temp[c("state","fips","z.mask","mask_percent")] 
+temp <- read.csv(paste(rootdir,"cases_and_policies/data/1976-2016-president.csv",sep="/"))
+temp <- temp[ which(temp$year==2016 & temp$candidate=="Trump, Donald J." & temp$party=="republican" & 
+                       temp$writein=="FALSE"), ]
+temp$voteshare <- temp$candidatevotes/temp$totalvotes
+temp$fips <- temp$state_fips
+temp <- temp[c("state","fips","voteshare")]
+temp <- merge(temp,temp1,by=c("state","fips")) 
+df <- merge(df,temp,by=c("state","fips"))
+
 
 
 covidstates <- df

### 2020-07-25 cases_and_policies/rmd/generatetables.R

@@ -141,8 +141,18 @@ mainregressions <- function(df, # data
                   xlist[[k]]),
                 interactions[[k]], "0", L=0)
     })
-  }
-  return(list(pib=pib, pbiy=pbiy, piy=piy))
+  } 
+  
+   
+  ijs <- expand.grid(1:length(pols), 1:length(xlist))
+  ip <- apply(ijs, 1, function(ij) {
+    policyreg(df, pols[[ij[1]]],  NULL,  bvars,
+              c(infovars[[ij[2]]],
+                xlist[[ij[2]]]),
+              interactions[[ij[2]]], "0", L=L)
+  })
+    
+  return(list(pib=pib, pbiy=pbiy, piy=piy, ip=ip))
 }
 
 
@@ -155,7 +165,7 @@ printstars <- function(est, se, starp=c(0.1, 0.05, 0.01)) {
   return(sprintf("%.3f$^{%s}$", est, stars))
 }
 
-showhtmltables <- function(pib, pbiy, piy) {
+showhtmltables <- function(pib, pbiy, piy, ip=NULL) {
 
   omit <- c(statevarregexp,
             "^month") #"testratedc:",)
@@ -183,7 +193,31 @@ showhtmltables <- function(pib, pbiy, piy) {
                                "\\hline")
               )
 
- }
+  }
+  
+  
+  if (!(is.null(ip))) {
+    cat("\n### Policies and Information\n")
+    
+    peff <- sapply(ip, function(x) x$peff[1])
+    sep <- sapply(ip, function(x) x$peff[2])
+    cnames <- sapply(ip, function(x) colnames(x$reg$response))
+    stargazer(lapply(ip, function(x) x$reg),
+              type="html",
+              title="Policies and Information",
+              #dep.var.labels=cnames,
+              dep.var.labels.include=FALSE,
+              column.labels=cnames,
+              omit=omit,
+              omit.labels=omit.labels,
+              omit.stat=c("f", "ser"), model.names=FALSE,
+              model.numbers=TRUE,
+              add.lines = list(c("sum Policy",sprintf("%.3f",peff)),
+                               c("",sprintf("(%.3f)",sep)),
+                               "\\hline")
+    )
+    
+  }
 
   if (!(is.null(pbiy))) {
     ylbl <- colnames(pbiy[[1]]$reg$response)
@@ -235,6 +269,9 @@ showhtmltables <- function(pib, pbiy, piy) {
               model.numbers=TRUE)
 
   }
+  
+  
+ 
 
   return(NULL)
 }

### 2020-07-25 cases_and_policies/rmd/regprep.R

@@ -78,12 +78,18 @@ df$pmaskany <- apply(df[,c("pmaskbus","pmaskall")], 1, max)
 pols <- c("pmaskbus", "pk12","pshelter","pmovie","pgyms","prestaurant","pnonessential")
 
 
+df$vote <- df$voteshare
+df$logvote <- log(df$vote)
+
 statevars <- c("log(Population.2018)",
                "log(Square.Miles)",
                "Percent.Unemployed..2018.",
                "Percent.living.under.the.federal.poverty.line..2018.",
-               "Percent.at.risk.for.serious.illness.due.to.COVID")
+               "Percent.at.risk.for.serious.illness.due.to.COVID", "logvote")
 bvars <- c("workplaces","retail","grocery","transit")
+
+for(i in 1:length(bvars)) df[,bvars[i]] <- df[,bvars[i]]/100 
+
 
 #df$udistance <- panelma(df$daily_distance_diff,df$state, df$date, 6)
 #df$uvisits <- panelma(df$daily_visitation_diff,df$state, df$date, 6)
@@ -114,6 +120,9 @@ ndf <- merge(ndf, nc, by="date", all=TRUE)
 ndf <- pdata.frame(ndf, index=c("state","t"))
 df <- ndf[order(ndf$state, ndf$date),]
 
+df$z.mask <- df$z.mask/100
+df$mask_percent <- df$mask_percent/100
+
 tilt = ifelse(df$date> as.Date("2020-05-01"),1,0 )
 month <- month(df$date)
 #month[month==6] <- 5
@@ -121,4 +130,4 @@ month <- month(df$date)
 df$month = as.factor(month)
 df$pmask.may <- df$pmaskbus*tilt
 df$pmask.april <- df$pmaskbus*(1-tilt)
-pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential")
+pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 

### 2020-07-25 cases_and_policies/rmd/varlabels.R

@@ -21,6 +21,7 @@ varlabels <-
     "log(deaths[t]-deaths[t-7])", "NYT", TRUE,
 
     "logt", "log(days since 2020-01-15)", "log(days since January 15, 2020)", "", TRUE,
+    
 
     "testratedc", "$\\frac{\\Delta{D}}{\\Delta{C}} T_{it}$",
     "tests in past week time deaths in past week divided by cases in past week",
@@ -58,6 +59,15 @@ varlabels <-
 
     "psoe","state of emergency","state of emergency", "Raifman et al",TRUE,
 
+    
+    "party","Governor's party","Governor's party", "OPENICPSR",TRUE,
+    
+    "vote","Trump voting shares","Trump voting shares", "MIT Election Lab",TRUE,
+    
+    "logvote","log(Trump voting shares)","log(Trump voting shares)", "MIT Election Lab",TRUE,
+    
+    "z.mask","mask wearing rates","z-score for mask wearing rates", "YouGov",TRUE,
+    
     "pmaskbus", "masks for employees",
     "mandate face mask use by employees in public facing businesses",
     "Raifman et al",TRUE,

### 2020-07-28 cases_and_policies/rmd/counterfactuals.Rmd
update

--- 2020-07-16 cases_and_policies/rmd/counterfactuals.Rmd
@@
 
-```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
-if (!knitr::is_latex_output()) {
-  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-    f <- cfplots(st, cfnop, data=df)
-    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-  }
-}
-```
-
-```{r nop-state-figures, cache=TRUE, warning=FALSE}
-states <- c("Washington","Illinois","Massachusetts")
-for(st in states) {
-  f <- cfplots(st, cfnop, data=df)
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figdp)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figd + ylim(c(-1,5)))
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figl)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figp)
-  dev.off()
-}
-```
+# ```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
+# if (!knitr::is_latex_output()) {
+#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
+#     f <- cfplots(st, cfnop, data=df)
+#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
+#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
+#   }
+# }
+# ```
+
+# ```{r nop-state-figures, cache=TRUE, warning=FALSE}
+# states <- c("Washington","Illinois","Massachusetts")
+# for(st in states) {
+#   f <- cfplots(st, cfnop, data=df)
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figdp)
+#   dev.off()
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figd + ylim(c(-1,5)))
+#   dev.off()
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figl)
+#   dev.off()
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figp)
+#   dev.off()
+# }
+# ```
 
@@
 
-```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
-if (!knitr::is_latex_output()) {
-  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-    f <- cfplots(st, cfnop, yvar="deaths", data=df)
-    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-  }
-}
-```
-
-```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
-states <- c("Washington","Illinois","Massachusetts")
-for(st in states) {
-  f <- cfplots(st, cfnop, yvar="deaths", data=df)
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figdp)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figd + ylim(c(-1,5)))
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figl)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figp)
-  dev.off()
-}
-```
+# ```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
+# if (!knitr::is_latex_output()) {
+#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
+#     f <- cfplots(st, cfnop, yvar="deaths", data=df)
+#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
+#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
+#   }
+# }
+```
+
+# ```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
+# states <- c("Washington","Illinois","Massachusetts")
+# for(st in states) {
+#   f <- cfplots(st, cfnop, yvar="deaths", data=df)
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figdp)
+#   dev.off()
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figd + ylim(c(-1,5)))
+#   dev.off()
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figl)
+#   dev.off()
+# 
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figp)
+#   dev.off()
+# }
+# ```
 

### 2020-07-28 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

@@
 
-```{r casedieff, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+```{r casedieff, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
 for (use.national in c(TRUE,FALSE)) {
@@
 
-## Deaths (with cases as information)
+<!-- ## Deaths (with cases as information) -->
 
-```{r deathtables, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
-infovars <- list(c("dlogdc", "logdc"),
-                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
-df$zero <- 0
-L <- 21
-sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-regs <- mainregressions(sdf,
-                        "dlogdd", pols, bvars, infovars, "zero", xlist, ilist, iv, L=L)
-showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
-savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-casesinfo")
-```
+<!-- ```{r deathtables, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'} -->
+<!-- infovars <- list(c("dlogdc", "logdc"), -->
+<!--                  c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #, -->
+<!-- df$zero <- 0 -->
+<!-- L <- 21 -->
+<!-- sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) -->
+<!-- regs <- mainregressions(sdf, -->
+<!--                         "dlogdd", pols, bvars, infovars, "zero", xlist, ilist, iv, L=L) -->
+<!-- showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy); -->
+<!-- savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-casesinfo") -->
+<!-- ``` -->
 
@@
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+#sdf <- subset(df, df$date>=as.Date("2020-02-29")) #+L)
 regs <- mainregressions(sdf,
                         "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L) 
-
 showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
@@
 savetextables(NULL, regs$pbiy[c(1,3)], regs$piy[c(1,3)], prefix="deaths-deathsinfo-nosplit") 
-
 ```
 
-```{r dieff-deaths, eval = TRUE, cache=FALSE, results='asis', warning=FALSE}
+```{r dieff-deaths, eval = FALSE, cache=FALSE, results='asis', warning=FALSE}
 for (use.national in c(TRUE,FALSE)) {
@@
 tvars <- "dlogtests"
-# xlist <-list(c("z.mask","vote"), c("z.mask","vote")) 
-xlist <-list(c("z.mask"), c("z.mask")) 
-#xlist <-list(c("mask_percent","vote"), c("mask_percent","vote")) 
 interactions <-list("month", statevars)
@@
 iv <- list("0","0") #,"(testratedc ~ lag(testratedc,14))")
+
 L <- 14
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-regs <- mainregressions(sdf,
-                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
+piy <- list()
+yvar <- "dlogdc"
+for (k in 1:length(infovars)) {
+  # (1) excluding NY 
+   xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
+  sdf2 <- sdf[which(sdf$state!='New York'),]
+  piy[[(1+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (2) adding mask wearing rate and log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+  piy[[(2+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (3) lagged behavior variables
+  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
+  xlist <-list("","")
+  piy[[(3+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (4) (1)-(3)
+   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
+   sdf2 <- sdf[which(sdf$state!='New York'),]
+   xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+   piy[[(4+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
 
-showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
-savetextables(NULL, regs$pbiy, regs$piy, prefix="cases-robust")
+}
 
-df$zero <- 0
-infovarsd <- list(c("dlogdd", "logdd"),
-                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
-# infovarsd <- list(c("logdd")
+showhtmltables(NULL, NULL, piy);
+# savetextables(NULL, NULL,  piy, prefix="cases-robust")
+
 L <- 21
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-regs <- mainregressions(sdf,
-                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)
-showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
+piy <- list()
+yvar <- "dlogdd"
+infovars <- list(c("dlogdd", "logdd"),
+                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
+#xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+for (k in 1:length(infovars)) {
+  # (1) excluding NY 
+  xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
+  sdf2 <- sdf[which(sdf$state!='New York'),]
+  piy[[(1+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (2) adding mask wearing rate and log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+  piy[[(2+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (3) lagged behavior variables
+  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
+  xlist <-list("","")
+  piy[[(3+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovars[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (4) (1)-(3)
+   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
+   # sdf2 <- sdf[which(sdf$state!='New York' & sdf$state!='New Jersey' & sdf$state!='Connecticut' & sdf$state!='Massachusetts'),]
+   sdf2 <- sdf[which(sdf$state!='New York'),]
+   xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+   piy[[(4+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovars[[k]], L),
+                xlist[[k]]),
+              ilist[[k]], 0, L=L)
 
+}
+showhtmltables(NULL, NULL, piy);
+# savetextables(NULL, NULL,  piy, prefix="deaths-robust")
+
+```
+
+## Fixed Effects
+```{r regsfe, cache=FALSE, warning=FALSE, results='asis'}
+
+source(paste(rootdir,"cases_and_policies/rmd/debiasedfe.R", sep="/"))
+
+#df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
+df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
+
+yvar <- "dlogdc"
+infovars <- list(c("dlogdc", "logdc"),
+                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
+tvars <- "dlogtests"  
+L <- 14
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+piy <- list()
+for (k in 1:length(infovars)) { 
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential") 
+  m <-  reg_fe(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars), 0, L, "state + month")
+  piy[[(1+4*(k-1))]] <- m
+  m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))] 
+  #m$reg$beta[1:(6+2*(k-1))] <- m$bc[1:(6+2*(k-1))]  
+  piy[[(2+4*(k-1))]] <- m
+  pols <- c("pmaskbus","pk12","pindex") 
+  # pols <- c("pmaskbus","pk12","pmovie","pindex") 
+  m <-  reg_fe(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars), 0, L, "state + month")
+  piy[[(3+4*(k-1))]] <- m 
+  m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
+  #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
+  piy[[(4+4*(k-1))]] <- m 
+}
+
+ showhtmltables(NULL, NULL,  piy);
+# savetextables(NULL, NULL, piy, prefix="cases-fe")
+
+yvar <- "dlogdd"
+infovars <- list(c("dlogdd", "logdd"),
+                  c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
+L <- 21
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+piy <- list() 
+for (k in 1:length(infovars)) { 
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential") 
+  m <-  reg_fe(sdf, yvar, pols, NULL,
+              sprintf("lag(%s, %d)", infovars[[k]], L), 0, L, "state + month")
+  piy[[(1+4*(k-1))]] <- m
+  m$reg$beta[1:(8+2*(k-1))] <- m$bc2[1:(8+2*(k-1))] 
+  #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
+  piy[[(2+4*(k-1))]] <- m
+  pols <- c("pmaskbus","pk12","pindex") 
+  #pols <- c("pmaskbus","pk12","pmovie","pindex") 
+  m <-  reg_fe(sdf, yvar, pols, NULL,
+              sprintf("lag(%s, %d)", infovars[[k]], L), 0, L, "state + month")
+  piy[[(3+4*(k-1))]] <- m 
+  m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))] 
+  #m$reg$beta[1:(4+2*(k-1))] <- m$bc[1:(4+2*(k-1))] 
+  piy[[(4+4*(k-1))]] <- m
+}
+
+ showhtmltables(NULL, NULL,  piy);
+# savetextables(NULL, NULL, piy, prefix="deaths-fe")
 ```

### 2020-07-28 cases_and_policies/rmd/debiasedfe.R

@@ -1,0 +1,320 @@
+# 7/27/2020 Hiro Kasahara
+# Modification of the code written by Shuowen Chen ``debiasedfe_shuowen.R''
+# 7/3/2020 Shuowen Chen
+# The script contains code that implement debiased fixed effects and panel bootstrap 
+# standard errors for dataset in Chernozhukov, Kasahara and Schrimpf (2020)
+# Serves as application for Chen, Chernozhukov and Fernandez-Val (2020)
+# Acknowledgements: functions extend those written by Paul Schrimpf. 
+
+library(lfe)
+library(speedglm)
+library(boot)
+# load data (sdf)
+# load("~/Desktop/Research/Panel-Cross-over-jackknife/application/covidsince0307.RData")
+
+# add weight to sdf
+sdf$sweight <- 1
+# create week factor to the data
+sdf$week <- as.factor(strftime(sdf$date, format = "%V"))
+
+#### Functions ####
+# Create formula 
+# The function extends the createfmla to accommodate FE specifications
+# yvar: dependent variable for bpiy and piy
+# xvars: regressors. 
+#  (a) For bpiy, regressors are behavior, policy and information
+#  (b) For pib, regressors are policy and information, dependent variable is 1 
+#      of the 4 behavior variables
+#  (c) For piy, regressors are policy and information variables
+# iv: This is to accommodate the syntax of felm. "0" means no iv. 
+# cluster: cluster at state level (treatment). 
+# Note: interacted fixed effect hasn't been tested. 
+createfmla_fe <- function(yvar, xvars, iv = "0", cluster = "state",
+                          fe = c("state + month", "state * month", 
+                                 "week + state", "date + state")) {
+  rhs <- paste(xvars, collapse = " + ")
+  if (fe == "state * month") {
+    # we will use speedlm, which is faster
+    rhs <- paste(rhs, "+", fe, collapse = " ")
+  } else {
+    rhs <- paste(rhs," | ", fe, " | ", iv, " | ", paste(cluster, collapse = " + ")) 
+  }
+  return(list(as.formula(paste(yvar, "~", rhs, sep = " ")), rhs)) 
+}
+
+# a function to conduct multiple split (along the cross section) bias correction
+# to be called in the policyreg function
+# s: number of splits
+# n: number of cross section units
+# dat: data in use
+# uc: uncorrected estimators
[…]

### 2020-07-28 cases_and_policies/rmd/debiasedfe_ shuowen.R

@@ -1,0 +1,313 @@
+# 7/3/2020 Shuowen Chen
+# The script contains code that implement debiased fixed effects and panel bootstrap 
+# standard errors for dataset in Chernozhukov, Kasahara and Schrimpf (2020)
+# Serves as application for Chen, Chernozhukov and Fernandez-Val (2020)
+# Acknowledgements: functions extend those written by Paul Schrimpf. 
+
+library(lfe)
+library(speedglm)
+library(boot)
+# load data (sdf)
+load("~/Desktop/Research/Panel-Cross-over-jackknife/application/covidsince0307.RData")
+
+# add weight to sdf
+sdf$sweight <- 1
+# create week factor to the data
+sdf$week <- as.factor(strftime(sdf$date, format = "%V"))
+
+#### Functions ####
+# Create formula 
+# The function extends the createfmla to accommodate FE specifications
+# yvar: dependent variable for bpiy and piy
+# xvars: regressors. 
+#  (a) For bpiy, regressors are behavior, policy and information
+#  (b) For pib, regressors are policy and information, dependent variable is 1 
+#      of the 4 behavior variables
+#  (c) For piy, regressors are policy and information variables
+# iv: This is to accommodate the syntax of felm. "0" means no iv. 
+# cluster: cluster at state level (treatment). 
+# Note: interacted fixed effect hasn't been tested. 
+createfmla_fe <- function(yvar, xvars, iv = "0", cluster = "state",
+                          fe = c("state + month", "state * month", 
+                                 "week + state", "date + state")) {
+  rhs <- paste(xvars, collapse = " + ")
+  if (fe == "state * month") {
+    # we will use speedlm, which is faster
+    rhs <- paste(rhs, "+", fe, collapse = " ")
+  } else {
+    rhs <- paste(rhs," | ", fe, " | ", iv, " | ", paste(cluster, collapse = " + ")) 
+  }
+  return(list(as.formula(paste(yvar, "~", rhs, sep = " ")), rhs)) 
+}
+
+# a function to conduct multiple split (along the cross section) bias correction
+# to be called in the policyreg function
+# s: number of splits
+# n: number of cross section units
+# dat: data in use
+# uc: uncorrected estimators
+# fm: regression formula
+multiple_split <- function(s, n, unit, time, dat, uc, fm) {
+  across <- 0 * uc
+  for (k in 1:s) {
[…]

### 2020-07-28 cases_and_policies/rmd/generatetables.R

@@ -100,7 +100,10 @@ mainregressions <- function(df, # data
                        L=14) {
 
 
-  plist <- list(pols, c("pmask.april","pmask.may", pols[-1]))
+  # plist <- list(pols, c("pmask.april","pmask.may", pols[-1]))
+  plist <- list(pols, c("pmaskbus","pk12","pindex"))
+  
+  
 
   stopifnot(length(xlist)==length(interactions))
   stopifnot(length(xlist)==length(ivlist))
@@ -212,11 +215,15 @@ showhtmltables <- function(pib, pbiy, piy, ip=NULL) {
               omit.labels=omit.labels,
               omit.stat=c("f", "ser"), model.names=FALSE,
               model.numbers=TRUE,
-              add.lines = list(c("sum Policy",sprintf("%.3f",peff)),
-                               c("",sprintf("(%.3f)",sep)),
-                               "\\hline")
-    )
-    
+              # add.lines = list(c("sum Policy",sprintf("%.3f",peff)),
+              #                  c("",sprintf("(%.3f)",sep)),
+              #                  "\\hline")
+              add.lines=list(c("sum behavior",
+                               sprintf("%.3f",sapply(ip, function(x) x$beff[1]))),
+                             c("",
+                               sprintf("(%.3f)",sapply(ip, function(x) x$beff[2]))),
+                             "\\hline")
+    ) 
   }
 
   if (!(is.null(pbiy))) {
@@ -352,9 +359,10 @@ savetextables <- function(pib, pbiy, piy, prefix,
 
   if (!(is.null(piy))) {
 
-    yvar <- colnames(pbiy[[1]]$reg$response)
-    if (yvar=="dlogdd") ylbl <- "Death Growth"
-    if (yvar=="dlogdc") ylbl <- "Case Growth"
+   #  yvar <- colnames(pbiy[[1]]$reg$response)
+    # yvar <- colnames(piy$reg$response)
+    # if (yvar=="dlogdd") ylbl <- "Death Growth"
+    # if (yvar=="dlogdc") ylbl <- "Case Growth"
 
     tbl <- capture.output(stargazer(
         lapply(piy, function(x) x$reg),

### 2020-07-28 cases_and_policies/rmd/regprep.R

@@ -2,7 +2,8 @@ cs$t <- as.numeric(cs$date)
 
 warning("Using data only up to 2020-06-02 as in submitted version of paper. Modify lines 4-5 of regprep.R to change")
 df <- pdata.frame(subset(cs, cs$date>=as.Date("2020-01-01") &
-                             cs$date<=as.Date("2020-06-02")),
+                                   cs$date<=as.Date("2020-06-02")),
+                  # cs$date<=as.Date("2020-06-30")),
                   index=c("state","t"), stringsAsFactors=FALSE)
 
 #df$tests[df$testmissing] <- 0
@@ -85,7 +86,7 @@ statevars <- c("log(Population.2018)",
                "log(Square.Miles)",
                "Percent.Unemployed..2018.",
                "Percent.living.under.the.federal.poverty.line..2018.",
-               "Percent.at.risk.for.serious.illness.due.to.COVID", "logvote")
+               "Percent.at.risk.for.serious.illness.due.to.COVID") #, "logvote")
 bvars <- c("workplaces","retail","grocery","transit")
 
 for(i in 1:length(bvars)) df[,bvars[i]] <- df[,bvars[i]]/100 
@@ -123,6 +124,9 @@ df <- ndf[order(ndf$state, ndf$date),]
 df$z.mask <- df$z.mask/100
 df$mask_percent <- df$mask_percent/100
 
+df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
+#df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
+
 tilt = ifelse(df$date> as.Date("2020-05-01"),1,0 )
 month <- month(df$date)
 #month[month==6] <- 5
@@ -130,4 +134,5 @@ month <- month(df$date)
 df$month = as.factor(month)
 df$pmask.may <- df$pmaskbus*tilt
 df$pmask.april <- df$pmaskbus*(1-tilt)
-pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential") 
+#pols <- c("pmaskbus","pk12","pindex") 

### 2020-07-29 cases_and_policies/rmd/regressionsWithPlots.Rmd
update on us governor party
(i) add republican party dummy to state-level characteristics
(ii) change robustness table

@@
 tvars <- "dlogtests"  
-xlist <-list("", "") 
+# xlist <-list("party", "party") 
+xlist <-list(c("party","logvote"), c("party","logvote")) 
 interactions <-list("month", statevars)
@@
   sdf2 <- sdf[which(sdf$state!='New York'),]
-  piy[[(1+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+  piy[[(1+5*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
               ilist[[k]], 0, L=L)
-  # (2) adding mask wearing rate and log(Trump's vote)
+  # (2) adding mask wearing rate  
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-  piy[[(2+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  xlist <-list(c("z.mask"), c("z.mask"))
+  piy[[(2+5*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
               ilist[[k]], 0, L=L)
-  # (3) lagged behavior variables
+  # (2) adding log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("logvote","party"), c("logvote","party"))
+  piy[[(3+5*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (4) lagged behavior variables
   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
   xlist <-list("","")
-  piy[[(3+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(4+5*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L)
-  # (4) (1)-(3)
+  # (5) (1)-(4)
    sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
    sdf2 <- sdf[which(sdf$state!='New York'),]
-   xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-   piy[[(4+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+  xlist <-list(c("z.mask","logvote","party"), c("z.mask","logvote","party"))
+   piy[[(5+5*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L)
-
 }
@@
               ilist[[k]], 0, L=L)
-  # (2) adding mask wearing rate and log(Trump's vote)
+  # (2) adding mask wearing rate 
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-  piy[[(2+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  xlist <-list(c("z.mask"), c("z.mask"))
+  piy[[(2+5*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
               ilist[[k]], 0, L=L)
-  # (3) lagged behavior variables
+  # (3) adding log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("logvote"), c("logvote"))
+  piy[[(3+5*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (4) lagged behavior variables
   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
   xlist <-list("","")
-  piy[[(3+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(4+5*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L)
-  # (4) (1)-(3)
+  # (5) (1)-(4)
    sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
@@
    xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-   piy[[(4+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+   piy[[(5+5*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),

### 2020-07-29 tex/tables_and_figures/policies.tex
update
update on us governor party
(i) add republican party dummy to state-level characteristics
(ii) change robustness table

@@
 % latex table generated in R 4.0.0 by xtable 1.8-4 package
-% Sat Jul 25 16:38:15 2020
+% Wed Jul 29 09:27:24 2020
 \begin{table}[ht]
@@
   \hline
-State of emergency & 50 & 2020-02-29 & 2020-03-11 & 2020-03-16 \\ 
-  Date closed K 12 schools & 50 & 2020-03-13 & 2020-03-17 & 2020-04-03 \\ 
+State of emergency & 49 & 2020-02-29 & 2020-03-11 & 2020-03-16 \\ 
+  Date closed K 12 schools & 49 & 2020-03-13 & 2020-03-17 & 2020-04-03 \\ 
   Closed day cares & 14 & 2020-03-16 & 2020-03-23 & 2020-04-06 \\ 
   Date banned visitors to nursing homes & 29 & 2020-03-09 & 2020-03-16 & 2020-04-06 \\ 
-  Closed non essential businesses & 42 & 2020-03-19 & 2020-03-25 & 2020-04-06 \\ 
-  Closed restaurants except take out & 47 & 2020-03-15 & 2020-03-17 & 2020-04-03 \\ 
-  Closed gyms & 45 & 2020-03-16 & 2020-03-20 & 2020-04-03 \\ 
-  Closed movie theaters & 48 & 2020-03-16 & 2020-03-20 & 2020-04-06 \\ 
-  Stay at home  shelter in place & 41 & 2020-03-19 & 2020-03-28 & 2020-04-07 \\ 
-  End relax stay at home shelter in place & 37 & 2020-04-24 & 2020-05-16 & 2020-06-19 \\ 
-  Began to reopen businesses statewide & 49 & 2020-04-20 & 2020-05-07 & 2020-06-05 \\ 
-  Reopen restaurants & 47 & 2020-04-24 & 2020-05-18 & 2020-06-22 \\ 
-  Reopened gyms & 43 & 2020-04-24 & 2020-05-18 & 2020-07-03 \\ 
-  Reopened movie theaters & 36 & 2020-04-27 & 2020-05-31 & 2020-07-03 \\ 
+  Closed non essential businesses & 41 & 2020-03-19 & 2020-03-25 & 2020-04-06 \\ 
+  Closed restaurants except take out & 46 & 2020-03-15 & 2020-03-17 & 2020-04-03 \\ 
+  Closed gyms & 44 & 2020-03-16 & 2020-03-20 & 2020-04-03 \\ 
+  Closed movie theaters & 47 & 2020-03-16 & 2020-03-21 & 2020-04-06 \\ 
+  Stay at home  shelter in place & 40 & 2020-03-19 & 2020-03-28 & 2020-04-07 \\ 
+  End relax stay at home shelter in place & 36 & 2020-04-24 & 2020-05-15 & 2020-06-19 \\ 
+  Began to reopen businesses statewide & 48 & 2020-04-20 & 2020-05-06 & 2020-06-05 \\ 
+  Reopen restaurants & 46 & 2020-04-24 & 2020-05-17 & 2020-06-22 \\ 
+  Reopened gyms & 42 & 2020-04-24 & 2020-05-18 & 2020-07-03 \\ 
+  Reopened movie theaters & 35 & 2020-04-27 & 2020-05-30 & 2020-07-03 \\ 
   Resumed elective medical procedures & 34 & 2020-04-20 & 2020-04-29 & 2020-05-29 \\ 
-  Mandate face mask use by all individuals in public spaces & 31 & 2020-04-08 & 2020-06-18 & 2020-07-27 \\ 
-  Mandate face mask use by employees in public facing businesses & 41 & 2020-04-03 & 2020-05-01 & 2020-07-03 \\ 
+  Mandate face mask use by all individuals in public spaces & 30 & 2020-04-08 & 2020-06-22 & 2020-07-27 \\ 
+  Mandate face mask use by employees in public facing businesses & 40 & 2020-04-03 & 2020-05-01 & 2020-07-03 \\ 
   Stop Initiation of Evictions overall or due to COVID related issues & 29 & 2020-03-16 & 2020-03-24 & 2020-05-30 \\ 
-  Stop enforcement of evictions overall or due to COVID related issues & 28 & 2020-03-15 & 2020-03-24 & 2020-06-08 \\ 
+  Stop enforcement of evictions overall or due to COVID related issues & 27 & 2020-03-17 & 2020-03-24 & 2020-06-08 \\ 
   Renter grace period or use of security deposit to pay rent & 3 & 2020-04-10 & 2020-04-24 & 2020-05-20 \\ 
-  Order freezing utility shut offs & 33 & 2020-03-12 & 2020-03-19 & 2020-04-13 \\ 
+  Order freezing utility shut offs & 32 & 2020-03-12 & 2020-03-19 & 2020-04-13 \\ 
   Froze mortgage payments & 1 & 2020-03-21 & 2020-03-21 & 2020-03-21 \\ 

### 2020-07-29 cases_and_policies/rmd/debiasedfe.R

@@ -105,7 +105,9 @@ reg_fe <- function(df, # data
                  weights = df[subsample1, ]$sweight)
   cross2 <- felm(fmla[[1]], data = df[subsample2, ], 
                  weights = df[subsample2, ]$sweight)
-  sdf <- rbind(df[subsample1, ], df[subsample2, ])
+  sdf2 <- df[subsample2, ]
+  sdf2$state <- paste(sdf2$state,2, sep = "-")
+  sdf <- rbind(df[subsample1, ], sdf2)
   cross3 <- felm(fmla[[1]], data = sdf, 
                  weights = sdf$sweight)
   # Debiased estimates

### 2020-07-29 cases_and_policies/rmd/regprep.R

@@ -86,7 +86,7 @@ statevars <- c("log(Population.2018)",
                "log(Square.Miles)",
                "Percent.Unemployed..2018.",
                "Percent.living.under.the.federal.poverty.line..2018.",
-               "Percent.at.risk.for.serious.illness.due.to.COVID") #, "logvote")
+               "Percent.at.risk.for.serious.illness.due.to.COVID","party")#, "logvote")
 bvars <- c("workplaces","retail","grocery","transit")
 
 for(i in 1:length(bvars)) df[,bvars[i]] <- df[,bvars[i]]/100 
@@ -123,6 +123,8 @@ df <- ndf[order(ndf$state, ndf$date),]
 
 df$z.mask <- df$z.mask/100
 df$mask_percent <- df$mask_percent/100
+
+df$party <- as.factor(df$party)
 
 df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
 #df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3

### 2020-07-31 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

--- 2020-07-28 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 tvars <- "dlogtests"  
-xlist <-list("", "") 
+# xlist <-list("party", "party") 
+# xlist <-list(c("party","logvote"), c("party","logvote")) 
+xlist <- list("","")
 interactions <-list("month", statevars)
@@
 
+# pi-y with log(vote)
+xlist <-list(c("logvote"), c("logvote")) 
+L <- 14
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+regs <- mainregressions(sdf,
+                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
+showhtmltables(NULL, NULL, NULL, regs$ip)
+savetextables(NULL, NULL, NULL, prefix="policy",ip = regs$ip[1:6])
+
 L <- 14
@@
 for (k in 1:length(infovars)) { 
-  # (1) excluding NY 
+  # (1) policy index 
+  pols <- c("pmaskbus","pk12","pindex") 
+   xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
+  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L) 
+  # (2) excluding NY 
+  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential") 
    xlist <-list("","")
@@
   sdf2 <- sdf[which(sdf$state!='New York'),]
-  piy[[(1+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
               ilist[[k]], 0, L=L)
-  # (2) adding mask wearing rate and log(Trump's vote)
+  # (3) adding mask wearing rate  
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-  piy[[(2+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  xlist <-list(c("z.mask"), c("z.mask"))
+  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
               ilist[[k]], 0, L=L)
-  # (3) lagged behavior variables
+  # (4) adding log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("logvote","party"), c("logvote","party"))
+  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (5) lagged behavior variables
   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
   xlist <-list("","")
-  piy[[(3+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L)
-  # (4) (1)-(3)
+  # (6) (2)-(5)
    sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
    sdf2 <- sdf[which(sdf$state!='New York'),]
-   xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-   piy[[(4+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+  xlist <-list(c("z.mask","logvote","party"), c("z.mask","logvote","party"))
+   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L)
-
 }
 
+colnames(piy[[1]]$reg$response) <- "dlogdc" 
 showhtmltables(NULL, NULL, piy);
-# savetextables(NULL, NULL,  piy, prefix="cases-robust")
+savetextables(NULL, NULL,  piy[1:6], prefix="cases-robust")
 
@@
 for (k in 1:length(infovars)) {
-  # (1) excluding NY 
+
+  # (1) policy index
+  pols <- c("pmaskbus","pk12","pindex") 
+  xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
+  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L) 
+  # (2) excluding NY 
+  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential") 
   xlist <-list("","")
@@
               ilist[[k]], 0, L=L)
-  # (2) adding mask wearing rate and log(Trump's vote)
+
+  # (3) adding mask wearing rate 
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-  piy[[(2+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  xlist <-list(c("z.mask"), c("z.mask"))
+  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
               ilist[[k]], 0, L=L)
-  # (3) lagged behavior variables
+  # (4) adding log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("logvote"), c("logvote"))
+  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (5) lagged behavior variables
   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
   xlist <-list("","")
-  piy[[(3+4*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L)
-  # (4) (1)-(3)
+  # (6) (2)-(5)
    sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
@@
    xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
-   piy[[(4+4*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
 }
+colnames(piy[[1]]$reg$response) <- "dlogdd"  
 showhtmltables(NULL, NULL, piy);
-# savetextables(NULL, NULL,  piy, prefix="deaths-robust")
+savetextables(NULL, NULL,  piy[1:6], prefix="deaths-robust")
 
@@
   piy[[(1+4*(k-1))]] <- m
+  #m$reg$coefficients[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))] 
   m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))] 
-  #m$reg$beta[1:(6+2*(k-1))] <- m$bc[1:(6+2*(k-1))]  
+  # m$reg$tval <- m$reg$beta/m$reg$se
+  # m$reg$rpval <- pnorm(-abs(m$reg$beta), sd=m$reg$se)
   piy[[(2+4*(k-1))]] <- m
@@
   piy[[(3+4*(k-1))]] <- m 
+  #m$reg$coefficients[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
   m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))]
+  # m$reg$tval <- m$reg$beta/m$reg$se
+  # m$reg$pval <- pnorm(-abs(m$reg$beta), sd=m$reg$se) 
   #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
@@
 
+colnames(piy[[1]]$reg$response) <- "dlogdc"  
 showhtmltables(NULL, NULL,  piy);
-# savetextables(NULL, NULL, piy, prefix="cases-fe")
+prefix="cases-fe"
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+   ylbl <- colnames(piy[[1]]$reg$response)
+    if (ylbl=="dlogdd") ylbl <- "Death Growth"
+    if (ylbl=="dlogdc") ylbl <- "Case Growth"
+
+    tbl <- capture.output(stargazer(
+        lapply(piy, function(x) x$reg),
+        type="latex",
+        dep.var.labels.include=FALSE,
+        title=ylbl,
+        # omit=omit,
+        # omit.labels=omit.labels,
+        column.labels=c(yvar),
+        column.separate=c(length(piy)),
+        # add.lines=list(c("$\\sum_j \\mathrm{Policy}_j$",
+        #                  sapply(piy, function(x) printstars(x$peff[1], x$peff[2]))),
+        #                c("",
+        #                  sprintf("(%.3f)",sapply(piy, function(x) x$peff[2])))
+        #                ),
+        omit.stat=c("f", "ser"), model.names=FALSE,
+        no.space=TRUE,
+        column.sep.width="1pt",
+        df=FALSE, header=FALSE,
+        model.numbers=TRUE))
+    tbl <- relabel(tbl)
+    tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
+    #tbl <- gsub("No","Yes", tbl)
+    texfile <- sprintf("%s/tex/tables_and_figures/%s-piy.tex",rootdir,prefix)
+    cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
 
@@
   #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
+  # m$reg$tval <- m$reg$beta/m$reg$se
+  # m$reg$pval <- pnorm(-abs(m$reg$beta), sd=m$reg$se)
   piy[[(2+4*(k-1))]] <- m
@@
   #m$reg$beta[1:(4+2*(k-1))] <- m$bc[1:(4+2*(k-1))] 
+  # m$reg$tval <- m$reg$beta/m$reg$se
+  # m$reg$pval <- pnorm(-abs(m$reg$beta), sd=m$reg$se) 
   piy[[(4+4*(k-1))]] <- m
@@
 
+colnames(piy[[1]]$reg$response) <- "dlogdd"  
 showhtmltables(NULL, NULL,  piy);
-# savetextables(NULL, NULL, piy, prefix="deaths-fe")
+#savetextables(NULL, NULL, piy[1:4], prefix="deaths-fe")
+
 ```

### 2020-07-31 cases_and_policies/rmd/generatetables.R

@@ -100,8 +100,8 @@ mainregressions <- function(df, # data
                        L=14) {
 
 
-  # plist <- list(pols, c("pmask.april","pmask.may", pols[-1]))
-  plist <- list(pols, c("pmaskbus","pk12","pindex"))
+  plist <- list(pols, c("pmask.april","pmask.may", pols[-1]))
+  # plist <- list(pols, c("pmaskbus","pk12","pindex"))
   
   
 
@@ -284,7 +284,7 @@ showhtmltables <- function(pib, pbiy, piy, ip=NULL) {
 }
 
 
-savetextables <- function(pib, pbiy, piy, prefix,
+savetextables <- function(pib, pbiy, piy, prefix, ip = NULL,
                           rootdir=system("git rev-parse --show-toplevel", intern=TRUE)[1])
 {
   omit <- c(statevarregexp,
@@ -360,9 +360,9 @@ savetextables <- function(pib, pbiy, piy, prefix,
   if (!(is.null(piy))) {
 
    #  yvar <- colnames(pbiy[[1]]$reg$response)
-    # yvar <- colnames(piy$reg$response)
-    # if (yvar=="dlogdd") ylbl <- "Death Growth"
-    # if (yvar=="dlogdc") ylbl <- "Case Growth"
+    ylbl <- colnames(piy[[1]]$reg$response)
+    if (ylbl=="dlogdd") ylbl <- "Death Growth"
+    if (ylbl=="dlogdc") ylbl <- "Case Growth"
 
     tbl <- capture.output(stargazer(
         lapply(piy, function(x) x$reg),
@@ -388,6 +388,38 @@ savetextables <- function(pib, pbiy, piy, prefix,
     #tbl <- gsub("No","Yes", tbl)
     texfile <- sprintf("%s/tex/tables_and_figures/%s-piy.tex",rootdir,prefix)
     cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
+  }
+  
+  
+  if (!(is.null(ip))) {
+    
+  beff <- sapply(ip, function(x) x$beff[1])
+  sep <- sapply(ip, function(x) x$beff[2])
+  cnames <- sapply(ip, function(x) colnames(x$reg$response))
+  tbl <- capture.output(stargazer(
+    lapply(ip, function(x) x$reg),
+    type="latex",
+    title="Inofrmation and Policies",
[…]

### 2020-07-31 cases_and_policies/rmd/regressionsWithPlots.Rmd
Some simple updates.

--- 2020-07-28 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
-rootdir <- "/Users/Hiro/Documents/GitHub/covid-impact" 
 source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
@@
 
-<!-- ## Deaths (with cases as information) -->
-
-<!-- ```{r deathtables, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'} -->
-<!-- infovars <- list(c("dlogdc", "logdc"), -->
-<!--                  c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #, -->
-<!-- df$zero <- 0 -->
-<!-- L <- 21 -->
-<!-- sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) -->
-<!-- regs <- mainregressions(sdf, -->
-<!--                         "dlogdd", pols, bvars, infovars, "zero", xlist, ilist, iv, L=L) -->
-<!-- showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy); -->
-<!-- savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-casesinfo") -->
-<!-- ``` -->
-
 ## Deaths (with deaths as information)
@@
 # savetextables(NULL, NULL,  piy, prefix="deaths-robust")
-
 ```
@@
   piy[[(1+4*(k-1))]] <- m
-  m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))] 
+  #m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))]
+  m$reg$beta <- m$bc2
   #m$reg$beta[1:(6+2*(k-1))] <- m$bc[1:(6+2*(k-1))]
@@
   piy[[(3+4*(k-1))]] <- m
-  m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
-  #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
+  #m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))]
+  m$reg$beta <- m$bc2
   piy[[(4+4*(k-1))]] <- m

### 2020-07-31 cases_and_policies/rmd/debiasedfe.R

@@ -1,10 +1,10 @@
 # 7/27/2020 Hiro Kasahara
 # Modification of the code written by Shuowen Chen ``debiasedfe_shuowen.R''
 # 7/3/2020 Shuowen Chen
-# The script contains code that implement debiased fixed effects and panel bootstrap 
+# The script contains code that implement debiased fixed effects and panel bootstrap
 # standard errors for dataset in Chernozhukov, Kasahara and Schrimpf (2020)
 # Serves as application for Chen, Chernozhukov and Fernandez-Val (2020)
-# Acknowledgements: functions extend those written by Paul Schrimpf. 
+# Acknowledgements: functions extend those written by Paul Schrimpf.
 
 library(lfe)
 library(speedglm)
@@ -18,28 +18,28 @@ sdf$sweight <- 1
 sdf$week <- as.factor(strftime(sdf$date, format = "%V"))
 
 #### Functions ####
-# Create formula 
+# Create formula
 # The function extends the createfmla to accommodate FE specifications
 # yvar: dependent variable for bpiy and piy
-# xvars: regressors. 
+# xvars: regressors.
 #  (a) For bpiy, regressors are behavior, policy and information
-#  (b) For pib, regressors are policy and information, dependent variable is 1 
+#  (b) For pib, regressors are policy and information, dependent variable is 1
 #      of the 4 behavior variables
 #  (c) For piy, regressors are policy and information variables
-# iv: This is to accommodate the syntax of felm. "0" means no iv. 
-# cluster: cluster at state level (treatment). 
-# Note: interacted fixed effect hasn't been tested. 
+# iv: This is to accommodate the syntax of felm. "0" means no iv.
+# cluster: cluster at state level (treatment).
+# Note: interacted fixed effect hasn't been tested.
 createfmla_fe <- function(yvar, xvars, iv = "0", cluster = "state",
-                          fe = c("state + month", "state * month", 
+                          fe = c("state + month", "state * month",
                                  "week + state", "date + state")) {
   rhs <- paste(xvars, collapse = " + ")
   if (fe == "state * month") {
     # we will use speedlm, which is faster
     rhs <- paste(rhs, "+", fe, collapse = " ")
   } else {
[…]

### 2020-08-01 cases_and_policies/rmd/regressionsWithPlots.Rmd

--- 2020-07-31 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 tvars <- "dlogtests"
-# xlist <-list("party", "party") 
-# xlist <-list(c("party","logvote"), c("party","logvote")) 
-xlist <- list("","")
+xlist <-list("", "")
 interactions <-list("month", statevars)
@@
 ```
-
-<!-- ## Deaths (with cases as information) -->
-
-<!-- ```{r deathtables, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'} -->
-<!-- infovars <- list(c("dlogdc", "logdc"), -->
-<!--                  c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #, -->
-<!-- df$zero <- 0 -->
-<!-- L <- 21 -->
-<!-- sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) -->
-<!-- regs <- mainregressions(sdf, -->
-<!--                         "dlogdd", pols, bvars, infovars, "zero", xlist, ilist, iv, L=L) -->
-<!-- showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy); -->
-<!-- savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-casesinfo") -->
-<!-- ``` -->
 
@@
 
-#df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
-df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
+df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
+#df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
 
@@
 piy <- list()
+
+#' Update the coefficients of felm object.  Also updates the t stats
+#' and p values, so that summary and stargazer work correctly after.
+replacecoef <- function(reg, newcoef) {
+  reg$beta[] <- newcoef
+  reg$coefficients[] <- newcoef
+  reg$tval <- reg$beta/reg$se
+  reg$pval <- pt(-abs(reg$tval),df=reg$df)*2
+  reg$rtval <- reg$beta/reg$rse
+  reg$rpval <- pt(-abs(reg$rtval),df=reg$df)*2
+  reg$ctval <- reg$beta/reg$cse
+  reg$cpval <- pt(-abs(reg$ctval),df=reg$df)*2
+  return(reg)
+}
+
 for (k in 1:length(infovars)) {
@@
   piy[[(1+4*(k-1))]] <- m
-  #m$reg$coefficients[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))] 
-  m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))] 
-  # m$reg$tval <- m$reg$beta/m$reg$se
-  # m$reg$rpval <- pnorm(-abs(m$reg$beta), sd=m$reg$se)
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(2+4*(k-1))]] <- m
   pols <- c("pmaskbus","pk12","pindex")
-  # pols <- c("pmaskbus","pk12","pmovie","pindex") 
   m <-  reg_fe(sdf, yvar, pols, NULL,
@@
   piy[[(3+4*(k-1))]] <- m
-  #m$reg$coefficients[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
-  m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))]
-  # m$reg$tval <- m$reg$beta/m$reg$se
-  # m$reg$pval <- pnorm(-abs(m$reg$beta), sd=m$reg$se) 
-  #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(4+4*(k-1))]] <- m
@@
 
-colnames(piy[[1]]$reg$response) <- "dlogdc"  
 showhtmltables(NULL, NULL,  piy);
 prefix="cases-fe"
-#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe") copying and
+
+# pasting this is bad style and seems like it might create maintenance
+# problems ...
    ylbl <- colnames(piy[[1]]$reg$response)
@@
   piy[[(1+4*(k-1))]] <- m
-  m$reg$beta[1:(8+2*(k-1))] <- m$bc2[1:(8+2*(k-1))] 
-  #m$reg$beta[1:(5+2*(k-1))] <- m$bc[1:(5+2*(k-1))] 
-  # m$reg$tval <- m$reg$beta/m$reg$se
-  # m$reg$pval <- pnorm(-abs(m$reg$beta), sd=m$reg$se)
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(2+4*(k-1))]] <- m
   pols <- c("pmaskbus","pk12","pindex")
-  #pols <- c("pmaskbus","pk12","pmovie","pindex") 
   m <-  reg_fe(sdf, yvar, pols, NULL,
@@
   piy[[(3+4*(k-1))]] <- m
-  m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))] 
-  #m$reg$beta[1:(4+2*(k-1))] <- m$bc[1:(4+2*(k-1))] 
-  # m$reg$tval <- m$reg$beta/m$reg$se
-  # m$reg$pval <- pnorm(-abs(m$reg$beta), sd=m$reg$se) 
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(4+4*(k-1))]] <- m

### 2020-08-03 cases_and_policies/rmd/regressionsWithPlots.Rmd
update on DML

--- 2020-07-31 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 library(reshape2)
+library(AER)
+library(hdm)
+library(randomForest)
 
@@
 source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
+source(paste(rootdir,"cases_and_policies/rmd/rlassoPanel.R", sep="/"))
 ```
@@
 ```{r corr, cache=FALSE}
-cmat <-  cor(df[,c(bvars,pols)], use="pairwise.complete.obs")
+
+cmat <-  cor(df[,c(bvars,pols,"pindex")], use="pairwise.complete.obs")
 scmat <-
-  matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)))
+  matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)+1))
 
@@
 
-```{r casedieff, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
+```{r casedieff, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 for (use.national in c(TRUE,FALSE)) {
@@
 
+# L <- 14
+#   xlist <- list("","")
+#   plist <- list(pols, pols)
+#   ijs <- expand.grid(1:length(bvars), 1:length(plist))
+#   pib <- list()
+#   for (k in 1:length(xlist)) {
+#     pib[[k]] <- apply(ijs, 1, function(ij) {
+#       policyreg(df, bvars[ij[1]], NULL,  bvars,
+#                 c(plist[[ij[2]]],infovars[[k]],
+#                   xlist[[k]]),
+#                 ilist[[k]], "0", L=14)
+#     })
+#   }
+
 # pi-y with log(vote)
@@
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("z.mask"), c("z.mask"))
+  xlist <-list(c("mask_percent"), c("mask_percent"))
   piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
@@
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("logvote","party"), c("logvote","party"))
+  xlist <-list(c("logvote"), c("logvote"))
   piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
@@
    sdf2 <- sdf[which(sdf$state!='New York'),]
-  xlist <-list(c("z.mask","logvote","party"), c("z.mask","logvote","party"))
+  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
    piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
@@
    sdf2 <- sdf[which(sdf$state!='New York'),]
-   xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
    piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
@@
     tbl <- capture.output(stargazer(
-        lapply(piy, function(x) x$reg),
+        lapply(piy[1:4], function(x) x$reg),
         type="latex",
@@
         column.labels=c(yvar),
-        column.separate=c(length(piy)),
+        column.separate=c(length(piy[1:4])),
         # add.lines=list(c("$\\sum_j \\mathrm{Policy}_j$",
@@
 showhtmltables(NULL, NULL,  piy); 
-#savetextables(NULL, NULL, piy[1:4], prefix="deaths-fe")
+prefix="deaths-fe"
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+   ylbl <- colnames(piy[[1]]$reg$response)
+    if (ylbl=="dlogdd") ylbl <- "Death Growth"
+    if (ylbl=="dlogdc") ylbl <- "Case Growth"
+
+    tbl <- capture.output(stargazer(
+        lapply(piy[1:4], function(x) x$reg),
+        type="latex",
+        dep.var.labels.include=FALSE,
+        title=ylbl,
+        # omit=omit,
+        # omit.labels=omit.labels,
+        column.labels=c(yvar),
+        column.separate=c(length(piy[1:4])),
+        # add.lines=list(c("$\\sum_j \\mathrm{Policy}_j$",
+        #                  sapply(piy, function(x) printstars(x$peff[1], x$peff[2]))),
+        #                c("",
+        #                  sprintf("(%.3f)",sapply(piy, function(x) x$peff[2])))
+        #                ),
+        omit.stat=c("f", "ser"), model.names=FALSE,
+        no.space=TRUE,
+        column.sep.width="1pt",
+        df=FALSE, header=FALSE,
+        model.numbers=TRUE))
+    tbl <- relabel(tbl)
+    tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
+    #tbl <- gsub("No","Yes", tbl)
+    texfile <- sprintf("%s/tex/tables_and_figures/%s-piy.tex",rootdir,prefix)
+    cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
+
+```
+
+## DML 
+```{r regsdml, cache=FALSE, warning=FALSE, results='asis'}
+
+library(glmnet)
+
+df$logpop <- log(df$Population.2018)
+df$logsq <- log(df$Square.Miles) 
+df$repub <- as.factor(df$party)
+df$party_dum <- ifelse(df$party == "republican", 0, 1) 
+df$month3 <- ifelse(df$month==3,0,1)
+df$month4 <- ifelse(df$month==4,0,1)
+df$month5 <- ifelse(df$month==5,0,1)
+df$month6 <- ifelse(df$month==6,0,1)
+L <- 14
+df$dlogdc.L14 <- panellag(df$dlogdc,df$state,df$date, lag=L)
+df$logdc.L14 <- panellag(df$logdc,df$state,df$date, lag=L)
+df$dlogdd.L14 <- panellag(df$dlogdd,df$state,df$date, lag=L)
+df$logdd.L14 <- panellag(df$logdd,df$state,df$date, lag=L)  
+df$pmaskbus.L14 <- panellag(df$pmaskbus,df$state,df$date, lag=L) 
+df$pk12.L14 <- panellag(df$pk12,df$state,df$date, lag=L) 
+df$pmovie.L14 <- panellag(df$pmovie,df$state,df$date, lag=L) 
+df$pshelter.L14 <- panellag(df$pshelter,df$state,df$date, lag=L) 
+df$prestaurant.L14 <- panellag(df$prestaurant,df$state,df$date, lag=L) 
+df$pnonessential.L14 <- panellag(df$pnonessential,df$state,df$date, lag=L)  
+df$pindex.L14 <- panellag(df$pindex,df$state,df$date, lag=L)
+
+xvars <- c("logpop",
+               "logsq",
+               "Percent.Unemployed..2018.",
+               "Percent.living.under.the.federal.poverty.line..2018.",
+               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
+               "mask_percent", "logvote",
+               "month3","month4","month5","month6","dlogdc.L14","logdc.L14")
+
+pols.L14 <- c("pmaskbus.L14","pk12.L14","pmovie.L14","pshelter.L14","prestaurant.L14","pnonessential.L14")
+
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+
+DML2.for.PLM <- function(x, d, y, dreg, yreg, sdf, nfold=5) {   
+  unit <- as.double(sdf$state)
+  j <-  quantile(unit, c(1:(nfold-1))/nfold)
+  I <- vector(mode = "list", length = nfold)
+  I[[1]] <- unit <= j[1] 
+  for (b in 2:(nfold-1)) {
+    I[[b]] <- (unit <= j[b])&(unit > j[b-1])
+  } 
+  I[[nfold]] <- unit > j[nfold-1]  
+  nobs <- nrow(x)
+  ytil <- dtil <- rep(NA, nobs)
+  for(b in 1:nfold){
+    I1 <- (I[[b]]==FALSE)
+    I2 <- I[[b]]
+    dfit <- dreg(x[I1,], d[I1]) # take a fold out
+    yfit <- yreg(x[I1,], y[I1]) # take a folot out
+    dhat <- predict(dfit, x[I2,], type="response") #predict the fold out
+    yhat <- predict(yfit, x[I2,], type="response") #predict the fold out
+    dtil[I2] <- (d[I2] - dhat) #record residual
+    ytil[I2] <- (y[I2] - yhat) #record residual
+    cat(b," ")
+  }  
+sdf$ytil <- ytil
+sdf$dtil <- dtil 
+rfit <- felm(ytil ~ dtil| 0| 0 | state, data=sdf) #estimate the main parameter by regressing one residual on the other
+coef.est <- coef(rfit)[2] #extract coefficient 
+se <- sqrt(vcov(rfit)[2,2]) 
+cat(sprintf("\ncoef (se) = %g (%g)\n", coef.est , se))
+return( list(coef.est = coef.est , se=se) )
+}
+
+# dreg <- function(x,d){ rlasso(x, d, c=2, POST=FALSE) } #ML method = rlasso
+# yreg <- function(x,y){ rlasso(x, y, c=2, POST=FALSE) } #ML method = rlasso 
+# DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf)
+
+x = as.matrix(cbind(sdf[,pols.L14[-1]],sdf[,xvars]))
+d = sdf$pmaskbus.L14
+y = sdf$dlogdc
+
+dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)
+
+dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
+yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
+
+x = as.matrix(cbind(sdf[,pols.L14[-4]],sdf[,xvars]))
+d = sdf$pshelter.L14
+y = sdf$dlogdc
+
+dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)
+
+dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
+yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
+
+x = as.matrix(cbind(sdf[,pols.L14[1:2]],sdf[,xvars]))
+d = sdf$pindex.L14
+y = sdf$dlogdc
+
+dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)
+
+dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
+yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
+
+xvars <- c("logpop",
+               "logsq",
+               "Percent.Unemployed..2018.",
+               "Percent.living.under.the.federal.poverty.line..2018.",
+               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
+               "mask_percent", "logvote",
+               "month3","month4","month5","month6","dlogdd.L14","logdd.L14")
+
+y = sdf$dlogdd  
+dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)
+
+dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
+yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
+
+x = as.matrix(cbind(sdf[,pols.L14[-4]],sdf[,xvars]))
+d = sdf$pshelter.L14
+
+dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)
+
+dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
+yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
+
+x = as.matrix(cbind(sdf[,pols.L14[1:2]],sdf[,xvars]))
+d = sdf$pindex.L14
+y = sdf$dlogdd
+
+dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 5)
+set.seed(1)
+DML2.lasso <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold = 10)
+
+dreg <- function(x,d){ randomForest(x, d) } #ML method=Forest
+yreg <- function(x,y){ randomForest(x, y) } #ML method=Forest
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=5)
+set.seed(1)
+DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
 

### 2020-08-03 tex/robustness.tex
update on DML

--- 2020-07-16 tex/covidEffectOfNPI-COVIDEconomics.tex
@@
 \begin{abstract}
-This paper evaluates the dynamic impact of various policies adopted by US states on the growth rates of confirmed Covid-19  cases and deaths as well as social distancing behavior measured by Google Mobility Reports, where we take into consideration people's voluntarily behavioral response to new information of transmission risks.
+%This paper evaluates the dynamic impact of various policies adopted by US states on the growth rates of confirmed Covid-19  cases and deaths as well as social distancing behavior measured by Google Mobility Reports, where we take into consideration people's voluntarily behavioral response to new information of transmission risks.
 Our analysis finds that both policies and information on transmission risks are important determinants of Covid-19  cases and deaths and shows that a change in policies explains a large fraction of observed changes in social distancing behavior.
@@
 %\tableofcontents
-\thispagestyle{empty}
 
-\pagebreak
+\section{Robustness}
 
-\pagestyle{empty}
+In this section, we examine the robustness of our results by including additional controls and examining the sub-sample that excludes the state of New York.
 
- \section{Introduction}
+Table \ref{tab:PtoY-robust} presents the regression result for robustness check, where dependent variable is the weekly growth rate of
+  cases  (left panel) or deaths (right panel).
+Column (1) replaces four policy variables  (stay-at-home, closure of movie theaters, closure of restaurants, and closure of non-essential businesses) that are highly correlated to each other with the variable of their average values.
+Column (2) excludes the state of New York, which could be viewed as an outlier in the early pandemic period, from the sample.
+ Column (3) adds the percentage of people who wears masks to protect themselves between March 26-April 29, 2020, which controls for people's attitude toward mask wearing in the first half of sample.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.}  Column (5) controls for the log of Trump voting shares in 2016 presidential election.
+Column (6) includes four weeks lagged behavior variables as an additional set of controls; under this specification, our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables.
+Finally, column (6) includes all additional controls in column (3)-(5) using the sample that excludes New York.
+The estimated coefficients of mask mandates on case or death growths are significantly negative in all specifications, confirming the importance of mask policy on reducing case and death growths.
+\textbf{[here, instead of table, we may present ``dot-and-whisker plot'' for mask coefficients as well as stay-at-home and pindex.]}
 
-Accumulating evidence suggests that various policies in the US have reduced social interactions and slowed down the growth of Covid-19  infections.\footnote{
-See
-\citet{courtemanche2020}, \citet{hsiang2020}, \citet{pei2020},  \citet{abouk2020}, and \cite{wright2020}.} An important outstanding issue, however, is how much of the observed slow down in the spread is attributable to the effect of policies as opposed to a voluntarily change in people's behavior out of fear of being infected.
-This question is critical for evaluating the effectiveness of restrictive policies  in the US relative to an alternative policy of just providing recommendations and information such as the one adopted by Sweden.
-More generally, understanding people's dynamic  behavioral response to policies and information  is indispensable for properly evaluating the effect of policies on the spread of Covid-19.
+In Table \ref{tab:PtoY-fe} of the appendix,   we find that the estimated coefficients of mask mandates are significantly negative even when we include state fixed effects.
+In contrast,  the estimated coefficients of stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses under fixed effects specification are sometimes different from those under random effects specification, especially when bias correction is applied to the fixed effects estimator.
+Non-robustness of the estimated coefficients of these four  policy variables
+ is likely due to high correlations among them (see Table \ref{corr}). In fact, as shown in columns (3)-(4) of Table \ref{tab:PtoY-fe}, the  estimated effects of the average of four policy variables 
+ under fixed effects specification are negative and significant, indicating that  the total effect of four policies is robustly estimated.
 
-This paper  quantitatively assesses the impact of various policies adopted by US states on the spread of Covid-19, such as non-essential business closure and mandatory  face masks, paying particular attention to how people adjust their behavior in response to policies as well as new information on cases and deaths.
+\subsection{Fixed Effects Regressions (in Appendix)}
 
-We present a conceptual framework that spells out the causal structure on how the Covid-19 spread  is dynamically determined by policies and human behavior.
-Our approach explicitly recognizes that policies not only directly affect the spread of Covid-19 (e.g., mask requirement) but also indirectly affect its spread by changing people's behavior (e.g., stay-at-home order).
-It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
-Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007}; see \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book} for  modern developments, especially in computer science and epidemiology), with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-As such it naturally converses in all languages for causal inference.}
-% It also allows us  to explicitly define counterfactual scenarios to properly evaluate the effect of alternative policies on the spread of Covid-19.
+Table \ref{tab:PtoY-fe} in the appendix reports the regression result when we include state fixed effects and month dummies.
+Column (1) reports our baseline estimation when we include state fixed effects while column (2) report the estimate when we apply the bias correction to the fixed effects estimator.
+Columns (3) and (4) are similar to columns (1) and (2) except that we replace  four policy variables  (stay-at-home, closure of movie theaters, closure of restaurants, and closure of non-essential businesses) with their average values.
+Again, the estimated coefficients of mask mandates are significantly negative across different specifications, confirming the importance of mask policy on reducing case and death growths.
+On the other hand,  the estimated coefficients of stay-at-home order as well as closures of movie theaters, restaurants, and non-essential businesses under fixed effects specification are substantially different from the estimates under random effects specification.
+This is likely due to muti-colinearity problem as implied by  high correlations among these four policy variables as shown in Table \ref{corr}.
+When we replace these four policy variables with their average variable in columns (3) and (4), the estimated effect of the sum of four policy variables under fixed effects specification is similar to that under random effects specification in column (1) of Table \ref{tab:PtoY-robust}.
 
-Guided by the causal model, our empirical analysis examines how the weekly growth rates of confirmed Covid-19 cases and deaths are determined by (the lags of) policies and behavior using US state-level data.
-To examine how policies and information affect people's behavior, we also regress social distancing measures  on policy and information variables.
-Our regression specification for case and death growths is explicitly guided by a SIR model although  our causal approach does not hinge on the validity of a SIR model.
-
-As policy variables, we consider  mandatory face masks for employees in public businesses, stay-at-home orders (or shelter-in-place orders), closure of K-12 schools, closure of restaurants except take out, closure of movie theaters, and closure of non-essential businesses.
-Our behavior variables are four mobility measures that capture the intensity of visits to ``transit,'' ``grocery,'' ``retail,'' and ``workplaces''  from Google Mobility Reports.
-We take the lagged growth rate of cases and deaths and the log of lagged cases and deaths at both state-level and national-level as our measures of information on infection risks that affects people's  behavior.
-We also consider the growth rate of tests, month dummies, and state-level characteristics (e.g., population size and total area) as confounders that have to be controlled for in order to identify the causal relationship between policy/behavior and the growth rate of cases and deaths.
-
-Our key findings from regression analysis are as follows.  We find
-that both policies and information on past cases and deaths are important
-determinants of people's social distancing behavior, where policy
-effects explain more than $50\%$ of the observed decline in the four
-behavior variables.\footnote{The behavior accounts for the other half. This is in line with theoretical study by \cite{gitmez2020} that investigates the role of private behavior and
-negative external effects for individual decisions over policy compliance as well as information acquisition during pandemics.} Our estimates suggest that there are both large policy effects and large behavioral effects on the growth of cases and deaths.
-Except for mandatory masks, the effect of policies on cases and deaths is indirectly materialized through their impact on behavior; the effect of mandatory mask policy is direct without affecting behavior.
-
-Using the estimated model, we evaluate the dynamic impact of the following counterfactual policies on Covid-19 cases and deaths: mandating face masks, allowing non-essential businesses to open, not implementing a stay-at-home order, and removing all policies.
-The counterfactual experiments show a large impact of  those policies on  the number of cases and deaths.
-They also highlight the importance of voluntary behavioral response to infection risks for evaluating the dynamic policy effects.
-
-Figure \ref{fig:US-mask-dgrowth} shows that nationally  implementing mandatory
-face masks  for employees in public businesses on April 1st would have reduced the growth rate of cases (top panel)
-and that of deaths (bottom panel) by more than 10 percentage points in late April.
-This leads to reductions of $25$\% and $35$\% in reported cases and deaths, respectively,  by the end of May with a 90 percent confidence interval of $[10,45]$\% and $[17,55]$\%, which  roughly implies that as many as 17 to 55
-  thousand lives could have been saved.\footnote{As of May 27, 2020,
-   the US Centers for Disease Control and Prevention  reports 99,031
-   deaths in the US.}
-This finding is significant: given this potentially large  benefit of reducing the spread of Covid-19, mandating masks is an attractive policy instrument especially because it involves relatively little economic disruption.
-These estimates contribute to the ongoing efforts towards designing approaches to minimize risks from reopening \citep{stock2020b}.
-
-Figure \ref{fig:US-nb-dgrowth} illustrates how allowing non-essential businesses to remain open could
-have affected the growth of cases.  We estimate that
-non-essential business closures have a small impact on growth rates,
-with a 90\% confidence interval that includes both negative and
-positive effects. When this effect on growth rates is converted to a
-change in levels, the point estimates indicate that keeping
-non-essential businesses open (other than movie theaters, gyms, and
-keeping restaurants in the ``take-out" mode) could have increased
-cases and  deaths by  $15\%$ (with a 90 percent confidence interval of
-$-20\%$ to $60\%$).  These estimates contribute to the ongoing efforts of evaluating various reopening approaches.
-
-In Figure \ref{fig:US-shelter-dgrowth}, we find that, without stay-at-home orders, the case
-growth rate would have been nearly 10 percentage points higher in late April.
-No stay-at-home orders could have led to 80\% more cases by the start of June with a 90 precent confidence interval given by  $25\%$ to $170\%$.
-This implies that 0.5 to 3.4 million more Americans would have been infected without stay-at-home orders, providing suggestive evidence that reopening via removal of stay-at-home orders could lead to a substantial increase in cases and deaths.
-
-In our counterfactual experiment of removing all policies, we find that the results are sensitive to whether  the number of past national cases/deaths is included  in a specification or not.
-This sensitivity arises because there is little variation across states in the timing of school closures.
-This makes the effect of school closures difficult to identify.
-In Figure \ref{fig:US-nop-SI}, we show that in a specification that excludes past national cases (which allow for greater attribution of effects to school closures), the number of cases by the end of May could have increased 7-fold or more with a very large upper bound.
-On the other hand, as shown in Figure \ref{fig:US-nop}, under a specification with past national cases, our counterfactual experiment implies a 0 to 10 fold increase in cases by the end of May.
-This highlights the uncertainty regarding the impact of all policies versus private behavioral responses to information.
-Evaluation of re-opening policies needs to be aware of this uncertainty.
-
-%Figure \ref{fig:US-nop-dgrowth} shows how not having implementing any policies would have
-%changed the national average growth rate of cases  over
-%time. The effect of removing all policies on the growth rate of cases
-% peaks in April at 0.1,  and then
-%plateaus. The estimate implies a 0 to 10 fold increase in cases by the end of May.
-%
-%Even this wide confidence interval arguably understates uncertainty regarding the effect of all policies.
-%There is little variation across states in the timing of school closures.  This makes the effect of school closures difficult to identify.
-%To illustrate this sensitivity of conclusions regarding school closures, in Figure \ref{fig:US-nop-SI} we show that with a less flexible parameterization of time effects (which allow for greater attribution of effects to schooling), the number of cases by the end of May could have been even larger (the confidence band is wide, ranging from at least 7 fold increase to a very large upper bound).
-This highlights the uncertainty regarding the impact of all policies versus private behavioral responses to information.
-%Evaluation of re-opening policies needs to be aware of this uncertainty.
-
-%, where about 80 \% of states closed schools within a day or two of March 15th.
-
-\begin{figure}[ht]
-  \caption{Effect  of nationally mandating masks for employees on April
-    1st in the US \label{fig:US-mask-dgrowth}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \medskip
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-%    \begin{flushleft}
-%      \footnotesize
-%      In the left column, the dots are the average change in growth in
-%      each state. The blue line is the average across states of the
-%      change in growth. The shaded region is a point-wise 90\% confidence
-%      interval. The right column shows the change in cases or deaths
-%      relative to the baseline of actual policies.
-%    \end{flushleft}
-\end{figure}
-
-\begin{figure}[ht]
- % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of   leaving non-essential businesses open on cases in the US  \label{fig:US-nb-dgrowth}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
- % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-%\begin{figure}[ht]
-% \caption{Average effect  of removing
-%   all policies on cases \label{fig:US-nop-dgrowth}}
-%%  \caption{Effect of removing policies: national average \label{fig:US-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
-%%      \\
-%%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
-%%      &
-%%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-
-%\subsection{Related literature}
-%
-%
-%... and why/how we are different.
-
-A growing number of other papers have examined the link between non-pharmaceutical interventions and Covid-19  cases.\footnote{We refer the reader to \cite{avery2020policy} for a comprehensive review of  a larger body of work researching Covid-19; here we focus on few quintessential comparisons on our work with other works that we are aware of.} \cite{hsiang2020} estimate the effect of policies on the growth rate of cases using
-data from the United States, China, Iran, Italy, France, and South
-Korea. In the United States, they find that the combined effect of all policies they consider on the
-growth rate is $-0.347$ $(0.061)$. \cite{courtemanche2020} use US county level data to analyze the effect
-of interventions on case growth rates. They find that the combination
-of policies they study reduced growth rates by 9.1 percentage points
-16-20 days after implementation, out of which 5.9 percentage points are attributed to shelter in place orders.
-Both \cite{hsiang2020} and \cite{courtemanche2020} adopted a reduced-form approach to estimate the total policy effect on case growth without using any social distancing behavior measures.
-In contrast, our study highlights the role of behavioral response to policies and information.
-
-Existing evidence for the impact of social distancing policies on behavior in the US is mixed.
-\cite{abouk2020} employ a difference-in-differences methodology to  find that statewide stay-at-home orders  have  strong causal impacts on reducing social interactions.
-In contrast, using data from Google Mobility Reports,
-\cite{maloney2020}  find that the increase in social distancing is
-largely voluntary and driven by information.\footnote{Specifically, they find
-that of the 60 percentage point drop in workplace intensity, 40
-percentage points can be explained by changes in information as
-proxied by case numbers, while roughly 8 percentage points can be
-explained by policy changes.} Another study by  \cite{gupta2020} also found little evidence that stay-at-home mandates induced distancing by
-using mobility measures from PlaceIQ and SafeGraph.
-Using data from SafeGraph, \cite{anderson2020} show that there has been substantial voluntary social distancing but also provide evidence that mandatory measures such as stay-at-home orders have  been effective at reducing the frequency of visits  outside of one’s home.
-% Overall, these evidence suggests the importance of voluntary social distancing behavior when we evaluate the effect of polices.
-
- \cite{pei2020} use county-level observations of reported infections and deaths in conjunction with mobility  data from SafeGraph to estimate how effective reproductive numbers in major metropolitan areas change over time.
-They conduct simulation of implementing all policies 1-2 weeks earlier and found that it would have resulted in reducing the number of cases and deaths by more than half.
-However, their study does not  explicitly analyze how policies are related to
-the effective reproduction numbers. % Our study  analyzes how policies  affects the reproduction numbers through their impact on people's behavior.
-
-Epidemiologists use model simulations to predict how cases and deaths evolve for the purpose of policy recommendation.
-As reviewed by \cite{avery2020policy}, there exist substantial uncertainty about the values of key epidimiological parameters \citep[see also][]{atkeson2020b,stock2020}.
-Simulations are often done under strong assumptions about the impact of social distancing policies without connecting to the relevant data  \citep[e.g.,][]{ferguson2020}.
-Furthermore, simulated models do not take into account that people may limit their contact with other people in response to higher transmission risks.\footnote{See \cite{atkeson2020a} and \cite{stock2020} for the implications of the SIR model for Covid-19 in the US.
-\cite{NBERw27128} estimate a SIRD model in which time-varying reproduction numbers depend on the daily deaths to capture feedback from daily deaths
-to future behavior and infections.}  When such a voluntary behavioral response is ignored, simulations would produce exponential spread of disease and would over-predict  cases and deaths.
-Our counterfactual experiments  illustrate the importance of this voluntary behavioral change.
-
-Whether wearing  masks in public place should be mandatory or not has been one of the most contested policy issues with health authorities of different countries providing contradiction recommendations.
-Reviewing evidence, \cite{Greenhalghm2020}  recognize that there is no randomized controlled trial evidence for the effectiveness of face masks,  but they state ``indirect evidence exists to support the argument for the public wearing masks in the Covid-19 pandemic."\footnote{The virus remains viable in the air for several hours, for which surgical masks may be effective.
-Also, a substantial fraction of individual who are infected become infectious before showing symptom onset.}
-\cite{howard2020} also review available medical evidence and conclude that ``mask wearing reduces the transmissibility per contact by reducing transmission of infected droplets in both laboratory and clinical contexts.''  The laboratory findings in \cite{hou2020} suggest that the nasal cavity may be the initial site of infection followed by aspiration to the lung, supporting the argument  ``for the widespread use of masks to prevent aersol, large droplet, and/or mechanical exposure to the nasal passages.''
-
-Given the lack of experimental evidence on the effect of masks, conducting observational studies  is useful and important.
-To the best of our knowledge, our paper is the first  empirical study that shows the effectiveness of  mask mandates on reducing the spread of Covid-19 by analyzing the US state-level data.
-This finding corroborates and is complementary to the medical observational evidence in \cite{howard2020}.
-Analyzing mitigation measures in New York, Wuhan, and Italy, \cite{zhangr2020} conclude that mandatory face coverings substantially reduced infections.
-\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.
-Our finding
-is also corroborated by a completely different causal methodology based on synthetic control 
-using German data in \cite{Mitze2020}.\footnote{Our study was first released in ArXiv on May 28, 2020 whereas
- \cite{Mitze2020} was released at SSRN on June 8, 2020.}
-
-Our empirical results contribute to informing the economic-epidemiological models that combine economic models with variants of SIR models to evaluate the efficiency of various economic policies aimed at gradual ``reopening" of various sectors of economy.\footnote{\cite{adda2016} analyzes the effect of policies reducing interpersonal contacts such as school closures or the closure of public transportation networks on the spread of  influenza, gastroenteritis, and chickenpox using high frequency data from France.} For example, the estimated effects of masks, stay-home mandates, and various other policies on behavior, and of behavior on infection can serve as useful inputs and validation checks in the calibrated macro, sectoral, and micro models (see, e.g., \cite{alvarez2020simple,baqaee2020reopening,NBERw27128,acemoglu2020multi,lsmith,nashSIR} and references therein).
-Furthermore, the causal framework developed in this paper could be applicable, with appropriate extensions, to the impact of policies on economic outcomes replacing health outcomes (see, e.g., \cite{chetty2020real,coibion2020labor}).
-
-\section{The Causal Model for the Effect of Policies, Behavior, and Information on Growth of Infection}\label{sec:causal-model}
-
-\subsection{The Causal Model and Its Structural Equation Form}
-We introduce our approach through the Wright-style causal diagram shown in Figure \ref{Wright}.  The diagram describes how policies, behavior, and information interact together:
-\begin{itemize}
-\item The \textit{forward} health outcome,
-$Y_{i,t+\ell}$, is determined last, after all other variables have been determined;
-\item The  adopted policies, $P_{it}$,  affect health outcome $Y_{i,t+\ell}$ either directly, or indirectly by altering  human behavior  $B_{it}$;
-\item  Information variables, $I_{it}$, such as lagged values of outcomes can affect human behavior and  policies, as well as  outcomes;
-\item The confounding factors $W_{it}$, which vary across states and time, affect all other variables.
-\end{itemize}
-The index $i$ denotes observational unit, the state, and $t$ and $t+\ell$ denotes the time, where  $\ell$ represents the time lag  between infection and case confirmation or death.
-\begin{figure}[ht]
-\begin{center}
-\begin {tikzpicture}[-latex, auto, node distance =1.5cm and 3cm, on grid, thick,
-  empty/.style ={circle, top color=white, bottom color = white, draw, white, text=white , minimum width =1.25 cm},
-  policy/.style ={circle, top color=white, bottom color = blue, draw, black, text=black , minimum width =1.25 cm},
-    behavior/.style ={circle, top color=white, bottom color = ForestGreen, draw, black, text=black , minimum width =1.25 cm},
-  observed/.style ={circle, top color=white, bottom color = magenta, draw, black, text=black , minimum width =1.25 cm} ,
-   confounder/.style ={circle, top color=white, bottom color =  gray, draw, black, text=black , minimum width =1.25 cm} ,
-outcome/.style ={circle, top color=white, bottom color = red, draw, black, text=black , minimum width =1.25 cm} ]
-
-\node[policy]   (P) {\tiny $P_{it}$};
-\node[empty] (E) [below=of P] {\tiny $I_{it}$};
-\node[outcome]  (Y)[ right =of E] {\tiny $Y_{i,t+\ell}$};
-\node[behavior] (B) [below =of E] {\tiny $B_{it}$};
-\node[observed] (I) [left =of P] {\tiny $I_{it}$};
-\node[confounder] (W) [left=of B] {\tiny  $W_{it}$};
-\path[->] (P) edge (Y);
-\path[->] (B) edge (Y);
-\path[->] (P) edge (B);
-\path[->] (I) edge (B);
-\path[->] (I) edge (P);
-\path[->] (W) edge (Y);
-\path[->] (W) edge (B);
-\path[->] (W) edge (P);
-\path[->] (W) edge (I);
-\path[->] (I) edge (Y);
-
-\end{tikzpicture} \end{center}
-\caption{S. \& P. Wright type causal path diagram for our model. }\label{Wright}
-\end{figure}
-
-Our main outcomes of interest are the growth rates in Covid-19 cases and deaths,  behavioral variables include proportion of time spent in transit or shopping and others, policy variables include stay-at-home orders and school and business closures, and the information variables include lagged values of outcome.
-We provide a detailed description of these variables and their timing in the next section.
-
-The causal structure  allows for  the effect of the policy to be either direct or indirect -- through-behavior or through dynamics; and all of these effects are not mutually exclusive.
-The structure also allows for changes in behavior to be brought by change in policies and information.
-These are all realistic properties that we expect from the contextual knowledge of the problem.
-Policies such as closures of schools, non-essential business, and restaurants, alter and constrain behavior in strong ways.
-In contrast, policies such as mandating employees to wear masks can potentially affect the Covid-19 transmission directly.
-The information variables, such as recent growth in the number of cases, can cause people to spend more time at home, regardless of adopted state policies; these changes in
-behavior in turn affect the transmission of Covid-19.
-
-The causal ordering induced by this directed acyclical graph is determined by the following
-timing sequence: % \textit{within a period}:%\footnote{Mathematically this corresponds to the factorization of the data likelihood, omitting indexes, $$f_{Y,P,B, I, W} = f_{Y\mid P, B,I, W} f_{B \mid P,I,W } f_{P \mid I, W} f_{I,W}.$$ }
-\begin{itemize}
-\item[(1)]  information and confounders get determined at $t$,
-\item[(2)] policies are set in place, given information and confounders at $t$;
-\item[(3)] behavior is realized, given policies, information, and confounders at $t$;
-\item[(4)] outcomes get realized at $t+\ell$ given policies, behavior, information, and confounders.
-\end{itemize}
-
-The model also allows for direct dynamic effects of information variables on the outcome through autoregressive structures that capture persistence in growth patterns.
-As  highlighted below, realized outcomes may become new information for future periods, inducing dynamics over multiple periods.
-
-Our quantitative model for causal structure in Figure \ref{Wright} is given by the following econometric structural equation model:\begin{equation} \label{eqn:SEM1} \tag{SEM}
-  \begin{aligned}
-   &  Y_{i,t+\ell} (b,p,\iota) &:=& {\bcolor \alpha ' b}  +  {\pcolor \pi 'p} +
-    {\icolor \mu'\iota } + {\wcolor \delta_Y 'W_{it}} + \varepsilon^y_{it}, \\
-   &  B_{it} (p,\iota) &:= & {\pcolor \beta'p } + {\icolor \gamma'\iota} +      {\wcolor\delta_B 'W_{it} } + \varepsilon^b_{it},
-      \end{aligned}
- \end{equation}
-which is a collection of functional relations with stochastic shocks, decomposed into observable part $\delta' W$ and unobservable part $\varepsilon$.
-The terms $\varepsilon^y_{it}$ and  $\varepsilon^b_{it} $  are the centered stochastic shocks that obey the orthogonality restrictions posed below.
-
-The policies can be modeled via a linear form as well,
-\begin{equation}\label{eq:Policy} \tag{P}
- P_{it}   (\iota) :=  {\icolor\eta'\iota} + {\wcolor \delta_P' W_{it}} +   \varepsilon^p_{it},   \end{equation}
-although  linearity is not critical.\footnote{Under some additional independence conditions, this
-can be replaced by an arbitrary non-additive function $P_{it}(\iota) = p (\iota, W_{it},  \varepsilon^p_{it})$, such that the unconfoundedness condition stated in the next footnote holds.}
-
-The orthogonality restrictions on the stochastic components are as follows:
-\begin{equation}\label{eqn:SEM2} \tag{O}
-\begin{aligned}
-   & \varepsilon^y_{it} &  \perp &  \quad (\varepsilon^b_{it}, \varepsilon^p_{it}, {\wcolor W_{it}}, {\icolor I_{it}}), \\
-&  \varepsilon^b_{it}  & \perp & \quad  (\varepsilon^p_{it}, {\wcolor W_{it}}, {\icolor I_{it}}), \\
-&   \varepsilon^p_{it} &  \perp &  \quad ({\wcolor W_{it}}, {\icolor I_{it}}),
-\end{aligned}
-\end{equation}
-where we say that $V \perp U$ if $\Ep VU = 0$. This is a standard way of representing restrictions on errors in structural equation modeling.\footnote{ An alternative useful
-starting point is to impose the Rubin-Rosenbaum type unconfoudedness condition:
-$$
-Y_{i,t+\ell} (\cdot,\cdot,\cdot)  \indep (P_{it}, B_{it}, I_{it})  \mid W_{it}, \
-B_{it} (\cdot,\cdot)  \indep (P_{it}, I_{it})  \mid W_{it}, \
- P_{it} (\cdot)  \indep  I_{it}  \mid W_{it},
-$$
-which imply, with treating stochastic errors as independent additive components, the orthogonal conditions stated above.
- }\footnote{The structural equations of this form are connected to triangular structural equation models, appearing
- in microeconometrics and macroeconometrics (SVARs), going back to the work of  \cite{strotz1960recursive}. .}
-
- The observed variables are generated by setting $\iota = I_{it}$ and propagating
-the system from the last equation to the first:
-%\begin{equation} \label{eqn:SEM2}
-\[
-  \begin{aligned}
-& {\ycolor Y_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B_{it} } ,{\pcolor P_{it}}, {\icolor I_{it}}), \\
-& {\bcolor B_{it} } & := &   B_{it}({\pcolor P_{it} } ,{\icolor I_{it}}), \\
-& {\pcolor P_{it} }& := &  P_{it}({\icolor I_{it}}). \end{aligned}
-\]
-%\end{equation}
-
-The system above together with  orthogonality restrictions (\ref{eqn:SEM2}) implies the following collection of stochastic equations for realized variables:
-\begin{align}
-   &  {\ycolor  Y_{i,t+\ell}}
-    = {\bcolor\alpha ' B_{it}} + {\pcolor\pi 'P_{it}} + {\icolor\mu'I_{it}} + {\wcolor\delta_Y 'W_{it}}  + \varepsilon^y_{it},
-    &  & \varepsilon^y_{it} \perp {\bcolor B_{it}}, {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R1} \tag{BPI$\to$Y} \\
-    &  {\bcolor B_{it}}
-     =  {\pcolor \beta' P_{it}} + {\icolor \gamma'I_{it}} +  {\wcolor \delta_B' W_{it}} + \varepsilon^b_{it},
-   & & \varepsilon^b_{it} \perp {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}}  \label{eq:R2} \tag{PI$\to$B}  \\
-    & {\pcolor P_{it}}
-    =  {\icolor\eta'I_{it}} + {\wcolor \delta_P' W_{it}} +   \varepsilon^p_{it},   & & \varepsilon^p_{it} \perp   {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R3}  \tag{I$\to$P}
-       \end{align}
-and
-\begin{equation}
-   {\ycolor  Y_{i,t+\ell}}
-   = ( {\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'} )
-    {\pcolor P_{it}} + ( {\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})
-    {\icolor I_{it} }+ {\wcolor \bar{\delta} '}{\wcolor W_{it}}  + {\bar \varepsilon}_{it},  \quad   {\bar \varepsilon}_{it} \perp
-  {\pcolor P_{it}},  {\icolor I_{it}}, {\wcolor W_{it}}.  \label{eq:R4} \tag{PI$\to$Y}
-\end{equation}
-These equations form the basis of our empirical analysis.
-
-As discussed below, the information variable includes case growth.
-Therefore,  an orthogonality restriction  $ \varepsilon^y_{it} \perp   {\pcolor P_{it}}$ holds if  the government does not have knowledge on future case growth beyond what is predicted by today's case growth, policies,  behavior, and  confounders; even when the government has some knowledge on $\varepsilon^y_{it}$, the orthogonality restriction may hold if there is a time lag for the government to implement its policies based on $\varepsilon^y_{it}$.
-
-The orthogonality condition in (\ref{eq:R4}) is weaker than the orthogonality conditions in  (\ref{eq:R1})-(\ref{eq:R2}) in that the former is implied by the latter but not vice versa.
-The system over-identifies the regression coefficients because  $({\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'})$ and $({\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})$ in  (\ref{eq:R4}) can be also identified from  ${\bcolor\alpha ' }$,  $ {\pcolor\pi '}$, ${\icolor\mu'}$,   ${\pcolor \beta' }$, and ${\icolor \gamma'}$ in (\ref{eq:R1})-(\ref{eq:R2}).
-Comparing the estimates of $({\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'})$ and $({\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})$ from  (\ref{eq:R4})  with  those implied by  the estimates of ${\bcolor\alpha ' }$,  $ {\pcolor\pi '}$, ${\icolor\mu'}$,   ${\pcolor \beta' }$, and ${\icolor \gamma'}$ from   (\ref{eq:R1})-(\ref{eq:R2}) provides a useful specification test.
-
-\subsection*{Identification and Parameter Estimation}
-
-The orthogonality equations imply that these are all projection equations, and the parameters of the SEM are identified by the parameters of these regression equation, provided the latter are identified by sufficient joint variation of these variables across states and time.
-
-The last point can be stated formally as follows. Consider the previous system of equations, after partialling out the confounders:
-\begin{equation}\label{eqn:SEM-PO}
- \begin{aligned}
-   & {\ycolor { \tilde Y_{i,t+\ell}}}  & =& {\bcolor {\alpha ' \tilde B_{it}}} +{\pcolor {\pi '\tilde P_{it}}} +{\icolor \mu'\tilde I_{it} } + \varepsilon^y_{it},&\varepsilon^y_{it} &\perp {\tilde B_{it}}, {\tilde  P_{it}}, { \tilde I_{it}},  \\
-   & {\bcolor { \tilde B_{it}}} & = &{\pcolor { \beta' \tilde P_{it}} } + {\icolor { \gamma' \tilde I_{it}} } + \varepsilon^b_{it},
-   &\quad \varepsilon^b_{it} &\perp { \tilde P_{it}}, { \tilde I_{it}},  \\
-   & {\pcolor { \tilde P_{it}} } &= &  {\icolor {\eta' \tilde I_{it}} } +   \varepsilon^p_{it}, &\quad \varepsilon^p_{it} &\perp {\tilde I_{it}}  \\
-  \end{aligned}
-\end{equation}
-where $ \tilde V_{it} = V_{it}   -     {\wcolor W_{it}'} \Ep[{\wcolor W_{it}W_{it}'}]^{-} \Ep[{\wcolor W_{it}} V_{it}]$ denotes
-the residual after removing the orthogonal projection of $V_{it}$ on ${\wcolor W_{it}}$.
-The residualization is a linear operator, implying that (\ref{eqn:SEM-PO}) follows immediately from the above.
-The parameters of (\ref{eqn:SEM-PO})  are identified as projection coefficients in these equations, provided that residualized vectors appearing in each of the equations have non-singular variance, that is
- \begin{equation}
- \Var ( {\pcolor \tilde P_{it}'} ,{\bcolor \tilde B_{it}'},{\icolor \tilde I_{it}'})>0,
- \ \Var ({\pcolor \tilde P_{it}'}, {\icolor \tilde I_{it}'})> 0 , \ \text{ and  }  \Var ( {\icolor \tilde I_{it}'}) >0.
- \end{equation}
-
-Our main estimation method is the standard correlated random effects estimator, where the random effects
-are parameterized as functions of observable characteristic, ${\wcolor W_{it}}$, which include both state-level and time random effects.  The state-level random effects
-are modeled as a function of state level characteristics, and the time random effects are modeled
-by including month dummies and their interactions with state level characteristics.  The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$
-are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.
-
-A secondary estimation method is the fixed effects estimator, where ${\wcolor W_{it}}$ includes
-latent (unobserved) state level effects ${\wcolor W_i}$ and  and time level effects ${\wcolor W_t}$,
-which must be estimated from the data.  This approach is much more demanding of the data and relies on long cross-sectional
-and time histories. When histories are relatively short,  large biases emerge and
-they need to be removed using debiasing methods. In our context, debiasing
-  changes the estimates substantially, often changing the sign of coefficients.\footnote{This is a pre-cautionary
-message that may be useful for other researchers using fixed effects estimators in the context
-of Covid-19 analysis. We recommend using debiased fixed effects estimators, see e.g., \cite{chen2019mastering}
-for expository treatment. }
-However, we find the debiased fixed effect estimates are
-qualitatively and quantitatively similar to the correlated random effects estimates.
-Given this finding, we chose to focus on the latter, as it is a more standard and familiar method, and report the former
-estimates in the supplementary materials for this paper.\footnote{The similarity of the debiased fixed
-effects and correlated random effects served as a useful specification check. Moreover, using the fixed
-effects estimators only yielded minor gains in predictive performances, as judging by the adjusted
-$R^2$'s, providing another useful specification check.}
-
-\subsection{Information Structures and Induced Dynamics}  We consider three examples of information
-structures: Information variable is  a function of time:
-$${\icolor I_{it}} = g(t);$$
- Information variable is lagged value of outcome:
-$$
-{\icolor I_{it}} =  {\ycolor Y_{it}};
- $$
- and finally:
- \begin{itemize}
- \item[(\textsc{I})] Information variables include time, lagged and integrated values of outcome:
- $$
-{\icolor I_{it}} = \left (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{t/\ell}
-{\ycolor Y_{i,t - \ell m}} \right) ',
- $$
- with the convention that ${\ycolor Y_{it}} = 0$  for $t \leq 0$.
-\end{itemize}
-The first information structure captures the  basic idea that, as individuals discover more information about covid over time, they adapt to safer modes of  behavior (stay-at-home, wear masks, wash hands).
-{Under this structure, information is common across states and exogenously evolves over time, independent of the number of cases.}
-The second  structure  arises from considering
-autoregressive components {and captures people's behavioral response to information on cases in the state they reside}.
-{Specifically,} we model persistence in growth rates, ${\ycolor Y_{i,t+\ell}}$, through an AR(1) model, which leads to ${\icolor I_{it}}= {\ycolor Y_{it}}$.
-This provides useful \textit{local}, state-specific, information about the forward growth rate and people may adjust their behavior to safer modes when they see a high value.
-We model this adjustment via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.
-The {third information} structure
-is the combination of the first two structures plus an additional  term representing the log of the total number of new cases in the state.
-We use this information structure in our empirical specification.
-In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases.
-The last element of the information set can be thought of as a local stochastic trend in cases.
-
-All of these examples fold into a specification of the form:
-\begin{equation}\label{eq:I}\tag{I}
- {\icolor I_{it} } :=   I_{it}({\icolor I_{i,t-\ell}}, {\ycolor Y_{it}}, t),  \quad t =1,..., T,
- \end{equation}
- with the initialization ${\icolor I_{i 0}} = 0$ and ${\ycolor Y_{i0}} = 0$.\footnote{This initialization is appropriate
- in our context for analyzing pandemics from the very beginning, but other initializations could be appropriate in other contexts.
-The  lagged values of behavior variable may be also included in the information set.
- }
-
-With any structure of this form, realized outcomes may become new information for future periods, inducing a dynamical system over multiple periods.
-We show the resulting dynamical system in a causal diagram of Figure \ref{fig:DynS}.
-Specification
-of this system is useful for studying delayed effects of policies and behaviors and in considering the counterfactual policy analysis.
-
-\begin{figure}[ht]
-\begin{center}
-\begin {tikzpicture}[-latex, auto, node distance =3cm and 3cm, on grid, thick,
-  empty/.style ={circle, top color=white, bottom color = white, draw, white, text=white , minimum width =1.25 cm},
-  policy/.style ={circle, top color=white, bottom color = blue, draw, black, text=black , minimum width =1.25 cm},
-    behavior/.style ={circle, top color=white, bottom color = ForestGreen, draw, black, text=black , minimum width =1.25 cm},
-  observed/.style ={circle, top color=white, bottom color = magenta, draw, black, text=black , minimum width =1.25 cm} ,
-   confounder/.style ={circle, top color=white, bottom color =  gray, draw, black, text=black , minimum width =1.25 cm} ,
-outcome/.style ={circle, top color=white, bottom color = red, draw, black, text=black , minimum width =1.25 cm},
-myarrow/.style={-Stealth}]
-
-\node[observed] (I)  {\tiny $I_{i,t-\ell}$};
-\node[outcome]  (Y)[below=of I ]{\tiny $Y_{it}$};
-\node[observed] (In)  [right=of I ]{\tiny $I_{it}$};
-\node[outcome]  (Yn)[below=of In ]{\tiny $Y_{i,t+\ell}$};
-\node[observed] (Inn)  [right=of In ]{\tiny $I_{i,t+\ell}$};
-\node[outcome]  (Ynn)[below=of Inn ]{\tiny $Y_{i,t+2\ell}$};
- \draw [->] (I) -- node[sloped,font=\small,below] {\tiny SEM(t-$\ell$)} (Y);
- \draw [->] (I) --  (In);
- \draw [->] (In) -- node[sloped,font=\small,below] {\tiny SEM(t)} (Yn);
-\draw [->] (Y) --  (In);
-\draw [->] (In) --  (Inn);
- \draw [->] (Inn) -- node[sloped,font=\small,below] {\tiny SEM(t+$\ell$) } (Ynn);
-\draw [->] (Yn) --  (Inn);
-
-\end{tikzpicture} \end{center}
-\caption{ Dynamic System Induced by Information Structure and SEM}\label{fig:DynS}
-\end{figure}
-
-%
-%\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel}
-%Letting $C_{it}$ denotes cumulative number of  confirmed cases in state $i$ at time $t$, our outcome
-%\begin{equation} \label{eq:y}
-%{\ycolor Y_{it}} :=\Delta \log(\Delta C_{it}):= \log( \Delta C_{it} ) -
-%\log( \Delta C_{i,t-7})
-%\end{equation}
-%approximates the weekly growth rate in new cases from $t-7$ to $t$.\footnote{We may show that $ \log( \Delta C_{it} ) -
-%\log( \Delta C_{i,t-7})$ approximates the average
-%growth rate of cases from $t-7$ to $t$.} Here $\Delta$ denotes the differencing operator over 7 days from $t$ to $t-7$, so that $\Delta C_{it}:=C_{it}-C_{i,t-7}$ is the number of new confirmed cases in the past 7 days.
-%
-%
-%We chose this metric as this is the key metric for policy makers deciding when to relax Covid mitigation policies.  The U.S. government's guidelines for state reopening
-%recommend that states display a
-%``downward trajectory of documented cases within a 14-day period''
-%\citep{whitehouse2020}. A negative value of
-%$Y_{it}$ is an indication of meeting this
-%criteria for reopening. By focusing on weekly cases  rather than daily cases, we smooth idiosyncratic daily fluctuations as well as periodic fluctuations associated with days of the week.
-%
-%Our measurement equation for estimating equations (\ref{eq:R1}) and (\ref{eq:R4}) will take the form:
-%\begin{align}
-%{\ycolor Y_{it}}  =      \theta' X_{it}  + \delta_T \Delta   \log(T)_{it}  +
-% \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  + \epsilon_{it},
-% \label{eq:M} \tag{M}
-%\end{align}
-%where $i$ is state, $t$ is day, $C_{it}$ is cumulative confirmed
-%cases, $D_{it}$ is deaths, $T_{it}$ is the number of tests over 7 days, $\Delta$ is
-%a 7-days differencing operator, $\epsilon_{it}$ is an unobserved error term,
-%where $X_{it}$ collects other behavioral, policy, and confounding variables, depending
-%on whether we estimate (\ref{eq:R1}) or (\ref{eq:R4}).  Here $$\Delta   \log(T)_{it} \ \text {and } \ \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}$$ are the key confounding variables,
-%derived from considering the SIR model below. We describe other confounders in the empirical section.
-%
-% Our main estimating equation (M) is motivated by a variant of a SIR
-%model, where we incorporate a
-%delay between infection and death instead of a constant death rate,
-%and we add confirmed cases to the model. Let $S$, $I$, $\mathcal{R}$, and $D$ denote the number of susceptible,
-%infected, recovered, and dead individuals in a given state or
-%province. Each of these variables are a function of time. We model
-%them as evolving as
-%\begin{align}
-%  \dot{S}(t) & = -\frac{S(t)}{N} \beta(t) I(t) \label{eq:s} \\
-%  \dot{I}(t) & = \frac{S(t)}{N} \beta(t) I(t) - \frac{S(t-\ell)}{N}
-%               \beta(t-\ell) I(t-\ell) \label{eq:i} \\
-%  \dot{\mathcal{R}}(t) & = (1-\pi) \frac{S(t-\ell)}{N} \beta(t-\ell)
-%                         I(t-\ell) \label{eq:r} \\
-%  \dot{D}(t) & = \pi \frac{S(t-\ell)}{N} \beta(t-\ell) I(t-\ell) \label{eq:d}
-%\end{align}
-%where $N$ is the population, $\beta(t)$ is the rate of infection
-%spread, $\ell$ is the duration between infection and
-%death, and $\pi$ is the
-%probability of death conditional on infection.\footnote{The model can easily be extended to allow a
-%  non-deterministic disease duration. This leaves our main estimating
-%  equation (\ref{eq:c2}) unchanged, as long as the distribution of
-%  duration between infection and death is equal to the distribution of
-%  duration between infection and recovery.}
-%
-%Confirmed cases, $C(t)$, evolve as
-%\begin{equation}
-%  \dot{C}(t) = \tau(t) I(t), \label{eq:c}
-%\end{equation}
-%where $\tau(t)$ is the rate that infections are detected.
-%
-%Our goal is to examine how the rate of infection $\beta(t)$ varies with observed policies
-%and measures of social distancing behavior. A key challenge is that we only
-%observed $C(t)$ and $D(t)$, but not $I(t)$. The unobserved $I(t)$ can
-%be eliminated by differentiating (\ref{eq:c}) and using (\ref{eq:i}) and (\ref{eq:d}) as
-%\begin{align}
-%%  \ddot{C}(t) & = \dot{\tau}(t) + \dot{I}(t) \notag \\
-%%  & = \dot{\tau}(t)I(t) + \frac{S(t)}{N} \beta(t) I(t) -
-%%    \frac{S(t-\ell)}{N} \beta(t-\ell) I(t-\ell)  \notag \\
-%%  & = \dot{\tau}(t) I(t) + \frac{S(t)}{N} \beta(t) I(t) -
-%%    \frac{1}{\pi} \dot{D}(t) \notag \\
-%  \frac{\ddot{C}(t)}{\dot{C}(t)}
-%              & =
-%                \frac{S(t)}{N} \beta(t)  + \frac{\dot{\tau}(t)}{\tau(t)}  -
-%                \frac{1}{\pi}
-%                \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}. \label{eq:c2}
-%\end{align}
-%We consider a discrete-time analogue of equation (\ref{eq:c2}) to motivate our empirical
-%specification by relating the detection rate $\tau(t)$  to the number of tests $T_{it}$ while specifying $\frac{S(t)}{N}\beta(t)$ as a linear function of variables $X_{it}$.
-%This results in
-%\begin{align}
-%  \underbracket{Y_{it}}_{\frac{\ddot{C}(t)}{\dot{C}(t)}}
-%  =
-%      \underbracket{X_{it}' \theta + \epsilon_{it}}_{\frac{S(t)}{N}\beta(t)} +
-%       & \underbracket{\delta_T \Delta
-%      \log(T)_{it}}_{\frac{\dot{\tau}(t)}{\tau(t)} } +
-%      \underbracket{\delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  }_{\frac{1}{\pi}
-%      \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}}, \nonumber % \label{eq:reg2}
-%\end{align}
-%which is equation (\ref{eq:M}), where $X_{it}$ captures a vector of variables related to $\beta(t)$.
-%
-%\begin{quote}
-%\textsc{Structural Interpretation}. In the early pandemics, when
-%the number of susceptibles is approximately the same as the entire population, i.e. $S_{it}/N_{it} \approx 1$, the component $X_{it}' \theta$
-%is the projection of infection rate $ \beta_i(t)$ on   $X_{it}$ (policy, behavioral, information, and confounders other than testing rate), provided
-%the stochastic component $\epsilon_{it}$ is orthogonal
-%to $X_{it}$ and the testing variables:
-%$$
-%\beta_i(t)S_{it}/N_{it}  = X_{it}' \theta + \epsilon_{it}, \quad \epsilon_{it} \perp X_{it}.
-%$$
-%\end{quote}
-
-\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel}
-Letting $C_{it}$ denote the cumulative number of  confirmed cases in state $i$ at time $t$, our outcome
-\begin{equation} \label{eq:y}
-{\ycolor Y_{it}} =\Delta \log(\Delta C_{it}):= \log( \Delta C_{it} ) -
-\log( \Delta C_{i,t-7})
-\end{equation}
-approximates the weekly growth rate in new cases from $t-7$ to $t$.\footnote{We may show that $ \log( \Delta C_{it} ) -
-\log( \Delta C_{i,t-7})$ approximates the average
-growth rate of cases from $t-7$ to $t$.} Here $\Delta$ denotes the differencing operator over 7 days from $t$ to $t-7$, so that $\Delta C_{it}:=C_{it}-C_{i,t-7}$ is the number of new confirmed cases in the past 7 days.
-
-We chose this metric as this is the key metric for policy makers deciding when to relax Covid mitigation policies.  The U.S. government's guidelines for state reopening
-recommend that states display a
-``downward trajectory of documented cases within a 14-day period''
-\citep{whitehouse2020}. A negative value of
-$Y_{it}$ is an indication of meeting this
-criteria for reopening. By focusing on weekly cases  rather than daily cases, we smooth idiosyncratic daily fluctuations as well as periodic fluctuations associated with days of the week.
-
-Our measurement equation for estimating equations (\ref{eq:R1}) and (\ref{eq:R4}) will take the form:
-\begin{align}
-{\ycolor \Delta \log(\Delta C_{it})}  =    X_{i,t-14} '   \theta  -\gamma +  \delta_T \Delta   \log(T_{it})  + \epsilon_{it},
-% \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  + \epsilon_{it},
- \label{eq:M} \tag{M-C}
-\end{align}
-where $i$ is state, $t$ is day, $C_{it}$ is cumulative confirmed
-cases, $T_{it}$ is the number of tests over 7 days, $\Delta$ is
-a 7-days differencing operator, $\epsilon_{it}$ is an unobserved error term.
- $X_{i,t-14}$  collects other behavioral, policy, and confounding variables, depending
-on whether we estimate (\ref{eq:R1}) or (\ref{eq:R4}), where the lag of $14$ days captures the time lag between infection and confirmed case (see the Appendix \ref{sec:timing}).
-%for available evidence on a delay between infection and confirmation of cases or deaths.
-   Here
-$$\Delta   \log(T_{it} ):=  \log(T_{it}) - \log(T_{i,t-7})  $$ %\ \text {and } \ \delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}$$
-is the key confounding variable,
-derived from considering the SIR model below. We describe other confounders in the empirical section.
-
-% Our main estimating equation (M) is motivated by a variant of a SIR
-%model, where we incorporate a
-%delay between infection and death instead of a constant death rate,
-%and we add confirmed cases to the model.
-
- Our main estimating equation (\ref{eq:M}) is motivated by a variant of SIR
-model, where we add confirmed cases and infection detection via testing.
-Let $S$, $\Infected$, $\Recovered$, and $D$ denote the number of susceptible,
-infected, recovered, and dead individuals in a given state. Each of these variables are a function of time. We model
-them as evolving as
-\begin{align}
-  \dot{S}(t) & = -\frac{S(t)}{N} \beta(t) \Infected(t) \label{eq:s} \\
-  \dot{\Infected}(t) & = \frac{S(t)}{N} \beta(t) \Infected(t) - \gamma  \Infected(t) \label{eq:i}\\
-  % \frac{S(t-\ell)}{N} \beta(t-\ell)  \Infected(t-\ell) \label{eq:i} \\
-  \dot{\Recovered}(t) & = (1-\kappa) \gamma  \Infected(t) \label{eq:r}\\ %  \frac{S(t-\ell)}{N} \beta(t-\ell)    \Infected(t-\ell) \label{eq:r} \\
-  \dot{D}(t) & = \kappa \gamma \Infected(t) %  \frac{S(t-\ell)}{N} \beta(t-\ell)   \Infected(t-\ell)
-  \label{eq:d}
-\end{align}
-where $N$ is the population, $\beta(t)$ is the rate of infection
-spread, $\gamma$ is the rate of recovery or death, and $\kappa$ is the
-probability of death conditional on infection.
-%\footnote{The model can easily be extended to allow a
-%  non-deterministic disease duration. This leaves our main estimating
-%  equation (\ref{eq:c2}) unchanged, as long as the distribution of
-%  duration between infection and death is equal to the distribution of
-%  duration between infection and recovery.}
-
-Confirmed cases, $C(t)$, evolve as
-\begin{equation}
-  \dot{C}(t) = \tau(t) \Infected(t), \label{eq:c}
-\end{equation}
-where $\tau(t)$ is the rate that infections are detected.
-
-Our goal is to examine how the rate of infection $\beta(t)$ varies with observed policies
-and measures of social distancing behavior. A key challenge is that we only
-observed $C(t)$ and $D(t)$, but not $\Infected(t)$. The unobserved $\Infected(t)$ can
-be eliminated by differentiating (\ref{eq:c}) and using (\ref{eq:i})  as
-\begin{align}
-%  \ddot{C}(t) & = \dot{\tau}(t) + \dot{\Infected}(t) \notag \\
-%  & = \dot{\tau}(t)\Infected(t) + \frac{S(t)}{N} \beta(t) \Infected(t) -
-%    \frac{S(t-\ell)}{N} \beta(t-\ell) \Infected(t-\ell)  \notag \\
-%  & = \dot{\tau}(t) \Infected(t) + \frac{S(t)}{N} \beta(t) \Infected(t) -
-%    \frac{1}{\kappa} \dot{D}(t) \notag \\
-  \frac{\ddot{C}(t)}{\dot{C}(t)}
-              & =
-                \frac{S(t)}{N} \beta(t) -\gamma  + \frac{\dot{\tau}(t)}{\tau(t)}. \label{eq:c2}
-%                 -
-%                \frac{1}{\kappa}
-%                \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}. \label{eq:c2}
-\end{align}
-We consider a discrete-time analogue of equation (\ref{eq:c2}) to motivate our empirical
-specification by relating the detection rate $\tau(t)$  to the number of tests $T_{it}$ while specifying $\frac{S(t)}{N}\beta(t)$ as a linear function of variables $X_{i,t-14}$.
-This results in
-\begin{align}
-  \underbracket{\Delta \log(\Delta C_{it})}_{\frac{\ddot{C}(t)}{\dot{C}(t)}}
-  =
-      \underbracket{X_{i,t-14}' \theta + \epsilon_{it}}_{\frac{S(t)}{N}\beta(t) -\gamma}
-       +
-       & \underbracket{\delta_T \Delta
-      \log(T)_{it}}_{\frac{\dot{\tau}(t)}{\tau(t)} } \nonumber
-      %+
-%      \underbracket{\delta_D \frac{T_{it}\Delta D_{it}}{\Delta C_{it}}  }_{-\frac{1}{\kappa}
-%      \frac{\tau(t)\dot{D}(t)}{\dot{C}(t)}}, \nonumber % \label{eq:reg2}
-\end{align}
-which is equation (\ref{eq:M}), where $X_{i,t-14}$ captures a vector of variables related to $\beta(t)$.
-
-\begin{quote}
-\textsc{Structural Interpretation}. Early in the pandemic, when
-the number of susceptibles is approximately the same as the entire population, i.e. $S_{it}/N_{it} \approx 1$, the component $X_{i,t-14}' \theta$
-is the projection of infection rate $ \beta_i(t)$ on   $X_{i,t-14}$ (policy, behavioral, information, and confounders other than testing rate), provided
-the stochastic component $\epsilon_{it}$ is orthogonal
-to $X_{i,t-14}$ and the testing variables:
-$$
-\beta_i(t)S_{it}/N_{it}  - \gamma = X_{i,t-14}' \theta + \epsilon_{it}, \quad \epsilon_{it} \perp X_{i,t-14}.
-$$
-\end{quote}
-
-\subsection{Growth Rate in Deaths as Outcome}\label{sec:sirmodel-death}
-By differentiating (\ref{eq:d}) and (\ref{eq:c}) with respect to $t$ and using (\ref{eq:c2}), we obtain
-\begin{align}
-\frac{\ddot{D}(t) }{\dot D(t)}& = \frac{\ddot{C}(t) }{\dot C(t)}  - \frac{\dot{\tau}(t) }{ \tau(t)}    =  \frac{S(t)}{N}\beta(t)  -   \gamma.\label{eq:d2}
-\end{align}
-Our measurement equation for the growth rate of deaths is based on equation (\ref{eq:d2}) but   account for a $21$ day lag between infection and death as
-\begin{align}
-{\ycolor \Delta \log(\Delta D_{it})}  = X_{i,t-21}' \theta + \epsilon_{it},\label{eq:M-D} \tag{M-D}
-\end{align}
-where
-\begin{equation} \label{eq:y-d}
- \Delta \log(\Delta D_{it}):= \log( \Delta D_{it} ) -
-\log( \Delta D_{i,t-7})
-\end{equation}
-approximates the weekly growth rate in deaths from $t-7$ to $t$ in state $i$.
-% In our baseline specification, we consider 7 days lag between case confirmation and death by setting   $\ell=7$ in (\ref{eq:M-D}).
-%The appendix provides additional results under alternative assumptions including non-deterministic disease durations.
-
-\section{Decomposition and Counterfactual Policy Analysis}
-
-\subsection{Total Change Decomposition}
-Given the SEM formulation above, we can carry out the following decomposition analysis, after removing the effects of confounders. For example,
-we can decompose the total change $\Ep \tilde Y_{i,t+\ell} - \Ep \tilde Y_{io}$ in the expected outcome, measured at two different time points $t+\ell$ and $o$ into the sum of three components:
-\begin{equation} \label{eqn:DC}
-  \begin{aligned}
-\underbracket{\Ep \tilde Y_{i,t+\ell} - \Ep \tilde Y_{io}}_{\text{Total Change}}& =
-{\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left( \Ep \tilde P_{it} - \Ep \tilde P_{io} \right)}}_{\text{Policy Effect via Behavior}}} \\
-& +    {\underbracket{\pcolor\pi' \left( \Ep \tilde P_{it} - \Ep \tilde P_{io} \right)}_{\text{Direct Policy Effect}} } \\
-& +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( \Ep \tilde I_{it} - \Ep \tilde I_{io} \right)}+ {\icolor{\mu}'  \left( \Ep \tilde I_{it} - \Ep \tilde I_{io} \right)}}_{\text{ Dynamic Effect}}} \\
-& =:   \mathrm{PEB}_t + \mathrm{PED}_t + \mathrm{DynE}_t,
-  \end{aligned}
-\end{equation}
-where the first two components capture the immediate effect and the third represents the delayed or dynamic effect.
-
-In the three examples of information structure given earlier, we have the following forms for the dynamic effect: for the trend model,
-$$
-   \mathrm{DynE}_t   = (\gamma \alpha + \mu) \Delta g_t, \quad  \Delta g_t = (g(t) - g(t-\ell))\\
-$$
-and for the lag model, $$
-   \mathrm{DynE}_t  =    \sum_{m=1}^{t/\ell} \left( \gamma \alpha+ \mu   \right)^m (\mathrm{PEB}_{t-m \ell }  +\mathrm{PED}_{t-m \ell } ),
- $$
-  interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
- For the general model we use, the dynamic effect is
- $$
- \begin{array}{lll}
-(\mathrm{I})    & \mathrm{DynE}_t  & = \sum_{m=0}^{t/\ell} \left(( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m  \ \ ( (\gamma \alpha)_{1} + \mu_1) \Delta g_t  \\
-& & +   \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}_{t-m \ell } +  \mathrm{DPE}_{t- m \ell} \right)
-  \\
- &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}_{t-(m +1) \ell } +  \mathrm{DPE}_{t- (m+1) \ell}  \right).
- \end{array}
- $$
- The effects can be decomposed into (a) delayed policy effects via behavior by
-summing terms containing $PEB$, (b) delayed policy effects via direct impact
-by summing terms containing $DPE$,   (c) pure behavior effects, and (d)  pure dynamic feedback effects.
-
-\subsection{Counterfactuals}
-
-We  also consider simple counterfactual exercises, where we examine the effects of setting
-a sequence of counterfactual policies for each state:
-$$
-\{{\pcolor P^\star_{it}} \}_{t=1}^T, \quad i=1, \ldots N.
-$$
-We assume that the SEM remains invariant, except of course for the policy equation.\footnote{It is possible to consider counterfactual exercises in which policy  responds to information through the policy equation if we are interested in endogenous policy responses to information.
-Although this is beyond the scope of the current paper, counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information.
-} Given the policies,
-we propagate the dynamic equations:
-\begin{equation} \label{eqn:CF-SEM} \tag{CEF-SEM}
-  \begin{aligned}
-& {\ycolor Y^\star_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B^\star_{it} } , {\pcolor P^\star_{it}}, {\icolor I^\star_{it}}), \\
-& {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
-& {\icolor I^\star_{it} }& := &  I_{it}({\icolor I^\star_{i,t-\ell}}, {\ycolor Y^*_{it}}, t),  \end{aligned}
-\end{equation}
-with the initialization $I^\star_{i0} = 0$, $ Y^\star_{i0} =0$, $B^\star_{i0} = 0$, $P^\star_{i0} = 0$.
-In stating this counterfactual system of equations, we make the following invariance assumption
-\begin{quote}
-\textsc{Invariance Assumption.} The equations of (CF-SEM) remain exactly of the same form
-as in the (SEM) and (I).
-That is, under the policy intervention $\{{\pcolor P^\star_{it}} \}$,  parameters and stochastic shocks in (SEM) and (I) remain the same as under the original policy intervention $\{{\pcolor P_{it}} \}$.
-\end{quote}
-
-Let $\mathcal{P} Y^\star_{i,t+\ell}$  and  $\mathcal{P} Y_{i,t+\ell}$ denote the predicted values produced
-by working with the counterfactual system (\ref{eqn:CF-SEM}) and the factual system (SEM):
-$$
-\mathcal{P} Y^\star_{i,t+\ell}  = ( {\bcolor\alpha '} {\pcolor \beta'}  + {\pcolor\pi'})
-    {\pcolor P^\star_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
-    {\icolor I^\star_{it} }+ {\wcolor \bar \delta 'W_{it}},$$$$\mathcal{P} Y_{i,t+\ell} \ =
-    ( {\bcolor\alpha '}  {\pcolor \beta'} + {\pcolor\pi'})
-    {\pcolor P_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
-    {\icolor I_{it} }+ {\wcolor \bar \delta 'W_{it}}.
-$$
-In generating these predictions, we make the assumption of invariance stated above.
-
-Then we can write  the difference into the sum of three components:
-  \begin{align}
-\underbracket{\mathcal{P} Y^\star_{i,t+\ell} - \mathcal{P} Y_{i,t+\ell} }_{\text{Predicted CF Change}} & =
-\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}  +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \nonumber\\
-&\qquad +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \nonumber \\
- & =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}. \label{eqn:DC2}
-  \end{align}
-%\begin{equation} \label{eqn:DC2}
-%  \begin{aligned}
-%\underbracket{\mathcal{P} Y^\star_{it} - \mathcal{P} Y_{it} }_{\text{Predicted CF Change}}& =
-%{\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}} \\
-%& +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \\
-%& +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \\
-%& =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}.
-%  \end{aligned}
-%\end{equation}
-
-Similarly to what we had before,  the counterfactual dynamic effects take the form:
-$$
-\begin{array}{lll}
-(\mathrm{I})  & \mathrm{DynE}^\star_{it} &=
-  \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}^\star_{i,t-m \ell } +  \mathrm{DPE}^\star_{i,t- m\ell} \right)
-  \\
- &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}^\star_{i,t-(m +1) \ell } +  \mathrm{DPE}^\star_{i,t- (m+1)\ell}  \right),
- \end{array}
- $$
- interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
- The effects can be decomposed into (a) delayed policy effects via behavior by
-summing terms containing $PEB$, (b) delayed policy effects via direct impact
-by summing terms containing $DPE$,  (c) pure behavior effects, and (d)  pure dynamic feedback effects.
-
-\section{Empirical Analysis}
-\subsection{Data}
-
-Our baseline measures for daily Covid-19   cases and deaths are from
-\href{https://github.com/nytimes/covid-19-data}{The New York Times (NYT)}.
-When there are missing values in NYT, we use reported cases and deaths from \href{https://github.com/CSSEGISandData/COVID-19}{JHU CSSE}, and then the
-\href{https://github.com/COVID19Tracking/covid-tracking-data}{Covid
-Tracking Project}.  The number of tests for each state is from  \href{https://github.com/COVID19Tracking/covid-tracking-data}{Covid
- Tracking Project}. As shown in Figure \ref{fig:test} in the appendix, there was a rapid increase in testing in the second half of March and then the
-number of tests increased very slowly in each state in April.
-
-We use the database on US state policies created by
-\cite{raifman2020}.
-%This database records that each state implemented
-%25 Covid-19  related policies.
-In our analysis, we focus on 6
-policies:  stay-at-home, closed nonessential
-businesses, closed K-12 schools, closed restaurants except takeout, closed movie theaters, and mandate face mask by  employees in public facing businesses.
-We believe that the first four of these
-policies are the most widespread and important.
-%, where state of emergency was declared first and the implementation of other policies followed with some time lag.
-Closed movie theaters is included because it captures common bans on gatherings of more than
-a handful of people.
-We also include mandatory face mask use by employees because its effectiveness on slowing down Covid-19 spread is a controversial policy issue \citep{howard2020,Greenhalghm2020,zhangr2020}.
-Table \ref{tab:policies_inreg} provides summary statistics, where $N$ is the number of states that have ever implemented the policy.
-We also obtain information on state-level covariates from \citet{raifman2020}, which include population size, total area, unemployment rate, poverty rate, and a percentage of people who are subject to illness.These confounders are motivated by \cite{wheaton2020} who finds that case growth is associated  with residential density and per capita income.
-
-\input{tables_and_figures/policies_inregs}
-
-We obtain social distancing behavior measures from``\href{https://www.google.com/covid19/mobility/}{Google COVID-19 Community Mobility Reports}''
-\citep{google2020}.
-The dataset provides six measures of ``mobility trends''  that report a percentage change in visits and length of stay at different places relative to a baseline computed by their median values of the same day of the week from January 3 to February 6, 2020.
-Our analysis focuses on the following four measures: ``Grocery \& pharmacy," ``Transit stations,'' ``Retail \& recreation,'' and ``Workplaces.''\footnote{The other two measures are ``Residential'' and  ``Parks.''  We drop ``Residential'' because it is highly correlated with both ``Workplaces'' and  ``Retail \& recreation''  at correlation coefficients of -0.98.
-We also drop ``Parks'' because it does not have clear implication on the spread of Covid-19.}
-
-Figure \ref{fig:transit-workplaces} shows the evolution of  ``Transit stations'' and ``Workplaces,'' where thin lines are the value in each state and date while thicker colored lines are quantiles  conditional on date.
-The figures illustrate a sharp decline in people's movements starting from mid-March as well as differences in their evolutions across states.
-They also reveal periodic fluctuations associated with days of the week, which motivates our use of weekly measures.
-
-\begin{figure}\caption{The Evolution of Google Mobility Measures: Transit stations and Workplaces\label{fig:transit-workplaces}}\vspace{0.1cm}
-    \begin{tabular}{cc}
-      \includegraphics[width=0.5\linewidth]{tables_and_figures/transit}
-      \includegraphics[width=0.5\linewidth]{tables_and_figures/workplaces}
-       \end{tabular}
-        \begin{flushleft}
-        \scriptsize{This figure shows the evolution of ``Transit stations'' and  ``Workplaces'' of Google Mobility
-      Reports. Thin gray lines are the value in each state and date. Thicker colored lines are
-      quantiles of the variables conditional on date.}
-      \end{flushleft}
-\end{figure}
-
-In our empirical analysis, we use weekly measures for cases, deaths,
-and tests by summing up their daily measures from day \(t\) to
-\(t-6\).  We focus on weekly cases and deaths because daily new cases and deaths are
-affected by the timing of reporting and testing, and are quite
-volatile as shown in Figure \ref{fig:dailycases} in the
-appendix.  Aggregating to weekly new cases/deaths/tests smooths
-out idiosyncratic daily noises as well as periodic fluctuations
-associated with days of the week.
-We also construct weekly policy and
-behavior variables by taking 7 day moving averages from day \(t-14\) to
-\(t-21\) for case growth, where the  delay reflects the time lag between infection and case confirmation. The four weekly behavior variables are referred as ``Transit
-Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
-Intensity.'' Consequently, our empirical analysis uses 7 day moving averages of all variables recorded at daily frequencies. Our sample period is from March 7, 2020 to June 3, 2020.
-
-\begin{table}\caption{Correlations among Policies and Behavior \label{tab:correlation}}\vspace{-0.2cm}
-  \begin{minipage}{\linewidth}
-    \resizebox{\linewidth}{!}{
-      \input{tables_and_figures/corr}
-    }\smallskip
-       \begin{flushleft}
-         \scriptsize
-         Each off-diagonal entry reports a correlation coefficient of
-         a pair of policy and behavior variables.
-       \end{flushleft}
-  \end{minipage}
-\end{table}
-
-Table \ref{tab:correlation} reports that weekly policy and behavior
-variables are highly correlated with each other, except for the``masks
-for employees'' policy.  High correlations may cause multicolinearity
-problems and could limit our ability to separately identify the
-effect of each policy or behavior variable on case growth, but this
-does not prevent us from identifying the aggregate effect of all
-policies and behavior variables on case or death growth.
-
-Figure \ref{fig:policyportion} shows the portion of states that have
-each policy in place at each date. For most policies, there is
-considerable variation across states in the time in which the policies
-are active. The one exception is K-12 school closures. About 80\% of
-states closed schools within a day or two of March 15th, and all
-states closed schools by early April. This makes the effect of school
-closings difficult to separate from aggregate time series variation.
-
-\begin{figure}[ht]\caption{Portion of states with each
-    policy \label{fig:policyportion}}
-  \begin{minipage}{\linewidth}
-    \begin{tabular}{ccc}
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/pmaskbus_p}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/pk12_p}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/pshelter_p}
-      \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/pmovie_p}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/prestaurant_p}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/pnonessential_p}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\subsection{The Effect of Policies and Information on Behavior\label{policies-and-behavior}}
-
-We first examine how policies and information affect social distancing behaviors by estimating a version of (\ref{eq:R2}):
-\begin{align}
-  {\bcolor B_{it}^j}
-  %& = {\pcolor (\beta^j)' P_{it} }  + {\icolor \gamma_{1}^j \log(t) + \gamma_{2}^j \log (\Delta C_{i,t-14}) +  \gamma_{3}^j Y_{i,t-7}} + {\wcolor (\delta_B^j)' W_{it}} + \varepsilon_{it}^j \nonumber\\
-  & = {\pcolor (\beta^j)' P_{it}} + {\icolor (\gamma^j)' I_{it}} +
-    {\wcolor (\delta_B^j)' W_{it}} + \varepsilon_{it}^{bj}, \notag
-\end{align}
-where ${\bcolor B_{it}^j}$ represents behavior variable $j$  in state  $i$ at time $t$.
-${\pcolor P_{it}}$ collects the Covid related policies  in state $i$ at $t$.
-Confounders, ${\wcolor W_{it}}$, include state-level covariates, month
-indicators, and their interactions.
-${\icolor I_{it}}$ is a set of information variables that affect
-people's behaviors at $t$. As information, we include each state's
-growth of cases (in panel \ref{tab:PItoBcases}) or deaths (in panel
-\ref{tab:PItoBdeaths}), and log cases or deaths.  Additionally, in
-columns (5)-(8) of each panel, we include national growth and log
-of cases or deaths.
-
-\afterpage{\clearpage
-  %\thispagestyle{empty}
+  \afterpage{
  \begin{landscape}
 \begin{table}[!htbp] \centering
-      \caption{The Effect of Policies and Information on Behavior ($P
-        I \to B$)}\vspace{-0.3cm}\label{tab:PItoB}
-
-      \begin{subtable}{\linewidth}\caption{Cases as Information\label{tab:PItoBcases}}
+ \caption{\label{tab:PtoY-robust}
+Robustness Check for the Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
  \begin{minipage}{\linewidth}
+   \resizebox{\textwidth}{!}{
    \centering
-        \resizebox{\textwidth}{!}{
-          \footnotesize
-          \input{tables_and_figures/behavior-NI-pib}
-        }
-      \begin{flushleft}
-        \scriptsize
-        Dependent variables are   ``Transit
-        Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
-        Intensity" defined as 7 days moving averages of corresponding daily measures obtained from Google Mobility Reports.
-All specifications include state-level characteristics (population, area, unemployment rate, poverty rate, and a percentage of people subject to illness) as well as their interactions with the log of days since Jan 15, 2020.
-The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients.
-Standard errors are clustered at the state level.
-        \end{flushleft}
-      \end{minipage}
-    \end{subtable}
-  \end{table}
-  \pagebreak
-  \begin{table}[!htbp] \centering
-    \ContinuedFloat
-    \begin{subtable}{\linewidth}\caption{Deaths as Information \label{tab:PItoBdeaths}}
-      \begin{minipage}{\linewidth}
-        \centering
-        \resizebox{\textwidth}{!}{
-          \footnotesize
-            \input{tables_and_figures/behavior-deathsinfo-NI-pib}
-          }
-        \begin{flushleft}
-          \scriptsize
-          Dependent variables are   ``Transit
-          Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
-          Intensity" defined as 7 days moving averages of corresponding daily measures obtained from Google Mobility Reports.
-All specifications include state-level characteristics (population, area, unemployment rate, poverty rate, and a percentage of people subject to illness) as well as their interactions with the log of days since Jan 15, 2020.
-The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients.
-Standard errors are clustered at the state level.
-        \end{flushleft}
-      \end{minipage}
-    \end{subtable}
-  \end{table}
-\end{landscape}
-\clearpage}
-
-Table \ref{tab:PItoB} reports the estimates with standard errors clustered at the state level.  Across different
-specifications, our results imply that policies have large effects on
-behavior. Comparing columns (1)-(4) with columns (5)-(8),
-the magnitude of policy effects are sensitive to whether national
-cases or deaths are included as information. The coefficient on school
-closures is particularly sensitive to the inclusion of national
-information variables.  As shown in Figure
-\ref{fig:policyportion}, there is little variation across states in
-the timing of school closures. Consequently, it is difficult to
-separate the effect of school closures from a behavioral response to the national
-trend in cases and deaths.
-
-The other policy coefficients are less sensitive to the inclusion of
-national case/death variables. After school closures, stay-at-home
-orders and restaurant closures have the next largest effect. Somewhat
-surprisingly, closure of nonessential businesses appears to have
-a modest effect on behavior. Closing movie theaters has a similar,
-small effect on behavior. The effect of masks for employees is also
-small. The comparison of effects across policies should be interpreted
-with caution.  Differences between policy effects are often
-statistically insignificant; except for masks for employees, the
-policies are highly correlated and it is difficult to separate their
-effects.
-
-The row ``$\sum_j \mathrm{Policy}_j$'' reports the sum of the estimated
-effect of all policies, which is substantial and can account for a
-large fraction of the observed declines in behavior variables.  For
-example,  in Figure \ref{fig:transit-workplaces}, transit intensity for a median state was approximately -50\% at its lowest point
- in early April. The estimated policy coefficients in columns imply
-that imposing all six policies would lead to roughly 85\% (in columns
-1-4) or roughly 50\% (in columns 5-8) of the observed decline.
-The large impact of policies on transit intensity suggests that the policies may have reduced the Covid-19 infection by reducing people's use of public transportation.\footnote{Analyzing the New York City's subway ridership,  \cite{NBERw27021} finds  a strong link between public transit and spread of infection.}
-
-%retail intensity was approximately -40\% at its lowest point
-%in mid April, and the estimated policy coefficients in columns imply
-%that imposing all six policies would lead to either all (in columns
-%1-4) or half (in columns 5-8) of the observed decline.
-
-In panel \ref{tab:PItoBdeaths}, estimated coefficients of deaths and
-death growth are generally negative. This suggests that the higher
-number of deaths reduces social interactions measured by Google
-Mobility Reports perhaps because people are increasingly aware of
-prevalence of Covid-19 \citep{maloney2020}.  The coefficients on log
-cases and case growth in panel \ref{tab:PItoBcases} are more mixed. In
-columns (5)-(8) of both panels, we see that national case/death
-variables have large, negative coefficients.
-This suggests that behavior responded to national conditions although it is also likely that national case/death variables capture unobserved aggregate time effects beyond information (e.g., latent policy variables and time-varying confounders that are common across states) which are not fully controlled by month dummies.
-
-%For grocery, the positive estimated coefficients of cases and case growth in columns (3) and (7) might be capturing stockpiling behavior, which could temporarily increase the visits to grocery.
-
-%In columns (5)-(8), the effects of masks for employees are allowed to be different before and after May 1st.
-The negative effect of mask mandates on mobility measures appears to be slightly larger in May than in April but the difference is not statistically significant.
-
-\begin{figure}[ht]
-  \caption{Case and death growth conditional on mask mandates \label{fig:masks}}\vspace{0.2cm}
- \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases}
-      &
-        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths}
-    \end{tabular}
-    \begin{flushleft}
-      \scriptsize In these figures, red points are the case or death
-      growth rate in states without a mask mandate. Blue points are
-      states with a mask mandate 14 (21 for deaths) days prior. The
-      red line is the average across states without a mask mandate 14
-      (21 for deaths) days earlier. The blue line is the average
-      across states with a mask mandate 14 (21 for deaths) earlier.
-    \end{flushleft}
-  \end{minipage}
-\end{figure}
-
-\afterpage{
- \clearpage
- %\thispagestyle{empty}
- \begin{landscape}
-\begin{table}[!htbp] \centering
- \caption{The Direct Effect of Behavior and Policies on Case and
-   Death Growth ($BPI \to Y$)}\vspace{-0.3cm}
- \label{tab:BPItoY}
- \begin{minipage}{\linewidth}
-   \centering
-   \resizebox{\textwidth}{!}{
    \tiny
    \begin{tabular}{c|c}
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.65\linewidth}
      \centering
-     \input{tables_and_figures/cases-pbiy}
+     \input{tables_and_figures/cases-robust-piy}
    \end{minipage}
-     %\hspace{1pt}
      &
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.65\linewidth}
      \centering
-     \input{tables_and_figures/deaths-deathsinfo-pbiy}
+     \input{tables_and_figures/deaths-robust-piy}
    \end{minipage}
+
    \end{tabular}
@@
      confirmed cases (in the left panel) or deaths (in the right
-     panel) as defined in equation (\ref{eq:y}). The covariates
-     include lagged policy and behavior variables, which are
+     panel) as defined in equation (\ref{eq:y}). Columns (2) and (6) exclude the observations of the state of New York.
+      The covariates
+     include lagged policy variables, which are
      constructed as 7 day moving averages between $t$ to $t-7$ of
@@
      ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
-     coefficients.  The row ``$\sum_k w_k \mathrm{Behavior}_k$''
-     reports the sum of four coefficients of behavior variables
-     weighted by the average of each behavioral variable from April
-     1st-10th.  Standard errors are clustered at the state level.
+     coefficients.  Standard errors are clustered at the state level.
    \end{flushleft}
@@
 \end{landscape}
-\clearpage
 }
-
-\subsection{The Direct Effect of Policies and Behavior on Case
-and Death Growth\label{policy-behavior-and-case-growth}}
-
-We now analyze how behavior and policies together influence case and death
-growth rates. We begin with some simple graphical evidence of the
-effect of policies on case and death growth. Figure \ref{fig:masks}
-shows average case and death growth conditional on date and whether
-masks are mandatory for employees.\footnote{We take 14 and 21 day lags of mask policies for case and death growths, respectively, to  identify the states with a mask mandate because policies affect cases and deaths with time lags.
-See our discussion in the Appendix \ref{sec:timing}.
-} The left panel of the figure shows
-that states with a mask mandate consistently have 0-0.2 lower case
-growth than states without. The right panel  also illustrates that
-states with a mask mandate tend to have lower average death growth than states without a mask mandate.
-
-Similar plots are shown for other policies in Figures
-\ref{fig:growthpolicies1} and \ref{fig:growthpolicies2} in the
-appendix. The  figures for stay-at-home orders and closure
-of nonessential businesses are qualitatively similar to that for
-masks. States with these two policies appear to have about 0.1 percentage point lower
-case growth  than states without. The effects of school closures, movie theater closures, and
-restaurant closures are not clearly visible in these figures. These
-figures are merely suggestive; the patterns observed in them may be
-driven by confounders.
-
-We more formally analyze the effect of policies by estimating
-regressions. We first look at the direct effect of policies on case and death growth
-conditional on behavior by estimating equation (\ref{eq:R1}):
-% We partition $X_{it}$ in (\ref{eq:M}) into behavior variables $B_{it}$,  policy variables $P_{it}$, information variables $I_{it}$, and other confounding variables $W_{X,it}$ as $X_{it}= (B_{it}',P_{it}', I_{it}', W_{X,it}')'$ with $\theta=(\alpha',\pi',\mu',\delta_X')'$ and rewrite  (\ref{eq:M}) as:
-\begin{align}
-  {\ycolor  Y_{i,t+\ell}}
-  & = {\bcolor \alpha ' B_{it}} + {\pcolor\pi 'P_{it}} + {\icolor \mu'I_{it}} +  {\wcolor\delta_Y 'W_{it}}  + \varepsilon^y_{it},
-  \label{eq:reg-y}
-\end{align}
-where the outcome variable, $\ycolor Y_{i,t+\ell}$, is either case growth  or death growth.
-
-For case growth as the outcome, we choose a lag length of $\ell=14$ days for behavior, policy, and information variables to reflect the delay between infection and
-confirmation of case.\footnote{As we review in the Appendix \ref{sec:timing},
-  a lag length of 14 days between exposure and case reporting, as well as a lag length of 21 days between exposure and deaths, is broadly consistent with currently
-  available evidence.  }   $\bcolor B_{it}=(B_{it}^1,...,B_{it}^4)'$ is a vector of four behavior
-variables  in state $i$. $\pcolor P_{it}$ includes the Covid-related policies
-in state $i$  that directly affect the spread of Covid-19
-after controlling for behavior variables (e.g., masks for employees).
-We include information variables,  $\icolor I_{it}$, that include the past cases and case growths
-because  the past cases may be
-correlated with (latent) government  policies or people's behaviors
-that are not fully captured by our observed policy and behavior
-variables.
-We also consider a specification that includes the past cases and case growth at the national level as additional information variables.
-$\wcolor W_{it}$ is a set of confounders that include month dummies, state-level covariates,  and the interaction terms between month dummies and state-level covariates.\footnote{Month dummies also represent the latent information that is not fully captured by the past cases and growths.}
-For case growth, $\wcolor W_{it}$ also includes the  test rate growth  $\Delta \log(T)_{it}$  to capture the effect of changing test rates on confirmed cases.
-Equation (\ref{eq:reg-y}) corresponds to (\ref{eq:M})  derived from the SIR model.
-
-%We also estimate the effect of policies on death growth by estimating  (\ref{eq:reg-y}) with death growth
-%$\Delta \log \Delta D_{it}$  as outcome variable.
-For death growth as the outcome, we take a lag length of $\ell=21$ days.
-The information variables $\icolor I_{it}$ include  past deaths and death growth rates;  $\wcolor W_{it}$ is the same as that of the case growth equation except that the growth rate of test rates is excluded from $\wcolor W_{it}$ as implied by equation (\ref{eq:M-D}).
-
-% Tables \ref{tab:BPItoY-c} and  \ref{tab:BPItoY-d} show
-Table \ref{tab:BPItoY} shows
-the results of estimating
-(\ref{eq:reg-y}) for case and death growth rates. %, respectively.
-Column (1) represents our baseline specification while column (2) allows the effect of masks to be different before and after May 1st.
-Columns (3) and (4) include past cases/deaths and growth rates at national level as additional regressors.
-
-The estimates indicate that mandatory face masks for employees
-reduce the growth rate of infections and deaths by 8-15 percent, while holding behavior constant. This suggests that
-requiring masks for employees in public-facing businesses may be an effective preventive measure.\footnote{Note that we are \textit{not} evaluating the effect of \textit{universal} mask-wearing for the public but that of mask-wearing for employees.
-The effect of \textit{universal} mask-wearing for the public could be larger if people comply with such a policy measure.
-\cite{tian2020calibrated} considered a  model in which mask wearing reduces  the reproduction number by a factor $(1-e \cdot pm)^2$, where $e$ is the efficacy of trapping viral particles inside the mask and $pm$ is the percentage of mask-wearing population.
-Given an estimate of $R_0=2.4$, \cite{howard2020} argue that  50\% mask usage and a 50\% mask efficacy level would reduce the reproduction number from 2.4 to 1.35, an order of magnitude
-impact.} The estimated effect of masks on death growth is larger than the effect on case growth, but this difference between the two estimated effects is not statistically significant.
-
-Except for mask requirements, policies appear to have little direct
-effect on case or death growth when behavior is held constant. The one
-exception is that closing schools has a large and statistically
-significant coefficient in the death growth regressions. As discussed
-above, there is little cross-state variation in the timing of school
-closures, making estimates of its effect less reliable.
-
-The row ``$\sum_k w_k \mathrm{Behavior}_k$'' reports the sum of estimated coefficients weighted by the average of the behavioral
-variables from April 1st-10th.
-The estimates of $-0.76$ and $-0.87$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 76 and 83 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
-When including national cases and deaths in information, as shown in columns (3) and (4), the estimated aggregate impact of behavior is substantially smaller, but remains large and statistically significant.
-
-A useful practical implication of these results are that Google Mobility
-Reports and similar data might be useful as a leading indicator of
-potential case or death growth. This should be done with caution, however, because other
-changes in the environment might alter the relationship between behavior
-and infections. Preventative
-measures, including mandatory face masks, and changes in habit that are not captured in our data might alter
-the future relationship between Google Mobility Reports and case/death growth.
-
-The negative coefficients of past cases or deaths in
-% tables \ref{tab:BPItoY-c} and  \ref{tab:BPItoY-d}  are
-Table \ref{tab:BPItoY} is
-consistent with a hypothesis that higher reported cases and deaths change people's behavior to reduce transmission risks. Such behavioral changes in response to new information   are
- partly captured by Google mobility measures, but the negative estimated coefficient of past cases or deaths imply that other latent behavioral changes that are not fully captured by Google mobility measures (e.g., frequent hand-washing, wearing masks, and keeping 6ft/2m distancing) are also important for reducing future cases and deaths.
-
-If policies are enacted and behavior changes, then future cases/deaths and
-information will change, which will induce further behavior changes.
-However, since the model includes lags of cases/deaths as well as their growth rates,  computing a long-run effect is not completely
-straightforward. We investigate dynamic effects that incorporate
-feedback through information in section \ref{counterfactuals}.
-
-% \begin{table}[!htbp] \centering
-%   \caption{\label{tab:PtoY-c}
-%     The Total Effect of Policies on Case Growth ($PI \to Y$)}\vspace{-0.3cm}
-%   \begin{minipage}{\linewidth}
-%     \centering
-%     \scriptsize
-%     \begin{tabular}{c}
-%     \begin{minipage}{\linewidth}
-%       \centering
-%       \input{tables_and_figures/cases-piy}
-%     \end{minipage}
-%     \end{tabular}
-%     \begin{flushleft}
-%       \scriptsize Dependent variable is the weekly growth rate of
-%       confirmed cases (in the left panel) or deaths (in the right
-%       panel) as defined in equation (\ref{eq:y}). The covariates
-%       include lagged policy variables, which are
-%       constructed as 7 day moving averages between $t$ to $t-7$ of
-%       corresponding daily measures.  The row
-%       ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
-%       coefficients.
-%     \end{flushleft}
-%   \end{minipage}
-% \end{table}
-
-% \begin{table}[!htbp] \centering
-%   \caption{\label{tab:PtoY-d}
-%     The Total Effect of Policies on Death Growth ($PI \to Y$)}\vspace{-0.3cm}
-%   \begin{minipage}{\linewidth}
-%     \centering
-%     \scriptsize
-%     \begin{tabular}{c}
-%     \begin{minipage}{\linewidth}
-%       \centering
-%       \input{tables_and_figures/deaths-deathsinfo-piy}
-%     \end{minipage}
-
-%     \end{tabular}
-%     \begin{flushleft}
-%       \scriptsize Dependent variable is the weekly growth rate of
-%       confirmed cases (in the left panel) or deaths (in the right
-%       panel) as defined in equation (\ref{eq:y}). The covariates
-%       include lagged policy variables, which are
-%       constructed as 7 day moving averages between $t$ to $t-7$ of
-%       corresponding daily measures.  The row
-%       ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
-%       coefficients.
-%     \end{flushleft}
-%   \end{minipage}
-% \end{table}
-
-%Here we present the results of counterfactual analyses that don't include the national cases/deaths as
-%the the information variables.  The growth in the national cases nearly perfectly predicts school closure
-%policies, which have little cross-sectional variation.  Here we remove the national cases from the information
-%and check what happens to the counterfactual results.  The counterfactual results of mask policies, shelter-in-place,
-%and closing
-%non-essential businesses remain robust. The results on school closures effects change substantially,
-%attributing a much bigger role to school closures. The resulting counterfactual of  removing
-%all policies has a much larger effect on cases, with the lower bound of the confidence band implying at
-%least a 7 fold increase in cases (see Figure \ref{fig:US-nop-SI} given in the main text).
 
@@
 \begin{table}[!htbp] \centering
- \caption{\label{tab:PtoY}
-   The Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
+ \caption{\label{tab:PtoY-fe}
+Fixed Effects Specification for  the Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
  \begin{minipage}{\linewidth}
@@
    \begin{tabular}{c|c}
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.5\linewidth}
      \centering
-     \input{tables_and_figures/cases-piy}
+     \input{tables_and_figures/cases-fe-piy}
    \end{minipage}
      &
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.5\linewidth}
      \centering
-     \input{tables_and_figures/deaths-deathsinfo-piy}
+     \input{tables_and_figures/deaths-fe-piy}
    \end{minipage}
@@
 
-\afterpage{
-\begin{table}[!b]
-  \caption{\label{tab:dieff-si}Direct and Indirect Policy Effects
-    without national case/death variables}
-\begin{minipage}{\linewidth}
-  \centering
-    \scriptsize
-  \begin{tabular}{c}
-    \textbf{Cases}
-    \\
-    \input{tables_and_figures/dieff-cases-SI}
-    \\
-    \textbf{Deaths}
-    \\
-    \input{tables_and_figures/dieff-deaths-deathsinfo-SI}
-  \end{tabular}
-  \smallskip
-  \begin{flushleft}
-    {\tiny Direct effects capture the effect of policy on case
-      growth holding behavior, information, and confounders
-      constant. Direct effects are given by ${\pcolor \pi}$ in
-      equation (\ref{eq:R1}). Indirect effects capture how policy
-      changes behavior and behavior shift case growth. They are given
-      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
-      from (\ref{eq:R2}). The total effect is
-      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
-      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
-      \ref{eq:R4}. Columns ``Difference''    are  the differences between
-      the estimates from (\ref{eq:R4}) and the combination of
-      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
-      Standard errors are computed by
-      bootstrap and clustered on state.}
-    %  \ref{tab:PtoY}.
-    \end{flushleft}
-  \end{minipage}
-\end{table}
-}
-
-\afterpage{
-\begin{table}
-  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
-\begin{minipage}{\linewidth}
-  \centering
-    \scriptsize
-  \begin{tabular}{c}
-    \textbf{Cases}
-    \\
-    \input{tables_and_figures/dieff-cases-NI}
-    \\ \\
-    \textbf{Deaths}
-    \\
-    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
-    \\
-  \end{tabular}
-  \smallskip
-  \begin{flushleft}
-      \scriptsize Direct effects capture the effect of policy on case
-      growth holding behavior, information, and confounders
-      constant. Direct effects are given by ${\pcolor \pi}$ in
-      equation (\ref{eq:R1}). Indirect effects capture how policy
-      changes behavior and behavior shift case growth. They are given
-      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
-      from (\ref{eq:R2}). The total effect is
-      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
-      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
-      \ref{eq:R4}. Columns ``Difference''    are  the differences between
-      the estimates from (\ref{eq:R4}) and the combination of
-      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
-      Standard errors are computed by
-      bootstrap and clustered on state.
-    %  \ref{tab:PtoY}.
-    \end{flushleft}
-  \end{minipage}
-\end{table}
-\clearpage}
-
-\subsection{The Total Effect of Policies on Case Growth\label{total-policy-effect}}
-
-In this section, we focus our analysis on policy effects when we hold
-information constant. The estimated effect of policy on
-behavior in Table \ref{tab:PItoB} and those of policies and
-behavior on case/death growth in Table  \ref{tab:BPItoY}  can be
-combined to calculate the total effect of policy as well as its
-decomposition into  direct and indirect effects.
-
-The first three columns of Table \ref{tab:dieff-si} show the direct
-(holding behavior constant) and indirect (through behavior changes)
-effects of policy under a specification that excludes national information variables. These are computed from the specification with
-national cases or deaths included as information (columns (1)-(4) of
-Table \ref{tab:PItoB} and column (1) of Table \ref{tab:BPItoY}). The estimates imply
-that all policies combined would reduce the growth rate of cases  and deaths by
-0.70 and 0.98, respectively,   out of which about two-third to
-three-fourth is attributable to the indirect effect through their impact on behavior.  The estimate also indicates that the
-effect of mandatory masks for employees is mostly direct.
-
-We can also examine the total effect of policies and information on
-case  or death growth, by estimating (\ref{eq:R4}). The coefficients on policy
-in this regression combine both the direct and indirect effects.
-
-Table \ref{tab:PtoY} shows
-the full set of coefficient estimates for
-(\ref{eq:R4}). The results are broadly consistent with what we found
-above.  As in Table \ref{tab:PItoB}, the effect of school closures is
-sensitive to the inclusion of national information variables.  Also as
-above, mask mandates have a significant negative effect on growth
-rates.
-
-In columns (2) and (4) of Table \ref{tab:PtoY}, we find that the estimated effect of mask mandates in April is larger than that in May for both case and death regressions.
-This may reflect a wider \textit{voluntary} adoption of masks in May than in April --- if more people wear masks even without mandatory mask policy,  the policy effect of mandating masks for employees becomes weaker.
-
-Table \ref{tab:dieff} presents the estimates for the specification with past national case/death variables.
-The effects of school closures and  the sum of policies are estimated substantially smaller in Table \ref{tab:dieff} when national case/death variables are included than in Table \ref{tab:dieff-si}.
-This sensitivity reflects the difficulty in identifying the aggregate time effect---which is largely captured by national cases/deaths---given little cross-sectional variation in the timing of school closures across states.
-On the other hand, the estimated effects of policies other than school closures are similar between Table \ref{tab:dieff-si} and Table \ref{tab:dieff}; the effect of other policies are well-identified from cross-sectional variations.
-
-Column ``Difference''  in Tables   \ref{tab:dieff-si} and \ref{tab:dieff}  show the
-difference between the estimate of (\ref{eq:R4}) in column  ``PI$\to$Y
-Coefficient''   and the implied estimate from
-(\ref{eq:R1})-(\ref{eq:R2}) in   column ``Total.''   Differences  are generally small and statistically insignificant, broadly supporting the validity of extra orthogonality condition in (\ref{eq:R1}).
-The difference for school closures as well as  the sum of all policies in Table   \ref{tab:dieff-si} is significantly different from zero, which may be due to the aforementioned difficulty in identifying the effect of school closures.
-There is substantial external epidemiological evidence that suggests that schooling closures may have substantial effects on the spread of the virus: studies like \cite{children:germany} and \cite{children:nature} establish that children carry substantial amounts of viral loads and can contribute to the transmission (due to higher contact rate than other age groups).\footnote{The evidence presented in \cite{children:germany}  has lead German to make the decision to close schools early.} The US data does not allow us to pint down the effect of closing schools reliably due to their approximate collinearity with trends in national cases.
-
-Column ``Average'' of  Tables \ref{tab:dieff-si} and \ref{tab:dieff}   reports the average of
-``Total'' and ``PI$\to$Y Coefficient'' columns.  The average is an appealing and simple way
-to combine the two estimates of the total effect: one relying on the causal structure and another inferred from a direct estimation of equation (PI $\to$ Y).\footnote{Averaging the two estimates theoretically reduces noise, albeit in our case the reductions are small.
-Another approach would be to use precision averaging, which would give similar result. Finally, another approach would be to use generalized method of moments to estimate
-all of the equations jointly.
-We don't pursue this approach since it is likely to be non-robust under local deviations from correct specification; simple model averaging is more appealing in this case.}  We shall be using the average estimate in generating the counterfactuals in the next section.
-Turning to the results, the estimates of Tables \ref{tab:dieff-si} and \ref{tab:dieff}  imply that all policies combined would reduce
-\(\Delta \log \Delta D\) by  0.97 and  0.40, respectively. For
-comparison, the median of \(\Delta \log \Delta D_{it}\) reached its
-peak in mid-March of about 1.3 (see Figure \ref{fig:growthq} in the
-appendix). Since then it has declined to near 0. Therefore,  -0.97 and -0.40 imply
-that policy changes can account for roughly one-third to two-third of the observed
-decrease in death growth.  The remainder of the decline is likely due
-to changes in behavior from information.
-
-\section{Empirical Evaluation of Counterfactual Policies}\label{counterfactuals}
-
-We now turn our focus to dynamic feedback effects. Policy
-and behavior changes that reduce case and death growth today can lead to a more
-optimistic, riskier behavior in the future, attenuating longer run effects.
-We perform the main counterfactual experiments using the average of two estimated coefficients as reported in column ``Average'' of Table \ref{tab:dieff-si} under a specification that excludes the number of past national cases and deaths from information variables.
-In the appendix, we also report additional counterfactual experiment results with the specification that includes the national information variables, and find that they are very similar.
-The results on mask policies, business closures, stay-at-home orders are robust with respect to this variation (see Figures \ref{fig:US-mask}-\ref{fig:US-shelter} in the appendix).
-On the other hand, the results on removing all policies, particularly closure of schools, reported in the next section, are sensitive to the inclusion of national information variables, highlighting the large uncertainty regarding the size of the effect.
-In Figures \ref{fig:WA-mask}-\ref{fig:US-shelter} below, the top panel presents the result on cases while the bottom panel presents the result on deaths.
-
-%Additional counterfactual experiments include the
-%} %\footnote{\textbf{The results of counterfactual experiments are not sensitive to whether we use the estimate of  table \ref{tab:dieff-si} or that of table \ref{tab:dieff} except for a counterfactual experiment of removing all policies  in section \ref{removing-policies}.}}
-
-\subsection{Business Mask Mandate}
-
-\begin{figure}[ht]
-  \caption{Effect of mandating masks on April 1st in Washington State \label{fig:WA-mask}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_v1}
-      \\
-      \\
-      \multicolumn{3}{c}{\textbf{Deaths}}
-      \\      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_v1}
-    \end{tabular}
-    \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff-si}. We set initial
-      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-      observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
-      point-wise 90\% confidence interval.
-    \end{flushleft}
-  \end{minipage}
-\end{figure}
-
-We first consider the impact of a nationwide mask mandate for
-employees beginning on April 1st. As discussed earlier, we find that
-mask mandates reduce case and death growth even when holding behavior
-constant. In other words, mask mandates may reduce infections with
-relatively little economic disruption. This makes mask mandates a
-particularly attractive policy instrument. In this section we examine
-what would have happened to the number of cases if all states had imposed a mask
-mandate on April 1st.\footnote{We feel this is a very plausible
-  counterfactual policy. In a paper made publicly available on
-  April 1st, \cite{abaluck2020} argued for mask usage based on
-  comparisons between countries with and without pre-existing norms of
-  widespread mask usage.}
-
-For illustrative purpose, we begin by focusing on Washington State.
-The
-left column of  Figure \ref{fig:WA-mask} shows the observed, estimated
-average, and counterfactual average of $\Delta \log \Delta C$ (top panel) and
-$\Delta \log \Delta D$ (bottom panel). To compute the estimated and counterfactual
-paths, we use the   estimate   in
-column ``Average'' of Table \ref{tab:dieff-si}. We set initial
-$\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-observed in the state we are simulating. We hold all other regressors
-at their observed values. Error terms are drawn with replacement from
-the residuals. We do this many times and report the average over draws
-of the residuals. The shaded region is a point-wise 90\% confidence
-interval. The left column shows that the fit of the estimated and
-observed growth rate is quite good.
-
-The middle column of Figure \ref{fig:WA-mask} shows the change in
-growth rate from mandating masks on April 1st. The shaded region is a
-90\% pointwise confidence interval. As shown, mandating masks on April
-1st lowers the growth of cases or deaths 14 or 21 days later by 0.1 to
-0.15. This effect then gradually declines due to information
-feedback. Mandatory masks reduce past cases or deaths, which leads to
-less cautious behavior, attenuating the impact of the policy. The
-reversal of the decrease in growth in late April is due to our
-comparison of a mask mandate on April 1st with Washington's actual
-mask mandate in early May. By late April, the counterfactual mask
-effect has decayed through information feedback, and we are comparing
-it the undecayed impact of Washington's actual, later mask mandate.
-
-The right column of Figure \ref{fig:WA-mask} shows how the changes in
-case and death growth translate into changes in cases and deaths. The
-estimates imply that mandating masks on April 1st would have led to
-500 fewer cases and 250 fewer deaths in Washington by the start of
-June.
-
-\begin{figure}[ht]
-  \caption{Effect of nationally mandating masks for employees on April
-    1st in the US\label{fig:US-mask}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \medskip
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-    \begin{flushleft}
-      \footnotesize In the left column, the dots are the average
-      change in growth in each state. The blue line is the average
-      across states of the change in growth. The shaded region is a
-      point-wise 90\% confidence interval. The right column shows the
-      change in cases or deaths relative to the baseline of actual
-      policies.
-    \end{flushleft}
-\end{figure}
-
-The results for other states are similar to those for Washington. In
-the appendix, Figures  \ref{fig:MA-mask} and \ref{fig:IL-mask} display
-similar results for Massachusetts and Illinois.
-Figure \ref{fig:US-mask} shows the average change in cases and
-deaths across states, where the top panel shows the effect on cases and the bottom panel shows the effect on deaths.
-The point estimates indicate that mandating masks on April 1st
-could have led to 25\% fewer cumulative cases and 37\% fewer
-cumulative deaths by the end of May with their 90 percent intervals given by $[10,47]$\% and $[18,55]$\%, respectively. The result roughly translates into $18$ to $55$ thousand saved lives.
-
-\subsection{Non-essential Business Closures}
-
-A particularly controversial policy is the closure of non-essential
-businesses. We now examine a counterfactual where non-essential
-businesses are never closed. Figure \ref{fig:WA-nb} shows the effect
-of leaving non-essential businesses open in Washington. The point
-estimate implies that the closure of non-essential businesses reduced cases and
-deaths by a small amount. However, this estimate is relatively
-imprecise; 90\% confidence intervals for the change in cases and deaths from
-leaving non-essential businesses open by the end of May are [-250,700] and
-[-100,1200], respectively.
-
-\begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open in Washington \label{fig:WA-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_v1}\\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_deaths_v1}
-      & \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-Figure \ref{fig:US-nb} shows the national effect of leaving non
-essential businesses open on cases and deaths. For cases, the
-estimates imply that with non-essential businesses open, cases would
-be about -15 to 60\% higher in late May. The results for deaths are
-similar but less precise.
-
-\begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open in the US\label{fig:US-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\subsection{Stay-at-home orders}
-
-We next examine a counterfactual where stay-at-home orders had been never
-issued.  Figure \ref{fig:US-shelter} shows the average effect of no
-stay-at-home orders. On average, without stay-at-home orders, case
-growth rate would have been nearly 0.1 higher in late April. This
-translates to 80\% [25\%,170\%] more cases by the start of June. The
-results for deaths are similar, but slightly less precise, with no
-increase included in a 90 percent confidence interval.
-
-\begin{figure}[ht]
-  \caption{Effect of having no stay-at-home orders in the US\label{fig:US-shelter}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\section{Counterfactual Effect of Removing All Policies and its Sensitivity}\label{removing-policies}
-
-%We now consider the impact of changing from the observed policies to
-%none.  Figure \ref{fig:WA-nb} shows the effect of removing all
-%policies in Washington. The left column shows that without any policy
-%intervention, case and death growth would have been higher by
-%about 0.1 and 0.2. The confidence intervals are wide relative to the
-%mean effect. In the right column, we see that the upper bounds of the
-%confidence intervals include large increases in cases, and especially
-%deaths.
-
-%\begin{figure}[ht]
-%  \caption{Effect of removing all policies in
-%    Washington \label{fig:WA-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dgrowth}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dcases}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dgrowth_deaths}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dcases_deaths}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-
-We now consider the impact of changing from the observed policies to none.
-Figure \ref{fig:US-nop-SI} shows the average across states of the change
-in case growth and relative increase in cases under a specification without past national case variables.
-Removing policies leads to an increase of above 0.2 in
-case growth throughout April and May. The confidence interval is
-fairly wide, and its upper bound includes a very large increase in
-cases by the end of May.  The right panel displays the national
-increase in aggregate cases without any policy intervention. The
-estimates imply at least a 7 fold increase in cases with a large upper bound by the end of May, or
-at least 14 million additional cases. The estimated impact on deaths is larger than
-cases, and even more imprecise.
-
-The effect of removing all policies includes the effect of school
-closures. The visual evidence on growth rates for states with and without school closures,
-presented blow, suggest that there may be a potentially large effect, though the history is very short.
-The main results presented in Section 3 also support the hypothesis that the school closures were important at lowering the growth rates.
-This evidence is consistent
-with the emerging evidence of prevalence of Covid-19 among children
-aged 10-17. \cite{children:nature} find that although children's
-transmission and susceptibility rates are half that of ages 20-30,
-children's contact rates are much higher.
-This type of evidence, as well as, evidence that children carry viral
-loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.
-
-\begin{figure}
-  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
-    \end{tabular}
-    \footnotetext{ In these figures, red points are the case or death
-      growth rate in states without each policy 14 (or 21 for deaths)
-      days earlier. Blue points are states with each policy 14 (or 21
-      for deaths) days earlier. The red line is the average across
-      states without each policy. The blue line is the average across
-      states with each policy.}
-  \end{minipage}
-\end{figure}
-
-As discussed above, there is little variation across states
-in the timing of school closures. Consequently, the effect of school
-closures is difficult to identify statistically, because it is hard to separate it from aggregate time effect, and its estimate is sensitive to an inclusion of some aggregate variables such as national cases.
-To support this point, Figure \ref{fig:US-nop} shows the
-effect of removing all policies on cases based on the estimates with national cases
-included as information. When national case variables are included
-in the specification, the estimated effect of school closures, and hence that of removing all
-policies, is much smaller with a 90\% confidence interval of [0,10] fold increases.
-
-Given this sensitivity, we conclude that there still exists a lot of uncertainty as to the effect of removing all policies, especially schooling.
-The impact of not implementing any policies on cases and deaths can be quite large, but  the effect of school closures, hence that of removing all policies, is not well identified statistically from the US state-level data alone, because of the lack of cross-sectional variations.
-Any analyses of re-opening plans need to be aware of this uncertainty.
-An important research question is how to resolve this uncertainty using additional data sources.
-
-\begin{figure}[ht]
-  \caption{Effect of removing policies on cases in the US under a specification with only state-level cases/deaths as information \label{fig:US-nop-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
-  \caption{Effect of removing policies on cases in the US under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nop}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
-
-\afterpage{\clearpage}
-
-\section{Conclusion}
-
-This paper assesses the effects of policies on the spread of Covid-19 in the US using state-level data on cases, tests, policies, and social distancing behavior measures from Google Mobility Reports.
-Our findings are summarized as follows.
-
-First, our empirical analysis indicates that mandating face masks has reduced the spread of Covid-19 without affecting people's social distancing behavior measured by Google Mobility Reports.
-Our counterfactual experiment based on the estimated model suggests that if all states had have adopted mandatory face mask policies on April 1st of 2020, then the number of deaths by the end of May would have been smaller by as much as  $17$ to $55$\%, which roughly translates to $17$ to $55$ thousand saved lives.
-
-Second, we find that keeping non-essential businesses open would have led to -20 to 60\% more cases  while not implementing stay-at-home orders would have increased cases by 25 to 170 \% by the start of June.
-% Except for mask mandates, the effect of all other policies on case and death growth is realized through their impact on social distancing behavior.
-
-Third, we find considerable uncertainty over  the impact of all policies combined on case or death growth because it is difficult to identify the effect of school closures from the US state-level data due to the lack of variation in the timing of school closures across states.
-
-Fourth, our analysis shows that people voluntarily reduce their visits to workplace, retails, grocery stores, and limit their use of public transit when they receive  information on a higher number of new cases and deaths.
-This suggests that individuals make decisions to voluntarily limit their contact with others in response to greater transmission risks, leading to an important feedback mechanism that affects future cases and deaths.
-Model simulations that ignore this voluntary behavioral response to information on transmission risks would over-predict the future number of cases and deaths.
-
-Beyond these findings, our paper presents a useful conceptual framework to investigate the relative roles of policies and information on determining the spread of Covid-19 through their impact on people's behavior.
-Our causal model allows us  to explicitly define counterfactual scenarios to properly evaluate the effect of alternative policies on the spread of Covid-19.
-%, and can be used to quantitatively examine various Covid-19 related issues  for future research.
- More broadly, our  causal framework can be useful for quantitatively analyzing not only health outcomes but also various economic outcomes \citep{bartik2020, chetty2020real}.
-
-\FloatBarrier
-
-\begin{footnotesize}
-
-\bibliographystyle{jpe}
-\bibliography{covid}
-
-\end{footnotesize}
-
-\newpage
-
-\appendix
-
-\section{Data Construction}
-
-\subsection{Measuring $\Delta C$ and $\Delta\log \Delta C$}
-
-We have three data sets with information on daily cumulative confirmed
-cases in each state. As shown in Table \ref{tab:casecor}, these
-cumulative case numbers are very highly correlated. However, Table
-\ref{tab:casediff} shows that the numbers are different more often
-than not.
-
-\input{tables_and_figures/casecor.tex}
-
-\input{tables_and_figures/casediff.tex}
-
-Figure \ref{fig:dailycases} shows the evolution of new cases in each
-of these three datasets. In all cases, daily changes in cumulative
-cases displays some excessive volatility. This is likely due to delays
-and bunching in testing and reporting of results. Table
-\ref{tab:newcasecor} shows the variance of log new cases in each data
-set, as well as their correlations. As shown, the correlations are
-approximately $0.9$. The NYT new case
-numbers have the lowest variance.\footnote{This comparison is somewhat
-  sensitive to how you handle negative and zero cases when taking
-  logs. Here, we replaced $\log(0)$ with $-1$. In our main
-  results, we work with weekly new cases, which are very rarely zero.}
-In our subsequent results, we will primarily use the case numbers from
-The New York Times.
-
-\begin{figure}[!ht]\caption{Daily cases \label{fig:dailycases}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/newcases}
-    \footnotetext{Each line shows daily new cases in a state.}
-  \end{minipage}
-\end{figure}
-
-\input{tables_and_figures/newcasecor.tex}
-
-For most of our results, we focus on new cases in a week instead of in
-a day. We do this for two reasons as discussed in the main text. First, a decline of new cases over
-two weeks has become a key metric for decision makers. Secondly, aggregating to weekly new cases smooths out the noise associated with  the timing of
-reporting and testing.
-
-Table \ref{tab:weekcasecor} reports the correlation and variance of
-weekly log new cases across the three data sets. Figure
-\ref{fig:weekcases} shows the evolution of weekly new cases in each
-state over time.
-
-\input{tables_and_figures/weekcasecor}
-
-\begin{figure}[!ht]\caption{Weekly Cases \label{fig:weekcases}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekcases}
-    \footnotetext{Each line shows weekly new cases in a state.}
-  \end{minipage}
-\end{figure}
-
-\subsection{Deaths}
-
-Table \ref{tab:weekdeathcor} reports the correlation and variance of
-weekly deaths in the three data sets. Figure \ref{fig:weekdeaths}
-shows the evolution of weekly deaths in each state. As with cases, we
-use death data from The New York Times in our main results.
-
-\input{tables_and_figures/weekdeathcor}
-
-\begin{figure}[!ht]\caption{Weekly Deaths \label{fig:weekdeaths}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekdeaths}
-    \footnotetext{Each line shows weekly deaths in a state.}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[!ht]\caption{Case and death growth \label{fig:growthq}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \begin{tabular}{c}
-      \includegraphics[width=0.75\textwidth]{tables_and_figures/casequantiles}
-      \\
-      \includegraphics[width=0.75\textwidth]{tables_and_figures/deathquantiles}
-      \footnotetext{Thin gray lines are case or death growth in each
-      state and date. Thicker colored lines are quantiles of case or
-      death growth conditional on date.}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\subsection{Tests}
-
-Our test data comes from The Covid Tracking Project. Figure
-\ref{fig:test} shows the evolution of tests over time.
-
-\begin{figure}[!ht]\caption{Number of Tests \label{fig:test}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/test}
-    \footnotetext{These figures use the ``total test results''
-      reported by The Covid Tracking Project. This is meant to reflect
-      the number of people tested
-      (\href{https://covidtracking.com/about-data/faq}{as opposed to
-        the number of specimens tested}).}
-  \end{minipage}
-\end{figure}
-
-\FloatBarrier
-
-\subsection{Social Distancing Measures}
-
-In measuring social distancing, we focus on Google Mobility
-Reports. This data has international coverage and is publicly
-available. Figure \ref{fig:gmr} shows the evolution of the four Google
-Mobility Reports variables that we use in our analysis.
-
-%% The following is true but we haven't included any of the
-%% results. We can add if needed.
-% We also compare Google Mobility Reports to other social
-% distancing measures from SafeGraph, Unacast, and PlaceIQ. Each of
-% these other data sources are limited to the United States. SafeGraph
-% and Unacast data is not publicly available (both companies make the
-% data available to researchers at no cost).
-\begin{figure}[ht]\caption{Evolution of Google Mobility Reports \label{fig:gmr}}
-  \begin{minipage}{\linewidth}
-    \begin{tabular}{cc}
-      \includegraphics[width=0.5\linewidth]{tables_and_figures/workplaces}
-      &
-      \includegraphics[width=0.5\linewidth]{tables_and_figures/retail}
-      \\
-      \includegraphics[width=0.5\linewidth]{tables_and_figures/grocery}
-      &
-      \includegraphics[width=0.5\linewidth]{tables_and_figures/transit}
-    \end{tabular}
-    \footnotetext{This figure shows the evolution of Google Mobility
-      Reports over time. Thin gray lines are the value of the
-      variables in each state and date. Thicker colored lines are
-      quantiles of the variables conditional on date.}
-  \end{minipage}
-\end{figure}
-
-\subsection{Policy Variables}
-
-We use the database on US state policies created by
-\cite{raifman2020}. %This database records that each state implemented 25 Covid-19 related policies.
-As discussed in the main text, our analysis focuses on seven policies. For stay-at-home orders, closed nonessential
-businesses, closed K-12 schools, closed restaurants except takeout,
-and closed movie theaters, we double-checked any
-state for which \cite{raifman2020} does not record a date. We filled
-in a few missing dates. Our modified data is available
-\href{"https://docs.google.com/spreadsheets/d/1E6HRkgbdSnZ9ZxrneydU6q4hhOCCt9oTl_5fa3OFVZE/edit?usp=sharing}{here}. Our
-modifications fill in 1 value for school closures, 2 for stay-at-home orders, 3 for movie theater closure, and 4 for
-non-essential business closures. Table \ref{tab:policies} displays all 25 dated policy variables in
-\cite{raifman2020}'s database with our modifications described above.
-
-\afterpage{%
-  \clearpage% Flush earlier floats (otherwise order might not be correct)
-  \thispagestyle{empty}% empty page style (?)
-  \begin{landscape}% Landscape page
-    \centering % Center table
-    \input{tables_and_figures/policies}
-  \end{landscape}
-  \clearpage% Flush page
-}
-
-\begin{figure}
-  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-deaths}
-      \\
-    \end{tabular}
-    \footnotetext{ In these figures, red points are the case or death
-      growth rate in states without each policy 14 (or 21 for deaths)
-      days earlier. Blue points are states with each policy 14 (or 21
-      for deaths) days earlier. The red line is the average across
-      states without each policy. The blue line is the average across
-      states with each policy.}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}
-  \caption{Case and death growth conditional on policies \label{fig:growthpolicies2}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-deaths}
-    \end{tabular}
-    \footnotetext{ In these figures, red points are the case or death
-      growth rate in states without each policy 14 (or 21 for deaths)
-      days earlier.  Blue points are states with each policy14 (or 21
-      for deaths) days earlier.  The red line is the average across
-      states without each policy. The blue line is the average across
-      states with each policy.}
-  \end{minipage}
-\end{figure}
-
-\subsection{Timing\label{sec:timing}}
-
-There is a delay between infection and when a person is tested and
-appears in our case data. \cite{midas2020} maintain a list of
-estimates of the duration of various stages of Covid-19
-infections. The incubation period, the time from infection to symptom
-onset, is widely believed to be 5 days. For example, using data from
-Wuhan, \cite{li2020} estimate a mean incubation period of 5.2 days.
-\cite{siorda2020} reviews the literature and concludes the mean
-incubation period is 3-9 days.
-
-Estimates of the time between symptom onset and case reporting or death
-are less common. Using Italian data, \cite{cereda2020} estimate an
-average of 7.3 days between symptom onset and
-reporting. \cite{zhang2020} find an average of 7.4 days using Chinese data
-from December to early February, but they find this period declined
-from 8.9 days in January to 5.4 days in the first week of
-February. Both of these papers on time from symptom onset to reporting
-have large confidence intervals covering approximately 1 to 20 days.
-
-Studying publicly available data on infected persons diagnosed outside of Wuhan,
-\cite{linton2020}  estimate an average of 15 days  from  onset to death.
-Similarly, using publicly available reports of 140 confirmed Covid-19 cases in China, mostly outside Hubei Province, \cite{sanche2020} estimate the time from onset to death to be  16.1 days.
-
-Based on the above, we expect a delay of roughly two weeks between
-changes in behavior or policies, and changes in reported
-cases while a corresponding delay of roughly three weeks for deaths.
-
-% Our growth rate variable,
-% $$
-% {\ycolor Y_{it}} = \Delta \log \Delta C_{t} = \log(C_{t} - C_{t-7}) - \log(C_{t-7}-C_{t-14})
-% $$
-% reflects an average growth rate of cases from time $t-7$ to $t$.\footnote{This
-% can be seen applying the intermediate value theorem to the dynamics
-% of the SIR model from section \label{sec:sirmodel}. Specifically,
-% \begin{align*}
-%   C_{t} - C_{t-7} = & \int_{t-7}^t \dot{C}(s) ds \\
-%   = & \int_{t-7}^t \tau(s) I(s) ds \\
-%   = & \overline{\tau}(t-7,t) I(t-7) \int_{t-7}^t \exp(\int_{t-7}^s \frac{S(r)}{N}
-%       \beta(r) - \gamma(r) dr \right) ds \\
-%   = & \overline{\tau}(t-7,t) \overline{\exp(\frac{S}{N} \beta - \gamma)}(t-7,t) I(t-7) \\
-%   = & \overline{\tau}(t-7,t) \overline{\exp(\frac{S}{N} \beta -
-%       \gamma)}(t-7,t)
-%       \overline{\exp(\frac{S}{N} \beta - \gamma)}(t-14,t-7) I(t-14)
-% \end{align*}
-% where $\gamma(r) = \frac{S(r-\ell)}{N}\beta(r-\ell)$ and expressions
-% with a line over them are constants that exist by the intermediate
-% value theorem. $\overline{\exp(\frac{S}{N} \beta - \gamma)}(t-7,t)$ is
-% a constant growth rate from $t-7$ to $t$ that gives rise to the same
-% number of infections at time $t$ as the original, potentially time
-% varying, growth rate. Applying similar reasoning to
-% $C_{t-7} - C_{t-14}$ and combining gives
-% \begin{align*}
-%   {\ycolor Y_{it}} = \log(C_{t} - C_{t-7}) - \log(C_{t-7}-C_{t-14})
-%   = & \overline{\exp(\frac{S}{N} \beta -  \gamma)}(t-7,t) +
-%       \log(\overline{\tau}(t-7,t)) -  \log(\overline{\tau}(t-14,t-7)).
-% \end{align*}
-% }
-
-\FloatBarrier
-
-\subsection{Counterfactuals for Massachusetts and
-  Illinois}
- Figures \ref{fig:MA-mask} and \ref{fig:IL-mask} present the fit of estimated cases as well as the counterfactual effect of mandating masks on April 1st in Massachusetts and Illinois, respectively.
-Figures \ref{fig:MA-nb} and \ref{fig:IL-nb}   show the counterfactual effect of leaving non-essential business open in Massachusetts and Illinois, respectively.
-
-\FloatBarrier
-
-\begin{figure}[b]
-  \caption{Effect of mandating masks on April 1st in Massachusetts \label{fig:MA-mask}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_v1}
-\\
-      \multicolumn{3}{c}{\textbf{Deaths}} \\
-     \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_deaths_v1}
-    \end{tabular}
-    \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff}. We set initial
-      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-      observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
-      point-wise 90\% confidence interval.
-    \end{flushleft}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \caption{Effect of mandating masks on April 1st in Illinois \label{fig:IL-mask}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_v1}      \\
-      \multicolumn{3}{c}{\textbf{Deaths}} \\
-    \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_deaths_v1}
-    \end{tabular}
-    \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff}. We set initial
-      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-      observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
-      point-wise 90\% confidence interval.
-    \end{flushleft}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \caption{Effect of leaving businesses open in Massachusetts \label{fig:MA-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_v1}
-       &
-         \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_v1}
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_deaths_v1}
-       &      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \caption{Effect of leaving businesses open in Illinois \label{fig:IL-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_v1} \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-           \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_deaths_v1}
-      &       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_deaths_v1}    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-%\begin{figure}[h]
-%  \caption{Effect of removing all policies in
-%    Illinois \label{fig:IL-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dgrowth_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dcases_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dgrowth_deaths_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dcases_deaths_v1}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-%
-%\begin{figure}[h]
-%  \caption{Effect of removing all policies in
-%    Massachusetts \label{fig:MA-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dgrowth_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dcases_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dgrowth_deaths_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dcases_deaths_v1}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-
-\FloatBarrier
-
-%\subsection{Policy Effects without National Cases as Information Variables}
-%
-%Table \ref{tab:dieff-si} presents the estimates under a specification that excludes national information variables.
-Comparing the estimates in table \ref{tab:dieff}, the effects of school closures and  the sum of policies are estimated substantially larger in table \ref{tab:dieff-si} when national information variables are excluded.
-%
-%%Here we present the results of counterfactual analyses that don't include the national cases/deaths as
-%%the the information variables.  The growth in the national cases nearly perfectly predicts school closure
-%%policies, which have little cross-sectional variation.  Here we remove the national cases from the information
-%%and check what happens to the counterfactual results.  The counterfactual results of mask policies, shelter-in-place,
-%%and closing
-%%non-essential businesses remain robust. The results on school closures effects change substantially,
-%%attributing a much bigger role to school closures. The resulting counterfactual of  removing
-%%all policies has a much larger effect on cases, with the lower bound of the confidence band implying at
-%%least a 7 fold increase in cases (see Figure \ref{fig:US-nop-SI} given in the main text).
-%
-%\begin{table}[!b]
-%  \caption{\label{tab:dieff-si}Direct and Indirect Policy Effects
-%    (without national information variables)}
-%\begin{minipage}{\linewidth}
-%  \centering
-%    \scriptsize
-%  \begin{tabular}{c}
-%    \textbf{Cases}
-%    \\
-%    \input{tables_and_figures/dieff-cases-SI}
-%    \\
-%    \textbf{Deaths}
-%    \\
-%    \input{tables_and_figures/dieff-deaths-deathsinfo-SI}
-%  \end{tabular}
-%  \smallskip
-%  \begin{flushleft}
-%    {\tiny Direct effects capture the effect of policy on case
-%      growth holding behavior, information, and confounders
-%      constant. Direct effects are given by ${\pcolor \pi}$ in
-%      equation (\ref{eq:R1}). Indirect effects capture how policy
-%      changes behavior and behavior shift case growth. They are given
-%      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
-%      from (\ref{eq:R2}). The total effect is
-%      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
-%      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
-%      \ref{eq:R4}. Columns ``Difference''    are  the differences between
-%      the estimates from (\ref{eq:R4}) and the combination of
-%      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
-%      Standard errors are computed by
-%      bootstrap and clustered on state.}
-%    %  \ref{tab:PtoY}.
-%    \end{flushleft}
-%  \end{minipage}
-%\end{table}
-
-\FloatBarrier
-
-\subsection{Counterfactuals with National Cases as Information
-  Variables}
-
-Figures \ref{fig:US-mask-SI}-\ref{fig:US-shelter-SI}  present the results of counterfactual analyses that include the national cases/deaths as the information variables.
-To create this figure, we repeat the same counterfactual
-simulation that we did for Washington with each state. For each state,
-we hold national cases constant, but endogenize state specific
-information. Thus, these figures should be interpreted as an average
-of state specific counterfactuals, and not a national counterfactual.
-
- The counterfactual results of mask policies, shelter-in-place,
- and closing non-essential businesses remain robust with respect to the inclusion of national case/death variables.
-This contrasts to the resulting counterfactual of  removing all policies  discussed in section \ref{removing-policies}.
-
-\begin{figure}[!b]
-  \caption{Effect of mandating masks for employees on April
-    1st
-     under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information
-    \label{fig:US-mask-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \medskip
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open      under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nb-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
-  \caption{Effect of having no stay-at-home orders     under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-shelter-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\FloatBarrier
-
-$\;$
-
 \end{document}

### 2020-08-03 cases_and_policies/rmd/rlassoPanel.R

@@ -1,0 +1,207 @@
+#' Rigorous Lasso for panel data
+#'
+#' Estimate the Lasso for panel data with fixed effects using the
+#' approach of Belloni, Chernozhukov, Hansen, & Kozbur (2016). Based
+#' on rlasso from the hdm package.
+#'
+#'
+#' @param x matrix of regressors
+#' @param y dependent variable
+#' @param f group identifier for fixed effects
+#' @param post Whether to use post lasso
+#' @param penalty list with options for the calculation of the penalty.
+#' \itemize{
+#' \item{\code{c} and \code{gamma}}{ constants for the penalty with default \code{c=1.1} and \code{gamma=0.1}}
+#' \item{\code{homoscedastic}}{ logical, if homoscedastic errors are considered (default \code{FALSE}). Option \code{none} is described below.}
+#' \item{\code{X.dependent.lambda}}{ logical,  \code{TRUE}, if the penalization parameter depends on the the design of the matrix \code{x}. \code{FALSE}, if independent of the design matrix  (default).}
+#' \item{\code{numSim}}{ number of simulations for the dependent methods, default=5000}
+#' \item{\code{lambda.start}}{ initial penalization value, compulsory for method "none"}
+#' }
+#' @param control list with control values.
+#' \code{numIter} number of iterations for the algorithm for
+#' the estimation of the variance and data-driven penalty, ie. loadings,
+#' \code{tol} tolerance for improvement of the estimated variances.
+#'\code{threshold} is applied to the final estimated lasso
+#' coefficients. Absolute values below the threshold are set to zero.
+#'
+#' @return
+rlassoPanel <-
+  function(x, y, f, post=TRUE,
+           penalty = list(homoscedastic=FALSE,
+                          X.dependent.lambda=FALSE,
+                          lambda.start = NULL, c = 1.1,
+                          gamma=.1/log(n)),
+           control = list(numIter = 15, tol = 10^-5, threshold =
+                            NULL))
+  {
+    x <- as.matrix(x)
+    y <- as.matrix(y)
+    
+    n <- dim(x)[1]
+    p <- dim(x)[2]
+    
[…]

### 2020-08-11 cases_and_policies/rmd/counterfactuals.Rmd

@@
 
-# ```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
-# if (!knitr::is_latex_output()) {
-#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-#     f <- cfplots(st, cfnop, yvar="deaths", data=df)
-#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-#   }
-# }
-```
-
-# ```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
-# states <- c("Washington","Illinois","Massachusetts")
-# for(st in states) {
-#   f <- cfplots(st, cfnop, yvar="deaths", data=df)
-# 
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figdp)
-#   dev.off()
-# 
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figd + ylim(c(-1,5)))
-#   dev.off()
-# 
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figl)
-#   dev.off()
-# 
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figp)
-#   dev.off()
-# }
-# ```
-
 ## National

### 2020-08-11 cases_and_policies/rmd/regressionsWithPlots.Rmd

@@
 
+scmat <-
 cmat <-  cor(df[,c(bvars,pols,"pindex")], use="pairwise.complete.obs")
-scmat <-
   matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)+1))
@@
 showhtmltables(NULL, NULL,  piy);
-prefix="deaths-fe"
-#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+colnames(piy[[1]]$reg$response) <- "dlogdd"
+showhtmltables(NULL, NULL,  piy);
+#savetextables(NULL, NULL, piy[1:4], prefix="deaths-fe")
+prefix <- "deaths-fe"
    ylbl <- colnames(piy[[1]]$reg$response)
@@
     cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
-
 ```
@@
 ## DML
+
 ```{r regsdml, cache=FALSE, warning=FALSE, results='asis'}
-
 library(glmnet)
@@
 DML2.RF <- DML2.for.PLM(x, d, y, dreg, yreg, sdf, nfold=10)
-
 ```

### 2020-08-11 cases_and_policies/rmd/counterfactuals.Rmd

--- 2020-07-23 cases_and_policies/rmd/counterfactuals.Rmd
@@
 
-```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
-if (!knitr::is_latex_output()) {
-  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-    f <- cfplots(st, cfnop, data=df)
-    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-  }
-}
-```
+# ```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
+# if (!knitr::is_latex_output()) {
+#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
+#     f <- cfplots(st, cfnop, data=df)
+#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
+#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
+#   }
+# }
+# ```
 
-```{r nop-state-figures, cache=TRUE, warning=FALSE}
-states <- c("Washington","Illinois","Massachusetts")
-for(st in states) {
-  f <- cfplots(st, cfnop, data=df)
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figdp)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figd + ylim(c(-1,5)))
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figl)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figp)
-  dev.off()
-}
-```
+# ```{r nop-state-figures, cache=TRUE, warning=FALSE}
+# states <- c("Washington","Illinois","Massachusetts")
+# for(st in states) {
+#   f <- cfplots(st, cfnop, data=df)
+#
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figdp)
+#   dev.off()
+#
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figd + ylim(c(-1,5)))
+#   dev.off()
+#
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figl)
+#   dev.off()
+#
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+#   print(f$figp)
+#   dev.off()
+# }
+# ```
 
@@
 
-```{r otherstates-deaths, cache=TRUE, fig.cap="Effect of Removing Policies"}
-if (!knitr::is_latex_output()) {
-  for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-    f <- cfplots(st, cfnop, yvar="deaths", data=df)
-    grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-    grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-  }
-}
-```
-
-```{r nop-state-figures-deaths, cache=TRUE, warning=FALSE}
-states <- c("Washington","Illinois","Massachusetts")
-for(st in states) {
-  f <- cfplots(st, cfnop, yvar="deaths", data=df)
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figdp)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figd + ylim(c(-1,5)))
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figl)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figp)
-  dev.off()
-}
-```
-
 ## National

### 2020-08-12 cases_and_policies/rmd/cf-pindex.Rmd
Added counterfactuals with pindex

--- 2020-08-11 cases_and_policies/rmd/counterfactuals.Rmd
@@
 title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
-subtitle    : "Counterfactuals"
-author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
+subtitle    : "Counterfactuals (with policies averaged)"
+author      :  Victor Chernozhukov,  Hiro Kasahara, Paul Schrimpf
 job         :
@@
 source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
+pols <- c("pmaskbus","pk12","pindex")
 ```
@@
 rerunsim=FALSE
-if (!file.exists("cfnop.Rda") || rerunsim) {
+if (!file.exists("cfnop_idx.Rda") || rerunsim) {
   cfnoplist <- lapply(unique(sdf$state), function(s)
@@
   allmeans <- rbindlist(lapply(cfnoplist, getmeans))
-  save(cfnoplist,allmeans, file="cfnop.Rda")
+  save(cfnoplist,allmeans, file="cfnop_idx.Rda")
 }
-load("cfnop.Rda")
+load("cfnop_idx.Rda")
 cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
@@
 #
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
 #   print(f$figdp)
@@
 #
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_idx.pdf",rootdir,st), width=5,height=3.75)
 #   print(f$figd + ylim(c(-1,5)))
@@
 #
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_idx.pdf",rootdir,st), width=5,height=3.75)
 #   print(f$figl)
@@
 #
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
 #   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_v1.pdf",sep="/"), width=5,height=3.75)
-figs$r
-dev.off()
-
-for (f in figs) print(f)
-```
-
-## Non-Essential Business Closures
-
-```{r business_sim, cache=FALSE}
-# remove non essential business closure
-cf <- sdf
-cf[,"pnonessential"] <- 0
-nobmodel <- createforwardsim(m, cf)
-
-if (!file.exists("cfnb.Rda") || rerunsim) {
-  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf))
-  nball <- rbindlist(lapply(cfnblist, getmeans))
-  save(cfnblist,nball, file="cfnb.Rda")
-}
-load("cfnb.Rda")
-cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
-```
-
-```{r, cache=TRUE}
-cfnb <- merge(cfnb, abbrv, by="state")
-for (st in c("Washington","Illinois","Massachusetts")) {
-  f <- cfplots(st, cfnb, data=df)
-  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
-  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
-  #textGrob(st,gp=gpar(fontsize=20,font=2)))
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figdp)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figp)
-  dev.off()
-}
-
-figs <- nationalplots(nball, cfnb,
-                      "never shutting down non-essential\nbusinesses")
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
-figs$g
-dev.off()
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_v1.pdf",sep="/"), width=5,height=3.75)
-figs$c
-dev.off()
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfmask.Rda") || rerunsim) {
+if (!file.exists("cfmask_idx.Rda") || rerunsim) {
   cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
-  save(cfmasklist, maskall, file="cfmask.Rda")
+  save(cfmasklist, maskall, file="cfmask_idx.Rda")
 }
-load("cfmask.Rda")
+load("cfmask_idx.Rda")
 cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 rerunsim <- FALSE
-if (!file.exists("cfmask1.Rda") || rerunsim) {
+if (!file.exists("cfmask1_idx.Rda") || rerunsim) {
   cfmasklist1 <- lapply(unique(df$state),
@@
   maskall1 <- rbindlist(lapply(cfmasklist1, getmeans))
-  save(cfmasklist1, maskall1, file="cfmask1.Rda")
+  save(cfmasklist1, maskall1, file="cfmask1_idx.Rda")
 }
-load("cfmask1.Rda")
+load("cfmask1_idx.Rda")
 cfmask1 <- do.call(rbind, lapply(cfmasklist1, function(x) x$edf))
@@
 
-if (!file.exists("cfshelter.Rda") || rerunsim) {
+if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
   cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfshelter.Rda")
+  save(cfshelterlist, shelterall, file="cfshelter_idx.Rda")
 }
-load("cfshelter.Rda")
+load("cfshelter_idx.Rda")
 cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 rerunsim=FALSE
-if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
+if (!file.exists("cfnop_deaths_idx.Rda") || rerunsim) {
   cfnoplist <- lapply(unique(df$state), function(s)
@@
   allmeans <- rbindlist(lapply(cfnoplist, getmeans))
-  save(cfnoplist,allmeans, file="cfnop_deaths.Rda")
+  save(cfnoplist,allmeans, file="cfnop_deaths_idx.Rda")
 }
-load("cfnop_deaths.Rda")
+load("cfnop_deaths_idx.Rda")
 cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
-figs$r
-dev.off()
-
-for (f in figs) print(f)
-```
-
-## Non Essential Business Closures
-
-```{r business_sim-deaths, cache=FALSE}
-# remove non essential business closure
-cf <- sdf
-cf[,"pnonessential"] <- 0
-nobmodel <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")
-
-if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
-  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
-  nball <- rbindlist(lapply(cfnblist, getmeans))
-  save(cfnblist,nball, file="cfnb_deaths.Rda")
-}
-load("cfnb_deaths.Rda")
-cfnb <- do.call(rbind, lapply(cfnblist, function(x) x$edf))
-```
-
-```{r, cache=TRUE}
-cfnb <- merge(cfnb, abbrv, by="state")
-for (st in c("Washington","Illinois","Massachusetts")) {
-  f <- cfplots(st, cfnb, yvar="deaths", data=df)
-  f$figp + ggtitle(sprintf("Effect of never shutting down non-essential businesses in %s",st))
-  #grid.arrange(f$figd + ggtitle(), f$figdp + ggtitle(), nrow=2, top =
-  #textGrob(st,gp=gpar(fontsize=20,font=2)))
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figdp)
-  dev.off()
-
-  figdev(sprintf("%s/tex/tables_and_figures/%s-nb-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
-  print(f$figp)
-  dev.off()
-}
-
-figs <- nationalplots(nball, cfnb,
-                      "never shutting down non-essential\nbusinesses", yvar="deaths")
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
-figs$g
-dev.off()
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
-figs$c
-dev.off()
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nb-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
+if (!file.exists("cfmask_deaths_idx.Rda") || rerunsim) {
   cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
-  save(cfmasklist, maskall, file="cfmask_deaths.Rda")
+  save(cfmasklist, maskall, file="cfmask_deaths_idx.Rda")
 }
-load("cfmask_deaths.Rda")
+load("cfmask_deaths_idx.Rda")
 cfmask <- do.call(rbind, lapply(cfmasklist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
+if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
   cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfshelter_deaths.Rda")
+  save(cfshelterlist, shelterall, file="cfshelter_deaths_idx.Rda")
 }
-load("cfshelter_deaths.Rda")
+load("cfshelter_deaths_idx.Rda")
 cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_v1.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_v1.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r

### 2020-08-12 cases_and_policies/rmd/cf-pindex.Rmd
Fix some formatting of rmd->html site

@@
 ```
-
-# ```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
-# if (!knitr::is_latex_output()) {
-#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-#     f <- cfplots(st, cfnop, data=df)
-#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-#   }
-# }
-# ```
-
-# ```{r nop-state-figures, cache=TRUE, warning=FALSE}
-# states <- c("Washington","Illinois","Massachusetts")
-# for(st in states) {
-#   f <- cfplots(st, cfnop, data=df)
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figdp)
-#   dev.off()
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_idx.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figd + ylim(c(-1,5)))
-#   dev.off()
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_idx.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figl)
-#   dev.off()
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figp)
-#   dev.off()
-# }
-# ```
 

### 2020-08-12 cases_and_policies/rmd/generatetables.R

@@ -163,11 +163,6 @@ statevarregexp <- gsub("\\)","\\\\)",gsub("\\(","\\\\(",statevars))
 statevarregexp <- paste("^(",paste(c("Constant",statevarregexp), collapse="|"),")",sep="")
 
 printstars <- function(est, se, starp=c(0.1, 0.05, 0.01)) {
-  foo()
-  print(est)
-  print(se)
-  print(starp)
-  print(sum(pnorm(-abs(est), sd=se)<starp/2))
   stars <- paste(rep("*", sum(pnorm(-abs(est), sd=se)<starp/2)), collapse="")
   if (length(stars)==0 || stars=="") return(sprintf("%.3f",est))
   return(sprintf("%.3f$^{%s}$", est, stars))

### 2020-08-12 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

--- 2020-08-01 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 library(reshape2)
+library(AER)
+library(hdm)
+library(randomForest)
+library(glmnet)
 
@@
 tvars <- "dlogtests"  
-xlist <-list("", "")
+# xlist <-list("party", "party") 
+# xlist <-list(c("party","logvote"), c("party","logvote")) 
+xlist <- list("","")
 interactions <-list("month", statevars)
@@
 
-L <- 14
+# Set time lag:
+# L.c <- 14
+# L.d <- 21
+L.c <- 9
+L.d <- 24
+
+L <- L.c
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
@@
 savetextables(c(regs$pib[[1]][1:4], regs$pib[[2]][1:4]), NULL, NULL, prefix="behavior-NI") 
-
 savetextables(c(regs$pib[[1]][1:4]), NULL, NULL, prefix="behavior-cases-stateinfo")
@@
 
+<!-- ## Deaths (with cases as information) -->
+
+<!-- ```{r deathtables, eval = FALSE, cache=FALSE, warning=FALSE, results='asis'} -->
+<!-- infovars <- list(c("dlogdc", "logdc"), -->
+<!--                  c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #, -->
+<!-- df$zero <- 0 -->
+<!-- L <- 21 -->
+<!-- sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) -->
+<!-- regs <- mainregressions(sdf, -->
+<!--                         "dlogdd", pols, bvars, infovars, "zero", xlist, ilist, iv, L=L) -->
+<!-- showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy); -->
+<!-- savetextables(NULL, regs$pbiy, regs$piy, prefix="deaths-casesinfo") -->
+<!-- ``` -->
+
 ## Deaths (with deaths as information)
@@
                   c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
-# infovarsd <- list(c("logdd")
-L <- 21
+L <- L.d
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
-#sdf <- subset(df, df$date>=as.Date("2020-02-29")) #+L)
 regs <- mainregressions(sdf,
@@
 
-<!-- ## $\log(\Delta D/\Delta C)$ -->
-
-<!-- ```{r logdddc, cache=FALSE, warning=FALSE, results='asis'} -->
-<!-- df$logdddc <- log(df$ddeath/panellag(df$cases, df$state, df$date, 7)) -->
-<!-- df$logdddc[df$ddeath<=0] <- NA -->
-<!-- df$logdddc[is.infinite(df$logdddc)] <- NA -->
-<!-- df$logtestrate <- log(df$testrate) -->
-<!-- df$logtestrate[df$testrate<=0] <- NA -->
-<!-- regs <- mainregressions(df, "logdddc", pols, bvars, infovars, "lag(logtestrate,7)", xlist[1], ilist[1], list("0"), L=21) -->
-
-<!-- showhtmltables(NULL, regs$pbiy, regs$piy); -->
-<!-- ``` -->
-
 ## Robustness Check
-```{r regsrobust, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsrobust,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 infovars <- list(c("dlogdc", "logdc"),
-                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
-#c("dlogdc", "logdc"))
-#infovars2 <- list(c( "logdc"))
+                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))  
 tvars <- "dlogtests"
@@
 xlist <-list(c("logvote"), c("logvote")) 
-L <- 14
+L <- L.c
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
@@
 
-L <- 14
+coef.vec.1 <- coef.vec.2 <- coef.vec.3 <- coef.vec.4 <-  rep(0.0, 40)
+se.vec.1 <- se.vec.2 <- se.vec.3 <- se.vec.4 <- rep(0.0, 40)
+
+# case growth robustness 
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
@@
 yvar <- "dlogdc"
+pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
 for (k in 1:length(infovars)) { 
-  # (1) policy index
-  pols <- c("pmaskbus","pk12","pindex")
+  # (1) baseline 
    xlist <-list("","")
@@
   # (2) excluding NY 
-  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
    xlist <-list("","")
@@
 
+for (k in 1:6) { 
+  for (j in 1:3) {
+  coef.vec.1[[10*(j-1)+k]] <- piy[[k]]$reg$coefficients[[1+j]]
+  se.vec.1[[10*(j-1)+k]] <- piy[[k]]$reg$se[[1+j]]
+  coef.vec.2[[10*(j-1)+k]] <- piy[[6+k]]$reg$coefficients[[1+j]]
+  se.vec.2[[10*(j-1)+k]] <- piy[[6+k]]$reg$se[[1+j]] 
+  }
+}
+
 colnames(piy[[1]]$reg$response) <- "dlogdc" 
@@
 
-L <- 21
-sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-piy <- list()
-yvar <- "dlogdd"
-infovars <- list(c("dlogdd", "logdd"),
-                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
-#xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+# case growth with pindex
+pols <- c("pmaskbus","pk12","pindex") 
 for (k in 1:length(infovars)) { 
-
-  # (1) policy index
-  pols <- c("pmaskbus","pk12","pindex")
+  # (1) baseline 
    xlist <-list("","")
@@
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                 xlist[[k]]),
+                tvars, xlist[[k]]),
               ilist[[k]], 0, L=L) 
   # (2) excluding NY 
-  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
    xlist <-list("","")
@@
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (3) adding mask wearing rate  
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("mask_percent"), c("mask_percent"))
+  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (4) adding log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("logvote"), c("logvote"))
+  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (5) lagged behavior variables
+  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
+  xlist <-list("","")
+  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (6) (2)-(5)
+   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
+   sdf2 <- sdf[which(sdf$state!='New York'),]
+  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
+   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L)
+}
+for (k in 1:6) {  
+  coef.vec.1[[30+k]] <- piy[[k]]$reg$coefficients[[4]]
+  se.vec.1[[30+k]] <- piy[[k]]$reg$se[[4]]
+  coef.vec.2[[30+k]] <- piy[[6+k]]$reg$coefficients[[4]]
+  se.vec.2[[30+k]] <- piy[[6+k]]$reg$se[[4]]  
+}
+
+# death growth
+L <- L.d 
+piy <- list()
+yvar <- "dlogdd"
+infovarsd <- list(c("dlogdd", "logdd"),
+                  c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
+pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+for (k in 1:length(infovars)) {  
+  # (1) baseline
+  xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
+  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
                  xlist[[k]]),
               ilist[[k]], 0, L=L)   
-
+  # (2) excluding NY  
+  xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
+  sdf2 <- sdf[which(sdf$state!='New York'),]
+  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L) 
   # (3) adding mask wearing rate 
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("z.mask"), c("z.mask"))
+  xlist <-list(c("mask_percent"), c("mask_percent"))
   piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
                  xlist[[k]]),
@@
   piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
                  xlist[[k]]),
@@
               c(sprintf("lag(%s, %d)", bvars, L+14),
-                sprintf("lag(%s, %d)", infovars[[k]], L),
+                sprintf("lag(%s, %d)", infovarsd[[k]], L),
                  xlist[[k]]),
@@
    sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
-   # sdf2 <- sdf[which(sdf$state!='New York' & sdf$state!='New Jersey' & sdf$state!='Connecticut' & sdf$state!='Massachusetts'),]
    sdf2 <- sdf[which(sdf$state!='New York'),]
-   xlist <-list(c("z.mask","logvote"), c("z.mask","logvote"))
+  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
    piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
-                sprintf("lag(%s, %d)", infovars[[k]], L),
+                sprintf("lag(%s, %d)", infovarsd[[k]], L),
                 xlist[[k]]),
               ilist[[k]], 0, L=L) 
+}
+for (k in 1:6) { 
+  for (j in 1:3) {
+  coef.vec.3[[10*(j-1)+k]] <- piy[[k]]$reg$coefficients[[1+j]]
+  se.vec.3[[10*(j-1)+k]] <- piy[[k]]$reg$se[[1+j]]
+  coef.vec.4[[10*(j-1)+k]] <- piy[[6+k]]$reg$coefficients[[1+j]]
+  se.vec.4[[10*(j-1)+k]] <- piy[[6+k]]$reg$se[[1+j]] 
+  }
+}
 
-}
 colnames(piy[[1]]$reg$response) <- "dlogdd"  
@@
 
+# death growth with pindex
+pols <- c("pmaskbus","pk12","pindex") 
+for (k in 1:length(infovars)) {  
+  # (1) baseline
+  xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
+  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)   
+  # (2) excluding NY  
+  xlist <-list("","")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
+  sdf2 <- sdf[which(sdf$state!='New York'),]
+  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L) 
+  # (3) adding mask wearing rate 
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("mask_percent"), c("mask_percent"))
+  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (4) adding log(Trump's vote)
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list(c("logvote"), c("logvote"))
+  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (5) lagged behavior variables
+  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
+  xlist <-list("","")
+  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)
+  # (6) (2)-(5)
+   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
+   sdf2 <- sdf[which(sdf$state!='New York'),]
+  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
+   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", bvars, L+14),
+                sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                xlist[[k]]),
+              ilist[[k]], 0, L=L) 
+}
+for (k in 1:6) {  
+  coef.vec.3[[30+k]] <- piy[[k]]$reg$coefficients[[4]]
+  se.vec.3[[30+k]] <- piy[[k]]$reg$se[[4]]
+  coef.vec.4[[30+k]] <- piy[[6+k]]$reg$coefficients[[4]]
+  se.vec.4[[30+k]] <- piy[[6+k]]$reg$se[[4]]  
+}
+
 ```
@@
 ## Fixed Effects
-```{r regsfe, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsfe,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 
@@
 df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
-#df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
 
@@
 tvars <- "dlogtests"  
-L <- 14
+L <- L.c
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
 piy <- list()
-
-#' Update the coefficients of felm object.  Also updates the t stats
-#' and p values, so that summary and stargazer work correctly after.
-replacecoef <- function(reg, newcoef) {
-  reg$beta[] <- newcoef
-  reg$coefficients[] <- newcoef
-  reg$tval <- reg$beta/reg$se
-  reg$pval <- pt(-abs(reg$tval),df=reg$df)*2
-  reg$rtval <- reg$beta/reg$rse
-  reg$rpval <- pt(-abs(reg$rtval),df=reg$df)*2
-  reg$ctval <- reg$beta/reg$cse
-  reg$cpval <- pt(-abs(reg$ctval),df=reg$df)*2
-  return(reg)
-}
-
 for (k in 1:length(infovars)) { 
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
+  pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
   m <-  reg_fe(sdf, yvar, pols, NULL,
@@
   piy[[(1+4*(k-1))]] <- m 
-  m$reg <- replacecoef(m$reg, m$bc2)
+  m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))]  
   piy[[(2+4*(k-1))]] <- m
@@
   piy[[(3+4*(k-1))]] <- m  
-  m$reg <- replacecoef(m$reg, m$bc2)
+  m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
   piy[[(4+4*(k-1))]] <- m 
@@
 
+for (j in 1:3) {
+  coef.vec.1[[10*(j-1)+9]] <- piy[[1]]$reg$beta[[j]]
+  se.vec.1[[10*(j-1)+9]] <- piy[[1]]$reg$se[[j]]
+  coef.vec.1[[10*(j-1)+10]] <- piy[[2]]$reg$beta[[j]]
+  se.vec.1[[10*(j-1)+10]] <- piy[[1]]$reg$se[[j]]
+  coef.vec.2[[10*(j-1)+9]] <- piy[[5]]$reg$beta[[j]]
+  se.vec.2[[10*(j-1)+9]] <- piy[[5]]$reg$se[[j]]
+  coef.vec.2[[10*(j-1)+10]] <- piy[[6]]$reg$beta[[j]]
+  se.vec.2[[10*(j-1)+10]] <- piy[[5]]$reg$se[[j]]
+}
+coef.vec.1[[30+9]] <- piy[[3]]$reg$beta[[3]]
+se.vec.1[[30+9]] <- piy[[3]]$reg$se[[3]]
+coef.vec.1[[30+10]] <- piy[[4]]$reg$beta[[3]]
+se.vec.1[[30+10]] <- piy[[3]]$reg$se[[3]]
+coef.vec.2[[30+9]] <- piy[[7]]$reg$beta[[3]]
+se.vec.2[[30+9]] <- piy[[7]]$reg$se[[3]]
+coef.vec.2[[30+10]] <- piy[[8]]$reg$beta[[3]]
+se.vec.2[[30+10]] <- piy[[7]]$reg$se[[3]]
+
+colnames(piy[[1]]$reg$response) <- "dlogdc"  
 showhtmltables(NULL, NULL,  piy);
 prefix="cases-fe"
-#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe") copying and
-
-# pasting this is bad style and seems like it might create maintenance
-# problems ...
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
    ylbl <- colnames(piy[[1]]$reg$response)
@@
                   c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
-L <- 21
+L <- L.d
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
@@
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  pols <- c("pmaskbus","pk12","pmovie","pshelter","prestaurant","pnonessential")
+  pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
   m <-  reg_fe(sdf, yvar, pols, NULL,
@@
   piy[[(1+4*(k-1))]] <- m
-  m$reg <- replacecoef(m$reg, m$bc2)
+  m$reg$beta[1:(8+2*(k-1))] <- m$bc2[1:(8+2*(k-1))]  
   piy[[(2+4*(k-1))]] <- m
@@
   piy[[(3+4*(k-1))]] <- m 
-  m$reg <- replacecoef(m$reg, m$bc2)
+  m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))]  
   piy[[(4+4*(k-1))]] <- m
@@
 
+for (j in 1:3) {
+  coef.vec.3[[10*(j-1)+9]] <- piy[[1]]$reg$beta[[j]]
+  se.vec.3[[10*(j-1)+9]] <- piy[[1]]$reg$se[[j]]
+  coef.vec.3[[10*(j-1)+10]] <- piy[[2]]$reg$beta[[j]]
+  se.vec.3[[10*(j-1)+10]] <- piy[[1]]$reg$se[[j]]
+  coef.vec.4[[10*(j-1)+9]] <- piy[[5]]$reg$beta[[j]]
+  se.vec.4[[10*(j-1)+9]] <- piy[[5]]$reg$se[[j]]
+  coef.vec.4[[10*(j-1)+10]] <- piy[[6]]$reg$beta[[j]]
+  se.vec.4[[10*(j-1)+10]] <- piy[[5]]$reg$se[[j]]
+}
+coef.vec.3[[30+9]] <- piy[[3]]$reg$beta[[3]]
+se.vec.3[[30+9]] <- piy[[3]]$reg$se[[3]]
+coef.vec.3[[30+10]] <- piy[[4]]$reg$beta[[3]]
+se.vec.3[[30+10]] <- piy[[3]]$reg$se[[3]]
+coef.vec.4[[30+9]] <- piy[[7]]$reg$beta[[3]]
+se.vec.4[[30+9]] <- piy[[7]]$reg$se[[3]]
+coef.vec.4[[30+10]] <- piy[[8]]$reg$beta[[3]]
+se.vec.4[[30+10]] <- piy[[7]]$reg$se[[3]]
+
 colnames(piy[[1]]$reg$response) <- "dlogdd"  
 showhtmltables(NULL, NULL,  piy); 
-#savetextables(NULL, NULL, piy[1:4], prefix="deaths-fe")
+prefix="deaths-fe"
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+   ylbl <- colnames(piy[[1]]$reg$response)
+    if (ylbl=="dlogdd") ylbl <- "Death Growth"
+    if (ylbl=="dlogdc") ylbl <- "Case Growth"
+
+    tbl <- capture.output(stargazer(
+        lapply(piy[1:4], function(x) x$reg),
+        type="latex",
+        dep.var.labels.include=FALSE,
+        title=ylbl,
+        # omit=omit,
+        # omit.labels=omit.labels,
+        column.labels=c(yvar),
+        column.separate=c(length(piy[1:4])),
+        # add.lines=list(c("$\\sum_j \\mathrm{Policy}_j$",
+        #                  sapply(piy, function(x) printstars(x$peff[1], x$peff[2]))),
+        #                c("",
+        #                  sprintf("(%.3f)",sapply(piy, function(x) x$peff[2])))
+        #                ),
+        omit.stat=c("f", "ser"), model.names=FALSE,
+        no.space=TRUE,
+        column.sep.width="1pt",
+        df=FALSE, header=FALSE,
+        model.numbers=TRUE))
+    tbl <- relabel(tbl)
+    tbl <- gsub("Model (\\d)", "\\(\\1\\)", tbl)
+    #tbl <- gsub("No","Yes", tbl)
+    texfile <- sprintf("%s/tex/tables_and_figures/%s-piy.tex",rootdir,prefix)
+    cat(paste(tbl[c(-1,-2,-3,-4, -length(tbl))], collapse="\n"), file=texfile)
+
+```
+
+## DML 
+```{r regsdml,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+
+df$logpop <- log(df$Population.2018)
+df$logsq <- log(df$Square.Miles) 
+df$repub <- as.factor(df$party)
+df$party_dum <- ifelse(df$party == "republican", 0, 1) 
+df$month3 <- ifelse(df$month==3,0,1)
+df$month4 <- ifelse(df$month==4,0,1)
+df$month5 <- ifelse(df$month==5,0,1)
+df$month6 <- ifelse(df$month==6,0,1)
+L <- L.c
+
+pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential")  
+infovars <- list(c("dlogdc", "logdc"),
+                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
+infovarsd <- list(c("dlogdd", "logdd"),
+                  c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
+vars <- c(infovars[[2]],infovarsd[[2]],pols,"pindex") 
+for(i in 1:length(vars)) {
+  L<-L.c
+  v <- sprintf("%s.L14",vars[i])
+  df[,v] <- panellag(df[,vars[i]],df$state,df$date, lag=L)
+  L<-L.d
+  v <- sprintf("%s.L21",vars[i])
+  df[,v] <- panellag(df[,vars[i]],df$state,df$date, lag=L)
+}
+
+for(i in 1:length(bvars)) {
+  L<-L.c
+  v <- sprintf("%s.L14",bvars[i])
+  df[,v] <- panellag(df[,bvars[i]],df$state,df$date, lag=L+14)
+  L<-L.d
+  v <- sprintf("%s.L21",bvars[i])
+  df[,v] <- panellag(df[,bvars[i]],df$state,df$date, lag=L+14)
+}
+
+df$test <- df$dlogtests
+df$test[is.na(df$test)] <- 0
+
+wvars <- c("logpop","logsq","Percent.Unemployed..2018.",
+               "Percent.living.under.the.federal.poverty.line..2018.",
+               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
+               "mask_percent", "logvote","test","month3","month4","month5","month6") 
+wvars.d <- c("logpop", "logsq","Percent.Unemployed..2018.",
+               "Percent.living.under.the.federal.poverty.line..2018.",
+               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
+               "mask_percent", "logvote","month3","month4","month5","month6")
+
+bvars.c <- c("workplaces.L14","retail.L14","grocery.L14","transit.L14") 
+bvars.d <- c("workplaces.L21","retail.L21","grocery.L21","transit.L21") 
+xvars <- c(bvars.c,"dlogdc.L14","logdc.L14")  
+xvars.national <- c(bvars.c,"dlogdc.L14","logdc.L14","dlogdc.national.L14","logdc.national.L14") 
+xvars.d <- c(bvars.d,"dlogdd.L21","logdd.L21")  
+xvars.national.d <- c(bvars.d,"dlogdd.L21","logdd.L21","dlogdd.national.L21","logdd.national.L21")
+
+pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pmovie.L14","prestaurant.L14","pnonessential.L14") 
+pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pmovie.L21","prestaurant.L21","pnonessential.L21")
+
+AM<- function(yv, xv, wv, sdf, reg,  steps=5){  
+ y <- sdf[,yv]
+ w <- sdf[,wv]
+ sdf[,"y"] <- y
+ # Initialize
+ fmla <- as.formula(paste("y", "~",  paste(xv, collapse=" + "), sep=" "))
+ g1 <- predict(lm(fmla, data=sdf)) 
+ m1 <- predict(reg(w,as.vector(y- g1)))
+ i<-1
+ while(i<= steps) {
+  sdf$y <- y-m1
+  g1 <- predict(lm(fmla, data=sdf))  
+  m1 <- predict(reg(w,as.vector(y-g1)))
+  i <- i+1 }  
+ fitx <- lm(fmla, data=sdf)
+ fitw <- reg(w,as.vector(y-g1))
+ return(list(x=fitx , w = fitw))
+}
+
+DML2.for.PLM <- function(xv, wv, dv, yv, reg, sdf, nfold=5) {   
+  unit <- as.double(sdf$state)
+  j <-  quantile(unit, c(1:(nfold-1))/nfold)
+  I <- vector(mode = "list", length = nfold)
+  I[[1]] <- unit <= j[1] 
+  for (b in 2:(nfold-1)) {
+    I[[b]] <- (unit <= j[b])&(unit > j[b-1])
+  } 
+  I[[nfold]] <- unit > j[nfold-1]  
+  nobs <- nrow(sdf)
+  ytil <- dtil <- rep(NA, nobs)
+  w <- as.matrix(sdf[,wv])
+  x <- as.matrix(sdf[,xv])
+  d <- as.vector(sdf[,dv])
+  y <- as.vector(sdf[,yv])
+  for(b in 1:nfold){
+    I1 <- (I[[b]]==FALSE)
+    I2 <- I[[b]]  
+    yfit <- AM(yv, xv, wv, sdf[I1,], reg)   
+    dfit <- AM(dv, xv, wv, sdf[I1,], reg)   
+    yhat <- predict(yfit$x,newdata=sdf[I2,]) + predict(yfit$w, w[I2,], type="response") 
+    dhat <- predict(dfit$x,newdata=sdf[I2,])+ predict(dfit$w, w[I2,], type="response")  
+    dtil[I2] <- (d[I2] - dhat) #record residual
+    ytil[I2] <- (y[I2] - yhat) #record residual
+    cat(b," ")
+  }  
+sdf$ytil <- ytil
+sdf$dtil <- dtil 
+rfit <- felm(ytil ~ dtil| 0| 0 | state, data=sdf) #estimate the main parameter by regressing one residual on the other
+coef.est <- coef(rfit)[2] #extract coefficient 
+se <- sqrt(vcov(rfit)[2,2]) 
+cat(sprintf("\ncoef (se) = %g (%g)\n", coef.est , se))
+return( list(coef.est = coef.est , se=se) )
+}
+
+# DML2.for.PLM <- function(x, d, y, dreg, yreg, sdf, nfold=5) {   
+#   unit <- as.double(sdf$state)
+#   j <-  quantile(unit, c(1:(nfold-1))/nfold)
+#   I <- vector(mode = "list", length = nfold)
+#   I[[1]] <- unit <= j[1] 
+#   for (b in 2:(nfold-1)) {
+#     I[[b]] <- (unit <= j[b])&(unit > j[b-1])
+#   } 
+#   I[[nfold]] <- unit > j[nfold-1]  
+#   nobs <- nrow(x)
+#   ytil <- dtil <- rep(NA, nobs)
+#   for(b in 1:nfold){
+#     I1 <- (I[[b]]==FALSE)
+#     I2 <- I[[b]]
+#     dfit <- dreg(x[I1,], d[I1]) # take a fold out
+#     yfit <- yreg(x[I1,], y[I1]) # take a fold out
+#     dhat <- predict(dfit, x[I2,], type="response") #predict the fold out
+#     yhat <- predict(yfit, x[I2,], type="response") #predict the fold out
+#     dtil[I2] <- (d[I2] - dhat) #record residual
+#     ytil[I2] <- (y[I2] - yhat) #record residual
+#     cat(b," ")
+#   }  
+# sdf$ytil <- ytil
+# sdf$dtil <- dtil 
+# rfit <- felm(ytil ~ dtil| 0| 0 | state, data=sdf) #estimate the main parameter by regressing one residual on the other
+# coef.est <- coef(rfit)[2] #extract coefficient 
+# se <- sqrt(vcov(rfit)[2,2]) 
+# cat(sprintf("\ncoef (se) = %g (%g)\n", coef.est , se))
+# return( list(coef.est = coef.est , se=se) )
+# }
+
+nfold <- 5
+
+penalty = list(homoscedastic = FALSE, X.dependent.lambda =
+  FALSE, lambda.start = NULL, c = 1.7, gamma = 0.1)
+reg.lasso <- function(x,y){ rlasso(x, y, penalty=penalty) } #ML method= rlasso 
+# dreg <- function(x,d){ cv.glmnet(x, d) } #ML method= rlasso
+# yreg <- function(x,y){ cv.glmnet(x, y) } #ML method = rlasso 
+reg.RF <- function(x,y){ randomForest(x, y, maxnodes=8) } #ML method=Forest 
+for (j in 1:4){  
+  sdf <- subset(df, df$date>=as.Date(sprintf("2020-03-%s",7+L.c))) 
+  # y = sdf$dlogdc 
+  yv <- "dlogdc" 
+  wv <- wvars         
+  if (j==1){ 
+    dv <- "pmaskbus.L14"
+    p <- pols.L14[-1]
+    # d = sdf$pmaskbus.L14
+  } else if (j==2) {
+    dv <- "pk12.L14"  
+    p <- pols.L14[-2]
+    # d = sdf$pk12.L14
+  } else if (j==3) {
+    dv <- "pshelter.L14" 
+    p <- pols.L14[-3]
+    # d = sdf$pshelter.L14 
+  } else if (j==4) {
+    dv <- "pindex.L14" 
+    p <- c("pmaskbus.L14","pk12.L14")
+    # d = sdf$pindex.L14 
+  } 
+  # case growth without national
+  # x = as.matrix(cbind(sdf[,p],sdf[,xvars]))  
+  xv <- c(xvars,p)        
+  set.seed(1)
+  DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
+  coef.vec.1[[10*(j-1)+7]] <- DML2.lasso$coef.est
+  se.vec.1[[10*(j-1)+7]] <- DML2.lasso$se
+  set.seed(1)
+  DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold=10) 
+  coef.vec.1[[10*(j-1)+8]] <- DML2.RF$coef.est
+  se.vec.1[[10*(j-1)+8]] <- DML2.RF$se 
+  # case growth with national
+  xv <- c(xvars.national,p)
+  # x = as.matrix(cbind(sdf[,p],sdf[,xvars.national]))  
+  set.seed(1)
+  DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
+  coef.vec.2[[10*(j-1)+7]] <- DML2.lasso$coef.est
+  se.vec.2[[10*(j-1)+7]] <- DML2.lasso$se 
+  set.seed(1)
+  DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold = nfold) 
+  coef.vec.2[[10*(j-1)+8]] <- DML2.RF$coef.est
+  se.vec.2[[10*(j-1)+8]] <- DML2.RF$se 
+  ## death growth
+  sdf <- subset(df, df$date>=as.Date(sprintf("2020-03-%s",7+L.d))) 
+  yv <- "dlogdd" 
+  wv <- wvars.d         
+  # y = sdf$dlogdd
+  # sdf <- sdf[which(sdf$state!='New York'),]
+    if (j==1){
+    dv <- "pmaskbus.L21"
+    p <- pols.L21[-1]
+    # d = sdf$pmaskbus.L21
+  } else if (j==2) {
+    dv <- "pk12.L21"
+    p <- pols.L21[-2]
+    # d = sdf$pk12.L21
+  } else if (j==3) {
+    dv <- "pshelter.L21"
+    p <- pols.L21[-3]
+    # d = sdf$pshelter.L21 
+  } else if (j==4) {
+    dv <- "pindex.L21"
+    p <- pols.L21[1:2]
+    # d = sdf$pindex.L21 
+  } 
+  # death growth without national
+  # x = as.matrix(cbind(sdf[,p],sdf[,xvars.d]))  
+  xv <- c(xvars.d,p)           
+  set.seed(1)
+  DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
+  coef.vec.3[[10*(j-1)+7]] <- DML2.lasso$coef.est
+  se.vec.3[[10*(j-1)+7]] <- DML2.lasso$se  
+  set.seed(1)
+  DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold = nfold)
+  coef.vec.3[[10*(j-1)+8]] <- DML2.RF$coef.est
+  se.vec.3[[10*(j-1)+8]] <- DML2.RF$se  
+  # death growth with national vars
+  # x = as.matrix(cbind(sdf[,p],sdf[,xvars.national.d]))    
+  xv <- c(xvars.national.d,p)           
+  set.seed(1)
+  DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
+  coef.vec.4[[10*(j-1)+7]] <- DML2.lasso$coef.est
+  se.vec.4[[10*(j-1)+7]] <- DML2.lasso$se  
+  set.seed(1)
+  DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold = nfold)
+  coef.vec.4[[10*(j-1)+8]] <- DML2.RF$coef.est
+  se.vec.4[[10*(j-1)+8]] <- DML2.RF$se 
+}
+
+# WARNINGS!  dplyr and dotwhisker seem to interfare with our regression codes --- need to call them only after calling regression codes; otherwise, they will improperly change our regression results 
+library(dplyr)
+library(dotwhisker) 
+var.names <- c("(1) baseline    ","(2) excluding NY","(3) mask wearing","(4) Trump's vote ","(5) behavior vars"," (6) all of (2)-(5)  ",  "(7) DML (Lasso)", "(8) DML (RF)  ", "(9) FE with BC")
+
+# mask coefficient
+ij <- c(1:8,10)
+ results_df <- data.frame(term = rep(var.names, times = 4),
+                         estimate = c(coef.vec.1[ij], coef.vec.2[ij],coef.vec.3[ij], coef.vec.4[ij]),
+                         std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
+                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
+                                   rep("death w/o national",9), rep("death wi/ national", 9)),
+                         stringsAsFactors = FALSE)
+ # Draw dot-and-whisker plot
+fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
+    theme(       legend.justification=c(.01, .999), legend.position=c(.63, .99), 
+           legend.title = element_blank(), legend.background =
+              element_rect(color="gray90")) + 
+    xlab("coefficients of masks for employees") +
+    geom_vline(xintercept = 0, colour = "grey60", linetype = 2) + 
+  theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
+   xlim(-0.55, 0.4) +theme(
+  legend.title = element_text(size = 6),
+  legend.text = element_text(size = 6)
+  )# + ggtitle("Mask Mandates")  
+p <- "pmaskbus"   
+figdev(sprintf("%s/tex/tables_and_figures/%s-whisker-%s.pdf", rootdir, p, L.c), width=4, height=3)
+print(fig)
+dev.off()
+fig
+
+# school coefficient
+ij <- c(11:18,20)
+ results_df <- data.frame(term = rep(var.names, times = 4),
+                         estimate = c(coef.vec.1[ij], coef.vec.2[ij],coef.vec.3[ij], coef.vec.4[ij]),
+                         std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
+                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
+                                   rep("death w/o national",9), rep("death w/ national", 9)),
+                         stringsAsFactors = FALSE) 
+fig <- dwplot(results_df, conf.level = .90) + theme_bw() +  
+    theme(
+       legend.justification=c(.01, .999), legend.position=c(.01,.99), 
+    # legend.position = c(.95, .95),
+    # legend.justification = c("right", "top"),
+    # legend.box.just = "right", 
+           legend.title = element_blank(), legend.background =
+              element_rect(color="gray90")) + 
+    xlab("coefficients of closed K-12 schools") +
+    geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
+  theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
+  xlim(-1.4, 0.5)  +theme(
+  legend.title = element_text(size = 6),
+  legend.text = element_text(size = 5)
+  )# + ggtitle("School Closures")  
+p <- "pk12" 
+figdev(sprintf("%s/tex/tables_and_figures/%s-whisker-%s.pdf", rootdir, p, L.c), width=4, height=3)
+print(fig)
+dev.off()
+fig
+
+# stay-at-home coefficient
+ij <- c(21:28,30)
+ results_df <- data.frame(term = rep(var.names, times = 4),
+                         estimate = c(coef.vec.1[ij], coef.vec.2[ij],coef.vec.3[ij], coef.vec.4[ij]),
+                         std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
+                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
+                                   rep("death w/o national",9), rep("death w/ national", 9)),
+                         stringsAsFactors = FALSE) 
+fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
+    theme(       legend.justification=c(.01, .999), legend.position=c(.63, .99), 
+           legend.title = element_blank(), legend.background =
+              element_rect(color="gray90")) + 
+    xlab("coefficients of stay-at-home") +
+    geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
+  theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
+   xlim(-0.55, 0.4) +theme(
+  legend.title = element_text(size = 6),
+  legend.text = element_text(size = 6)
+  )
+# + ggtitle("Stay-at-home Orders")  
+p <- "pshelter" 
+figdev(sprintf("%s/tex/tables_and_figures/%s-whisker-%s.pdf", rootdir, p, L.c), width=4, height=3)
+print(fig)
+dev.off()
+fig
+
+# pindex coefficient
+ij <- c(31:38,40)
+ results_df <- data.frame(term = rep(var.names, times = 4),
+                         estimate = c(coef.vec.1[ij], coef.vec.2[ij],coef.vec.3[ij], coef.vec.4[ij]),
+                         std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
+                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
+                                   rep("death w/o national",9), rep("death w/ national", 9)),
+                         stringsAsFactors = FALSE) 
+fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
+    theme(       legend.justification=c(.01, .999), legend.position=c(.67, .99), 
+           legend.title = element_blank(), legend.background =
+              element_rect(color="gray90")) + 
+    xlab("coefficients of average of four policy variables") +
+    geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
+  theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
+  xlim(-0.55, 0.4) +theme(
+  legend.title = element_text(size = 6),
+  legend.text = element_text(size = 5)
+  )# + ggtitle("Average of Four Policy Variables")  
+p <- "pindex" 
+figdev(sprintf("%s/tex/tables_and_figures/%s-whisker-%s.pdf", rootdir, p, L.c), width=4, height=3)
+print(fig)
+dev.off()
+fig
 
@@
     xlab("") +
-    ylim(c(-1.2,1.2))
+    ylim(c(-0.65,0.4))
   if (!is.null(ldf))
@@
 }
-fig <- pfigure(df, "pmaskbus", stlab=-1) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("dlogdc")))
+fig <- pfigure(df, "pmaskbus", stlab=-1,L=L.c) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("dlogdc")))
 
 p <- "pmaskbus"
-figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
+figdev(sprintf("%s/tex/tables_and_figures/%s-cases-%s.pdf", rootdir, p, L.c), width=4, height=3)
 print(fig)
@@
 ```{r polplots, cache=FALSE, fig_width=10, fig_height=6}
-fig <- pfigure(df, "pk12", stlab=-1) +
+fig <- pfigure(df, "pk12", stlab=-1,L=L.c) +
   ggtitle(TeX(relabel("dlogdc given pk12"))) +
-  xlim(c(as.Date("2020-03-15"), max(df$date))) +
-  ylim(c(-1,3)) + ylab(TeX(relabel("dlogdc")))
+   xlim(c(as.Date("2020-03-21"), as.Date("2020-05-01"))) +
+  ylim(c(-0.6,2)) + ylab(TeX(relabel("dlogdc")))
 
 p <- "pk12"
-figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
+figdev(sprintf("%s/tex/tables_and_figures/%s-cases-%s.pdf", rootdir, p, L.c), width=4, height=3)
 print(fig)
@@
     ylim(c(-1,1)) + ylab(TeX(relabel("dlogdc")))
-  figdev(sprintf("%s/tex/tables_and_figures/%s-cases.pdf", rootdir, p), width=4, height=3)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-cases-%s.pdf", rootdir, p, L.c), width=4, height=3)
   print(fig)
@@
 ```{r poldeathsplots, cache=FALSE, fig_width=10, fig_height=6, warning=FALSE}
-fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=21) +
+fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=L.d) +
   ggtitle(TeX(relabel("dlogdd given pmaskbus"))) +
@@
 p <- "pmaskbus"
-figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
+figdev(sprintf("%s/tex/tables_and_figures/%s-deaths-%s.pdf", rootdir, p, L.d),width=4, height=3)
 print(fig)
@@
 
-# fig <- pfigure(df, "pmaskbus", stlab=-1, outcome="dlogdd", L=14) +
-#   ggtitle(TeX(relabel("dlogdd given pmaskbus"))) +
-#   ylab(TeX(relabel("dlogdd")))
-# p <- "pmaskbus"
-# figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
-# print(fig)
-# dev.off()
-# fig
-
-fig <- pfigure(df, "pk12", stlab=-1, outcome="dlogdd", L=21) +
+fig <- pfigure(df, "pk12", stlab=-1, outcome="dlogdd", L=L.d) +
   ggtitle(TeX(relabel("dlogdd given pk12"))) +
-  xlim(c(as.Date("2020-03-15"), max(df$date))) +
-  ylim(c(-1,3)) + ylab(TeX(relabel("dlogdd")))
+   xlim(c(as.Date("2020-03-21"), as.Date("2020-05-01"))) +
+  ylim(c(-0.6,2)) + ylab(TeX(relabel("dlogdd")))
 
 p <- "pk12"
-figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p),width=4, height=3)
+figdev(sprintf("%s/tex/tables_and_figures/%s-deaths-%s.pdf", rootdir, p, L.d),width=4, height=3)
 print(fig)
@@
 for (p in pols[3:length(pols)]) {
-  fig <- pfigure(df, p, stlab=-1, outcome="dlogdd", L=21) +
+  fig <- pfigure(df, p, stlab=-1, outcome="dlogdd", L=L.d) +
     ggtitle(TeX(relabel(sprintf("dlogdd given %s",p)))) +
@@
     ylim(c(-1,1)) + ylab(TeX(relabel("dlogdd")))
-  figdev(sprintf("%s/tex/tables_and_figures/%s-deaths.pdf", rootdir, p), width=4, height=3)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-deaths-%s.pdf", rootdir, p, L.d), width=4, height=3)
   print(fig)

### 2020-08-12 tex/covidEffectOfNPI-revision.tex
update on DML
update

--- 2020-07-16 tex/covidEffectOfNPI.tex
@@
 {Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.}\thanks{We are grateful to Daron Acemoglu, V.V.
-Chari, Raj Chetty,  Christian Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green,  Ido Rosen, Konstantin Sonin, James Stock, and Ivan Werning for helpful comments.
+Chari, Raj Chetty,  Christian Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green,  C.
+Jessica E.
+Metcalf, Ido Rosen, Konstantin Sonin, James Stock, and Ivan Werning for helpful comments.
 We also thank Chiyoung Ahn,
@@
     \begin{tabular}{cc}
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases}
-      &
-        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases-14}
+      &
+        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths-21}
     \end{tabular}
@@
 
+\section{Sensitivity Analysis}
+
+In this section, we provide sensitivity analysis by estimating (\ref{eq:R4}) with alternative specifications and methods as follows:
+  \begin{itemize}
+  \item Exclude  the state of New York from the sample because it may be viewed as an outlier in the early pandemic period.
+  \item Add private mask wearing rates from survey in March and April as confounder for unobserved  personal risk-aversion and initial attitude toward mask wearing.
+   \item Add the log of Trump's vote share  in 2016 presidential election as confounder for unobserved private behavioral response.
+   \item Add  past behavior variables  as information used to set policies.
+   \item All of the above, estimated using linear regression model.
+   \item All of the above, estimated by Double Machine Learning (DML) with Lasso to reduce dimensionality.
+   \item All of the above, estimated by  DML with Random Forest to reduce dimensionality and capture some nonlinearities.
+   \item Latent state confounders as components of $W_{it}$, estimated using fixed effects with bias correction.
+   \item Alternative Timing assumption ($\ell=10$ for cases and $\ell=23$ for deaths) on all of the above.\footnote{This alternative timing assumption is motivated by the estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
+We thank Jessica Metclaf for recommending us to do a sensitivity analysis on the timing assumption while suggesting this reference to us.}
+  \end{itemize}
+
+%We examine the robustness of our results by adding more controls, excluding the state of New York, and applying  estimation methods of double/debiased machine learning (DML) of \cite{chernozhukov18} and bias corrected fixed effects estimator.
+We also provide a sensitivity analysis regarding the assumption on the times from exposure to case confirmation and death reporting.
+
+Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses across different specifications, estimation methods, and timing assumptions.
+Here, ``red'', ``green'', ``blue'', and ``purple" indicate the regression models for case growth without national information variables, case growth with national information variables, death growth without national information variables, and death growth with national information variables, respectively.
+The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 10 and 23 days, respectively, in the right panel of ``(ii) alternative timing.''
+
+In each of four regression models, we report: (1) baseline specification  in Table xx; (2) excluding New York from the sample; (3) adding  the percentage of people who wears masks to protect themselves between March 26-April 29, 2020 to control for people's attitude toward mask wearing in the first half of sample;\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.} (4) adding the log of Trump voting shares;   (5)  including four weeks lagged behavior variables as an additional set of controls, where our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables under this specification; (6) including  all additional controls in (3)-(5) with the sample that excludes New York; (7) applying the DML  with Lasso to (6); (8) applying the DML with Random Forest to (6); (9) including state fixed effects while applying bias correction based on cross-splitting.\footnote{In the appendix, Table \ref{tab:PtoY-robust} presents the estimates for (1)-(6) while Table \ref{tab:PtoY-fe} compares bias corrected fixed effect estimates with those without bias correction.
+The appendix also provides more detailed discussion on how we implement DML procedures as well as bias correction to fixed effects estimator.}
+
+Panel (A) of Figure  \ref{fig:whisker} illustrates that the estimated coefficients of mask mandates are negative and significant across different specifications, methods, and timing assumptions, confirming the importance of mask policy on reducing case and death growths.
+% Comparing (i) baseline timing on the left with (ii) alternative timing on the right,  we notice that the estimated magnitudes of masks for employees are similar to each other.
+
+In Panel (B) of  Figure  \ref{fig:whisker}, many estimates of closures of K-12 schools  suggest that  the effect of school closures is large.
+The visual evidence on growth rates for states with and without school closures in Figure \ref{fig:growthpolicies1} also suggests that there may be a potentially large effect, though the history is very short.
+This evidence is consistent
+with the emerging evidence of prevalence of Covid-19 among children  \citep{Lee2020jama,Szablewski2020cdc}. \cite{children:nature} find that although children's
+transmission and susceptibility rates are half that of ages 20-30,
+children's contact rates are much higher.
+This type of evidence, as well as, evidence that children carry viral
+loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.
+
+Our estimates of school closures substantially vary  across  specifications, however, and the estimated effects of school closures on case growth are not significant once national cases are controlled for.
+In the US state-level data,  there is little variation across states in the timing of school closures.  
+Consequently, its estimate is particularly sensitive to an inclusion of some aggregate variables such as national cases.
+Given this sensitivity,  there still exists a lot of uncertainty as to the magnitude of the effect of school closures.
+Any analyses of re-opening plans need to be aware of this uncertainty.
+An important research question is how to resolve this uncertainty using additional data sources.
+
+Panel (C) of Figure  \ref{fig:whisker} indicates that the estimated coefficients of stay-at-home order  are generally negative and often significant 
+ except for fixed effects estimator with bias correction.
+The sensitivity under fixed effects estimator may be due to high correlations between stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses (see Table \ref{tab:correlation}).
+As shown in Panel (D), the  estimated coefficient of the average of  stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses   are  negative and often significant even under fixed effect specification; similar to school closure estimates, however,  its estimate is sensitive to controlling for national cases and deaths because the timing of implementing these policies is similar across states.
+
+\begin{figure}[ht]
+  \caption{Estimated Coefficients for  Policy Variables: Sensitivity Analysis \label{fig:whisker}}\bigskip
+  \begin{minipage}{\linewidth}
+    \centering
+   {\textbf{(A)  masks for employees}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-10} 
+    \end{tabular}
+  \end{minipage} \\\smallskip
+    \begin{minipage}{\linewidth}
+    \centering 
+     {\textbf{(B)  closed K-12 Schools}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-10}
+          \end{tabular}
+  \end{minipage}  
+    \begin{flushleft}
+      \footnotesize
+      $^\dagger$The times from exposure to case confirmation and death reporting  are assumed to be 14 and 21 days, respectively.
+$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 10 and 23 days, respectively.
+    \end{flushleft}
+\end{figure}
+
+   \addtocounter{figure}{-1}
+\begin{figure}[ht]
+  \caption{Estimated Coefficients for Policy Variables:  Sensitivity Analysis (cont.) \label{fig:whisker-2}}\bigskip
+  \begin{minipage}{\linewidth}
+    \centering
+   {\textbf{(C)  stay-at-home orders}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-10} 
+    \end{tabular}
+  \end{minipage} \\ \smallskip
+    \begin{minipage}{\linewidth}
+    \centering 
+     {\textbf{(D)  Average of stay-at-home, closed movie theaters, closed restaurants, and closed  businesses}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-10}
+          \end{tabular}
+  \end{minipage}   
+    \begin{flushleft}
+      \footnotesize
+      $^\dagger$The times from exposure to case confirmation and death reporting  are assumed to be 14 and 21 days, respectively.
+$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 10 and 23 days, respectively.
+    \end{flushleft}
+\end{figure}
+
+ \begin{figure}
+  \caption{Case and death growth conditional on school closures \label{fig:growthpolicies1}}
+  \begin{minipage}{\linewidth}
+    \centering
+    \begin{tabular}{cc}
+       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths-21} 
+    \end{tabular} 
+      \footnotesize In these figures, red points are the case or death
+      growth rate in states without each policy 14 (or 21 for deaths)
+      days earlier. Blue points are states with each policy 14 (or 21
+      for deaths) days earlier. The red line is the average across
+      states without each policy. The blue line is the average across
+      states with each policy.
+  \end{minipage}
+\end{figure}
+
 \section{Empirical Evaluation of Counterfactual Policies}\label{counterfactuals}
@@
     \begin{tabular}{cc}
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-deaths}
+      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmaskbus-deaths-21}
+      \\
+      \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths-21}
+      \\
+      \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pshelter-deaths-21}
       \\
@@
     \begin{tabular}{cc}
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-deaths}
-      \\
-      \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-deaths}
+      \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pmovie-deaths-21}
+      \\
+      \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/prestaurant-deaths-21}
+      \\
+      \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pnonessential-deaths-21}
     \end{tabular}
@@
 
+\end{document}
+
 \subsection{Counterfactuals for Massachusetts and
@@
 
-%\begin{figure}[h]
-%  \caption{Effect of removing all policies in
-%    Illinois \label{fig:IL-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dgrowth_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dcases_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dgrowth_deaths_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nop-dcases_deaths_v1}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-%
-%\begin{figure}[h]
-%  \caption{Effect of removing all policies in
-%    Massachusetts \label{fig:MA-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dgrowth_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dcases_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dgrowth_deaths_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nop-dcases_deaths_v1}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-
-\FloatBarrier
-
-%\subsection{Policy Effects without National Cases as Information Variables}
-%
-%Table \ref{tab:dieff-si} presents the estimates under a specification that excludes national information variables.
-Comparing the estimates in table \ref{tab:dieff}, the effects of school closures and  the sum of policies are estimated substantially larger in table \ref{tab:dieff-si} when national information variables are excluded.
-%
-%%Here we present the results of counterfactual analyses that don't include the national cases/deaths as
-%%the the information variables.  The growth in the national cases nearly perfectly predicts school closure
-%%policies, which have little cross-sectional variation.  Here we remove the national cases from the information
-%%and check what happens to the counterfactual results.  The counterfactual results of mask policies, shelter-in-place,
-%%and closing
-%%non-essential businesses remain robust. The results on school closures effects change substantially,
-%%attributing a much bigger role to school closures. The resulting counterfactual of  removing
-%%all policies has a much larger effect on cases, with the lower bound of the confidence band implying at
-%%least a 7 fold increase in cases (see Figure \ref{fig:US-nop-SI} given in the main text).
-%
-%\begin{table}[!b]
-%  \caption{\label{tab:dieff-si}Direct and Indirect Policy Effects
-%    (without national information variables)}
-%\begin{minipage}{\linewidth}
-%  \centering
-%    \scriptsize
-%  \begin{tabular}{c}
-%    \textbf{Cases}
-%    \\
-%    \input{tables_and_figures/dieff-cases-SI}
-%    \\
-%    \textbf{Deaths}
-%    \\
-%    \input{tables_and_figures/dieff-deaths-deathsinfo-SI}
-%  \end{tabular}
-%  \smallskip
-%  \begin{flushleft}
-%    {\tiny Direct effects capture the effect of policy on case
-%      growth holding behavior, information, and confounders
-%      constant. Direct effects are given by ${\pcolor \pi}$ in
-%      equation (\ref{eq:R1}). Indirect effects capture how policy
-%      changes behavior and behavior shift case growth. They are given
-%      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
-%      from (\ref{eq:R2}). The total effect is
-%      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
-%      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
-%      \ref{eq:R4}. Columns ``Difference''    are  the differences between
-%      the estimates from (\ref{eq:R4}) and the combination of
-%      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
-%      Standard errors are computed by
-%      bootstrap and clustered on state.}
-%    %  \ref{tab:PtoY}.
-%    \end{flushleft}
-%  \end{minipage}
-%\end{table}
-
 \FloatBarrier

### 2020-08-12 tex/robustness.tex
update

@@
 
-\section{Robustness}
+\section{Sensitivity Analysis}
 
-In this section, we examine the robustness of our results by including additional controls and examining the sub-sample that excludes the state of New York.
+In this section, we provide sensitivity analysis by estimating (\ref{PtoY}) with alternative specifications and methods.
+Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses across different specifications and estimation methods as follows:
+  \begin{itemize}
+  \item[(1)] Baseline specification  in Table  \ref{tab:PtoY}.
+  \item[(2)]  Exclude  the state of New York from the sample because it may be viewed as an outlier in the early pandemic period.
+  \item[(3)]   Add  the percentage of people who wears masks to protect themselves in March and April as confounder for unobserved  personal risk-aversion and initial attitude toward mask wearing.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.}
+   \item[(4)]   Add the log of Trump's vote share  in 2016 presidential election as confounder for unobserved private behavioral response.
+   \item[(5)]   Add  past behavior variables  as information used to set policies.
+Under this specification, our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables.
+   \item[(6)]   Include all additional controls in (3)-(5) with the sample that excludes New York as in (2).
+   \item[(7)]   Estimated by Double Machine Learning (DML) of \cite{chernozhukov18} with Lasso to reduce dimensionality while adding private mask wearing rates and  the log of Trump's vote share as additional controls.
+\item[(8)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while adding private mask wearing rates and  the log of Trump's vote share as additional controls.
+   \item[(9)]   Add latent state confounders as components of $W_{it}$, estimated using fixed effects with bias correction.
+  % \item Alternative Timing assumption ($\ell=10$ for cases and $\ell=23$ for deaths) on all of the above.
+  \end{itemize} 
+In Figure \ref{fig:whisker}, ``red'', ``green'', ``blue'', and ``purple" indicate the regression models for case growth without national information variables, case growth with national information variables, death growth without national information variables, and death growth with national information variables, respectively.
+The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 9 and 24 days, respectively, in the right panel of ``(ii) alternative timing.''\footnote{This alternative timing assumption is motivated by the lower bound estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
+We thank Jessica Metclaf for recommending us to do a sensitivity analysis on the timing assumption while suggesting this reference to us.}
 
-Table \ref{tab:PtoY-robust} presents the regression result for robustness check, where dependent variable is the weekly growth rate of
-  cases  (left panel) or deaths (right panel).
+%We examine the robustness of our results by adding more controls, excluding the state of New York, and applying  estimation methods of double/debiased machine learning (DML) of \cite{chernozhukov18} and bias corrected fixed effects estimator.
+We also provide a sensitivity analysis regarding the assumption on the times from exposure to case confirmation and death reporting.
+
+%  
+%In each of four regression models, we report: (1) baseline specification  in Table xx; (2) excluding New York from the sample; (3) adding  the percentage of people who wears masks to protect themselves between March 26-April 29, 2020 to control for people's attitude toward mask wearing in the first half of sample;\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.} (4) adding the log of Trump voting shares;   (5)  including four weeks lagged behavior variables as an additional set of controls, where our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables under this specification; (6) including  all additional controls in (3)-(5) with the sample that excludes New York; (7) applying the DML  with Lasso to (6); (8) applying the DML with Random Forest to (6); (9) including state fixed effects while applying bias correction based on cross-splitting.\footnote{In the appendix, Table \ref{tab:PtoY-robust} presents the estimates for (1)-(6) while Table \ref{tab:PtoY-fe} compares bias corrected fixed effect estimates with those without bias correction.
+The appendix also provides more detailed discussion on how we implement DML procedures as well as bias correction to fixed effects estimator.}
+
+Panel (A) of Figure  \ref{fig:whisker} illustrates that the estimated coefficients of mask mandates are negative and significant across different specifications, methods, and timing assumptions, confirming the importance of mask policy on reducing case and death growths.
+% Comparing (i) baseline timing on the left with (ii) alternative timing on the right,  we notice that the estimated magnitudes of masks for employees are similar to each other.
+
+In Panel (B) of  Figure  \ref{fig:whisker}, many estimates of closures of K-12 schools  suggest that  the effect of school closures is large.
+The visual evidence on growth rates for states with and without school closures in Figure \ref{fig:growthpolicies1} also suggests that there may be a potentially large effect, though the history is very short.
+This evidence is consistent
+with the emerging evidence of prevalence of Covid-19 among children  \citep{Lee2020jama,Szablewski2020cdc}. \cite{children:nature} find that although children's
+transmission and susceptibility rates are half that of ages 20-30,
+children's contact rates are much higher.
+This type of evidence, as well as, evidence that children carry viral
+loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.
+
+Our estimates of school closures substantially vary  across  specifications, however, and the estimated effects of school closures on case growth are not significant once national cases are controlled for.
+In the US state-level data,  there is little variation across states in the timing of school closures.  
+Consequently, its estimate is particularly sensitive to an inclusion of some aggregate variables such as national cases.
+Given this sensitivity,  there still exists a lot of uncertainty as to the magnitude of the effect of school closures.
+Any analyses of re-opening plans need to be aware of this uncertainty.
+An important research question is how to resolve this uncertainty using additional data sources.
+
+Panel (C) of Figure  \ref{fig:whisker} indicates that the estimated coefficients of stay-at-home order  are generally negative and often significant 
+ except for fixed effects estimator with bias correction.
+The sensitivity under fixed effects estimator may be due to high correlations between stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses (see Table \ref{corr}).
+As shown in Panel (D), the  estimated coefficient of the average of  stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses   are  negative and often significant even under fixed effect specification; similar to school closure estimates, however,  its estimate is sensitive to controlling for national cases and deaths because the timing of implementing these policies is similar across states.
+
+\begin{figure}[ht]
+  \caption{Estimated Coefficients for  Policy Variables: Sensitivity Analysis \label{fig:whisker}}\bigskip
+  \begin{minipage}{\linewidth}
+    \centering
+   {\textbf{(A)  masks for employees}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-9} 
+    \end{tabular}
+  \end{minipage} \\\smallskip
+    \begin{minipage}{\linewidth}
+    \centering 
+     {\textbf{(B)  closed K-12 Schools}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-9}
+          \end{tabular}
+  \end{minipage}  
+    \begin{flushleft}
+      \footnotesize
+      $^\dagger$The times from exposure to case confirmation and death reporting  are assumed to be 14 and 21 days, respectively.
+$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 9 and 24 days, respectively.
+    \end{flushleft}
+\end{figure}
+
+   \addtocounter{figure}{-1}
+\begin{figure}[ht]
+  \caption{Estimated Coefficients for Policy Variables:  Sensitivity Analysis (cont.) \label{fig:whisker-2}}\bigskip
+  \begin{minipage}{\linewidth}
+    \centering
+   {\textbf{(C)  stay-at-home orders}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-9} 
+    \end{tabular}
+  \end{minipage} \\ \smallskip
+    \begin{minipage}{\linewidth}
+    \centering 
+     {\textbf{(D)  Average of stay-at-home, closed movie theaters, closed restaurants, and closed  businesses}}\\
+    \medskip
+    \begin{tabular}{cc}  
+ $\quad$  (i) baseline timing$^\dagger$ &$\quad$ (ii) alternative timing$^\ddagger$\\ 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-14}
+      &
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-9}
+          \end{tabular}
+  \end{minipage}   
+    \begin{flushleft}
+      \footnotesize
+      $^\dagger$The times from exposure to case confirmation and death reporting  are assumed to be 14 and 21 days, respectively.
+$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 9 and 24 days, respectively.
+    \end{flushleft}
+\end{figure}
+
+ \begin{figure}
+  \caption{Case and death growth conditional on school closures \label{fig:growthpolicies1}}
+  \begin{minipage}{\linewidth}
+    \centering
+    \begin{tabular}{cc}
+       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases-14}
+      &
+        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths-21} 
+    \end{tabular} 
+      \footnotesize In these figures, red points are the case or death
+      growth rate in states without each policy 14 (or 21 for deaths)
+      days earlier. Blue points are states with each policy 14 (or 21
+      for deaths) days earlier. The red line is the average across
+      states without each policy. The blue line is the average across
+      states with each policy.
+  \end{minipage}
+\end{figure}
+
+%Table \ref{tab:PtoY-robust} presents the regression result for robustness check, where dependent variable is the weekly growth rate of
+%  cases  (left panel) or deaths (right panel).
 Column (1) replaces four policy variables  (stay-at-home, closure of movie theaters, closure of restaurants, and closure of non-essential businesses) that are highly correlated to each other with the variable of their average values.
 Column (2) excludes the state of New York, which could be viewed as an outlier in the early pandemic period, from the sample.
- Column (3) adds the percentage of people who wears masks to protect themselves between March 26-April 29, 2020, which controls for people's attitude toward mask wearing in the first half of sample.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+% Column (3) adds the percentage of people who wears masks to protect themselves between March 26-April 29, 2020, which controls for people's attitude toward mask wearing in the first half of sample.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
 The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.}  Column (5) controls for the log of Trump voting shares in 2016 presidential election.
@@
 The estimated coefficients of mask mandates on case or death growths are significantly negative in all specifications, confirming the importance of mask policy on reducing case and death growths.
-\textbf{[here, instead of table, we may present ``dot-and-whisker plot'' for mask coefficients as well as stay-at-home and pindex.]}
-
-In Table \ref{tab:PtoY-fe} of the appendix,   we find that the estimated coefficients of mask mandates are significantly negative even when we include state fixed effects.
+% 
+%   
+%In Table \ref{tab:PtoY-fe} of the appendix,   we find that the estimated coefficients of mask mandates are significantly negative even when we include state fixed effects.
 In contrast,  the estimated coefficients of stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses under fixed effects specification are sometimes different from those under random effects specification, especially when bias correction is applied to the fixed effects estimator.
 Non-robustness of the estimated coefficients of these four  policy variables
- is likely due to high correlations among them (see Table \ref{corr}). In fact, as shown in columns (3)-(4) of Table \ref{tab:PtoY-fe}, the  estimated effects of the average of four policy variables 
- under fixed effects specification are negative and significant, indicating that  the total effect of four policies is robustly estimated.
+% is likely due to high correlations among them (see Table \ref{corr}).
+In fact, as shown in columns (3)-(4) of Table \ref{tab:PtoY-fe}, the  estimated effects of the average of four policy variables  under fixed effects specification are negative and significant, indicating that  the total effect of four policies is robustly estimated.
+%
 
@@
  \caption{\label{tab:PtoY-robust}
-Robustness Check for the Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
+Sensitivity Check for the Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
  \begin{minipage}{\linewidth}
@@
 
+\begin{footnotesize}
+
+\bibliographystyle{jpe}
+\bibliography{covid}
+
+\end{footnotesize}
+
 \end{document}

### 2020-08-13 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

@@
 # Set time lag:
-# L.c <- 14
-# L.d <- 21
-L.c <- 9
-L.d <- 24
+L.c <- 14
+L.d <- 21
+# L.c <- 9
+# L.d <- 24
 
@@
 
-```{r dieff-deaths, eval = FALSE, cache=FALSE, results='asis', warning=FALSE}
+```{r dieff-deaths, eval = TRUE, cache=FALSE, results='asis', warning=FALSE}
 for (use.national in c(TRUE,FALSE)) {

### 2020-08-13 tex/robustness.tex
update

@@
 In this section, we provide sensitivity analysis by estimating (\ref{PtoY}) with alternative specifications and methods.
-Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses across different specifications and estimation methods as follows:
+Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses for the following specifications and estimation methods:
   \begin{itemize}
@@
 
-%We examine the robustness of our results by adding more controls, excluding the state of New York, and applying  estimation methods of double/debiased machine learning (DML) of \cite{chernozhukov18} and bias corrected fixed effects estimator.
-We also provide a sensitivity analysis regarding the assumption on the times from exposure to case confirmation and death reporting.
-
-%  
-%In each of four regression models, we report: (1) baseline specification  in Table xx; (2) excluding New York from the sample; (3) adding  the percentage of people who wears masks to protect themselves between March 26-April 29, 2020 to control for people's attitude toward mask wearing in the first half of sample;\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
-The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.} (4) adding the log of Trump voting shares;   (5)  including four weeks lagged behavior variables as an additional set of controls, where our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables under this specification; (6) including  all additional controls in (3)-(5) with the sample that excludes New York; (7) applying the DML  with Lasso to (6); (8) applying the DML with Random Forest to (6); (9) including state fixed effects while applying bias correction based on cross-splitting.\footnote{In the appendix, Table \ref{tab:PtoY-robust} presents the estimates for (1)-(6) while Table \ref{tab:PtoY-fe} compares bias corrected fixed effect estimates with those without bias correction.
-The appendix also provides more detailed discussion on how we implement DML procedures as well as bias correction to fixed effects estimator.}
-
 Panel (A) of Figure  \ref{fig:whisker} illustrates that the estimated coefficients of mask mandates are negative and significant across different specifications, methods, and timing assumptions, confirming the importance of mask policy on reducing case and death growths.
@@
 
-Our estimates of school closures substantially vary  across  specifications, however, and the estimated effects of school closures on case growth are not significant once national cases are controlled for.
+Our estimates of school closures substantially vary  across  specifications, however, and the estimated effects of school closures on case/death growth are often not significantly different from zero once national cases are controlled for.
 In the US state-level data,  there is little variation across states in the timing of school closures.  

### 2020-08-14 cases_and_policies/rmd/cf-pindex.Rmd
Add new figures.

@@
 
-## Stay at home
+## Averaged policies
 
-We now look at what happens if stay at home orders were never issued.
+We now look at what happens if none of the four averaged policies
+(stay at home, close movie theaters, close restaurants, close
+nonessential businesses) were implemented.
 
@@
 cf <- df
-cf[,"pshelter"] <- 0
+cf[,"pindex"] <- 0
 cf <- subset(cf, cf$date>=datelims[1]) #+L)
@@
 
-if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
+if (!file.exists("cfindex_idx.Rda") || rerunsim) {
   cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfshelter_idx.Rda")
+  save(cfshelterlist, shelterall, file="cfindex_idx.Rda")
 }
-load("cfshelter_idx.Rda")
+load("cfindex_idx.Rda")
 cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
@@
   f <- cfplots(st, cfshelter, data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))
+  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")
+figs <- nationalplots(shelterall, cfshelter, "no shutdown policies")
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_idx.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_idx.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-## Stay at home
+## Shutdown policies
 
-We now look at what happens if stay at home orders were never issued.
+We now look at what happens if the four shutdown policies in pindex were never implemented.
 
@@
 
-if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
+if (!file.exists("cfindex_deaths_idx.Rda") || rerunsim) {
   cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfshelter_deaths_idx.Rda")
+  save(cfshelterlist, shelterall, file="cfindex_deaths_idx.Rda")
 }
-load("cfshelter_deaths_idx.Rda")
+load("cfindex_deaths_idx.Rda")
 cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
@@
   f <- cfplots(st, cfshelter, data=df, yvar="deaths")
-  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))
+  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figdp)
@@
 
-  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
   print(f$figp)
@@
 
-figs <- nationalplots(shelterall, cfshelter, "no stay at home orders", yvar="deaths")
+figs <- nationalplots(shelterall, cfshelter, "no shutdown policies", yvar="deaths")
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$g
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$c
@@
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
 figs$r

### 2020-08-14 tex/reply-to-referees.tex
update after including D.C. and starting date of June 3

\newtheorem{proposition}{Proposition}
\newtheorem{remark}{Remark}
\newtheorem{lemma}{Lemma}
\newtheorem{corollary}{Corollary}
\newtheorem{example}{Example}

\documentclass[11pt]{article}

\usepackage{amssymb}
\usepackage{graphics}
\usepackage{amsmath}
\usepackage{verbatim}
\usepackage{setspace}
\usepackage{color}
\usepackage{ulem}
\usepackage{sectsty}
\usepackage{hyperref}
\sectionfont{\large}
\usepackage{graphicx}
\def\bs{\boldsymbol}

%\renewcommand{\sout}[1]{}
%\renewcommand{\bf}{}

\newtheorem{assumption}{Assumption}
\setlength{\topmargin}{-1cm}
\setlength{\oddsidemargin}{0.25cm}
\setlength{\evensidemargin}{-0.2cm} \setlength{\textheight}{22cm}
\setlength{\textwidth}{16cm}
\onehalfspacing
\def\thebibliography#1{\section*{References\markboth
  {REFERENCES}{REFERENCES}}\list
  {}{\settowidth\labelwidth{}\leftmargin\labelwidth
  \advance\leftmargin\labelsep
  \usecounter{enumi}}
  \def\newblock{\hskip .11em plus .33em minus -.07em}
  \sloppy
  \sfcode`\.=1000\relax}

  \pdfminorversion=4

\begin{document}

\newcommand{\replytitle}
{
\subsection*{Reply to the Comments on  ``Causal Impact of Masks, Policies, Behavior on Early COVID-19 Pandemic in The U.S.''}
}

\newcommand{\refreplyheading}
{Thank you very much for your helpful and constructive comments on our paper. We appreciate the time you spent in reviewing this paper. We have followed your recommendations, and those of the other reviewers', as far as possible.
Major changes are:}

\newcommand{\changes}
{
\begin{enumerate}

\item   
%\item We made some grammatical adjustments.
\end{enumerate}

%No other changes of substance are made.
}

\replytitle
\subsection*{Reply to Editor}
\refreplyheading

\changes

\newpage

\replytitle
\subsection*{Reply to Referee 1}
\refreplyheading

\changes
\subsection*{Reply to Comments}
\begin{itemize}

\item  \textit{``I think it would be helpful if the authors clearly state the limitations of their work and
cautiously warn that their casual impact cannot be interpreted too literally.''}

\item[1.]   \textit{(Source of state-level variation) ``The main specification is based on random effects panel data
models. It is unclear what is the source of state-level variation. In particular, what is behind
the difference in mask policies across the states? Is this mainly driven by political partisanship?
To argue that the current paper credibly identifies the casual impact of masks for
employees, could the economics profession be persuaded that adopting masks for employees
is sufficiently random conditional on observed confounders? The paper provides some
discussions; however, more thorough discussions would be highly valuable.''}

We provide various sensitivity checks. First, we now include governor's party affiliation as an additional state-level variable in random effects specification. 
Second, we exclude the state of New York from the sample because it may be viewed as an
outlier in the early pandemic period and it is  one of the states that adopted mandatory mask policies in April. Third, we added mask wearing rates in Mach and April from survey as an additional regressor to control for unobserved personal risk-aversion and people's initial attitude for mask wearing. Fourth,  we added the log of Trump's vote share in 2016 presidential election as confounder for
unobserved private behavioral response. Fifth, we report the fixed effects estimator that controls for latent state confounders as components of $W_{it}$.  
As shown in Figures xx(2)-(4) and (9), the estimated coefficient of masks for employees is not so sensitive with respect to excluding NY as well as adding more controls that are possibly important confounders.

As stated below, we also examine a specification that includes lagged behavior variables as information; employed the Double Machine Learning (DML) with Lasso for dimensionality reduction of confounders as well as the DML with Random Forest for dimensionality reduction and capturing potential nonlinearity of confounders. We find that the estimated coefficients of masks for employees are robust with respect to alternative specifications and methods.

\item[2. ]  \textit{(Information structure) ``I think it is really nice that the paper emphasizes the importance of
information structure. Having said this, section 2.2 is quite limited because the most extensive
form of information structure consists of time effects and lagged and integrated effects
of state-level cases and deaths. I am puzzled what sense these form information structure. It
is unclear to me why lagged values of policies and behavior are not part of the information 
structure. I think it would be useful to provide remarks regarding the limitation of the current
information structure and acknowledge that the estimation results and in particular the
simulated counterfactuals are crucially dependent on it.''}

We chose information structure that is parsimonious and yet we believe captures the most important part of information that is relevant for people's behavior.

We think that conditioning on current policy variables that is included in all specifications, lagged values of policies will  add little additional information and it is natural to exclude lagged values of policies from the information structure --- or we may re-interpret that the coefficient of current policy variables partially reflects its effect through providing information.

As suggested, it is possible to add lagged values of policies and behavior as part of information. If we specify that information structure contains two weeks lagged behavior variables, then a specification of $(PI\rightarrow Y)$ equation contains four weeks lagged behavior variables as additional regressors. As a sensitivity check, we estimated this specification with lagged behavior variables in Section 5. As shown in specification (5) of Figures xx-xx, the estimated coefficients of mask for employees, closed K-12 schools, stay-at-home orders, and the average of stay-at-home orders and closures of movie theaters, restaurants, and non-essential businesses are not so sensitive to an inclusion of lagged behavior variables as information.

\item[3.]  \textit{(Expectation and Lucas critique) ``Related to the previous comment, what would be the link
between the current paper’s setup and how expectations are formed given the information
structure? For example, is this paper safe from Lucas critique? Some comments would be
helpful.''}

\item[4.]  \textit{(Causal ordering) ``It is assumed in the paper (on page 8) that behavior is affected by policies
but not the other way. This is certainly an important exclusion restriction that the paper
depends on. This may be a reasonable first-order approximation, but it could be that some
polices are formed by behavior (such as online petition and so on). It would be useful if
the authors could add some more remarks why the causal chain on pages 8-10 is plausible
and acknowledge that the results in the paper are highly conditional on this particular causal
ordering.''}

\item[5.]  \textit{(Details about confidence intervals) ``Starting from Figure 1, confidence intervals are shown
in the shaded region. It is not clearly stated in the paper how they are obtained. On page 32,
in Figure 9, the authors state:
We set initial $\Delta\log\Delta C$ and $\log\Delta C$  to their values first observed in the state we
are simulating. We hold all other regressors at their observed values. Error terms
are drawn with replacement from the residuals. We do this many times and report
the average over draws of the residuals. The shaded region is a point-wise 90\%
confidence interval. It is still unclear how a pointwise confidence interval is obtained. Does this mean that a pointwise
90\% confidence interval is constructed by simulation or bootstrap? It would be useful if the authors describe how to construct a confidence interval and provide a remark why it is valid.''}

\item[6.]  \textit{(Replication files) ``It would be great if the authors could post their dataset and replication
files on public domain such as a GitHub page or ICPSR at https://www.openicpsr.org/
openicpsr/covid19.''}

\item[7.]  \textit{(Debiased fixed effect estimates) ``On page 11, the authors state:
However, we find the debiased fixed effect estimates are qualitatively and quantitatively
similar to the correlated random effects estimates. Given this finding,
we chose to focus on the latter, as it is a more standard and familiar method, and
report the former estimates in the supplementary materials for this paper.
Could the authors kindly point out where the supplementary materials are? Perhaps I did not
locate them properly, but the appendices do not seem to include the results from debiased
fixed effect estimates.''}

\item[8.]  \textit{(Random effects vs. fixed effects) ``In addition to the previous comment, another reason for
preferring the random effects model could be that there could be substantial measurement
errors in behavior and policy variables. Fixed effects estimators could suffer more from measurement
errors.''}

\item[9.]  \textit{(Multiple testing) ``There are lots of hypotheses and stars in Tables 3-7. Perhaps it might be
useful to consider multiple testing corrections by controlling the familywise error rate or a
related concept.''}

\item[10.]  \textit{(Table 3A: Cases as Information) ``It is puzzling why the coefficients on  $\Delta\log\Delta C$ are positive
and large. How would the authors interpret these numbers? Would this mean that information
on lagged growth rates in new cases encourages people to become more mobile?
Probably I am mistaken. It would be helpful if the authors could clarify this.''}
\item[11.]  \textit{Figure 8 is very informative. Perhaps the authors could show this in introduction as well.''}

  \end{itemize}

\subsection*{Reply to Minor Comments}
\begin{itemize}

\item[1.]  \textit{(Figure 1) ``What are black and grey dotted lines? It would be helpful if the authors add notes
to the figure. The same questions apply to Figures 2 and 3 and other figures.''}
\item[2.]  \textit{(Figure 4) ``What do S. \& P. stand for in Figure 4?''}
\item[3.]  \textit{(Footnote 17) ``Is this necessary? This seems obvious. The authors could simply drop this
footnote or make it more informative.''}
\item[4.]  \textit{(Footnote 19) ``On page 17, in footnote 19, the authors state:
We drop “Residential” because it is highly correlated with both “Workplaces” and
“Retail \& recreation” at correlation coefficients of0:98.
This sentence is ambiguous. Do the authors mean that both correlation coefficients are identically 0:98? Please rephrase this sentence.''}
\item[5.]  \textit{(Figures 9 and 11) ``On page 32, in Figure 9, the subfigures in the middle and right panels
contain non-legible legend. Please improve the figures. The same applies to Figure 11 and
other similar figures in the appendix.''}
\item[6. ]  \textit{(Intepretation) ``On page 13, it is state that
The estimates imply that mandating masks on April 1st would have led to 500
fewer cases and 250 fewer deaths in Washington by the start of June.
What is the time scale of this statement? Does this mean 500 fewer cases weekly or in total
by the start of June? Please clarify the quantitative statement.''}
\item[7.]  \textit{``Some of the figures are difficult to read because there are too many curves. In particular,
Figures 18, 19, and 21 convey very little information. Please improve the figures and delete
some of them if they are less important.''}

\end{itemize}

\newpage

\replytitle
\subsection*{Reply to Comments by C. Jessica E. Metcalf }
\refreplyheading

\changes
\subsection*{Reply to Comments}
\begin{enumerate}

\item       
\end{enumerate}

\end{document}

### 2020-08-14 cases_and_policies/rmd/regressionsWithPlots.Rmd
update after including D.C. and starting date of June 3

@@
 df$testratedc.ma[df$dcases==0] <- 0
+
+df$lagtests <- panellag(df$dlogtests, df$state, df$date, lag=1)  
+df$dlogtests[is.na(df$dlogtests)] <- df$lagtests[is.na(df$dlogtests)]
+
 ```
@@
 # case growth with pindex
-pols <- c("pmaskbus","pk12","pindex") 
+pols <- c("pmaskbus","pk12","pshelter","pindex") 
 for (k in 1:length(infovars)) { 
@@
 # death growth with pindex
-pols <- c("pmaskbus","pk12","pindex") 
+pols <- c("pmaskbus","pk12","pshelter","pindex") 
 for (k in 1:length(infovars)) {  
@@
 ## Fixed Effects
-```{r regsfe,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsfe,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
 
@@
 
-df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
+df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/4
 
@@
   piy[[(2+4*(k-1))]] <- m
-  pols <- c("pmaskbus","pk12","pindex")  
+  pols <- c("pmaskbus","pk12","pshelter","pindex")  
   m <-  reg_fe(sdf, yvar, pols, NULL,
@@
 ## DML 
-```{r regsdml,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsdml,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
 
@@
 
-df$test <- df$dlogtests
-df$test[is.na(df$test)] <- 0
-
-wvars <- c("logpop","logsq","Percent.Unemployed..2018.",
+# df$test <- df$dlogtests
+# df$test[is.na(df$test)] <- 0
+
+svars <- c("logpop","logsq","Percent.Unemployed..2018.",
                "Percent.living.under.the.federal.poverty.line..2018.",
-               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
-               "mask_percent", "logvote","test","month3","month4","month5","month6") 
-wvars.d <- c("logpop", "logsq","Percent.Unemployed..2018.",
-               "Percent.living.under.the.federal.poverty.line..2018.",
-               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum",
-               "mask_percent", "logvote","month3","month4","month5","month6")
+               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum")
+mvars <- c("month4","month5","month6") 
+ijs <- expand.grid(1:length(svars), 1:length(mvars))
+msvars <- vector(mode="character", length=lengths(ijs)[[1]])
+for (ij in (1:lengths(ijs)[[1]])){
+  v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars[ijs[[2]][ij]])
+  msvars[ij] <- v
+  df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars[ijs[[2]][ij]]]
+}
+
+wvars <- c(svars,mvars,msvars,"month3","mask_percent","logvote","test") 
+wvars.d <- c(svars,mvars,msvars,"month3","mask_percent","logvote")
 
@@
     dv <- "pindex.L14" 
-    p <- c("pmaskbus.L14","pk12.L14")
+    p <- pols.L14[1:3]
     # d = sdf$pindex.L14 
@@
     dv <- "pindex.L21"
-    p <- pols.L21[1:2]
+    p <- pols.L21[1:3]
     # d = sdf$pindex.L21 

### 2020-08-14 tex/tables_and_figures/policies.tex
update
Some simple updates.
update on DML
Fix some formatting of rmd->html site
update after including D.C. and starting date of June 3

@@
 % latex table generated in R 4.0.0 by xtable 1.8-4 package
-% Wed Jul 29 09:27:24 2020
+% Fri Aug 14 14:53:13 2020
 \begin{table}[ht]
@@
   \hline
-State of emergency & 49 & 2020-02-29 & 2020-03-11 & 2020-03-16 \\ 
-  Date closed K 12 schools & 49 & 2020-03-13 & 2020-03-17 & 2020-04-03 \\ 
-  Closed day cares & 14 & 2020-03-16 & 2020-03-23 & 2020-04-06 \\ 
-  Date banned visitors to nursing homes & 29 & 2020-03-09 & 2020-03-16 & 2020-04-06 \\ 
-  Closed non essential businesses & 41 & 2020-03-19 & 2020-03-25 & 2020-04-06 \\ 
-  Closed restaurants except take out & 46 & 2020-03-15 & 2020-03-17 & 2020-04-03 \\ 
-  Closed gyms & 44 & 2020-03-16 & 2020-03-20 & 2020-04-03 \\ 
-  Closed movie theaters & 47 & 2020-03-16 & 2020-03-21 & 2020-04-06 \\ 
-  Stay at home  shelter in place & 40 & 2020-03-19 & 2020-03-28 & 2020-04-07 \\ 
-  End relax stay at home shelter in place & 36 & 2020-04-24 & 2020-05-15 & 2020-06-19 \\ 
-  Began to reopen businesses statewide & 48 & 2020-04-20 & 2020-05-06 & 2020-06-05 \\ 
-  Reopen restaurants & 46 & 2020-04-24 & 2020-05-17 & 2020-06-22 \\ 
-  Reopened gyms & 42 & 2020-04-24 & 2020-05-18 & 2020-07-03 \\ 
-  Reopened movie theaters & 35 & 2020-04-27 & 2020-05-30 & 2020-07-03 \\ 
-  Resumed elective medical procedures & 34 & 2020-04-20 & 2020-04-29 & 2020-05-29 \\ 
-  Mandate face mask use by all individuals in public spaces & 30 & 2020-04-08 & 2020-06-22 & 2020-07-27 \\ 
-  Mandate face mask use by employees in public facing businesses & 40 & 2020-04-03 & 2020-05-01 & 2020-07-03 \\ 
-  Stop Initiation of Evictions overall or due to COVID related issues & 29 & 2020-03-16 & 2020-03-24 & 2020-05-30 \\ 
-  Stop enforcement of evictions overall or due to COVID related issues & 27 & 2020-03-17 & 2020-03-24 & 2020-06-08 \\ 
+State of emergency & 51 & 2020-02-29 & 2020-03-11 & 2020-03-16 \\ 
+  Date closed K 12 schools & 51 & 2020-03-13 & 2020-03-17 & 2020-04-03 \\ 
+  Closed day cares & 15 & 2020-03-16 & 2020-03-23 & 2020-04-06 \\ 
+  Date banned visitors to nursing homes & 30 & 2020-03-09 & 2020-03-16 & 2020-04-06 \\ 
+  Closed non essential businesses & 43 & 2020-03-19 & 2020-03-25 & 2020-04-06 \\ 
+  Closed restaurants except take out & 48 & 2020-03-15 & 2020-03-17 & 2020-04-03 \\ 
+  Closed gyms & 49 & 2020-03-16 & 2020-03-20 & 2020-04-03 \\ 
+  Closed movie theaters & 49 & 2020-03-16 & 2020-03-21 & 2020-04-06 \\ 
+  Stay at home  shelter in place & 42 & 2020-03-19 & 2020-03-28 & 2020-04-07 \\ 
+  End relax stay at home shelter in place & 39 & 2020-04-24 & 2020-05-18 & 2020-06-19 \\ 
+  Began to reopen businesses statewide & 50 & 2020-04-20 & 2020-05-07 & 2020-06-05 \\ 
+  Reopen restaurants & 48 & 2020-04-24 & 2020-05-18 & 2020-06-22 \\ 
+  Reopened gyms & 44 & 2020-04-24 & 2020-05-18 & 2020-07-13 \\ 
+  Reopened movie theaters & 37 & 2020-04-27 & 2020-06-01 & 2020-07-13 \\ 
+  Resumed elective medical procedures & 35 & 2020-04-20 & 2020-04-30 & 2020-05-29 \\ 
+  Mandate face mask use by all individuals in public spaces & 35 & 2020-04-08 & 2020-06-26 & 2020-08-05 \\ 
+  Mandate face mask use by employees in public facing businesses & 44 & 2020-04-03 & 2020-05-01 & 2020-08-03 \\ 
+  Stop Initiation of Evictions overall or due to COVID related issues & 30 & 2020-03-16 & 2020-03-24 & 2020-05-30 \\ 
+  Stop enforcement of evictions overall or due to COVID related issues & 29 & 2020-03-15 & 2020-03-24 & 2020-06-08 \\ 
   Renter grace period or use of security deposit to pay rent & 3 & 2020-04-10 & 2020-04-24 & 2020-05-20 \\ 
-  Order freezing utility shut offs & 32 & 2020-03-12 & 2020-03-19 & 2020-04-13 \\ 
+  Order freezing utility shut offs & 34 & 2020-03-12 & 2020-03-19 & 2020-04-13 \\ 
   Froze mortgage payments & 1 & 2020-03-21 & 2020-03-21 & 2020-03-21 \\ 
-  Waived one week waiting period for unemployment insurance & 35 & 2020-03-08 & 2020-03-18 & 2020-04-06 \\ 
+  Waived one week waiting period for unemployment insurance & 36 & 2020-03-08 & 2020-03-18 & 2020-04-06 \\ 
    \hline

### 2020-08-14 cases_and_policies/R/cases_and_policies.R

@@ -25,7 +25,7 @@ gitrootdir <- function() {
   rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
   if (is.na(rootdir)) {
     thisfilepath <- thisPath()
-    rootdir <- normalizePath(paste(thifilepath,"../..",sep="/"))
+    rootdir <- normalizePath(paste(thisfilepath,"../..",sep="/"))
   }
   validationfile <- normalizePath(paste(rootdir,"cases_and_policies/",sep="/"))
   if (!file.exists(validationfile)) {
@@ -35,8 +35,6 @@ gitrootdir <- function() {
 }
 
 rootdir <- gitrootdir()
-
-rootdir <- "/Users/Hiro/Documents/GitHub/covid-impact"
 
 # Data on policies from Raifman and collaborators
 urls <- c("https://docs.google.com/spreadsheets/d/1zu9qEWI8PsOI_i8nI_S29HDGHlIp2lfVMsGxpQ5tvAQ/edit#gid=993060716",

### 2020-08-14 cases_and_policies/rmd/generatetables.R

@@ -102,6 +102,7 @@ mainregressions <- function(df, # data
 
   plist <- list(pols, c("pmask.april","pmask.may", pols[-1]))
   plist <- list(pols,c("pmaskbus","pk12","pindex"))
+  plist <- list(pols,c("pmaskbus","pk12","pshelter","pindex"))
   
   
 
@@ -135,12 +136,16 @@ mainregressions <- function(df, # data
               interactions[[ij[2]]], iv[[ij[2]]], L=L)
   })
 
+   
+ # sdf <- subset(df, df$date<=(max(df$date)-L)) 
+ sdf <- df
   ijs <- expand.grid(1:length(bvars), 1:length(plist))
   pib <- list()
   for (k in 1:length(xlist)) {
     pib[[k]] <- apply(ijs, 1, function(ij) {
-      policyreg(df, bvars[ij[1]], plist[[ij[2]]], NULL,
+      policyreg(sdf, bvars[ij[1]], plist[[ij[2]]], NULL,
                 c(infovars[[k]],
+                 # tvars,
                   xlist[[k]]),
                 interactions[[k]], "0", L=0)
     })

### 2020-08-14 cases_and_policies/rmd/regprep.R

@@ -2,7 +2,7 @@ cs$t <- as.numeric(cs$date)
 
 warning("Using data only up to 2020-06-02 as in submitted version of paper. Modify lines 4-5 of regprep.R to change")
 df <- pdata.frame(subset(cs, cs$date>=as.Date("2020-01-01") &
-                                   cs$date<=as.Date("2020-06-02")),
+                                   cs$date<=as.Date("2020-06-03")),
                   # cs$date<=as.Date("2020-06-30")),
                   index=c("state","t"), stringsAsFactors=FALSE)
 
@@ -126,8 +126,11 @@ df$mask_percent <- df$mask_percent/100
 
 df$party <- as.factor(df$party)
 
-df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
+# df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
 #df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
+
+df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/3
+
 
 tilt = ifelse(df$date> as.Date("2020-05-01"),1,0 )
 month <- month(df$date)

### 2020-08-14 cases_and_policies/rmd/cf-pindex.Rmd

@@
 
-To compute the estimated and counterfactual paths we use the estimates
-from column 2 of table \ref{tab:policy-cases}. We set initial $\Delta
-\log \Delta C$ and $\log \Delta C$ to their values first observed in
-the state we are simulating. We hold all other regressors at their
-observed values. Error terms are drawn with replacement from the
-residuals. We do this many times and report the average over draws of
-the residuals. To calculate confidence intervals, we repeat the above
-with coefficients drawn randomly from their asymptotic
-distribution. The shaded region is a pointwise 90\% confidence
-interval.
-
-To be more precise, we can rewrite our regression model as:
-
-$$
-  Y_{it} = P_{it}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \epsilon_{it}
-$$
-
-where $Y_{it} = \log \Delta C_{it}$, $P_{it}$ are policies, and
-$X_{it}$ are controls. Each counterfactual involves changing $P_{it}$
-to some other value, $P_{it}^\ast$. We then generate counterfactual paths
-for cases by in state $i$ conditional on
-$Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast \}_{t=1}^T, \theta=(\alpha, \beta, \rho)$ by setting:
-
-$$
-  Y_{it}^{\ast,r} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}^{\ast,r}
-$$
-
-where $\hat{\epsilon}_{it}^{\ast,r}$ are resampled with replacement
-from $\{\hat{\epsilon}_{it}\}_{i=1,t=1}^{N,T}$. We do resample the
-residuals $R$ times, repeating the above each time. We then compute
-
-$$
-\hat{E}[Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
-\frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}
-$$
-
-Finally, we repeatedly draw $\theta$ from its asymptotic distribution
-and repeat all of the above to compute confidence intervals.
-
-For case growth, the same procedure is used, but with $\Delta Y_{it} = \Delta \log \Delta C_{it}$ in place of $Y_{it}$. The top panel of the following figure shows
-
-$$
-E[\Delta Y_{it} | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}\}_{t=1}^T]
-$$
-
-and
-
-$$
-E[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T]
-$$
-
-along with the observed $Y_{it}$, and the bottom panel shows the difference.
-
 ```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
@@
 ```
-
-To look at cases we instead compute,
-
-$$
-\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
-\frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r})
-$$
 
@@
 
-### Holding Residuals fixed
-
-In this subsection, instead of resampling residuals, we condition on them. That is we look at
-
-$$
-  Y_{it}^{\ast} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}
-$$
-
-instead of
-
-$$
-  \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}.
-$$
-
-```{r masks-resids-fixed}
-simmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
-  simmodel(st, samplecoefs=samplecoefs, nsim=nsim,
-           shockgen=NULL)
-maskmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
-  maskmodel(st, samplecoefs=samplecoefs, nsim=nsim,
-            shockgen=NULL)
-
-rerunsim <- FALSE
-if (!file.exists("cfmask1_idx.Rda") || rerunsim) {
-  cfmasklist1 <- lapply(unique(df$state),
-                        function(s) nopstatesim(s,simobs=simmodel1,simcf=maskmodel1, data=sdf, nresid=1))
-  maskall1 <- rbindlist(lapply(cfmasklist1, getmeans))
-  save(cfmasklist1, maskall1, file="cfmask1_idx.Rda")
-}
-load("cfmask1_idx.Rda")
-cfmask1 <- do.call(rbind, lapply(cfmasklist1, function(x) x$edf))
-
-for (st in c("Washington","Illinois","Massachusetts")) {
-  f <- cfplots(st, cfmask1, yvar="deaths", data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))
-
-}
-
-figs1 <- nationalplots(maskall1, cfmask1, "mandating masks on April 1st\n")
-
-for (f in figs1) print(f)
-```
-
 ## Averaged policies
@@
 cf <- subset(cf, cf$date>=datelims[1]) #+L)
-noshelter <- createforwardsim(m, cf)
+noindex <- createforwardsim(m, cf)
 
 if (!file.exists("cfindex_idx.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
-  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfindex_idx.Rda")
+  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
+  indexall <- rbindlist(lapply(cfindexlist, getmeans))
+  save(cfindexlist, indexall, file="cfindex_idx.Rda")
 }
 load("cfindex_idx.Rda")
-cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
+cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
 ```
@@
 ```{r, cache=TRUE}
-cfshelter <- merge(cfshelter, abbrv, by="state")
+cfindex <- merge(cfindex, abbrv, by="state")
 for (st in c("Washington","Illinois","Massachusetts")) {
-  f <- cfplots(st, cfshelter, data=df)
+  f <- cfplots(st, cfindex, data=df)
   print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))
@@
 
-figs <- nationalplots(shelterall, cfshelter, "no shutdown policies")
+figs <- nationalplots(indexall, cfindex, "no shutdown policies")
 
@@
 cf <- subset(cf, cf$date>=datelims[1]) #+L)
-noshelter <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")
+noindex <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")
 
 if (!file.exists("cfindex_deaths_idx.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
-  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
-  save(cfshelterlist, shelterall, file="cfindex_deaths_idx.Rda")
+  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
+  indexall <- rbindlist(lapply(cfindexlist, getmeans))
+  save(cfindexlist, indexall, file="cfindex_deaths_idx.Rda")
 }
 load("cfindex_deaths_idx.Rda")
-cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
+cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
 ```
@@
 ```{r, cache=TRUE}
-cfshelter <- merge(cfshelter, abbrv, by="state")
+cfindex <- merge(cfindex, abbrv, by="state")
 for (st in c("Washington","Illinois","Massachusetts")) {
-  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
+  f <- cfplots(st, cfindex, data=df, yvar="deaths")
   print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))
@@
 
-figs <- nationalplots(shelterall, cfshelter, "no shutdown policies", yvar="deaths")
+figs <- nationalplots(indexall, cfindex, "no shutdown policies", yvar="deaths")
 

### 2020-08-14 cases_and_policies/rmd/counterfactuals.Rmd

@@
   cfnoplist <- lapply(unique(sdf$state), function(s)
-    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
+    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf, nresid=1))
   allmeans <- rbindlist(lapply(cfnoplist, getmeans))
@@
 
-To compute the estimated and counterfactual paths we use the estimates
-from column 2 of table \ref{tab:policy-cases}. We set initial $\Delta
-\log \Delta C$ and $\log \Delta C$ to their values first observed in
-the state we are simulating. We hold all other regressors at their
-observed values. Error terms are drawn with replacement from the
-residuals. We do this many times and report the average over draws of
-the residuals. To calculate confidence intervals, we repeat the above
-with coefficients drawn randomly from their asymptotic
-distribution. The shaded region is a pointwise 90\% confidence
-interval.
+We hold residuals fixed throughout.
+<!-- To compute the estimated and counterfactual paths we use the estimates -->
+<!-- from column 2 of table \ref{tab:policy-cases}. We set initial $\Delta -->
+<!-- \log \Delta C$ and $\log \Delta C$ to their values first observed in -->
+<!-- the state we are simulating. We hold all other regressors at their -->
+<!-- observed values. Error terms are drawn with replacement from the -->
+<!-- residuals. We do this many times and report the average over draws of -->
+<!-- the residuals. To calculate confidence intervals, we repeat the above -->
+<!-- with coefficients drawn randomly from their asymptotic -->
+<!-- distribution. The shaded region is a pointwise 90\% confidence -->
+<!-- interval. -->
 
-To be more precise, we can rewrite our regression model as:
+<!-- To be more precise, we can rewrite our regression model as: -->
 
-$$
-  Y_{it} = P_{it}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \epsilon_{it}
-$$
+<!-- $$ -->
+<!--   Y_{it} = P_{it}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \epsilon_{it} -->
+<!-- $$ -->
 
-where $Y_{it} = \log \Delta C_{it}$, $P_{it}$ are policies, and
-$X_{it}$ are controls. Each counterfactual involves changing $P_{it}$
-to some other value, $P_{it}^\ast$. We then generate counterfactual paths
-for cases by in state $i$ conditional on
-$Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast \}_{t=1}^T, \theta=(\alpha, \beta, \rho)$ by setting:
+<!-- where $Y_{it} = \log \Delta C_{it}$, $P_{it}$ are policies, and -->
+<!-- $X_{it}$ are controls. Each counterfactual involves changing $P_{it}$ -->
+<!-- to some other value, $P_{it}^\ast$. We then generate counterfactual paths -->
+<!-- for cases by in state $i$ conditional on -->
+<!-- $Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast \}_{t=1}^T, \theta=(\alpha, \beta, \rho)$ by setting: -->
 
-$$
-  Y_{it}^{\ast,r} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}^{\ast,r}
-$$
+<!-- $$ -->
+<!--   Y_{it}^{\ast,r} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}^{\ast,r} -->
+<!-- $$ -->
 
-where $\hat{\epsilon}_{it}^{\ast,r}$ are resampled with replacement
-from $\{\hat{\epsilon}_{it}\}_{i=1,t=1}^{N,T}$. We do resample the
-residuals $R$ times, repeating the above each time. We then compute
+<!-- where $\hat{\epsilon}_{it}^{\ast,r}$ are resampled with replacement -->
+<!-- from $\{\hat{\epsilon}_{it}\}_{i=1,t=1}^{N,T}$. We do resample the -->
+<!-- residuals $R$ times, repeating the above each time. We then compute -->
 
-$$
-\hat{E}[Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
-\frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}
-$$
+<!-- $$ -->
+<!-- \hat{E}[Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] = -->
+<!-- \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r} -->
+<!-- $$ -->
 
-Finally, we repeatedly draw $\theta$ from its asymptotic distribution
-and repeat all of the above to compute confidence intervals.
+<!-- Finally, we repeatedly draw $\theta$ from its asymptotic distribution -->
+<!-- and repeat all of the above to compute confidence intervals. -->
 
-For case growth, the same procedure is used, but with $\Delta Y_{it} = \Delta \log \Delta C_{it}$ in place of $Y_{it}$. The top panel of the following figure shows
+<!-- For case growth, the same procedure is used, but with $\Delta Y_{it} = \Delta \log \Delta C_{it}$ in place of $Y_{it}$. The top panel of the following figure shows -->
 
-$$
-E[\Delta Y_{it} | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}\}_{t=1}^T]
-$$
+<!-- $$ -->
+<!-- E[\Delta Y_{it} | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}\}_{t=1}^T] -->
+<!-- $$ -->
 
-and
+<!-- and -->
 
-$$
-E[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T]
-$$
+<!-- $$ -->
+<!-- E[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T] -->
+<!-- $$ -->
 
-along with the observed $Y_{it}$, and the bottom panel shows the difference.
+<!-- along with the observed $Y_{it}$, and the bottom panel shows the difference. -->
+
+<!-- In this subsection, instead of resampling residuals, we condition on them. That is we look at -->
+
+<!-- $$ -->
+<!--   Y_{it}^{\ast} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it} -->
+<!-- $$ -->
+
+<!-- instead of -->
+
+<!-- $$ -->
+<!--   \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}. -->
+<!-- $$ -->
 
@@
 if (!file.exists("cfnb.Rda") || rerunsim) {
-  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf))
+  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, nresid=1))
   nball <- rbindlist(lapply(cfnblist, getmeans))
@@
 if (!file.exists("cfmask.Rda") || rerunsim) {
-  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
+  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, nresid=1))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
@@
 
-### Holding Residuals fixed
-
-In this subsection, instead of resampling residuals, we condition on them. That is we look at
-
-$$
-  Y_{it}^{\ast} = {P_{it}^\ast}'\alpha + \sum_{\ell=1}^L \rho_\ell Y_{it-\ell} + X_{it}'\beta + \hat{\epsilon}_{it}
-$$
-
-instead of
-
-$$
-  \frac{1}{R} \sum_{r=1}^R Y_{it}^{\ast,r}.
-$$
-
-```{r masks-resids-fixed}
-simmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
-  simmodel(st, samplecoefs=samplecoefs, nsim=nsim,
-           shockgen=NULL)
-maskmodel1 <- function(st, samplecoefs=FALSE, nsim=1)
-  maskmodel(st, samplecoefs=samplecoefs, nsim=nsim,
-            shockgen=NULL)
-
-rerunsim <- FALSE
-if (!file.exists("cfmask1.Rda") || rerunsim) {
-  cfmasklist1 <- lapply(unique(df$state),
-                        function(s) nopstatesim(s,simobs=simmodel1,simcf=maskmodel1, data=sdf, nresid=1))
-  maskall1 <- rbindlist(lapply(cfmasklist1, getmeans))
-  save(cfmasklist1, maskall1, file="cfmask1.Rda")
-}
-load("cfmask1.Rda")
-cfmask1 <- do.call(rbind, lapply(cfmasklist1, function(x) x$edf))
-
-for (st in c("Washington","Illinois","Massachusetts")) {
-  f <- cfplots(st, cfmask1, yvar="deaths", data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))
-
-}
-
-figs1 <- nationalplots(maskall1, cfmask1, "mandating masks on April 1st\n")
-
-for (f in figs1) print(f)
-```
-
 ## Stay at home
@@
 if (!file.exists("cfshelter.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
+  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
@@
 figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_v1.pdf",sep="/"), width=5,height=3.75)
+figs$r
+dev.off()
+
+for (f in figs) print(f)
+```
+
+## Shutdown policies
+
+We now look at what happens if the four shutdown policies in pindex were never implemented.
+
+```{r shutdown, cache=FALSE}
+cf <- df
+cf[,"pshelter"] <- 0
+cf[,"pmovie"] <- 0
+cf[,"prestaurant"] <- 0
+cf[,"pnonessential"] <- 0
+cf <- subset(cf, cf$date>=datelims[1]) #+L)
+noindex <- createforwardsim(m, cf)
+
+if (!file.exists("cfindex.Rda") || rerunsim) {
+  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf, nresid=1))
+  indexall <- rbindlist(lapply(cfindexlist, getmeans))
+  save(cfindexlist, indexall, file="cfindex.Rda")
+}
+load("cfindex.Rda")
+cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
+```
+
+```{r, cache=TRUE}
+cfindex <- merge(cfindex, abbrv, by="state")
+for (st in c("Washington","Illinois","Massachusetts")) {
+  f <- cfplots(st, cfindex, data=df, yvar="deaths")
+  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdp)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figp)
+  dev.off()
+}
+
+figs <- nationalplots(indexall, cfindex, "no shutdown policies", yvar="deaths")
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth.pdf",sep="/"), width=5,height=3.75)
+figs$g
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases.pdf",sep="/"), width=5,height=3.75)
+figs$c
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel.pdf",sep="/"), width=5,height=3.75)
 figs$r
@@
 
-### Removing Policies
+## Removing Policies
 
@@
 
-## National
+### National
 
@@
 
-## Non Essential Business Closures
+## Non-essential Business Closures
 
@@
 ```
+
+## Shutdown policies
+
+We now look at what happens if the four shutdown policies in pindex were never implemented.
+
+```{r shutdown-deaths, cache=FALSE}
+cf <- df
+cf[,"pshelter"] <- 0
+cf[,"pmovie"] <- 0
+cf[,"prestaurant"] <- 0
+cf[,"pnonessential"] <- 0
+cf <- subset(cf, cf$date>=datelims[1]) #+L)
+noindex <- createforwardsim(m, cf, dvar="dlogdd", var="logdd")
+
+if (!file.exists("cfindex_deaths.Rda") || rerunsim) {
+  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
+  indexall <- rbindlist(lapply(cfindexlist, getmeans))
+  save(cfindexlist, indexall, file="cfindex_deaths.Rda")
+}
+load("cfindex_deaths.Rda")
+cfindex <- do.call(rbind, lapply(cfindexlist, function(x) x$edf))
+```
+
+```{r, cache=TRUE}
+cfindex <- merge(cfindex, abbrv, by="state")
+for (st in c("Washington","Illinois","Massachusetts")) {
+  f <- cfplots(st, cfindex, data=df, yvar="deaths")
+  print(f$figp +  ggtitle(sprintf("Effect of no shutdown policies in %s",st)))
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dgrowth_deaths.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdp)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcases_deaths.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figp)
+  dev.off()
+}
+
+figs <- nationalplots(indexall, cfindex, "no shutdown policies", yvar="deaths")
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dgrowth_deaths.pdf",sep="/"), width=5,height=3.75)
+figs$g
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-dcases_deaths.pdf",sep="/"), width=5,height=3.75)
+figs$c
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-rel_deaths.pdf",sep="/"), width=5,height=3.75)
+figs$r
+dev.off()
+
+for (f in figs) print(f)
+```

### 2020-08-14 cases_and_policies/rmd/regressionsWithPlots.Rmd

--- 2020-08-13 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
     ti <- t[id==i]
-    xi[is.na(xi)] <- approx(ti[!is.na(xi)], xi[!is.na(xi)], ti[is.na(xi)])$y
+    xi[is.na(xi)] <- approx(ti[!is.na(xi)], xi[!is.na(xi)], ti[is.na(xi)], rule=2)$y
     x[id==i] <- xi
@@
 df$testrate  <- df$tests/df$Population.2018*1000
-df$testrate[df$testrate<0] <- NA
+#df$testrate[df$testrate<0] <- NA
 df$dlogtests <- paneldiff(log(sapply(df$testrate, function(x)
@@
 xlist <-list(c("logvote"), c("logvote"))
-L <- L.c
+L <- 14
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
@@
   piy[[(1+4*(k-1))]] <- m
-  m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))]  
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(2+4*(k-1))]] <- m
@@
 prefix="cases-fe"
-#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe") copying and
+
+# pasting this is bad style and seems like it might create maintenance
+# problems ...
    ylbl <- colnames(piy[[1]]$reg$response)
@@
   piy[[(1+4*(k-1))]] <- m
-  m$reg$beta[1:(8+2*(k-1))] <- m$bc2[1:(8+2*(k-1))]  
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(2+4*(k-1))]] <- m
@@
   piy[[(3+4*(k-1))]] <- m
-  m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))]  
+  m$reg <- replacecoef(m$reg, m$bc2)
   piy[[(4+4*(k-1))]] <- m

### 2020-08-14 tex/covidEffectOfNPI-revision.tex

@@
 Table \ref{tab:policies_inreg} provides summary statistics, where $N$ is the number of states that have ever implemented the policy.
-We also obtain information on state-level covariates from \citet{raifman2020}, which include population size, total area, unemployment rate, poverty rate, and a percentage of people who are subject to illness.These confounders are motivated by \cite{wheaton2020} who finds that case growth is associated  with residential density and per capita income.
+We also obtain information on state-level covariates mostly from \citet{raifman2020}, which include population size, total area, unemployment rate, poverty rate,  a percentage of people who are subject to illness, and state governor's party affiliations.
+These confounders are motivated by \cite{wheaton2020} who finds that case growth is associated  with residential density and per capita income.
 
@@
          a pair of policy and behavior variables.
+``ave of four policy vars'' is defined by the average of stay-at-home, closured movie theaters, closured restaurants, and closured non-essential businesses.
        \end{flushleft}
@@
 problems and could limit our ability to separately identify the
-effect of each policy or behavior variable on case growth, but this
-does not prevent us from identifying the aggregate effect of all
-policies and behavior variables on case or death growth.
+effect of each policy or behavior variable on case growth. 
+%, but this does not prevent us from identifying the aggregate effect of all policies and behavior variables on case or death growth. 
+For this reason, we also consider a specification that includes
+the average of stay-at-home, closured movie theaters, closed  restaurants, and closured non-essential businesses in place of these four policy variables separately.
 
@@
         Intensity" defined as 7 days moving averages of corresponding daily measures obtained from Google Mobility Reports.
-All specifications include state-level characteristics (population, area, unemployment rate, poverty rate, and a percentage of people subject to illness) as well as their interactions with the log of days since Jan 15, 2020.
+All specifications include state-level characteristics (population, area, unemployment rate, poverty rate,  a percentage of people subject to illness, and governor's party) as well as their interactions with the log of days since Jan 15, 2020.
 The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients.
@@
           Intensity" defined as 7 days moving averages of corresponding daily measures obtained from Google Mobility Reports.
-All specifications include state-level characteristics (population, area, unemployment rate, poverty rate, and a percentage of people subject to illness) as well as their interactions with the log of days since Jan 15, 2020.
+All specifications include state-level characteristics (population, area, unemployment rate, poverty rate,  a percentage of people subject to illness, and governor's party) as well as their interactions with the log of days since Jan 15, 2020.
 The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients.
@@
      weighted by the average of each behavioral variable from April
-     1st-10th.  Standard errors are clustered at the state level.
+     1st-10th.
+``ave of four policy vars'' is defined by the average of stay-at-home orders, closured movie theaters, closured restaurants, and closured non-essential businesses.
+Standard errors are clustered at the state level.
    \end{flushleft}
@@
   a lag length of 14 days between exposure and case reporting, as well as a lag length of 21 days between exposure and deaths, is broadly consistent with currently
-  available evidence.  }   $\bcolor B_{it}=(B_{it}^1,...,B_{it}^4)'$ is a vector of four behavior
+  available evidence. Section \ref{sec:sensitivity} provides a sensitivity analysis under different timing assumptions.  }   $\bcolor B_{it}=(B_{it}^1,...,B_{it}^4)'$ is a vector of four behavior
 variables  in state $i$. $\pcolor P_{it}$ includes the Covid-related policies
@@
 (\ref{eq:reg-y}) for case and death growth rates. %, respectively.
-Column (1) represents our baseline specification while column (2) allows the effect of masks to be different before and after May 1st.
+
+Column (1) represents our baseline specification while column (2) replaces stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses with the average of these four policy variables.
 Columns (3) and (4) include past cases/deaths and growth rates at national level as additional regressors.
@@
 
-Except for mask requirements, policies appear to have little direct
-effect on case or death growth when behavior is held constant. The one
-exception is that closing schools has a large and statistically
+The estimates also imply that stay-at-home orders reduce the growth rate of cases and deaths by 8-10 percent directly.
+This can be a reflection of multi-colinearity problem given high correlations among policy variables; in columns (2) and (4), the estimates of the average of stay-at-home orders, closed movie theaters, closed restaurants, and closed non-essential businesses are not significantly zero, indicating that the four policies in total may have little direct
+effect on case or death growth when behavior is held constant.
+
+Closing schools has a large and statistically
 significant coefficient in the death growth regressions. As discussed
-above, there is little cross-state variation in the timing of school
+above, however, there is little cross-state variation in the timing of school
 closures, making estimates of its effect less reliable.
@@
 variables from April 1st-10th.
-The estimates of $-0.76$ and $-0.87$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 76 and 83 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
+The estimates of $-0.73$ and $-0.79$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 73 and 79 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
 When including national cases and deaths in information, as shown in columns (3) and (4), the estimated aggregate impact of behavior is substantially smaller, but remains large and statistically significant.
@@
      ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
-     coefficients.  Standard errors are clustered at the state level.
+     coefficients.
+``ave of four policy vars'' is defined by the average of stay-at-home orders, closured movie theaters, closured restaurants, and closured non-essential businesses.
+Standard errors are clustered at the state level.
    \end{flushleft}
@@
 that all policies combined would reduce the growth rate of cases  and deaths by
-0.70 and 0.98, respectively,   out of which about two-third to
-three-fourth is attributable to the indirect effect through their impact on behavior.  The estimate also indicates that the
+0.75 and 1.01, respectively,   out of which more than one-half to
+two-third is attributable to the indirect effect through their impact on behavior.  The estimate also indicates that the
 effect of mandatory masks for employees is mostly direct.
@@
 
-In columns (2) and (4) of Table \ref{tab:PtoY}, we find that the estimated effect of mask mandates in April is larger than that in May for both case and death regressions.
+%In columns (2) and (4) of Table \ref{tab:PtoY}, we find that the estimated effect of mask mandates in April is larger than that in May for both case and death regressions.
 This may reflect a wider \textit{voluntary} adoption of masks in May than in April --- if more people wear masks even without mandatory mask policy,  the policy effect of mandating masks for employees becomes weaker.
@@
 Turning to the results, the estimates of Tables \ref{tab:dieff-si} and \ref{tab:dieff}  imply that all policies combined would reduce
-\(\Delta \log \Delta D\) by  0.97 and  0.40, respectively. For
+\(\Delta \log \Delta D\) by  0.97 and  0.48, respectively. For
 comparison, the median of \(\Delta \log \Delta D_{it}\) reached its
 peak in mid-March of about 1.3 (see Figure \ref{fig:growthq} in the
-appendix). Since then it has declined to near 0. Therefore,  -0.97 and -0.40 imply
+appendix). Since then it has declined to near 0. Therefore,  -0.97 and -0.48 imply
 that policy changes can account for roughly one-third to two-third of the observed
@@
 
-\section{Sensitivity Analysis}
-
-In this section, we provide sensitivity analysis by estimating (\ref{eq:R4}) with alternative specifications and methods as follows:
+\section{Sensitivity Analysis}\label{sec:sensitivity}
+
+In this section, we provide sensitivity analysis by estimating (\ref{eq:R4}) with alternative specifications and methods.
+Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses for the following specifications and estimation methods:
   \begin{itemize}
-  \item Exclude  the state of New York from the sample because it may be viewed as an outlier in the early pandemic period.
-  \item Add private mask wearing rates from survey in March and April as confounder for unobserved  personal risk-aversion and initial attitude toward mask wearing.
-   \item Add the log of Trump's vote share  in 2016 presidential election as confounder for unobserved private behavioral response.
-   \item Add  past behavior variables  as information used to set policies.
-   \item All of the above, estimated using linear regression model.
-   \item All of the above, estimated by Double Machine Learning (DML) with Lasso to reduce dimensionality.
-   \item All of the above, estimated by  DML with Random Forest to reduce dimensionality and capture some nonlinearities.
-   \item Latent state confounders as components of $W_{it}$, estimated using fixed effects with bias correction.
-   \item Alternative Timing assumption ($\ell=10$ for cases and $\ell=23$ for deaths) on all of the above.\footnote{This alternative timing assumption is motivated by the estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
+  \item[(1)] Baseline specification  in Table  \ref{tab:PtoY}.
+  \item[(2)]  Exclude  the state of New York from the sample because it may be viewed as an outlier in the early pandemic period.
+  \item[(3)]   Add  the percentage of people who wears masks to protect themselves in March and April as confounder for unobserved  personal risk-aversion and initial attitude toward mask wearing.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.}
+   \item[(4)]   Add the log of Trump's vote share  in 2016 presidential election as confounder for unobserved private behavioral response.
+   \item[(5)]   Add  past behavior variables  as information used to set policies.
+Under this specification, our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables.
+   \item[(6)]   Include all additional controls in (3)-(5) with the sample that excludes New York as in (2).
+   \item[(7)]   Estimated by Double Machine Learning (DML) of \cite{chernozhukov18} with Lasso to reduce dimensionality while adding private mask wearing rates and  the log of Trump's vote share as additional controls.
+\item[(8)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while adding private mask wearing rates and  the log of Trump's vote share as additional controls.
+   \item[(9)]   Add latent state confounders as components of $W_{it}$, estimated using fixed effects with bias correction.
+  % \item Alternative Timing assumption ($\ell=10$ for cases and $\ell=23$ for deaths) on all of the above.
+  \end{itemize} 
+In Figure \ref{fig:whisker}, ``red'', ``green'', ``blue'', and ``purple" indicate the regression models for case growth without national information variables, case growth with national information variables, death growth without national information variables, and death growth with national information variables, respectively.
+The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 9 and 24 days, respectively, in the right panel of ``(ii) alternative timing.''\footnote{This alternative timing assumption is motivated by the lower bound estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
 We thank Jessica Metclaf for recommending us to do a sensitivity analysis on the timing assumption while suggesting this reference to us.}
-  \end{itemize}
-
-%We examine the robustness of our results by adding more controls, excluding the state of New York, and applying  estimation methods of double/debiased machine learning (DML) of \cite{chernozhukov18} and bias corrected fixed effects estimator.
-We also provide a sensitivity analysis regarding the assumption on the times from exposure to case confirmation and death reporting.
-
-Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses across different specifications, estimation methods, and timing assumptions.
-Here, ``red'', ``green'', ``blue'', and ``purple" indicate the regression models for case growth without national information variables, case growth with national information variables, death growth without national information variables, and death growth with national information variables, respectively.
-The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 10 and 23 days, respectively, in the right panel of ``(ii) alternative timing.''
-
-In each of four regression models, we report: (1) baseline specification  in Table xx; (2) excluding New York from the sample; (3) adding  the percentage of people who wears masks to protect themselves between March 26-April 29, 2020 to control for people's attitude toward mask wearing in the first half of sample;\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
-The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.} (4) adding the log of Trump voting shares;   (5)  including four weeks lagged behavior variables as an additional set of controls, where our causal interpretation is valid when policy variables are sufficiently random conditional on past behavior variables under this specification; (6) including  all additional controls in (3)-(5) with the sample that excludes New York; (7) applying the DML  with Lasso to (6); (8) applying the DML with Random Forest to (6); (9) including state fixed effects while applying bias correction based on cross-splitting.\footnote{In the appendix, Table \ref{tab:PtoY-robust} presents the estimates for (1)-(6) while Table \ref{tab:PtoY-fe} compares bias corrected fixed effect estimates with those without bias correction.
-The appendix also provides more detailed discussion on how we implement DML procedures as well as bias correction to fixed effects estimator.}
 
@@
  except for fixed effects estimator with bias correction.
-The sensitivity under fixed effects estimator may be due to high correlations between stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses (see Table \ref{tab:correlation}).
+The sensitivity under fixed effects estimator may be due to high correlations between stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses.
 As shown in Panel (D), the  estimated coefficient of the average of  stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses   are  negative and often significant even under fixed effect specification; similar to school closure estimates, however,  its estimate is sensitive to controlling for national cases and deaths because the timing of implementing these policies is similar across states.
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-10} 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-9} 
     \end{tabular}
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-10}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-9}
           \end{tabular}
@@
       $^\dagger$The times from exposure to case confirmation and death reporting  are assumed to be 14 and 21 days, respectively.
-$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 10 and 23 days, respectively.
+$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 9 and 24 days, respectively.
     \end{flushleft}
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-10} 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-9} 
     \end{tabular}
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-10}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-9}
           \end{tabular}
@@
 
-The effect of removing all policies includes the effect of school
-closures. The visual evidence on growth rates for states with and without school closures,
-presented blow, suggest that there may be a potentially large effect, though the history is very short.
+[Discussion on the effect of school closures has been moved to sensitivity analysis section.]
+
+%The effect of removing all policies includes the effect of school
+%closures. The visual evidence on growth rates for states with and without school closures,
+%presented blow, suggest that there may be a potentially large effect, though the history is very short.
 The main results presented in Section 3 also support the hypothesis that the school closures were important at lowering the growth rates.
 This evidence is consistent
-with the emerging evidence of prevalence of Covid-19 among children
-aged 10-17. \cite{children:nature} find that although children's
-transmission and susceptibility rates are half that of ages 20-30,
-children's contact rates are much higher.
-This type of evidence, as well as, evidence that children carry viral
-loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.
-
-\begin{figure}
-  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
-      &
-        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
-    \end{tabular}
-    \footnotetext{ In these figures, red points are the case or death
-      growth rate in states without each policy 14 (or 21 for deaths)
-      days earlier. Blue points are states with each policy 14 (or 21
-      for deaths) days earlier. The red line is the average across
-      states without each policy. The blue line is the average across
-      states with each policy.}
-  \end{minipage}
-\end{figure}
-
-As discussed above, there is little variation across states
-in the timing of school closures. Consequently, the effect of school
-closures is difficult to identify statistically, because it is hard to separate it from aggregate time effect, and its estimate is sensitive to an inclusion of some aggregate variables such as national cases.
+%with the emerging evidence of prevalence of Covid-19 among children
+%aged 10-17. \cite{children:nature} find that although children's
+%transmission and susceptibility rates are half that of ages 20-30,
+%children's contact rates are much higher.
+%This type of evidence, as well as, evidence that children carry viral
+%loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.
+%
+%\begin{figure}
+%  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
+%  \begin{minipage}{\linewidth}
+%    \centering
+%    \begin{tabular}{cc}
+%       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
+%      &
+%        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
+%    \end{tabular}
+%    \footnotetext{ In these figures, red points are the case or death
+%      growth rate in states without each policy 14 (or 21 for deaths)
+%      days earlier. Blue points are states with each policy 14 (or 21
+%      for deaths) days earlier. The red line is the average across
+%      states without each policy. The blue line is the average across
+%      states with each policy.}
+%  \end{minipage}
+%\end{figure}
+
+%As discussed above, there is little variation across states
+%in the timing of school closures. Consequently, the effect of school
+%closures is difficult to identify statistically, because it is hard to separate it from aggregate time effect, and its estimate is sensitive to an inclusion of some aggregate variables such as national cases.
 To support this point, Figure \ref{fig:US-nop} shows the
-effect of removing all policies on cases based on the estimates with national cases
-included as information. When national case variables are included
-in the specification, the estimated effect of school closures, and hence that of removing all
-policies, is much smaller with a 90\% confidence interval of [0,10] fold increases.
-
-Given this sensitivity, we conclude that there still exists a lot of uncertainty as to the effect of removing all policies, especially schooling.
+%effect of removing all policies on cases based on the estimates with national cases
+%included as information. When national case variables are included
+%in the specification, the estimated effect of school closures, and hence that of removing all
+%policies, is much smaller with a 90\% confidence interval of [0,10] fold increases.
+%
+%Given this sensitivity, we conclude that there still exists a lot of uncertainty as to the effect of removing all policies, especially schooling.
 The impact of not implementing any policies on cases and deaths can be quite large, but  the effect of school closures, hence that of removing all policies, is not well identified statistically from the US state-level data alone, because of the lack of cross-sectional variations.
@@
 An important research question is how to resolve this uncertainty using additional data sources.
-
-\begin{figure}[ht]
-  \caption{Effect of removing policies on cases in the US under a specification with only state-level cases/deaths as information \label{fig:US-nop-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths_v1}
+%
+%\begin{figure}[ht]
+%  \caption{Effect of removing policies on cases in the US under a specification with only state-level cases/deaths as information \label{fig:US-nop-SI}}
+%  \begin{minipage}{\linewidth}
+%    \centering
+%    \begin{tabular}{cc}
+%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
 %      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
-  \caption{Effect of removing policies on cases in the US under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nop}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
+%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
+%%      \\
+%%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths_v1}
+%%      &
+%%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths_v1}
+%    \end{tabular}
+%  \end{minipage}
+%\end{figure}
+%
+%\begin{figure}[ht]
+%  \caption{Effect of removing policies on cases in the US under a specification with both
+%  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nop}}
+%  \begin{minipage}{\linewidth}
+%    \centering
+%    \begin{tabular}{cc}
+%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth}
 %      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
+%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel}
+%%      \\
+%%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
+%%      &
+%%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
+%    \end{tabular}
+%  \end{minipage}
+%\end{figure}
 

### 2020-08-14 tex/tables_and_figures/policies 2.tex

% latex table generated in R 4.0.0 by xtable 1.8-4 package
% Thu Aug 13 10:34:03 2020
\begin{table}[ht]
\centering
\begin{tabular}{rllll}
  \hline
 & N & Min & Median & Max \\ 
  \hline
State of emergency & 49 & 2020-02-29 & 2020-03-11 & 2020-03-16 \\ 
  Date closed K 12 schools & 49 & 2020-03-13 & 2020-03-17 & 2020-04-03 \\ 
  Closed day cares & 14 & 2020-03-16 & 2020-03-23 & 2020-04-06 \\ 
  Date banned visitors to nursing homes & 29 & 2020-03-09 & 2020-03-16 & 2020-04-06 \\ 
  Closed non essential businesses & 41 & 2020-03-19 & 2020-03-25 & 2020-04-06 \\ 
  Closed restaurants except take out & 46 & 2020-03-15 & 2020-03-17 & 2020-04-03 \\ 
  Closed gyms & 44 & 2020-03-16 & 2020-03-20 & 2020-04-03 \\ 
  Closed movie theaters & 47 & 2020-03-16 & 2020-03-21 & 2020-04-06 \\ 
  Stay at home  shelter in place & 40 & 2020-03-19 & 2020-03-28 & 2020-04-07 \\ 
  End relax stay at home shelter in place & 36 & 2020-04-24 & 2020-05-15 & 2020-06-19 \\ 
  Began to reopen businesses statewide & 48 & 2020-04-20 & 2020-05-06 & 2020-06-05 \\ 
  Reopen restaurants & 46 & 2020-04-24 & 2020-05-17 & 2020-06-22 \\ 
  Reopened gyms & 42 & 2020-04-24 & 2020-05-18 & 2020-07-03 \\ 
  Reopened movie theaters & 35 & 2020-04-27 & 2020-05-30 & 2020-07-03 \\ 
  Resumed elective medical procedures & 34 & 2020-04-20 & 2020-04-29 & 2020-05-29 \\ 
  Mandate face mask use by all individuals in public spaces & 30 & 2020-04-08 & 2020-06-22 & 2020-07-27 \\ 
  Mandate face mask use by employees in public facing businesses & 40 & 2020-04-03 & 2020-05-01 & 2020-07-03 \\ 
  Stop Initiation of Evictions overall or due to COVID related issues & 29 & 2020-03-16 & 2020-03-24 & 2020-05-30 \\ 
  Stop enforcement of evictions overall or due to COVID related issues & 27 & 2020-03-17 & 2020-03-24 & 2020-06-08 \\ 
  Renter grace period or use of security deposit to pay rent & 3 & 2020-04-10 & 2020-04-24 & 2020-05-20 \\ 
  Order freezing utility shut offs & 32 & 2020-03-12 & 2020-03-19 & 2020-04-13 \\ 
  Froze mortgage payments & 1 & 2020-03-21 & 2020-03-21 & 2020-03-21 \\ 
  Waived one week waiting period for unemployment insurance & 35 & 2020-03-08 & 2020-03-18 & 2020-04-06 \\ 
   \hline
\end{tabular}
\caption{State Policies \label{tab:policies}} 
\end{table}

### 2020-08-14 cases_and_policies/rmd/cf-pindex.Rmd

@@
 source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
-pols <- c("pmaskbus","pk12","pindex")
+pols <- c("pmaskbus","pk12","pshelter","pindex")
 ```
@@
 
+## Stay at home
+
+We now look at what happens if stay at home orders were never issued.
+
+```{r home, cache=FALSE}
+cf <- df
+cf[,"pshelter"] <- 0
+cf <- subset(cf, cf$date>=datelims[1]) #+L)
+noshelter <- createforwardsim(m, cf)
+
+if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
+  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
+  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
+  save(cfshelterlist, shelterall, file="cfshelter_idx.Rda")
+}
+load("cfshelter_idx.Rda")
+cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
+```
+
+```{r, cache=TRUE}
+cfshelter <- merge(cfshelter, abbrv, by="state")
+for (st in c("Washington","Illinois","Massachusetts")) {
+  f <- cfplots(st, cfshelter, data=df)
+  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdp)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figp)
+  dev.off()
+}
+
+figs <- nationalplots(shelterall, cfshelter, "no stay at home orders")
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
+figs$g
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_idx.pdf",sep="/"), width=5,height=3.75)
+figs$c
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_idx.pdf",sep="/"), width=5,height=3.75)
+figs$r
+dev.off()
+
+for (f in figs) print(f)
+```
+
 ## Averaged policies
 
-We now look at what happens if none of the four averaged policies
-(stay at home, close movie theaters, close restaurants, close
+We now look at what happens if none of the three averaged policies
+(close movie theaters, close restaurants, close
 nonessential businesses) were implemented.
 
-```{r home, cache=FALSE}
+```{r index, cache=FALSE}
 cf <- df
@@
 
+## Stay at home
+
+We now look at what happens if stay at home orders were never issued.
+
+```{r home-deaths, cache=FALSE}
+cf <- df
+cf[,"pshelter"] <- 0
+cf <- subset(cf, cf$date>=datelims[1]) #+L)
+noshelter <-  createforwardsim(m, cf, dvar="dlogdd", var="logdd")
+
+if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
+  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
+  shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
+  save(cfshelterlist, shelterall, file="cfshelter_deaths_idx.Rda")
+}
+load("cfshelter_deaths_idx.Rda")
+cfshelter <- do.call(rbind, lapply(cfshelterlist, function(x) x$edf))
+```
+
+```{r, cache=TRUE}
+cfshelter <- merge(cfshelter, abbrv, by="state")
+for (st in c("Washington","Illinois","Massachusetts")) {
+  f <- cfplots(st, cfshelter, data=df, yvar="deaths")
+  print(f$figp +  ggtitle(sprintf("Effect of no stay at home order in %s",st)))
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dgrowth_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdp)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcases_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figp)
+  dev.off()
+}
+
+figs <- nationalplots(shelterall, cfshelter, "no stay at home orders",yvar="deaths")
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figs$g
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figs$c
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figs$r
+dev.off()
+
+for (f in figs) print(f)
+```
+
 ## Shutdown policies
 
-We now look at what happens if the four shutdown policies in pindex were never implemented.
+We now look at what happens if the three business related shutdown policies in pindex were never implemented.
 
-```{r home-deaths, cache=FALSE}
+```{r index-deaths, cache=FALSE}
 cf <- df

### 2020-08-14 cases_and_policies/rmd/regressionsWithPlots.Rmd
cleaned up incomplete merge.

--- 2020-08-14 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 cmat <-  cor(df[,c(bvars,pols,"pindex")], use="pairwise.complete.obs")
-scmat <-
-  matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)+1))
+scmat <- matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)+1))
 
@@
 regs <- mainregressions(sdf,
-                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L) 
+                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L,
+                        sameobservations=FALSE)
 showhtmltables(regs$pib[[1]], regs$pbiy, regs$piy);
@@
 
-df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/4
+df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/3
 
@@
 ## DML
-```{r regsdml,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsdml,  eval = TRUE, cache=TRUE, warning=FALSE, results='asis'}
 
@@
 xvars.national.d <- c(bvars.d,"dlogdd.L21","logdd.L21","dlogdd.national.L21","logdd.national.L21")
+
+pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pmovie.L14","prestaurant.L14","pnonessential.L14")
+pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pmovie.L21","prestaurant.L21","pnonessential.L21")
 

### 2020-08-21 tex/reply-to-referees.tex
update

@@
 
+ This point is well taken. We added the following paragraph to Section 2.2. 
+ \begin{quote}
+ Validation of the model by (TR) allows us to check exclusion restrictions brought by contextual knowledge, and check stability of the model by using different data subsets.
+However, passing the (TR) does not, of course, guarantee that the model is necessarily valid for recovery of causal effects.
+The only fundamental way to truly validate a causal model for observational data is through a controlled experiment, which is impossible to carry in our setting.
+We shall however cross-check our results against other scientific evidence and, for example, results from RCT on effectiveness of masks on respiratory deceases (though not in the Covid-19 context).
+ \end{quote}
+
 \item[1.]   \textit{(Source of state-level variation) ``The main specification is based on random effects panel data
@@
 
-We chose information structure that is parsimonious and yet we believe captures the most important part of information that is relevant for people's behavior.
+We chose information structure that is parsimonious and yet, we believe, captures the  information that is most relevant for people's behavior.
 
-We think that conditioning on current policy variables that is included in all specifications, lagged values of policies will  add little additional information and it is natural to exclude lagged values of policies from the information structure --- or we may re-interpret that the coefficient of current policy variables partially reflects its effect through providing information.
+Conditioning on current policy variables that is included in all specifications, the lagged values of policies will  likely add little additional information and, therefore, we decided to exclude lagged values of policies from the information structure.
+We may re-interpret that the coefficient of current policy variables partially reflects its effect of providing information.
 
-As suggested, it is possible to add lagged values of policies and behavior as part of information.
-If we specify that information structure contains two weeks lagged behavior variables, then a specification of $(PI\rightarrow Y)$ equation contains four weeks lagged behavior variables as additional regressors.
-As a sensitivity check, we estimated this specification with lagged behavior variables in Section 5.
+On the other hand, the lagged values of  behaviors may be a part of the information structure.
+As a sensitivity check, we estimated a specification that includes  lagged behavior variables as information variables in Section 5.
 As shown in specification (5) of Figures xx-xx, the estimated coefficients of mask for employees, closed K-12 schools, stay-at-home orders, and the average of stay-at-home orders and closures of movie theaters, restaurants, and non-essential businesses are not so sensitive to an inclusion of lagged behavior variables as information.
@@
 helpful.''}
+
+In our view, people's mobility decisions are largely based on the current information and people's expectation of future cases/deaths is not the first order importance when people make their mobility decisions because there is no obvious inter-temporal substitution with respect to their mobility decisions.
+
+Also, given a lack of expectation data, statistically testing whether expectation over future cases/deaths is important as a determinant of people's behavior is difficult in this context.
+For example, including future realized cases/deaths as an additional regressor in the behavior equation leads to the positive estimate because of the reverse causality, i.e., an increase in mobility leads to an increase in future cases.
 
@@
 ordering.''}
+
+Thank you for your comment.
+As a part of sensitivity checks, we now examine if the estimates are sensitive when we include past behavior variables as information variables, under which policies depend on past behavior.
+We find that the estimated coefficients of masks for employees are not sensitive to an inclusion of past behavior variables.
 
@@
 
+We now present the results of the debiased fixed effect estimates in Table 8 in Section 4.2.
+
 \item[8.]  \textit{(Random effects vs. fixed effects) ``In addition to the previous comment, another reason for
@@
 errors.''}
+
+Following your suggestion, we have added a comment on this on page xx in Section 4.2.
 
@@
 Probably I am mistaken. It would be helpful if the authors could clarify this.''}
+
+In columns (1)-(4) of Table 3(A), the estimated coefficients on $\log\Delta C_{it}$ is negative and the sum of estimated coefficients of $\Delta \log\Delta C_{it} $ and $\log\Delta C_{it}$ are also negative except for grocery.
+
+To interpret these results, rewrite a regression specification after omitting other variables as 
+$B_{it} =  \gamma_1 \Delta \log\Delta C_{it} + \gamma_2   \log\Delta C_{it} =  (\gamma_1 +\gamma_2) \log\Delta C_{it} - \gamma_1   \log\Delta C_{i,t-7}$.
+In columns (1)-(4) of Table 3(A),  the estimated values of both $(\gamma_1 +\gamma_2)$ and $-\gamma_1$ are negative except for grocery.
+This suggests that  a higher \textit{level} of confirmed cases reduces people's mobility in workplaces, retails, and transit.
+For grocery, the positive estimated coefficient of $(\gamma_1 +\gamma_2)$ may reflect stock-piling behavior in early pandemic periods.
+
+ We also note that the estimated coefficients of  $\Delta \log\Delta D_{it} $ and $\log\Delta D_{it}$ are negative in columns (1)-(4) of  Table 3(B), consistent with a hypothesis that a higher number of deaths as well as a higher growth rate of deaths reduces people's mobility.
+
+Following your suggestion, we added the above discussion in footnote 20 (check!) when we discuss in Table 3A.
+
 \item[11.]  \textit{Figure 8 is very informative. Perhaps the authors could show this in introduction as well.''}
@@
 
-\item       
+\item  \textit{``My sense is that the delay between infection and case reporting should be closer to a week (more in specific comments below) and various lines of evidence suggest two waves of death following infection, one early and one late (e.g., in https://science.sciencemag.org/content/369/6500/208: "We identify two clear subpopulations among hospitalized cases: individuals that die quickly after hospital admission (15\% of fatal cases, with a mean time to death of 0.67 days) and individuals who die after longer time periods (85\% of fatal cases, with a mean time to death of 13.2 days) (fig.
+S3).").
+Presumably the preoponderance of weight at the later time should minimize the effects on your estimates...
+I think a sensitivity analysis to this would be desirable - as well as a reference to some of the literature beyond that written by economists...
+My sense is that the number is closer to 6 days or a week?
+But I have not reviewed that literature recently (Table 2 in this seems to suggest that aligns...
+https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html)"}
+
+Following your suggestion,   we provide a sensitivity analysis on the delay between infection and case reporting (or death reporting) in Section 4.2.
+
+\item \textit{``the collinearities, and lack of agreement in cases/deaths makes the separate estimates for different policies really rather uncertain?
+We would all love a crisp statement about the effects of masks....
+but I wonder if Figure 1 gives a sufficiently robust sense of the uncertainty  based on the estimates provided?''}
+
+Given  your suggestion, Figure 9 presents whisker-dot-plots for confidence intervals of mask coefficients (as well as other policy variables) across different specifications, timing assumptions, and estimation methods.
+We believe this figure will provide a sufficient robust sense of the uncertainty on the  estimates.
+
+\item \textit{`` Figure 1 - are the points on the left hand figures the state specific estimates?
+or?
+What is the blue enveloppe?
+Mildly distracting having nothing happen till halfway through April (I know this is purely illustration, but an 'all else equal' in the caption might help?) And then, for what happens next, crudely, log(It/It-1) is diminshed by a factor by masks, etc, and the lag is simply due to the delays until cases?
+''}
+
+\item \textit{`` p5 - perhaps some papers from beyond the Econ literature would be relevant for a sense of the status of our knowledge of the epidemiogogy?  ''}
+
+\item \textit{`` p6 - simulations would produce exponential spread of disease - well, yes?
+The mechanism is reasonably well described, and the conditions to elude exponential growth are fairly crisp - even if you include 'a voluntary behavioral response' the outcome will be exponential growth unless you are in a regime where susceptible depletion is dominating dynamics, or you have shifted Re<1.
+''}
+
+\item \textit{`` p20 log delta C\_it is obviously strongly autocorrelated (on top of the other collinearities detailed) - I also wondered if time since start of the outbreak might not do just as well at capturing the dirve in change in behavior?
+Do numbers of cases and deaths do better?
+''}
+
+\item \textit{`` Figure 8 - I do like seeing the data, but it also just raises questions, as you say, confounders + many other things at play....
+I wondered if there would be a way to create a visual that captured a bit more of this, and pinpointed the anomalies which are where the inferential power lies.
+''}
+
 \end{enumerate}

### 2020-08-21 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

--- 2020-08-14 cases_and_policies/rmd/regressionsWithPlots.Rmd
@@
 
+df$logtests <- log(sapply(df$testrate, function(x)
+  max(x,exp(-1))))
+
 df$lagtests <- panellag(df$dlogtests, df$state, df$date, lag=1)  
@@
 #infovars2 <- list(c( "logdc"))
-tvars <-  "dlogtests"  
+# tvars <-  "dlogtests"  
 # xlist <-list("party", "party") 
 # xlist <-list(c("party","logvote"), c("party","logvote")) 
+# xlist <- list("week","week")
+
 xlist <- list("","")
@@
 ilist <- list(interactions, interactions)
-iv <- list("0","0") #,"(testratedc ~ lag(testratedc,14))")
+# iv <- list("(dlogtests ~ lag(logtests,7))","(dlogtests ~ lag(logtests,7))")
+# tvars <- ""
+iv <- list("0","0")
+tvars <- "dlogtests"
+
+pols = c("pmaskbus","pk12","pshelter","pindex")
+#pols = c("pmaskbus","pmaskall","pk12","pshelter","pindex")
 
 # Set time lag:
+timing <- 0 # 0 or 1
+if (timing==0){
   L.c <- 14
   L.d <- 21
-# L.c <- 9
-# L.d <- 24
+} else if (timing==1) {
+  L.c <- 7
+  L.d <- 24
+}
 
 L <- L.c
-sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+sdf <- subset(df, df$date>=as.Date("2020-03-07"))
 regs <- mainregressions(sdf,
                         "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
+summary(regs$piy[[1]]$reg)
 ```
@@
 ```{r casedieff, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+
 for (use.national in c(TRUE,FALSE)) {
@@
                        sep="/")) 
-
+  kable(tbl)
+}
+```
+
+```{r casedieff2, eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+
+# source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
+# source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
+
+# pols <- c("pmaskbus","pk12","pshelter","pindex")
+L <- L.c
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+regs <- mainregressions2(sdf,
+                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
+for (use.national in c(TRUE,FALSE)) {
+  sb <- ifelse(use.national, 2, 1)
+  sy <- ifelse(use.national, 3, 1)
+  pib <- lapply(regs$pib[[sb]][1:4], function(x) x$reg)
+  pbiy <- regs$pbiy[[sy]]$reg
+  piy <- regs$piy[[sy]]$reg
+
+  for (i in 1:4) {
+    coef <- pib[[i]]$coefficients
+    coef[is.na(coef)] <- 0
+    pib[[i]]$coefficients <- coef
+    pib[[i]]$beta <- coef
+  }
+  coef <- pbiy$coefficients 
+  coef[is.na(coef)] <- 0
+  pbiy$coefficients <- coef
+  pbiy$beta <- coef
+
+  S <- 999
+  bs <- bootstrap_felm2(sdf, c(pib,list(pbiy,piy)), S=S)
+  for (i in 1:4){
+    bs[[i]][2,]<-0.0
+  }
+  bs[[5]][5,]<-0.0
+
+  addse <- function(di, bdi) {
+    se <- di*NA
+    for (i in 1:length(se)) {
+      se[i] <- sd(sapply(bdi, function(b) b[i]))
+    }
+
+    tbl <- matrix(NA, nrow=2*nrow(di), ncol=ncol(di))
+    for (i in 1:nrow(di)) {
+      tbl[2*(i-1) + 1,] <- sapply(1:ncol(di), function(j) printstars(di[i,j], se[i,j]))
+      tbl[2*i,] <- sprintf("(%.3f)", se[i,])
+    }
+    colnames(tbl) <- colnames(di)
+    rownames(tbl) <- rep("", nrow(tbl))
+    rownames(tbl)[seq(1,nrow(tbl),by=2)] <- relabel(rownames(di))
+  return(tbl)
+  }
+
+  diall <- dieff_table2(lapply(regs$pib[[sb]][1:4], function(x) x$reg$coef[,1]),
+                       regs$pbiy[[sy]]$reg$coef[,1],
+                       regs$piy[[sy]]$reg$coef[,1],
+                       policies=c(pols, infovars[[sb]]), nsum=length(pols))
+
+  bdiall <- lapply(1:S, function(i)
+    dieff_table2(lapply(1:4, function(j) bs[[j]][,i]),
+                bs[[5]][,i], bs[[6]][,i],
+                policies=c(pols, infovars[[sb]]), nsum=length(pols))
+    )
+
+  tbl <- addse(diall, bdiall)
+
+  #tbl <- tbl[-c(nrow(tbl), nrow(tbl)-1),]
+
+  ltbl <- kable(tbl, digits=2, format="latex", booktabs=TRUE, escape=FALSE,
+                linesep="",
+                align=(rep("c",ncol(tbl))))  %>%
+    column_spec(column=ncol(tbl)+1, border_left=TRUE)
+  cat(ltbl, file=paste(rootdir,
+                       sprintf("tex/tables_and_figures/dieff-cases-constrain-%s.tex",ifelse(use.national, "NI","SI")),
+                       sep="/")) 
   kable(tbl)
@@
                   c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
+pols = c("pmaskbus","pk12","pshelter","pindex")
+iv <- list("0","0")
+#pols = c("pmaskbus","pmaskall","pk12","pshelter","pindex")
 L <- L.d
@@
   bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
+  for (i in 1:4){
+    bs[[i]][2,]<-0.0
+  }
+  bs[[5]][5,]<-0.0
 
@@
 
+```{r dieff-deaths2, eval = TRUE, cache=FALSE, results='asis', warning=FALSE}
+
+pols <- c("pmaskbus","pk12","pshelter","pindex")
+L <- L.d
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+regs <- mainregressions2(sdf,
+                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)
+
+for (use.national in c(TRUE,FALSE)) {
+  sb <- ifelse(use.national, 2, 1)
+  sy <- ifelse(use.national, 3, 1)
+  pib <- lapply(regs$pib[[sb]][1:4], function(x) x$reg)
+  pbiy <- regs$pbiy[[sy]]$reg
+  piy <- regs$piy[[sy]]$reg
+
+  S <- 999
+  bs <- bootstrap_felm2(sdf, c(pib,list(pbiy,piy)), S=S)
+
+  diall <- dieff_table2(lapply(regs$pib[[sb]][1:4], function(x) x$reg$coef[,1]),
+                       regs$pbiy[[sy]]$reg$coef[,1],
+                       regs$piy[[sy]]$reg$coef[,1],
+                       policies=c(pols, infovarsd[[sb]]),
+                       nsum=length(pols)
+                       )
+
+  bdiall <- lapply(1:S, function(i)
+    dieff_table2(lapply(1:4, function(j) bs[[j]][,i]),
+                bs[[5]][,i], bs[[6]][,i],
+                policies=c(pols, infovarsd[[sb]]),
+                nsum=length(pols)
+                ))
+
+  tbl <- addse(diall, bdiall)
+
+  #tbl <- tbl[-c(nrow(tbl), nrow(tbl)-1),]
+
+  ltbl <- kable(tbl, digits=2, format="latex", booktabs=TRUE, escape=FALSE,
+                linesep="",
+                align=(rep("c",ncol(tbl))))  %>%
+    column_spec(column=ncol(tbl)+1, border_left=TRUE)
+  cat(ltbl, file=paste(rootdir,
+                       sprintf("tex/tables_and_figures/dieff-deaths-deathsinfo-constrain-%s.tex",ifelse(use.national, "NI","SI")),
+                       sep="/"))
+  kable(tbl)
+}
+```
+
 ## Robustness Check
@@
 
+# interactions <-list("lag(month,L)", statevars)
+# ilist <- list(interactions, interactions)
+
 coef.vec.1 <- coef.vec.2 <- coef.vec.3 <- coef.vec.4 <-  rep(0.0, 40)
@@
 yvar <- "dlogdc" 
-pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+# pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+pols <- c("pmaskbus","pk12","pshelter","pindex") 
 for (k in 1:length(infovars)) { 
@@
    xlist <-list("","")
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
-  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+   tvars <- "dlogtests" 
+   iv <- list("0","0")
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
+  piy[[(1+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
   sdf2 <- sdf[which(sdf$state!='New York'),]
-  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+  piy[[(2+8*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
   xlist <-list(c("mask_percent"), c("mask_percent"))
-  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(3+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
   xlist <-list(c("logvote"), c("logvote"))
-  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(4+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovars[[k]], L),
@@
   xlist <-list("","")
-  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(5+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
   xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
-   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+   piy[[(6+8*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", bvars, L+14),
@@
               ilist[[k]], 0, L=L) 
-}
-
-for (k in 1:6) { 
-  for (j in 1:3) {
+   # (7) weekly dummy
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list("week", "week")
+  piy[[(7+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], 0, L=L) 
+   # (8) iv
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list("", "")
+  iv <- list("(dlogtests ~ lag(logtests,7))","(dlogtests ~ lag(logtests,7))")
+  tvars <- ""
+  piy[[(8+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovars[[k]], L),
+                tvars, xlist[[k]]),
+              ilist[[k]], iv, L=L) 
+}
+
+for (j in 1:4) {
+  for (k in c(1:8)) { 
     coef.vec.1[[10*(j-1)+k]] <- piy[[k]]$reg$coefficients[[1+j]]
     se.vec.1[[10*(j-1)+k]] <- piy[[k]]$reg$se[[1+j]]
-  coef.vec.2[[10*(j-1)+k]] <- piy[[6+k]]$reg$coefficients[[1+j]]
-  se.vec.2[[10*(j-1)+k]] <- piy[[6+k]]$reg$se[[1+j]] 
+    coef.vec.2[[10*(j-1)+k]] <- piy[[7+k]]$reg$coefficients[[1+j]]
+    se.vec.2[[10*(j-1)+k]] <- piy[[7+k]]$reg$se[[1+j]] 
   } 
@@
 savetextables(NULL, NULL,  piy[1:6], prefix="cases-robust")
-
-# case growth with pindex
-pols <- c("pmaskbus","pk12","pshelter","pindex") 
-for (k in 1:length(infovars)) { 
-  # (1) baseline 
-   xlist <-list("","")
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
-  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars, xlist[[k]]),
-              ilist[[k]], 0, L=L) 
-  # (2) excluding NY 
-   xlist <-list("","")
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
-  sdf2 <- sdf[which(sdf$state!='New York'),]
-  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars, xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (3) adding mask wearing rate  
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("mask_percent"), c("mask_percent"))
-  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars, xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (4) adding log(Trump's vote)
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("logvote"), c("logvote"))
-  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars, xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (5) lagged behavior variables
-  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
-  xlist <-list("","")
-  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", bvars, L+14),
-                sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars, xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (6) (2)-(5)
-   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
-   sdf2 <- sdf[which(sdf$state!='New York'),]
-  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
-   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", bvars, L+14),
-                sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars, xlist[[k]]),
-              ilist[[k]], 0, L=L)
-}
-for (k in 1:6) {  
-  coef.vec.1[[30+k]] <- piy[[k]]$reg$coefficients[[4]]
-  se.vec.1[[30+k]] <- piy[[k]]$reg$se[[4]]
-  coef.vec.2[[30+k]] <- piy[[6+k]]$reg$coefficients[[4]]
-  se.vec.2[[30+k]] <- piy[[6+k]]$reg$se[[4]]  
-}
 
@@
                   c("dlogdd", "logdd","dlogdd.national", "logdd.national")) 
-pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+#pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+pols <- c("pmaskbus","pk12","pshelter","pindex") 
 for (k in 1:length(infovars)) {  
@@
   sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
-  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(1+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
@@
   sdf2 <- sdf[which(sdf$state!='New York'),]
-  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
+  piy[[(2+8*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
@@
   xlist <-list(c("mask_percent"), c("mask_percent"))
-  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(3+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
@@
   xlist <-list(c("logvote"), c("logvote"))
-  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+  piy[[(4+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
               c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
@@
               ilist[[k]], 0, L=L) 
-}
-for (k in 1:6) { 
-  for (j in 1:3) {
+   # (7) weekly dummy
+  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  xlist <-list("week", "week")
+  piy[[(7+8*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
+              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
+                 xlist[[k]]),
+              ilist[[k]], 0, L=L)  
+   # (8) iv = baseline because there is no testing vars in death equation
+  piy[[(8+8*(k-1))]] <- piy[[(1+8*(k-1))]] 
+} 
+for (j in 1:4) {
+  for (k in c(1:8)) { 
     coef.vec.3[[10*(j-1)+k]] <- piy[[k]]$reg$coefficients[[1+j]]
     se.vec.3[[10*(j-1)+k]] <- piy[[k]]$reg$se[[1+j]]
-  coef.vec.4[[10*(j-1)+k]] <- piy[[6+k]]$reg$coefficients[[1+j]]
-  se.vec.4[[10*(j-1)+k]] <- piy[[6+k]]$reg$se[[1+j]] 
+    coef.vec.4[[10*(j-1)+k]] <- piy[[7+k]]$reg$coefficients[[1+j]]
+    se.vec.4[[10*(j-1)+k]] <- piy[[7+k]]$reg$se[[1+j]] 
   } 
@@
 
-# death growth with pindex
-pols <- c("pmaskbus","pk12","pshelter","pindex") 
-for (k in 1:length(infovars)) {  
-  # (1) baseline
-  xlist <-list("","")
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)  
-  piy[[(1+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
-                 xlist[[k]]),
-              ilist[[k]], 0, L=L)   
-  # (2) excluding NY  
-  xlist <-list("","")
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L) 
-  sdf2 <- sdf[which(sdf$state!='New York'),]
-  piy[[(2+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
-                 xlist[[k]]),
-              ilist[[k]], 0, L=L) 
-  # (3) adding mask wearing rate 
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("mask_percent"), c("mask_percent"))
-  piy[[(3+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
-                 xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (4) adding log(Trump's vote)
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
-  xlist <-list(c("logvote"), c("logvote"))
-  piy[[(4+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovarsd[[k]], L),
-                 xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (5) lagged behavior variables
-  sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L)
-  xlist <-list("","")
-  piy[[(5+6*(k-1))]] <- policyreg(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", bvars, L+14),
-                sprintf("lag(%s, %d)", infovarsd[[k]], L),
-                 xlist[[k]]),
-              ilist[[k]], 0, L=L)
-  # (6) (2)-(5)
-   sdf <- subset(df, df$date>=as.Date("2020-02-22")) #+L) 
-   sdf2 <- sdf[which(sdf$state!='New York'),]
-  xlist <-list(c("mask_percent","logvote"), c("mask_percent","logvote"))
-   piy[[(6+6*(k-1))]] <- policyreg(sdf2, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", bvars, L+14),
-                sprintf("lag(%s, %d)", infovarsd[[k]], L),
-                xlist[[k]]),
-              ilist[[k]], 0, L=L) 
-}
-for (k in 1:6) {  
-  coef.vec.3[[30+k]] <- piy[[k]]$reg$coefficients[[4]]
-  se.vec.3[[30+k]] <- piy[[k]]$reg$se[[4]]
-  coef.vec.4[[30+k]] <- piy[[6+k]]$reg$coefficients[[4]]
-  se.vec.4[[30+k]] <- piy[[6+k]]$reg$se[[4]]  
-}
-
 ```
@@
 ## Fixed Effects
-```{r regsfe,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsfe,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 
@@
 
-df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/4
-
+# Call boot command to conduct bootstrap
+set.seed(88) # seed for replication
+num_boot <- 500 # number of bootstraps
+ncores <- 1 # number of cpus (speed)
+# # nonpar will throw warning wrt multiple splitting bias correction
+
+# # # Compute standard errors
+# result <- structure(vapply(result_boot$t, as.double, numeric(1)), dim = dim(result_boot$t))
+# tot_bse <- apply(result, 2, function(x) {
+#   # Normal scaled IQR
+#   return((quantile(x, .75, na.rm = TRUE) - quantile(x, .25, na.rm = TRUE))/(qnorm(.75) - qnorm(.25)))
+# })
+
+# df$month.c <- panellag(df$month,df$state,df$date,L.c)
+# df$month.d <- panellag(df$month,df$state,df$date,L.d)
+df$month.c <- df$month.d <- df$month
+
+bse <- function(sdf,  yvar,  pols, bv, x,  iv,  L, fe) {
+   bootstat_fe <- function(data, form) { 
+    m <- reg_fe(data,  yvar,  pols, bv, x,  iv,  L, fe)   
+    return(as.vector(m$bc2))
+  }
+  result_boot <- boot(data = sdf, statistic = bootstat_fe, sim = "parametric", ran.gen = data_wb, # data_rg,
+                      mle = 0, parallel = "multicore", ncpus = ncores, R = num_boot)
+  # Compute standard errors
+  result <- structure(vapply(result_boot$t, as.double, numeric(1)), dim = dim(result_boot$t))
+  tot_bse <- apply(result, 2, function(x) { 
+    return((quantile(x, .75, na.rm = TRUE) - quantile(x, .25, na.rm = TRUE))/(qnorm(.75) - qnorm(.25)))
+  })
+  return(tot_bse)
+}
+
+# df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/3
+
+fe <- "state + week"
+# fe <- "state + month" 
+iv <- 0
 yvar <- "dlogdc"
+bv <- NULL
 infovars <- list(c("dlogdc", "logdc"),
@@
 L <- L.c
+
 sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
@@
 for (k in 1:length(infovars)) {  
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  # pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
+  pols = c("pmaskbus","pk12","pshelter","pindex")
+  x <- c(sprintf("lag(%s, %d)", infovars[[k]], L), tvars) 
+  m <-  reg_fe(sdf, yvar, pols, bv, x, iv, L, fe)
+  piy[[(1+4*(k-1))]] <- m 
+  m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))]  
+  se <- bse(sdf,  yvar,  pols, bv, x,  iv,  L, fe) 
+  z <- m$bc2/se
+  pval <- 2*pnorm(-abs(z))
+  m$reg$cse[1:(6+2*(k-1))] <- se[1:(6+2*(k-1))] 
+  m$reg$ctval[1:(6+2*(k-1))] <- z[1:(6+2*(k-1))]   
+  m$reg$cpval[1:(6+2*(k-1))] <- pval[1:(6+2*(k-1))]
+  # m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))]  
+  piy[[(2+4*(k-1))]] <- m
+  #pols <- c("pmaskbus","pk12","pshelter","pindex")  
   pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
-  m <-  reg_fe(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars), 0, L, "state + month")
-  piy[[(1+4*(k-1))]] <- m 
+  m <-  reg_fe(sdf, yvar, pols, bv, x, iv, L, fe) 
+  piy[[(3+4*(k-1))]] <- m  
+  # m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
   m$reg$beta[1:(9+2*(k-1))] <- m$bc2[1:(9+2*(k-1))]
-  piy[[(2+4*(k-1))]] <- m
-  pols <- c("pmaskbus","pk12","pshelter","pindex")  
-  m <-  reg_fe(sdf, yvar, pols, NULL,
-              c(sprintf("lag(%s, %d)", infovars[[k]], L),
-                tvars), 0, L, "state + month")
-  piy[[(3+4*(k-1))]] <- m  
-  m$reg$beta[1:(6+2*(k-1))] <- m$bc2[1:(6+2*(k-1))] 
+  se <- bse(sdf,  yvar,  pols, bv, x,  iv,  L, fe) 
+  z <- m$bc2/se
+  pval <- 2*pnorm(-abs(z))
+  m$reg$cse[1:(9+2*(k-1))] <- se[1:(9+2*(k-1))] 
+  m$reg$ctval[1:(9+2*(k-1))] <- z[1:(9+2*(k-1))]   
+  m$reg$cpval[1:(9+2*(k-1))] <- pval[1:(9+2*(k-1))]  
   piy[[(4+4*(k-1))]] <- m 
@@
 
-for (j in 1:3) {
-  coef.vec.1[[10*(j-1)+9]] <- piy[[1]]$reg$beta[[j]]
-  se.vec.1[[10*(j-1)+9]] <- piy[[1]]$reg$se[[j]]
-  coef.vec.1[[10*(j-1)+10]] <- piy[[2]]$reg$beta[[j]]
-  se.vec.1[[10*(j-1)+10]] <- piy[[1]]$reg$se[[j]]
-  coef.vec.2[[10*(j-1)+9]] <- piy[[5]]$reg$beta[[j]]
-  se.vec.2[[10*(j-1)+9]] <- piy[[5]]$reg$se[[j]]
-  coef.vec.2[[10*(j-1)+10]] <- piy[[6]]$reg$beta[[j]]
-  se.vec.2[[10*(j-1)+10]] <- piy[[5]]$reg$se[[j]]
-}
-coef.vec.1[[30+9]] <- piy[[3]]$reg$beta[[3]]
-se.vec.1[[30+9]] <- piy[[3]]$reg$se[[3]]
-coef.vec.1[[30+10]] <- piy[[4]]$reg$beta[[3]]
-se.vec.1[[30+10]] <- piy[[3]]$reg$se[[3]]
-coef.vec.2[[30+9]] <- piy[[7]]$reg$beta[[3]]
-se.vec.2[[30+9]] <- piy[[7]]$reg$se[[3]]
-coef.vec.2[[30+10]] <- piy[[8]]$reg$beta[[3]]
-se.vec.2[[30+10]] <- piy[[7]]$reg$se[[3]]
+# for (j in 1:4) {
+#   # coef.vec.1[[10*(j-1)+9]] <- piy[[1]]$reg$beta[[j]]
+#   # se.vec.1[[10*(j-1)+9]] <- piy[[1]]$reg$se[[j]]
+#   coef.vec.1[[10*(j-1)+10]] <- piy[[2]]$reg$beta[[j]]
+#   se.vec.1[[10*(j-1)+10]] <- piy[[2]]$reg$cse[[j]]
+#   # coef.vec.2[[10*(j-1)+9]] <- piy[[5]]$reg$beta[[j]]
+#   # se.vec.2[[10*(j-1)+9]] <- piy[[5]]$reg$se[[j]]
+#   coef.vec.2[[10*(j-1)+10]] <- piy[[6]]$reg$beta[[j]]
+#   se.vec.2[[10*(j-1)+10]] <- piy[[6]]$reg$cse[[j]]
+# }
 
@@
 for (k in 1:length(infovars)) {  
-  sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+  pols = c("pmaskbus","pk12","pshelter","pindex") 
+  x <- sprintf("lag(%s, %d)", infovars[[k]], L)
+  m <-  reg_fe(sdf, yvar, pols, bv, x, iv, L, fe)
+  piy[[(1+4*(k-1))]] <- m
+  m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))]  
+  se <- bse(sdf,  yvar,  pols, bv, x,  iv,  L, fe)  
+  z <- m$bc2/se
+  pval <- 2*pnorm(-abs(z))
+  m$reg$cse[1:(5+2*(k-1))]  <- se[1:(5+2*(k-1))] 
+  m$reg$ctval[1:(5+2*(k-1))]  <- z[1:(5+2*(k-1))]    
+  m$reg$cpval[1:(5+2*(k-1))]  <- pval[1:(5+2*(k-1))]  
+  #m$reg$beta[1:(8+2*(k-1))] <- m$bc2[1:(8+2*(k-1))]  
+  piy[[(2+4*(k-1))]] <- m
+  # pols <- c("pmaskbus","pk12","pshelter","pindex")  
   pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
-  m <-  reg_fe(sdf, yvar, pols, NULL,
-              sprintf("lag(%s, %d)", infovars[[k]], L), 0, L, "state + month")
-  piy[[(1+4*(k-1))]] <- m
+  m <-  reg_fe(sdf, yvar, pols, bv, x, iv, L, fe)  
+  piy[[(3+4*(k-1))]] <- m 
+  #m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))]  
   m$reg$beta[1:(8+2*(k-1))] <- m$bc2[1:(8+2*(k-1))]  
-  piy[[(2+4*(k-1))]] <- m
-  pols <- c("pmaskbus","pk12","pshelter","pindex")  
-  m <-  reg_fe(sdf, yvar, pols, NULL,
-              sprintf("lag(%s, %d)", infovars[[k]], L), 0, L, "state + month")
-  piy[[(3+4*(k-1))]] <- m 
-  m$reg$beta[1:(5+2*(k-1))] <- m$bc2[1:(5+2*(k-1))]  
+  se <- bse(sdf,  yvar,  pols, bv, x,  iv,  L, fe)  
+  z <- m$bc2/se
+  pval <- 2*pnorm(-abs(z))
+  m$reg$cse[1:(8+2*(k-1))]  <- se[1:(8+2*(k-1))] 
+  m$reg$ctval[1:(8+2*(k-1))]  <- z[1:(8+2*(k-1))]    
+  m$reg$cpval[1:(8+2*(k-1))]  <- pval[1:(8+2*(k-1))]  
   piy[[(4+4*(k-1))]] <- m
@@
 
-for (j in 1:3) {
-  coef.vec.3[[10*(j-1)+9]] <- piy[[1]]$reg$beta[[j]]
-  se.vec.3[[10*(j-1)+9]] <- piy[[1]]$reg$se[[j]]
-  coef.vec.3[[10*(j-1)+10]] <- piy[[2]]$reg$beta[[j]]
-  se.vec.3[[10*(j-1)+10]] <- piy[[1]]$reg$se[[j]]
-  coef.vec.4[[10*(j-1)+9]] <- piy[[5]]$reg$beta[[j]]
-  se.vec.4[[10*(j-1)+9]] <- piy[[5]]$reg$se[[j]]
-  coef.vec.4[[10*(j-1)+10]] <- piy[[6]]$reg$beta[[j]]
-  se.vec.4[[10*(j-1)+10]] <- piy[[5]]$reg$se[[j]]
-}
-coef.vec.3[[30+9]] <- piy[[3]]$reg$beta[[3]]
-se.vec.3[[30+9]] <- piy[[3]]$reg$se[[3]]
-coef.vec.3[[30+10]] <- piy[[4]]$reg$beta[[3]]
-se.vec.3[[30+10]] <- piy[[3]]$reg$se[[3]]
-coef.vec.4[[30+9]] <- piy[[7]]$reg$beta[[3]]
-se.vec.4[[30+9]] <- piy[[7]]$reg$se[[3]]
-coef.vec.4[[30+10]] <- piy[[8]]$reg$beta[[3]]
-se.vec.4[[30+10]] <- piy[[7]]$reg$se[[3]]
+# for (j in 1:4) {
+#   # coef.vec.3[[10*(j-1)+9]] <- piy[[1]]$reg$beta[[j]]
+#   # se.vec.3[[10*(j-1)+9]] <- piy[[1]]$reg$se[[j]]
+#   coef.vec.3[[10*(j-1)+10]] <- piy[[2]]$reg$beta[[j]]
+#   se.vec.3[[10*(j-1)+10]] <- piy[[2]]$reg$cse[[j]]
+#   # coef.vec.4[[10*(j-1)+9]] <- piy[[5]]$reg$beta[[j]]
+#   # se.vec.4[[10*(j-1)+9]] <- piy[[5]]$reg$se[[j]]
+#   coef.vec.4[[10*(j-1)+10]] <- piy[[6]]$reg$beta[[j]]
+#   se.vec.4[[10*(j-1)+10]] <- piy[[6]]$reg$cse[[j]]
+# }
 
@@
 ## DML 
-```{r regsdml,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsdml,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 
@@
                   c("dlogdd", "logdd","dlogdd.national", "logdd.national"))  
+bvars <- c("workplaces","retail","grocery","transit")
 vars <- c(infovars[[2]],infovarsd[[2]],pols,"pindex") 
@@
 
-# df$test <- df$dlogtests
-# df$test[is.na(df$test)] <- 0
-
 svars <- c("logpop","logsq","Percent.Unemployed..2018.",
@@
                "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum")
-mvars <- c("month4","month5","month6") 
-ijs <- expand.grid(1:length(svars), 1:length(mvars))
-msvars <- vector(mode="character", length=lengths(ijs)[[1]])
+mvars.c <- mvars.d <- mvars <- c("month4","month5","month6")  
+# mvars.c <- c("month4.c","month5.c","month6.c") 
+# mvars.d <- c("month4.d","month5.d","month6.d") 
+ijs <- expand.grid(1:length(svars), 1:length(mvars.c))
+msvars.c <- vector(mode="character", length=lengths(ijs)[[1]])
+msvars.d <- vector(mode="character", length=lengths(ijs)[[1]])
 for (ij in (1:lengths(ijs)[[1]])){
-  v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars[ijs[[2]][ij]])
-  msvars[ij] <- v
-  df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars[ijs[[2]][ij]]]
-}
-
-wvars <- c(svars,mvars,msvars,"month3","mask_percent","logvote","test") 
-wvars.d <- c(svars,mvars,msvars,"month3","mask_percent","logvote")
+  v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars.c[ijs[[2]][ij]])
+  msvars.c[ij] <- v
+  df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars.c[ijs[[2]][ij]]]
+  v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars.d[ijs[[2]][ij]])
+  msvars.d[ij] <- v
+  df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars.d[ijs[[2]][ij]]]
+}
 
@@
 bvars.d <- c("workplaces.L21","retail.L21","grocery.L21","transit.L21")
+
+wvars <- c(svars,mvars.c,msvars.c,"month3","mask_percent","logvote","dlogtests") 
+wvars.d <- c(svars,mvars.d,msvars.d,"month3","mask_percent","logvote")  
 xvars <- c(bvars.c,"dlogdc.L14","logdc.L14")  
@@
 
-pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pmovie.L14","prestaurant.L14","pnonessential.L14") 
-pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pmovie.L21","prestaurant.L21","pnonessential.L21")
+# wvars <- c(bvars.c,svars,mvars.c,msvars.c,"month3","mask_percent","logvote","dlogtests") 
+# wvars.d <- c(bvars.d,svars,mvars.d,msvars.d,"month3","mask_percent","logvote") 
+# xvars <- c("dlogdc.L14","logdc.L14")  
+# xvars.national <- c("dlogdc.L14","logdc.L14","dlogdc.national.L14","logdc.national.L14") 
+# xvars.d <- c("dlogdd.L21","logdd.L21")  
+# xvars.national.d <- c("dlogdd.L21","logdd.L21","dlogdd.national.L21","logdd.national.L21")
+
+# pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pmovie.L14","prestaurant.L14","pnonessential.L14") 
+# pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pmovie.L21","prestaurant.L21","pnonessential.L21")
+pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pindex.L14") 
+pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pindex.L21")
 
@@
 
-nfold <- 5
+nfold <- 10
 
@@
   DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
-  coef.vec.1[[10*(j-1)+7]] <- DML2.lasso$coef.est
-  se.vec.1[[10*(j-1)+7]] <- DML2.lasso$se
+  coef.vec.1[[10*(j-1)+9]] <- DML2.lasso$coef.est
+  se.vec.1[[10*(j-1)+9]] <- DML2.lasso$se
   set.seed(1)
-  DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold=10) 
-  coef.vec.1[[10*(j-1)+8]] <- DML2.RF$coef.est
-  se.vec.1[[10*(j-1)+8]] <- DML2.RF$se 
+  DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold=nfold) 
+  coef.vec.1[[10*(j-1)+10]] <- DML2.RF$coef.est
+  se.vec.1[[10*(j-1)+10]] <- DML2.RF$se 
   # case growth with national
@@
   DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
-  coef.vec.2[[10*(j-1)+7]] <- DML2.lasso$coef.est
-  se.vec.2[[10*(j-1)+7]] <- DML2.lasso$se 
+  coef.vec.2[[10*(j-1)+9]] <- DML2.lasso$coef.est
+  se.vec.2[[10*(j-1)+9]] <- DML2.lasso$se 
   set.seed(1)
   DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold = nfold) 
-  coef.vec.2[[10*(j-1)+8]] <- DML2.RF$coef.est
-  se.vec.2[[10*(j-1)+8]] <- DML2.RF$se 
+  coef.vec.2[[10*(j-1)+10]] <- DML2.RF$coef.est
+  se.vec.2[[10*(j-1)+10]] <- DML2.RF$se 
   ## death growth
@@
   DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
-  coef.vec.3[[10*(j-1)+7]] <- DML2.lasso$coef.est
-  se.vec.3[[10*(j-1)+7]] <- DML2.lasso$se  
+  coef.vec.3[[10*(j-1)+9]] <- DML2.lasso$coef.est
+  se.vec.3[[10*(j-1)+9]] <- DML2.lasso$se  
   set.seed(1)
   DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold = nfold)
-  coef.vec.3[[10*(j-1)+8]] <- DML2.RF$coef.est
-  se.vec.3[[10*(j-1)+8]] <- DML2.RF$se  
+  coef.vec.3[[10*(j-1)+10]] <- DML2.RF$coef.est
+  se.vec.3[[10*(j-1)+10]] <- DML2.RF$se  
   # death growth with national vars
@@
   DML2.lasso <- DML2.for.PLM(xv, wv, dv, yv, reg.lasso, sdf, nfold = nfold)
-  coef.vec.4[[10*(j-1)+7]] <- DML2.lasso$coef.est
-  se.vec.4[[10*(j-1)+7]] <- DML2.lasso$se  
+  coef.vec.4[[10*(j-1)+9]] <- DML2.lasso$coef.est
+  se.vec.4[[10*(j-1)+9]] <- DML2.lasso$se  
   set.seed(1)
   DML2.RF <- DML2.for.PLM(xv, wv, dv, yv, reg.RF, sdf, nfold = nfold)
-  coef.vec.4[[10*(j-1)+8]] <- DML2.RF$coef.est
-  se.vec.4[[10*(j-1)+8]] <- DML2.RF$se 
+  coef.vec.4[[10*(j-1)+10]] <- DML2.RF$coef.est
+  se.vec.4[[10*(j-1)+10]] <- DML2.RF$se 
 }
@@
 library(dotwhisker) 
-var.names <- c("(1) baseline    ","(2) excluding NY","(3) mask wearing","(4) Trump's vote ","(5) behavior vars"," (6) all of (2)-(5)  ",  "(7) DML (Lasso)", "(8) DML (RF)  ", "(9) FE with BC")
+var.names <- c("(1) baseline    ","(2) excluding NY","(3) mask wearing","(4) Trump's vote ","(5) behavior vars"," (6) all of (2)-(5)  ","(7) Weekly dummy","(8) IV for dlogtest","(9) DML (Lasso)", "(10) DML (RF)  ") #, "(10) FE with BC")
 
 # mask coefficient
-ij <- c(1:8,10)
+k <- 10
+ij <- c(1:k)
  results_df <- data.frame(term = rep(var.names, times = 4),
@@
                          std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
-                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
-                                   rep("death w/o national",9), rep("death wi/ national", 9)),
+                         model = c(rep("case w/o national", k), rep("case w/ national", k),
+                                   rep("death w/o national",k), rep("death w/ national", k)),
                          stringsAsFactors = FALSE)
@@
 fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
-    theme(       legend.justification=c(.01, .999), legend.position=c(.63, .99), 
+    theme(       legend.justification=c(.01, .999), legend.position=c(.61, .99), 
            legend.title = element_blank(), legend.background =
@@
   theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
-   xlim(-0.55, 0.4) +theme(
+   xlim(-0.35, 0.35) +theme(
   legend.title = element_text(size = 6),
@@
 # school coefficient
-ij <- c(11:18,20)
+ij <- c(11:(10+k))
  results_df <- data.frame(term = rep(var.names, times = 4),
@@
                          std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
-                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
-                                   rep("death w/o national",9), rep("death w/ national", 9)),
+                         model = c(rep("case w/o national", k), rep("case w/ national", k),
+                                   rep("death w/o national",k), rep("death w/ national", k)),
                          stringsAsFactors = FALSE) 
@@
   theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
-  xlim(-1.4, 0.5)  +theme(
+  xlim(-1.3, 0.3)  +theme(
   legend.title = element_text(size = 6),
@@
 # stay-at-home coefficient
-ij <- c(21:28,30)
+ij <- c(21:(20+k))
  results_df <- data.frame(term = rep(var.names, times = 4),
@@
                          std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
-                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
-                                   rep("death w/o national",9), rep("death w/ national", 9)),
+                         model = c(rep("case w/o national", k), rep("case w/ national", k),
+                                   rep("death w/o national",k), rep("death w/ national", k)),
                          stringsAsFactors = FALSE) 
 fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
-    theme(       legend.justification=c(.01, .999), legend.position=c(.63, .99), 
+    theme(       legend.justification=c(.01, .999), legend.position=c(.61, .99), 
            legend.title = element_blank(), legend.background =
@@
 # pindex coefficient
-ij <- c(31:38,40)
+ij <- c(31:(30+k))
  results_df <- data.frame(term = rep(var.names, times = 4),
@@
                          std.error = c(se.vec.1[ij], se.vec.2[ij],se.vec.3[ij], se.vec.4[ij]),
-                         model = c(rep("case w/o national", 9), rep("case w/ national", 9),
-                                   rep("death w/o national",9), rep("death w/ national", 9)),
+                         model = c(rep("case w/o national", k), rep("case w/ national", k),
+                                   rep("death w/o national",k), rep("death w/ national", k)),
                          stringsAsFactors = FALSE) 
 fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
-    theme(       legend.justification=c(.01, .999), legend.position=c(.67, .99), 
+    theme(       legend.justification=c(.01, .999), legend.position=c(.61, .99), 
            legend.title = element_blank(), legend.background =
               element_rect(color="gray90")) + 
-    xlab("coefficients of average of four policy variables") +
-    geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
+    xlab("coefficients of average of three policy variables") +
+    # geom_vline(xintercept = 0, colour = "grey60", linetype = 2) +
   theme(text=element_text(family="Times New Roman", face="bold", size=8)) + 
-  xlim(-0.55, 0.4) +theme(
+  xlim(-0.35, 0.35) +theme(
   legend.title = element_text(size = 6),

### 2020-08-21 tex/covidEffectOfNPI-revision.tex
update

@@
 {Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.}\thanks{We are grateful to Daron Acemoglu, V.V.
-Chari, Raj Chetty,  Christian Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green,  C.
-Jessica E.
-Metcalf, Ido Rosen, Konstantin Sonin, James Stock, and Ivan Werning for helpful comments.
+Chari, Raj Chetty,  Christian Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green,  Ido Rosen, Konstantin Sonin, James Stock, Elie Tamer, and Ivan Werning for helpful comments.
 We also thank Chiyoung Ahn,
 Joshua Catalano,  Jason Chau, Samuel Gyetvay, Sev Chenyu Hou,
-Jordan Hutchings, and Dongxiao Zhang  for excellent research assistance.  All mistakes are our own. }
+Jordan Hutchings, and Dongxiao Zhang  for excellent research assistance.  We are especially grateful
+to   C. Jessica E. Metcalf and an anonymous referee for detailed and insightful comments,
+which considerable improved the paper. All mistakes are our own. }
 
@@
 \begin{abstract}
-This paper evaluates the dynamic impact of various policies adopted by US states on the growth rates of confirmed Covid-19  cases and deaths as well as social distancing behavior measured by Google Mobility Reports, where we take into consideration people's voluntarily behavioral response to new information of transmission risks.
+The paper evaluates the dynamic impact of various policies adopted by US states on the growth rates of confirmed Covid-19  cases and deaths as well as social distancing behavior measured by Google Mobility Reports, where we take into consideration people's voluntarily behavioral response to new information of transmission risks in a causal structural model framework.
 Our analysis finds that both policies and information on transmission risks are important determinants of Covid-19  cases and deaths and shows that a change in policies explains a large fraction of observed changes in social distancing behavior.
-Our counterfactual experiments suggest that nationally mandating face masks for employees on April 1st could have reduced the growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 17 to 55 percent less deaths nationally by the end of May, which roughly translates into 17 to 55 thousand saved lives.
-Our estimates imply that removing non-essential business closures  (while maintaining school closures, restrictions on  movie theaters and restaurants)  could have led to -20 to 60 percent more cases and deaths by the end of May.
+Our main Lcounterfactual experiments suggest that nationally mandating face masks for employees on April 1st could have reduced the growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 17 to 55 percent less deaths nationally by the end of May, which roughly translates into 17 to 55 thousand saved lives.
 We also find that, without stay-at-home orders, cases would have been larger by  25 to 170 percent, which implies that 0.5 to 3.4 million more Americans could have been infected if stay-at-home orders had not been implemented.
-Finally,  we find considerable uncertainty over the effects of removing all policies because the timing of school closures had little cross-sectional variation.%  (the potential range of effects could be even larger the effect of school closures on behavior is hard to identify separately from aggregate time effects).
+We find considerable uncertainty over the effects of school closures due to lack of cross-sectional variation;  we could not robustly rule out either large or small effects.
+We also carry out sensitivity analyses to find neighborhoods of
+the models under which the results hold robustly.
+Overall, substantial declines in growth rates are attributable to private behavioral response, but policies played an important role as well.
+We stress that our study is observational and should be interpreted with great caution.
+From a completely agnostic point of view,
+our findings uncover predictive effects (association) of observed policies and behavioral changes  on future health outcomes,
+ controlling for informational and other confounding variables.
+
+%  (the potential range of effects could be even larger the effect of school closures on behavior is hard to identify separately from aggregate time effects).
 \end{abstract}
@@
 Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{
-The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007} and genetics \citep{wright1923}, with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-See \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book}  for modern developments, especially in computer
-science and epidemiology. The particular causal diagram has several “mediation” components, where variables affect
-outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
+The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, tinbergen1940, wold1954} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+See \cite{pearl:biometrika},\cite{pearl:robins},  \cite{pearl:causality}, \cite{white:chalak}, \cite{peters2020book}, \cite{heckman:pinto}, and \cite{hernanrobins2020book}  for modern developments, in computer
+science, econometrics, and epidemiology.
+The particular causal diagram has several “mediation” components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
 The father and son,
@@
     \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
+%     \multicolumn{2}{c}{\textbf{Cases}}     \multicolumn{2}{c}{\textbf{Deaths}}
+ %  \\
+     \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
     \end{tabular}
@@
     \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
-      &
+    %  \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
+%
         \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
@@
     \begin{tabular}{cc}
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
-      &
+  %    \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
+    %  &
         \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
@@
 In contrast, policies such as mandating employees to wear masks can potentially affect the Covid-19 transmission directly.
-The information variables, such as recent growth in the number of cases, can cause people to spend more time at home, regardless of adopted state policies; these changes in
-behavior in turn affect the transmission of Covid-19.
+The information variables, such as recent growth in the number of cases, can cause people to spend more time at home, regardless of adopted state policies; these changes in behavior in turn affect the transmission of Covid-19.
 
@@
 
-Our quantitative model for causal structure in Figure \ref{Wright} is given by the following econometric structural equation model:\begin{equation} \label{eqn:SEM1} \tag{SEM}
+Our quantitative model for causal structure in Figure \ref{Wright} is given by the following econometric structural (or potential) outcomes model:\begin{equation} \label{eqn:SEM1} \tag{SO}
   \begin{aligned}
@@
 
-The orthogonality restrictions on the stochastic components are as follows:
-\begin{equation}\label{eqn:SEM2} \tag{O}
+The exogeneity restrictions on the stochastic components are as follows:
+\begin{equation}\label{eqn:SEM2} \tag{E}
 \begin{aligned}
@@
 the system from the last equation to the first:
-%\begin{equation} \label{eqn:SEM2}
-\[
+\begin{equation} \label{eqn:SEMO}\tag{O}
 \begin{aligned}
@@
 & {\pcolor P_{it} }& := &  P_{it}({\icolor I_{it}}). \end{aligned}
-\]
-%\end{equation}
-
-The system above together with  orthogonality restrictions (\ref{eqn:SEM2}) implies the following collection of stochastic equations for realized variables:
-\begin{align}
-   &  {\ycolor  Y_{i,t+\ell}}
-    = {\bcolor\alpha ' B_{it}} + {\pcolor\pi 'P_{it}} + {\icolor\mu'I_{it}} + {\wcolor\delta_Y 'W_{it}}  + \varepsilon^y_{it},
-    &  & \varepsilon^y_{it} \perp {\bcolor B_{it}}, {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R1} \tag{BPI$\to$Y} \\
-    &  {\bcolor B_{it}}
-     =  {\pcolor \beta' P_{it}} + {\icolor \gamma'I_{it}} +  {\wcolor \delta_B' W_{it}} + \varepsilon^b_{it},
-   & & \varepsilon^b_{it} \perp {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}}  \label{eq:R2} \tag{PI$\to$B}  \\
-    & {\pcolor P_{it}}
-    =  {\icolor\eta'I_{it}} + {\wcolor \delta_P' W_{it}} +   \varepsilon^p_{it},   & & \varepsilon^p_{it} \perp   {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R3}  \tag{I$\to$P}
-       \end{align}
-and
-\begin{equation}
-   {\ycolor  Y_{i,t+\ell}}
-   = ( {\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'} )
-    {\pcolor P_{it}} + ( {\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})
-    {\icolor I_{it} }+ {\wcolor \bar{\delta} '}{\wcolor W_{it}}  + {\bar \varepsilon}_{it},  \quad   {\bar \varepsilon}_{it} \perp
-  {\pcolor P_{it}},  {\icolor I_{it}}, {\wcolor W_{it}}.  \label{eq:R4} \tag{PI$\to$Y}
 \end{equation}
-These equations form the basis of our empirical analysis.
-
-As discussed below, the information variable includes case growth.
-Therefore,  an orthogonality restriction  $ \varepsilon^y_{it} \perp   {\pcolor P_{it}}$ holds if  the government does not have knowledge on future case growth beyond what is predicted by today's case growth, policies,  behavior, and  confounders; even when the government has some knowledge on $\varepsilon^y_{it}$, the orthogonality restriction may hold if there is a time lag for the government to implement its policies based on $\varepsilon^y_{it}$.
-
-The orthogonality condition in (\ref{eq:R4}) is weaker than the orthogonality conditions in  (\ref{eq:R1})-(\ref{eq:R2}) in that the former is implied by the latter but not vice versa.
-The system over-identifies the regression coefficients because  $({\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'})$ and $({\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})$ in  (\ref{eq:R4}) can be also identified from  ${\bcolor\alpha ' }$,  $ {\pcolor\pi '}$, ${\icolor\mu'}$,   ${\pcolor \beta' }$, and ${\icolor \gamma'}$ in (\ref{eq:R1})-(\ref{eq:R2}).
-Comparing the estimates of $({\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'})$ and $({\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'})$ from  (\ref{eq:R4})  with  those implied by  the estimates of ${\bcolor\alpha ' }$,  $ {\pcolor\pi '}$, ${\icolor\mu'}$,   ${\pcolor \beta' }$, and ${\icolor \gamma'}$ from   (\ref{eq:R1})-(\ref{eq:R2}) provides a useful specification test.
-
-\subsection*{Identification and Parameter Estimation}
-
-The orthogonality equations imply that these are all projection equations, and the parameters of the SEM are identified by the parameters of these regression equation, provided the latter are identified by sufficient joint variation of these variables across states and time.
-
-The last point can be stated formally as follows. Consider the previous system of equations, after partialling out the confounders:
-\begin{equation}\label{eqn:SEM-PO}
- \begin{aligned}
-   & {\ycolor { \tilde Y_{i,t+\ell}}}  & =& {\bcolor {\alpha ' \tilde B_{it}}} +{\pcolor {\pi '\tilde P_{it}}} +{\icolor \mu'\tilde I_{it} } + \varepsilon^y_{it},&\varepsilon^y_{it} &\perp {\tilde B_{it}}, {\tilde  P_{it}}, { \tilde I_{it}},  \\
-   & {\bcolor { \tilde B_{it}}} & = &{\pcolor { \beta' \tilde P_{it}} } + {\icolor { \gamma' \tilde I_{it}} } + \varepsilon^b_{it},
-   &\quad \varepsilon^b_{it} &\perp { \tilde P_{it}}, { \tilde I_{it}},  \\
-   & {\pcolor { \tilde P_{it}} } &= &  {\icolor {\eta' \tilde I_{it}} } +   \varepsilon^p_{it}, &\quad \varepsilon^p_{it} &\perp {\tilde I_{it}}  \\
-  \end{aligned}
-\end{equation}
-where $ \tilde V_{it} = V_{it}   -     {\wcolor W_{it}'} \Ep[{\wcolor W_{it}W_{it}'}]^{-} \Ep[{\wcolor W_{it}} V_{it}]$ denotes
-the residual after removing the orthogonal projection of $V_{it}$ on ${\wcolor W_{it}}$.
-The residualization is a linear operator, implying that (\ref{eqn:SEM-PO}) follows immediately from the above.
-The parameters of (\ref{eqn:SEM-PO})  are identified as projection coefficients in these equations, provided that residualized vectors appearing in each of the equations have non-singular variance, that is
- \begin{equation}
- \Var ( {\pcolor \tilde P_{it}'} ,{\bcolor \tilde B_{it}'},{\icolor \tilde I_{it}'})>0,
- \ \Var ({\pcolor \tilde P_{it}'}, {\icolor \tilde I_{it}'})> 0 , \ \text{ and  }  \Var ( {\icolor \tilde I_{it}'}) >0.
- \end{equation}
-
-Our main estimation method is the standard correlated random effects estimator, where the random effects
-are parameterized as functions of observable characteristic, ${\wcolor W_{it}}$, which include both state-level and time random effects.  The state-level random effects
-are modeled as a function of state level characteristics, and the time random effects are modeled
-by including month dummies and their interactions with state level characteristics.  The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$
-are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.
-
-A secondary estimation method is the fixed effects estimator, where ${\wcolor W_{it}}$ includes
-latent (unobserved) state level effects ${\wcolor W_i}$ and  and time level effects ${\wcolor W_t}$,
-which must be estimated from the data.  This approach is much more demanding of the data and relies on long cross-sectional
-and time histories. When histories are relatively short,  large biases emerge and
-they need to be removed using debiasing methods. In our context, debiasing
-  changes the estimates substantially, often changing the sign of coefficients.\footnote{This is a pre-cautionary
-message that may be useful for other researchers using fixed effects estimators in the context
-of Covid-19 analysis. We recommend using debiased fixed effects estimators, see e.g., \cite{chen2019mastering}
-for expository treatment. }
-However, we find the debiased fixed effect estimates are
-qualitatively and quantitatively similar to the correlated random effects estimates.
-Given this finding, we chose to focus on the latter, as it is a more standard and familiar method, and report the former
-estimates in the supplementary materials for this paper.\footnote{The similarity of the debiased fixed
-effects and correlated random effects served as a useful specification check. Moreover, using the fixed
-effects estimators only yielded minor gains in predictive performances, as judging by the adjusted
-$R^2$'s, providing another useful specification check.}
-
-\subsection{Information Structures and Induced Dynamics}  We consider three examples of information
-structures: Information variable is  a function of time:
-$${\icolor I_{it}} = g(t);$$
- Information variable is lagged value of outcome:
-$$
-{\icolor I_{it}} =  {\ycolor Y_{it}};
- $$
- and finally:
- \begin{itemize}
- \item[(\textsc{I})] Information variables include time, lagged and integrated values of outcome:
- $$
-{\icolor I_{it}} = \left (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{t/\ell}
-{\ycolor Y_{i,t - \ell m}} \right) ',
- $$
- with the convention that ${\ycolor Y_{it}} = 0$  for $t \leq 0$.
-\end{itemize}
-The first information structure captures the  basic idea that, as individuals discover more information about covid over time, they adapt to safer modes of  behavior (stay-at-home, wear masks, wash hands).
-{Under this structure, information is common across states and exogenously evolves over time, independent of the number of cases.}
-The second  structure  arises from considering
-autoregressive components {and captures people's behavioral response to information on cases in the state they reside}.
-{Specifically,} we model persistence in growth rates, ${\ycolor Y_{i,t+\ell}}$, through an AR(1) model, which leads to ${\icolor I_{it}}= {\ycolor Y_{it}}$.
-This provides useful \textit{local}, state-specific, information about the forward growth rate and people may adjust their behavior to safer modes when they see a high value.
-We model this adjustment via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.
-The {third information} structure
-is the combination of the first two structures plus an additional  term representing the log of the total number of new cases in the state.
-We use this information structure in our empirical specification.
+
+The specification of the model above grasps one-period responses, but does not allow
+to answer about multi-period responses.  The dynamics will be induced by the evolution
+of information variables: Information variables include time, lagged and integrated values of outcome:
+\begin{equation}\label{eq:I}\tag{I}
+ {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{t/\ell}
+{\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}
+ for  each  $t \in [T]$, where $g$ is deterministic function of time, e.g., month indicators,
+ and $t/\ell$ is defined as $\lfloor {t/\ell} \rfloor$.
+
 In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases.
-The last element of the information set can be thought of as a local stochastic trend in cases.
-
-All of these examples fold into a specification of the form:
-\begin{equation}\label{eq:I}\tag{I}
- {\icolor I_{it} } :=   I_{it}({\icolor I_{i,t-\ell}}, {\ycolor Y_{it}}, t),  \quad t =1,..., T,
- \end{equation}
- with the initialization ${\icolor I_{i 0}} = 0$ and ${\ycolor Y_{i0}} = 0$.\footnote{This initialization is appropriate
- in our context for analyzing pandemics from the very beginning, but other initializations could be appropriate in other contexts.
-The  lagged values of behavior variable may be also included in the information set.
- }
+The local information also captures the persistence of the growth rate process.
+We model the reaction of people's behavior via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.
+The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$ in our context, since we are analyzing the pandemics from the very beginning, but other initializations could be appropriate in other contexts.
+The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
 
 With any structure of this form, realized outcomes may become new information for future periods, inducing a dynamical system over multiple periods.
-We show the resulting dynamical system in a causal diagram of Figure \ref{fig:DynS}.
+We show the resulting dynamical system in a diagram of Figure \ref{fig:DynS}.
 Specification
@@
 \end{figure}
+
+Next we combine the above parts together with an appropriate
+initializations to give a formal definition of the model we use.
+
+\begin{quote}
+\textsc{Structural Equations Model (SEM)}.
+Let $i \in \{1,..., N\}$ denote the observational init and $t$ be the time periods and $\ell$ the time delay, a non-negative integer.
+(1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined outside of the model for $t \leq 0$.
+(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are determined, then information variables $I_{it}$ are determined  by (I), and policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) and realizing the shochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E),  the behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) and realizing  the shochastic shock $\varepsilon^b_{it}$ that obeys the exogeneity condition (E), the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) and realizing the shochastic shock $\varepsilon^y_{it}$ that obeys the exogeneity condition (E).
+\end{quote}
+
+%This formulation uses traditional econometrics language of causal structural equation models.
+It appears to fit in well with general formulations of structural causal models developed \cite{pearl:why} and \cite{white:chalak}.
+
+\subsection{Main Testable Implication, Identification, Parameter Estimation}
+
+The system above together with  orthogonality restrictions (\ref{eqn:SEM2}) implies the following collection of projection equations for realized variables:
+\begin{align}
+   &  {\ycolor  Y_{i,t+\ell}}
+    = {\bcolor\alpha ' B_{it}} + {\pcolor\pi 'P_{it}} + {\icolor\mu'I_{it}} + {\wcolor\delta_Y 'W_{it}}  + \varepsilon^y_{it},
+    &  & \varepsilon^y_{it} \perp {\bcolor B_{it}}, {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}} \label{eq:R1} \tag{BPI$\to$Y} \\
+    &  {\bcolor B_{it}}
+     =  {\pcolor \beta' P_{it}} + {\icolor \gamma'I_{it}} +  {\wcolor \delta_B' W_{it}} + \varepsilon^b_{it},
+   & & \varepsilon^b_{it} \perp {\pcolor P_{it}}, {\icolor I_{it}}, {\wcolor W_{it}}  \label{eq:R2} \tag{PI$\to$B}  \\
+    & {\pcolor P_{it}}
+    =  {\icolor\eta'I_{it}} + {\wcolor \delta_P' W_{it}} +   \varepsilon^p_{it},   & & \varepsilon^p_{it} \perp   {\icolor I_{it}}, {\wcolor W_{it}}  \label{eq:R3}  \tag{I$\to$P} \\ 
+    &  {\ycolor  Y_{i,t+\ell}}
+   =     ( {\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'} )   {\pcolor P_{it}} +    ( {\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'}){\icolor I_{it} }+ {\wcolor \bar{\delta} '}{\wcolor W_{it}}  + {\bar \varepsilon}_{it},  &&  {\bar \varepsilon}_{it} \perp
+  {\pcolor P_{it}},  {\icolor I_{it}}, {\wcolor W_{it}}.  \label{eq:R4} \tag{PI$\to$Y}
+      \end{align}
+
+Therefore the projection equation:
+\begin{equation}
+   {\ycolor  Y_{i,t+\ell}}
+   = \mathsf{a}'
+    {\pcolor P_{it}} + \mathsf{b}'    {\icolor I_{it} }+ {\wcolor \tilde {\delta} '}{\wcolor W_{it}}  + {\bar \varepsilon}_{it},  \quad   {\bar \varepsilon}_{it} \perp
+  {\pcolor P_{it}},  {\icolor I_{it}}, {\wcolor W_{it}}.  \label{eq:PR} \tag{Y$\sim$PI}
+\end{equation}
+should obey:
+\begin{equation}\label{eq:TI} \tag{TR}
+ \mathsf{a}'  = ( {\bcolor\alpha '}  {\pcolor \beta' }+{\pcolor\pi'} ) \text{ and } 
+\mathsf{b}'  = ( {\bcolor\alpha '}  {\icolor \gamma'} + {\icolor \mu'}).
+\end{equation}
+
+Without any exclusion restrictions, this equality is just a decomposition of total effects
+into direct and indirect components, and is not a testable restriction. However, in our case we rely on the SIR
+model with testing to motivate the presence of change in testing rate as a confounder in the outcome
+equations, but not in the behavior equation (implying that a component of $\delta_B$ and $\delta_P$ is set to 0), implying that (TR) does not necessarily hold and is testable.
+Furthermore, we estimate (\ref{eq:R2}) on the data set that has many more observations than the data set used to estimate the outcome equations, implying that (TR) is again testable.
+Later we  shall also try to
+utilize the contextual knowledge that masks business mandates only affect outcome directly and not by changing mobility, implying again that (TR) is testable.
+If not rejected by the data, (TR) can be used to sharpen the estimate of the causal effect of mask policies on the outcomes.
+
+Validation of the model by (TR) allows us to check exclusion restrictions brought by contextual knowledge,
+and check stability of the model by using different data subsets.  However, passing the (TR) does 
+not, of course, guarantee that the model is necessarily valid for recovery of causal effects.
+The only fundamental way to truly validate a causal model for observational data is through a controlled experiment, which is impossible to carry  in our setting.
+We shall however cross-check our results against other scientific evidence and, for example, results from RCT on effectiveness of masks on respiratory deceases (though not in the Covid-19 context).
+
+The parameters of the SEM are identified by the projection equation set above, provided the latter are identified by sufficient joint variation of these variables across states and time.
+We can develop this point formally as follows.
+Consider the previous system of equations, after partialling out the confounders:
+\begin{equation}\label{eqn:SEM-PO}
+ \begin{aligned}
+   & {\ycolor { \tilde Y_{i,t+\ell}}}  & =& {\bcolor {\alpha ' \tilde B_{it}}} +{\pcolor {\pi '\tilde P_{it}}} +{\icolor \mu'\tilde I_{it} } + \varepsilon^y_{it},&\varepsilon^y_{it} &\perp {\tilde B_{it}}, {\tilde  P_{it}}, { \tilde I_{it}},  \\
+   & {\bcolor { \tilde B_{it}}} & = &{\pcolor { \beta' \tilde P_{it}} } + {\icolor { \gamma' \tilde I_{it}} } + \varepsilon^b_{it},
+   &\quad \varepsilon^b_{it} &\perp { \tilde P_{it}}, { \tilde I_{it}},  \\
+   & {\pcolor { \tilde P_{it}} } &= &  {\icolor {\eta' \tilde I_{it}} } +   \varepsilon^p_{it}, &\quad \varepsilon^p_{it} &\perp {\tilde I_{it}}  \\
+     & {\ycolor { \tilde Y_{i,t+\ell}}}  & =&  \mathsf{a} ' {\pcolor {\tilde P_{it}}} + \mathsf{b}' {\icolor \tilde I_{it} } + \bar \varepsilon^y_{it},&\bar \varepsilon^y_{it} &\perp  {\tilde  P_{it}}, { \tilde I_{it}},  \\
+   \end{aligned}
+\end{equation}
+where $ \tilde V_{it} = V_{it}   -     {\wcolor W_{it}'} \Ep[{\wcolor W_{it}W_{it}'}]^{-} \Ep[{\wcolor W_{it}} V_{it}]$ denotes
+the residual after removing the orthogonal projection of $V_{it}$ on ${\wcolor W_{it}}$.
+The residualization is a linear operator, implying that (\ref{eqn:SEM-PO}) follows immediately from the above.
+The parameters of (\ref{eqn:SEM-PO})  are identified as projection coefficients in these equations, provided that residualized vectors appearing in each of the equations have non-singular variance, that is
+ \begin{equation}
+ \Var ( {\pcolor \tilde P_{it}'} ,{\bcolor \tilde B_{it}'},{\icolor \tilde I_{it}'})>0,
+ \ \Var ({\pcolor \tilde P_{it}'}, {\icolor \tilde I_{it}'})> 0 , \ \text{ and  }  \Var ( {\icolor \tilde I_{it}'}) >0.
+ \end{equation}
+
+Our main estimation method is the standard correlated random effects estimator, where the random effects
+are parameterized as functions of observable characteristic, ${\wcolor W_{it}}$, which include both state-level and time random effects.  The state-level random effects
+are modeled as a function of state level characteristics, and the time random effects are modeled
+by including month dummies and their interactions with state level characteristics.
+The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$ are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.
+
+Another modeling approach is the fixed effects panel data model, where ${\wcolor W_{it}}$ includes
+\text{latent} (unobserved) state level confonders ${\wcolor W_i}$ and  and time level effects ${\wcolor W_t}$,
+which must be estimated from the data.
+This approach is much more demanding of the data and relies on long time and cross-sectional histories to estimate ${\wcolor W_i}$ and ${\wcolor W_t}$, resulting in amplification of uncertainty.
+In addition, when histories are relatively short,  large biases emerge and they need to be removed using debiasing methods, see e.g., \cite{chen2019mastering} for overview.
+We present the results
+on debiased fixed effect estimation as parts of our sensitivity analysis. Other approaches we have explored
+is to consider more flexible nonlinear model $f(W_{it}, \delta)$ for the way observed confounders enter the model,
+for example partialling out these effects using Random Forest, using the debiased machine learning approach.
+All of these results are reported as parts of the robustness checks.
+
+With exclusion restrictions there are multiple approaches to estimation, for example,
+via generalized method of moments. We shall take a more pragmatic approach where we estimate
+the parameters of equations separately, and then compute 
+\begin{equation}
+ \frac{1}{2} \hat {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\pcolor \hat \beta' }+{\pcolor\hat \pi'} ) \text{ and } 
+ \frac{1}{2} \hat{\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\icolor \hat \gamma'} + {\icolor \hat \mu'}),
+\end{equation}
+as the estimator of the total policy effect, that concentrates around  their population analogues
+\begin{equation}
+ \frac{1}{2}  {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \alpha '}  {\pcolor \beta' }+{\pcolor \pi'} ) \text{ and } 
+ \frac{1}{2} {\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor  \alpha '}  {\icolor \gamma'} + {\icolor \mu'}),
+\end{equation}
+with approximate deviations controlled by the normal laws, with standard deviations that can be approximated by the bootsrap
+resampling of observational units $i$. Under correct specification the target quantities
+reduce to $  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} $ and ${\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'}$, respectively.
+This construction is not as efficient as generalized method of moments,
+but has a nicer interpretation under possible misspecification of the model: we are combining predictions
+from two models, one motivated via the causal path, reflecting contextual knowledge, and another from a ``reduced form" model not exploiting the path.
+The combined estimator can improve on precision of either estimator.
 
@@
 
-\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel}
+\subsection{Counterfactual Policy Analysis}
+
+We  also consider simple counterfactual exercises, where we examine the effects of setting
+a sequence of counterfactual policies for each state:
+$$
+\{{\pcolor P^\star_{it}} \}_{t=1}^T, \quad i=1, \ldots N.
+$$
+We assume that the SEM remains invariant, except for the policy equation.\footnote{It is possible to consider counterfactual exercises in which policy  responds to information through the policy equation if we are interested in endogenous policy responses to information.
+Although this is beyond the scope of the current paper, counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information.
+} Given the policies, we generate the counterfactual
+outcomes, behavior, information by propaging the dynamic equations:
+\begin{equation} \label{eqn:CF-SEM} \tag{CEF-SEM}
+  \begin{aligned}
+& {\ycolor Y^\star_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B^\star_{it} } , {\pcolor P^\star_{it}}, {\icolor I^\star_{it}}), \\
+& {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
+& {\icolor I^\star_{it} }& := &  I_{it}({\icolor I^\star_{i,t-\ell}}, {\ycolor Y^*_{it}}, t),  \end{aligned}
+\end{equation}
+with the same initialization as the factual system up to $t \leq 0$.
+In stating this counterfactual system of equations, we assume that structural outcome equations (SO) and information equations (I) remain invariant, and so do the stochastic shocks, decomposed
+into observable and unobservable parts. Formally, we record this assumption and above discussion as follows.
+
+\begin{quote}
+\textsc{Counterfactual Structural Equations Model (CF-SEM)}.
+Let $i \in \{1,..., N\}$ be the observational init and $t$ be time periods and $\ell$ the time delay, a positive integer.
+(1) For each $i$ and $t \leq 0$, the confounder, information, behavior, policy, and outcome variables are determined as in SEM: $W^\star_{it}= W_{it}$, $I^\star_{it}= I_{it}$,  $B^\star_{it} = B_{it}$, $P^\star_{it}=P_{it}$, $Y^\star_{it} =Y_{it}$.
+(2) For each $i$ and $t \geq 0$, confounders $W^\star_{it} =W_{it}$ are determined, then information variables $I^\star_{it}$ are determined  by (I), and policy variables $P^\star_{it}$ are set,  the behavior variables $B^\star_{it}$ are determined by setting $\iota = I^\star_{it}$ and $p= P^\star_{it}$ in  (SO) and keeping  the same stochastic shock $\varepsilon^b_{it}$ in (SO), the outcome $Y^\star_{i, t + \ell}$ is realized by setting $\iota = I^\star_{it}$, $p= P^\star_{it}$, and $b = B^\star_{it}$ in  (SO) and keeping the same shochastic shock $\varepsilon^y_{it}$  (SO).
+\end{quote}
+
+The counterfactual outcome $Y^\star_{i,t+\ell}$  and  factual outcome $Y_{i,t+\ell}$  are given by:
+$$
+ {\ycolor Y^\star_{i,t+\ell}} = ( {\bcolor\alpha '} {\pcolor \beta'}  + {\pcolor\pi'})
+    {\pcolor P^\star_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
+    {\icolor I^\star_{it} }+ {\wcolor \bar \delta 'W_{it}} + \bar \varepsilon_{it}^y, $$$$\ {\ycolor Y_{i,t+\ell}} \ =
+    ( {\bcolor\alpha '}  {\pcolor \beta'} + {\pcolor\pi'})
+    {\pcolor P_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
+    {\icolor I_{it} }+ {\wcolor \bar \delta 'W_{it}} + \bar \varepsilon_{it}^y.
+$$
+In generating these predictions, we explore the assumption of invariance stated above.
+We can write  the counterfactual contrast into the sum of three components:
+  \begin{align}
+ \underbracket{{\ycolor Y^\star_{i,t+\ell} -  Y_{i,t+\ell} }}_{\text{Predicted CF Change}} & =
+\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}  +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \nonumber\\
+&\qquad +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \nonumber \\
+ & =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}, \label{eqn:DC2}
+  \end{align}
+ which describe the immediate indirect effect of the policy via behavior, the direct effect of the policy, and the dynamic effect
+ of the policy.
+
+The dynamic effects take the form:
+$$
+\begin{array}{lll}
+  & \mathrm{DynE}^\star_{it} &=
+  \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}^\star_{i,t-m \ell } +  \mathrm{PED}^\star_{i,t- m\ell} \right)
+  \\
+ &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}^\star_{i,t-(m +1) \ell } +  \mathrm{PED}^\star_{i,t- (m+1)\ell}  \right),
+ \end{array}
+ $$
+ interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
+ The effects can be decomposed into (a) delayed policy effects via behavior by
+summing terms containing $PEB$ and (b) delayed policy effects via direct impact
+by summing terms containing $PED$.
+
+In order to compute the counterfactual relative contrasts for levels variables, for example, confirmed cases $C_{it}$ 
+we would use the formula:
+\begin{equation}
+C^{\star}_{it}/C_{it} = \exp(  \sum_{\bar t=1}^t ( Y^\star_{i\bar t} -  Y_{i\bar t}) ),
+\end{equation}
+and similarly for fatalities. The counterfactual relative contrast for national level variables takes the form:
+\begin{equation}
+C^{\star}_{it}/C_{it} =  \sum_{i=1}^N w_i \exp(  \sum_{\bar t=1}^t ( Y^\star_{i\bar t} -  Y_{i\bar t}) ), \quad w_i = C_{it}/\sum_{i=1}^N C_{it}.
+\end{equation}
+Note this counterfactual contrast is conditional on the actual history $\{C_{it}\}$.  If we'd want unconditional counterfactual,
+we'd need to make assumptions about the evolution of stochastic shocks appearing in $Y_{it}$.\footnote{See, e.g, previous
+versions of our paper in ArXiv, where unconditional counterfactuals were calculated by assuming stochastic shocks
+are i.i.d. and resampling them from the empirical distribution. The differences between conditional and unconditional 
+contrasts were small in our empirical analysis.}
+
+\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel} We next provide details
+of our key measurement equations, defining the outcomes and key confounders. We motivate the structural
+outcome equations via the fundamental epidemiological model for the spread of infectious decease called
+the SIRD (susceptible - infectected - recovered-deceased) model with testing.
+
 Letting $C_{it}$ denote the cumulative number of  confirmed cases in state $i$ at time $t$, our outcome
@@
 is the key confounding variable,
-derived from considering the SIR model below. We describe other confounders in the empirical section.
+derived from considering the SIR model below. We are treating the change in testing
+rate as exogenous.\footnote{To check sensitivity to this assumption
+we performed robustness checks, where we used
+the further lag of $\Delta   \log(T_{it} )$ as a proxy for exogenous change in the testing rate, and we also
+used that as an instrument for $\Delta   \log(T_{it} )$; this did not affect the results on policy effects, although the instrument
+was not sufficiently strong.} We describe other confounders in the empirical section.
 
@@
 
- Our main estimating equation (\ref{eq:M}) is motivated by a variant of SIR
+ Our main measurement equation (\ref{eq:M}) is motivated by a variant of SIRD
 model, where we add confirmed cases and infection detection via testing.
 Let $S$, $\Infected$, $\Recovered$, and $D$ denote the number of susceptible,
-infected, recovered, and dead individuals in a given state. Each of these variables are a function of time. We model
+infected, recovered, and deceased individuals in a given state. Each of these variables are a function of time. We model
 them as evolving as
@@
 
-\subsection{Growth Rate in Deaths as Outcome}\label{sec:sirmodel-death}
-By differentiating (\ref{eq:d}) and (\ref{eq:c}) with respect to $t$ and using (\ref{eq:c2}), we obtain
+The specification for growth rate in deaths as outcome  is motivated by SIRD as follows. By differentiating (\ref{eq:d}) and (\ref{eq:c}) with respect to $t$ and using (\ref{eq:c2}), we obtain
 \begin{align}
@@
 %The appendix provides additional results under alternative assumptions including non-deterministic disease durations.
-
-\section{Decomposition and Counterfactual Policy Analysis}
-
-\subsection{Total Change Decomposition}
-Given the SEM formulation above, we can carry out the following decomposition analysis, after removing the effects of confounders. For example,
-we can decompose the total change $\Ep \tilde Y_{i,t+\ell} - \Ep \tilde Y_{io}$ in the expected outcome, measured at two different time points $t+\ell$ and $o$ into the sum of three components:
-\begin{equation} \label{eqn:DC}
-  \begin{aligned}
-\underbracket{\Ep \tilde Y_{i,t+\ell} - \Ep \tilde Y_{io}}_{\text{Total Change}}& =
-{\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left( \Ep \tilde P_{it} - \Ep \tilde P_{io} \right)}}_{\text{Policy Effect via Behavior}}} \\
-& +    {\underbracket{\pcolor\pi' \left( \Ep \tilde P_{it} - \Ep \tilde P_{io} \right)}_{\text{Direct Policy Effect}} } \\
-& +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( \Ep \tilde I_{it} - \Ep \tilde I_{io} \right)}+ {\icolor{\mu}'  \left( \Ep \tilde I_{it} - \Ep \tilde I_{io} \right)}}_{\text{ Dynamic Effect}}} \\
-& =:   \mathrm{PEB}_t + \mathrm{PED}_t + \mathrm{DynE}_t,
-  \end{aligned}
-\end{equation}
-where the first two components capture the immediate effect and the third represents the delayed or dynamic effect.
-
-In the three examples of information structure given earlier, we have the following forms for the dynamic effect: for the trend model,
-$$
-   \mathrm{DynE}_t   = (\gamma \alpha + \mu) \Delta g_t, \quad  \Delta g_t = (g(t) - g(t-\ell))\\
-$$
-and for the lag model, $$
-   \mathrm{DynE}_t  =    \sum_{m=1}^{t/\ell} \left( \gamma \alpha+ \mu   \right)^m (\mathrm{PEB}_{t-m \ell }  +\mathrm{PED}_{t-m \ell } ),
- $$
-  interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
- For the general model we use, the dynamic effect is
- $$
- \begin{array}{lll}
-(\mathrm{I})    & \mathrm{DynE}_t  & = \sum_{m=0}^{t/\ell} \left(( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m  \ \ ( (\gamma \alpha)_{1} + \mu_1) \Delta g_t  \\
-& & +   \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}_{t-m \ell } +  \mathrm{DPE}_{t- m \ell} \right)
-  \\
- &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}_{t-(m +1) \ell } +  \mathrm{DPE}_{t- (m+1) \ell}  \right).
- \end{array}
- $$
- The effects can be decomposed into (a) delayed policy effects via behavior by
-summing terms containing $PEB$, (b) delayed policy effects via direct impact
-by summing terms containing $DPE$,   (c) pure behavior effects, and (d)  pure dynamic feedback effects.
-
-\subsection{Counterfactuals}
-
-We  also consider simple counterfactual exercises, where we examine the effects of setting
-a sequence of counterfactual policies for each state:
-$$
-\{{\pcolor P^\star_{it}} \}_{t=1}^T, \quad i=1, \ldots N.
-$$
-We assume that the SEM remains invariant, except of course for the policy equation.\footnote{It is possible to consider counterfactual exercises in which policy  responds to information through the policy equation if we are interested in endogenous policy responses to information.
-Although this is beyond the scope of the current paper, counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information.
-} Given the policies,
-we propagate the dynamic equations:
-\begin{equation} \label{eqn:CF-SEM} \tag{CEF-SEM}
-  \begin{aligned}
-& {\ycolor Y^\star_{i,t+\ell}}  & := & Y_{i,t+\ell} ( {\bcolor B^\star_{it} } , {\pcolor P^\star_{it}}, {\icolor I^\star_{it}}), \\
-& {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
-& {\icolor I^\star_{it} }& := &  I_{it}({\icolor I^\star_{i,t-\ell}}, {\ycolor Y^*_{it}}, t),  \end{aligned}
-\end{equation}
-with the initialization $I^\star_{i0} = 0$, $ Y^\star_{i0} =0$, $B^\star_{i0} = 0$, $P^\star_{i0} = 0$.
-In stating this counterfactual system of equations, we make the following invariance assumption
-\begin{quote}
-\textsc{Invariance Assumption.} The equations of (CF-SEM) remain exactly of the same form
-as in the (SEM) and (I).
-That is, under the policy intervention $\{{\pcolor P^\star_{it}} \}$,  parameters and stochastic shocks in (SEM) and (I) remain the same as under the original policy intervention $\{{\pcolor P_{it}} \}$.
-\end{quote}
-
-Let $\mathcal{P} Y^\star_{i,t+\ell}$  and  $\mathcal{P} Y_{i,t+\ell}$ denote the predicted values produced
-by working with the counterfactual system (\ref{eqn:CF-SEM}) and the factual system (SEM):
-$$
-\mathcal{P} Y^\star_{i,t+\ell}  = ( {\bcolor\alpha '} {\pcolor \beta'}  + {\pcolor\pi'})
-    {\pcolor P^\star_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
-    {\icolor I^\star_{it} }+ {\wcolor \bar \delta 'W_{it}},$$$$\mathcal{P} Y_{i,t+\ell} \ =
-    ( {\bcolor\alpha '}  {\pcolor \beta'} + {\pcolor\pi'})
-    {\pcolor P_{it}} + ({\bcolor\alpha '}  {\icolor \gamma' + \mu'})
-    {\icolor I_{it} }+ {\wcolor \bar \delta 'W_{it}}.
-$$
-In generating these predictions, we make the assumption of invariance stated above.
-
-Then we can write  the difference into the sum of three components:
-  \begin{align}
-\underbracket{\mathcal{P} Y^\star_{i,t+\ell} - \mathcal{P} Y_{i,t+\ell} }_{\text{Predicted CF Change}} & =
-\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}  +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \nonumber\\
-&\qquad +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \nonumber \\
- & =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}. \label{eqn:DC2}
-  \end{align}
-%\begin{equation} \label{eqn:DC2}
-%  \begin{aligned}
-%\underbracket{\mathcal{P} Y^\star_{it} - \mathcal{P} Y_{it} }_{\text{Predicted CF Change}}& =
-%{\underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}} \\
-%& +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \\
-%& +  {\underbracket{{\bcolor\alpha}' {\icolor{\gamma}'  \left( I^\star_{it} - I_{it} \right)}+ {\icolor{\mu}'  \left( I^\star_{it} -  I_{it} \right)}}_{\text{ CF Dynamic Effect}}} \\
-%& =:   \mathrm{PEB}^\star_{it} + \mathrm{PED}^\star_{it} + \mathrm{DynE}^\star_{it}.
-%  \end{aligned}
-%\end{equation}
-
-Similarly to what we had before,  the counterfactual dynamic effects take the form:
-$$
-\begin{array}{lll}
-(\mathrm{I})  & \mathrm{DynE}^\star_{it} &=
-  \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}^\star_{i,t-m \ell } +  \mathrm{DPE}^\star_{i,t- m\ell} \right)
-  \\
- &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}^\star_{i,t-(m +1) \ell } +  \mathrm{DPE}^\star_{i,t- (m+1)\ell}  \right),
- \end{array}
- $$
- interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
- The effects can be decomposed into (a) delayed policy effects via behavior by
-summing terms containing $PEB$, (b) delayed policy effects via direct impact
-by summing terms containing $DPE$,  (c) pure behavior effects, and (d)  pure dynamic feedback effects.
 
@@
          Each off-diagonal entry reports a correlation coefficient of
-         a pair of policy and behavior variables.
-``ave of four policy vars'' is defined by the average of stay-at-home, closured movie theaters, closured restaurants, and closured non-essential businesses.
+         a pair of policy and behavior variables. ``business closure policies'' is defined by the average of  closured movie theaters, closured restaurants, and closured non-essential businesses. 
        \end{flushleft}
@@
 %, but this does not prevent us from identifying the aggregate effect of all policies and behavior variables on case or death growth. 
-For this reason, we also consider a specification that includes
-the average of stay-at-home, closured movie theaters, closed  restaurants, and closured non-essential businesses in place of these four policy variables separately.
+For this reason, we define  ``business closure policies'' variable by the average of closured movie theaters, closed  restaurants, and closured non-essential businesses variables and consider a specification that includes business closure policies   in place of these three policy variables separately.
 
@@
 people's behaviors at $t$. As information, we include each state's
-growth of cases (in panel \ref{tab:PItoBcases}) or deaths (in panel
-\ref{tab:PItoBdeaths}), and log cases or deaths.  Additionally, in
-columns (5)-(8) of each panel, we include national growth and log
+growth of cases (in panel A) or deaths (in panel B), and log cases or deaths.  Additionally, in
+columns (5)-(8) of Table \ref{tab:PItoB}, we include national growth and log
 of cases or deaths.
@@
 All specifications include state-level characteristics (population, area, unemployment rate, poverty rate,  a percentage of people subject to illness, and governor's party) as well as their interactions with the log of days since Jan 15, 2020.
-The row ``$\sum_j \text{Policy}_j$'' reports the sum of six policy coefficients.
+The row ``$\sum_j \text{Policy}_j$'' reports the sum of four policy coefficients.
 Standard errors are clustered at the state level.
@@
 The other policy coefficients are less sensitive to the inclusion of
-national case/death variables. After school closures, stay-at-home
-orders and restaurant closures have the next largest effect. Somewhat
-surprisingly, closure of nonessential businesses appears to have
-a modest effect on behavior. Closing movie theaters has a similar,
-small effect on behavior. The effect of masks for employees is also
-small. The comparison of effects across policies should be interpreted
-with caution.  Differences between policy effects are often
-statistically insignificant; except for masks for employees, the
-policies are highly correlated and it is difficult to separate their
-effects.
+national case/death variables. After school closures, business closure policies has the next largest effect followed by stay-at-home
+orders.  The effect of masks for employees is  
+small. 
+%The comparison of effects across policies should be interpreted
+%with caution.  Differences between policy effects are often
+%statistically insignificant; except for masks for employees, the
+%policies are highly correlated and it is difficult to separate their
+%effects.
 
@@
  in early April. The estimated policy coefficients in columns imply
-that imposing all six policies would lead to roughly 85\% (in columns
-1-4) or roughly 50\% (in columns 5-8) of the observed decline.
+that imposing all policies would lead to roughly 75\% (in column 
+4) or roughly 35\% (in column 8) of the observed decline.
 The large impact of policies on transit intensity suggests that the policies may have reduced the Covid-19 infection by reducing people's use of public transportation.\footnote{Analyzing the New York City's subway ridership,  \cite{NBERw27021} finds  a strong link between public transit and spread of infection.}
@@
 
-In panel \ref{tab:PItoBdeaths}, estimated coefficients of deaths and
+In Table \ref{tab:PItoB}(B), estimated coefficients of deaths and
 death growth are generally negative. This suggests that the higher
@@
 prevalence of Covid-19 \citep{maloney2020}.  The coefficients on log
-cases and case growth in panel \ref{tab:PItoBcases} are more mixed. In
+cases and case growth in Table \ref{tab:PItoB}(A) are more mixed.\footnote{Rewrite a regression specification after omitting other variables as 
+$B_{it} =  \gamma_1 \Delta \log\Delta C_{it} + \gamma_2   \log\Delta C_{it} =  (\gamma_1 +\gamma_2) \log\Delta C_{it} - \gamma_1   \log\Delta C_{i,t-7}$.
+In columns (1)-(4) of Table \ref{tab:PItoB}(A),  the estimated values of both $(\gamma_1 +\gamma_2)$ and $-\gamma_1$ are negative except for grocery.
+This suggests that  a higher \textit{level} of confirmed cases reduces people's mobility in workplaces, retails, and transit.
+For grocery, the positive estimated coefficient of $(\gamma_1 +\gamma_2)$ may reflect stock-piling behavior in early pandemic periods.} In
 columns (5)-(8) of both panels, we see that national case/death
 variables have large, negative coefficients.
-This suggests that behavior responded to national conditions although it is also likely that national case/death variables capture unobserved aggregate time effects beyond information (e.g., latent policy variables and time-varying confounders that are common across states) which are not fully controlled by month dummies.
+This suggests that behavior responded to national conditions although it is also likely that national case/death variables capture unobserved aggregate time effects beyond information which are not fully controlled by month dummies (e.g., latent policy variables and time-varying confounders that are common across states).
 
@@
    \begin{tabular}{c|c}
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.5\linewidth}
      \centering
@@
      &
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.5\linewidth}
      \centering
@@
      corresponding daily measures.  The row
-     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
+     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of all policy
      coefficients.  The row ``$\sum_k w_k \mathrm{Behavior}_k$''
@@
      1st-10th.
-``ave of four policy vars'' is defined by the average of stay-at-home orders, closured movie theaters, closured restaurants, and closured non-essential businesses.
+``business closure policies'' is defined by the average of   closured movie theaters, closured restaurants, and closured non-essential businesses.
 Standard errors are clustered at the state level.
@@
 
-Column (1) represents our baseline specification while column (2) replaces stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses with the average of these four policy variables.
+Column (1) represents our baseline specification while column (2) replaces business closure policies with closed movie theaters, closed restaurants, and closed non-essential businesses.
 Columns (3) and (4) include past cases/deaths and growth rates at national level as additional regressors.
@@
 The estimates indicate that mandatory face masks for employees
-reduce the growth rate of infections and deaths by 8-15 percent, while holding behavior constant. This suggests that
+reduce the growth rate of infections and deaths by 9-15 percent, while holding behavior constant. This suggests that
 requiring masks for employees in public-facing businesses may be an effective preventive measure.\footnote{Note that we are \textit{not} evaluating the effect of \textit{universal} mask-wearing for the public but that of mask-wearing for employees.
@@
 
-The estimates also imply that stay-at-home orders reduce the growth rate of cases and deaths by 8-10 percent directly.
-This can be a reflection of multi-colinearity problem given high correlations among policy variables; in columns (2) and (4), the estimates of the average of stay-at-home orders, closed movie theaters, closed restaurants, and closed non-essential businesses are not significantly zero, indicating that the four policies in total may have little direct
-effect on case or death growth when behavior is held constant.
-
 Closing schools has a large and statistically
@@
 
+Neither the effect of stay-at-home orders nor that of business closure policies is estimated significantly different from zero, suggesting that these lockdown policies may  have little direct
+effect on case or death growth when behavior is held constant.
+
+%The estimates also imply that stay-at-home orders reduce the growth rate of cases and deaths by 5-8 percent directly.
+This can be a reflection of multi-colinearity problem given high correlations among policy variables; in columns (2) and (4), the estimates of the average of stay-at-home orders, closed movie theaters, closed restaurants, and closed non-essential businesses are not significantly zero, indicating that the four policies in total may have little direct
+%effect on case or death growth when behavior is held constant.
+
 The row ``$\sum_k w_k \mathrm{Behavior}_k$'' reports the sum of estimated coefficients weighted by the average of the behavioral
 variables from April 1st-10th.
-The estimates of $-0.73$ and $-0.79$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 73 and 79 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
+The estimates of $-0.80$ and $-0.84$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 80 and 84 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
 When including national cases and deaths in information, as shown in columns (3) and (4), the estimated aggregate impact of behavior is substantially smaller, but remains large and statistically significant.
 
-A useful practical implication of these results are that Google Mobility
-Reports and similar data might be useful as a leading indicator of
-potential case or death growth. This should be done with caution, however, because other
-changes in the environment might alter the relationship between behavior
-and infections. Preventative
-measures, including mandatory face masks, and changes in habit that are not captured in our data might alter
-the future relationship between Google Mobility Reports and case/death growth.
-
-The negative coefficients of past cases or deaths in
-% tables \ref{tab:BPItoY-c} and  \ref{tab:BPItoY-d}  are
+A useful practical implication of these results are that Google Mobility Reports and similar data might be useful as a leading indicator of potential case or death growth.
+This should be done with caution, however, because other changes in the environment might alter the relationship between behavior and infections.
+Preventative measures, including mandatory face masks, and changes in habit that are not captured in our data might alter the future relationship between Google Mobility Reports and case/death growth.
+
+The negative coefficients of the log of past cases or deaths in 
 Table \ref{tab:BPItoY} is
-consistent with a hypothesis that higher reported cases and deaths change people's behavior to reduce transmission risks. Such behavioral changes in response to new information   are
- partly captured by Google mobility measures, but the negative estimated coefficient of past cases or deaths imply that other latent behavioral changes that are not fully captured by Google mobility measures (e.g., frequent hand-washing, wearing masks, and keeping 6ft/2m distancing) are also important for reducing future cases and deaths.
-
-If policies are enacted and behavior changes, then future cases/deaths and
-information will change, which will induce further behavior changes.
-However, since the model includes lags of cases/deaths as well as their growth rates,  computing a long-run effect is not completely
-straightforward. We investigate dynamic effects that incorporate
-feedback through information in section \ref{counterfactuals}.
+consistent with a hypothesis that higher reported cases and deaths change people's behavior to reduce transmission risks.
+Such behavioral changes in response to new information   are  partly captured by Google mobility measures, but the negative estimated coefficient of past cases or deaths imply that other latent behavioral changes that are not fully captured by Google mobility measures (e.g., frequent hand-washing, wearing masks, and keeping 6ft/2m distancing) are also important for reducing future cases and deaths.
+
+If policies are enacted and behavior changes, then future cases/deaths and information will change, which will induce further behavior changes.
+However, since the model includes lags of cases/deaths as well as their growth rates,  computing a long-run effect is not completely straightforward.
+We investigate dynamic effects that incorporate feedback through information in section \ref{counterfactuals}.
 
@@
      corresponding daily measures.  The row
-     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
+     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of all policy
      coefficients.
-``ave of four policy vars'' is defined by the average of stay-at-home orders, closured movie theaters, closured restaurants, and closured non-essential businesses.
+``business closure policies'' is defined by the average of   closured movie theaters, closured restaurants, and closured non-essential businesses.
 Standard errors are clustered at the state level.
@@
 \afterpage{
-\begin{table}[!b]
+\begin{table}[!h]
   \caption{\label{tab:dieff-si}Direct and Indirect Policy Effects
@@
 \afterpage{
-\begin{table}
-  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
+\begin{table}[!h]
+  \caption{\label{tab:dieff-con-si}Direct and Indirect Policy Effects
+    without national case/death variables, \textbf{constraints imposed}}
 \begin{minipage}{\linewidth}
@@
     \\
-    \input{tables_and_figures/dieff-cases-NI}
-    \\ \\
+    \input{tables_and_figures/dieff-cases-constrain-SI}
+    \\
     \textbf{Deaths}
     \\
-    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
-    \\
+    \input{tables_and_figures/dieff-deaths-deathsinfo-constrain-SI}
   \end{tabular}
@@
   \begin{flushleft}
-      \scriptsize Direct effects capture the effect of policy on case
+    {\tiny Direct effects capture the effect of policy on case
       growth holding behavior, information, and confounders
@@
       Standard errors are computed by
-      bootstrap and clustered on state.
+      bootstrap and clustered on state.}
     %  \ref{tab:PtoY}.
@@
 \end{table}
-\clearpage}
+}
+
+%
+%
+%\afterpage{
+%\begin{table}
+%  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
+%\begin{minipage}{\linewidth}
+%  \centering
+%    \scriptsize
+%  \begin{tabular}{c}
+%    \textbf{Cases}
+%    \\
+%    \input{tables_and_figures/dieff-cases-NI}
+%    \\ \\
+%    \textbf{Deaths}
+%    \\
+%    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
+%    \\
+%  \end{tabular}
+%  \smallskip
+%  \begin{flushleft}
+%      \scriptsize Direct effects capture the effect of policy on case
+%      growth holding behavior, information, and confounders
+%      constant. Direct effects are given by ${\pcolor \pi}$ in
+%      equation (\ref{eq:R1}). Indirect effects capture how policy
+%      changes behavior and behavior shift case growth. They are given
+%      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
+%      from (\ref{eq:R2}). The total effect is
+%      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
+%      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
+%      \ref{eq:R4}. Columns ``Difference''    are  the differences between
+%      the estimates from (\ref{eq:R4}) and the combination of
+%      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
+%      Standard errors are computed by
+%      bootstrap and clustered on state.
+%    %  \ref{tab:PtoY}.
+%    \end{flushleft}
+%  \end{minipage}
+%\end{table}
+%\clearpage}
+%
+%
+%\afterpage{
+%\begin{table}
+%  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables, \textbf{Constraints imposed}}
+%\begin{minipage}{\linewidth}
+%  \centering
+%    \scriptsize
+%  \begin{tabular}{c}
+%    \textbf{Cases}
+%    \\
+%    \input{tables_and_figures/dieff-cases-constrain-NI}
+%    \\ \\
+%    \textbf{Deaths}
+%    \\
+%    \input{tables_and_figures/dieff-deaths-deathsinfo-constrain-NI}
+%    \\
+%  \end{tabular}
+%  \smallskip
+%  \begin{flushleft}
+%      \scriptsize Direct effects capture the effect of policy on case
+%      growth holding behavior, information, and confounders
+%      constant. Direct effects are given by ${\pcolor \pi}$ in
+%      equation (\ref{eq:R1}). Indirect effects capture how policy
+%      changes behavior and behavior shift case growth. They are given
+%      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
+%      from (\ref{eq:R2}). The total effect is
+%      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
+%      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
+%      \ref{eq:R4}. Columns ``Difference''    are  the differences between
+%      the estimates from (\ref{eq:R4}) and the combination of
+%      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
+%      Standard errors are computed by
+%      bootstrap and clustered on state.
+%    %  \ref{tab:PtoY}.
+%    \end{flushleft}
+%  \end{minipage}
+%\end{table}
+%\clearpage}
 
@@
 that all policies combined would reduce the growth rate of cases  and deaths by
-0.75 and 1.01, respectively,   out of which more than one-half to
+0.69 and 0.90, respectively,   out of which more than one-half to
 two-third is attributable to the indirect effect through their impact on behavior.  The estimate also indicates that the
@@
 
+Similarly to Table  \ref{tab:dieff-si}, the first three columns of  Table \ref{tab:dieff-con-si} report  the estimated direct and indirect effects of policy but imposing that masks for employees only affect cases/deaths directly without affecting behavior and that business closure policies only affects cases/deaths indirectly through its effect on behavior.\footnote{These are computed from estimating the specification in columns (1)-(4) of
+Table \ref{tab:PItoB} and column (1) of Table \ref{tab:BPItoY} but imposing that the coefficient of masks for employees is zero in (\ref{eq:R1}) and that the coefficient of business closure policies is zero in (\ref{eq:R2}).} The estimated total effect of masks for employees in the third column of Table \ref{tab:dieff-con-si} is higher than that of Table  \ref{tab:dieff-si}.
+Similarly, the total effect of business closure policies is estimated larger in Table \ref{tab:dieff-con-si} than in  Table  \ref{tab:dieff-si}.
+
+Column ``Difference''  in Tables   \ref{tab:dieff-si} and \ref{tab:dieff-con-si}  show the
+difference between the estimate of (\ref{eq:R4}) in column  ``PI$\to$Y Coefficient''   and the implied estimate from
+(\ref{eq:R1})-(\ref{eq:R2}) in   column ``Total.''   Differences  are generally small except for the coefficient of closed K-12 schools and  the sum of all policies in Table   \ref{tab:dieff-si}.
+The large differences in school closures  may be due to the  difficulty in identifying the effect of school closures given a lack of cross-sectional variations.
+Imposing the coefficients of masks for employees and business closure policies to be zero in (\ref{eq:R1}) and  (\ref{eq:R2}), respectively, does not increase differences between  ``Total" and  ``PI$\to$Y Coefficient''  as reported in the last column of  Table \ref{tab:dieff-con-si}; if anything the differences in mask coefficients become smaller in Table \ref{tab:dieff-con-si}, suggesting a validity of zero coefficients imposed in estimating the total effect.
+
+%and statistically insignificant, broadly supporting the validity of extra orthogonality condition in (\ref{eq:R1}).
+The difference for school closures as well as  the sum of all policies in Table   \ref{tab:dieff-si} is significantly different from zero, which may be due to the aforementioned difficulty in identifying the effect of school closures.
+There is substantial external epidemiological evidence that suggests that schooling closures may have substantial effects on the spread of the virus: studies like \cite{children:germany} and \cite{children:nature} establish that children carry substantial amounts of viral loads and can contribute to the transmission (due to higher contact rate than other age groups).\footnote{The evidence presented in \cite{children:germany}  has lead German to make the decision to close schools early.} The US data does not allow us to pint down the effect of closing schools reliably due to their approximate collinearity with trends in national cases.
+
 %In columns (2) and (4) of Table \ref{tab:PtoY}, we find that the estimated effect of mask mandates in April is larger than that in May for both case and death regressions.
 This may reflect a wider \textit{voluntary} adoption of masks in May than in April --- if more people wear masks even without mandatory mask policy,  the policy effect of mandating masks for employees becomes weaker.
-
-Table \ref{tab:dieff} presents the estimates for the specification with past national case/death variables.
-The effects of school closures and  the sum of policies are estimated substantially smaller in Table \ref{tab:dieff} when national case/death variables are included than in Table \ref{tab:dieff-si}.
-This sensitivity reflects the difficulty in identifying the aggregate time effect---which is largely captured by national cases/deaths---given little cross-sectional variation in the timing of school closures across states.
-On the other hand, the estimated effects of policies other than school closures are similar between Table \ref{tab:dieff-si} and Table \ref{tab:dieff}; the effect of other policies are well-identified from cross-sectional variations.
-
-Column ``Difference''  in Tables   \ref{tab:dieff-si} and \ref{tab:dieff}  show the
-difference between the estimate of (\ref{eq:R4}) in column  ``PI$\to$Y
-Coefficient''   and the implied estimate from
-(\ref{eq:R1})-(\ref{eq:R2}) in   column ``Total.''   Differences  are generally small and statistically insignificant, broadly supporting the validity of extra orthogonality condition in (\ref{eq:R1}).
-The difference for school closures as well as  the sum of all policies in Table   \ref{tab:dieff-si} is significantly different from zero, which may be due to the aforementioned difficulty in identifying the effect of school closures.
-There is substantial external epidemiological evidence that suggests that schooling closures may have substantial effects on the spread of the virus: studies like \cite{children:germany} and \cite{children:nature} establish that children carry substantial amounts of viral loads and can contribute to the transmission (due to higher contact rate than other age groups).\footnote{The evidence presented in \cite{children:germany}  has lead German to make the decision to close schools early.} The US data does not allow us to pint down the effect of closing schools reliably due to their approximate collinearity with trends in national cases.
 
@@
 Turning to the results, the estimates of Tables \ref{tab:dieff-si} and \ref{tab:dieff}  imply that all policies combined would reduce
-\(\Delta \log \Delta D\) by  0.97 and  0.48, respectively. For
+\(\Delta \log \Delta D\) by  0.93 and  0.39, respectively. For
 comparison, the median of \(\Delta \log \Delta D_{it}\) reached its
 peak in mid-March of about 1.3 (see Figure \ref{fig:growthq} in the
-appendix). Since then it has declined to near 0. Therefore,  -0.97 and -0.48 imply
+appendix). Since then it has declined to near 0. Therefore,  -0.93 and -0.39 imply
 that policy changes can account for roughly one-third to two-third of the observed
@@
 
+\subsection{Specifications, Timings, and Double Machine Learning}
 In this section, we provide sensitivity analysis by estimating (\ref{eq:R4}) with alternative specifications and methods.
-Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of stay-at-home, closed movie theaters, closed restaurants, and closed non-essential businesses for the following specifications and estimation methods:
+Figure \ref{fig:whisker} shows  the 90\% confidence intervals of   coefficients of (A) masks for employees, (B)  closed K-12 school, (C) stay-at-home, and (D) the average variable of  closed movie theaters, closed restaurants, and closed non-essential businesses for the following specifications and estimation methods:
   \begin{itemize}
-  \item[(1)] Baseline specification  in Table  \ref{tab:PtoY}.
+  \item[(1)] Baseline specification  in columns (1) and (3) of Table  \ref{tab:PtoY}.
   \item[(2)]  Exclude  the state of New York from the sample because it may be viewed as an outlier in the early pandemic period.
@@
    \item[(6)]   Include all additional controls in (3)-(5) with the sample that excludes New York as in (2). 
-   \item[(7)]   Estimated by Double Machine Learning (DML) of \cite{chernozhukov18} with Lasso to reduce dimensionality while adding private mask wearing rates and  the log of Trump's vote share as additional controls.
-\item[(8)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while adding private mask wearing rates and  the log of Trump's vote share as additional controls.
-   \item[(9)]   Add latent state confounders as components of $W_{it}$, estimated using fixed effects with bias correction.
+   \item[(7)] Add weekly dummies fo  to the baseline specification.
+   \item[(8)] Baseline specification is estimated by instrumenting $\Delta \log T_{it}$ with one week lagged  logarithm  value of the number of tests per 1000 people.
+   \item[(9)]   Estimated by Double Machine Learning (DML) of \cite{chernozhukov18} with Lasso to reduce dimensionality while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables  as additional controls.
+\item[(10)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables as additional controls.
+ %  \item[(10)]   Add latent state confounders and weekly dummies as components of $W_{it}$, estimated using fixed effects with bias correction.
   % \item Alternative Timing assumption ($\ell=10$ for cases and $\ell=23$ for deaths) on all of the above.
@@
 In Figure \ref{fig:whisker}, ``red'', ``green'', ``blue'', and ``purple" indicate the regression models for case growth without national information variables, case growth with national information variables, death growth without national information variables, and death growth with national information variables, respectively.
-The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 9 and 24 days, respectively, in the right panel of ``(ii) alternative timing.''\footnote{This alternative timing assumption is motivated by the lower bound estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
+The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 7 and 24 days, respectively, in the right panel of ``(ii) alternative timing.''\footnote{This alternative timing assumption is motivated by the lower bound estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
 We thank Jessica Metclaf for recommending us to do a sensitivity analysis on the timing assumption while suggesting this reference to us.}
 
-Panel (A) of Figure  \ref{fig:whisker} illustrates that the estimated coefficients of mask mandates are negative and significant across different specifications, methods, and timing assumptions, confirming the importance of mask policy on reducing case and death growths.
+Panel (A) of Figure  \ref{fig:whisker} illustrates that the estimated coefficients of mask mandates are negative and significant in most specifications, methods, and timing assumptions, confirming the importance of mask policy on reducing case and death growths.
 % Comparing (i) baseline timing on the left with (ii) alternative timing on the right,  we notice that the estimated magnitudes of masks for employees are similar to each other.
@@
 
-Our estimates of school closures substantially vary  across  specifications, however, and the estimated effects of school closures on case/death growth are often not significantly different from zero once national cases are controlled for.
+Our estimates of school closures substantially vary  across  specifications, however.
+In particular, the estimated effects of school closures on case/death growth become notably smaller once national cases/deaths or weekly dummies are controlled for.
 In the US state-level data,  there is little variation across states in the timing of school closures.  
-Consequently, its estimate is particularly sensitive to an inclusion of some aggregate variables such as national cases.
+Consequently, its estimate is particularly sensitive to an inclusion of some aggregate variables such as national cases or weekly dummies.
 Given this sensitivity,  there still exists a lot of uncertainty as to the magnitude of the effect of school closures.
@@
 
-Panel (C) of Figure  \ref{fig:whisker} indicates that the estimated coefficients of stay-at-home order  are generally negative and often significant 
- except for fixed effects estimator with bias correction.
-The sensitivity under fixed effects estimator may be due to high correlations between stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses.
-As shown in Panel (D), the  estimated coefficient of the average of  stay-at-home order  and closures of movie theaters, restaurants, and non-essential businesses   are  negative and often significant even under fixed effect specification; similar to school closure estimates, however,  its estimate is sensitive to controlling for national cases and deaths because the timing of implementing these policies is similar across states.
+Panel (C) of Figure  \ref{fig:whisker} indicates that the estimated coefficients of stay-at-home order  are generally negative and often significant, providing evidence that stay-at-home orders reduce the spread of COVID-19, although the estimates are sometimes sensitive to timing assumptions.
+Panel (D) shows that the estimated coefficients of business closure policies substantially vary across specifications, providing a mixed evidence for the effect of closures of businesses on case/death growth.
 
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-9} 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-whisker-7} 
     \end{tabular}
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-9}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pk12-whisker-7}
           \end{tabular}
@@
       $^\dagger$The times from exposure to case confirmation and death reporting  are assumed to be 14 and 21 days, respectively.
-$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 9 and 24 days, respectively.
+$^\ddagger$The times from exposure to case confirmation and death reporting  are assumed to be 7 and 24 days, respectively.
     \end{flushleft}
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-9} 
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pshelter-whisker-7} 
     \end{tabular}
@@
     \centering 
-     {\textbf{(D)  Average of stay-at-home, closed movie theaters, closed restaurants, and closed  businesses}}\\
+     {\textbf{(D)  Average of  closed movie theaters, closed restaurants, and closed  businesses}}\\
     \medskip
@@
       &
-      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-9}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pindex-whisker-7}
           \end{tabular}
@@
 \end{figure}
+
+\subsection{Fixed Effects Specification}
+
+Table \ref{tab:PtoY-fe} presents the results of estimating (\ref{eq:R4})  with  state fixed effects and weekly dummies.
+Because fixed effects estimator could be substantially biased when the time dimension is relatively short, we report not only the standard fixed effects estimator in columns (1) and (3) but also the debiased fixed effects estimator \citep[e.g.,][]{chen2019mastering} in columns (2) and (4), where the state-level clustered bootstrap standard errors are reported in all columns.
+
+The estimated coefficients of masks for employees largely confirms our finding that mandatory mask policy reduces the case and death growths.
+The estimated coefficients of stay-at-home orders and business closures are negative in columns (1) and (2) but their magnitudes as well as statistical significance are somewhat sensitive to whether bias corrections are applied or not.
+This sensitivity may be partly due to a high correction between stay-at-home orders variable and business closure policies.
+
+The results in Table \ref{tab:PtoY-fe} should be interpreted with caution.
+The fixed effects approach may not be preferred to random effects approach here because the former relies on long time and cross-sectional histories but, in our data, the effective time-dimension is short and a number of states is not large.
+Furthermore, the fixed effects approach could suffer more from  measurement errors, which could be a concern for our behavior and policy variables.
+
+  \afterpage{
+ \begin{landscape}
+\begin{table}[!htbp] \centering
+ \caption{\label{tab:PtoY-fe}
+Fixed Effects Specification: the Total Effect of Policies on Case and Death Growth ($PI \to Y$)}\vspace{-0.3cm}
+ \begin{minipage}{\linewidth}
+   \resizebox{\textwidth}{!}{
+   \centering
+   \tiny
+   \begin{tabular}{c|c}
+   \begin{minipage}{0.5\linewidth}
+     \centering
+     \input{tables_and_figures/cases-fe-piy}
+   \end{minipage}
+     &
+   \begin{minipage}{0.5\linewidth}
+     \centering
+     \input{tables_and_figures/deaths-fe-piy}
+   \end{minipage}
+
+   \end{tabular}
+   }
+   \begin{flushleft}
+     \scriptsize   All specifications include weekly dummies and state fixed effects. Dependent variable is the weekly growth rate of
+     confirmed cases (in the left panel) or deaths (in the right
+     panel) as defined in equation (\ref{eq:y}).
+Columns (1) and (3) report the standard fixed effects estimator  while Columns (2) and (4) report the fixed effects estimator with bias corrections.
+Standard errors are computed by multiplier bootstrap clustered at the state level.
+
+   \end{flushleft}
+ \end{minipage}
+\end{table}
+\end{landscape}
+}
 
@@
 
+\subsection{Direct and Indirect Policy Effects with national case/death variables}
+Tables \ref{tab:dieff} and \ref{tab:dieff-con}  present  the estimates for the specification with past national case/death variables.
+The effects of school closures and  the sum of policies are estimated substantially smaller in Table \ref{tab:dieff} when national case/death variables are included than in Table \ref{tab:dieff-si}.
+This sensitivity reflects the difficulty in identifying the aggregate time effect---which is largely captured by national cases/deaths---given little cross-sectional variation in the timing of school closures across states.
+On the other hand, the estimated effects of policies other than school closures are similar between Table \ref{tab:dieff-si} and Table \ref{tab:dieff}; the effect of other policies are well-identified from cross-sectional variations.
+
+\afterpage{
+\begin{table}
+  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
+\begin{minipage}{\linewidth}
+  \centering
+    \scriptsize
+  \begin{tabular}{c}
+    \textbf{Cases}
+    \\
+    \input{tables_and_figures/dieff-cases-NI}
+    \\ \\
+    \textbf{Deaths}
+    \\
+    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
+    \\
+  \end{tabular}
+  \smallskip
+  \begin{flushleft}
+      \scriptsize Direct effects capture the effect of policy on case
+      growth holding behavior, information, and confounders
+      constant. Direct effects are given by ${\pcolor \pi}$ in
+      equation (\ref{eq:R1}). Indirect effects capture how policy
+      changes behavior and behavior shift case growth. They are given
+      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
+      from (\ref{eq:R2}). The total effect is
+      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
+      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
+      \ref{eq:R4}. Columns ``Difference''    are  the differences between
+      the estimates from (\ref{eq:R4}) and the combination of
+      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
+      Standard errors are computed by
+      bootstrap and clustered on state.
+    %  \ref{tab:PtoY}.
+    \end{flushleft}
+  \end{minipage}
+\end{table}
+\clearpage}
+
+\afterpage{
+\begin{table}
+  \caption{\label{tab:dieff-con}Direct and Indirect Policy Effects with national case/death variables, \textbf{constraints imposed}}
+\begin{minipage}{\linewidth}
+  \centering
+    \scriptsize
+  \begin{tabular}{c}
+    \textbf{Cases}
+    \\
+    \input{tables_and_figures/dieff-cases-constrain-NI}
+    \\ \\
+    \textbf{Deaths}
+    \\
+    \input{tables_and_figures/dieff-deaths-deathsinfo-constrain-NI}
+    \\
+  \end{tabular}
+  \smallskip
+  \begin{flushleft}
+      \scriptsize Direct effects capture the effect of policy on case
+      growth holding behavior, information, and confounders
+      constant. Direct effects are given by ${\pcolor \pi}$ in
+      equation (\ref{eq:R1}). Indirect effects capture how policy
+      changes behavior and behavior shift case growth. They are given
+      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
+      from (\ref{eq:R2}). The total effect is
+      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
+      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
+      \ref{eq:R4}. Columns ``Difference''    are  the differences between
+      the estimates from (\ref{eq:R4}) and the combination of
+      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
+      Standard errors are computed by
+      bootstrap and clustered on state.
+    %  \ref{tab:PtoY}.
+    \end{flushleft}
+  \end{minipage}
+\end{table}
+\clearpage}
+
 \end{document}

### 2020-08-21 tex/robustness.tex
update

@@
    \begin{flushleft}
-     \scriptsize Dependent variable is the weekly growth rate of
+     \scriptsize   All specifications include weekly dummies and state fixed effects. Dependent variable is the weekly growth rate of
      confirmed cases (in the left panel) or deaths (in the right
-     panel) as defined in equation (\ref{eq:y}).        The covariates
-     include lagged policy variables, which are
-     constructed as 7 day moving averages between $t$ to $t-7$ of
-     corresponding daily measures.  The row
-     ``$\sum_j \mathrm{Policies}_j$'' reports the sum of six policy
-     coefficients.  Standard errors are clustered at the state level.
+     panel) as defined in equation (\ref{eq:y}).
+Columns (1) and (3) report the standard fixed effects estimator  while Columns (2) and (4) report the fixed effects estimator with bias corrections.
+Standard errors are computed by multiplier bootstrap clustered at the state level.
+
    \end{flushleft}

### 2020-08-21 cases_and_policies/rmd/bootstrap_felm.R

@@ -56,3 +56,73 @@ bootstrap_felm <- function(data, models, S=999) {
 
   return(betastar)
 }
+
+
+bootstrap_felm2 <- function(data, models, S=999) {
+  
+  
+  data[,"rowid"] <- 1:nrow(data)
+  fmlas <- lapply(models, function(x) x$formula)
+  
+  # find cluster identifier
+  cvar <- unique(sapply(fmlas, function(x) {
+    trimws(strsplit(as.character(x)[3], "\\|")[[1]][4])
+  }))
+  cvar[is.na(cvar)] <- "rowid" # no clustering
+  stopifnot(length(cvar)==1) # formulas have different clustering
+  
+  # # make sure keepCX==TRUE
+  # models <- lapply(models, function(m) {
+  #   if (m$keepCX) return(m)
+  #   m <- felm(m$formula, data=data, keepCX=TRUE)
+  # })
+  for (i in 1:length(models)){
+    sdf <- data
+    if (i <= 4){
+      sdf$pmaskbus <- 0*sdf$pmaskbus
+    } else if (i==5) {
+      sdf$pindex <- 0*sdf$pindex
+    }
+    m <- models[[i]]
+    if (m$keepCX) return(m)
+    models[[i]] <- felm(m$formula, data=sdf, keepCX=TRUE)
+  }
+  
+  
+  # extract index of observations in each model
+  idx <- lapply(models, function(m) {
+    yvar <- as.character(m$formula)[2]
+    data[,"rid"] <- data$rowid
+    data$rid[is.na(data[,yvar])] <- NA
+    f <- update.felm.formula(m$formula, rid ~ .)
+    return(felm(f, data=data)$response)
+  })
+  for (i in 1:length(models)) stopifnot(length(idx[[i]])==length(models[[i]]$cY))
+  
+  
+  betahat <- lapply(models, function(m) m$coef)
+  betastar <- list()
+  
+  # draw multipliers
+  cvals <- unique(data[,cvar])
+  gc <- matrix(rnorm(S*length(cvals)), nrow=S)
+  colnames(gc) <- cvals
+  gi <- gc[,data[,cvar]]
+  
+  # bootstrap
+  for (i in 1:length(models)) {
+    X <- models[[i]]$cX[,!is.na(models[[i]]$coef)]
+    score <- matrix(0, nrow=nrow(data), ncol=ncol(X))
[…]

### 2020-08-21 cases_and_policies/rmd/debiasedfe.R

@@ -68,6 +68,7 @@ multiple_split <- function(s, n, unit, time, dat, uc, fm) {
   return(acbc)
 }
 
+
 ##### DFE estimation based on cross-over jackknife
 # Note: The output of this function facilitates calculation of bootstrap se
 reg_fe <- function(df, # data
@@ -76,20 +77,20 @@ reg_fe <- function(df, # data
                    bvars, # behavior (NULL if not behavior)
                    x, # controls 
                    iv, # iv for felm, if not needed just input "0" 
-                   l = 14,
-                   fixedeffect) {
-  if (l == 0) {
+                   L,
+                   fe) {
+  if (L == 0) {
     # This is for pib
     p <- pols
     b <- bvars
   } else {
     # This is for pbiy and piy
-    p <- sprintf("lag(%s, %d)", pols, l)
-    b <- sprintf("lag(%s, %d)", bvars, l)
+    p <- sprintf("lag(%s, %d)", pols, L)
+    b <- sprintf("lag(%s, %d)", bvars, L)
   }
   # create regression formula
-  xvars <- c(p, b, x)
-  fmla <- createfmla_fe(yvar, xvars, fe = fixedeffect, iv = iv)
+  xvars <- c(p, b, x) 
+  fmla <- createfmla_fe(yvar, xvars, fe = fe, iv = iv)
   # two cross-over subsamples
   unit <- as.double(df$state)
   # Split by day. Data is linear so no bias coming from the time effects,
@@ -128,20 +129,63 @@ reg_fe <- function(df, # data
   } else {
     # for piy and pib no such metric
     beff <- NULL
-  } 
+  }
   # outputs: noncorrected, corrected, sum of policy and behavioral coefficients
   return(list(reg = whole, nobc = coef_nobc, bc = coef_cbc, bc2 = coef_cbc2, peff=peff, beff=beff, acbc = coef_acbc))
+  # return(list(reg = whole, nobc = coef_nobc, bc = coef_cbc, bc2 = coef_cbc2, peff=peff, acbc = coef_acbc))
 }
-
  
+　
+data_rg <- function(data, mle) {
+  n <- length(unique(data$state))
+  t <- length(unique(data$date))
+  # swap state index
+  ids <- kronecker(sample.int(n, n, replace = TRUE), rep(1, t))
+  data_b <- data[(ids - 1)*t + rep(c(1:t), n), ]
[…]

### 2020-08-21 cases_and_policies/rmd/generatetables.R

@@ -103,6 +103,8 @@ mainregressions <- function(df, # data
   plist <- list(pols, c("pmask.april","pmask.may", pols[-1]))
   plist <- list(pols,c("pmaskbus","pk12","pindex"))
   plist <- list(pols,c("pmaskbus","pk12","pshelter","pindex"))
+  plist <- list(pols,c("pmaskbus","pmaskall","pk12","pshelter","pmovie","prestaurant","pnonessential"))
+  plist <- list(pols,c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential"))
   
   
 
@@ -111,11 +113,14 @@ mainregressions <- function(df, # data
   ijs <- expand.grid(1:length(plist), 1:length(xlist))
 
   pbiy <- apply(ijs, 1, function(ij) {
+    # ia <- interactions[[ij[2]]]
+    # ia[[1]] <-  "lag(month,L)"
     policyreg(df, yvar, plist[[ij[1]]], bvars,
               c(sprintf("lag(%s, %d)", infovars[[ij[2]]], L),
                 tvars,
                 xlist[[ij[2]]]),
-              interactions[[ij[2]]], iv[[ij[2]]], L=L)
+            #  ia, iv[[ij[2]]], L=L)
+               interactions[[ij[2]]], iv[[ij[2]]], L=L)
   })
 
   ## # Get subset included in regressions
@@ -127,29 +132,55 @@ mainregressions <- function(df, # data
   ## df$insample <- NA
   ## df$insample[tmp$reg$response] <- 1
 
+  
 
-  piy <- apply(ijs, 1, function(ij) {
+  piy <- apply(ijs, 1, function(ij) {  
+    # ia <- interactions[[ij[2]]]
+    # ia[[1]] <-  "lag(month,L)"
     policyreg(df, yvar, plist[[ij[1]]], NULL,
               c(sprintf("lag(%s, %d)", infovars[[ij[2]]], L),
                 tvars,
                 xlist[[ij[2]]]),
-              interactions[[ij[2]]], iv[[ij[2]]], L=L)
+           #   ia, iv[[ij[2]]], L=L)
+                interactions[[ij[2]]], iv[[ij[2]]], L=L)
   })
+  # piy <- apply(ijs, 1, function(ij) {
+  #   policyreg(df, yvar, plist[[ij[1]]], NULL,
[…]

### 2020-08-21 cases_and_policies/rmd/regprep.R

@@ -134,9 +134,11 @@ df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/3
 
 tilt = ifelse(df$date> as.Date("2020-05-01"),1,0 )
 month <- month(df$date)
+week <- week(df$date)
 #month[month==6] <- 5
 #month[month<3] <- 3
 df$month = as.factor(month)
+df$week = as.factor(week)
 df$pmask.may <- df$pmaskbus*tilt
 df$pmask.april <- df$pmaskbus*(1-tilt)
 pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 

### 2020-08-21 cases_and_policies/rmd/varlabels.R

@@ -54,12 +54,15 @@ varlabels <-
     "prestaurant","closed restaurants", "closed restaurants except takeout",
     "Raifman et al",TRUE,
 
-    "pnonessential","closed businesses", "closed non-essential businesses",
+    "pnonessential","closed non-essent bus", "closed non-essential businesses",
     "Raifman et al",TRUE,
 
     "psoe","state of emergency","state of emergency", "Raifman et al",TRUE,
     
-    "pindex","ave of four policy vars","the average of stay-at-hom and closures of movie theaters, 
+    # "pindex","ave of four policy vars","the average of stay-at-hom and closures of movie theaters, 
+    # reataurants, and non-essential businesses", "Raifman et al",TRUE,
+    
+    "pindex","business closure policies","the average of closures of movie theaters, 
     reataurants, and non-essential businesses", "Raifman et al",TRUE,
     
     "party","Governor's party","Governor's party", "OPENICPSR",TRUE,

### 2020-08-22 tex/reply-to-referees.tex
update

@@
 \changes
+\subsection*{Reply to Comments}
+\begin{itemize}
+
+\item   \textit{``I particularly like your mapping of your estimating equation to the SIR model.
+But, it would be nice to discuss further the assumptions in light of the SIR model.
+For example, how do the orthogonality assumptions in (1) relate to the assumptions on top of page (14)?
+and just to be clear, you have a "causal model" yet for identification you rely on projections in the random effects specification (and no exclusion restriction).
+For example, what do you rule out vis a vis "diff in diff" type event study models with fixed effects?
+How about a brief discussion of the "invariance assumption" and why you think it would hold in your counterfactual exercise.
+''}
+
+\item   \textit{``It might make sense to reorganize Section 2 (which is the heart of the theory in the paper) into a simpler to read section  where the assumptions that are required for causal interpretation are laid out.
+Also, one easier threshold is that your system of estimating equations provide a direct "predictive effect" of various policies.
+These predictive (not causal) in my opinion would still be very useful.
+''}
+
+Following your suggestion, we have reorganized Section 2. 
+ We also state the following in abstract:
+ \begin{quote}
+ We stress that our study is observational and should be interpreted with great caution. 
+ From a completely agnostic point of view,
+our findings uncover predictive effects (association) of observed policies and behavioral changes  on future health outcomes,
+ controlling for informational and other confounding variables.
+\end{quote}
+
+\end{itemize}
 
@@
 We shall however cross-check our results against other scientific evidence and, for example, results from RCT on effectiveness of masks on respiratory deceases (though not in the Covid-19 context).
+ \end{quote}
+
+ We also state the following in abstract:
+ \begin{quote}
+ From a completely agnostic point of view,
+our findings uncover predictive effects (association) of observed policies and behavioral changes  on future health outcomes,
+ controlling for informational and other confounding variables.
 \end{quote}
@@
 \item[3.]  \textit{(Expectation and Lucas critique) ``Related to the previous comment, what would be the link
-between the current paper’s setup and how expectations are formed given the information
+between the current paper‚Äôs setup and how expectations are formed given the information
 structure? For example, is this paper safe from Lucas critique? Some comments would be
@@
 
-Thank you for your comment.
-As a part of sensitivity checks, we now examine if the estimates are sensitive when we include past behavior variables as information variables, under which policies depend on past behavior.
+Thank you for your comment. Following your comment, we added the following sentence:
+\begin{quote}
+The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
+\end{quote} 
+As a part of sensitivity checks, we now examine if the estimates are sensitive when we include two weeks lagged value of behavior variables in the information set, under which policies depend on past behavior.
 We find that the estimated coefficients of masks for employees are not sensitive to an inclusion of past behavior variables.
+
+We think it is reasonable to assume that people's behavior today will not immediately affect today's policy because it takes time for policy makers to gather information on people's behavior and make a policy decision based on gathered information, especially in the early pandemic period.
 
@@
 
+We have the following paragraph:
+\begin{quote}
+ To compute the estimated and counterfactual
+paths, we use the   estimate   in
+column ``Average'' of Table \ref{tab:dieff-si}. We set initial
+$\Delta \log \Delta C$ and $\log \Delta C$ to their values first
+observed in the state we are simulating. We hold all other regressors
+at their observed values. Error terms are drawn with replacement from
+the residuals. We do this many times and report the average over draws
+of the residuals. The shaded region is a point-wise 90\% confidence
+interval.
+\end{quote}
+
 \item[6.]  \textit{(Replication files) ``It would be great if the authors could post their dataset and replication
@@
 openicpsr/covid19.''}
+
+All programs and dataset will be available at our GitHub page, which is currently set to be private but we will plan to make it publicly accessible once the paper is accepted.
 
@@
 
+[Not sure of what to do here]
+
 \item[10.]  \textit{(Table 3A: Cases as Information) ``It is puzzling why the coefficients on  $\Delta\log\Delta C$ are positive
@@
 
+[We can move this to introduction. Or we can keep it as it is by saying that there could be confounders and many other things at play.]
+
   \end{itemize}
@@
 to the figure. The same questions apply to Figures 2 and 3 and other figures.''}
+
+Black and grey dots represent each state's change in growth rate over time.
+Given that it may be difficult to digest a counterfactual change in case/death growth in introduction, we decided to focus on relative effect of policies in Figures 1-3.
+
 \item[2.]  \textit{(Figure 4) ``What do S. \& P. stand for in Figure 4?''}
+
+It stands for 
+P. Wright and S. Wright as discussed in the footnote 2:
+\begin{quote}
+The father and son,
+P. Wright and S. Wright closely collaborated to develop structural equation models and causal path diagrams; P. Wright'Äôs
+key work represented supply-demand system as a directed acyclical graph and established its identification using
+exclusion restrictions on instrumental variables.
+\end{quote}
+
 \item[3.]  \textit{(Footnote 17) ``Is this necessary? This seems obvious. The authors could simply drop this
 footnote or make it more informative.''}
+
+Following your suggestion, we have dropped this footnote.
+
 \item[4.]  \textit{(Footnote 19) ``On page 17, in footnote 19, the authors state:
-We drop “Residential” because it is highly correlated with both “Workplaces” and
-“Retail \& recreation” at correlation coefficients of0:98.
-This sentence is ambiguous. Do the authors mean that both correlation coefficients are identically 0:98? Please rephrase this sentence.''}
+We drop ``Residential‚Ä'' because it is highly correlated with both ‚ÄúWorkplaces‚Äù and
+``Retail \& recreation‚Ä'' at correlation coefficients of 0.98.
+This sentence is ambiguous. Do the authors mean that both correlation coefficients are identically 0.98? Please rephrase this sentence.''}
+
+We rephrased this sentence as follows: `` We drop ``Residential'' because it is highly correlated with  ``Workplaces'' and  ``Retail \& recreation''  at correlation coefficients of -0.98 and -0.97, respectively.
+''
+
 \item[5.]  \textit{(Figures 9 and 11) ``On page 32, in Figure 9, the subfigures in the middle and right panels
@@
 other similar figures in the appendix.''}
+
+We have decided to drop these legends to improve the readability of those figures.
+
 \item[6. ]  \textit{(Intepretation) ``On page 13, it is state that
@@
 by the start of June? Please clarify the quantitative statement.''}
+
+It is in total by the start of June. We have modify the sentence as suggested.
+
 \item[7.]  \textit{``Some of the figures are difficult to read because there are too many curves. In particular,
@@
 some of them if they are less important.''}
+
+Following your suggestion, we have consolidated these figures and present figures with quantiles to improve the readability.
 
@@
 
+Black and grey dots represent each state's change in growth rate over time.
+Given that it may be difficult to digest a counterfactual change in case/death growth in introduction, we decided to focus on relative effect of policies in Figures 1-3.
+
+[We may change the starting date for counterfactual figures.]
+
 \item \textit{`` p5 - perhaps some papers from beyond the Econ literature would be relevant for a sense of the status of our knowledge of the epidemiogogy?  ''}
 
+We have reviewed and added more references from the epidemiology literature.
+
 \item \textit{`` p6 - simulations would produce exponential spread of disease - well, yes?
-The mechanism is reasonably well described, and the conditions to elude exponential growth are fairly crisp - even if you include 'a voluntary behavioral response' the outcome will be exponential growth unless you are in a regime where susceptible depletion is dominating dynamics, or you have shifted Re<1.
+The mechanism is reasonably well described, and the conditions to elude exponential growth are fairly crisp - even if you include 'a voluntary behavioral response' the outcome will be exponential growth unless you are in a regime where susceptible depletion is dominating dynamics, or you have shifted Re$<$1.
 ''}
+
+ Thank you for raising this issue. Yes, you are right. Following your comment, we have changed the sentence as follows:
+\begin{quote}
+When such a voluntary behavioral response is ignored, simulations would necessarily exhibit exponential spread and over-predict  cases and deaths.
+In contrast, as cases and deaths rise, a voluntary behavioral response may possibly reduce the effective reproduction number  below 1,  potentially preventing exponential spread.
+Our counterfactual experiments  illustrate the importance of this voluntary behavioral change.
+\end{quote}
 
@@
 
+We believe that it is better to use the number of cases and deaths rather than time since start of the outbreak as a  general framework to analyze the spread of COVID-19.
+This is especially true if we analyze the second wave of COVID-19, where time since start of the first outbreak may not be so relevant, and identifying the start of the second outbreak may be difficult.
+
 \item \textit{`` Figure 8 - I do like seeing the data, but it also just raises questions, as you say, confounders + many other things at play....
@@
 
+[This is something we don't do given the time/space constraint.]
+
 \end{enumerate}

### 2020-08-22 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

@@
 source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
+
 ```
@@
 
+# cmat <-  cor(df[,c(bvars,"residential",pols,"pindex")], use="pairwise.complete.obs")
+# scmat <-
+#   matrix(sprintf("%.2f",cmat),nrow=(length(bvars)+length(pols)+2))
 cmat <-  cor(df[,c(bvars,pols,"pindex")], use="pairwise.complete.obs")
@@
 infovars <- list(c("dlogdc", "logdc"),
-                 c("dlogdc", "logdc","dlogdc.national", "logdc.national")) #,
-#c("dlogdc", "logdc"))
-#infovars2 <- list(c( "logdc"))
-# tvars <-  "dlogtests"  
-# xlist <-list("party", "party") 
-# xlist <-list(c("party","logvote"), c("party","logvote")) 
-# xlist <- list("week","week")
-
+                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))  
 xlist <- list("","")
@@
 ilist <- list(interactions, interactions) 
-# iv <- list("(dlogtests ~ lag(logtests,7))","(dlogtests ~ lag(logtests,7))")
-# tvars <- ""
 iv <- list("0","0")
 tvars <- "dlogtests" 
-
 pols = c("pmaskbus","pk12","pshelter","pindex")
-#pols = c("pmaskbus","pmaskall","pk12","pshelter","pindex")
 
@@
 fig <- dwplot(results_df, conf.level = .90) + theme_bw() +
-    theme(       legend.justification=c(.01, .999), legend.position=c(.61, .99), 
+    theme(       legend.justification=c(.01, .999), legend.position=c(.62, .99), 
            legend.title = element_blank(), legend.background =
@@
   legend.title = element_text(size = 6),
-  legend.text = element_text(size = 5)
+  legend.text = element_text(size = 6)
   )# + ggtitle("School Closures")  
@@
   legend.title = element_text(size = 6),
-  legend.text = element_text(size = 5)
+  legend.text = element_text(size = 6)
   )# + ggtitle("Average of Four Policy Variables")  

### 2020-08-22 tex/covidEffectOfNPI-revision.tex
update

@@
 It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
-Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{
-The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, tinbergen1940, wold1954} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, tinbergen1940, wold1954} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
 See \cite{pearl:biometrika},\cite{pearl:robins},  \cite{pearl:causality}, \cite{white:chalak}, \cite{peters2020book}, \cite{heckman:pinto}, and \cite{hernanrobins2020book}  for modern developments, in computer
 science, econometrics, and epidemiology.
-The particular causal diagram has several “mediation” components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
-The father and son,
-P. Wright and S. Wright closely collaborated to develop structural equation models and causal path diagrams; P.Wright’s
-key work represented supply-demand system as a directed acyclical graph and established its identification using
-exclusion restrictions on instrumental variables.
-}
+The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
+The father and son, P.
+Wright and S.
+Wright closely collaborated to develop structural equation models and causal path diagrams; P.
+Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables.}
 %The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007}.
@@
 using mobility measures from PlaceIQ and SafeGraph.
-Using data from SafeGraph, \cite{anderson2020} show that there has been substantial voluntary social distancing but also provide evidence that mandatory measures such as stay-at-home orders have  been effective at reducing the frequency of visits  outside of one’s home.
+Using data from SafeGraph, \cite{anderson2020} show that there has been substantial voluntary social distancing but also provide evidence that mandatory measures such as stay-at-home orders have  been effective at reducing the frequency of visits  outside of one's home.
 % Overall, these evidence suggests the importance of voluntary social distancing behavior when we evaluate the effect of polices.
@@
 Furthermore, simulated models do not take into account that people may limit their contact with other people in response to higher transmission risks.\footnote{See \cite{atkeson2020a} and \cite{stock2020} for the implications of the SIR model for Covid-19 in the US.
-\cite{NBERw27128} estimate a SIRD model in which time-varying reproduction numbers depend on the daily deaths to capture feedback from daily deaths
-to future behavior and infections.}  When such a voluntary behavioral response is ignored, simulations would produce exponential spread of disease and would over-predict  cases and deaths.
+\cite{NBERw27128} estimate a SIRD model in which time-varying reproduction numbers depend on the daily deaths to capture feedback from daily deaths to future behavior and infections.}  When such a voluntary behavioral response is ignored, simulations would necessarily exhibit exponential spread and over-predict  cases and deaths.
+In contrast, as cases and deaths rise, a voluntary behavioral response may possibly reduce the effective reproduction number  below 1,  potentially preventing exponential spread.
 Our counterfactual experiments  illustrate the importance of this voluntary behavioral change.
@@
 \end{equation}
-approximates the weekly growth rate in new cases from $t-7$ to $t$.\footnote{We may show that $ \log( \Delta C_{it} ) -
-\log( \Delta C_{i,t-7})$ approximates the average
-growth rate of cases from $t-7$ to $t$.} Here $\Delta$ denotes the differencing operator over 7 days from $t$ to $t-7$, so that $\Delta C_{it}:=C_{it}-C_{i,t-7}$ is the number of new confirmed cases in the past 7 days.
+approximates the weekly growth rate in new cases from $t-7$ to $t$.
+Here $\Delta$ denotes the differencing operator over 7 days from $t$ to $t-7$, so that $\Delta C_{it}:=C_{it}-C_{i,t-7}$ is the number of new confirmed cases in the past 7 days.
 
@@
 Tracking Project}.  The number of tests for each state is from  \href{https://github.com/COVID19Tracking/covid-tracking-data}{Covid
- Tracking Project}. As shown in Figure \ref{fig:test} in the appendix, there was a rapid increase in testing in the second half of March and then the
+ Tracking Project}. As shown in the lower right panel of Figure \ref{fig:casedeathtest} in the appendix, there was a rapid increase in testing in the second half of March and then the
 number of tests increased very slowly in each state in April.
@@
 The dataset provides six measures of ``mobility trends''  that report a percentage change in visits and length of stay at different places relative to a baseline computed by their median values of the same day of the week from January 3 to February 6, 2020.
-Our analysis focuses on the following four measures: ``Grocery \& pharmacy," ``Transit stations,'' ``Retail \& recreation,'' and ``Workplaces.''\footnote{The other two measures are ``Residential'' and  ``Parks.''  We drop ``Residential'' because it is highly correlated with both ``Workplaces'' and  ``Retail \& recreation''  at correlation coefficients of -0.98.
+Our analysis focuses on the following four measures: ``Grocery \& pharmacy," ``Transit stations,'' ``Retail \& recreation,'' and ``Workplaces.''\footnote{The other two measures are ``Residential'' and  ``Parks.''  We drop ``Residential'' because it is highly correlated with  ``Workplaces'' and  ``Retail \& recreation''  at correlation coefficients of -0.98 and -0.97, respectively.
 We also drop ``Parks'' because it does not have clear implication on the spread of Covid-19.}
@@
 affected by the timing of reporting and testing, and are quite
-volatile as shown in Figure \ref{fig:dailycases} in the
+volatile as shown in the upper right panel of Figure \ref{fig:casedeathtest} in the
 appendix.  Aggregating to weekly new cases/deaths/tests smooths
@@
 %, but this does not prevent us from identifying the aggregate effect of all policies and behavior variables on case or death growth. 
-For this reason, we define  ``business closure policies'' variable by the average of closured movie theaters, closed  restaurants, and closured non-essential businesses variables and consider a specification that includes business closure policies   in place of these three policy variables separately.
+For this reason, we define  ``business closure policies'' variable by the average of closured movie theaters, closed  restaurants, and closured non-essential businesses variables, and consider a specification that includes business closure policies   in place of these three policy variables separately.
 
@@
    \begin{tabular}{c|c}
-   \begin{minipage}{0.5\linewidth}
+   \begin{minipage}{0.51\linewidth}
      \centering
@@
      &
-   \begin{minipage}{0.5\linewidth}
+   \begin{minipage}{0.51\linewidth}
      \centering
@@
    \begin{tabular}{c|c}
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.51\linewidth}
      \centering
@@
      &
-   \begin{minipage}{0.49\linewidth}
+   \begin{minipage}{0.51\linewidth}
      \centering
@@
    \item[(6)]   Include all additional controls in (3)-(5) with the sample that excludes New York as in (2). 
-   \item[(7)] Add weekly dummies fo  to the baseline specification.
+   \item[(7)] Add weekly dummies  to the baseline specification.
    \item[(8)] Baseline specification is estimated by instrumenting $\Delta \log T_{it}$ with one week lagged  logarithm  value of the number of tests per 1000 people.
@@
 estimates imply that mandating masks on April 1st would have led to
-500 fewer cases and 250 fewer deaths in Washington by the start of
+500 fewer cases and 250 fewer deaths in Washington in total by the start of
 June.
@@
 
-Figure \ref{fig:dailycases} shows the evolution of new cases in each
-of these three datasets. In all cases, daily changes in cumulative
+%Figure \ref{fig:dailycases} shows the evolution of new cases in each
+%of these three datasets. In all cases, daily changes in cumulative
+%cases displays some excessive volatility. This is likely due to delays
+%and bunching in testing and reporting of results. 
+The upper left panel of Figure \ref{fig:casedeathtest} shows the evolution of new cases in the NYT data, where daily changes in cumulative
 cases displays some excessive volatility. This is likely due to delays
-and bunching in testing and reporting of results. Table
+and bunching in testing and reporting of results. 
+Table
 \ref{tab:newcasecor} shows the variance of log new cases in each data
@@
 
-\begin{figure}[!ht]\caption{Daily cases \label{fig:dailycases}}
+\begin{figure}[!ht]\caption{Daily Cases, Weekly Cases,  Weekly Deaths from NYT Data, and Weekly Tests from JHU\label{fig:casedeathtest}}\smallskip
   \centering
@@
     \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/newcases}
-    \footnotetext{Each line shows daily new cases in a state.}
-  \end{minipage}
+    \includegraphics[width=0.8\textwidth,height=0.7\textwidth]{tables_and_figures/casesdeaths_q}
+%      \footnotetext{Thin gray lines are case or death growth in each
+%      state and date. Thicker colored lines are quantiles of case or
+%      death growth conditional on date.}
+  \end{minipage}
+     \begin{flushleft}
+      \footnotesize Thin gray lines are the log of cases, death, and tests  in each
+      state and date. Thicker colored lines are their quantiles conditional on date.
+      \end{flushleft}
 \end{figure}
+
+%\begin{figure}[!ht]\caption{Daily cases \label{fig:dailycases}}
+%  \centering
+%  \begin{minipage}{\textwidth}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{tables_and_figures/newcases}
+%    \footnotetext{Each line shows daily new cases in a state.}
+%  \end{minipage}
+%\end{figure}
 
@@
 Table \ref{tab:weekcasecor} reports the correlation and variance of
-weekly log new cases across the three data sets. Figure
-\ref{fig:weekcases} shows the evolution of weekly new cases in each
+weekly log new cases across the three data sets. The upper right panel of Figure
+\ref{fig:casedeathtest} shows the evolution of weekly new cases in each
+state over time  from the NYT data. The upper panel of Figure \ref{fig:growthq} shows the evolution of the weekly growth rate of new cases in each
 state over time.
@@
 
-\begin{figure}[!ht]\caption{Weekly Cases \label{fig:weekcases}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekcases}
-    \footnotetext{Each line shows weekly new cases in a state.}
-  \end{minipage}
-\end{figure}
+%\begin{figure}[!ht]\caption{Weekly Cases \label{fig:weekcases}}
+%  \centering
+%  \begin{minipage}{\textwidth}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekcases}
+%    \footnotetext{Each line shows weekly new cases in a state.}
+%  \end{minipage}
+%\end{figure}
 
@@
 Table \ref{tab:weekdeathcor} reports the correlation and variance of
-weekly deaths in the three data sets. Figure \ref{fig:weekdeaths}
-shows the evolution of weekly deaths in each state. As with cases, we
-use death data from The New York Times in our main results.
+weekly deaths in the three data sets. The lower left panel of Figure \ref{fig:casedeathtest}
+shows the evolution of weekly deaths in each state from the NYT data. As with cases, we
+use death data from The New York Times in our main results. The lower panel of Figure \ref{fig:growthq} shows the evolution of the weekly growth rate of new deaths in each
+state over time.
 
@@
 
-\begin{figure}[!ht]\caption{Weekly Deaths \label{fig:weekdeaths}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekdeaths}
-    \footnotetext{Each line shows weekly deaths in a state.}
-  \end{minipage}
-\end{figure}
+%\begin{figure}[!ht]\caption{Weekly Deaths \label{fig:weekdeaths}}
+%  \centering
+%  \begin{minipage}{\textwidth}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{tables_and_figures/weekdeaths}
+%    \footnotetext{Each line shows weekly deaths in a state.}
+%  \end{minipage}
+%\end{figure}
 
@@
       \includegraphics[width=0.75\textwidth]{tables_and_figures/deathquantiles}
-      \footnotetext{Thin gray lines are case or death growth in each
+%      \footnotetext{Thin gray lines are case or death growth in each
+%      state and date. Thicker colored lines are quantiles of case or
+%      death growth conditional on date.}
+    \end{tabular}
+  \end{minipage}
+     \begin{flushleft}
+      \footnotesize Thin gray lines are case or death growth in each
       state and date. Thicker colored lines are quantiles of case or
-      death growth conditional on date.}
-    \end{tabular}
-  \end{minipage}
+      death growth conditional on date.      \end{flushleft}
 \end{figure}
@@
 
-Our test data comes from The Covid Tracking Project. Figure
-\ref{fig:test} shows the evolution of tests over time.
-
-\begin{figure}[!ht]\caption{Number of Tests \label{fig:test}}
-  \centering
-  \begin{minipage}{\textwidth}
-    \centering
-    \includegraphics[width=0.45\textwidth]{tables_and_figures/test}
-    \footnotetext{These figures use the ``total test results''
-      reported by The Covid Tracking Project. This is meant to reflect
-      the number of people tested
-      (\href{https://covidtracking.com/about-data/faq}{as opposed to
-        the number of specimens tested}).}
-  \end{minipage}
-\end{figure}
+Our test data comes from The Covid Tracking Project. The lower right panel of Figure
+\ref{fig:casedeathtest} shows the evolution of the weekly number of tests over time.
+
+%\begin{figure}[!ht]\caption{Number of Tests \label{fig:test}}
+%  \centering
+%  \begin{minipage}{\textwidth}
+%    \centering
+%    \includegraphics[width=0.45\textwidth]{tables_and_figures/test}
+%    \footnotetext{These figures use the ``total test results''
+%      reported by The Covid Tracking Project. This is meant to reflect
+%      the number of people tested
+%      (\href{https://covidtracking.com/about-data/faq}{as opposed to
+%        the number of specimens tested}).}
+%  \end{minipage}
+%\end{figure}
 
@@
     \end{tabular}
-    \footnotetext{This figure shows the evolution of Google Mobility
+%    \footnotetext{This figure shows the evolution of Google Mobility
+%      Reports over time. Thin gray lines are the value of the
+%      variables in each state and date. Thicker colored lines are
+%      quantiles of the variables conditional on date.}
+  \end{minipage}
+     \begin{flushleft}
+      \footnotesize This figure shows the evolution of Google Mobility
       Reports over time. Thin gray lines are the value of the
       variables in each state and date. Thicker colored lines are
-      quantiles of the variables conditional on date.}
-  \end{minipage}
+      quantiles of the variables conditional on date.     \end{flushleft}
 \end{figure}
@@
 \cite{raifman2020}. %This database records that each state implemented 25 Covid-19 related policies.
-As discussed in the main text, our analysis focuses on seven policies. For stay-at-home orders, closed nonessential
+As discussed in the main text, our analysis focuses on six policies. For stay-at-home orders, closed nonessential
 businesses, closed K-12 schools, closed restaurants except takeout,
@@
 \cite{raifman2020}'s database with our modifications described above.
+
+Figures \ref{fig:growthpolicies1}-\ref{fig:growthpolicies2} show the  average case and death growth conditional on date and whether each of six policies  is implemented or not.
 
@@
     \end{tabular}
-    \footnotetext{ In these figures, red points are the case or death
+%    \footnotetext{ In these figures, red points are the case or death
+%      growth rate in states without each policy 14 (or 21 for deaths)
+%      days earlier. Blue points are states with each policy 14 (or 21
+%      for deaths) days earlier. The red line is the average across
+%      states without each policy. The blue line is the average across
+%      states with each policy.}
+  \end{minipage}
+     \begin{flushleft}
+      \footnotesize In these figures, red points are the case or death
       growth rate in states without each policy 14 (or 21 for deaths)
@@
       states without each policy. The blue line is the average across
-      states with each policy.}
-  \end{minipage}
-\end{figure}
+      states with each policy.    \end{flushleft}\end{figure}
 
@@
     \end{tabular}
-    \footnotetext{ In these figures, red points are the case or death
+%    \footnotetext{ In these figures, red points are the case or death
+%      growth rate in states without each policy 14 (or 21 for deaths)
+%      days earlier.  Blue points are states with each policy14 (or 21
+%      for deaths) days earlier.  The red line is the average across
+%      states without each policy. The blue line is the average across
+%      states with each policy.}
+  \end{minipage}
+     \begin{flushleft}
+      \footnotesize  In these figures, red points are the case or death
       growth rate in states without each policy 14 (or 21 for deaths)
@@
       states without each policy. The blue line is the average across
-      states with each policy.}
-  \end{minipage}
+      states with each policy.    \end{flushleft}
 \end{figure}
@@
 \subsection{Direct and Indirect Policy Effects with national case/death variables}
-Tables \ref{tab:dieff} and \ref{tab:dieff-con}  present  the estimates for the specification with past national case/death variables.
+Tables \ref{tab:dieff} and \ref{tab:dieff-con}  present  the estimates of direct and indirect effects of policies for the specification with past national case/death variables.
 The effects of school closures and  the sum of policies are estimated substantially smaller in Table \ref{tab:dieff} when national case/death variables are included than in Table \ref{tab:dieff-si}.
@@
 
+\subsection{Double Machine Learning}
+
+\subsection{Debiased Fixed Effects Estimator}
+
 \end{document}

### 2020-08-22 cases_and_policies/R/exploratory.R

@@ -17,7 +17,33 @@ colors <-  scale_color_solarized
 
 cs <- covidstates
 cs <- cs[cs$fips<60, ] # drop territories (Puerto Rico, Guam, etc)
+# cs <- cs[cs$date<=as.Date("2020-06-03"), ] # drop territories (Puerto Rico, Guam, etc)
 cs <- cs[order(cs$state, cs$date),]
+ 
+quantileplot <- function(v, p=c(0.05, 0.25, 0.5, 0.75, 0.95), data=cs) {
+  data$y <- data[,v]
+  adf <- aggregate(y ~ date, data=data,
+                   FUN=function(x) quantile(x, probs=p, na.rm=TRUE))
+  p <- as.factor(p)
+  s <- 1.2
+  txt <- data.frame(date=max(adf$date), y=adf$y[nrow(adf$y),], p=p)
+  names(adf)[1] <- "date"
+  figm <- ggplot(adf) +
+    geom_line(aes(x=date, y=adf$y[,1], color=p[1]), size=s) +
+    geom_line(aes(x=date, y=adf$y[,2], color=p[2]), size=s) +
+    geom_line(aes(x=date, y=adf$y[,3], color=p[3]), size=s) +
+    geom_line(aes(x=date, y=adf$y[,4], color=p[4]), size=s) +
+    geom_line(aes(x=date, y=adf$y[,5], color=p[5]), size=s) +
+    geom_text(data=txt,aes(x=date,y=y,color=p, label=p,
+                           hjust="left",vjust="center", size=2)) +
+    plottheme +
+    colors() + 
+    #ggtitle(TeX(paste("Evolution of", varlabels[varlabels$Variable==v,"Label"]))) +
+    geom_line(data=data, aes(x=date, y=y, group=state), alpha=0.2) +
+    theme(legend.position="none") +
+     labs(color='Quantile')  #+
+    # xlim(startday, max(data$date))
+}
 
 ################################################################################
 ## Case Measures
@@ -58,6 +84,16 @@ fjhu <- ggplot(cs, aes(x=date, y=newcases.jhu, colour=state)) + geom_line(alpha=
 pdf(paste(outdir,"newcases.pdf", sep="/"))
 grid.arrange(fnyt,fjhu,fctp)
 dev.off()
+
[…]

### 2020-08-22 tex/reply-to-referees.tex
update

@@
 ''}
+
+The testable restrictions of the model are now described better.  Please see the discussion around TR. The SIR model
+for example implies that testing rate is a confounder for case growth rate, but the causal model implies  that testing rate
+does not affect the behavior or policy variables (other than through cases). This is an overidentification restriction.
+To show the power of this approach, we also looked at imposition of other "contextual information": for example, mask
+having only direct effect on growth rates (and not via behavior) or business closure policies affecting case growth rates
+only through behavhior.
+
+The fixed effect estimates lump state-level confounders in one unobservable component.  Our approach does not preclude
+the use of fixed effects to estimate the model. We present fixed effects as part of Sensitivity analysis.
+
+The invariance assumption is similar to standard assumptions in Structural VAR literature, starting with Sims (1972).
 
@@
 \item[3.]  \textit{(Expectation and Lucas critique) ``Related to the previous comment, what would be the link
-between the current paper‚Äôs setup and how expectations are formed given the information
+between the current paper's setup and how expectations are formed given the information
 structure? For example, is this paper safe from Lucas critique? Some comments would be
@@
 $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-observed in the state we are simulating. We hold all other regressors
-at their observed values. Error terms are drawn with replacement from
-the residuals. We do this many times and report the average over draws
-of the residuals. The shaded region is a point-wise 90\% confidence
+observed in the state we are simulating. We now provide
+more details on how counterfactuals are constructed -- see Section 2.
+The shaded region is a point-wise 90\% confidence
 interval.
@@
 
-All programs and dataset will be available at our GitHub page, which is currently set to be private but we will plan to make it publicly accessible once the paper is accepted.
+All programs and dataset will be available at our GitHub page, which is currently set to be private but we will plan to make it publicly accessible once the final version of the paper is accepted.
 
@@
 
-[Not sure of what to do here]
+We are not doing any hypothesis testing in this paper: we are constructing point and interval predictions for 2-3 key policies. This
+is in line with with IHME institute is doing (our analysis on masks predates theirs, as we posted our findings
+at the end of May, and they announced their findings in middle of July). The 
+"stars" are reported as a matter of statistical convention, but we never use them in any formal way.
+
+Furthermore, we believe that testing "no mask effect" hypothesis is actually not an interesting exercise for policy analysis
+or actual policy decisions.   From a decision-theoretic point of view the cost of accepting the false null is extremely high,
+which could result in huge losses. The Neyman-Pearson test with $\alpha=$ small, corresponds to the idea that the cost
+of the null is small (little harm done if the "no mask effect" is falsely accepted), e.g.
+Amemiya "Introduction to Statistics and Econometrics".
+So doing "no mask effect" testing with very low $\alpha$ would correspond to assuming that the cost of deciding that masks are not effective is very low.
+Multiple hypothesis testing would further drive $\alpha$ to smaller values, and we are not sure how this would align with the decision-theoretic framework.
+
+%Remark. Elaborating further on decision-theoretic point, with $H_0$ masks being useless, the policy makers problem is to make a decision 
+%to achieve minimal loss:
+%$$
+%\min (E [Loss (NoMask) | H_0] P[H_0], E [Loss (Mask) | H_A] P[H_A]).
+%$$
+%The optimal decision is not to use Masks if $$P[H_0]> \alpha = P[H_A] \cdot E [Loss (Mask) | H_A] /E [Loss (NoMask) | H_0] ),$$
+%which gives the relevant $\alpha$.
+E.g, using $\alpha =.1$ results from the information that $P[H_A]=1/2$ and assuming that the loss from using masks has to be 5 times higher than the loss from not using masks.
+
+ For masks, given the decision-theoretic framework, given prior evidence from RCT on efficacy of masks for respiratory deceases, (unfortunately not including corona), the interesting nulls would be "mask effects  have -.15, -.1, -.05" effects on growth rate.
+This null would not be  rejected no matter what adjustments to p-values we make or what sensitivity analysis we make.
+Moreover, after doing all of the sensitivity exercises, we would not be able to reject this null.
+This corresponds to the idea that the null hypothesis
+ here is the one that is very costly if falsely rejected.
 
@@
 The father and son,
-P. Wright and S. Wright closely collaborated to develop structural equation models and causal path diagrams; P. Wright'Äôs
+P. Wright and S. Wright closely collaborated to develop structural equation models and causal path diagrams; P. Wright's
 key work represented supply-demand system as a directed acyclical graph and established its identification using
@@
 \item[4.]  \textit{(Footnote 19) ``On page 17, in footnote 19, the authors state:
-We drop ``Residential‚Ä'' because it is highly correlated with both ‚ÄúWorkplaces‚Äù and
-``Retail \& recreation‚Ä'' at correlation coefficients of 0.98.
+We drop ``Residential' because it is highly correlated with both ``Workplace'' and
+``Retail \& recreation" at correlation coefficients of 0.98.
 This sentence is ambiguous. Do the authors mean that both correlation coefficients are identically 0.98? Please rephrase this sentence.''}

### 2020-08-24 cases_and_policies/rmd/generatetables.R
Committing changes before mergning master.

@@ -140,7 +140,7 @@ mainregressions <- function(df, # data
               c(sprintf("lag(%s, %d)", infovars[[ij[2]]], L),
                 tvars,
                 xlist[[ij[2]]]),
-              interactions[[ij[2]]], iv[[ij[2]]], L=L)
+              ia, iv[[ij[2]]], L=L)
   })
 
   ijs <- expand.grid(1:length(bvars), 1:length(plist))

### 2020-08-24 cases_and_policies/rmd/regprep.R

@@ -1,8 +1,8 @@
 cs$t <- as.numeric(cs$date)
 
-warning("Using data only up to 2020-06-02 as in submitted version of paper. Modify lines 4-5 of regprep.R to change")
+warning("Using data only up to 2020-06-30. Modify lines 4-5 of regprep.R to change")
 df <- pdata.frame(subset(cs, cs$date>=as.Date("2020-01-01") &
-                                   cs$date<=as.Date("2020-06-03")),
+                                   cs$date<=as.Date("2020-06-30")),
                   # cs$date<=as.Date("2020-06-30")),
                   index=c("state","t"), stringsAsFactors=FALSE)
 
@@ -89,7 +89,7 @@ statevars <- c("log(Population.2018)",
                "Percent.at.risk.for.serious.illness.due.to.COVID","party")#, "logvote")
 bvars <- c("workplaces","retail","grocery","transit")
 
-for(i in 1:length(bvars)) df[,bvars[i]] <- df[,bvars[i]]/100 
+for(i in 1:length(bvars)) df[,bvars[i]] <- df[,bvars[i]]/100
 
 
 #df$udistance <- panelma(df$daily_distance_diff,df$state, df$date, 6)
@@ -139,5 +139,5 @@ month <- month(df$date)
 df$month = as.factor(month)
 df$pmask.may <- df$pmaskbus*tilt
 df$pmask.april <- df$pmaskbus*(1-tilt)
-pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential") 
-#pols <- c("pmaskbus","pk12","pindex") 
+pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential")
+#pols <- c("pmaskbus","pk12","pindex")

### 2020-08-26 tex/covidEffectOfNPI-revision.tex
update

@@
 It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
+Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.
+\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+See, e.g., \cite{pearl:biometrika},\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak}, \cite{peters2020book}, \cite{heckman:pinto}, and \cite{hernanrobins2020book} for modern developments, in computer science, econometrics, and epidemiology.
+The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986},\cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
+The father and son, P.
+Wright and S.
+Wright closely collaborated to develop structural equation models and causal path diagrams; P.
+Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables.}
+
+%We present a conceptual framework that spells out the causal structure on how the Covid-19 spread  is dynamically determined by policies and human behavior.
+Our approach explicitly recognizes that policies not only directly affect the spread of Covid-19 (e.g., mask requirement) but also indirectly affect its spread by changing people's behavior (e.g., stay-at-home order).
+It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
 Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, tinbergen1940, wold1954} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
 See \cite{pearl:biometrika},\cite{pearl:robins},  \cite{pearl:causality}, \cite{white:chalak}, \cite{peters2020book}, \cite{heckman:pinto}, and \cite{hernanrobins2020book}  for modern developments, in computer
-science, econometrics, and epidemiology.
+%science, econometrics, and epidemiology.
 The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
@@
 
- \cite{pei2020} use county-level observations of reported infections and deaths in conjunction with mobility  data from SafeGraph to estimate how effective reproductive numbers in major metropolitan areas change over time.
-They conduct simulation of implementing all policies 1-2 weeks earlier and found that it would have resulted in reducing the number of cases and deaths by more than half.
-However, their study does not  explicitly analyze how policies are related to
-the effective reproduction numbers. % Our study  analyzes how policies  affects the reproduction numbers through their impact on people's behavior.
+ \cite{pei2020} use county-level observations of reported infections and deaths in conjunction with mobility  data from SafeGraph to conduct simulation of implementing all policies 1-2 weeks earlier and found that it would have resulted in reducing the number of cases and deaths by more than half.
+However, their study does not  explicitly analyze how policies are related to the effective reproduction numbers.
+% Our study  analyzes how policies  affects the reproduction numbers through their impact on people's behavior.
 
@@
 Also, a substantial fraction of individual who are infected become infectious before showing symptom onset.}
-\cite{howard2020} also review available medical evidence and conclude that ``mask wearing reduces the transmissibility per contact by reducing transmission of infected droplets in both laboratory and clinical contexts.''  The laboratory findings in \cite{hou2020} suggest that the nasal cavity may be the initial site of infection followed by aspiration to the lung, supporting the argument  ``for the widespread use of masks to prevent aersol, large droplet, and/or mechanical exposure to the nasal passages.''
-
-Given the lack of experimental evidence on the effect of masks, conducting observational studies  is useful and important.
+\cite{howard2020} also review available medical evidence and conclude that ``mask wearing reduces the transmissibility per contact by reducing transmission of infected droplets in both laboratory and clinical contexts.''  The laboratory findings in \cite{hou2020} suggest that the nasal cavity may be the initial site of infection followed by aspiration to the lung, supporting the argument  ``for the widespread use of masks to prevent aersol, large droplet, and/or mechanical exposure to the nasal passages.''   \cite{He2020} examined temporal patterns of viral shedding in COVID-19 patients and found the highest viral load at the time of symptom onset; this suggests that a significant portion of transmission may have occurred before symptom onset and that universal face masks may be an effective control measure to reduce transmission.\footnote{\cite{Lee2020} find evidence that viral loads in asymptomatic patients are similar to those in symptomatic patients.
+Aerosol transmission of viruses may occur through aerosols particles released during breathing and speaking by asymptomatic infected individuals; masks reduce such airborne transmission \citep{Prather2020}.
+\cite{Anfinrud2020} provide visual evidence of speech-generated droplet as well as the effectiveness of cloth masks to reduce the emission of droplets.
+\cite{Chu2020} conduct a meta-analysis of observational studies on transmission of the viruses that cause COVID-19 and related diseases and find the effectiveness of mask use for reducing transmission.
+} \cite{ollila2020} provide a meta-analyhsis of randomized controlled trials of non-surgical face masks in preventing viral respiratory infections in non-hospital and non-household settings, finding that face masks decreased infections across all five studies they reviewed.\footnote{Whether wearing masks creates a false sense of security and leads to decrease in social distancing is also a hotly debated topic.
+A randomized field experiment in Berlin, Germany, conducted by \cite{seres2020}  finds that wearing masks actually increases social distancing, providing no evidence that mandatory masks leads to decrease in social distancing.}
+
+Given the lack of experimental evidence on the effect of masks in the context of COVID-19, conducting observational studies  is useful and important.
 To the best of our knowledge, our paper is the first  empirical study that shows the effectiveness of  mask mandates on reducing the spread of Covid-19 by analyzing the US state-level data.
@@
 Analyzing mitigation measures in New York, Wuhan, and Italy, \cite{zhangr2020} conclude that mandatory face coverings substantially reduced infections.
-\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.
-Our finding
-is also corroborated by a completely different causal methodology based on synthetic control 
+\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} Our finding is also corroborated by a completely different causal methodology based on synthetic control
 using German data in \cite{Mitze2020}.\footnote{Our study was first released in ArXiv on May 28, 2020 whereas
@@
 orders.  The effect of masks for employees is  
-small. 
+small.\footnote{Similar to our finding, \cite{kovacs2020} find no evidence that introduction of compulsory face mask policies affect community mobility in Germany.}
+
 %The comparison of effects across policies should be interpreted
@@
   \caption{\label{tab:dieff-con-si}Direct and Indirect Policy Effects
-    without national case/death variables, \textbf{constraints imposed}}
+    without national case/death variables, \textbf{restrictions imposed}}
 \begin{minipage}{\linewidth}
@@
 }
+
+\afterpage{
+\begin{table}
+  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
+\begin{minipage}{\linewidth}
+  \centering
+    \scriptsize
+  \begin{tabular}{c}
+    \textbf{Cases}
+    \\
+    \input{tables_and_figures/dieff-cases-NI}
+    \\ \\
+    \textbf{Deaths}
+    \\
+    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
+    \\
+  \end{tabular}
+  \smallskip
+  \begin{flushleft}
+      \scriptsize Direct effects capture the effect of policy on case
+      growth holding behavior, information, and confounders
+      constant. Direct effects are given by ${\pcolor \pi}$ in
+      equation (\ref{eq:R1}). Indirect effects capture how policy
+      changes behavior and behavior shift case growth. They are given
+      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
+      from (\ref{eq:R2}). The total effect is
+      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
+      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
+      \ref{eq:R4}. Columns ``Difference''    are  the differences between
+      the estimates from (\ref{eq:R4}) and the combination of
+      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
+      Standard errors are computed by
+      bootstrap and clustered on state.
+    %  \ref{tab:PtoY}.
+    \end{flushleft}
+  \end{minipage}
+\end{table}
+\clearpage}
+
+\afterpage{
+\begin{table}
+  \caption{\label{tab:dieff-con}Direct and Indirect Policy Effects with national case/death variables, \textbf{restrictions imposed}}
+\begin{minipage}{\linewidth}
+  \centering
+    \scriptsize
+  \begin{tabular}{c}
+    \textbf{Cases}
+    \\
+    \input{tables_and_figures/dieff-cases-constrain-NI}
+    \\ \\
+    \textbf{Deaths}
+    \\
+    \input{tables_and_figures/dieff-deaths-deathsinfo-constrain-NI}
+    \\
+  \end{tabular}
+  \smallskip
+  \begin{flushleft}
+      \scriptsize Direct effects capture the effect of policy on case
+      growth holding behavior, information, and confounders
+      constant. Direct effects are given by ${\pcolor \pi}$ in
+      equation (\ref{eq:R1}). Indirect effects capture how policy
+      changes behavior and behavior shift case growth. They are given
+      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
+      from (\ref{eq:R2}). The total effect is
+      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
+      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
+      \ref{eq:R4}. Columns ``Difference''    are  the differences between
+      the estimates from (\ref{eq:R4}) and the combination of
+      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
+      Standard errors are computed by
+      bootstrap and clustered on state.
+    %  \ref{tab:PtoY}.
+    \end{flushleft}
+  \end{minipage}
+\end{table}
+\clearpage}
 
@@
 %\begin{table}
-%  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables, \textbf{Constraints imposed}}
+%  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables, \textbf{restrictions imposed}}
 %\begin{minipage}{\linewidth}
@@
 
-Similarly to Table  \ref{tab:dieff-si}, the first three columns of  Table \ref{tab:dieff-con-si} report  the estimated direct and indirect effects of policy but imposing that masks for employees only affect cases/deaths directly without affecting behavior and that business closure policies only affects cases/deaths indirectly through its effect on behavior.\footnote{These are computed from estimating the specification in columns (1)-(4) of
+Similarly to Table  \ref{tab:dieff-si}, the first three columns of  Table \ref{tab:dieff-con-si} report  the estimated direct and indirect effects of policy but imposing that masks for employees only affect cases/deaths directly without affecting behavior and that business closure policies only affect  cases/deaths indirectly through their effects on behavior.\footnote{These are computed from estimating the specification in columns (1)-(4) of
 Table \ref{tab:PItoB} and column (1) of Table \ref{tab:BPItoY} but imposing that the coefficient of masks for employees is zero in (\ref{eq:R1}) and that the coefficient of business closure policies is zero in (\ref{eq:R2}).} The estimated total effect of masks for employees in the third column of Table \ref{tab:dieff-con-si} is higher than that of Table  \ref{tab:dieff-si}.
@@
 The large differences in school closures  may be due to the  difficulty in identifying the effect of school closures given a lack of cross-sectional variations.
-Imposing the coefficients of masks for employees and business closure policies to be zero in (\ref{eq:R1}) and  (\ref{eq:R2}), respectively, does not increase differences between  ``Total" and  ``PI$\to$Y Coefficient''  as reported in the last column of  Table \ref{tab:dieff-con-si}; if anything the differences in mask coefficients become smaller in Table \ref{tab:dieff-con-si}, suggesting a validity of zero coefficients imposed in estimating the total effect.
+Imposing the coefficients of masks for employees and business closure policies to be zero in (\ref{eq:R1}) and  (\ref{eq:R2}), respectively, does not increase differences between  ``Total" and  ``PI$\to$Y Coefficient''  as reported in the last column of  Table \ref{tab:dieff-con-si}.
+
+ Tables  \ref{tab:dieff} and \ref{tab:dieff-con}  present the results for a specification that includes national information variables, where the estimates on masks for employees are similar to those in  Tables  \ref{tab:dieff-si} and \ref{tab:dieff-con-si}.
+Column ``Difference''  of Table \ref{tab:dieff-con} indicates that the restrictions of zero coefficients for masks for employees in (\ref{eq:R1}) and for business closure policies in  (\ref{eq:R2}) are not statistically rejected.
 
@@
 ``Total'' and ``PI$\to$Y Coefficient'' columns.  The average is an appealing and simple way
-to combine the two estimates of the total effect: one relying on the causal structure and another inferred from a direct estimation of equation (PI $\to$ Y).\footnote{Averaging the two estimates theoretically reduces noise, albeit in our case the reductions are small.
-Another approach would be to use precision averaging, which would give similar result. Finally, another approach would be to use generalized method of moments to estimate
-all of the equations jointly.
-We don't pursue this approach since it is likely to be non-robust under local deviations from correct specification; simple model averaging is more appealing in this case.}  We shall be using the average estimate in generating the counterfactuals in the next section.
+to combine the two estimates of the total effect: one relying on the causal structure and another inferred from a direct estimation of equation (PI $\to$ Y).
+%\footnote{Averaging the two estimates theoretically reduces noise, albeit in our case the reductions are small.
+%Another approach would be to use precision averaging, which would give similar result. Finally, another approach would be to use generalized method of moments to estimate
+%all of the equations jointly.
+We don't pursue this approach since it is likely to be non-robust under local deviations from correct specification; simple model averaging is more appealing in this case.}
+We shall be using the average estimate in generating the counterfactuals in the next section.
 Turning to the results, the estimates of Tables \ref{tab:dieff-si} and \ref{tab:dieff}  imply that all policies combined would reduce
@@
 The left panel of ``(i) baseline timing'' assume that the times from exposure to case confirmation and death reporting are 14 and 21 days, respectively,  while they are 7 and 24 days, respectively, in the right panel of ``(ii) alternative timing.''\footnote{This alternative timing assumption is motivated by the lower bound estimate of median times from exposure to case confirmation or death reporting in Table 2 of \url{https://www.cdc.gov/coronavirus/2019-ncov/hcp/planning-scenarios.html}, which are based on data received by CDC through June 29, 2020.
-We thank Jessica Metclaf for recommending us to do a sensitivity analysis on the timing assumption while suggesting this reference to us.}
+We thank C.
+Jessica E.
+Metcalf for recommending us to do a sensitivity analysis on the timing assumption while suggesting this reference to us.}
 
@@
 
-The estimated coefficients of masks for employees largely confirms our finding that mandatory mask policy reduces the case and death growths.
+The estimated coefficients of masks for employees largely confirm our finding that mandatory mask policy reduces the case and death growths.
 The estimated coefficients of stay-at-home orders and business closures are negative in columns (1) and (2) but their magnitudes as well as statistical significance are somewhat sensitive to whether bias corrections are applied or not.
@@
 The results in Table \ref{tab:PtoY-fe} should be interpreted with caution.
-The fixed effects approach may not be preferred to random effects approach here because the former relies on long time and cross-sectional histories but, in our data, the effective time-dimension is short and a number of states is not large.
+The fixed effects approach may not be preferred to random effects approach here because the former relies on long time and cross-sectional histories but, in our data, the effective time-dimension is short and the number of states is not large.
 Furthermore, the fixed effects approach could suffer more from  measurement errors, which could be a concern for our behavior and policy variables.
@@
 
-\afterpage{
-\begin{table}
-  \caption{\label{tab:dieff}Direct and Indirect Policy Effects with national case/death variables}
-\begin{minipage}{\linewidth}
-  \centering
-    \scriptsize
-  \begin{tabular}{c}
-    \textbf{Cases}
-    \\
-    \input{tables_and_figures/dieff-cases-NI}
-    \\ \\
-    \textbf{Deaths}
-    \\
-    \input{tables_and_figures/dieff-deaths-deathsinfo-NI}
-    \\
-  \end{tabular}
-  \smallskip
-  \begin{flushleft}
-      \scriptsize Direct effects capture the effect of policy on case
-      growth holding behavior, information, and confounders
-      constant. Direct effects are given by ${\pcolor \pi}$ in
-      equation (\ref{eq:R1}). Indirect effects capture how policy
-      changes behavior and behavior shift case growth. They are given
-      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
-      from (\ref{eq:R2}). The total effect is
-      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
-      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
-      \ref{eq:R4}. Columns ``Difference''    are  the differences between
-      the estimates from (\ref{eq:R4}) and the combination of
-      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
-      Standard errors are computed by
-      bootstrap and clustered on state.
-    %  \ref{tab:PtoY}.
-    \end{flushleft}
-  \end{minipage}
-\end{table}
-\clearpage}
-
-\afterpage{
-\begin{table}
-  \caption{\label{tab:dieff-con}Direct and Indirect Policy Effects with national case/death variables, \textbf{constraints imposed}}
-\begin{minipage}{\linewidth}
-  \centering
-    \scriptsize
-  \begin{tabular}{c}
-    \textbf{Cases}
-    \\
-    \input{tables_and_figures/dieff-cases-constrain-NI}
-    \\ \\
-    \textbf{Deaths}
-    \\
-    \input{tables_and_figures/dieff-deaths-deathsinfo-constrain-NI}
-    \\
-  \end{tabular}
-  \smallskip
-  \begin{flushleft}
-      \scriptsize Direct effects capture the effect of policy on case
-      growth holding behavior, information, and confounders
-      constant. Direct effects are given by ${\pcolor \pi}$ in
-      equation (\ref{eq:R1}). Indirect effects capture how policy
-      changes behavior and behavior shift case growth. They are given
-      by ${\bcolor \alpha}$ from (\ref{eq:R1}) times ${\pcolor \beta}$
-      from (\ref{eq:R2}). The total effect is
-      ${\pcolor \pi} + {\pcolor \beta} {\bcolor \alpha}$. Column
-      ``PI$\to$Y Coefficients'' shows the coefficient estimates from
-      \ref{eq:R4}. Columns ``Difference''    are  the differences between
-      the estimates from (\ref{eq:R4}) and the combination of
-      (\ref{eq:R1}) and (\ref{eq:R2}) while column ``Average'' are their averages.
-      Standard errors are computed by
-      bootstrap and clustered on state.
-    %  \ref{tab:PtoY}.
-    \end{flushleft}
-  \end{minipage}
-\end{table}
-\clearpage}
-
 \subsection{Double Machine Learning}
 
+To estimate the coefficient of the $j$-th policy variable, $P_{it}^j$, using Double Machine Learning of \cite{chernozhukov18}, we consider a version of (\ref{eq:R4}) as follows:
+\begin{align*}
+Y_{i,t+\ell} & = P_{it}^j \theta_0 + \phi' X_{it} + g_0(W_{it}) + \xi_{it}, && E[\xi_{it}|P_{it}^j,X_{it},W_{it}]=0,\\
+P_{it}^j &= \psi' X_{it} +m_0(W_{it}) +V_{it},&& E[V_{it}|X_{it},W_{it}]=0,
+\end{align*}
+where $X_{it}=((P_{it}^{-j})',B_{it}',I_{it}')'$ collects  policy variables except for the $j$-th policy variables $P_{it}^{-j}$, behavior variables $B_{it}$, and information variables $I_{it}$.
+The confounding factors $W_{it}$ affect the policy variable $P_{it}^j$ via the function $m_0(W)$ and the outcome variable $Y_{i,t+\ell}$ via the function $g_0(W)$.
+We apply Lasso  or  Random Forests to estimate $g_0(W)$  for dimension reductions or for capturing non-linearity while the coefficients of $X_{it}$ are estimated under linearity without imposing any dimension reductions.
+We omit the details of estimation procedure here.
+Please see the discussion in Example 1.1 of  \cite{chernozhukov18} for reference.
+
+%We also assume that  $E[Y_{it}|X_{it},W_{it}]$  takes a partial linear form with respect to $X_{it}$ given by 
+%$E[Y_{it}|X_{it},W_{it}]= \varphi' X_{it} +q_0(W_{it})$. 
+%Then, we can rewrite the above partially linear model as
+%\begin{align*}
+%\tilde Y_{i,t+\ell} & = V_{it} \theta_0 +  \xi_{it},&& E[\xi_{it}|P_{it}^j,X_{it},W_{it}]=0,\\
+%\tilde Y_{i,t+\ell} & = Y_{i,t+\ell}-(\varphi' X_{it} +q_0(W_{it})),&&   E[Y_{i,t+\ell}|X_{it},W_{it}]=\varphi' X_{it}+q_0(W_{it}),\\
+%V_{it} & = P_{it}^j - (\psi' X_{it} + m_0(W_{it})) &&    E[ P_{it}^j |X_{it},W_{it}]=\psi' X_{it} +m_0(W_{it}) .
+%\end{align*} 
+%
+%We consider a 10-fold partition $({\bf S}_{k})_{k=1}^{10}$ of observations across states such that each fold $\bf S_k$ contains the  panel data set of  $n=5$ or $6$ states.
+For each $k=1,...,10$, define $\bf S_{-k}=\cup_{j\neq k} \bf S_{j}$.
+%\begin{enumerate}
+%\item For $k=1,...,10$,  iterate on:
+%\begin{enumerate}
+%\item Apply Lasso or Random Forests while imposing partial linear structure to the subsample $\bf S_{-k}$ to obtain an estimate   $\hat \varphi^{(k)}$, $\hat \psi^{(k)} $, $\hat q^{(k)} _0(W)$, and $\hat m^{(k)} _0(W)$.\footnote{In practice, we obtain an estimator $\hat \varphi^{(k)}$ and $\hat q^{(k)} _0(W)$ by the following iterative procedure.
+Let $L(\xi)$ denote a loss function (e.g., $L(\xi)=\xi^2$).
+%\begin{enumerate}
+%\item[(i)] Initialize $(\hat \varphi^0,\hat q^{0} _0(W))$.
+%\item[(ii)] For $j=1,2,...,J$ for $J=5$, iterate on:
+%\begin{enumerate} 
+%\item[(a)] Update $\hat \varphi^j = \arg\min_\varphi \sum_{(i,t)\in \bf S_{-k} } L( Y_{i,t+\ell} - \varphi-  \hat q^{j-1}_0(W_{it}))$.
+%\item[(b)] Update $\hat q^{j}_0(W) = \arg\min_{q_0} \sum_{(i,t)\in \bf S_{-k} } L(Y_{i,t+\ell}- \hat \varphi^j - q_0(W_{it}))$.
+%\end{enumerate} 
+%\item[(iii)] Set $\hat \varphi^{(k)}=\hat \varphi^J$ and $\hat q^{(k)} _0(W)=\hat q^{J}_0(W) $.
+%\end{enumerate}
+%We obtain $\hat \psi^{(k)} $ and $\hat m^{(k)} _0(W)$ by a similar iterative procedure.
+%} 
+%\item Using $\hat \varphi^{(k)}$, $\hat \psi^{(k)} $, $\hat q^{(k)} _0(W)$, and $\hat m^{(k)} _0(W)$, compute the predicted residuals for the observations in $\bf S_{k}$ as
+%\begin{align*}
+%\hat{\tilde Y}_{i,t+\ell}    = Y_{i,t+\ell}-((\hat \varphi^{(k)})' X_{it} +\hat q^{(k)}_0(W_{it})),\quad 
+%\hat V_{it}    = P_{it}^j - ((\hat \psi^{(k)})' X_{it} + \hat m^{(k)}_0(W_{it})).
+%\end{align*} 
+%\end{enumerate}
+%\item Given $\{(\hat{\tilde Y}_{i,t+\ell} ,\hat V_{it}) \}$ in (1), our DML estimator for $\theta_0$ is obtained by regressing $\hat{\tilde Y}_{i,t+\ell}$ on $\hat V_{it}$ with no intercept term.
+We compute the asymptotic standard errors for $\hat\theta_0$ clustered at the state-level using $\{(\hat{\tilde Y}_{i,t+\ell} ,\hat V_{it}): i = 1,...,N; t=1,...,T \}$.
+\end{enumerate}
+
 \subsection{Debiased Fixed Effects Estimator}
+
+We apply cross-over Jackknife bias correction as discussed in \cite{chen2020} in details. Here, we briefly describe our debiased fixed effects estimator.
+
+Given our panel data with $N=51$ states and $T=75$ days  for case growth equation (\ref{eq:R4}), consider two subpanels as follows:
+$$
+{\bf S}_1 =  \{(i,t) :  i \leq \lceil N/2 \rceil,  t \leq \lceil T/2 \rceil \} \cup  \{(i,t) : i  \geq \lfloor N/2 + 1 \rfloor, t  \geq \lfloor T/2 + 1\rfloor \}
+$$
+and
+$$
+{\bf S}_2 =  \{(i,t) :  i \leq \lceil N/2 \rceil, t  \geq \lfloor T/2 + 1\rfloor \} \cup  \{(i,t) : i  \geq \lfloor N/2 + 1 \rfloor, t  \leq \lceil  T/2 \rceil \}
+$$
+where $\lceil . \rceil$  and  $\lfloor . \rfloor$ are the ceiling and floor functions.  Each of these two subpanels includes observations for all cross-sectional units and time periods.
+
+We form the debiased fixed effects estimator as
+$$
+\widehat \beta_{\rm BC}  =  2 \widehat \beta -  \widetilde \beta_{{\bf S}_1\cup{\bf S}_2},
+$$
+where $\widehat \beta$ is the standard fixed effects estimator while $\widetilde \beta_{{\bf S}_1\cup{\bf S}_2}$ denotes the fixed effects estimator using the data set  ${\bf S}_1\cup{\bf S}_2$ but treats the states in  ${\bf S}_1$ differently from those in  ${\bf S}_2$ to form the fixed effects estimator; namely, we include approximately twice more state fixed effects to compute $\widetilde \beta_{{\bf S}_1\cup{\bf S}_2}$.\footnote{Alternatively, we may form the cross-over jackknife corrected estimator as $\widehat \beta_{\rm CBC}  =  2 \widehat \beta -  (\widetilde \beta_{{\bf S_1}}+\widetilde \beta_{{\bf S_2}})/2$, where $\widetilde \beta_{{\bf S}_j}$ denotes the fixed effect estimator using the subpanel  ${\bf S}_j$ for $j=1,2$.
+In our empirical analysis, these two cross-over jackknife bias corrected estimators give  similar result.} We obtain bootstrap standard errors by using multipler bootstrap with state-level clustering.
 

### 2020-08-26 cases_and_policies/rmd/cf-pindex.Rmd

@@
 
+```{r testsmooth, cache=FALSE}
+source("smoothtests.R")
+```
+
 ```{r statesim, cache=FALSE}
@@
 sdf <- subset(df, df$date>=datelims[1]) #+L)
-regs <- mainregressions(sdf,
+regs <- mainregressions2(sdf,
                         "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
@@
 cnames <- c(pols, infovars[[1]])
-di <- dieff_table(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
+di <- dieff_table2(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
                   regs$pbiy[[1]]$reg$coef[,1],
@@
 S <- 999
-bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
+bs <- bootstrap_felm2(sdf, c(pib,list(pbiy,piy)), S=S)
 bdi <- lapply(1:S, function(i)
-  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
+  dieff_table2(lapply(1:4, function(j) bs[[j]][,i]),
                bs[[5]][,i], bs[[6]][,i],
@@
 sdf <- subset(df, df$date>=datelims[1]) #+L)
-regs <- mainregressions(sdf,
+regs <- mainregressions2(sdf,
                         "dlogdd", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
@@
 S <- 999
-bs <- bootstrap_felm(sdf, c(pib,list(pbiy,piy)), S=S)
+bs <- bootstrap_felm2(sdf, c(pib,list(pbiy,piy)), S=S)
 bdi <- lapply(1:S, function(i)
-  dieff_table(lapply(1:4, function(j) bs[[j]][,i]),
+  dieff_table2(lapply(1:4, function(j) bs[[j]][,i]),
               bs[[5]][,i], bs[[6]][,i],

### 2020-08-26 cases_and_policies/rmd/counterfactuals.Rmd

@@
 # Data Preparation
+
 ```{r setup, cache=FALSE, warning=FALSE, include=FALSE}
@@
 library(reshape2)
+
+# stupid tidyverse stealing base package names ...
+if("dplyr" %in% (.packages())){
+  detach("package:dplyr", unload=TRUE)
+}
 
@@
 source(paste(rootdir,"cases_and_policies/rmd/counterfactuals.R", sep="/"))
+```
+
+```{r testsmooth, cache=FALSE}
+source("smoothtests.R")
 ```
@@
 ```
-
-# ```{r otherstates, cache=TRUE, fig.cap="Effect of Removing Policies"}
-# if (!knitr::is_latex_output()) {
-#   for( st in c("New York", "Florida", "Michigan", "Georgia", "Massachusetts")) {
-#     f <- cfplots(st, cfnop, data=df)
-#     grid.arrange(f$figd, f$figdp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-#     grid.arrange(f$figl, f$figp, nrow=2, top =  textGrob(st,gp=gpar(fontsize=20, font=2)))
-#   }
-# }
-# ```
-
-# ```{r nop-state-figures, cache=TRUE, warning=FALSE}
-# states <- c("Washington","Illinois","Massachusetts")
-# for(st in states) {
-#   f <- cfplots(st, cfnop, data=df)
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dgrowth_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figdp)
-#   dev.off()
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-growth_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figd + ylim(c(-1,5)))
-#   dev.off()
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-cases_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figl)
-#   dev.off()
-#
-#   figdev(sprintf("%s/tex/tables_and_figures/%s-nop-dcases_v1.pdf",rootdir,st), width=5,height=3.75)
-#   print(f$figp)
-#   dev.off()
-# }
-# ```
 

### 2020-08-26 cases_and_policies/rmd/datasummary.Rmd

@@
 library(kableExtra)
+library(latex2exp)
+library(plm)
+library(lubridate)
 
@@
 
-```{r gplots, cache=TRUE, warning=FALSE}
+```{r gplots, cache=FALSE, warning=FALSE}
 ## fign <-
@@
     #labs(color='Quantile')  +
-    xlim(startday, max(data$date))
+   # xlim(startday, max(data$date))
+  xlim(startday,as.Date("2020-06-03"))
 }

### 2020-08-26 tex/covidEffectOfNPI-revision.tex

--- 2020-08-22 tex/covidEffectOfNPI-revision.tex
@@
 
-We now turn our focus to dynamic feedback effects. Policy
-and behavior changes that reduce case and death growth today can lead to a more
-optimistic, riskier behavior in the future, attenuating longer run effects.
-We perform the main counterfactual experiments using the average of two estimated coefficients as reported in column ``Average'' of Table \ref{tab:dieff-si} under a specification that excludes the number of past national cases and deaths from information variables.
-In the appendix, we also report additional counterfactual experiment results with the specification that includes the national information variables, and find that they are very similar.
-The results on mask policies, business closures, stay-at-home orders are robust with respect to this variation (see Figures \ref{fig:US-mask}-\ref{fig:US-shelter} in the appendix).
-On the other hand, the results on removing all policies, particularly closure of schools, reported in the next section, are sensitive to the inclusion of national information variables, highlighting the large uncertainty regarding the size of the effect.
-In Figures \ref{fig:WA-mask}-\ref{fig:US-shelter} below, the top panel presents the result on cases while the bottom panel presents the result on deaths.
-
-%Additional counterfactual experiments include the
-%} %\footnote{\textbf{The results of counterfactual experiments are not sensitive to whether we use the estimate of  table \ref{tab:dieff-si} or that of table \ref{tab:dieff} except for a counterfactual experiment of removing all policies  in section \ref{removing-policies}.}}
+We now turn our focus to dynamic feedback effects. Policy and behavior
+changes that reduce case and death growth today can lead to a more
+optimistic, riskier behavior in the future, attenuating longer run
+effects.  We perform the main counterfactual experiments using the
+average of two estimated coefficients as reported in column
+``Average'' of Table \ref{tab:dieff-con-si} under a specification that
+excludes the number of past national cases and deaths from information
+variables and constrains masks to have only direct effects and
+business policies only indirect effects.\footnote{Results (not shown)
+  using the unconstrained estimates in Table \ref{tab:dieff-si} are similar.}
+%In the appendix, we also report additional counterfactual
+%experiment results with the specification that includes the national
+%information variables, and find that they are very similar.
+% The results on mask policies, business closures, stay-at-home orders
+% are robust with respect to this variation (see Figures
+% \ref{fig:US-mask}-\ref{fig:US-shelter} in the appendix). On the other
+% hand, the results on removing all policies, particularly closure of
+% schools, reported in the next section, are sensitive to the inclusion
+% of national information variables, highlighting the large uncertainty
+% regarding the size of the effect.  In Figures
+% \ref{fig:WA-mask}-\ref{fig:US-shelter} below, the top panel presents
+% the result on cases while the bottom panel presents the result on
+% deaths.
 
@@
     \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_v1}
-      \\
-      \\
-      \multicolumn{3}{c}{\textbf{Deaths}}
-      \\      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_v1}
+    \begin{tabular}{cc}
+      \multicolumn{2}{c}{\textbf{Cases}} \\
+      %\includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
+      % &
+        \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dgrowth_idx}
+      &        \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dcases_idx}
+      \\
+      \\
+      \multicolumn{2}{c}{\textbf{Deaths}}
+      \\
+      % \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths} &
+                                                                                                   \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_idx}
+      &      \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_idx}
     \end{tabular}
+
     \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff-si}. We set initial
+      \footnotesize
+      The vertical (often truncated) text denotes the
+      dates that Washington imposed each policy.  To compute the
+      estimated and counterfactual paths we use the average of two
+      estimated coefficients as reported in column ``Average'' of
+      Table \ref{tab:dieff-con-si}. We set initial
       $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
       observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
+      regressors at their observed values. The shaded region is a
       point-wise 90\% confidence interval.
@@
 For illustrative purpose, we begin by focusing on Washington State.
-The
-left column of  Figure \ref{fig:WA-mask} shows the observed, estimated
-average, and counterfactual average of $\Delta \log \Delta C$ (top panel) and
-$\Delta \log \Delta D$ (bottom panel). To compute the estimated and counterfactual
-paths, we use the   estimate   in
-column ``Average'' of Table \ref{tab:dieff-si}. We set initial
-$\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-observed in the state we are simulating. We hold all other regressors
-at their observed values. Error terms are drawn with replacement from
-the residuals. We do this many times and report the average over draws
-of the residuals. The shaded region is a point-wise 90\% confidence
-interval. The left column shows that the fit of the estimated and
-observed growth rate is quite good.
-
-The middle column of Figure \ref{fig:WA-mask} shows the change in
+% The
+% left column of  Figure \ref{fig:WA-mask} shows the observed, estimated
+% average, and counterfactual average of $\Delta \log \Delta C$ (top panel) and
+% $\Delta \log \Delta D$ (bottom panel). To compute the estimated and counterfactual
+% paths, we use the   estimate   in
+% column ``Average'' of Table \ref{tab:dieff-si}. We set initial
+% $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
+% observed in the state we are simulating. We hold all other regressors
+% at their observed values. Error terms are drawn with replacement from
+% the residuals. We do this many times and report the average over draws
+% of the residuals. The shaded region is a point-wise 90\% confidence
+% interval. The left column shows that the fit of the estimated and
+% observed growth rate is quite good.
+The left column of Figure \ref{fig:WA-mask} shows the change in
 growth rate from mandating masks on April 1st. The shaded region is a
@@
 estimates imply that mandating masks on April 1st would have led to
-500 fewer cases and 250 fewer deaths in Washington in total by the start of
-June.
+about 300 fewer cases and 10 fewer deaths in Washington in total by
+the start of June.
 
@@
       \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}
+      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_idx}
+      &
+        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_idx}
       \\
@@
       \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
+      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_idx}
     \end{tabular}
-  \end{minipage}
+
     \begin{flushleft}
@@
       across states of the change in growth. The shaded region is a
-      point-wise 90\% confidence interval. The right column shows the
-      change in cases or deaths relative to the baseline of actual
-      policies.
+      point-wise 90\% confidence interval for the national average
+      change in growth. The right column shows the change in cases or
+      deaths relative to the baseline of actual policies.
     \end{flushleft}
+  \end{minipage}
 \end{figure}
 
-The results for other states are similar to those for Washington. In
-the appendix, Figures  \ref{fig:MA-mask} and \ref{fig:IL-mask} display
-similar results for Massachusetts and Illinois.
+The results for other states are similar to those for Washington.
+%In
+%the appendix, Figures  \ref{fig:MA-mask} and \ref{fig:IL-mask} display
+%similar results for Massachusetts and Illinois.
 Figure \ref{fig:US-mask} shows the average change in cases and
@@
 The point estimates indicate that mandating masks on April 1st
-could have led to 25\% fewer cumulative cases and 37\% fewer
-cumulative deaths by the end of May with their 90 percent intervals given by $[10,47]$\% and $[18,55]$\%, respectively. The result roughly translates into $18$ to $55$ thousand saved lives.
-
-\subsection{Non-essential Business Closures}
-
-A particularly controversial policy is the closure of non-essential
-businesses. We now examine a counterfactual where non-essential
-businesses are never closed. Figure \ref{fig:WA-nb} shows the effect
-of leaving non-essential businesses open in Washington. The point
-estimate implies that the closure of non-essential businesses reduced cases and
-deaths by a small amount. However, this estimate is relatively
-imprecise; 90\% confidence intervals for the change in cases and deaths from
-leaving non-essential businesses open by the end of May are [-250,700] and
-[-100,1200], respectively.
+could have led to 30\% fewer cumulative cases and 34\% fewer
+cumulative deaths by the end of May with their 90 percent intervals given by $[13,46]$\% and $[12,47]$\%, respectively. The result roughly translates into $30$ thousand saved lives.
+
+\subsection{Business Closure Policies}
+
+Business closures are particularly controversial. We now examine a
+counterfactual where there are no business closure policies (no
+restaurant, movie theater, or non-essential business closure).  Figure
+\ref{fig:US-business} shows the national effect of never restricting
+businesses on cases and deaths. The point estimates imply that,
+without business closures, cases and deaths would be about 60\% at the
+end of May. The confidence intervals are wide. A 90 percent confidence
+interval for the increase in cases at the end of May is 0 to
+150\%. The confidence interval for deaths is even wider.
 
 \begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open in Washington \label{fig:WA-nb}}
+  \caption{Effect of having no business closure policies in the US\label{fig:US-business}}
   \begin{minipage}{\linewidth}
@@
       \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_v1}\\
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-dgrowth_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_idx}
+      \\
       \\
       \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_deaths_v1}
-      & \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_deaths_v1}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_deaths_idx}
     \end{tabular}
-  \end{minipage}
-\end{figure}
-
-Figure \ref{fig:US-nb} shows the national effect of leaving non
-essential businesses open on cases and deaths. For cases, the
-estimates imply that with non-essential businesses open, cases would
-be about -15 to 60\% higher in late May. The results for deaths are
-similar but less precise.
-
-\begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open in the US\label{fig:US-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths_v1}
-    \end{tabular}
+
+    \begin{flushleft}
+      \footnotesize In the left column, the dots are the average
+      change in growth in each state. The blue line is the average
+      across states of the change in growth. The shaded region is a
+      point-wise 90\% confidence interval for the national average
+      change in growth. The right column shows the change in cases or
+      deaths relative to the baseline of actual policies.
+    \end{flushleft}
   \end{minipage}
@@
       \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_idx}
       \\
@@
       \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths_v1}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths_idx}
     \end{tabular}
+
+    \begin{flushleft}
+      \footnotesize In the left column, the dots are the average
+      change in growth in each state. The blue line is the average
+      across states of the change in growth. The shaded region is a
+      point-wise 90\% confidence interval for the national average
+      change in growth. The right column shows the change in cases or
+      deaths relative to the baseline of actual policies.
+    \end{flushleft}
   \end{minipage}
@@
 
-[Discussion on the effect of school closures has been moved to sensitivity analysis section.]
+[Discussion on the effect of school closures has been moved to
+sensitivity analysis section.]
+
+[Do we even still need this section?]
 
@@
 
-\subsection{Counterfactuals for Massachusetts and
-  Illinois}
- Figures \ref{fig:MA-mask} and \ref{fig:IL-mask} present the fit of estimated cases as well as the counterfactual effect of mandating masks on April 1st in Massachusetts and Illinois, respectively.
+% \subsection{Counterfactuals for Massachusetts and
+%   Illinois}
+%  Figures \ref{fig:MA-mask} and \ref{fig:IL-mask} present the fit of estimated cases as well as the counterfactual effect of mandating masks on April 1st in Massachusetts and Illinois, respectively.
 Figures \ref{fig:MA-nb} and \ref{fig:IL-nb}   show the counterfactual effect of leaving non-essential business open in Massachusetts and Illinois, respectively.
 
-\FloatBarrier
-
-\begin{figure}[b]
-  \caption{Effect of mandating masks on April 1st in Massachusetts \label{fig:MA-mask}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_v1}
-\\
-      \multicolumn{3}{c}{\textbf{Deaths}} \\
-     \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_deaths_v1}
-    \end{tabular}
-    \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff}. We set initial
-      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-      observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
-      point-wise 90\% confidence interval.
-    \end{flushleft}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \caption{Effect of mandating masks on April 1st in Illinois \label{fig:IL-mask}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_v1}      \\
-      \multicolumn{3}{c}{\textbf{Deaths}} \\
-    \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_deaths_v1}
-    \end{tabular}
-    \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff}. We set initial
-      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-      observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
-      point-wise 90\% confidence interval.
-    \end{flushleft}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \caption{Effect of leaving businesses open in Massachusetts \label{fig:MA-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_v1}
-       &
-         \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_v1}
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_deaths_v1}
-       &      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[h]
-  \caption{Effect of leaving businesses open in Illinois \label{fig:IL-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_v1} \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-           \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_deaths_v1}
-      &       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_deaths_v1}    \end{tabular}
-  \end{minipage}
-\end{figure}
+% \FloatBarrier
+
+% \begin{figure}[b]
+%   \caption{Effect of mandating masks on April 1st in Massachusetts \label{fig:MA-mask}}
+%   \begin{minipage}{\linewidth}
+%     \centering
+%     \begin{tabular}{ccc}
+%       \multicolumn{3}{c}{\textbf{Cases}} \\
+%       \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth}
+%       &
+%         \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_v1}
+%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_v1}
+% \\
+%       \multicolumn{3}{c}{\textbf{Deaths}} \\
+%      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth_deaths}
+%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_deaths_v1}
+%       &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_deaths_v1}
+%     \end{tabular}
+%     \begin{flushleft}
+%       \footnotesize To compute the estimated and counterfactual paths
+%       we use the average of two estimated coefficients as reported in
+%       column ``Average'' of Table \ref{tab:dieff}. We set initial
+%       $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
+%       observed in the state we are simulating. We hold all other
+%       regressors at their observed values. Error terms are drawn with
+%       replacement from the residuals. We do this many times and report
+%       the average over draws of the residuals. The shaded region is a
+%       point-wise 90\% confidence interval.
+%     \end{flushleft}
+%   \end{minipage}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \caption{Effect of mandating masks on April 1st in Illinois \label{fig:IL-mask}}
+%   \begin{minipage}{\linewidth}
+%     \centering
+%     \begin{tabular}{ccc}
+%       \multicolumn{3}{c}{\textbf{Cases}} \\
+%       \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth}
+%       &
+%         \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_v1}
+%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_v1}      \\
+%       \multicolumn{3}{c}{\textbf{Deaths}} \\
+%     \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth_deaths}
+%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_deaths_v1}
+%       &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_deaths_v1}
+%     \end{tabular}
+%     \begin{flushleft}
+%       \footnotesize To compute the estimated and counterfactual paths
+%       we use the average of two estimated coefficients as reported in
+%       column ``Average'' of Table \ref{tab:dieff}. We set initial
+%       $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
+%       observed in the state we are simulating. We hold all other
+%       regressors at their observed values. Error terms are drawn with
+%       replacement from the residuals. We do this many times and report
+%       the average over draws of the residuals. The shaded region is a
+%       point-wise 90\% confidence interval.
+%     \end{flushleft}
+%   \end{minipage}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \caption{Effect of leaving businesses open in Massachusetts \label{fig:MA-nb}}
+%   \begin{minipage}{\linewidth}
+%     \centering
+%     \begin{tabular}{cc}
+%       \multicolumn{2}{c}{\textbf{Cases}} \\
+%       \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_v1}
+%        &
+%          \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_v1}
+%       \\
+%       \multicolumn{2}{c}{\textbf{Deaths}} \\
+%         \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_deaths_v1}
+%        &      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_deaths_v1}
+%     \end{tabular}
+%   \end{minipage}
+% \end{figure}
+
+% \begin{figure}[h]
+%   \caption{Effect of leaving businesses open in Illinois \label{fig:IL-nb}}
+%   \begin{minipage}{\linewidth}
+%     \centering
+%     \begin{tabular}{cc}
+%       \multicolumn{2}{c}{\textbf{Cases}} \\
+%       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_v1}
+%       &
+%         \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_v1} \\
+%       \multicolumn{2}{c}{\textbf{Deaths}} \\
+%            \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_deaths_v1}
+%       &       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_deaths_v1}    \end{tabular}
+%   \end{minipage}
+% \end{figure}
 
@@
   Variables}
+
+\fixme
+% DO WE STILL WANT THESE? THESE FIGURES HAVE NOT BEEN UPDATED.
 

### 2020-08-27 tex/reply-to-referees.tex

@@
 \def\bs{\boldsymbol}
+
+\usepackage{xr}
+\externaldocument{covidEffectOfNPI-revision}
 
@@
 
-\item   
+\item   We  state the limitations of our work with cautious notes on interpreting our causal model.
+\item   Following the editor's suggestion, Section 2 is reorganized and we improve our discussion on the testable restrictions.
+Tables 7 and 9 in Section 3 provide   an  over-identification test for no indirect effect of mask policy on case/death growth via behavior.
+\item   Following suggestions by the editor and the referees, we provide various sensitivity checks as well as the results of debiased fixed effects estimator  in Section 4.
+\item   We improve the figures for counterfactual experiments so that they are now easier to understand. 
+\item  We adjusted the timing of monthly dummies for estimating behavior equation ($PI \to B$)  in Table 3.
+In the previous version, the definition of monthly dummies for behavior equation ($PI \to B$) is different from that for case growth equation ($BPI \to Y$)  or ($PI \to Y$).
+This issue is now fixed.
 %\item We made some grammatical adjustments.
@@
 \changes
-\subsection*{Reply to Comments}
+\subsection*{Reply to Editor's Comments}
 \begin{itemize}
@@
 \changes
-\subsection*{Reply to Comments}
+\subsection*{Reply to Referee 1's Comments}
 \begin{itemize}
@@
 
-Following your suggestion, we have added a comment on this on page xx in Section 4.2.
+Following your suggestion, we have added the following paragraph:
+\begin{quote}
+The results in Table \ref{tab:PtoY-fe} should be interpreted with caution.
+The fixed effects approach may not be preferred to random effects approach here because the former relies on long time and cross-sectional histories but, in our data, the effective time-dimension is short and a number of states is not large.
+Furthermore, the fixed effects approach could suffer more from  measurement errors, which could be a concern for our behavior and policy variables.
+\end{quote}
 
@@
 
-\subsection*{Reply to Minor Comments}
+\subsection*{Reply to Referee 1's Minor Comments}
 \begin{itemize}
@@
 \changes
-\subsection*{Reply to Comments}
+\subsection*{Reply to Comments  by C. Jessica E. Metcalf }
 \begin{enumerate}
@@
 Given  your suggestion, Figure 9 presents whisker-dot-plots for confidence intervals of mask coefficients (as well as other policy variables) across different specifications, timing assumptions, and estimation methods.
-We believe this figure will provide a sufficient robust sense of the uncertainty on the  estimates.
+We believe these figures provide a sufficient robust sense of the uncertainty on the  estimates.
 
@@
 
-We have reviewed and added more references from the epidemiology literature.
+We have reviewed and added more references from the epidemiology literature. See, for example, footnote 9.
 
@@
 
-[This is something we don't do given the time/space constraint.]
+This is an interesting suggestion.
+We thought about the possibility of providing a version of Figure 8 based on the regression residuals after taking out the effect of confounders etc but, eventually, we decided not to present such results due to the space constraint.
 

### 2020-08-27 cases_and_policies/rmd/cf-pindex.Rmd

@@
 m$robustvcv <- V
+m$beta <- m$coefficients
+m$cse <- sqrt(diag(V))
 
@@
 ## National
-
-These figures show the average effect on $\Delta \log \Delta C$ across
-states, and the total national effect on cases in past week.
-
-That is for the first figure, we take:
-
-$$
-\hat{E}[\Delta Y_{it}^\ast | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
-\frac{1}{N} \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \Delta Y_{it}^{\ast,r}
-$$
-
-and plot the difference of the above between two different values of $P_{it}^\ast$
-
-The second figure shows the total change in cases, that is the change in:
-
-$$
-\hat{E}[\exp(Y_{it}^\ast) | Y_{i1}, ..., Y_{iL}, \{X_{it}, P_{it}^\ast\}_{t=1}^T, \theta] =
-\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}).
-$$
-
-The final figure shows
-$$
-\frac{\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{\ast,r}) - \sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^r)} {\sum_{i=1}^{N} \frac{1}{R} \sum_{r=1}^R \exp(Y_{it}^{r})}
-$$
-that is, the (average over residual draws) change in national cases relative to (average over residual draws) national cases.
 
@@
 m$coefficients <- subcoefs(di, m$coefficients)
+m$beta <- m$coefficients
 m$vcv <- V
@@
 m$robustvcv <- V
+m$cse <- sqrt(diag(V))
 

### 2020-08-27 cases_and_policies/rmd/regressionsWithPlots.Rmd
Committing changes before mergning master.

@@
         #lib_dir : deps
-        self_cononontained : true
+        self_contained : true
         fig_width: 8
@@
 source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
-
 ```
@@
 ```{r testsmooth, cache=FALSE}
-rearrange <- function(x, id) {
-  for (i in unique(id)) {
-    xi <- x[id==i]
-    xi[!is.na(xi)] <- sort(xi[!is.na(xi)])
-    x[id==i] <- xi
-  }
-  return(x)
-}
-
-dropnotincreasing <- function(x, id, t) {
-  const <- (x <= panellag(x, id, t))
-  const[is.na(const)] <- FALSE
-  x[const & x>0] <- NA
-  return(x)
-}
-
-lininterp <- function(x, id, t) {
-  for (i in unique(id)) {
-    xi <- x[id==i]
-    ti <- t[id==i]
-    xi[is.na(xi)] <- approx(ti[!is.na(xi)], xi[!is.na(xi)], ti[is.na(xi)])$y
-    x[id==i] <- xi
-  }
-  return(x)
-}
-
-smoothtests <- function(x, id, t) {
-  x[is.na(x)] <- 0
-  return(lininterp(dropnotincreasing(rearrange(x,id),id,t),id,t))
-}
-
-df$neg.interp <- smoothtests(df$negative, df$state, df$date)
-df$pos.interp <- df$cases #smoothtests(df$positive, df$state, df$date)
-df$tot.interp <- df$neg.interp + df$pos.interp
-
-L <- 7
-df$tests <- paneldiff(df$tot.interp, df$state, df$date, lag=L)
-df$testrate  <- df$tests/df$Population.2018*1000
-df$testrate[df$testrate<0] <- NA
-df$dlogtests <- paneldiff(log(sapply(df$testrate, function(x)
-  max(x,exp(-1)))), df$state, df$date, lag=L)
-df$testratedc <- df$ddeath/df$dcases*df$testrate
-df$testratedc[df$dcases==0] <- 0
-
-df$tests.ma <- paneldiff(panelma(df$neg.interp, df$state, df$date, 7) + df$cases,
-                         df$state, df$date, lag=L)
-df$testrate.ma  <- df$tests.ma/df$Population.2018*1000
-df$testrate.ma[df$testrate.ma<0] <- NA
-df$dlogtests.ma <- paneldiff(log(sapply(df$testrate.ma, function(x)
-  max(x,exp(-1)))), df$state, df$date, lag=L)
-df$testratedc.ma <- df$ddeath/df$dcases*df$testrate.ma
-df$testratedc.ma[df$dcases==0] <- 0
-
-df$logtests <- log(sapply(df$testrate, function(x)
-  max(x,exp(-1))))
-
-df$lagtests <- panellag(df$dlogtests, df$state, df$date, lag=1)  
-df$dlogtests[is.na(df$dlogtests)] <- df$lagtests[is.na(df$dlogtests)]
-
+source("smoothtests.R")
 ```
@@
 prefix="cases-fe"
-#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe")
+#savetextables(NULL, NULL, piy[1:4], prefix="cases-fe") copying and
+
+# pasting this is bad style and seems like it might create maintenance
+# problems ...
 ylbl <- colnames(piy[[1]]$reg$response)
@@
 ## DML
-```{r regsdml,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsdml,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
 

### 2020-08-27 tex/covidEffectOfNPI-revision.tex

--- 2020-08-26 tex/covidEffectOfNPI-revision.tex
@@
 
-Figure \ref{fig:US-mask-dgrowth} shows that nationally  implementing mandatory
-face masks  for employees in public businesses on April 1st would have reduced the growth rate of cases (top panel)
-and that of deaths (bottom panel) by more than 10 percentage points in late April.
-This leads to reductions of $25$\% and $35$\% in reported cases and deaths, respectively,  by the end of May with a 90 percent confidence interval of $[10,45]$\% and $[17,55]$\%, which  roughly implies that as many as 17 to 55
-  thousand lives could have been saved.\footnote{As of May 27, 2020,
-   the US Centers for Disease Control and Prevention  reports 99,031
-   deaths in the US.}
+Our estimates imply that nationally  implementing mandatory
+face masks  for employees in public businesses on April 1st would have reduced the growth rate of cases
+and that of deaths by more than 10 percentage points in late April. As
+shown in Figure \ref{fig:US-mask-dgrowth}, this leads to reductions of
+$20$\% and $35$\% in weekly reported cases and deaths, respectively,
+by the end of May with a 90 percent confidence interval of $[8,32]$\%
+and $[17,50]$\%, which roughly implies that 6 thousand lives could
+have been saved.\footnote{The US averaged roughly one thousand deaths
+  per day during May. Reducing one thousand deaths per day by 20\%
+  over 30 days gives a rough estimate of 6 thousand saved lives.}
 This finding is significant: given this potentially large  benefit of reducing the spread of Covid-19, mandating masks is an attractive policy instrument especially because it involves relatively little economic disruption.
@@
  %  \\
-     \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_v1}        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_v1}
+      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_idx}
+      & \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_idx}
     \end{tabular}
-  \end{minipage}
-%    \begin{flushleft}
-%      \footnotesize
-%      In the left column, the dots are the average change in growth in
-%      each state. The blue line is the average across states of the
-%      change in growth. The shaded region is a point-wise 90\% confidence
-%      interval. The right column shows the change in cases or deaths
-%      relative to the baseline of actual policies.
-%    \end{flushleft}
+    \begin{flushleft}
+      \footnotesize This figure shows the estimated relative change in
+      cases if all states had mandated masks on April 1st. The thick
+      blue line is the estimated change in weekly cases or deaths
+      relative to the observed number of weekly cases or deaths. The
+      shaded region is a point wise 90\% confidence band.
+    \end{flushleft}
+  \end{minipage}
 \end{figure}
@@
  % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of   leaving non-essential businesses open on cases in the US  \label{fig:US-nb-dgrowth}}
+  \caption{Effect  of no business closure policies on cases in the US  \label{fig:US-nb-dgrowth}}
   \begin{minipage}{\linewidth}
@@
 %
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_idx}
 %      \\
@@
     \end{tabular}
+    \begin{flushleft}
+      \footnotesize This figure shows the estimated relative change in
+      cases if no states had never implemented any business closure
+      policies. The thick blue line is the estimated change in weekly
+      cases or deaths relative to the observed number of weekly cases
+      or deaths. The shaded region is a point wise 90\% confidence
+      band.
+    \end{flushleft}
   \end{minipage}
@@
  % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth}}
+  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth_idx}}
   \begin{minipage}{\linewidth}
@@
     \end{tabular}
+    \begin{flushleft}
+      \footnotesize This figure shows the estimated relative change in
+      cases if no states had ever issued stay at home orders. The thick
+      blue line is the estimated change in weekly cases or deaths
+      relative to the observed number of weekly cases or deaths. The
+      shaded region is a point wise 90\% confidence band.
+    \end{flushleft}
   \end{minipage}
@@
 
-In order to compute the counterfactual relative contrasts for levels variables, for example, confirmed cases $C_{it}$ 
+[Doesn't this just give us weekly cases instead of cumulative
+cases? I.e. following the notation in \ref{eq:M}, it these $C_{it}$
+should be $\Delta C_{it}$.]
+
+In order to compute the counterfactual relative contrasts for levels
+variables, for example, confirmed cases $C_{it}$
 we would use the formula:
@@
 \begin{equation}
-C^{\star}_{it}/C_{it} =  \sum_{i=1}^N w_i \exp(  \sum_{\bar t=1}^t ( Y^\star_{i\bar t} -  Y_{i\bar t}) ), \quad w_i = C_{it}/\sum_{i=1}^N C_{it}.
+C^{\star}_{it}/C_{it} =  \sum_{i=1}^N w_i \exp(  \sum_{\bar t=1}^t (
+Y^\star_{i\bar t} -  Y_{i\bar t}) ), \quad w_i = C_{it}/\sum_{i=1}^N
+C_{it}.  \label{eqn:relative_national}
 \end{equation}
@@
 In Panel (B) of  Figure  \ref{fig:whisker}, many estimates of closures of K-12 schools  suggest that  the effect of school closures is large.
-The visual evidence on growth rates for states with and without school closures in Figure \ref{fig:growthpolicies1} also suggests that there may be a potentially large effect, though the history is very short.
+The visual evidence on growth rates for states with and without school closures in Figure \ref{fig:growthpolicies-school} also suggests that there may be a potentially large effect, though the history is very short.
 This evidence is consistent
@@
  \begin{figure}
-  \caption{Case and death growth conditional on school closures \label{fig:growthpolicies1}}
+  \caption{Case and death growth conditional on school closures \label{fig:growthpolicies-school}}
   \begin{minipage}{\linewidth}
@@
     \centering
-    \begin{tabular}{ccc}
-      \multicolumn{3}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
-      &
-        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_v1}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_v1}
-      \\
-      \\
-      \multicolumn{3}{c}{\textbf{Deaths}}
-      \\      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths}
-      &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_v1}
-      &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_v1}
+    \begin{tabular}{cc}
+      \multicolumn{2}{c}{\textbf{Cases}} \\
+      %\includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
+      % &
+        \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dgrowth_idx}
+      &        \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dcases_idx}
+      \\
+      \\
+      \multicolumn{2}{c}{\textbf{Deaths}}
+      \\
+      % \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths} &
+                                                                                                   \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_idx}
+      &      \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_idx}
     \end{tabular}
+
     \begin{flushleft}
-      \footnotesize To compute the estimated and counterfactual paths
-      we use the average of two estimated coefficients as reported in
-      column ``Average'' of Table \ref{tab:dieff-si}. We set initial
-      $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-      observed in the state we are simulating. We hold all other
-      regressors at their observed values. Error terms are drawn with
-      replacement from the residuals. We do this many times and report
-      the average over draws of the residuals. The shaded region is a
-      point-wise 90\% confidence interval.
+      \footnotesize The vertical text denotes the dates that
+      Washington imposed each policy.  To compute the estimated and
+      counterfactual paths we use the average of two estimated
+      coefficients as reported in column ``Average'' of Table
+      \ref{tab:dieff-con-si}. We set initial $\Delta \log \Delta C$
+      and $\log \Delta C$ to their values first observed in the state
+      we are simulating. We hold all other regressors at their
+      observed values. The shaded region is a point-wise 90\%
+      confidence interval.
     \end{flushleft}
@@
 For illustrative purpose, we begin by focusing on Washington State.
-The
-left column of  Figure \ref{fig:WA-mask} shows the observed, estimated
-average, and counterfactual average of $\Delta \log \Delta C$ (top panel) and
-$\Delta \log \Delta D$ (bottom panel). To compute the estimated and counterfactual
-paths, we use the   estimate   in
-column ``Average'' of Table \ref{tab:dieff-si}. We set initial
-$\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-observed in the state we are simulating. We hold all other regressors
-at their observed values. Error terms are drawn with replacement from
-the residuals. We do this many times and report the average over draws
-of the residuals. The shaded region is a point-wise 90\% confidence
-interval. The left column shows that the fit of the estimated and
-observed growth rate is quite good.
-
-The middle column of Figure \ref{fig:WA-mask} shows the change in
+% The
+% left column of  Figure \ref{fig:WA-mask} shows the observed, estimated
+% average, and counterfactual average of $\Delta \log \Delta C$ (top panel) and
+% $\Delta \log \Delta D$ (bottom panel). To compute the estimated and counterfactual
+% paths, we use the   estimate   in
+% column ``Average'' of Table \ref{tab:dieff-si}. We set initial
+% $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
+% observed in the state we are simulating. We hold all other regressors
+% at their observed values. Error terms are drawn with replacement from
+% the residuals. We do this many times and report the average over draws
+% of the residuals. The shaded region is a point-wise 90\% confidence
+% interval. The left column shows that the fit of the estimated and
+% observed growth rate is quite good.
+The left column of Figure \ref{fig:WA-mask} shows the change in
 growth rate from mandating masks on April 1st. The shaded region is a
 90\% pointwise confidence interval. As shown, mandating masks on April
-1st lowers the growth of cases or deaths 14 or 21 days later by 0.1 to
-0.15. This effect then gradually declines due to information
+1st lowers the growth of cases or deaths 14 or 21 days later by 0.09 or
+0.14. This effect then gradually declines due to information
 feedback. Mandatory masks reduce past cases or deaths, which leads to
@@
 effect has decayed through information feedback, and we are comparing
-it the undecayed impact of Washington's actual, later mask mandate.
+it to the undecayed impact of Washington's actual, later mask mandate.
 
 The right column of Figure \ref{fig:WA-mask} shows how the changes in
-case and death growth translate into changes in cases and deaths. The
-estimates imply that mandating masks on April 1st would have led to
-500 fewer cases and 250 fewer deaths in Washington in total by the start of
-June.
+case and death growth translate into changes in weekly cases and
+deaths. The estimates imply that mandating masks on April 1st would
+have led to about 150 fewer cases and 20 fewer deaths per week in
+Washington by the start of June.
 
@@
       across states of the change in growth. The shaded region is a
-      point-wise 90\% confidence interval. The right column shows the
-      change in cases or deaths relative to the baseline of actual
-      policies.
+      point-wise 90\% confidence interval for the national average
+      change in growth. The right column shows the change in cases or
+      deaths relative to the baseline of actual policies.
     \end{flushleft}
+  \end{minipage}
 \end{figure}
 
-The results for other states are similar to those for Washington. In
-the appendix, Figures  \ref{fig:MA-mask} and \ref{fig:IL-mask} display
-similar results for Massachusetts and Illinois.
-Figure \ref{fig:US-mask} shows the average change in cases and
-deaths across states, where the top panel shows the effect on cases and the bottom panel shows the effect on deaths.
-The point estimates indicate that mandating masks on April 1st
-could have led to 25\% fewer cumulative cases and 37\% fewer
-cumulative deaths by the end of May with their 90 percent intervals given by $[10,47]$\% and $[18,55]$\%, respectively. The result roughly translates into $18$ to $55$ thousand saved lives.
-
-\subsection{Non-essential Business Closures}
-
-A particularly controversial policy is the closure of non-essential
-businesses. We now examine a counterfactual where non-essential
-businesses are never closed. Figure \ref{fig:WA-nb} shows the effect
-of leaving non-essential businesses open in Washington. The point
-estimate implies that the closure of non-essential businesses reduced cases and
-deaths by a small amount. However, this estimate is relatively
-imprecise; 90\% confidence intervals for the change in cases and deaths from
-leaving non-essential businesses open by the end of May are [-250,700] and
-[-100,1200], respectively.
+The results for other states are similar to those for Washington.
+Figure \ref{fig:US-mask} shows the national effect of mask mandate on
+April 1st. The top panel shows the effect on cases and the bottom
+panel shows the effect on deaths. The left column shows the change in
+growth rates. Dots are the average change in growth rate in each state
+(i.e. the dots are identical to the solid line in the left panel of
+Figure \label{fig:WA-mask}, except for all states instead of just
+Washington). The solid blue is the national average change in growth
+rate. The shaded region is a 90\% pointwise confidence band for the
+national average change in growth rate. The right column of Figure
+\ref{fig:US-mask} shows the national relative change in weekly cases
+or deaths. The point estimates indicate that mandating masks on April
+1st could have led to 20\% fewer weekly cases and 35\% fewer
+weekly deaths by the end of May with their 90 percent intervals
+given by $[8,32]$\% and $[18,50]$\%, respectively.
+
+%The result roughly
+%translates into $30$ thousand saved lives.
+
+\subsection{Business Closure Policies}
+
+Business closures are particularly controversial. We now examine a
+counterfactual where there are no business closure policies (no
+restaurant, movie theater, or non-essential business closure).  Figure
+\ref{fig:US-business} shows the national effect of never restricting
+businesses on cases and deaths. The point estimates imply that,
+without business closures, cases and deaths would be about 75\% higher
+at the end of May. The confidence intervals are wide. A 90 percent
+confidence interval for the increase in cases at the end of May is 25\%
+to 140\%. The confidence interval for deaths is even wider, $[10, 250]$\%
 
 \begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open in Washington \label{fig:WA-nb}}
+  \caption{Effect of having no business closure policies in the US\label{fig:US-business}}
   \begin{minipage}{\linewidth}
@@
       \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_v1}\\
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-dgrowth_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_idx}
+      \\
       \\
       \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dgrowth_deaths_v1}
-      & \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nb-dcases_deaths_v1}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_deaths_idx}
     \end{tabular}
+
+    \begin{flushleft}
+      \footnotesize In the left column, the dots are the average
+      change in growth in each state. The blue line is the average
+      across states of the change in growth. The shaded region is a
+      point-wise 90\% confidence interval for the national average
+      change in growth. The right column shows the change in cases or
+      deaths relative to the baseline of actual policies.
+    \end{flushleft}
   \end{minipage}
@@
 
-Figure \ref{fig:US-nb} shows the national effect of leaving non
-essential businesses open on cases and deaths. For cases, the
-estimates imply that with non-essential businesses open, cases would
-be about -15 to 60\% higher in late May. The results for deaths are
-similar but less precise.
-
-\begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open in the US\label{fig:US-nb}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_v1}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths_v1}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths_v1}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
 \subsection{Stay-at-home orders}
@@
 We next examine a counterfactual where stay-at-home orders had been never
-issued.  Figure \ref{fig:US-shelter} shows the average effect of no
+issued.  Figure \ref{fig:US-shelter} shows the national effect of no
 stay-at-home orders. On average, without stay-at-home orders, case
 growth rate would have been nearly 0.1 higher in late April. This
-translates to 80\% [25\%,170\%] more cases by the start of June. The
+translates to 60\% [10\%,140\%] more cases per week by the start of June. The
 results for deaths are similar, but slightly less precise, with no
@@
 
-\section{Counterfactual Effect of Removing All Policies and its Sensitivity}\label{removing-policies}
+%\section{Counterfactual Effect of Removing All Policies and its Sensitivity}\label{removing-policies}
 
@@
 
-We now consider the impact of changing from the observed policies to none.
-Figure \ref{fig:US-nop-SI} shows the average across states of the change
-in case growth and relative increase in cases under a specification without past national case variables.
-Removing policies leads to an increase of above 0.2 in
-case growth throughout April and May. The confidence interval is
-fairly wide, and its upper bound includes a very large increase in
-cases by the end of May.  The right panel displays the national
-increase in aggregate cases without any policy intervention. The
-estimates imply at least a 7 fold increase in cases with a large upper bound by the end of May, or
-at least 14 million additional cases. The estimated impact on deaths is larger than
-cases, and even more imprecise.
-
-[Discussion on the effect of school closures has been moved to sensitivity analysis section.]
+% We now consider the impact of changing from the observed policies to none.
+% Figure \ref{fig:US-nop-SI} shows the average across states of the change
+% in case growth and relative increase in cases under a specification without past national case variables.
+% Removing policies leads to an increase of above 0.2 in
+% case growth throughout April and May. The confidence interval is
+% fairly wide, and its upper bound includes a very large increase in
+% cases by the end of May.  The right panel displays the national
+% increase in aggregate cases without any policy intervention. The
+% estimates imply at least a 7 fold increase in cases with a large upper bound by the end of May, or
+% at least 14 million additional cases. The estimated impact on deaths is larger than
+% cases, and even more imprecise.
+
+% [Discussion on the effect of school closures has been moved to
+% sensitivity analysis section.]
+
+% [Do we even still need this section?]
 
@@
 
-First, our empirical analysis indicates that mandating face masks has reduced the spread of Covid-19 without affecting people's social distancing behavior measured by Google Mobility Reports.
-Our counterfactual experiment based on the estimated model suggests that if all states had have adopted mandatory face mask policies on April 1st of 2020, then the number of deaths by the end of May would have been smaller by as much as  $17$ to $55$\%, which roughly translates to $17$ to $55$ thousand saved lives.
+First, our empirical analysis indicates that mandating face masks has
+reduced the spread of Covid-19 without affecting people's social
+distancing behavior measured by Google Mobility Reports. Our
+counterfactual experiment based on the estimated model suggests that
+if all states had have adopted mandatory face mask policies on April
+1st of 2020, then the number of deaths per week
+by the end of May would have been smaller by as much as  $18$ to $50$\%, which roughly translates to $6$ thousand saved lives.
 

### 2020-08-28 cases_and_policies/rmd/regressionsWithPlots.Rmd
Updated whiskerplots, which were previously out of date and incorrect.

@@
 ## DML
-```{r regsdml,  eval = FALSE, cache=FALSE, warning=FALSE, results='asis'}
+```{r regsdml,  eval = TRUE, cache=FALSE, warning=FALSE, results='asis'}
 

### 2020-08-29 cases_and_policies/rmd/cf-pindex.Rmd
Added cumulative cases and deaths to counterfactuals.

@@
 title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
-subtitle    : "Counterfactuals (with policies averaged)"
+subtitle    : "Counterfactuals"
 author      :  Victor Chernozhukov,  Hiro Kasahara, Paul Schrimpf
@@
 
-# Counterfactuals
-
 ```{r cffuncs, cache=FALSE}
@@
 
-```{r statesim, cache=FALSE}
+```{r regressions, cache=FALSE}
 infovars <- list(c("dlogdc", "logdc")) #,
@@
 cnames <- c(pols, infovars[[1]])
+
 di <- dieff_table2(lapply(regs$pib[[1]][1:4], function(x) x$reg$coef[,1]),
@@
 
-simmodel <- createforwardsim(m, sdf)
-
-# remove policies
-cf <- sdf
-cf[,pols] <- 0
-nopmodel <- createforwardsim(m, cf)
-
-rerunsim=FALSE
-if (!file.exists("cfnop_idx.Rda") || rerunsim) {
-  cfnoplist <- lapply(unique(sdf$state), function(s)
-    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
-  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
-  save(cfnoplist,allmeans, file="cfnop_idx.Rda")
-}
-load("cfnop_idx.Rda")
-cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
+rerunsim <- FALSE
 ```
 
-### Removing Policies
+# Method
 
-We now examine the impact changing from the observed policies to
-none. For illustrative purpose, we being by focusing on
-Washington. Figure \ref{fig:WA-nop-dlogdc} shows the observed,
-estimated average, and counterfactual without policies average of
-$\Delta \log \Delta C$ in Washington.
+We compute counterfactuals from the "total effect" version of the model, with behavior concentrated out.
 
-```{r WA-nop-dlogdc, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
-st <- "Washington"
-f <- cfplots(st, cfnop, data=df)
-grid.arrange(f$figd, f$figdp, nrow=2, top =
-                                        textGrob(st,gp=gpar(fontsize=20,font=2)))
+$$
+Y_{i,t+\ell}
+= \mathsf{a}'
+P_{it} + \mathsf{b}' I_{it} +  {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
+$$
 
-```
+We consider a counterfactual change of $P_{it}$ to $P_{it}^*$, while
+$W_{it}$ and $\bar \varepsilon_{it}$ are held constant. In response to
+the policy change, $Y_{i,t+\ell}$ and the part of $I_{it}$ that
+contains $Y_{it}$, change to $Y_{i,t+\ell}^*$ and $I_{it}^*$. To be
+specific, let $C_{it}$ denote cumulative cases in state $i$ on day
+$t$. Our outcome is:
 
-```{r WA-nop-cases, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
-grid.arrange(f$figl, f$figp, nrow=2) #,
-#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
-```
+$$
+Y_{i,t+\ell} \equiv \log
+\left(C_{i, t+\ell} - C_{i, t + \ell - 7} \right) - \log
+\left(C_{i, t+\ell-7} - C_{i, t + \ell - 14} \right) \equiv \Delta
+\log \Delta C_{i,t+\ell}
+$$
 
-## National
+where $\Delta$ is a 7 day difference operator. Writing the model in
+terms of $\Delta \log \Delta C$ we have:
 
-```{r nop-all, cache=TRUE}
-abbrv <- unique(df[,c("state","ST")])
-cfnop <- merge(cfnop, abbrv, by="state")
+$$
+\Delta \log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \mathsf{b}_D
+\Delta \log \Delta C_{i,t} + \mathsf{b}_L \log \Delta C_{i,t} +
+{\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
+$$
 
-figs <- nationalplots(allmeans,cfnop, "removing policies")
+To simplify computation, we rewrite this model in terms of $\log
+\Delta C$:
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_idx.pdf",sep="/"), width=5,height=3.75)
-figs$g
-dev.off()
+$$
+\log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \log \Delta
+C_{i,t+\ell-7} + (\mathsf{b}_D + \mathsf{b}_L) \log \Delta C_{i,t} -
+\mathsf{b}_D \log \Delta C_{i,t-7} + {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
+$$
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_idx.pdf",sep="/"), width=5,height=3.75)
-figs$c
-dev.off()
+This equation is used to iteratively compute
+$\log \Delta C_{i,t+\ell}^*$
+conditional on initial $\{\log \Delta C_{i,s}\}_{s=-\ell-7}^{0}$,
+and the entire sequence of $\{W_{i,t}, P_{i,t}^*, \bar \varepsilon_{it} \}_{t=0}^T$.
 
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_idx.pdf",sep="/"), width=5,height=3.75)
-figs$r
-dev.off()
+Weekly cases instead of log weekly cases are given simply by
 
-for (f in figs) print(f)
-```
+$$ \Delta C_{i,t+\ell}^* = \exp(\log \Delta C_{i,t+\ell}^*)$$
+
+Cumulative cases can be recursively computed as:
+
+$$
+  C_{i,t} = C_{i,t-7} + \Delta C_{i,t} = C_{i,t-7} + \exp(\log \Delta C_{i,t})
+$$
+
+given initial $\{C_{i,s}\}_{s=-7}^0$.
+
+Note that $\log \Delta C_{i,t+\ell}^*$ depends linearly on
+$\bar \varepsilon$, so the residuals (and our decision to condition on
+versus intergrate out) do not matter when considering linear
+constrasts of $\log \Delta C_{it}$ or $\Delta \log \Delta C_{it}$, or
+when considering relative constrasts of $\Delta C_{it}$.
+
+On the other hand, cumulative cases are a nonlinear function of
+$\bar\varepsilon$, so the conditioning on $\bar\varepsilon$ might
+matter (not that there's anything wrong with that).
+
+## Inference
+
+Let $S(\theta, \mathbf{\varepsilon})$ denote some counterfactual
+quantity or contrast of interest, where $\theta = (\mathsf{a},
+\mathsf{b}, \tilde{\delta})$ are the parameters, and
+$\mathbf{\varepsilon}$ is the vector of residuals. Examples of $S$
+that we compute include:
+
+- Contrasts of growth rates: $S(\theta, \mathbf{\varepsilon}) =
+  \Delta \log \Delta C_{it}^* - \Delta \log \Delta C_{it}$
+
+- Relative contrasts of weekly cases: $S(\theta, \mathbf{\varepsilon}) =
+  \Delta C_{it}^*/\Delta C_{it}$
+
+- Constrasts of cumulative cases: $S(\theta, \mathbf{\varepsilon}) =
+  C_{it}^*/C_{it}$
+
+The first two examples do not actually depend on
+$\mathbf{\varepsilon}$, but the third one does. Inference is by
+simulation. Let $\hat{\theta}$ denote our point estimates and
+$\hat{\mathbf{\varepsilon}}$ the associated residuals. We draw
+$\tilde{\theta}_j$ from the asymptotic distribution of
+$\hat{\theta}$. Currently the code then computes
+
+$$ \tilde{s}_j = S(\tilde{\theta}_j, \hat{\mathbf{\varepsilon}}) $$
+
+for $j=1,..., 200$ and plots the mean across $j$ as a point estimate
+and quantiles across $j$ for confidence intervals. This is fine when
+$\varepsilon$ does not affect the quantity of interest, but I think
+the correct thing to when $\varepsilon$ matters is to compute
+
+$$ \tilde{s}_j = S(\tilde{\theta}_j, \tilde{\mathbf{\varepsilon}}_j) $$
+
+where $\tilde{\mathbf{\varepsilon}}_j$ are the residuals from using
+$\tilde{\theta}_j$ as coefficients.
+
+# Cases
 
@@
 We now look at what happens when all states impose a mask mandate on
-April 1st.
+May 14th.
 
 ```{r mask_sim, cache=FALSE}
+sdf <- subset(df, df$date>=datelims[1]) #+L)
+simmodel <- createforwardsim(m, sdf)
+abbrv <- unique(df[,c("state","ST")])
+
 cf <- df
-cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
+cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-03-14")
 cf[,"pmaskbus"] <-
@@
 
-```{r, cache=TRUE}
+```{r, cache=FALSE}
 cfmask <- merge(cfmask, abbrv, by="state")
@@
   f <- cfplots(st, cfmask, data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))
+  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on May 14th in %s",st)))
 
@@
   dev.off()
+
+  if (st=="Washington") for (fig in f) print(fig)
 }
 
-figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n")
+figs <- nationalplots(maskall, cfmask, "mandating masks on May 14th\n")
 
@@
 if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
+  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
@@
 We now look at what happens when all states impose a mask mandate on
-April 1st.
+May 14th.
 
@@
 cf <- df
-cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-04-01")
+cf[,"Mandate.face.mask.use.by.employees.in.public.facing.businesses"]  <- as.Date("2020-03-14")
 cf[,"pmaskbus"] <-
@@
   f <- cfplots(st, cfmask, yvar="deaths", data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on April 1st in %s",st)))
+  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on May 14th in %s",st)))
 
@@
 
-figs <- nationalplots(maskall, cfmask, "mandating masks on April 1st\n", yvar="deaths")
+figs <- nationalplots(maskall, cfmask, "mandating masks on May 14th\n", yvar="deaths")
 
@@
 if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
+  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))

### 2020-08-29 cases_and_policies/rmd/counterfactuals.R

@@ -24,7 +24,7 @@ source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
 source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
 source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
 source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
-
+source(paste(rootdir,"cases_and_policies/rmd/smoothtests.R", sep="/"))
 figdatelims <- c(as.Date("2020-03-07"), max(df$date))
 ################################################################################
 
@@ -53,16 +53,12 @@ figdatelims <- c(as.Date("2020-03-07"), max(df$date))
 #' df <- df[order(df$id,$df$t),]
 #' df$dy <- paneldiff(df$y, df$id, df$t, lag=2)
 #' fmla <- dy ~ lag(y,3) + lag(dy,1) + x
-#' # hack to get get id
 #'
 #' sim <- createforwardsim(fmla, df, "dy", "y", 2, TRUE)
-#' # simulate with errors set to 0's
-#' yhat <- sim(df$id[1], function(n) fill(0, n))
-#' #' # simulate sampling residuals with replacement
-#' yhat <- sim(df$id[1])
 createforwardsim <- function(m, df, id="state",
                        dvar="dlogdc", var="logdc",
                        difflag=7, lhsdiff=TRUE) {
+  stopifnot(m$keepX)
   m$coef[is.nan(m$coef)] <- 0
 
   # get index of observations used in mdl
@@ -101,9 +97,7 @@ createforwardsim <- function(m, df, id="state",
     stopifnot(all(is.pconsecutive(df)))
   }
 
-  sim <- function(idvar, shockgen=NULL, #shockgen=function(n=1) { sample(m$residuals, size=n,
-                         #                                replace=TRUE)},
-                  samplecoefs=FALSE, nsim=1) {
+  sim <- function(idvar, samplecoefs=FALSE) {
     ids <- df[idx,id]
     if (samplecoefs) {
       V <- vcov(m)
@@ -112,23 +106,29 @@ createforwardsim <- function(m, df, id="state",
       Vnona <- V[!nac,!nac]
[…]

### 2020-09-01 tex/covidEffectOfNPI-revision.tex
Corrected counterfactual calculations.
Added appendix section on counterfactual calculation.
Updated counterfactual section of paper. (Switched from emphasizing relative weekly effects to relative cumulative effects).

@@
 \begin{figure}[ht]
-  \caption{Effect  of nationally mandating masks for employees on April
-    1st in the US \label{fig:US-mask-dgrowth}}
+  \caption{Effect  of nationally mandating masks for employees on March
+    14th in the US \label{fig:US-mask-dgrowth}}
   \begin{minipage}{\linewidth}
@@
  %  \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_idx}
-      & \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_idx}
+      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rcumu_idx}
+      & \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rcumu_deaths_idx}
     \end{tabular}
@@
       \footnotesize This figure shows the estimated relative change in
-      cases if all states had mandated masks on April 1st. The thick
-      blue line is the estimated change in weekly cases or deaths
-      relative to the observed number of weekly cases or deaths. The
-      shaded region is a point wise 90\% confidence band.
+      cumulative cases and deaths if all states had mandated masks on
+      March 14th. The thick blue line is the estimated change in
+      cumulative cases or deaths relative to the observed number of
+      cases or deaths. The shaded region is a point wise 90\%
+      confidence band.
     \end{flushleft}
@@
 %
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_idx}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rcumu_idx}
+      & \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rcumu_deaths_idx}
 %      \\
@@
       \footnotesize This figure shows the estimated relative change in
-      cases if no states had never implemented any business closure
-      policies. The thick blue line is the estimated change in weekly
-      cases or deaths relative to the observed number of weekly cases
+      cases and deaths if no states had never implemented any business closure
+      policies. The thick blue line is the estimated change in cumulative
+      cases or deaths relative to the observed number of cumulative cases
       or deaths. The shaded region is a point wise 90\% confidence
@@
  % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth_idx}}
+  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth}}
   \begin{minipage}{\linewidth}
@@
     %  &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_v1}
+         \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rcumu_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rcumu_deaths_idx}
     \end{tabular}
@@
       \footnotesize This figure shows the estimated relative change in
-      cases if no states had ever issued stay at home orders. The thick
-      blue line is the estimated change in weekly cases or deaths
-      relative to the observed number of weekly cases or deaths. The
-      shaded region is a point wise 90\% confidence band.
+      cases and deaths if no states had ever issued stay at home
+      orders. The thick blue line is the estimated change in cumulative
+      cases or deaths relative to the observed number of cumulative cases
+      or deaths. The shaded region is a point wise 90\% confidence
+      band.
     \end{flushleft}
@@
 business policies only indirect effects.\footnote{Results (not shown)
-  using the unconstrained estimates in Table \ref{tab:dieff-si} are similar.}
+  using the unconstrained estimates in Table \ref{tab:dieff-si} are
+  very similar.} Details of the counterfactual computations are in
+appendix \ref{app:counterfactuals}.
 %In the appendix, we also report additional counterfactual
@@
     \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      %\includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth}
-      % &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dgrowth_idx}
-      &        \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dcases_idx}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}}
-      \\
-      % \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-growth_deaths} &
-                                                                                                   \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_idx}
-      &      \includegraphics[width=0.49\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_idx}
+    \begin{tabular}{ccc}
+      \multicolumn{3}{c}{\textbf{Cases}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-rcumu_idx}
+      \\
+      \\
+      \multicolumn{3}{c}{\textbf{Deaths}}
+      \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-dcases_deaths_idx}
+      & \includegraphics[width=0.31\textwidth]{tables_and_figures/Washington-mask-rcumu_deaths_idx}
     \end{tabular}
@@
       observed values. The shaded region is a point-wise 90\%
-      confidence interval.
+      confidence interval. See appendix \ref{app:counterfactuals} for
+      more details.
     \end{flushleft}
@@
 We first consider the impact of a nationwide mask mandate for
-employees beginning on April 1st. As discussed earlier, we find that
+employees beginning on March 14th. As discussed earlier, we find that
 mask mandates reduce case and death growth even when holding behavior
@@
 particularly attractive policy instrument. In this section we examine
-what would have happened to the number of cases if all states had imposed a mask
-mandate on April 1st.\footnote{We feel this is a very plausible
-  counterfactual policy. In a paper made publicly available on
-  April 1st, \cite{abaluck2020} argued for mask usage based on
-  comparisons between countries with and without pre-existing norms of
-  widespread mask usage.}
+what would have happened to the number of cases if all states had
+imposed a mask mandate on March 14th.\footnote{We feel this is a
+  plausible counterfactual policy. Many states began implement
+  business restrictions and school closures in mid March. In a paper
+  made publicly available on April 1st, \cite{abaluck2020} argued for
+  mask usage based on comparisons between countries with and without
+  pre-existing norms of widespread mask usage.}
 
@@
 The left column of Figure \ref{fig:WA-mask} shows the change in
-growth rate from mandating masks on April 1st. The shaded region is a
+growth rate from mandating masks on March 14th. The shaded region is a
 90\% pointwise confidence interval. As shown, mandating masks on April
@@
 less cautious behavior, attenuating the impact of the policy. The
-reversal of the decrease in growth in late April is due to our
-comparison of a mask mandate on April 1st with Washington's actual
-mask mandate in early May. By late April, the counterfactual mask
+reversal of the decrease in growth in late May is due to our
+comparison of a mask mandate on March 14th with Washington's actual
+mask mandate in early May. By late May, the counterfactual mask
 effect has decayed through information feedback, and we are comparing
@@
 
-The right column of Figure \ref{fig:WA-mask} shows how the changes in
-case and death growth translate into changes in weekly cases and
-deaths. The estimates imply that mandating masks on April 1st would
-have led to about 150 fewer cases and 20 fewer deaths per week in
-Washington by the start of June.
+The middle column of Figure \ref{fig:WA-mask} shows how the changes in
+case and death growth translate into changes in daily cases and
+deaths. The estimates imply that mandating masks on March 14th would
+have led to about 70 fewer cases and 5 fewer deaths per day throughout
+April and May. Cumulatively, this implies 19\% [9\%,29\%]fewer cases and 26\%
+[14\%,37\%] fewer deaths in Washington by the start of June.
 
 \begin{figure}[ht]
-  \caption{Effect of nationally mandating masks for employees on April
-    1st in the US\label{fig:US-mask}}
+  \caption{Effect of nationally mandating masks for employees on March
+    14th in the US\label{fig:US-mask}}
   \begin{minipage}{\linewidth}
@@
       \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_idx}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_idx}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-dgrowth_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-rcumu_idx}
       \\
@@
       \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_idx}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths_idx}
+      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-rcumu_deaths_idx}
     \end{tabular}
@@
       point-wise 90\% confidence interval for the national average
-      change in growth. The right column shows the change in cases or
-      deaths relative to the baseline of actual policies.
+      change in growth. The right column shows the change in
+      cumulative cases or deaths relative to the baseline of actual policies.
     \end{flushleft}
@@
 Figure \ref{fig:US-mask} shows the national effect of mask mandate on
-April 1st. The top panel shows the effect on cases and the bottom
+March 14th. The top panel shows the effect on cases and the bottom
 panel shows the effect on deaths. The left column shows the change in
@@
 national average change in growth rate. The right column of Figure
-\ref{fig:US-mask} shows the national relative change in weekly cases
-or deaths. The point estimates indicate that mandating masks on April
-1st could have led to 20\% fewer weekly cases and 35\% fewer
-weekly deaths by the end of May with their 90 percent intervals
-given by $[8,32]$\% and $[18,50]$\%, respectively.
-
-%The result roughly
-%translates into $30$ thousand saved lives.
+\ref{fig:US-mask} shows the national relative change in cumulative cases
+or deaths. The point estimates indicate that mandating masks on March
+14th could have led to 21\% fewer cumulative cases and 34\% fewer
+deaths by the end of May with their 90 percent intervals
+given by $[9,32]$\% and $[19,47]$\%, respectively.
+The result roughly translates into $30$ thousand saved lives.
 
@@
 businesses on cases and deaths. The point estimates imply that,
-without business closures, cases and deaths would be about 75\% higher
+without business closures, cases and deaths would be about 40\% higher
 at the end of May. The confidence intervals are wide. A 90 percent
-confidence interval for the increase in cases at the end of May is 25\%
-to 140\%. The confidence interval for deaths is even wider, $[10, 250]$\%
+confidence interval for the increase in cases at the end of May is
+$[17,78]$\%.
+The confidence interval for deaths is even wider, $[1, 97]$\%
 
@@
       &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_idx}
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rcumu_idx}
       \\
@@
       &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rel_deaths_idx}
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rcumu_deaths_idx}
     \end{tabular}
@@
 stay-at-home orders. On average, without stay-at-home orders, case
-growth rate would have been nearly 0.1 higher in late April. This
-translates to 60\% [10\%,140\%] more cases per week by the start of June. The
-results for deaths are similar, but slightly less precise, with no
-increase included in a 90 percent confidence interval.
+growth rate would have been nearly 0.075 higher in late April. This
+translates to 37\% $[6,63]$\% more cases per week by the start of June. The
+estimated effect for deaths is a bit smaller, with no
+increase included in a 90 percent confidence interval, $[-7,50]$\%.
 
@@
       &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_idx}
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rcumu_idx}
       \\
@@
       &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths_idx}
+        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rcumu_deaths_idx}
     \end{tabular}
@@
       point-wise 90\% confidence interval for the national average
-      change in growth. The right column shows the change in cases or
+      change in growth. The right column shows the change in
+      cumulative cases or
       deaths relative to the baseline of actual policies.
@@
 \end{figure}
-
-%\section{Counterfactual Effect of Removing All Policies and its Sensitivity}\label{removing-policies}
-
-%We now consider the impact of changing from the observed policies to
-%none.  Figure \ref{fig:WA-nb} shows the effect of removing all
-%policies in Washington. The left column shows that without any policy
-%intervention, case and death growth would have been higher by
-%about 0.1 and 0.2. The confidence intervals are wide relative to the
-%mean effect. In the right column, we see that the upper bounds of the
-%confidence intervals include large increases in cases, and especially
-%deaths.
-
-%\begin{figure}[ht]
-%  \caption{Effect of removing all policies in
-%    Washington \label{fig:WA-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dgrowth}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dcases}
-%      \\
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dgrowth_deaths}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/Washington-nop-dcases_deaths}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-
-% We now consider the impact of changing from the observed policies to none.
-% Figure \ref{fig:US-nop-SI} shows the average across states of the change
-% in case growth and relative increase in cases under a specification without past national case variables.
-% Removing policies leads to an increase of above 0.2 in
-% case growth throughout April and May. The confidence interval is
-% fairly wide, and its upper bound includes a very large increase in
-% cases by the end of May.  The right panel displays the national
-% increase in aggregate cases without any policy intervention. The
-% estimates imply at least a 7 fold increase in cases with a large upper bound by the end of May, or
-% at least 14 million additional cases. The estimated impact on deaths is larger than
-% cases, and even more imprecise.
-
-% [Discussion on the effect of school closures has been moved to
-% sensitivity analysis section.]
-
-% [Do we even still need this section?]
-
-%The effect of removing all policies includes the effect of school
-%closures. The visual evidence on growth rates for states with and without school closures,
-%presented blow, suggest that there may be a potentially large effect, though the history is very short.
-The main results presented in Section 3 also support the hypothesis that the school closures were important at lowering the growth rates.
-This evidence is consistent
-%with the emerging evidence of prevalence of Covid-19 among children
-%aged 10-17. \cite{children:nature} find that although children's
-%transmission and susceptibility rates are half that of ages 20-30,
-%children's contact rates are much higher.
-%This type of evidence, as well as, evidence that children carry viral
-%loads similar to older people (\cite{children:germany}), led Germany to make the early decision of closing schools.
-%
-%\begin{figure}
-%  \caption{Case and death growth conditional on policies \label{fig:growthpolicies1}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%       \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-cases}
-%      &
-%        \includegraphics[width=0.483\textwidth]{tables_and_figures/pk12-deaths}
-%    \end{tabular}
-%    \footnotetext{ In these figures, red points are the case or death
-%      growth rate in states without each policy 14 (or 21 for deaths)
-%      days earlier. Blue points are states with each policy 14 (or 21
-%      for deaths) days earlier. The red line is the average across
-%      states without each policy. The blue line is the average across
-%      states with each policy.}
-%  \end{minipage}
-%\end{figure}
-
-%As discussed above, there is little variation across states
-%in the timing of school closures. Consequently, the effect of school
-%closures is difficult to identify statistically, because it is hard to separate it from aggregate time effect, and its estimate is sensitive to an inclusion of some aggregate variables such as national cases.
-To support this point, Figure \ref{fig:US-nop} shows the
-%effect of removing all policies on cases based on the estimates with national cases
-%included as information. When national case variables are included
-%in the specification, the estimated effect of school closures, and hence that of removing all
-%policies, is much smaller with a 90\% confidence interval of [0,10] fold increases.
-%
-%Given this sensitivity, we conclude that there still exists a lot of uncertainty as to the effect of removing all policies, especially schooling.
-The impact of not implementing any policies on cases and deaths can be quite large, but  the effect of school closures, hence that of removing all policies, is not well identified statistically from the US state-level data alone, because of the lack of cross-sectional variations.
-Any analyses of re-opening plans need to be aware of this uncertainty.
-An important research question is how to resolve this uncertainty using additional data sources.
-%
-%\begin{figure}[ht]
-%  \caption{Effect of removing policies on cases in the US under a specification with only state-level cases/deaths as information \label{fig:US-nop-SI}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_v1}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_v1}
-%%      \\
-%%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths_v1}
-%%      &
-%%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths_v1}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-%
-%\begin{figure}[ht]
-%  \caption{Effect of removing policies on cases in the US under a specification with both
-%  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nop}}
-%  \begin{minipage}{\linewidth}
-%    \centering
-%    \begin{tabular}{cc}
-%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth}
-%      &
-%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel}
-%%      \\
-%%      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-dgrowth_deaths}
-%%      &
-%%        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nop-rel_deaths}
-%    \end{tabular}
-%  \end{minipage}
-%\end{figure}
-
-%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
 
@@
 
+\subsection{Counterfactuals \label{app:counterfactuals}}
+
+We compute counterfactuals from the "total effect" version of the
+model, with behavior concentrated out.
+
+$$
+Y_{i,t+\ell}
+= \mathsf{a}'
+P_{it} + \mathsf{b}' I_{it} +  {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
+$$
+
+We consider a counterfactual change of $P_{it}$ to $P_{it}^*$, while
+$W_{it}$ and $\bar \varepsilon_{it}$ are held constant. In response to
+the policy change, $Y_{i,t+\ell}$ and the part of $I_{it}$ that
+contains $Y_{it}$, change to $Y_{i,t+\ell}^*$ and $I_{it}^*$. To be
+specific, let $C_{it}$ denote cumulative cases in state $i$ on day
+$t$. Our outcome is:
+
+$$
+Y_{i,t+\ell} \equiv \log
+\left(C_{i, t+\ell} - C_{i, t + \ell - 7} \right) - \log
+\left(C_{i, t+\ell-7} - C_{i, t + \ell - 14} \right) \equiv \Delta
+\log \Delta C_{i,t+\ell}
+$$
+
+where $\Delta$ is a 7 day difference operator. Writing the model in
+terms of $\Delta \log \Delta C$ we have:
+
+$$
+\Delta \log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \mathsf{b}_D
+\Delta \log \Delta C_{i,t} + \mathsf{b}_L \log \Delta C_{i,t} +
+{\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
+$$
+
+To simplify computation, we rewrite this model in terms of $\log
+\Delta C$:
+
+$$
+\log \Delta C_{i,t+\ell} = \mathsf{a}' P_{it} + \log \Delta
+C_{i,t+\ell-7} + (\mathsf{b}_D + \mathsf{b}_L) \log \Delta C_{i,t} -
+\mathsf{b}_D \log \Delta C_{i,t-7} + {\tilde {\delta}}' W_{it}   + {\bar \varepsilon}_{it}
+$$
+
+This equation is used to iteratively compute
+$\log \Delta C_{i,t+\ell}^*$
+conditional on initial $\{\log \Delta C_{i,s}\}_{s=-\ell-7}^{0}$,
+and the entire sequence of $\{W_{i,t}, P_{i,t}^*, \bar \varepsilon_{it} \}_{t=0}^T$.
+
+Weekly cases instead of log weekly cases are given simply by
+
+$$ \Delta C_{i,t+\ell}^* = \exp(\log \Delta C_{i,t+\ell}^*)$$
+
+Cumulative cases can be recursively computed as:
+
+$$
+  C_{i,t} = C_{i,t-7} + \Delta C_{i,t} = C_{i,t-7} + \exp(\log \Delta C_{i,t})
+$$
+
+given initial $\{C_{i,s}\}_{s=-7}^0$.
+
+Note that $\log \Delta C_{i,t+\ell}^*$ depends linearly on
+$\bar \varepsilon$, so the residuals (and our decision to condition on
+versus integrate them out) do not matter when considering linear
+constrasts of $\log \Delta C_{it}$ or $\Delta \log \Delta C_{it}$, or
+when considering relative constrasts of $\Delta C_{it}$.
+
+%On the other hand, cumulative cases are a nonlinear function of
+%$\bar\varepsilon$, so conditioning on $\bar\varepsilon$
+
+\subsubsection*{Inference}
+
+Let $S(\theta, \mathbf{\varepsilon})$ denote some counterfactual
+quantity or contrast of interest, where $\theta = (\mathsf{a},
+\mathsf{b}, \tilde{\delta})$ are the parameters, and
+$\mathbf{\varepsilon}$ is the vector of residuals. Examples of $S$
+that we compute include:
+
+\begin{itemize}
+\item Contrasts of growth rates: $S(\theta, \mathbf{\varepsilon}) =
+  \Delta \log \Delta C_{it}^* - \Delta \log \Delta C_{it}$
+
+\item Relative contrasts of weekly cases: $S(\theta, \mathbf{\varepsilon}) =
+  \Delta C_{it}^*/\Delta C_{it}$
+
+\item Relative contrasts of cumulative cases: $S(\theta, \mathbf{\varepsilon}) =
+  \frac{C_{it}^* - C_{it}}{C_{it}}$
+\end{itemize}
+
+The first two examples do not actually depend on
+$\mathbf{\varepsilon}$, but the third one does. Inference is by
+simulation. Let $\hat{\theta}$ denote our point estimates and
+$\hat{\mathbf{\varepsilon}}$ the associated residuals. We draw
+$\tilde{\theta}_j$ from the asymptotic distribution of
+$\hat{\theta}$. Let $\tilde{\varepsilon}_j$ denote the
+residuals associated with $\tilde{\theta}_j$. We then compute
+
+$$ \tilde{s}_j = S(\tilde{\theta}_j, \tilde{\varepsilon}_j) $$
+
+for $j=1,..., 200$ and plots the mean across $j$ as a point estimate
+and quantiles across $j$ for confidence intervals.
+
 \end{document}
-
-% \subsection{Counterfactuals for Massachusetts and
-%   Illinois}
-%  Figures \ref{fig:MA-mask} and \ref{fig:IL-mask} present the fit of estimated cases as well as the counterfactual effect of mandating masks on April 1st in Massachusetts and Illinois, respectively.
-Figures \ref{fig:MA-nb} and \ref{fig:IL-nb}   show the counterfactual effect of leaving non-essential business open in Massachusetts and Illinois, respectively.
-
-% \FloatBarrier
-
-% \begin{figure}[b]
-%   \caption{Effect of mandating masks on April 1st in Massachusetts \label{fig:MA-mask}}
-%   \begin{minipage}{\linewidth}
-%     \centering
-%     \begin{tabular}{ccc}
-%       \multicolumn{3}{c}{\textbf{Cases}} \\
-%       \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth}
-%       &
-%         \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_v1}
-%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_v1}
-% \\
-%       \multicolumn{3}{c}{\textbf{Deaths}} \\
-%      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-growth_deaths}
-%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dgrowth_deaths_v1}
-%       &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Massachusetts-mask-dcases_deaths_v1}
-%     \end{tabular}
-%     \begin{flushleft}
-%       \footnotesize To compute the estimated and counterfactual paths
-%       we use the average of two estimated coefficients as reported in
-%       column ``Average'' of Table \ref{tab:dieff}. We set initial
-%       $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-%       observed in the state we are simulating. We hold all other
-%       regressors at their observed values. Error terms are drawn with
-%       replacement from the residuals. We do this many times and report
-%       the average over draws of the residuals. The shaded region is a
-%       point-wise 90\% confidence interval.
-%     \end{flushleft}
-%   \end{minipage}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \caption{Effect of mandating masks on April 1st in Illinois \label{fig:IL-mask}}
-%   \begin{minipage}{\linewidth}
-%     \centering
-%     \begin{tabular}{ccc}
-%       \multicolumn{3}{c}{\textbf{Cases}} \\
-%       \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth}
-%       &
-%         \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_v1}
-%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_v1}      \\
-%       \multicolumn{3}{c}{\textbf{Deaths}} \\
-%     \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-growth_deaths}
-%       &        \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dgrowth_deaths_v1}
-%       &      \includegraphics[width=0.31\textwidth]{tables_and_figures/Illinois-mask-dcases_deaths_v1}
-%     \end{tabular}
-%     \begin{flushleft}
-%       \footnotesize To compute the estimated and counterfactual paths
-%       we use the average of two estimated coefficients as reported in
-%       column ``Average'' of Table \ref{tab:dieff}. We set initial
-%       $\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-%       observed in the state we are simulating. We hold all other
-%       regressors at their observed values. Error terms are drawn with
-%       replacement from the residuals. We do this many times and report
-%       the average over draws of the residuals. The shaded region is a
-%       point-wise 90\% confidence interval.
-%     \end{flushleft}
-%   \end{minipage}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \caption{Effect of leaving businesses open in Massachusetts \label{fig:MA-nb}}
-%   \begin{minipage}{\linewidth}
-%     \centering
-%     \begin{tabular}{cc}
-%       \multicolumn{2}{c}{\textbf{Cases}} \\
-%       \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_v1}
-%        &
-%          \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_v1}
-%       \\
-%       \multicolumn{2}{c}{\textbf{Deaths}} \\
-%         \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dgrowth_deaths_v1}
-%        &      \includegraphics[width=0.45\textwidth]{tables_and_figures/Massachusetts-nb-dcases_deaths_v1}
-%     \end{tabular}
-%   \end{minipage}
-% \end{figure}
-
-% \begin{figure}[h]
-%   \caption{Effect of leaving businesses open in Illinois \label{fig:IL-nb}}
-%   \begin{minipage}{\linewidth}
-%     \centering
-%     \begin{tabular}{cc}
-%       \multicolumn{2}{c}{\textbf{Cases}} \\
-%       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_v1}
-%       &
-%         \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_v1} \\
-%       \multicolumn{2}{c}{\textbf{Deaths}} \\
-%            \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dgrowth_deaths_v1}
-%       &       \includegraphics[width=0.45\textwidth]{tables_and_figures/Illinois-nb-dcases_deaths_v1}    \end{tabular}
-%   \end{minipage}
-% \end{figure}
-
-\FloatBarrier
-
-\subsection{Counterfactuals with National Cases as Information
-  Variables}
-
-\fixme
-% DO WE STILL WANT THESE? THESE FIGURES HAVE NOT BEEN UPDATED.
-
-Figures \ref{fig:US-mask-SI}-\ref{fig:US-shelter-SI}  present the results of counterfactual analyses that include the national cases/deaths as the information variables.
-To create this figure, we repeat the same counterfactual
-simulation that we did for Washington with each state. For each state,
-we hold national cases constant, but endogenize state specific
-information. Thus, these figures should be interpreted as an average
-of state specific counterfactuals, and not a national counterfactual.
-
- The counterfactual results of mask policies, shelter-in-place,
- and closing non-essential businesses remain robust with respect to the inclusion of national case/death variables.
-This contrasts to the resulting counterfactual of  removing all policies  discussed in section \ref{removing-policies}.
-
-\begin{figure}[!b]
-  \caption{Effect of mandating masks for employees on April
-    1st
-     under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information
-    \label{fig:US-mask-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \medskip
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-dgrowth_deaths}
-      &
-        \includegraphics[width=0.49\textwidth]{tables_and_figures/us-mask-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
-  \caption{Effect of leaving non-essential businesses open      under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-nb-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-dgrowth_deaths}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-nb-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\begin{figure}[ht]
-  \caption{Effect of having no stay-at-home orders     under a specification with both
-  state-level cases/deaths and national-level cases/deaths as information \label{fig:US-shelter-SI}}
-  \begin{minipage}{\linewidth}
-    \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rel_deaths}
-    \end{tabular}
-  \end{minipage}
-\end{figure}
-
-\FloatBarrier
-
-$\;$
-
-\end{document}

### 2020-09-01 cases_and_policies/rmd/counterfactuals.R

@@ -178,7 +178,7 @@ meanandci <- function(yi, cfyi, date, p=0.9, difflag=7) {
         statframe(dy,"dy"), # growth
         statframe(exp(cfyi),"cf"), # cf weekly cases
         statframe(dcfy,"dcf"), # cf growth
-        statframe(exp(cfyi)-exp(yi),"p"), # change in weekly cases
+        statframe((exp(cfyi)-exp(yi)),"p"), # change in weekly cases
         statframe(dcfy-dy,"dp"), # change in growth
         statframe(exp(cfyi)/exp(yi) - 1, "rel"), # relative weekly
         statframe(apply(exp(yi), 2, cumcases), "cum"), # cumulative cases
@@ -261,12 +261,12 @@ cfplots <- function(st, cfdata=NULL, simobs=NULL, simcf=NULL,
     theme(legend.position=c(0.2, 0.8)) +
     ggtitle(TeX(sprintf("%s in past week",yvar))) + xlim(xlims)
 
-  ylo <- min(est$clp)
-  figp <- ggplot(data=est, aes(x=date, y=mp)) +
+  ylo <- min(est$clp/7)
+  figp <- ggplot(data=est, aes(x=date, y=mp/7)) +
     geom_line(size=1.2) +
-    geom_ribbon(aes(ymin=clp, ymax=chp), alpha=0.3) +
-    ylab(sprintf("Change in %s in past week", yvar)) + figtheme + colors() + colors_fill() +
-    ggtitle(TeX(sprintf("Change in %s",yvar))) +
+    geom_ribbon(aes(ymin=clp/7, ymax=chp/7), alpha=0.3) +
+    ylab(sprintf("Change in %s per day", yvar)) + figtheme + colors() + colors_fill() +
+    ggtitle(TeX(sprintf("Change in daily %s",yvar))) +
     xlim(xlims) #+
     #annotate("text",x=dt, y=rep(ylo,length(dt)),
     #         label=gsub("\\."," ", labels), size=4, angle=90, hjust=0)
@@ -329,8 +329,8 @@ cfplots <- function(st, cfdata=NULL, simobs=NULL, simcf=NULL,
   figr <- ggplot(data=est, aes(x=date, y=mrel)) +
     geom_line(size=1.2) +
     geom_ribbon(aes(ymin=clrel, ymax=chrel), alpha=0.3) +
[…]

### 2020-09-01 cases_and_policies/rmd/cf-pindex.Rmd

@@
 
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcumu_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdcum)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-rcumu_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figrcum)
+  dev.off()
+
   if (st=="Washington") for (fig in f) print(fig)
@@
 figs$r
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rcumu_idx.pdf",sep="/"), width=5,height=3.75)
+figs$cr
 dev.off()
@@
   dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcumu_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdcum)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-rcumu_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figrcum)
+  dev.off()
+
 }
@@
 figs$r
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rcumu_idx.pdf",sep="/"), width=5,height=3.75)
+figs$cr
 dev.off()
@@
   dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcumu_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdcum)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-rcumu_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figrcum)
+  dev.off()
+
 }
@@
 figs$r
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-rcumu_idx.pdf",sep="/"), width=5,height=3.75)
+figs$cr
 dev.off()
@@
 rerunsim=FALSE
-if (!file.exists("cfnop_deaths_idx.Rda") || rerunsim) {
-  cfnoplist <- lapply(unique(df$state), function(s)
-    nopstatesim(s,simobs=simmodel,simcf=nopmodel, data=sdf, var="logdd", dvar="dlogdd"))
-  allmeans <- rbindlist(lapply(cfnoplist, getmeans))
-  save(cfnoplist,allmeans, file="cfnop_deaths_idx.Rda")
-}
-load("cfnop_deaths_idx.Rda")
-cfnop <- do.call(rbind, lapply(cfnoplist, function(x) x$edf))
-```
-
-### Removing Policies
-
-```{r WA-nop-dlogdd, cache=TRUE,fig.cap="Case growth without policies in Washington \\label{fig:WA-nop-dlogdc}"}
-st <- "Washington"
-f <- cfplots(st, cfnop, yvar="deaths", data=df)
-grid.arrange(f$figd, f$figdp, nrow=2, top =
-                                        textGrob(st,gp=gpar(fontsize=20,font=2)))
-
-```
-
-```{r WA-nop-deaths, cache=TRUE, fig.cap="Cases without policies in Washington \\label{fig:WA-nop-cases}"}
-grid.arrange(f$figl, f$figp, nrow=2) #,
-#top = textGrob(st,gp=gpar(fontsize=20,font=2)))
-```
-
-## National
-
-```{r nop-all-deaths, cache=TRUE}
-abbrv <- unique(df[,c("state","ST")])
-cfnop <- merge(cfnop, abbrv, by="state")
-
-figs <- nationalplots(allmeans,cfnop, "removing policies", yvar="deaths")
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dgrowth_deaths_idx.pdf",sep="/"), width=5,height=3.75)
-figs$g
-dev.off()
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-dcases_deaths_idx.pdf",sep="/"), width=5,height=3.75)
-figs$c
-dev.off()
-
-figdev(paste(rootdir,"tex/tables_and_figures/us-nop-rel_deaths_idx.pdf",sep="/"), width=5,height=3.75)
-figs$r
-dev.off()
-
-for (f in figs) print(f)
 ```
@@
 We now look at what happens when all states impose a mask mandate on
-May 14th.
+March 14th.
 
@@
   f <- cfplots(st, cfmask, yvar="deaths", data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on May 14th in %s",st)))
+  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on March 14th in %s",st)))
 
@@
   dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-dcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdcum)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-mask-rcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figrcum)
+  dev.off()
+
 }
 
-figs <- nationalplots(maskall, cfmask, "mandating masks on May 14th\n", yvar="deaths")
+figs <- nationalplots(maskall, cfmask, "mandating masks on March 14th\n", yvar="deaths")
 
@@
 figs$r
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-mask-rcumu_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figs$cr
 dev.off()
@@
   dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-dcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdcum)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-shelter-rcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figrcum)
+  dev.off()
+
 }
@@
 figs$r
+dev.off()
+
+figdev(paste(rootdir,"tex/tables_and_figures/us-shelter-rcumu_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figs$cr
 dev.off()
@@
   dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-dcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figdcum)
+  dev.off()
+
+  figdev(sprintf("%s/tex/tables_and_figures/%s-index-rcumu_deaths_idx.pdf",rootdir,st), width=5,height=3.75)
+  print(f$figrcum)
+  dev.off()
+
 }
@@
 
+figdev(paste(rootdir,"tex/tables_and_figures/us-index-rcumu_deaths_idx.pdf",sep="/"), width=5,height=3.75)
+figs$cr
+dev.off()
+
 for (f in figs) print(f)

### 2020-09-01 cases_and_policies/rmd/cf-pindex.Rmd
Fixed titles of mask figures.

@@
   f <- cfplots(st, cfmask, data=df)
-  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on May 14th in %s",st)))
+  print(f$figp +  ggtitle(sprintf("Effect of mandating masks on March 14th in %s",st)))
 
@@
 
-figs <- nationalplots(maskall, cfmask, "mandating masks on May 14th\n")
+figs <- nationalplots(maskall, cfmask, "mandating masks on March 14th\n")
 

### 2020-09-02 tex/covidEffectOfNPI-revision.tex
Updated counterfactual figures.
Updated discussion of counterfactuals in the intro.

@@
 Our analysis finds that both policies and information on transmission risks are important determinants of Covid-19  cases and deaths and shows that a change in policies explains a large fraction of observed changes in social distancing behavior.
-Our main Lcounterfactual experiments suggest that nationally mandating face masks for employees on April 1st could have reduced the growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 17 to 55 percent less deaths nationally by the end of May, which roughly translates into 17 to 55 thousand saved lives.
-We also find that, without stay-at-home orders, cases would have been larger by  25 to 170 percent, which implies that 0.5 to 3.4 million more Americans could have been infected if stay-at-home orders had not been implemented.
+Our main Lcounterfactual experiments suggest that nationally mandating face masks for employees on April 1st could have reduced the growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 19 to 47 percent less deaths nationally by the end of May, which roughly translates into 19 to 47 thousand saved lives.
+We also find that, without stay-at-home orders, cases would have been larger by  6 to 63 percent and without business closures, cases would have been larger by 17 to 78 percent.
 We find considerable uncertainty over the effects of school closures due to lack of cross-sectional variation;  we could not robustly rule out either large or small effects.
@@
 
-%We present a conceptual framework that spells out the causal structure on how the Covid-19 spread  is dynamically determined by policies and human behavior.
-Our approach explicitly recognizes that policies not only directly affect the spread of Covid-19 (e.g., mask requirement) but also indirectly affect its spread by changing people's behavior (e.g., stay-at-home order).
-It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
-Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, tinbergen1940, wold1954} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-See \cite{pearl:biometrika},\cite{pearl:robins},  \cite{pearl:causality}, \cite{white:chalak}, \cite{peters2020book}, \cite{heckman:pinto}, and \cite{hernanrobins2020book}  for modern developments, in computer
-%science, econometrics, and epidemiology.
-The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986} and \cite{Hines2020} for modern treatments.
-The father and son, P.
-Wright and S.
-Wright closely collaborated to develop structural equation models and causal path diagrams; P.
-Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables.}
-%The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo1944, heckman2007}.
-See \cite{pearl:robins}, \cite{peters2020book}, and \cite{hernanrobins2020book} for  modern developments, especially in computer science and epidemiology), with natural unfolding potential outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-As such it naturally converses in all languages for causal inference.
-
-% It also allows us  to explicitly define counterfactual scenarios to properly evaluate the effect of alternative policies on the spread of Covid-19.
-
 Guided by the causal model, our empirical analysis examines how the weekly growth rates of confirmed Covid-19 cases and deaths are determined by (the lags of) policies and behavior using US state-level data.
@@
 
-Our estimates imply that nationally  implementing mandatory
-face masks  for employees in public businesses on April 1st would have reduced the growth rate of cases
-and that of deaths by more than 10 percentage points in late April. As
-shown in Figure \ref{fig:US-mask-dgrowth}, this leads to reductions of
-$20$\% and $35$\% in weekly reported cases and deaths, respectively,
-by the end of May with a 90 percent confidence interval of $[8,32]$\%
-and $[17,50]$\%, which roughly implies that 6 thousand lives could
-have been saved.\footnote{The US averaged roughly one thousand deaths
-  per day during May. Reducing one thousand deaths per day by 20\%
-  over 30 days gives a rough estimate of 6 thousand saved lives.}
-This finding is significant: given this potentially large  benefit of reducing the spread of Covid-19, mandating masks is an attractive policy instrument especially because it involves relatively little economic disruption.
-These estimates contribute to the ongoing efforts towards designing approaches to minimize risks from reopening \citep{stock2020b}.
-
-Figure \ref{fig:US-nb-dgrowth} illustrates how allowing non-essential businesses to remain open could
-have affected the growth of cases.  We estimate that
-non-essential business closures have a small impact on growth rates,
-with a 90\% confidence interval that includes both negative and
-positive effects. When this effect on growth rates is converted to a
-change in levels, the point estimates indicate that keeping
-non-essential businesses open (other than movie theaters, gyms, and
-keeping restaurants in the ``take-out" mode) could have increased
-cases and  deaths by  $15\%$ (with a 90 percent confidence interval of
-$-20\%$ to $60\%$).  These estimates contribute to the ongoing efforts of evaluating various reopening approaches.
-
-In Figure \ref{fig:US-shelter-dgrowth}, we find that, without stay-at-home orders, the case
-growth rate would have been nearly 10 percentage points higher in late April.
-No stay-at-home orders could have led to 80\% more cases by the start of June with a 90 precent confidence interval given by  $25\%$ to $170\%$.
-This implies that 0.5 to 3.4 million more Americans would have been infected without stay-at-home orders, providing suggestive evidence that reopening via removal of stay-at-home orders could lead to a substantial increase in cases and deaths.
-
-In our counterfactual experiment of removing all policies, we find that the results are sensitive to whether  the number of past national cases/deaths is included  in a specification or not.
-This sensitivity arises because there is little variation across states in the timing of school closures.
-This makes the effect of school closures difficult to identify.
-In Figure \ref{fig:US-nop-SI}, we show that in a specification that excludes past national cases (which allow for greater attribution of effects to school closures), the number of cases by the end of May could have increased 7-fold or more with a very large upper bound.
-On the other hand, as shown in Figure \ref{fig:US-nop}, under a specification with past national cases, our counterfactual experiment implies a 0 to 10 fold increase in cases by the end of May.
-This highlights the uncertainty regarding the impact of all policies versus private behavioral responses to information.
-Evaluation of re-opening policies needs to be aware of this uncertainty.
+Our estimates imply that nationally implementing mandatory face masks
+for employees in public businesses on March 14th would have reduced the
+growth rate of cases and that of deaths by approximately 10 percentage
+points in late April. As shown in Figure \ref{fig:US-mask-dgrowth},
+this leads to reductions of $21$\% and $34$\% in cumulative reported
+cases and deaths, respectively, by the end of May with 90 percent
+confidence intervals of $[9,32]$\% and $[19,47]$\%, which roughly
+implies that 34 thousand lives could have been saved.  This finding is
+significant: given this potentially large benefit of reducing the
+spread of Covid-19, mandating masks is an attractive policy instrument
+especially because it involves relatively little economic
+disruption. These estimates contribute to the ongoing efforts towards
+designing approaches to minimize risks from reopening
+\citep{stock2020b}.
+
+Figure \ref{fig:US-nb-dgrowth} illustrates how never closing any
+businesses (no movie theater closure, no non-essential business
+closure, and no closure of restaurants except take-out)) could have
+affected cases and deaths.  We estimate that business shutdowns have a
+roughly the same impact on growth rates as mask mandates, albeit with
+more uncertainty. The point estimates indicate that keeping all
+businesses open could have increased cumulative cases and deaths by
+$40\%$ at the end of May (with 90 percent confidence intervals of
+$[17,78]$\% for cases and $[1,97]$\% for deaths).
+
+Figure \ref{fig:US-shelter-dgrowth} shows that stay-at-home orders had
+effects of similar magnitude as business closures.  No stay-at-home
+orders could have led to $37$\% more cases by the start of June with a
+90 percent confidence interval given by $6$\% to $63$\%. The estimated
+effect of no stay-at-home orders on deaths is a slightly smaller with
+a 90 percent confidence interval of $-7$\% to $50$\%.  These estimates
+contribute to the ongoing efforts of evaluating various reopening
+approaches.
+
+The above findings are robust to changes to our regression specification, sample selection, methodology, and assumptions about delays between policy changes and changes in recorded cases.
+In all cases, we find that mask mandates, business closure policies, and stay-at-home orders all reduce growth rates by roughly 0.1.
+The impact of mask mandates are more precisely estimated than that of business closure policies or stay-at-home orders.
+This is largely due to the greater variation in the timing of mask mandates across states.
+
+We also examine the impact of school closures. Unfortunately, there is
+very little variation across states in school closure in the timing of
+school closures.  Across robustness specifications, we obtain point
+estimates of the effect of school closures as low as 0, and as high as
+0.6. In particular, we find that the results are sensitive to whether
+the number of past national cases/deaths is included in a
+specification or not. This highlights the uncertainty regarding the
+impact some policies versus private behavioral responses to
+information. Evaluation of re-opening policies needs to be aware of
+this uncertainty.
 
@@
 have led to about 70 fewer cases and 5 fewer deaths per day throughout
-April and May. Cumulatively, this implies 19\% [9\%,29\%]fewer cases and 26\%
-[14\%,37\%] fewer deaths in Washington by the start of June.
+April and May. Cumulatively, this implies 19\% $[9,29]$\% fewer cases and 26\%
+$[14,37]$\% fewer deaths in Washington by the start of June.
 
@@
     \medskip
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-dgrowth_idx}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-rcumu_idx}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_idx}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-mask-rcumu_deaths_idx}
+    \begin{tabular}{ccc}
+      \multicolumn{3}{c}{\textbf{Cases}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-mask-dgrowth_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-mask-dcases_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-mask-rcumu_idx}
+      \\
+      \\
+      \multicolumn{3}{c}{\textbf{Deaths}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-mask-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-mask-dcases_deaths_idx}
+     &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-mask-rcumu_deaths_idx}
     \end{tabular}
@@
       point-wise 90\% confidence interval for the national average
-      change in growth. The right column shows the change in
-      cumulative cases or deaths relative to the baseline of actual policies.
+      change in growth.  The middle column shows the total national
+      change in daily cases and deaths. Black dots are the changes in each
+      state.  The right column shows the change in cumulative cases or
+      deaths relative to the baseline of actual policies.
     \end{flushleft}
@@
 (i.e. the dots are identical to the solid line in the left panel of
-Figure \label{fig:WA-mask}, except for all states instead of just
+Figure \label{fig:WA-mask}, except for each state instead of just
 Washington). The solid blue is the national average change in growth
 rate. The shaded region is a 90\% pointwise confidence band for the
-national average change in growth rate. The right column of Figure
-\ref{fig:US-mask} shows the national relative change in cumulative cases
+national average change in growth rate.
+
+The middle column shows the changes in cases and deaths per day. Dots
+are the expected change in each state. The thick line is the national
+total change in daily cases or deaths. The shaded region is a
+pointwise 90\% confidence band for the national total change. The
+estimates that a mask mandate in mid March would have decreases cases
+and deaths by about 6000 and 600 respectively throughout April and
+May.
+
+The right column of Figure \ref{fig:US-mask} shows how these daily
+changes compare to the total cumulative cases and deaths. The right
+column shows the national relative change in cumulative cases
 or deaths. The point estimates indicate that mandating masks on March
@@
     \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-dgrowth_idx}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rcumu_idx}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-dgrowth_deaths_idx}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-index-rcumu_deaths_idx}
+    \begin{tabular}{ccc}
+      \multicolumn{3}{c}{\textbf{Cases}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-index-dgrowth_idx}
+      &
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-index-dcases_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-index-rcumu_idx}
+      \\
+      \\
+      \multicolumn{3}{c}{\textbf{Deaths}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-index-dgrowth_deaths_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-index-dcases_deaths_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-index-rcumu_deaths_idx}
     \end{tabular}
@@
     \centering
-    \begin{tabular}{cc}
-      \multicolumn{2}{c}{\textbf{Cases}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_idx}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rcumu_idx}
-      \\
-      \\
-      \multicolumn{2}{c}{\textbf{Deaths}} \\
-      \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths_idx}
-      &
-        \includegraphics[width=0.45\textwidth]{tables_and_figures/us-shelter-rcumu_deaths_idx}
+    \begin{tabular}{ccc}
+      \multicolumn{3}{c}{\textbf{Cases}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-shelter-dgrowth_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-shelter-dcases_idx}
+      &
+
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-shelter-rcumu_idx}
+      \\
+      \\
+      \multicolumn{3}{c}{\textbf{Deaths}} \\
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-shelter-dgrowth_deaths_idx}
+      &
+      \includegraphics[width=0.31\textwidth]{tables_and_figures/us-shelter-dcases_deaths_idx}
+      &
+        \includegraphics[width=0.31\textwidth]{tables_and_figures/us-shelter-rcumu_deaths_idx}
     \end{tabular}
@@
       point-wise 90\% confidence interval for the national average
-      change in growth. The right column shows the change in
-      cumulative cases or
+      change in growth.  The middle column shows the total national
+      change in daily cases and deaths. Black dots are the changes in each
+      state.  The right column shows the change in cumulative cases or
       deaths relative to the baseline of actual policies.

### 2020-09-03 tex/covidEffectOfNPI-revision.tex
updated CF discussion

@@
 Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.
-\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto} and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-See, e.g., \cite{pearl:biometrika},\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak}, \cite{peters2020book}, \cite{heckman:pinto}, and \cite{hernanrobins2020book} for modern developments, in computer science, econometrics, and epidemiology.
+\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto}) and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+The work on causal graphs has been
+modernized and put on rigorous footing by  \cite{pearl:biometrika},\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak}, \cite{peters2020book}, and \cite{hernanrobins2020book}, with major applications taking place in
+in computer science, econometrics, and epidemiology.
 The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986},\cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
@@
 Analyzing mitigation measures in New York, Wuhan, and Italy, \cite{zhangr2020} conclude that mandatory face coverings substantially reduced infections.
-\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} Our finding is also corroborated by a completely different causal methodology based on synthetic control
+\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} Our finding is also independently corroborated by a completely different causal methodology based on synthetic control
 using German data in \cite{Mitze2020}.\footnote{Our study was first released in ArXiv on May 28, 2020 whereas
@@
 \end{tikzpicture} \end{center}
-\caption{ Dynamic System Induced by Information Structure and SEM}\label{fig:DynS}
+\caption{ Diagram for Information Dynamics in SEM}\label{fig:DynS}
 \end{figure}
@@
 & {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
-& {\icolor I^\star_{it} }& := &  I_{it}({\icolor I^\star_{i,t-\ell}}, {\ycolor Y^*_{it}}, t),  \end{aligned}
+& {\icolor I^\star_{it} }& := &  I ({ t,  {\ycolor Y^*_{it}}, \icolor I^\star_{i,t-\ell}}),  \end{aligned}
 \end{equation}
@@
 \end{quote}
+
+Figures \ref{Wright2} and \ref{fig:DynS2} present the causal path diagram for CF-SEM as well as the dynamics
+of counterfactual information in CF-SEM.
+
+\begin{figure}[ht]
+\begin{center}
+\begin {tikzpicture}[-latex, auto, node distance =1.5cm and 3cm, on grid, thick,
+  empty/.style ={circle, top color=white, bottom color = white, draw, white, text=white , minimum width =1.25 cm},
+  policy/.style ={circle, top color=white, bottom color = blue, draw, black, text=black , minimum width =1.25 cm},
+    behavior/.style ={circle, top color=white, bottom color = ForestGreen, draw, black, text=black , minimum width =1.25 cm},
+  observed/.style ={circle, top color=white, bottom color = magenta, draw, black, text=black , minimum width =1.25 cm} ,
+   confounder/.style ={circle, top color=white, bottom color =  gray, draw, black, text=black , minimum width =1.25 cm} ,
+outcome/.style ={circle, top color=white, bottom color = red, draw, black, text=black , minimum width =1.25 cm} ]
+
+\node[policy]   (P) {\tiny $P^\star_{it}$};
+\node[empty] (E) [below=of P] {\tiny $I^\star_{it}$};
+\node[outcome]  (Y)[ right =of E] {\tiny $Y^\star_{i,t+\ell}$};
+\node[behavior] (B) [below =of E] {\tiny $B^\star_{it}$};
+\node[observed] (I) [left =of P] {\tiny $I^\star_{it}$};
+\node[confounder] (W) [left=of B] {\tiny  $W_{it}$};
+\path[->] (P) edge (Y);
+\path[->] (B) edge (Y);
+\path[->] (P) edge (B);
+\path[->] (I) edge (B);
+%\path[->] (I) edge (P);
+\path[->] (W) edge (Y);
+\path[->] (W) edge (B);
+%\path[->] (W) edge (P);
+\path[->] (W) edge (I);
+\path[->] (I) edge (Y);
+
+\end{tikzpicture} \end{center}
+\caption{Causal path diagram for CF-SEM. }\label{Wright2}
+\end{figure}
+
+\begin{figure}[ht]
+\begin{center}
+\begin {tikzpicture}[-latex, auto, node distance =3.25 cm and 3cm, on grid, thick,
+  empty/.style ={circle, top color=white, bottom color = white, draw, white, text=white , minimum width =1.25 cm},
+  policy/.style ={circle, top color=white, bottom color = blue, draw, black, text=black , minimum width =1.25 cm},
+    behavior/.style ={circle, top color=white, bottom color = ForestGreen, draw, black, text=black , minimum width =1.25 cm},
+  observed/.style ={circle, top color=white, bottom color = magenta, draw, black, text=black , minimum width =1.25 cm} ,
+   confounder/.style ={circle, top color=white, bottom color =  gray, draw, black, text=black , minimum width =1.25 cm} ,
+outcome/.style ={circle, top color=white, bottom color = red, draw, black, text=black , minimum width =1.25 cm},
+myarrow/.style={-Stealth}]
+
+\node[observed] (I)  {\tiny $I^\star_{i,t-\ell}$};
+\node[outcome]  (Y)[below=of I ]{\tiny $Y^\star_{it}$};
+\node[observed] (In)  [right=of I ]{\tiny $I^\star_{it}$};
+\node[outcome]  (Yn)[below=of In ]{\tiny $Y^\star_{i,t+\ell}$};
+\node[observed] (Inn)  [right=of In ]{\tiny $I^\star_{i,t+\ell}$};
+\node[outcome]  (Ynn)[below=of Inn ]{\tiny $Y^\star_{i,t+2\ell}$};
+ \draw [->] (I) -- node[sloped,font=\small,below] {\tiny CF-SEM(t-$\ell$)} (Y);
+ \draw [->] (I) --  (In);
+ \draw [->] (In) -- node[sloped,font=\small,below] {\tiny CF-SEM(t)} (Yn);
+\draw [->] (Y) --  (In);
+\draw [->] (In) --  (Inn);
+ \draw [->] (Inn) -- node[sloped,font=\small,below] {\tiny CF-SEM(t+$\ell$) } (Ynn);
+\draw [->] (Yn) --  (Inn);
+
+\end{tikzpicture} \end{center}
+\caption{ A Diagram for Counterfactual Information Dynamics in CF-SEM}\label{fig:DynS2}
+\end{figure}
 
@@
   \begin{align}
- \underbracket{{\ycolor Y^\star_{i,t+\ell} -  Y_{i,t+\ell} }}_{\text{Predicted CF Change}} & =
+ \underbracket{{\ycolor Y^\star_{i,t+\ell} -  Y_{i,t+\ell} }}_{\text{CF Change}} & =
 \underbracket{{\bcolor\alpha}' {\pcolor\beta}' {\pcolor\left(P^\star_{it} - P_{it} \right)}}_{\text{CF Policy Effect via Behavior}}  +    {\underbracket{\pcolor\pi' \left( P^\star_{it} -  P_{it} \right)}_{\text{CF Direct  Effect}} } \nonumber\\
@@
 
-[Doesn't this just give us weekly cases instead of cumulative
-cases? I.e. following the notation in \ref{eq:M}, it these $C_{it}$
-should be $\Delta C_{it}$.]
-
-In order to compute the counterfactual relative contrasts for levels
-variables, for example, confirmed cases $C_{it}$
-we would use the formula:
-\begin{equation}
-C^{\star}_{it}/C_{it} = \exp(  \sum_{\bar t=1}^t ( Y^\star_{i\bar t} -  Y_{i\bar t}) ),
-\end{equation}
-and similarly for fatalities. The counterfactual relative contrast for national level variables takes the form:
-\begin{equation}
-C^{\star}_{it}/C_{it} =  \sum_{i=1}^N w_i \exp(  \sum_{\bar t=1}^t (
-Y^\star_{i\bar t} -  Y_{i\bar t}) ), \quad w_i = C_{it}/\sum_{i=1}^N
-C_{it}.  \label{eqn:relative_national}
-\end{equation}
-Note this counterfactual contrast is conditional on the actual history $\{C_{it}\}$.  If we'd want unconditional counterfactual,
-we'd need to make assumptions about the evolution of stochastic shocks appearing in $Y_{it}$.\footnote{See, e.g, previous
+All counterfactual quantities and contrasts can be computed from the expressions given
+above. For examples, new cases are linked to growth rates via relation:
+$$
+\Delta C^{\star}_{it} = \exp\left (  \sum_{\bar t=1}^t  Y^\star_{i\bar t} \right) \text{ and } \Delta C_{it} = \exp
+\left (  \sum_{\bar t=1}^t Y_{i\bar t}\right).
+$$
+The cumulative cases can be constructed by summing over the new cases. Various contrasts are
+then calculated from these quantities. For example, the relative contrast of new counterfactual cases
+to the  factual cases is given by:
+$$
+\Delta C^{\star}_{it}/\Delta C_{it} = \exp\left(  \sum_{\bar t=1}^t ( Y^\star_{i\bar t} -  Y_{i\bar t}) \right).
+$$
+We refer to the appendix for further details. Similar calculations apply for fatalities. 
+Note that our analysis is conditional on the factual history and structural stochastic 
+shocks.\footnote{ If we'd want unconditional counterfactual,
+we'd need to make assumptions about the evolution of stochastic shocks appearing in $Y_{it}$. See, e.g, previous
 versions of our paper in ArXiv, where unconditional counterfactuals were calculated by assuming stochastic shocks
@@
 contrasts were small in our empirical analysis.}
+
+The estimated counterfactuals are smooth functionals of the underlying parameter estimates. Therefore, we
+construct the confidence intervals for counterfactual quantities and contrasts by bootstrapping the parameter
+estimates. We refer to the appendix for further details.
 
@@
 
-\subsection{Counterfactuals \label{app:counterfactuals}}
+\subsection{Details for Computing Counterfactuals \label{app:counterfactuals}}
 
@@
 model, with behavior concentrated out.
-
 $$
@@
 $$
-
 We consider a counterfactual change of $P_{it}$ to $P_{it}^*$, while
@@
 $t$. Our outcome is:
-
 $$
@@
 $$
-
 where $\Delta$ is a 7 day difference operator. Writing the model in
 terms of $\Delta \log \Delta C$ we have:
-
 $$
@@
 $$
-
 To simplify computation, we rewrite this model in terms of $\log
 \Delta C$:
-
 $$
@@
 $$
-
 This equation is used to iteratively compute
@@
 Weekly cases instead of log weekly cases are given simply by
-
 $$ \Delta C_{i,t+\ell}^* = \exp(\log \Delta C_{i,t+\ell}^*)$$
@@
 Cumulative cases can be recursively computed as:
-
 $$
@@
 $$
-
 given initial $\{C_{i,s}\}_{s=-7}^0$.
@@
 versus integrate them out) do not matter when considering linear
-constrasts of $\log \Delta C_{it}$ or $\Delta \log \Delta C_{it}$, or
-when considering relative constrasts of $\Delta C_{it}$.
+constasts of $\log \Delta C_{it}$ or $\Delta \log \Delta C_{it}$, or
+when considering relative contrasts of $\Delta C_{it}$.
 
@@
 residuals associated with $\tilde{\theta}_j$. We then compute
-
-$$ \tilde{s}_j = S(\tilde{\theta}_j, \tilde{\varepsilon}_j) $$
-
+$ \tilde{s}_j = S(\tilde{\theta}_j, \tilde{\varepsilon}_j) $
 for $j=1,..., 200$ and plots the mean across $j$ as a point estimate

### 2020-09-03 tex/covidEffectOfNPI-revision.tex
sensitivity discussion

@@
 Our analysis finds that both policies and information on transmission risks are important determinants of Covid-19  cases and deaths and shows that a change in policies explains a large fraction of observed changes in social distancing behavior.
-Our main Lcounterfactual experiments suggest that nationally mandating face masks for employees on April 1st could have reduced the growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 19 to 47 percent less deaths nationally by the end of May, which roughly translates into 19 to 47 thousand saved lives.
+Our main counterfactual experiments suggest that nationally mandating face masks for employees early in the pandemic could have reduced the weekly growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 19 to 47 percent less deaths nationally by the end of May, which roughly translates into 19 to 47 thousand saved lives.
 We also find that, without stay-at-home orders, cases would have been larger by  6 to 63 percent and without business closures, cases would have been larger by 17 to 78 percent.
 We find considerable uncertainty over the effects of school closures due to lack of cross-sectional variation;  we could not robustly rule out either large or small effects.
+Overall, substantial declines in growth rates are attributable to private behavioral response, but policies played an important role as well.
 We also carry out sensitivity analyses to find neighborhoods of
-the models under which the results hold robustly.
-Overall, substantial declines in growth rates are attributable to private behavioral response, but policies played an important role as well.
-We stress that our study is observational and should be interpreted with great caution.
+the models under which the results hold robustly:  the results on mask policies appear to be more much more robust than 
+than the results on business closures and stay-at-home orders.
+Finally, we stress that our study is observational and therefore should be interpreted with great caution.
 From a completely agnostic point of view,
@@
 confidence intervals of $[9,32]$\% and $[19,47]$\%, which roughly
-implies that 34 thousand lives could have been saved.  This finding is
+implies that 34 thousand lives could have been saved.\footnote{These estimates
+are slightly lower than those reported in earlier releases of the paper on ArXiv.
+One of the reason is that we chose to focus on a simplified version of the model 
+as our base model, by
+averaging business
+closure policies as measure of shutdown. This produced more stable results when
+conducting sensitivity results, in particular using fixed effects estimators.  
+We also present several variants of base model, including
+one with separate shutdown policies, in which the effects are slightly larger, in line
+with the previous findings, but we focus reporting counterfactuals only 
+for the base model. } This finding is
 significant: given this potentially large benefit of reducing the
@@
 effect of no stay-at-home orders on deaths is a slightly smaller with
-a 90 percent confidence interval of $-7$\% to $50$\%.  These estimates
-contribute to the ongoing efforts of evaluating various reopening
-approaches.
-
-The above findings are robust to changes to our regression specification, sample selection, methodology, and assumptions about delays between policy changes and changes in recorded cases.
-In all cases, we find that mask mandates, business closure policies, and stay-at-home orders all reduce growth rates by roughly 0.1.
-The impact of mask mandates are more precisely estimated than that of business closure policies or stay-at-home orders.
-This is largely due to the greater variation in the timing of mask mandates across states.
+a 90 percent confidence interval of $-7$\% to $50$\%.
+
+We also conducted sensitivity analysis with respect to changes to our regression specification, sample selection, methodology, and assumptions about delays between policy changes and changes in recorded cases.
+In particular, we examined whether certain effect
+sizes can be ruled by various more flexible models or by using alternative timing assumptions
+that define forward growth rates.
+The impact of mask mandates is more robustly and more precisely estimated than that of business closure policies or stay-at-home orders, and the effect of $-.1$ could not be ruled out by any of the models.\footnote{This null
+hypothesis can be generated by looking at the meta-evidence from RCTs on the efficacy of masks in preventing other respiratory
+cold-like deceases. Falsely rejecting this null is very costly in terms of potential loss of life, and so it is a reasonable null choice 
+for the mask policy from decision-theoretic point of view.} This is largely due to the greater variation in the timing of mask mandates across states.
+The findings of shelter-in-place and business closures policies are considerably less robust.
+For example, for stay-at-home mandates, models with alternative timing and richer specification for information set suggested smaller effects, albeit
+after application of machine learning tools to reduce dimensionality, the range of effects $[-.15, 0]$ could not be ruled out.  A similar
+wide range of effects could also be not ruled out for business closures.
 
@@
 estimates of the effect of school closures as low as 0, and as high as
-0.6. In particular, we find that the results are sensitive to whether
+-0.6. In particular, we find that the results are sensitive to whether
 the number of past national cases/deaths is included in a
@@
 \begin{figure}[ht]
-  \caption{Effect  of nationally mandating masks for employees on March
+  \caption{Relative cumulative effect  on confirmed cases and fatalities 
+  of nationally mandating masks for employees on March
     14th in the US \label{fig:US-mask-dgrowth}}
@@
  % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of no business closure policies on cases in the US  \label{fig:US-nb-dgrowth}}
+  \caption{Relative cumulative effect  of no business closure policies on cases and fatalities in the US  \label{fig:US-nb-dgrowth}}
   \begin{minipage}{\linewidth}
@@
  % \caption{National effect of leaving non-essential businesses open\label{fig:US-nb}}
-  \caption{Effect  of  not implementing stay-at-home order on cases in the US  \label{fig:US-shelter-dgrowth}}
+  \caption{Relative cumulative effect  of  not implementing stay-at-home order on cases and fatalities in the US  \label{fig:US-shelter-dgrowth}}
   \begin{minipage}{\linewidth}
@@
 
-\subsection{Specifications, Timings, and Double Machine Learning}
+\subsection{Specifications, Timings, and Flexible Controls via Machine Learning}
 In this section, we provide sensitivity analysis by estimating (\ref{eq:R4}) with alternative specifications and methods.
@@
    \item[(8)] Baseline specification is estimated by instrumenting $\Delta \log T_{it}$ with one week lagged  logarithm  value of the number of tests per 1000 people.
-   \item[(9)]   Estimated by Double Machine Learning (DML) of \cite{chernozhukov18} with Lasso to reduce dimensionality while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables  as additional controls.
+   \item[(9)]   Estimated by Double Machine Learning (DML) (e.g,,\cite{chernozhukov18}) with Lasso to reduce dimensionality while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables  as additional controls.
 \item[(10)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables as additional controls.

### 2020-09-03 tex/reply-to-referees.tex
updated replies still incomplete

@@
 
-The testable restrictions of the model are now described better.  Please see the discussion around TR. The SIR model
+Thank you. The testable restrictions of the model are now described better.  Please see the discussion around equation (TR). The SIR model
 for example implies that testing rate is a confounder for case growth rate, but the causal model implies  that testing rate
@@
  \begin{quote}
- We stress that our study is observational and should be interpreted with great caution. 
+ We stress that our study is observational and therefore should be interpreted with great caution. 
  From a completely agnostic point of view,
@@
 
- This point is well taken. We added the following paragraph to Section 2.2. 
+ This point is valid.  We added a cautionary message to the abstract. Furthermore,
+ we added the following paragraph to Section 2.2. 
  \begin{quote}
@@
 
-We provide various sensitivity checks. First, we now include governor's party affiliation as an additional state-level variable in random effects specification. 
+We provide various sensitivity checks: First, we now include governor's party affiliation as an additional state-level variable in random effects specification. 
 Second, we exclude the state of New York from the sample because it may be viewed as an
@@
 unobserved private behavioral response. Fifth, we report the fixed effects estimator that controls for latent state confounders as components of $W_{it}$.  
-As shown in Figures xx(2)-(4) and (9), the estimated coefficient of masks for employees is not so sensitive with respect to excluding NY as well as adding more controls that are possibly important confounders.
+As shown in the paper, the estimated coefficient of masks for employees is not so sensitive with respect to excluding NY as well as adding more controls that are possibly important confounders.
+The results of other policies are less robust/uncertain and we clearly communicate this in the abstract and in the introduction.
 
@@
 We find that the estimated coefficients of masks for employees are robust with respect to alternative specifications and methods.
+Other coefficients on other policies are less robust
+(but helpful effects of these policies can not be ruled out).
 
@@
 
-We chose information structure that is parsimonious and yet, we believe, captures the  information that is most relevant for people's behavior.
+We chose information structure that is parsimonious and yet, we believe, captures the  information that is most relevant for people's behavior (at the minimum, it captures our own behaviors).
+This information structure is also in line with CDC guidelines
+for policies, which
+is formulated in terms of observed infection of outcomes and not in terms of behavior.
 
 Conditioning on current policy variables that is included in all specifications, the lagged values of policies will  likely add little additional information and, therefore, we decided to exclude lagged values of policies from the information structure.
-We may re-interpret that the coefficient of current policy variables partially reflects its effect of providing information.
+We are
+completely free to interpret that the coefficient of current policy variables has both informational content and behavior-constraining effect.
 
@@
 As a sensitivity check, we estimated a specification that includes  lagged behavior variables as information variables in Section 5.
-As shown in specification (5) of Figures xx-xx, the estimated coefficients of mask for employees, closed K-12 schools, stay-at-home orders, and the average of stay-at-home orders and closures of movie theaters, restaurants, and non-essential businesses are not so sensitive to an inclusion of lagged behavior variables as information.
+The mask policies
+appear to be robust withe respect to the inclusion of these variables. The effects of business closures appear to be less
+robust -- the estimated effects become more attenuated, and we report these findings in the conclusion.
 
@@
 We think it is reasonable to assume that people's behavior today will not immediately affect today's policy because it takes time for policy makers to gather information on people's behavior and make a policy decision based on gathered information, especially in the early pandemic period.
+Moreover, CDC guidelines don't seem to rely on people's behavior but on measured infection outcomes.
 
@@
 
-We have the following paragraph:
-\begin{quote}
- To compute the estimated and counterfactual
-paths, we use the   estimate   in
-column ``Average'' of Table \ref{tab:dieff-si}. We set initial
-$\Delta \log \Delta C$ and $\log \Delta C$ to their values first
-observed in the state we are simulating. We now provide
-more details on how counterfactuals are constructed -- see Section 2.
-The shaded region is a point-wise 90\% confidence
-interval.
-\end{quote}
+Thank you! We have added a much more detailed discussion and also an appendix where the details
+are explained.
 
@@
 at the end of May, and they announced their findings in middle of July). The 
-"stars" are reported as a matter of statistical convention, but we never use them in any formal way.
+"asteriks" are reported as a matter of statistical convention, but we never use them in any formal way. We could probably
+erase them and proceed in asterikless fashion.
 
@@
 or actual policy decisions.   From a decision-theoretic point of view the cost of accepting the false null is extremely high,
-which could result in huge losses. The Neyman-Pearson test with $\alpha=$ small, corresponds to the idea that the cost
-of the null is small (little harm done if the "no mask effect" is falsely accepted), e.g.
+which could result in huge losses. The Neyman-Pearson test with level $\alpha=$ small, corresponds to the idea that the cost
+of falsely rejecting the null is small (little harm done if the "no mask effect" is falsely accepted), e.g.
 Amemiya "Introduction to Statistics and Econometrics".
 So doing "no mask effect" testing with very low $\alpha$ would correspond to assuming that the cost of deciding that masks are not effective is very low.
+The potential cost of accepting the null, if it is false, is just huge.
 Multiple hypothesis testing would further drive $\alpha$ to smaller values, and we are not sure how this would align with the decision-theoretic framework.
@@
 
-Following your suggestion, we added the above discussion in footnote 20 (check!) when we discuss in Table 3A.
+Following your suggestion, we added the above discussion in footnote X (check!) when we discuss in Table 3A.
 
@@
 
-It is in total by the start of June. We have modify the sentence as suggested.
+It is in total by the start of June. We have modified the sentence as suggested.  
+[CHECK]
 
@@
 Following your suggestion,   we provide a sensitivity analysis on the delay between infection and case reporting (or death reporting) in Section 4.2.
+The timing was motivated by a careful reading of Table 2 referenced above.
 
@@
 
-[We may change the starting date for counterfactual figures.]
+We changed the starting date for counterfactual figures.
 
@@
 We have reviewed and added more references from the epidemiology literature. See, for example, footnote 9.
+There are also many more references in the Into other than Econ literature.
 
@@
 
+The autocorrelation component is captured by the lags of this variable in the dynamic specification. In addition,
+we cluster the standard errors by state (capturing the impact of temporal dependence in the series, 
+not accounted for by the lag structure).
+
 We believe that it is better to use the number of cases and deaths rather than time since start of the outbreak as a  general framework to analyze the spread of COVID-19.
@@
 
+[Shall we include such figure here in the response?]
+
 \end{enumerate}

### 2020-09-04 tex/covidEffectOfNPI-revision.tex
revised abstract, intro, section 2, and conclusion by Hiro

@@
 the models under which the results hold robustly:  the results on mask policies appear to be more much more robust than 
-than the results on business closures and stay-at-home orders.
+ the results on business closures and stay-at-home orders.
 Finally, we stress that our study is observational and therefore should be interpreted with great caution.
@@
 It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
-Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.
-\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto}) and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-The work on causal graphs has been
-modernized and put on rigorous footing by  \cite{pearl:biometrika},\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak}, \cite{peters2020book}, and \cite{hernanrobins2020book}, with major applications taking place in
-in computer science, econometrics, and epidemiology.
-The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \cite[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986},\cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
+Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto}) and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika}, \cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak}, \cite{peters2020book}, and \cite{hernanrobins2020book}, with major applications taking place in computer science, econometrics, and epidemiology.
+The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
 The father and son, P.
@@
 
-Using the estimated model, we evaluate the dynamic impact of the following counterfactual policies on Covid-19 cases and deaths: mandating face masks, allowing non-essential businesses to open, not implementing a stay-at-home order, and removing all policies.
+Using the estimated model, we evaluate the dynamic impact of the following three counterfactual policies on Covid-19 cases and deaths: (1) mandating face masks, (2)  allowing all businesses to open, and (3) not implementing a stay-at-home order.
 The counterfactual experiments show a large impact of  those policies on  the number of cases and deaths.
@@
 businesses (no movie theater closure, no non-essential business
-closure, and no closure of restaurants except take-out)) could have
+closure, and no closure of restaurants except take-out) could have
 affected cases and deaths.  We estimate that business shutdowns have a
@@
 that define forward growth rates.
-The impact of mask mandates is more robustly and more precisely estimated than that of business closure policies or stay-at-home orders, and the effect of $-.1$ could not be ruled out by any of the models.\footnote{This null
+The impact of mask mandates is more robustly and more precisely estimated than that of business closure policies or stay-at-home orders, and an undesirable effect of increasing the weekly death growth by 5 percentage points is ruled out by any of the  models.\footnote{This null
 hypothesis can be generated by looking at the meta-evidence from RCTs on the efficacy of masks in preventing other respiratory
-cold-like deceases. Falsely rejecting this null is very costly in terms of potential loss of life, and so it is a reasonable null choice 
+cold-like deceases. Falsely rejecting this null is   costly in terms of potential loss of life, and so it is a reasonable null choice 
 for the mask policy from decision-theoretic point of view.} This is largely due to the greater variation in the timing of mask mandates across states.
@@
 For example, for stay-at-home mandates, models with alternative timing and richer specification for information set suggested smaller effects, albeit
-after application of machine learning tools to reduce dimensionality, the range of effects $[-.15, 0]$ could not be ruled out.  A similar
-wide range of effects could also be not ruled out for business closures.
+after application of machine learning tools to reduce dimensionality, the range of effects $[0,0.15]$ could not be ruled out.  A similar
+wide range of effects could  not be ruled out for business closures.
 
@@
       cumulative cases or deaths relative to the observed number of
-      cases or deaths. The shaded region is a point wise 90\%
+      cases or deaths. The shaded region is a pointwise 90\%
       confidence band.
@@
       \footnotesize This figure shows the estimated relative change in
-      cases and deaths if no states had never implemented any business closure
+      cases and deaths if no states had ever implemented any business closure
       policies. The thick blue line is the estimated change in cumulative
       cases or deaths relative to the observed number of cumulative cases
-      or deaths. The shaded region is a point wise 90\% confidence
+      or deaths. The shaded region is a pointwise 90\% confidence
       band.
@@
 
-Whether wearing  masks in public place should be mandatory or not has been one of the most contested policy issues with health authorities of different countries providing contradiction recommendations.
+Whether wearing  masks in public place should be mandatory or not has been one of the most contested policy issues with health authorities of different countries providing contradictory recommendations.
 Reviewing evidence, \cite{Greenhalghm2020}  recognize that there is no randomized controlled trial evidence for the effectiveness of face masks,  but they state ``indirect evidence exists to support the argument for the public wearing masks in the Covid-19 pandemic."\footnote{The virus remains viable in the air for several hours, for which surgical masks may be effective.
@@
 \end{itemize}
-The index $i$ denotes observational unit, the state, and $t$ and $t+\ell$ denotes the time, where  $\ell$ represents the time lag  between infection and case confirmation or death.
+The index $i$ denotes observational unit, the state, and $t$ and $t+\ell$ denotes the time, where $\ell$ is a positive integer that represents the time lag  between infection and case confirmation or death.
 \begin{figure}[ht]
@@
 \end{equation}
-where we say that $V \perp U$ if $\Ep VU = 0$. This is a standard way of representing restrictions on errors in structural equation modeling.\footnote{ An alternative useful
+where we say that $V \perp U$ if $\Ep VU = 0$.\footnote{ An alternative useful
 starting point is to impose the Rubin-Rosenbaum type unconfoudedness condition:
@@
 $$
-which imply, with treating stochastic errors as independent additive components, the orthogonal conditions stated above.
- }\footnote{The structural equations of this form are connected to triangular structural equation models, appearing
- in microeconometrics and macroeconometrics (SVARs), going back to the work of  \cite{strotz1960recursive}. .}
+which imply, with treating stochastic errors as independent additive components, the orthogonal conditions stated above.} This is a standard way of representing restrictions on errors in structural equation modeling.\footnote{The structural equations of this form are connected to triangular structural equation models, appearing in microeconometrics and macroeconometrics (SVARs), going back to the work of  \cite{strotz1960recursive}.}
 
@@
 to answer about multi-period responses.  The dynamics will be induced by the evolution
-of information variables: Information variables include time, lagged and integrated values of outcome:
+of information variables.
+Our baseline specification assumes that information variables include time, lagged and integrated values of outcome:\footnote{Our empirical analysis also considers a specification in which information variables include lagged national cases/deaths as well as lagged behavior variables.}
 \begin{equation}\label{eq:I}\tag{I}
@@
 {\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}
- for  each  $t \in [T]$, where $g$ is deterministic function of time, e.g., month indicators,
+ for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
  and $t/\ell$ is defined as $\lfloor {t/\ell} \rfloor$.
@@
 \textsc{Structural Equations Model (SEM)}.
-Let $i \in \{1,..., N\}$ denote the observational init and $t$ be the time periods and $\ell$ the time delay, a non-negative integer.
+Let $i \in \{1,..., N\}$ denote the observational unit,  $t$ be the time periods, and $\ell$ be the time delay.
 (1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined outside of the model for $t \leq 0$.
-(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are determined, then information variables $I_{it}$ are determined  by (I), and policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) and realizing the shochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E),  the behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) and realizing  the shochastic shock $\varepsilon^b_{it}$ that obeys the exogeneity condition (E), the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) and realizing the shochastic shock $\varepsilon^y_{it}$ that obeys the exogeneity condition (E).
+(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E).
 \end{quote}
@@
 model with testing to motivate the presence of change in testing rate as a confounder in the outcome
-equations, but not in the behavior equation (implying that a component of $\delta_B$ and $\delta_P$ is set to 0), implying that (TR) does not necessarily hold and is testable.
+equations, but not in the behavior equation (therefore, a component of $\delta_B$ and $\delta_P$ is set to 0), implying that (TR) does not necessarily hold and is testable.
 Furthermore, we estimate (\ref{eq:R2}) on the data set that has many more observations than the data set used to estimate the outcome equations, implying that (TR) is again testable.
 Later we  shall also try to
-utilize the contextual knowledge that masks business mandates only affect outcome directly and not by changing mobility, implying again that (TR) is testable.
+utilize the contextual knowledge that masks business mandates only affect outcome directly and not by changing mobility (i.e., $\beta=0$ for mask policy), implying again that (TR) is testable.
 If not rejected by the data, (TR) can be used to sharpen the estimate of the causal effect of mask policies on the outcomes.
@@
 We present the results
-on debiased fixed effect estimation as parts of our sensitivity analysis. Other approaches we have explored
+on debiased fixed effect estimation with weekly dummies as parts of our sensitivity analysis. Other approaches we have explored
 is to consider more flexible nonlinear model $f(W_{it}, \delta)$ for the way observed confounders enter the model,
@@
 the parameters of equations separately, and then compute
-\begin{equation}
+\[
  \frac{1}{2} \hat {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\pcolor \hat \beta' }+{\pcolor\hat \pi'} ) \text{ and }
  \frac{1}{2} \hat{\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor \hat \alpha '}  {\icolor \hat \gamma'} + {\icolor \hat \mu'}),
-\end{equation}
+\]
 as the estimator of the total policy effect, that concentrates around  their population analogues
-\begin{equation}
+\[
  \frac{1}{2}  {\mathsf{a}}'  + \frac{1}{2}  ( {\bcolor \alpha '}  {\pcolor \beta' }+{\pcolor \pi'} ) \text{ and }
  \frac{1}{2} {\mathsf{b}}'  + \frac{1}{2}  ( {\bcolor  \alpha '}  {\icolor \gamma'} + {\icolor \mu'}),
-\end{equation}
+\]
 with approximate deviations controlled by the normal laws, with standard deviations that can be approximated by the bootsrap
@@
 a sequence of counterfactual policies for each state:
-$$
+\begin{equation} \label{p-cf}\tag{P-CF}
 \{{\pcolor P^\star_{it}} \}_{t=1}^T, \quad i=1, \ldots N.
-$$
+\end{equation}
 We assume that the SEM remains invariant, except for the policy equation.\footnote{It is possible to consider counterfactual exercises in which policy  responds to information through the policy equation if we are interested in endogenous policy responses to information.
@@
 & {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
-& {\icolor I^\star_{it} }& := &  I ({ t,  {\ycolor Y^*_{it}}, \icolor I^\star_{i,t-\ell}}),  \end{aligned}
+& {\icolor I^\star_{it} }& := &  I ({ t,  {\ycolor Y^\star_{it}}, \icolor I^\star_{i,t-\ell}}),  \end{aligned}
 \end{equation}
@@
 \textsc{Counterfactual Structural Equations Model (CF-SEM)}.
-Let $i \in \{1,..., N\}$ be the observational init and $t$ be time periods and $\ell$ the time delay, a positive integer.
+Let $i \in \{1,..., N\}$ be the observational unit, $t$ be time periods, and $\ell$ be the time delay.
 (1) For each $i$ and $t \leq 0$, the confounder, information, behavior, policy, and outcome variables are determined as in SEM: $W^\star_{it}= W_{it}$, $I^\star_{it}= I_{it}$,  $B^\star_{it} = B_{it}$, $P^\star_{it}=P_{it}$, $Y^\star_{it} =Y_{it}$.
-(2) For each $i$ and $t \geq 0$, confounders $W^\star_{it} =W_{it}$ are determined, then information variables $I^\star_{it}$ are determined  by (I), and policy variables $P^\star_{it}$ are set,  the behavior variables $B^\star_{it}$ are determined by setting $\iota = I^\star_{it}$ and $p= P^\star_{it}$ in  (SO) and keeping  the same stochastic shock $\varepsilon^b_{it}$ in (SO), the outcome $Y^\star_{i, t + \ell}$ is realized by setting $\iota = I^\star_{it}$, $p= P^\star_{it}$, and $b = B^\star_{it}$ in  (SO) and keeping the same shochastic shock $\varepsilon^y_{it}$  (SO).
+(2) For each $i$ and $t \geq 0$, confounders $W^\star_{it} =W_{it}$ are determined, and information variables $I^\star_{it}$ are determined  by (I);  policy variables $P^\star_{it}$ are set in (\ref{p-cf}); behavior variables $B^\star_{it}$ are determined by setting $\iota = I^\star_{it}$ and $p= P^\star_{it}$ in  (SO) with  the same stochastic shock $\varepsilon^b_{it}$ in (SO); the  counterfactual outcome $Y^\star_{i, t + \ell}$ is realized by setting $\iota = I^\star_{it}$, $p= P^\star_{it}$, and $b = B^\star_{it}$ in  (SO) with the same stochastic shock $\varepsilon^y_{it}$ in  (SO).
 \end{quote}
@@
 All counterfactual quantities and contrasts can be computed from the expressions given
-above. For examples, new cases are linked to growth rates via relation:
+above. For examples, given $\Delta C_{i0}>0$, new cases are linked to growth rates via relation:
 $$
-\Delta C^{\star}_{it} = \exp\left (  \sum_{\bar t=1}^t  Y^\star_{i\bar t} \right) \text{ and } \Delta C_{it} = \exp
-\left (  \sum_{\bar t=1}^t Y_{i\bar t}\right).
+\frac{\Delta C^{\star}_{it}}{\Delta C_{i0}} = \exp\left (  \sum_{s=1}^t  Y^\star_{is} \right) \text{ and }\frac{ \Delta C_{it}}{\Delta C_{i0}} = \exp
+\left (  \sum_{s=1}^t Y_{is}\right).
 $$
@@
 $$
-\Delta C^{\star}_{it}/\Delta C_{it} = \exp\left(  \sum_{\bar t=1}^t ( Y^\star_{i\bar t} -  Y_{i\bar t}) \right).
+\Delta C^{\star}_{it}/\Delta C_{it} = \exp\left(  \sum_{s=1}^t ( Y^\star_{is} -  Y_{is}) \right).
 $$
@@
 Note that our analysis is conditional on the factual history and structural stochastic 
-shocks.\footnote{ If we'd want unconditional counterfactual,
-we'd need to make assumptions about the evolution of stochastic shocks appearing in $Y_{it}$. See, e.g, previous
+shocks.\footnote{For unconditional counterfactual,
+we need to make assumptions about the evolution of stochastic shocks appearing in $Y_{it}$. See, e.g, previous
 versions of our paper in ArXiv, where unconditional counterfactuals were calculated by assuming stochastic shocks
@@
 
-\subsection{Outcome and Key Confounders via SIR model}\label{sec:sirmodel} We next provide details
+\subsection{Outcome and Key Confounders via SIRD model}\label{sec:sirmodel} We next provide details
 of our key measurement equations, defining the outcomes and key confounders. We motivate the structural
 outcome equations via the fundamental epidemiological model for the spread of infectious decease called
-the SIRD (susceptible - infectected - recovered-deceased) model with testing.
+the Susceptible-Infected-Recovered-Dead (SIRD) model with testing.
 
@@
 is the key confounding variable,
-derived from considering the SIR model below. We are treating the change in testing
+derived from considering the SIRD model below. We are treating the change in testing
 rate as exogenous.\footnote{To check sensitivity to this assumption
@@
    \item[(7)] Add weekly dummies  to the baseline specification.
-   \item[(8)] Baseline specification is estimated by instrumenting $\Delta \log T_{it}$ with one week lagged  logarithm  value of the number of tests per 1000 people.
-   \item[(9)]   Estimated by Double Machine Learning (DML) (e.g,,\cite{chernozhukov18}) with Lasso to reduce dimensionality while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables  as additional controls.
-\item[(10)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while adding private mask wearing rates,  the log of Trump's vote share, and  past behavior variables as additional controls.
+   \item[(8)] Baseline specification estimated by instrumenting $\Delta \log T_{it}$ with one week lagged  logarithm  value of the number of tests per 1000 people.
+   \item[(9)]   Estimated by Double Machine Learning (DML) \citep[e.g.,][]{chernozhukov18} with Lasso to reduce dimensionality while including all additional controls in (3)-(5).
+\item[(10)]  Estimated by DML with Random Forest to reduce dimensionality  and capture some nonlinearities while including all additional controls in (3)-(5).
 %  \item[(10)]   Add latent state confounders and weekly dummies as components of $W_{it}$, estimated using fixed effects with bias correction.
@@
 growth rate from mandating masks on March 14th. The shaded region is a
-90\% pointwise confidence interval. As shown, mandating masks on April
-1st lowers the growth of cases or deaths 14 or 21 days later by 0.09 or
+90\% pointwise confidence interval. As shown, mandating masks on March 14th lowers the growth of cases or deaths 14 or 21 days later by 0.07 or
 0.14. This effect then gradually declines due to information
@@
 (i.e. the dots are identical to the solid line in the left panel of
-Figure \label{fig:WA-mask}, except for each state instead of just
-Washington). The solid blue is the national average change in growth
+Figure \ref{fig:WA-mask}, except for each state instead of just
+Washington). The solid blue line is the national average change in growth
 rate. The shaded region is a 90\% pointwise confidence band for the
@@
 given by $[9,32]$\% and $[19,47]$\%, respectively.
-The result roughly translates into $30$ thousand saved lives.
+The result roughly translates into $19$ to $47$ thousand saved lives.
 
@@
 $[17,78]$\%.
-The confidence interval for deaths is even wider, $[1, 97]$\%
+The confidence interval for deaths is even wider, $[1, 97]$\%.
 
@@
 
-First, our empirical analysis indicates that mandating face masks has
+First, our empirical analysis robustly indicates that mandating face masks has
 reduced the spread of Covid-19 without affecting people's social
 distancing behavior measured by Google Mobility Reports. Our
-counterfactual experiment based on the estimated model suggests that
-if all states had have adopted mandatory face mask policies on April
-1st of 2020, then the number of deaths per week
-by the end of May would have been smaller by as much as  $18$ to $50$\%, which roughly translates to $6$ thousand saved lives.
-
-Second, we find that keeping non-essential businesses open would have led to -20 to 60\% more cases  while not implementing stay-at-home orders would have increased cases by 25 to 170 \% by the start of June.
-% Except for mask mandates, the effect of all other policies on case and death growth is realized through their impact on social distancing behavior.
-
-Third, we find considerable uncertainty over  the impact of all policies combined on case or death growth because it is difficult to identify the effect of school closures from the US state-level data due to the lack of variation in the timing of school closures across states.
+counterfactual experiment suggests that
+if all states had have adopted mandatory face mask policies on March 14th of 2020, then the cumulative number of deaths by the end of May would have been smaller by as much as  $19$ to $47$\%, which roughly translates into $19$ to $47$  thousand saved lives.
+
+Second, our baseline counterfactual analysis suggests that keeping all businesses open would have led to 17 to 78\% more cases  while not implementing stay-at-home orders would have increased cases by 6 to 63\% by the end of May although we should interpret these numbers with some caution given that the estimated effect of business closures and stay-at-home orders vary across specifications in our sensitivity analysis.
+
+Third, we find considerable uncertainty over  the impact of school closures on case or death growth because it is difficult to identify the effect of school closures from the US state-level data due to the lack of variation in the timing of school closures across states.
 
@@
 
-We compute counterfactuals from the "total effect" version of the
+We compute counterfactuals from the ``total effect" version of the
 model, with behavior concentrated out.
@@
 $$
-We consider a counterfactual change of $P_{it}$ to $P_{it}^*$, while
+We consider a counterfactual change of $P_{it}$ to $P_{it}^\star$, while
 $W_{it}$ and $\bar \varepsilon_{it}$ are held constant. In response to
 the policy change, $Y_{i,t+\ell}$ and the part of $I_{it}$ that
-contains $Y_{it}$, change to $Y_{i,t+\ell}^*$ and $I_{it}^*$. To be
+contains $Y_{it}$, change to $Y_{i,t+\ell}^\star$ and $I_{it}^\star$. To be
 specific, let $C_{it}$ denote cumulative cases in state $i$ on day
@@
 This equation is used to iteratively compute
-$\log \Delta C_{i,t+\ell}^*$
+$\log \Delta C_{i,t+\ell}^\star$
 conditional on initial $\{\log \Delta C_{i,s}\}_{s=-\ell-7}^{0}$,
-and the entire sequence of $\{W_{i,t}, P_{i,t}^*, \bar \varepsilon_{it} \}_{t=0}^T$.
+and the entire sequence of $\{W_{i,t}, P_{i,t}^\star, \bar \varepsilon_{it} \}_{t=0}^T$.
 
 Weekly cases instead of log weekly cases are given simply by
-$$ \Delta C_{i,t+\ell}^* = \exp(\log \Delta C_{i,t+\ell}^*)$$
+$$ \Delta C_{i,t+\ell}^\star = \exp(\log \Delta C_{i,t+\ell}^\star)$$
 
@@
 
-Note that $\log \Delta C_{i,t+\ell}^*$ depends linearly on
+Note that $\log \Delta C_{i,t+\ell}^\star$ depends linearly on
 $\bar \varepsilon$, so the residuals (and our decision to condition on
@@
 \item Contrasts of growth rates: $S(\theta, \mathbf{\varepsilon}) =
-  \Delta \log \Delta C_{it}^* - \Delta \log \Delta C_{it}$
+  \Delta \log \Delta C_{it}^\star - \Delta \log \Delta C_{it}$
 
 \item Relative contrasts of weekly cases: $S(\theta, \mathbf{\varepsilon}) =
-  \Delta C_{it}^*/\Delta C_{it}$
+  \Delta C_{it}^\star/\Delta C_{it}$
 
 \item Relative contrasts of cumulative cases: $S(\theta, \mathbf{\varepsilon}) =
-  \frac{C_{it}^* - C_{it}}{C_{it}}$
+  \frac{C_{it}^\star - C_{it}}{C_{it}}$
 \end{itemize}

### 2020-09-04 tex/covidEffectOfNPI-revision.tex
update

@@
 We assume that the SEM remains invariant, except for the policy equation.\footnote{It is possible to consider counterfactual exercises in which policy  responds to information through the policy equation if we are interested in endogenous policy responses to information.
-Although this is beyond the scope of the current paper, counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information.
+Counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information.
 } Given the policies, we generate the counterfactual

### 2020-09-04 tex/covidEffectOfNPI-revision.tex
upted

@@
 Overall, substantial declines in growth rates are attributable to private behavioral response, but policies played an important role as well.
-We also carry out sensitivity analyses to find neighborhoods of
-the models under which the results hold robustly:  the results on mask policies appear to be more much more robust than 
- the results on business closures and stay-at-home orders.
+We also carry out sensitivity analyses to find neighborhoods of the models under which the results hold robustly:  the results on mask policies appear to be  much more robust than   the results on business closures and stay-at-home orders.
 Finally, we stress that our study is observational and therefore should be interpreted with great caution.
@@
 The information variables, such as recent growth in the number of cases, can cause people to spend more time at home, regardless of adopted state policies; these changes in behavior in turn affect the transmission of Covid-19.
+Importantly, policies can have the informational content as well, guiding behavior rather than constraining it.
 
@@
 
-Our quantitative model for causal structure in Figure \ref{Wright} is given by the following econometric structural (or potential) outcomes model:\begin{equation} \label{eqn:SEM1} \tag{SO}
+Our quantitative model for causal structure in Figure \ref{Wright} is given by the following econometric structural
+ (or potential) outcomes model:\begin{equation} \label{eqn:SEM1} \tag{SO}
   \begin{aligned}
@@
 
-The exogeneity restrictions on the stochastic components are as follows:
+The exogeneity restrictions on the stochastic shocks are as follows:
 \begin{equation}\label{eqn:SEM2} \tag{E}
@@
 are modeled as a function of state level characteristics, and the time random effects are modeled
-by including month dummies and their interactions with state level characteristics.
-The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$ are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.
+by including month dummies and their interactions with state level characteristics (in the sensitivity analysis,
+we also add weekly dummies).  The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$ are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.
 
@@
 \]
-as the estimator of the total policy effect, that concentrates around  their population analogues
+as the estimator of the total policy effect. Under standard regularity conditions,
+these estimators concentrate around  their population analogues
 \[
@@
 resampling of observational units $i$. Under correct specification the target quantities
-reduce to $  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} $ and ${\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'}$, respectively.
+reduce to $$  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} \text{ and } {\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'},$$ respectively.
 This construction is not as efficient as generalized method of moments,
@@
 a sequence of counterfactual policies for each state:
-\begin{equation} \label{p-cf}\tag{P-CF}
+\begin{equation} \label{p-cf}\tag{CF-P}
 \{{\pcolor P^\star_{it}} \}_{t=1}^T, \quad i=1, \ldots N.
@@
 outcomes, behavior, information by propaging the dynamic equations:
-\begin{equation} \label{eqn:CF-SEM} \tag{CEF-SEM}
+\begin{equation} \label{eqn:CF-SEM} \tag{CF-SEM}
   \begin{aligned}

### 2020-09-04 tex/reply-to-referees.tex
updating Figure 10 with residual versions and regressionWithPlots

@@
 \item   Following suggestions by the editor and the referees, we provide various sensitivity checks as well as the results of debiased fixed effects estimator  in Section 4.
+
+\item Given Referee 1's comment, all regressions now include a dummy for state governor's party and its interaction with month dummies as additional controls.
+
 \item   We improve the figures for counterfactual experiments so that they are now easier to understand.
+
+\item We now combine movie theater closures, restaurant closures, and non-essential business closures as one policy variable called ``business closure policies'' by taking the average of these three variables.
+Accordingly, we conduct a counterfactual of keeping all business open.
+
+\item We changed the starting date of mask mandates to March 14th in our counterfactual experiment of mandatory mask policy.
+
 \item  We adjusted the timing of monthly dummies for estimating behavior equation ($PI \to B$)  in Table 3.
@@
 
-[We can move this to introduction. Or we can keep it as it is by saying that there could be confounders and many other things at play.]
+We thought about this suggestion but, eventually, we can keep it as it is.
 
@@
 It is in total by the start of June. We have modified the sentence as suggested.
-[CHECK]
 
@@
 
-[Shall we include such figure here in the response?]
+The following figures are generated by first regressing the case or death growth rates on state characteristics, month dummies, and their interactions and then computing their average value for states with or without a mask mandate for each date.
+
+\begin{figure}[ht]
+  \caption{Average  residuals of regressing case and death growth on confounders among states with or without  mask mandates \label{fig:masks-res}}\vspace{0.2cm}
+ \begin{minipage}{\linewidth}
+    \centering
+    \begin{tabular}{cc}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases-res-14}
+      &
+        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths-res-21}
+    \end{tabular}
+    \begin{flushleft}
+      \scriptsize In these figures, red points are the regression residuals of case or death
+      growth rate on confounders in states without a mask mandate. Blue points are
+      states with a mask mandate 14 (21 for deaths) days prior. The
+      red line is the average across states without a mask mandate 14
+      (21 for deaths) days earlier. The blue line is the average
+      across states with a mask mandate 14 (21 for deaths) earlier.
+    \end{flushleft}
+  \end{minipage}
+\end{figure}
 

### 2020-09-04 cases_and_policies/rmd/regressionsWithPlots.Rmd
updating Figure 10 with residual versions and regressionWithPlots

@@
 rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
+
+# Set time lag:
+timing <- 0 # 0 or 1
+if (timing==0){
+  L.c <- 14
+  L.d <- 21
+} else if (timing==1) {
+  L.c <- 7
+  L.d <- 24
+}
+
 source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
@@
 pols = c("pmaskbus","pk12","pshelter","pindex")
-
-# Set time lag:
-timing <- 0 # 0 or 1
-if (timing==0){
-  L.c <- 14
-  L.d <- 21
-} else if (timing==1) {
-  L.c <- 7
-  L.d <- 24
-}
 
@@
 
-df$logpop <- log(df$Population.2018)
-df$logsq <- log(df$Square.Miles)
-df$repub <- as.factor(df$party)
-df$party_dum <- ifelse(df$party == "republican", 0, 1)
-df$month3 <- ifelse(df$month==3,0,1)
-df$month4 <- ifelse(df$month==4,0,1)
-df$month5 <- ifelse(df$month==5,0,1)
-df$month6 <- ifelse(df$month==6,0,1)
-L <- L.c
-
-pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential")
-infovars <- list(c("dlogdc", "logdc"),
-                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
-infovarsd <- list(c("dlogdd", "logdd"),
-                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
-bvars <- c("workplaces","retail","grocery","transit")
-vars <- c(infovars[[2]],infovarsd[[2]],pols,"pindex")
-for(i in 1:length(vars)) {
-  L<-L.c
-  v <- sprintf("%s.L14",vars[i])
-  df[,v] <- panellag(df[,vars[i]],df$state,df$date, lag=L)
-  L<-L.d
-  v <- sprintf("%s.L21",vars[i])
-  df[,v] <- panellag(df[,vars[i]],df$state,df$date, lag=L)
-}
-
-for(i in 1:length(bvars)) {
-  L<-L.c
-  v <- sprintf("%s.L14",bvars[i])
-  df[,v] <- panellag(df[,bvars[i]],df$state,df$date, lag=L+14)
-  L<-L.d
-  v <- sprintf("%s.L21",bvars[i])
-  df[,v] <- panellag(df[,bvars[i]],df$state,df$date, lag=L+14)
-}
-
-svars <- c("logpop","logsq","Percent.Unemployed..2018.",
-               "Percent.living.under.the.federal.poverty.line..2018.",
-               "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum")
-mvars.c <- mvars.d <- mvars <- c("month4","month5","month6")
-# mvars.c <- c("month4.c","month5.c","month6.c")
-# mvars.d <- c("month4.d","month5.d","month6.d")
-ijs <- expand.grid(1:length(svars), 1:length(mvars.c))
-msvars.c <- vector(mode="character", length=lengths(ijs)[[1]])
-msvars.d <- vector(mode="character", length=lengths(ijs)[[1]])
-for (ij in (1:lengths(ijs)[[1]])){
-  v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars.c[ijs[[2]][ij]])
-  msvars.c[ij] <- v
-  df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars.c[ijs[[2]][ij]]]
-  v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars.d[ijs[[2]][ij]])
-  msvars.d[ij] <- v
-  df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars.d[ijs[[2]][ij]]]
-}
-
-bvars.c <- c("workplaces.L14","retail.L14","grocery.L14","transit.L14")
-bvars.d <- c("workplaces.L21","retail.L21","grocery.L21","transit.L21")
-
-wvars <- c(svars,mvars.c,msvars.c,"month3","mask_percent","logvote","dlogtests")
-wvars.d <- c(svars,mvars.d,msvars.d,"month3","mask_percent","logvote")
-xvars <- c(bvars.c,"dlogdc.L14","logdc.L14")
-xvars.national <- c(bvars.c,"dlogdc.L14","logdc.L14","dlogdc.national.L14","logdc.national.L14")
-xvars.d <- c(bvars.d,"dlogdd.L21","logdd.L21")
-xvars.national.d <- c(bvars.d,"dlogdd.L21","logdd.L21","dlogdd.national.L21","logdd.national.L21")
-
-# wvars <- c(bvars.c,svars,mvars.c,msvars.c,"month3","mask_percent","logvote","dlogtests")
-# wvars.d <- c(bvars.d,svars,mvars.d,msvars.d,"month3","mask_percent","logvote")
-# xvars <- c("dlogdc.L14","logdc.L14")
-# xvars.national <- c("dlogdc.L14","logdc.L14","dlogdc.national.L14","logdc.national.L14")
-# xvars.d <- c("dlogdd.L21","logdd.L21")
-# xvars.national.d <- c("dlogdd.L21","logdd.L21","dlogdd.national.L21","logdd.national.L21")
-
-# pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pmovie.L14","prestaurant.L14","pnonessential.L14")
-# pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pmovie.L21","prestaurant.L21","pnonessential.L21")
-pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pindex.L14")
-pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pindex.L21")
+## the following part is moved to regprep.R
+# df$logpop <- log(df$Population.2018)
+# df$logsq <- log(df$Square.Miles)
+# df$repub <- as.factor(df$party)
+# df$party_dum <- ifelse(df$party == "republican", 0, 1)
+# df$month3 <- ifelse(df$month==3,0,1)
+# df$month4 <- ifelse(df$month==4,0,1)
+# df$month5 <- ifelse(df$month==5,0,1)
+# df$month6 <- ifelse(df$month==6,0,1)
+# L <- L.c 
+# pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential")
+# infovars <- list(c("dlogdc", "logdc"),
+#                  c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
+# infovarsd <- list(c("dlogdd", "logdd"),
+#                   c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
+# bvars <- c("workplaces","retail","grocery","transit")
+# vars <- c(infovars[[2]],infovarsd[[2]],pols,"pindex")
+# for(i in 1:length(vars)) {
+#   L<-L.c
+#   v <- sprintf("%s.L14",vars[i])
+#   df[,v] <- panellag(df[,vars[i]],df$state,df$date, lag=L)
+#   L<-L.d
+#   v <- sprintf("%s.L21",vars[i])
+#   df[,v] <- panellag(df[,vars[i]],df$state,df$date, lag=L)
+# } 
+# for(i in 1:length(bvars)) {
+#   L<-L.c
+#   v <- sprintf("%s.L14",bvars[i])
+#   df[,v] <- panellag(df[,bvars[i]],df$state,df$date, lag=L+14)
+#   L<-L.d
+#   v <- sprintf("%s.L21",bvars[i])
+#   df[,v] <- panellag(df[,bvars[i]],df$state,df$date, lag=L+14)
+# } 
+# svars <- c("logpop","logsq","Percent.Unemployed..2018.",
+#                "Percent.living.under.the.federal.poverty.line..2018.",
+#                "Percent.at.risk.for.serious.illness.due.to.COVID","party_dum")
+# mvars.c <- mvars.d <- mvars <- c("month4","month5","month6")
+# # mvars.c <- c("month4.c","month5.c","month6.c")
+# # mvars.d <- c("month4.d","month5.d","month6.d")
+# ijs <- expand.grid(1:length(svars), 1:length(mvars.c))
+# msvars.c <- vector(mode="character", length=lengths(ijs)[[1]])
+# msvars.d <- vector(mode="character", length=lengths(ijs)[[1]])
+# for (ij in (1:lengths(ijs)[[1]])){
+#   v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars.c[ijs[[2]][ij]])
+#   msvars.c[ij] <- v
+#   df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars.c[ijs[[2]][ij]]]
+#   v <-  sprintf("%s.%s",svars[ijs[[1]][ij]],mvars.d[ijs[[2]][ij]])
+#   msvars.d[ij] <- v
+#   df[,v] <- df[,svars[ijs[[1]][ij]]]*df[,mvars.d[ijs[[2]][ij]]]
+# } 
+# bvars.c <- c("workplaces.L14","retail.L14","grocery.L14","transit.L14")
+# bvars.d <- c("workplaces.L21","retail.L21","grocery.L21","transit.L21")  
+# wvars <- c(svars,mvars.c,msvars.c,"month3","mask_percent","logvote","dlogtests")
+# wvars.d <- c(svars,mvars.d,msvars.d,"month3","mask_percent","logvote")
+# xvars <- c(bvars.c,"dlogdc.L14","logdc.L14")
+# xvars.national <- c(bvars.c,"dlogdc.L14","logdc.L14","dlogdc.national.L14","logdc.national.L14")
+# xvars.d <- c(bvars.d,"dlogdd.L21","logdd.L21")
+# xvars.national.d <- c(bvars.d,"dlogdd.L21","logdd.L21","dlogdd.national.L21","logdd.national.L21") 
+# pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pindex.L14")
+# pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pindex.L21")
 
@@
 }
-
-# DML2.for.PLM <- function(x, d, y, dreg, yreg, sdf, nfold=5) {
-#   unit <- as.double(sdf$state)
-#   j <-  quantile(unit, c(1:(nfold-1))/nfold)
-#   I <- vector(mode = "list", length = nfold)
-#   I[[1]] <- unit <= j[1]
-#   for (b in 2:(nfold-1)) {
-#     I[[b]] <- (unit <= j[b])&(unit > j[b-1])
-#   }
-#   I[[nfold]] <- unit > j[nfold-1]
-#   nobs <- nrow(x)
-#   ytil <- dtil <- rep(NA, nobs)
-#   for(b in 1:nfold){
-#     I1 <- (I[[b]]==FALSE)
-#     I2 <- I[[b]]
-#     dfit <- dreg(x[I1,], d[I1]) # take a fold out
-#     yfit <- yreg(x[I1,], y[I1]) # take a fold out
-#     dhat <- predict(dfit, x[I2,], type="response") #predict the fold out
-#     yhat <- predict(yfit, x[I2,], type="response") #predict the fold out
-#     dtil[I2] <- (d[I2] - dhat) #record residual
-#     ytil[I2] <- (y[I2] - yhat) #record residual
-#     cat(b," ")
-#   }
-# sdf$ytil <- ytil
-# sdf$dtil <- dtil
-# rfit <- felm(ytil ~ dtil| 0| 0 | state, data=sdf) #estimate the main parameter by regressing one residual on the other
-# coef.est <- coef(rfit)[2] #extract coefficient
-# se <- sqrt(vcov(rfit)[2,2])
-# cat(sprintf("\ncoef (se) = %g (%g)\n", coef.est , se))
-# return( list(coef.est = coef.est , se=se) )
-# }
 
@@
 fig
+
+# use residuals after controlling for confounders 
+wv <- c(svars,mvars.c,msvars.c,"month3") 
+fmla <- as.formula(paste("dlogdc", "~",  paste(wv, collapse=" + "), sep=" "))
+g1 <- predict(lm(fmla, data=df))
+df$dlogdc.res <- df[,yv] - g1    
+fig <- pfigure(df, "pmaskbus", stlab=-1,outcome="dlogdc.res",L=L.c) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("residual of dlogdc"))) + 
+  ylim(c(-1.8,0.5))  
+p <- "pmaskbus"
+figdev(sprintf("%s/tex/tables_and_figures/%s-cases-res-%s.pdf", rootdir, p, L.c), width=4, height=3)
+print(fig)
+dev.off()
+fig
+
+wv <- c(svars,mvars.d,msvars.d,"month3") 
+fmla <- as.formula(paste("dlogdd", "~",  paste(wv, collapse=" + "), sep=" "))
+g1 <- predict(lm(fmla, data=df))
+df$dlogdd.res <- df[,yv] - g1  
+fig <- pfigure(df, "pmaskbus", stlab=-1,outcome="dlogdd.res",L=L.d) + ggtitle(TeX(relabel("dlogdd given pmaskbus"))) + ylab(TeX(relabel("residual of dlogdd")))+ 
+  ylim(c(-0.9,0.5))  
+p <- "pmaskbus"
+figdev(sprintf("%s/tex/tables_and_figures/%s-deaths-res-%s.pdf", rootdir, p, L.d), width=4, height=3)
+print(fig)
+dev.off()
+fig
+
 ```

### 2020-09-04 tex/covidEffectOfNPI-revision.tex
updating Figure 10 with residual versions and regressionWithPlots

@@
 
+\afterpage{
+ \clearpage
 \begin{figure}[ht]
@@
     \end{flushleft}
+  \end{minipage}\\ \medskip
+ \begin{minipage}{\linewidth}
+    \centering
+    \begin{tabular}{cc}
+      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases-res-14}
+      &
+        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths-res-21}
+    \end{tabular}
+    \begin{flushleft}
+      \scriptsize In these figures, red points are the regression residuals of case or death
+      growth rate on confounders in states without a mask mandate. Blue points are
+      states with a mask mandate 14 (21 for deaths) days prior. The
+      red line is the average across states without a mask mandate 14
+      (21 for deaths) days earlier. The blue line is the average
+      across states with a mask mandate 14 (21 for deaths) earlier.
+    \end{flushleft}
   \end{minipage}
 \end{figure} 
+\clearpage}
 
@@
 growth rates. We begin with some simple graphical evidence of the
-effect of policies on case and death growth. Figure \ref{fig:masks}
-shows average case and death growth conditional on date and whether
+effect of policies on case and death growth. The upper two panels of figure \ref{fig:masks}
+show average case and death growth conditional on date and whether
 masks are mandatory for employees.\footnote{We take 14 and 21 day lags of mask policies for case and death growths, respectively, to  identify the states with a mask mandate because policies affect cases and deaths with time lags.
-See our discussion in the Appendix \ref{sec:timing}.
-} The left panel of the figure shows
-that states with a mask mandate consistently have 0-0.2 lower case
-growth than states without. The right panel  also illustrates that
+See our discussion in the Appendix \ref{sec:timing}.} The upper left panel of the figure shows that states with a mask mandate consistently have 0-0.2 lower case
+growth than states without. The upper right panel also illustrates that
 states with a mask mandate tend to have lower average death growth than states without a mask mandate.
+The lower two panels of figure \ref{fig:masks} are similar to the upper two panels except that we use the residuals of  regressing case or death growth on state-level characteristics, month dummies, and their interactions in place of  case or death growth, indicating that states with a mask mandate tend to have lower case and death growths even after controlling for confounders.
 

### 2020-09-04 cases_and_policies/rmd/regprep.R

@@ -91,12 +91,7 @@ bvars <- c("workplaces","retail","grocery","transit")
 
 for(i in 1:length(bvars)) df[,bvars[i]] <- df[,bvars[i]]/100
 
-
-#df$udistance <- panelma(df$daily_distance_diff,df$state, df$date, 6)
-#df$uvisits <- panelma(df$daily_visitation_diff,df$state, df$date, 6)
-#df$sghometime <- panelma(df$median_home_dwell_time, df$state, df$date, 6)
-#df$sgpart <- panelma(df$part_time_work_behavior_devices/df$device_count, df$state, df$date, 6)
-#df$sgfull <- panelma(df$full_time_work_behavior_devices/df$device_count, df$state, df$date, 6)
+ 
 
 nc <- aggregate(cases ~ date, function(x) sum(x, na.rm=TRUE), data=df)
 names(nc) <- c("date","cases.national")
@@ -124,14 +119,9 @@ df <- ndf[order(ndf$state, ndf$date),]
 df$z.mask <- df$z.mask/100
 df$mask_percent <- df$mask_percent/100
 
-df$party <- as.factor(df$party)
-
-# df$pindex <- (df$pshelter+df$pmovie+df$prestaurant+df$pnonessential)/4
-#df$pindex <- (df$pshelter+df$prestaurant+df$pnonessential)/3
-
+df$party <- as.factor(df$party) 
 df$pindex <- (df$pmovie+df$prestaurant+df$pnonessential)/3
-
-
+ 
 tilt = ifelse(df$date> as.Date("2020-05-01"),1,0 )
 month <- month(df$date)
 week <- week(df$date)
@@ -142,4 +132,67 @@ df$week = as.factor(week)
 df$pmask.may <- df$pmaskbus*tilt
 df$pmask.april <- df$pmaskbus*(1-tilt)
 pols <- c("pmaskbus","pk12","pshelter","pmovie","prestaurant","pnonessential")
-#pols <- c("pmaskbus","pk12","pindex")
+
+
+
+## the following variables are used for DML
+df$logpop <- log(df$Population.2018)
+df$logsq <- log(df$Square.Miles)
+df$repub <- as.factor(df$party)
+df$party_dum <- ifelse(df$party == "republican", 0, 1)
+df$month3 <- ifelse(df$month==3,0,1)
[…]

### 2020-09-04 tex/reply-to-referees.tex
update

@@
 \changes
+
+In addition to our reply to the referee report, we also attached our reply to the comments we have received from
+C. Jessica E. Metcalf (\url{https://metcalflab.princeton.edu/people/metcalf-c-jessica/}) who gave us a set of detailed comments via private communication.
+
 \subsection*{Reply to Editor's Comments}
@@
 This is an interesting suggestion.
-We thought about the possibility of providing a version of Figure 8 based on the regression residuals after taking out the effect of confounders etc but, eventually, we decided not to present such results due to the space constraint.
+We thought about the possibility of providing a version of Figure 8 (now, Figure 10) based on the regression residuals after taking out the effect of confounders etc but, eventually, we decided not to present such results due to the space constraint.
 
-The following figures are generated by first regressing the case or death growth rates on state characteristics, month dummies, and their interactions and then computing their average value for states with or without a mask mandate for each date.
+For your reference, we have constructed a version of  figure based on the residuals of regressing the case or death growth rates on state characteristics, month dummies, and their interactions, which illustrates that states with a mask mandate tend to have lower case and death growths even after controlling for confounders.
+% and then computing their average value for states with or without a mask mandate for each date.
 

### 2020-09-04 tex/covidEffectOfNPI-revision.tex
update

--- 2020-09-04 tex/covidEffectOfNPI-revision.tex
@@
     \end{flushleft}
-  \end{minipage}
+  \end{minipage}%\\ \medskip
+% \begin{minipage}{\linewidth}
+%    \centering
+%    \begin{tabular}{cc}
+%      \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-cases-res-14}
+%      &
+%        \includegraphics[width=0.5\textwidth]{tables_and_figures/pmaskbus-deaths-res-21}
+%    \end{tabular}
+%    \begin{flushleft}
+%      \scriptsize In these figures, red points are the regression residuals of case or death
+%      growth rate on confounders in states without a mask mandate. Blue points are
+%      states with a mask mandate 14 (21 for deaths) days prior. The
+%      red line is the average across states without a mask mandate 14
+%      (21 for deaths) days earlier. The blue line is the average
+%      across states with a mask mandate 14 (21 for deaths) earlier.
+%    \end{flushleft}
+%  \end{minipage}
 \end{figure}
+
+\subsection{The Direct Effect of Policies and Behavior on Case
+and Death Growth\label{policy-behavior-and-case-growth}}
+
+We now analyze how behavior and policies together influence case and death
+growth rates. We begin with some simple graphical evidence of the
+effect of policies on case and death growth. Fgure \ref{fig:masks}
+shows average case and death growth conditional on date and whether
+masks are mandatory for employees.\footnote{We take 14 and 21 day lags of mask policies for case and death growths, respectively, to  identify the states with a mask mandate because policies affect cases and deaths with time lags.
+See our discussion in the Appendix \ref{sec:timing}.} The   left panel of the figure shows that states with a mask mandate consistently have 0-0.2 lower case
+growth than states without. The  right panel also illustrates that
+states with a mask mandate tend to have lower average death growth than states without a mask mandate.
+
+%The lower two panels of figure \ref{fig:masks} are similar to the upper two panels except that we use the residuals of  regressing case or death growth on state-level characteristics, month dummies, and their interactions in place of  case or death growth, indicating that states with a mask mandate tend to have lower case and death growths even after controlling for confounders.
+
+Similar plots are shown for other policies in Figures
+\ref{fig:growthpolicies1} and \ref{fig:growthpolicies2} in the
+appendix. The  figures for stay-at-home orders and closure
+of nonessential businesses are qualitatively similar to that for
+masks. States with these two policies appear to have about 0.1 percentage point lower
+case growth  than states without. The effects of school closures, movie theater closures, and
+restaurant closures are not clearly visible in these figures. These
+figures are merely suggestive; the patterns observed in them may be
+driven by confounders.
 
@@
 }
-
-\subsection{The Direct Effect of Policies and Behavior on Case
-and Death Growth\label{policy-behavior-and-case-growth}}
-
-We now analyze how behavior and policies together influence case and death
-growth rates. We begin with some simple graphical evidence of the
-effect of policies on case and death growth. Figure \ref{fig:masks}
-shows average case and death growth conditional on date and whether
-masks are mandatory for employees.\footnote{We take 14 and 21 day lags of mask policies for case and death growths, respectively, to  identify the states with a mask mandate because policies affect cases and deaths with time lags.
-See our discussion in the Appendix \ref{sec:timing}.
-} The left panel of the figure shows
-that states with a mask mandate consistently have 0-0.2 lower case
-growth than states without. The right panel  also illustrates that
-states with a mask mandate tend to have lower average death growth than states without a mask mandate.
-
-Similar plots are shown for other policies in Figures
-\ref{fig:growthpolicies1} and \ref{fig:growthpolicies2} in the
-appendix. The  figures for stay-at-home orders and closure
-of nonessential businesses are qualitatively similar to that for
-masks. States with these two policies appear to have about 0.1 percentage point lower
-case growth  than states without. The effects of school closures, movie theater closures, and
-restaurant closures are not clearly visible in these figures. These
-figures are merely suggestive; the patterns observed in them may be
-driven by confounders.
 

### 2020-09-06 tex/reply-to-referees.tex
update

@@
 
-For your reference, we have constructed a version of  figure based on the residuals of regressing the case or death growth rates on state characteristics, month dummies, and their interactions, which illustrates that states with a mask mandate tend to have lower case and death growths even after controlling for confounders.
+For your reference, we have constructed a version of  figures based on the residuals of regressing the case or death growth rates on state characteristics, month dummies,   their interactions,  the log of lagged cases/deaths, and the growth rate of lagged cases/deaths, which illustrates that states with a mask mandate tend to have lower case and death growths even after controlling for confounders and information variables as below.
 % and then computing their average value for states with or without a mask mandate for each date.
@@
       \scriptsize In these figures, red points are the regression residuals of case or death
-      growth rate on confounders in states without a mask mandate. Blue points are
+      growth rate on confounders and information variables in states without a mask mandate. Blue points are
       states with a mask mandate 14 (21 for deaths) days prior. The

### 2020-09-06 cases_and_policies/rmd/regressionsWithPlots.Rmd
update

@@
 # use residuals after controlling for confounders
-wv <- c(svars,mvars.c,msvars.c,"month3") 
+
+sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
+sdf$one <- 1
+wv <- c(svars,mvars.c,msvars.c,"month3","one")
+wv <- c(svars,mvars.d,msvars.d,"month3","one","dlogdc.L14","logdc.L14")#,"dlogdc.national.L14","logdc.national.L14")
 fmla <- as.formula(paste("dlogdc", "~",  paste(wv, collapse=" + "), sep=" "))
-g1 <- predict(lm(fmla, data=df))
-df$dlogdc.res <- df[,yv] - g1    
-fig <- pfigure(df, "pmaskbus", stlab=-1,outcome="dlogdc.res",L=L.c) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("residual of dlogdc"))) + 
-  ylim(c(-1.8,0.5))  
+g1 <- predict(lm(fmla, data=sdf))
+sdf$dlogdc.res <- sdf$dlogdc - g1
+mean(sdf$dlogdc.res)
+fig <- pfigure(sdf, "pmaskbus", stlab=-1,outcome="dlogdc.res",L=L.c) + ggtitle(TeX(relabel("dlogdc given pmaskbus"))) + ylab(TeX(relabel("residual of dlogdc"))) + 
+ #   xlim(c(as.Date("2020-04-01"), max(df$date))) +
+  ylim(c(-0.8,0.3))  
 p <- "pmaskbus"
@@
 
-wv <- c(svars,mvars.d,msvars.d,"month3") 
+wv <- c(svars,mvars.d,msvars.d,"month3","one")
+wv <- c(svars,mvars.d,msvars.d,"month3","one","dlogdd.L21","logdd.L21")#,"dlogdd.national.L21","logdd.national.L21")
 fmla <- as.formula(paste("dlogdd", "~",  paste(wv, collapse=" + "), sep=" "))
-g1 <- predict(lm(fmla, data=df))
-df$dlogdd.res <- df[,yv] - g1  
-fig <- pfigure(df, "pmaskbus", stlab=-1,outcome="dlogdd.res",L=L.d) + ggtitle(TeX(relabel("dlogdd given pmaskbus"))) + ylab(TeX(relabel("residual of dlogdd")))+ 
-  ylim(c(-0.9,0.5))  
+g1 <- predict(lm(fmla, data=sdf))
+sdf$dlogdd.res <- sdf$dlogdd - g1  
+fig <- pfigure(sdf, "pmaskbus", stlab=-1,outcome="dlogdd.res",L=L.d) + ggtitle(TeX(relabel("dlogdd given pmaskbus"))) + ylab(TeX(relabel("residual of dlogdd")))+ 
+ #   xlim(c(as.Date("2020-04-01"), max(df$date))) +
+  ylim(c(-0.8,0.3))  
 p <- "pmaskbus"

### 2020-09-06 tex/covidEffectOfNPI-revision.tex
update

@@
 16-20 days after implementation, out of which 5.9 percentage points are attributed to shelter in place orders.
-Both \cite{hsiang2020} and \cite{courtemanche2020} adopted a reduced-form approach to estimate the total policy effect on case growth without using any social distancing behavior measures.
-In contrast, our study highlights the role of behavioral response to policies and information.
+Both \cite{hsiang2020} and \cite{courtemanche2020} adopted a reduced-form approach to estimate the total policy effect on case growth without using any social distancing behavior measures.\footnote{Using
+a synthetic control approach, \cite{cho2020} finds that  the cases would have been lower by 75 percent  had Sweden adopted stricter lockdown policies.}
 
@@
 Analyzing mitigation measures in New York, Wuhan, and Italy, \cite{zhangr2020} conclude that mandatory face coverings substantially reduced infections.
-\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} Our finding is also independently corroborated by a completely different causal methodology based on synthetic control
+\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} The Institute for Health Metrics and Evaluation at the University of Washington  assesses that, if 95\% of the people in the US were to start wearing masks from early August of 2020, 66,000 lives would be saved by December 2020 \citep{IHME2020}.
+Our finding is also independently corroborated by a completely different causal methodology based on synthetic control
 using German data in \cite{Mitze2020}.\footnote{Our study was first released in ArXiv on May 28, 2020 whereas
- \cite{Mitze2020} was released at SSRN on June 8, 2020.}
+ \cite{Mitze2020} was released at SSRN on June 8, 2020. }
 

### 2020-09-07 tex/covidEffectOfNPI-revision.tex
update after english editing

@@
 Our analysis finds that both policies and information on transmission risks are important determinants of Covid-19  cases and deaths and shows that a change in policies explains a large fraction of observed changes in social distancing behavior.
-Our main counterfactual experiments suggest that nationally mandating face masks for employees early in the pandemic could have reduced the weekly growth rate of cases and deaths  by more than 10 percentage points in late April, and could have led to as much as 19 to 47 percent less deaths nationally by the end of May, which roughly translates into 19 to 47 thousand saved lives.
+Our main counterfactual experiments suggest that nationally mandating face masks for employees early in the pandemic could have reduced the weekly growth rate of cases and deaths  by more than 10 percentage points in late April and could have led to as much as 19 to 47 percent less deaths nationally by the end of May, which roughly translates into 19 to 47 thousand saved lives.
 We also find that, without stay-at-home orders, cases would have been larger by  6 to 63 percent and without business closures, cases would have been larger by 17 to 78 percent.
@@
 Our approach explicitly recognizes that policies not only directly affect the spread of Covid-19 (e.g., mask requirement) but also indirectly affect its spread by changing people's behavior (e.g., stay-at-home order).
-It also recognizes that people react to new information on Covid-19 cases and deaths, and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
+It also recognizes that people react to new information on Covid-19 cases and deaths and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
 Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto}) and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
@@
 Our behavior variables are four mobility measures that capture the intensity of visits to ``transit,'' ``grocery,'' ``retail,'' and ``workplaces''  from Google Mobility Reports.
-We take the lagged growth rate of cases and deaths and the log of lagged cases and deaths at both state-level and national-level as our measures of information on infection risks that affects people's  behavior.
+We take the lagged growth rate of cases and deaths and the log of lagged cases and deaths at both the state-level and the national-level as our measures of information on infection risks that affects people's  behavior.
 We also consider the growth rate of tests, month dummies, and state-level characteristics (e.g., population size and total area) as confounders that have to be controlled for in order to identify the causal relationship between policy/behavior and the growth rate of cases and deaths.
@@
 The counterfactual experiments show a large impact of  those policies on  the number of cases and deaths.
-They also highlight the importance of voluntary behavioral response to infection risks for evaluating the dynamic policy effects.
+They also highlight the importance of voluntary behavioral response to infection risks when evaluating the dynamic policy effects.
 
@@
 One of the reason is that we chose to focus on a simplified version of the model 
-as our base model, by
-averaging business
-closure policies as measure of shutdown. This produced more stable results when
+as our base model, by averaging business closure policies as measure of shutdown. This produced more stable results when
 conducting sensitivity results, in particular using fixed effects estimators.  
@@
 with the previous findings, but we focus reporting counterfactuals only 
-for the base model. } This finding is
+for the base model. } 
+This finding is
 significant: given this potentially large benefit of reducing the
@@
 closure, and no closure of restaurants except take-out) could have
-affected cases and deaths.  We estimate that business shutdowns have a
+affected cases and deaths.  We estimate that business shutdowns have 
 roughly the same impact on growth rates as mask mandates, albeit with
@@
 In particular, we examined whether certain effect
-sizes can be ruled by various more flexible models or by using alternative timing assumptions
+sizes can be ruled out by various more flexible models or by using alternative timing assumptions
 that define forward growth rates.
-The impact of mask mandates is more robustly and more precisely estimated than that of business closure policies or stay-at-home orders, and an undesirable effect of increasing the weekly death growth by 5 percentage points is ruled out by any of the  models.\footnote{This null
+The impact of mask mandates is more robustly and more precisely estimated than that of business closure policies or stay-at-home orders, and an undesirable effect of increasing the weekly death growth by 5 percentage points is ruled out by all of the  models we consider.\footnote{This null
 hypothesis can be generated by looking at the meta-evidence from RCTs on the efficacy of masks in preventing other respiratory
@@
 The findings of shelter-in-place and business closures policies are considerably less robust.
-For example, for stay-at-home mandates, models with alternative timing and richer specification for information set suggested smaller effects, albeit
+For example, for stay-at-home mandates, models with alternative timing and richer specification for information set suggested smaller effects.
+Albeit
 after application of machine learning tools to reduce dimensionality, the range of effects $[0,0.15]$ could not be ruled out.  A similar
@@
 We also examine the impact of school closures. Unfortunately, there is
-very little variation across states in school closure in the timing of
+very little variation across states in the timing of
 school closures.  Across robustness specifications, we obtain point
-estimates of the effect of school closures as low as 0, and as high as
+estimates of the effect of school closures as low as 0 and as high as
 -0.6. In particular, we find that the results are sensitive to whether
@@
 specification or not. This highlights the uncertainty regarding the
-impact some policies versus private behavioral responses to
-information. Evaluation of re-opening policies needs to be aware of
-this uncertainty.
+impact of some policies versus private behavioral responses to
+information. %Evaluation of re-opening policies needs to be aware of this uncertainty.
 
@@
 of policies they study reduced growth rates by 9.1 percentage points
-16-20 days after implementation, out of which 5.9 percentage points are attributed to shelter in place orders.
-Both \cite{hsiang2020} and \cite{courtemanche2020} adopted a reduced-form approach to estimate the total policy effect on case growth without using any social distancing behavior measures.\footnote{Using
+16-20 days after implementation, out of which 5.9 percentage points are attributable to shelter in place orders.
+Both \cite{hsiang2020} and \cite{courtemanche2020} adopt a reduced-form approach to estimate the total policy effect on case growth without using any social distancing behavior measures.\footnote{Using
 a synthetic control approach, \cite{cho2020} finds that  the cases would have been lower by 75 percent  had Sweden adopted stricter lockdown policies.}
@@
 using mobility measures from PlaceIQ and SafeGraph.
-Using data from SafeGraph, \cite{anderson2020} show that there has been substantial voluntary social distancing but also provide evidence that mandatory measures such as stay-at-home orders have  been effective at reducing the frequency of visits  outside of one's home.
+Using data from SafeGraph, \cite{anderson2020} shows that there has been substantial voluntary social distancing but also provide evidence that mandatory measures such as stay-at-home orders have  been effective at reducing the frequency of visits  outside of one's home.
 % Overall, these evidence suggests the importance of voluntary social distancing behavior when we evaluate the effect of polices.
@@
 Epidemiologists use model simulations to predict how cases and deaths evolve for the purpose of policy recommendation.
-As reviewed by \cite{avery2020}, there exist substantial uncertainty about the values of key epidimiological parameters \citep[see also][]{atkeson2020b,stock2020}.
+As reviewed by \cite{avery2020}, there exists substantial uncertainty about the values of key epidimiological parameters \citep[see also][]{atkeson2020b,stock2020}.
 Simulations are often done under strong assumptions about the impact of social distancing policies without connecting to the relevant data  \citep[e.g.,][]{ferguson2020}.
@@
 \cite{Chu2020} conduct a meta-analysis of observational studies on transmission of the viruses that cause COVID-19 and related diseases and find the effectiveness of mask use for reducing transmission.
-} \cite{ollila2020} provide a meta-analyhsis of randomized controlled trials of non-surgical face masks in preventing viral respiratory infections in non-hospital and non-household settings, finding that face masks decreased infections across all five studies they reviewed.\footnote{Whether wearing masks creates a false sense of security and leads to decrease in social distancing is also a hotly debated topic.
+} \cite{ollila2020} provide a meta-analysis of randomized controlled trials of non-surgical face masks in preventing viral respiratory infections in non-hospital and non-household settings, finding that face masks decreased infections across all five studies they reviewed.\footnote{Whether wearing masks creates a false sense of security and leads to decrease in social distancing is also a hotly debated topic.
 A randomized field experiment in Berlin, Germany, conducted by \cite{seres2020}  finds that wearing masks actually increases social distancing, providing no evidence that mandatory masks leads to decrease in social distancing.}
@@
 
-Our empirical results contribute to informing the economic-epidemiological models that combine economic models with variants of SIR models to evaluate the efficiency of various economic policies aimed at gradual ``reopening" of various sectors of economy.\footnote{\cite{adda2016} analyzes the effect of policies reducing interpersonal contacts such as school closures or the closure of public transportation networks on the spread of  influenza, gastroenteritis, and chickenpox using high frequency data from France.} For example, the estimated effects of masks, stay-home mandates, and various other policies on behavior, and of behavior on infection can serve as useful inputs and validation checks in the calibrated macro, sectoral, and micro models (see, e.g., \cite{alvarez2020simple,baqaee2020reopening,NBERw27128,acemoglu2020multi,lsmith,nashSIR} and references therein).
+Our empirical results contribute to informing the economic-epidemiological models that combine economic models with variants of SIR models to evaluate the efficiency of various economic policies aimed at  the gradual ``reopening" of various sectors of economy.\footnote{\cite{adda2016} analyzes the effect of policies reducing interpersonal contacts such as school closures or the closure of public transportation networks on the spread of  influenza, gastroenteritis, and chickenpox using high frequency data from France.} For example, the estimated effects of masks, stay-home mandates, and various other policies on behavior, and of behavior on infection can serve as useful inputs and validation checks in the calibrated macro, sectoral, and micro models (see, e.g., \cite{alvarez2020simple,baqaee2020reopening,NBERw27128,acemoglu2020multi,lsmith,nashSIR} and references therein).
 Furthermore, the causal framework developed in this paper could be applicable, with appropriate extensions, to the impact of policies on economic outcomes replacing health outcomes (see, e.g., \cite{chetty2020real,coibion2020labor}).
@@
 
-Our main outcomes of interest are the growth rates in Covid-19 cases and deaths,  behavioral variables include proportion of time spent in transit or shopping and others, policy variables include stay-at-home orders and school and business closures, and the information variables include lagged values of outcome.
+Our main outcomes of interest are the growth rates in Covid-19 cases and deaths,  behavioral variables include proportion of time spent in transit, shopping, and workplaces, policy variables include mask mandates, stay-at-home orders, and school and business closures, and the information variables include lagged values of outcome.
 We provide a detailed description of these variables and their timing in the next section.
 
-The causal structure  allows for  the effect of the policy to be either direct or indirect -- through-behavior or through dynamics; and all of these effects are not mutually exclusive.
+The causal structure  allows for  the effect of the policy to be either direct or indirect -- through behavior or through dynamics;  all of these effects are not mutually exclusive.
 The structure also allows for changes in behavior to be brought by change in policies and information.
 These are all realistic properties that we expect from the contextual knowledge of the problem.
-Policies such as closures of schools, non-essential business, and restaurants, alter and constrain behavior in strong ways.
+Policies such as closures of schools, non-essential business, and restaurants alter and constrain behavior in strong ways.
 In contrast, policies such as mandating employees to wear masks can potentially affect the Covid-19 transmission directly.
@@
 
-The specification of the model above grasps one-period responses, but does not allow
-to answer about multi-period responses.  The dynamics will be induced by the evolution
-of information variables.
-Our baseline specification assumes that information variables include time, lagged and integrated values of outcome:\footnote{Our empirical analysis also considers a specification in which information variables include lagged national cases/deaths as well as lagged behavior variables.}
+The specification of the model above grasps one-period responses.  The dynamics over multiple periods will be induced by the evolution
+of information variables, which include time, lagged and integrated values of outcome:\footnote{Our empirical analysis also considers a specification in which information variables include lagged national cases/deaths as well as lagged behavior variables.}
 \begin{equation}\label{eq:I}\tag{I}
@@
 We model the reaction of people's behavior via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.
-The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$ in our context, since we are analyzing the pandemics from the very beginning, but other initializations could be appropriate in other contexts.
-The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
+The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$ in our context, since we are analyzing the pandemic from the very beginning, but other initializations could be appropriate in other contexts.
+% The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
 
@@
 Without any exclusion restrictions, this equality is just a decomposition of total effects
-into direct and indirect components, and is not a testable restriction. However, in our case we rely on the SIR
+into direct and indirect components and is not a testable restriction. However, in our case we rely on the SIR
 model with testing to motivate the presence of change in testing rate as a confounder in the outcome
-equations, but not in the behavior equation (therefore, a component of $\delta_B$ and $\delta_P$ is set to 0), implying that (TR) does not necessarily hold and is testable.
+equations but not in the behavior equation (therefore, a component of $\delta_B$ and $\delta_P$ is set to 0), implying that (TR) does not necessarily hold and is testable.
 Furthermore, we estimate (\ref{eq:R2}) on the data set that has many more observations than the data set used to estimate the outcome equations, implying that (TR) is again testable.
 Later we  shall also try to
-utilize the contextual knowledge that masks business mandates only affect outcome directly and not by changing mobility (i.e., $\beta=0$ for mask policy), implying again that (TR) is testable.
+utilize the contextual knowledge that mask mandates  only affect the outcome directly and not by changing mobility (i.e., $\beta=0$ for mask policy), implying again that (TR) is testable.
 If not rejected by the data, (TR) can be used to sharpen the estimate of the causal effect of mask policies on the outcomes.
 
-Validation of the model by (TR) allows us to check exclusion restrictions brought by contextual knowledge,
+Validation of the model by (TR) allows us to check exclusion restrictions brought by contextual knowledge
 and check stability of the model by using different data subsets.  However, passing the (TR) does
-not, of course, guarantee that the model is necessarily valid for recovery of causal effects.
-The only fundamental way to truly validate a causal model for observational data is through a controlled experiment, which is impossible to carry  in our setting.
-We shall however cross-check our results against other scientific evidence and, for example, results from RCT on effectiveness of masks on respiratory deceases (though not in the Covid-19 context).
+not guarantee that the model is necessarily valid for recovering causal effects.
+The only fundamental way to truly validate a causal model for observational data is through a controlled experiment, which is impossible to carry out in our setting.
+%We shall, however, cross-check our results  against other scientific evidence, such as results from an RCT on the effectiveness of masks on respiratory deaths.
 
@@
 We present the results
-on debiased fixed effect estimation with weekly dummies as parts of our sensitivity analysis. Other approaches we have explored
-is to consider more flexible nonlinear model $f(W_{it}, \delta)$ for the way observed confounders enter the model,
-for example partialling out these effects using Random Forest, using the debiased machine learning approach.
-All of these results are reported as parts of the robustness checks.
+on debiased fixed effect estimation with weekly dummies as parts of our sensitivity analysis.
+Our sensitivity analysis also considers a debiased machine learning approach  using Random Forest in which observed confounders enter the model nonlinearly.
+%Other approaches we have explored
+%is to consider more flexible nonlinear model $f(W_{it}, \delta)$ for the way observed confounders enter the model,
+%for example partialling out these effects using Random Forest, using the debiased machine learning approach.
+%All of these results are reported as parts of the robustness checks.
 
@@
 via generalized method of moments. We shall take a more pragmatic approach where we estimate
-the parameters of equations separately, and then compute
+the parameters of equations separately and then compute
 \[
@@
 reduce to $$  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} \text{ and } {\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'},$$ respectively.
-This construction is not as efficient as generalized method of moments,
+This construction is not as efficient as generalized method of moments
 but has a nicer interpretation under possible misspecification of the model: we are combining predictions
@@
 } Given the policies, we generate the counterfactual
-outcomes, behavior, information by propaging the dynamic equations:
+outcomes, behavior, and information by propagating the dynamic equations:
 \begin{equation} \label{eqn:CF-SEM} \tag{CF-SEM}
@@
 with the same initialization as the factual system up to $t \leq 0$.
-In stating this counterfactual system of equations, we assume that structural outcome equations (SO) and information equations (I) remain invariant, and so do the stochastic shocks, decomposed
+In stating this counterfactual system of equations, we assume that structural outcome equations (SO) and information equations (I) remain invariant and so do the stochastic shocks, decomposed
 into observable and unobservable parts. Formally, we record this assumption and above discussion as follows.
@@
 cases, $T_{it}$ is the number of tests over 7 days, $\Delta$ is
-a 7-days differencing operator, $\epsilon_{it}$ is an unobserved error term.
+a 7-days differencing operator, and $\epsilon_{it}$ is an unobserved error term.
  $X_{i,t-14}$  collects other behavioral, policy, and confounding variables, depending
@@
 
-The specification for growth rate in deaths as outcome  is motivated by SIRD as follows. By differentiating (\ref{eq:d}) and (\ref{eq:c}) with respect to $t$ and using (\ref{eq:c2}), we obtain
+The specification for growth rate in deaths as the outcome  is motivated by SIRD as follows. By differentiating (\ref{eq:d}) and (\ref{eq:c}) with respect to $t$ and using (\ref{eq:c2}), we obtain
 \begin{align}
@@
 \end{align}
-Our measurement equation for the growth rate of deaths is based on equation (\ref{eq:d2}) but   account for a $21$ day lag between infection and death as
+Our measurement equation for the growth rate of deaths is based on equation (\ref{eq:d2}) but   accounts for a $21$ day lag between infection and death as
 \begin{align}
@@
 policies:  stay-at-home, closed nonessential
-businesses, closed K-12 schools, closed restaurants except takeout, closed movie theaters, and mandate face mask by  employees in public facing businesses.
+businesses, closed K-12 schools, closed restaurants except takeout, closed movie theaters, and face mask mandates for employees in public facing businesses.
 We believe that the first four of these
@@
 We also obtain information on state-level covariates mostly from \citet{raifman2020}, which include population size, total area, unemployment rate, poverty rate,  a percentage of people who are subject to illness, and state governor's party affiliations.
-These confounders are motivated by \cite{wheaton2020} who finds that case growth is associated  with residential density and per capita income.
+These confounders are motivated by \cite{wheaton2020} who find that case growth is associated  with residential density and per capita income.
 
@@
 \(t-6\).  We focus on weekly cases and deaths because daily new cases and deaths are
-affected by the timing of reporting and testing, and are quite
+affected by the timing of reporting and testing and are quite
 volatile as shown in the upper right panel of Figure \ref{fig:casedeathtest} in the
@@
 behavior variables by taking 7 day moving averages from day \(t-14\) to
-\(t-21\) for case growth, where the  delay reflects the time lag between infection and case confirmation. The four weekly behavior variables are referred as ``Transit
+\(t-21\) for case growth, where the  delay reflects the time lag between infection and case confirmation. The four weekly behavior variables are referred  to ``Transit
 Intensity,'' ``Workplace Intensity,'' ``Retail Intensity,'' and ``Grocery
@@
 %, but this does not prevent us from identifying the aggregate effect of all policies and behavior variables on case or death growth.
-For this reason, we define  ``business closure policies'' variable by the average of closured movie theaters, closed  restaurants, and closured non-essential businesses variables, and consider a specification that includes business closure policies   in place of these three policy variables separately.
+For this reason, we define  the ``business closure policies'' variable by the average of closed movie theaters, closed  restaurants, and closed non-essential businesses variables and consider a specification that includes business closure policies   in place of these three policy variables separately.
 
@@
 The other policy coefficients are less sensitive to the inclusion of
-national case/death variables. After school closures, business closure policies has the next largest effect followed by stay-at-home
+national case/death variables. After school closures, business closure policies have the next largest effect followed by stay-at-home
 orders.  The effect of masks for employees is
@@
 growth rates. We begin with some simple graphical evidence of the
-effect of policies on case and death growth. Fgure \ref{fig:masks}
+effect of policies on case and death growth. Figure \ref{fig:masks}
 shows average case and death growth conditional on date and whether
@@
 We also consider a specification that includes the past cases and case growth at the national level as additional information variables.
-$\wcolor W_{it}$ is a set of confounders that include month dummies, state-level covariates,  and the interaction terms between month dummies and state-level covariates.\footnote{Month dummies also represent the latent information that is not fully captured by the past cases and growths.}
+$\wcolor W_{it}$ is a set of confounders that includes month dummies, state-level covariates,  and the interaction terms between month dummies and state-level covariates.\footnote{Month dummies also represent the latent information that is not fully captured by the past cases and growths.}
 For case growth, $\wcolor W_{it}$ also includes the  test rate growth  $\Delta \log(T)_{it}$  to capture the effect of changing test rates on confirmed cases.
@@
 The estimates of $-0.80$ and $-0.84$ for ``$\sum_k w_k \mathrm{Behavior}_k$'' in column (1)  imply that a reduction in mobility measures relative to the baseline in January and February have induced a decrease in case and death growth rates by 80 and 84 percent, respectively, suggesting an importance of social distancing for reducing the spread of Covid-19.
-When including national cases and deaths in information, as shown in columns (3) and (4), the estimated aggregate impact of behavior is substantially smaller, but remains large and statistically significant.
-
-A useful practical implication of these results are that Google Mobility Reports and similar data might be useful as a leading indicator of potential case or death growth.
+When including national cases and deaths in information, as shown in columns (3) and (4), the estimated aggregate impact of behavior is substantially smaller but remains large and statistically significant.
+
+A useful practical implication of these results is that Google Mobility Reports and similar data might be useful as a leading indicator of potential case or death growth.
 This should be done with caution, however, because other changes in the environment might alter the relationship between behavior and infections.
@@
      coefficients.
-``business closure policies'' is defined by the average of   closured movie theaters, closured restaurants, and closured non-essential businesses.
+``business closure policies'' is defined by the average of   closed movie theaters, closed restaurants, and closed non-essential businesses.
 Standard errors are clustered at the state level.
@@
 
-Similarly to Table  \ref{tab:dieff-si}, the first three columns of  Table \ref{tab:dieff-con-si} report  the estimated direct and indirect effects of policy but imposing that masks for employees only affect cases/deaths directly without affecting behavior and that business closure policies only affect  cases/deaths indirectly through their effects on behavior.\footnote{These are computed from estimating the specification in columns (1)-(4) of
+Similarly to Table  \ref{tab:dieff-si}, the first three columns of  Table \ref{tab:dieff-con-si} report  the estimated direct and indirect effects of policy but impose that masks for employees only affect cases/deaths directly without affecting behavior and that business closure policies only affect  cases/deaths indirectly through their effects on behavior.\footnote{These are computed from estimating the specification in columns (1)-(4) of
 Table \ref{tab:PItoB} and column (1) of Table \ref{tab:BPItoY} but imposing that the coefficient of masks for employees is zero in (\ref{eq:R1}) and that the coefficient of business closure policies is zero in (\ref{eq:R2}).} The estimated total effect of masks for employees in the third column of Table \ref{tab:dieff-con-si} is higher than that of Table  \ref{tab:dieff-si}.
-Similarly, the total effect of business closure policies is estimated larger in Table \ref{tab:dieff-con-si} than in  Table  \ref{tab:dieff-si}.
+Similarly, the total effect of business closure policies is estimated to be larger in Table \ref{tab:dieff-con-si} than in  Table  \ref{tab:dieff-si}.
 
@@
 (\ref{eq:R1})-(\ref{eq:R2}) in   column ``Total.''   Differences  are generally small except for the coefficient of closed K-12 schools and  the sum of all policies in Table   \ref{tab:dieff-si}.
-The large differences in school closures  may be due to the  difficulty in identifying the effect of school closures given a lack of cross-sectional variations.
+The large differences in school closures  may be due to the  difficulty in identifying the effect of school closures given a lack of cross-sectional variation.
 Imposing the coefficients of masks for employees and business closure policies to be zero in (\ref{eq:R1}) and  (\ref{eq:R2}), respectively, does not increase differences between  ``Total" and  ``PI$\to$Y Coefficient''  as reported in the last column of  Table \ref{tab:dieff-con-si}.
@@
   \item[(2)]  Exclude  the state of New York from the sample because it may be viewed as an outlier in the early pandemic period.
-  \item[(3)]   Add  the percentage of people who wears masks to protect themselves in March and April as confounder for unobserved  personal risk-aversion and initial attitude toward mask wearing.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
+  \item[(3)]   Add  the percentage of people who wear masks to protect themselves in March and April as a confounder for unobserved  personal risk-aversion and initial attitude toward mask wearing.\footnote{ The survey is conducted online by YouGov and is based on the interviews of 89,347 US adults aged 18 and over between March 26-April 29, 2020.
 The survey question is ``Which, if any, of the following measures have you taken in the past 2 weeks to protect yourself from the Coronavirus (COVID-19)?''.}
-   \item[(4)]   Add the log of Trump's vote share  in 2016 presidential election as confounder for unobserved private behavioral response.
+   \item[(4)]   Add the log of Trump's vote share  in the 2016 presidential election as a confounder for unobserved private behavioral response.
    \item[(5)]   Add  past behavior variables  as information used to set policies.
@@
 
-Panel (C) of Figure  \ref{fig:whisker} indicates that the estimated coefficients of stay-at-home order  are generally negative and often significant, providing evidence that stay-at-home orders reduce the spread of COVID-19, although the estimates are sometimes sensitive to timing assumptions.
+Panel (C) of Figure  \ref{fig:whisker} indicates that the estimated coefficients of stay-at-home orders  are generally negative and often significant, providing evidence that stay-at-home orders reduce the spread of COVID-19, although the estimates are sometimes sensitive to timing assumptions.
 Panel (D) shows that the estimated coefficients of business closure policies substantially vary across specifications, providing a mixed evidence for the effect of closures of businesses on case/death growth.
@@
 
-The estimated coefficients of masks for employees largely confirm our finding that mandatory mask policy reduces the case and death growths.
+The estimated coefficients of masks for employees largely confirm our finding that mandatory mask policies reduce the case and death growths.
 The estimated coefficients of stay-at-home orders and business closures are negative in columns (1) and (2) but their magnitudes as well as statistical significance are somewhat sensitive to whether bias corrections are applied or not.
-This sensitivity may be partly due to a high correction between stay-at-home orders variable and business closure policies.
+% This sensitivity may be partly due to a high correction between stay-at-home orders variable and business closure policies.
 
@@
 pointwise 90\% confidence band for the national total change. The
-estimates that a mask mandate in mid March would have decreases cases
+estimates that a mask mandate in mid March would have decreased cases
 and deaths by about 6000 and 600 respectively throughout April and
@@
 counterfactual experiment suggests that
-if all states had have adopted mandatory face mask policies on March 14th of 2020, then the cumulative number of deaths by the end of May would have been smaller by as much as  $19$ to $47$\%, which roughly translates into $19$ to $47$  thousand saved lives.
+if all states had  adopted mandatory face mask policies on March 14th of 2020, then the cumulative number of deaths by the end of May would have been smaller by as much as  $19$ to $47$\%, which roughly translates into $19$ to $47$  thousand saved lives.
 
@@
 
-Fourth, our analysis shows that people voluntarily reduce their visits to workplace, retails, grocery stores, and limit their use of public transit when they receive  information on a higher number of new cases and deaths.
+Fourth, our analysis shows that people voluntarily reduce their visits to workplace, retail stores, grocery stores, and people limit their use of public transit when they receive  information on a higher number of new cases and deaths.
 This suggests that individuals make decisions to voluntarily limit their contact with others in response to greater transmission risks, leading to an important feedback mechanism that affects future cases and deaths.
@@
 The upper left panel of Figure \ref{fig:casedeathtest} shows the evolution of new cases in the NYT data, where daily changes in cumulative
-cases displays some excessive volatility. This is likely due to delays
+cases  exhibit some excessive volatility. This is likely due to delays
 and bunching in testing and reporting of results.
@@
 For most of our results, we focus on new cases in a week instead of in
-a day. We do this for two reasons as discussed in the main text. First, a decline of new cases over
+a day. We do this for two reasons as discussed in the main text. First, a decline in new cases over
 two weeks has become a key metric for decision makers. Secondly, aggregating to weekly new cases smooths out the noise associated with  the timing of
@@
       growth rate in states without each policy 14 (or 21 for deaths)
-      days earlier.  Blue points are states with each policy14 (or 21
+      days earlier.  Blue points are states with each policy 14 (or 21
       for deaths) days earlier.  The red line is the average across

### 2020-09-07 tex/covidEffectOfNPI-revision.tex
update

@@
 confidence intervals of $[9,32]$\% and $[19,47]$\%, which roughly
-implies that 34 thousand lives could have been saved.\footnote{These estimates
-are slightly lower than those reported in earlier releases of the paper on ArXiv.
-One of the reason is that we chose to focus on a simplified version of the model 
-as our base model, by averaging business closure policies as measure of shutdown. This produced more stable results when
-conducting sensitivity results, in particular using fixed effects estimators.  
-We also present several variants of base model, including
-one with separate shutdown policies, in which the effects are slightly larger, in line
-with the previous findings, but we focus reporting counterfactuals only 
-for the base model. } 
+implies that 34 thousand lives could have been saved.
+%\footnote{These estimates
+%are slightly lower than those reported in earlier releases of the paper on ArXiv.
+%One of the reason is that we chose to focus on a simplified version of the model 
+%as our base model, by averaging business closure policies as measure of shutdown. This produced more stable results when
+%conducting sensitivity results, in particular using fixed effects estimators.  
+%We also present several variants of base model, including
+%one with separate shutdown policies, in which the effects are slightly larger, in line
+%with the previous findings, but we focus reporting counterfactuals only 
+%for the base model. } 
 This finding is
@@
 \]
-with approximate deviations controlled by the normal laws, with standard deviations that can be approximated by the bootsrap
+with approximate deviations controlled by the normal laws, with standard deviations that can be approximated by the bootstrap
 resampling of observational units $i$. Under correct specification the target quantities
-reduce to $$  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} \text{ and } {\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'},$$ respectively.
-This construction is not as efficient as generalized method of moments
+reduce to $$  {\bcolor \alpha '}  {\pcolor  \beta' }+{\pcolor \pi'} \text{ and } {\bcolor  \alpha '}  {\icolor  \gamma'} + {\icolor  \mu'},$$ respectively.\footnote{This construction is not as efficient as generalized method of moments
 but has a nicer interpretation under possible misspecification of the model: we are combining predictions
 from two models, one motivated via the causal path, reflecting contextual knowledge, and another from a ``reduced form" model not exploiting the path.
-The combined estimator can improve on precision of either estimator.
+The combined estimator can improve on precision of either estimator.}
 

### 2020-09-07 tex/covidEffectOfNPI-revision.tex
update

@@
 Analyzing mitigation measures in New York, Wuhan, and Italy, \cite{zhangr2020} conclude that mandatory face coverings substantially reduced infections.
-\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} The Institute for Health Metrics and Evaluation at the University of Washington  assesses that, if 95\% of the people in the US were to start wearing masks from early August of 2020, 66,000 lives would be saved by December 2020 \citep{IHME2020}.
+\cite{abaluck2020}  find that the growth rates of cases and of deaths in countries with pre-existing norms that sick people should wear masks are lower by 8 to 10\% than those rates in countries with no pre-existing mask norms.\footnote{\cite{miyazawa2020} find that  country's COVID-19 death rates are negatively associated with mask wearing rates.} The Institute for Health Metrics and Evaluation at the University of Washington  assesses that, if 95\% of the people in the US were to start wearing masks from early August of 2020, 66,000 lives would be saved by December 2020 \citep{IHME2020}, which is largely consistent with our results.
 Our finding is also independently corroborated by a completely different causal methodology based on synthetic control

### 2020-09-07 tex/covidEffectOfNPI-revision.tex
minor update

@@
 It also recognizes that people react to new information on Covid-19 cases and deaths and voluntarily adjust their behavior (e.g., voluntary social distancing and hand washing) even without any policy in place.
-Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.\footnote{The causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika, heckman:pinto}) and genetics \citep{wright1923}, with natural unfolding potential/structiral outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika}, \cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak}, \cite{peters2020book}, and \cite{hernanrobins2020book}, with major applications taking place in computer science, econometrics, and epidemiology.
-The particular causal diagram has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
-The father and son, P.
-Wright and S.
-Wright closely collaborated to develop structural equation models and causal path diagrams; P.
-Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables.}
+Our casual model provides a framework to quantitatively decompose the growth of Covid-19 cases and deaths into three components: (1) direct policy effect, (2) policy effect through behavior, and (3) direct behavior effect in response to new information.
 
@@
 confidence intervals of $[9,32]$\% and $[19,47]$\%, which roughly
-implies that 34 thousand lives could have been saved.
-%\footnote{These estimates
-%are slightly lower than those reported in earlier releases of the paper on ArXiv.
+implies that 34 thousand lives could have been saved. %\footnote{These estimates
+%are slightly lower than those reported in earlier releases of the paper on ArXiv,
 %One of the reason is that we chose to focus on a simplified version of the model 
@@
 Furthermore, the causal framework developed in this paper could be applicable, with appropriate extensions, to the impact of policies on economic outcomes replacing health outcomes (see, e.g., \cite{chetty2020real,coibion2020labor}).
+
+Finally, our causal model is framed using the language of structural equations models and causal diagrams of econometrics (\cite{pwright, haavelmo:1944, tinbergen:1940, wold:1954, pearl:biometrika}) and genetics \citep{wright1923},\footnote{The father and son, P.
+Wright (economist) and S.
+Wright (geneticist) collaborated to develop structural equation models and causal path diagrams; P.
+Wright's key work represented supply-demand system as a directed acyclical graph and established its identification using exclusion restrictions on instrumental variables.
+We view our work as following this classical tradition.} with natural unfolding potential/structural outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
+The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika} and others (\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak},  \cite{heckman:pinto}, \cite{peters2020book}, and \cite{hernanrobins2020book}), with major applications taking place in computer science, epidemiology, and econometrics.
+The particular causal diagram we use has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
 
@@
 where we say that $V \perp U$ if $\Ep VU = 0$.\footnote{ An alternative useful
-starting point is to impose the Rubin-Rosenbaum type unconfoudedness condition:
+starting point is to impose the Rosenbaum-Rubin type unconfoundedness condition:
 $$
@@
 $$
-which imply, with treating stochastic errors as independent additive components, the orthogonal conditions stated above.} This is a standard way of representing restrictions on errors in structural equation modeling.\footnote{The structural equations of this form are connected to triangular structural equation models, appearing in microeconometrics and macroeconometrics (SVARs), going back to the work of  \cite{strotz1960recursive}.}
+which imply, with treating stochastic errors as independent additive components, the orthogonal conditions stated above.
+The same unconfoundedness restrictions can be formulated using formal causal DAGs, and also imply orthogonality restrictions stated above, once stochastic errors are modeled as independent additive components.} This is a standard way of representing restrictions on errors in structural equation modeling in econometrics.\footnote{The structural equations of this form are connected to triangular structural equation models, appearing in microeconometrics and macroeconometrics (SVARs), going back to the work of  \cite{strotz1960recursive}.}
 
@@
  for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
- and $t/\ell$ is defined as $\lfloor {t/\ell} \rfloor$.
+ and $t/\ell$ is defined as $\lceil {t/\ell} \rceil$.
 
@@
 Let $i \in \{1,..., N\}$ denote the observational unit,  $t$ be the time periods, and $\ell$ be the time delay.
-(1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined outside of the model for $t \leq 0$.
-(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E).
+(1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined by factors outside of the model for $t \leq 0$.
+(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined by factors outside of the mode, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E).
 \end{quote}
@@
 Counterfactual experiments with endogenous government policy would be important, for example, to understand the issues related to the lagged response of government policies to higher infection rates due to incomplete information.
-} Given the policies, we generate the counterfactual
+}
+The assumption  of invariance captures the idea that counterfactual policy interventions would not change the structural functions within the period of the study.
+The assumption is strong but is necessary to conduct
+counterfactual experiments, e.g.
+\cite{sims1972money} and \cite{strotz1960recursive}.
+It captures the idea that counterfactual policy interventions would not change the structural functions within the period of the study.
+To make the assumption more plausible we limited our study to the early pandemic period.\footnote{Furthermore, we conducted several stability checks, for example, checking if the coefficients on mask policies remain stable (reported in the previous version) and also looking at more recent data during reopening, beyond early pandemic, to examine the stability of the model.}
+
+Given the policies, we generate the counterfactual
 outcomes, behavior, and information by propagating the dynamic equations:
@@
 Let $i \in \{1,..., N\}$ be the observational unit, $t$ be time periods, and $\ell$ be the time delay.
-(1) For each $i$ and $t \leq 0$, the confounder, information, behavior, policy, and outcome variables are determined as in SEM: $W^\star_{it}= W_{it}$, $I^\star_{it}= I_{it}$,  $B^\star_{it} = B_{it}$, $P^\star_{it}=P_{it}$, $Y^\star_{it} =Y_{it}$.
-(2) For each $i$ and $t \geq 0$, confounders $W^\star_{it} =W_{it}$ are determined, and information variables $I^\star_{it}$ are determined  by (I);  policy variables $P^\star_{it}$ are set in (\ref{p-cf}); behavior variables $B^\star_{it}$ are determined by setting $\iota = I^\star_{it}$ and $p= P^\star_{it}$ in  (SO) with  the same stochastic shock $\varepsilon^b_{it}$ in (SO); the  counterfactual outcome $Y^\star_{i, t + \ell}$ is realized by setting $\iota = I^\star_{it}$, $p= P^\star_{it}$, and $b = B^\star_{it}$ in  (SO) with the same stochastic shock $\varepsilon^y_{it}$ in  (SO).
+(1) For each $i$ and $t \leq 0$, the confounder, information, behavior, policy, and outcome variables are determined as previously stated in SEM: $W^\star_{it}= W_{it}$, $I^\star_{it}= I_{it}$,  $B^\star_{it} = B_{it}$, $P^\star_{it}=P_{it}$, $Y^\star_{it} =Y_{it}$.
+(2) For each $i$ and $t \geq 0$, confounders $W^\star_{it} =W_{it}$ are determined as in SEM, and information variables $I^\star_{it}$ are determined  by (I);  policy variables $P^\star_{it}$ are set in (\ref{p-cf}); behavior variables $B^\star_{it}$ are determined by setting $\iota = I^\star_{it}$ and $p= P^\star_{it}$ in  (SO) with  the same stochastic shock $\varepsilon^b_{it}$ in (SO); the  counterfactual outcome $Y^\star_{i, t + \ell}$ is realized by setting $\iota = I^\star_{it}$, $p= P^\star_{it}$, and $b = B^\star_{it}$ in  (SO) with the same stochastic shock $\varepsilon^y_{it}$ in  (SO).
 \end{quote}
@@
  $$
- interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.
- The effects can be decomposed into (a) delayed policy effects via behavior by
+ interpreting $t/\ell$ as $\lceil t/\ell \rceil$, where $\mathrm{PEB}^\star_{i,t } =  \mathrm{PED}^\star_{i,t} = 0$
+ for $t \leq 0$.  The effects can be decomposed into (a) delayed policy effects via behavior by
 summing terms containing $PEB$ and (b) delayed policy effects via direct impact
@@
 All counterfactual quantities and contrasts can be computed from the expressions given
-above. For examples, given $\Delta C_{i0}>0$, new cases are linked to growth rates via relation:
+above. For examples, given $\Delta C_{i0}>0$, new confirmed cases are linked to growth rates via relation:
 $$
@@
 The cumulative cases can be constructed by summing over the new cases. Various contrasts are
-then calculated from these quantities. For example, the relative contrast of new counterfactual cases
-to the  factual cases is given by:
+then calculated from these quantities. For example, the relative contrast of counterfactual new confirmed cases
+to the  factual confirmed cases is given by:
 $$

### 2020-09-08 tex/covidEffectOfNPI-revision.tex
initialization update

@@
 
+%The specification of the model above grasps one-period responses.  The dynamics over multiple periods will be induced by the evolution
+%of information variables, which include time, lagged and integrated values of outcome:\footnote{Our empirical analysis also considers a specification in which information variables include lagged national cases/deaths as well as lagged behavior variables.}
+%\begin{equation}\label{eq:I}\tag{I}
+ %{\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  S_{i, t, \ell } +  \sum_{m=1}^{ \lfloor t/\ell \rfloor}
+%{\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}
+% for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
+%where $ S_{i, t, \ell }$ is the initial condition, the log of new cases at time $t - \lfloor t/\ell \rfloor$.
+
 The specification of the model above grasps one-period responses.  The dynamics over multiple periods will be induced by the evolution
@@
 \begin{equation}\label{eq:I}\tag{I}
- {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{t/\ell}
+ {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=1}^{ \lfloor t/\ell \rfloor}
 {\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}
  for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
- and $t/\ell$ is defined as $\lceil {t/\ell} \rceil$.
+ assuming that the log of new cases at time $ - t \ \mathrm{mod} \ \ell $ is zero, for notational convenience.
+ \footnote{The general formula  for  $I_{i, t-1}$ is $
+S_{i, t, \ell } + \sum_{m=1}^{ \lfloor t/\ell \rfloor} Y_{i,t - \ell m}$, where $ S_{i, t, \ell }$ is the initial condition, the log of new cases at time $- t \ \mathrm{ mod } \ \ell$.}
 
@@
 We model the reaction of people's behavior via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.
-The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$ in our context, since we are analyzing the pandemic from the very beginning, but other initializations could be appropriate in other contexts.
-% The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
+The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$.
+The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
 
@@
 we also add weekly dummies).  The stochastic shocks $\{ \varepsilon_{it}\}_{t=1}^T$ are treated as independent across states $i$ and can be arbitrarily dependent across time $t$ within a state.
+%:
 
@@
 counterfactual experiments, e.g.
-\cite{sims1972money} and \cite{strotz1960recursive}.
+\cite{sims1972} and \cite{strotz1960recursive}.
 It captures the idea that counterfactual policy interventions would not change the structural functions within the period of the study.
@@
  $$
- interpreting $t/\ell$ as $\lceil t/\ell \rceil$, where $\mathrm{PEB}^\star_{i,t } =  \mathrm{PED}^\star_{i,t} = 0$
- for $t \leq 0$.  The effects can be decomposed into (a) delayed policy effects via behavior by
+ interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.  %, where $\mathrm{PEB}^\star_{i,t } =  \mathrm{PED}^\star_{i,t} = 0$
+ %for $t \leq 0$.  
+ The effects can be decomposed into (a) delayed policy effects via behavior by
 summing terms containing $PEB$ and (b) delayed policy effects via direct impact
@@
 All counterfactual quantities and contrasts can be computed from the expressions given
-above. For examples, given $\Delta C_{i0}>0$, new confirmed cases are linked to growth rates via relation:
+above. For examples, given $\Delta C_{i0}>0$, new confirmed cases are linked to growth rates via relation
+(taking $t$ divisible by $\ell$ for simplicity):
 $$
-\frac{\Delta C^{\star}_{it}}{\Delta C_{i0}} = \exp\left (  \sum_{s=1}^t  Y^\star_{is} \right) \text{ and }\frac{ \Delta C_{it}}{\Delta C_{i0}} = \exp
-\left (  \sum_{s=1}^t Y_{is}\right).
+\frac{\Delta C^{\star}_{it}}{\Delta C_{i0}} = \exp\left (  \sum_{m=1}^{t/\ell}  Y^\star_{i,m\ell} \right) \text{ and }\frac{ \Delta C_{it}}{\Delta C_{i0}} = \exp
+\left (  \sum_{m=1}^{t/\ell}   Y_{i, m \ell }\right).
 $$
@@
 $$
-\Delta C^{\star}_{it}/\Delta C_{it} = \exp\left(  \sum_{s=1}^t ( Y^\star_{is} -  Y_{is}) \right).
+\Delta C^{\star}_{it}/\Delta C_{it} = \exp\left(  \sum_{m=1}^{t/\ell} ( Y^\star_{i,m\ell} -  Y_{i,m \ell}) \right).
 $$

### 2020-09-08 tex/covidEffectOfNPI-revision.tex
info  update

@@
 \begin{equation}\label{eq:I}\tag{I}
- {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=1}^{ \lfloor t/\ell \rfloor}
+ {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-\ell} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{ \lfloor t/\ell \rfloor}
 {\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}

### 2020-09-08 tex/covidEffectOfNPI-revision.tex
info  update

--- 2020-09-08 tex/covidEffectOfNPI-revision.tex
@@
 \begin{equation}\label{eq:I}\tag{I}
- {\icolor I_{it} } := I(t, {\ycolor Y_{it}},  { \icolor  I_{i, t-1} })' := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=1}^{ \lfloor t/\ell \rfloor}
-{\ycolor Y_{i,t - \ell m}} \Big )  \end{equation}
+ {\icolor I_{it} } := I_t( {\ycolor Y_{it}},  { \icolor  I_{i, t-\ell} }) := \Big (g(t) , {\ycolor Y_{it}},  \sum_{m=0}^{ \lfloor t/\ell \rfloor}
+{\ycolor Y_{i,t - \ell m}} \Big )'  \end{equation}
  for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
@@
  \footnote{The general formula  for  $I_{i, t-1}$ is $
-S_{i, t, \ell } + \sum_{m=1}^{ \lfloor t/\ell \rfloor} Y_{i,t - \ell m}$, where $ S_{i, t, \ell }$ is the initial condition, the log of new cases at time $- t \ \mathrm{ mod } \ \ell$.}
-
-In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases.
+S_{i, t, \ell } + \sum_{m=1}^{ \lfloor t/\ell \rfloor} Y_{i,t - \ell m}$, where $ S_{i, t, \ell }$ is the initial condition, the log of new cases at time $- t \ \mathrm{ mod } \ \ell$.} In this structure, people respond to both global information, captured by a function of time such as month dummies, and local information sources, captured by the local growth rate and the total number of cases.
 The local information also captures the persistence of the growth rate process.
 We model the reaction of people's behavior via the term ${\icolor \gamma'I_{t}}$ in the behavior equation.
-The dynamical system requires initialization, for example we can set  ${\ycolor Y_{it}} = 0$ for $t \leq 0$.
 The  lagged values of behavior variable may be also included in the information set, but we postpone this discussion after the main empirical results are presented.
@@
 (1) For each $i$ and $t \leq -\ell$, the confounder, information, behavior, and policy variables $W_{it}, I_{it},  B_{it}, P_{it}$ are determined outside of the model, and the outcome variable $Y_{i, t+\ell}$ is determined by factors outside of the model for $t \leq 0$.
-(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined by factors outside of the mode, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E).
+(2) For each $i$ and $t \geq -\ell$, confounders $W_{it}$ are  determined by factors outside of the model, and information variables $I_{it}$ are determined  by (I); policy variables $P_{it}$ are determined by setting $\iota = I_{it}$  in (P) with  a realized stochastic shock $\varepsilon^p_{it}$ that obeys the exogeneity condition (E);  behavior variables $B_{it}$ are determined by setting $\iota = I_{it}$ and $p= P_{it}$ in  (SO) with a shock $\varepsilon^b_{it}$ that obeys (E); finally, the outcome $Y_{i, t + \ell}$ is realized by setting $\iota = I_{it}$, $p= P_{it}$, and $b = B_{it}$ in  (SO) with a shock $\varepsilon^y_{it}$ that obeys (E).
 \end{quote}
@@
 & {\bcolor B^\star_{it} } & := &   B_{it}({\pcolor P^\star_{it} } ,{\icolor I^\star_{it}}), \\
-& {\icolor I^\star_{it} }& := &  I ({ t,  {\ycolor Y^\star_{it}}, \icolor I^\star_{i,t-\ell}}),  \end{aligned}
+& {\icolor I^\star_{it} }& := &  I_t ( {\ycolor Y^\star_{it}}, \icolor I^\star_{i,t-\ell}),  \end{aligned}
 \end{equation}

### 2020-09-09 tex/covidEffectOfNPI-revision.tex
tiny  update

@@
 %Letting $C_{it}$ denotes cumulative number of  confirmed cases in state $i$ at time $t$, our outcome
+%:
 %\begin{equation} \label{eq:y}
@@
 \cite{sims1972} and \cite{strotz1960recursive}.
-It captures the idea that counterfactual policy interventions would not change the structural functions within the period of the study.
 To make the assumption more plausible we limited our study to the early pandemic period.\footnote{Furthermore, we conducted several stability checks, for example, checking if the coefficients on mask policies remain stable (reported in the previous version) and also looking at more recent data during reopening, beyond early pandemic, to examine the stability of the model.}
@@
  of the policy.
-
-The dynamic effects take the form:
-$$
-\begin{array}{lll}
-  & \mathrm{DynE}^\star_{it} &=
-  \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}^\star_{i,t-m \ell } +  \mathrm{PED}^\star_{i,t- m\ell} \right)
-  \\
- &&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}^\star_{i,t-(m +1) \ell } +  \mathrm{PED}^\star_{i,t- (m+1)\ell}  \right),
- \end{array}
- $$
- interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.  %, where $\mathrm{PEB}^\star_{i,t } =  \mathrm{PED}^\star_{i,t} = 0$
+By recursive substitutions the dynamic effect  can be further decomposed into a weighted sum of delayed policy effects via behavior and a weighted sum of delayed policy effects via direct impact.
+
+%The dynamic effects take the form:
+%$$
+%\begin{array}{lll}
+ % & \mathrm{DynE}^\star_{it} &=
+ % \sum_{m=1}^{t/\ell} \left( (\gamma \alpha)_2 + \mu_2+  (\gamma \alpha)_3+ \mu_3    \right)^m \left(\mathrm{PEB}^\star_{i,t-m \ell } +  \mathrm{PED}^\star_{i,t- m\ell} \right)
+ % \\
+ %&&  + \sum_{m=1}^{t/\ell-1} \left( (\gamma \alpha)_3 + \mu_3    \right)^m \left (\mathrm{PEB}^\star_{i,t-(m +1) \ell } +  %\mathrm{PED}^\star_{i,t- (m+1)\ell}  \right),
+% \end{array}
+ %$$
+ %interpreting $t/\ell$ as $\lfloor t/\ell \rfloor$.  %, where $\mathrm{PEB}^\star_{i,t } =  \mathrm{PED}^\star_{i,t} = 0$
  %for $t \leq 0$.
- The effects can be decomposed into (a) delayed policy effects via behavior by
-summing terms containing $PEB$ and (b) delayed policy effects via direct impact
-by summing terms containing $PED$.
 

### 2020-09-09 tex/covidEffectOfNPI-revision.tex
tiny  update

@@
 We view our work as following this classical tradition.} with natural unfolding potential/structural outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika} and others (\cite{pearl:robins}, \cite{pearl:causality}, \cite{pearl:why}, \cite{white:chalak},  \cite{heckman:pinto}, \cite{peters2020book}, and \cite{hernanrobins2020book}), with major applications taking place in computer science, epidemiology, and econometrics.
+The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika} and others (\cite{pearl:robins,pearl:causality, pearl:why,white:chalak,heckman:pinto,hunermund,richardson:mediation,peters2020book,bareinboim:H,hernanrobins2020book}), with major applications taking place especially in computer science and epidemiology.
 The particular causal diagram we use has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
@@
 
-To estimate the coefficient of the $j$-th policy variable, $P_{it}^j$, using Double Machine Learning of \cite{chernozhukov18}, we consider a version of (\ref{eq:R4}) as follows:
+To estimate the coefficient of the $j$-th policy variable, $P_{it}^j$, using double machine learning (\cite{chernozhukov18}), we consider a version of (\ref{eq:R4}) as follows:
 \begin{align*}

### 2020-09-09 tex/covidEffectOfNPI-revision.tex
tiny  update

@@
 We view our work as following this classical tradition.} with natural unfolding potential/structural outcomes representation \citep{rubin1974,tinbergen1930,neyman:PO,imbens_rubin_2015}.
-The work on causal graphs has been modernized and put on rigorous footing by  \cite{pearl:biometrika} and others (\cite{pearl:robins,pearl:causality, pearl:why,white:chalak,heckman:pinto,hunermund,richardson:mediation,peters2020book,bareinboim:H,hernanrobins2020book}), with major applications taking place especially in computer science and epidemiology.
+The work on causal graphs has been modernized and developed by \cite{pearl:biometrika,pearl:robins,pearl:causality,pearl:why} and many others (e.g., \cite{pearl:why,white:chalak,richardson:mediation,peters2020book,bareinboim:H,hernanrobins2020book}), with applications in computer science, genetics, epidemiology, and econometrics (see, e.g., \cite{heckman:pinto,hunermund,white:chalak} for applications in econometrics).
 The particular causal diagram we use has several ``mediation" components, where variables affect outcomes directly and indirectly through other variables called mediators; these structures go back at least to \citet[][see Figure 6]{wright1923}; see, e.g., \cite{baron1986}, \cite{Hines2020}, \cite{richardson:mediation} for modern treatments.
@@
  for  each  $t \in \{0,1,...,T\}$, where $g$ is deterministic function of time, e.g., month indicators,
- assuming that the log of new cases at time $ - t \ \mathrm{mod} \ \ell $ is zero, for notational convenience.
+ assuming that the log of new cases at time $t \leq 0$ is zero, for notational convenience.
  \footnote{The general formula  for  $I_{i, t-1}$ is $

### 2020-09-11 cases_and_policies/R/countyData.R
Added county-level mask mandate data.

@@ -1,0 +1,157 @@
+## Data on state cases, deaths, etc from NYTimes
+nyt  <- read.csv("https://raw.githubusercontent.com/nytimes/covid-19-data/master/us-counties.csv", stringsAsFactors=FALSE)
+nyt$date <- as.Date(nyt$date)
+
+
+## County Mask Mandate info from Wright, Chawla, Chen, & Farmer
+library(readxl)
+if (!(file.exists("mm.xlsx")) {
+  download.file("https://drive.google.com/uc?export=download&id=1qVIhPaBQ-apdDjOaKV2eA9SgZNkLMLAm",destfile="mm.xlsx")
+} else {
+  warning("mm.xlsx exists. Not redownloading mask mandate data.")
+}
+
+mask <- data.frame(read_excel("mm.xlsx", col_names=FALSE))
+names(mask) <- c("state.fips", # 1
+                 "state",      # 2
+                 "fips",       # 3
+                 "county",     # 4
+                 "start_county", # 5
+                 "end_county",   # 6
+                 "scope_county", # 7
+                 "source_county", #8
+                 "note1_county", # 9
+                 "note2_county", # 10
+                 "code1", # 11  (this is some stata encoding of a date, I think)
+                 "start_state", # 12
+                 "end_state",  # 13
+                 "scope_state", # 14
+                 "source_state", # 15
+                 "code2",
+                 "code3",
+                 "code4",
+                 "start_mask") # min(start_county, start_state)
+
+for (v in grep("start|end",names(mask))) {
+  # let's enter dates in 3 different formats for some reason ...
+  if (is.character(mask[,v])) {
+    if (any(grepl("/",mask[,v])))
+      mask[,v] <- as.Date(mask[,v],format="%m/%d/%y")
+    else if (any(grepl("-",mask[,v])))
+      mask[,v] <- as.Date(mask[,v],format="%m-%d-%y")
+  } else
+    mask[,v] <- as.Date(mask[,v])
+}
+stopifnot(all(pmin(mask$start_county, mask$start_state) == mask$start_mask, na.rm=TRUE))
+
+## Corrections and additions to mask data
+
[…]

### 2020-09-11 cases_and_policies/R/utils.R

@@ -1,3 +1,5 @@
+library(ggplot2)
+library(ggthemes)
 #' Create lags for panel data.
 #'
 #' This function creates lags (or leads) of panel data variables.
@@ -70,7 +72,8 @@ panelma <- function(x, id, t, len=1) {
 }
 
 
-smoothedpolicy <- function(df, pdatevar, enddatevar=NA, type="pnorm", bw=7, lag=0) {
+smoothedpolicy <- function(df, pdatevar, enddatevar=NA, type="pnorm",
+  bw=7, lag=0, id="state") {
   if (type=="pnorm") {
     p <- as.numeric(pnorm(df$date, df[,pdatevar], bw))
     p[is.na(p)] <- 0.0
@@ -83,9 +86,9 @@ smoothedpolicy <- function(df, pdatevar, enddatevar=NA, type="pnorm", bw=7, lag=
       over[is.na(over)] <- FALSE
       d[over] <- FALSE
     }
-    p <- panelma(d, df$state, df$date, len=bw)
+    p <- panelma(d, df[,id], df$date, len=bw)
   }
-  if (lag != 0) p  <- panellag(p, df$state, df$date, lag=lag)
+  if (lag != 0) p  <- panellag(p, df[,id], df$date, lag=lag)
   return(p)
 }
 

### 2020-09-18 tex/covidEffectOfNPI-revision.tex
Changed tex file to use the 'elsarticle' class.

@@
-\documentclass[11pt,reqno,letter]{amsart}
+\documentclass[3p, longtitle]{elsarticle}
 %\pdfoutput=1
@@
 \usepackage{hyperref}
-\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
+%\usepackage[left=1.25in,right=1.25in,top=1.5in,bottom=1.25in]{geometry}
 \usepackage{multirow}
@@
 \usetikzlibrary{positioning}
-\geometry{letterpaper}
+%\geometry{letterpaper}
 %\usepackag\Ep[parfill]{parskip}    % Activate to begin paragraphs
@@
 
-\setlength{\parskip}{.4cm}
-\geometry{centering}
+%\setlength{\parskip}{.4cm}
+%\geometry{centering}
 \usepackage{mathtools}
@@
 
-\usepackage[authoryear]{natbib}
+%\usepackage[authoryear]{natbib}
 \usepackage{pdflscape}
@@
 
-\title[Causal Impact of Masks, Policies, Behavior]
-{Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.}\thanks{We are grateful to Daron Acemoglu, V.V.
+%\title[Causal Impact of Masks, Policies, Behavior]{Causal Impact of Masks, Policies, Behavior on Early Covid-19 Pandemic in the U.S.}
+\title{Causal Impact of Masks, Policies, Behavior on Early Covid-19
+  Pandemic in the U.S.}
+
+%% Formatting instructions move this to the end
+% \thanks{We are grateful to Daron Acemoglu, V.V.
 Chari, Raj Chetty,  Christian Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green,  Ido Rosen, Konstantin Sonin, James Stock, Elie Tamer, and Ivan Werning for helpful comments.
 We also thank Chiyoung Ahn,
-Joshua Catalano,  Jason Chau, Samuel Gyetvay, Sev Chenyu Hou,
-Jordan Hutchings, and Dongxiao Zhang  for excellent research assistance.  We are especially grateful
-to   C. Jessica E. Metcalf and an anonymous referee for detailed and insightful comments,
-which considerable improved the paper. All mistakes are our own. }
+% Joshua Catalano,  Jason Chau, Samuel Gyetvay, Sev Chenyu Hou,
+% Jordan Hutchings, and Dongxiao Zhang  for excellent research assistance.  We are especially grateful
+% to   C. Jessica E. Metcalf and an anonymous referee for detailed and insightful comments,
+% which considerable improved the paper. All mistakes are our own. }
 
 \author{Victor Chernozhukov}
-\address{Department of Economics and Center for Statistics and Data Science, MIT,  MA 02139}
-\email{vchern@mit.edu}
-%\urladdr{www.mit.edu/$\sim$vcherm} % Delete if not wanted.
-\author{Hiroyuki Kasahara}
-\address{ Vancouver School of Economics, UBC, 6000 Iona Drive, Vancouver, BC.}
-\email{hkasahar@mail.ubc.ca}
-
+\ead{vchern@mit.edu}
+%\urladdr{www.mit.edu/$\sim$vchern} % Delete if not wanted.
+\address{Department of Economics and Center for Statistics and Data
+  Science, MIT,  MA 02139}
+
+\author{Hiroyuki Kasahara \corref{cor1}}
+\ead{hkasahar@mail.ubc.ca}
+\address{Vancouver School of Economics, UBC, 6000 Iona Drive, Vancouver, BC.}
 \author{Paul Schrimpf}
-\address{ Vancouver School of Economics, UBC, 6000 Iona Drive, Vancouver, BC.}
-\email{schrimpf@mail.ubc.ca}
+\ead{schrimpf@mail.ubc.ca}
+\address{Vancouver School of Economics, UBC, 6000 Iona Drive, Vancouver, BC.}
+
+\cortext[cor1]{Corresponding author}
 
@@
 
-\keywords{Covid-19, causal impact, masks, non-essential business, policies, behavior}
+\begin{keyword}
+  Covid-19, causal impact, masks, policies, behavior
+\end{keyword}
 
@@
 
+\clearpage
 \begin{figure}[ht]
@@
 
+\section*{Acknowledgements}
+
+We are grateful to Daron Acemoglu, V.V. Chari, Raj Chetty, Christian
+Hansen, Glenn Ellison, Ivan Fernandez-Val, David Green, Ido Rosen,
+Konstantin Sonin, James Stock, Elie Tamer, and Ivan Werning for
+helpful comments. We also thank Chiyoung Ahn, Joshua Catalano, Jason
+Chau, Samuel Gyetvay, Sev Chenyu Hou, Jordan Hutchings, and Dongxiao
+Zhang for excellent research assistance.  We are especially grateful
+to C. Jessica E. Metcalf and an anonymous referee for detailed and
+insightful comments, which considerable improved the paper. All
+mistakes are our own.
+
 \begin{footnotesize}
@@
 \FloatBarrier
+
+\section{Additional Results}
 

### 2020-09-18 tex/covidEffectOfNPI-revision.tex
Added url of git repo to data section.
Fixed typo in acknowledgements.

@@
 % to   C. Jessica E. Metcalf and an anonymous referee for detailed and insightful comments,
-% which considerable improved the paper. All mistakes are our own. }
+% which considerably improved the paper. All mistakes are our own. }
 
@@
 \section{Empirical Analysis}
+
 \subsection{Data}
 
-Our baseline measures for daily Covid-19   cases and deaths are from
+\footnote{All code is available at \url{https://github.com/ubcecon/covid-impact}.}Our baseline measures for daily Covid-19  cases and deaths are from
 \href{https://github.com/nytimes/covid-19-data}{The New York Times (NYT)}.
@@
 to C. Jessica E. Metcalf and an anonymous referee for detailed and
-insightful comments, which considerable improved the paper. All
+insightful comments, which considerably improved the paper. All
 mistakes are our own.

### 2020-11-28 cases_and_policies/rmd/counterfactuals.Rmd
Minor changes in code to make sure it works with versions of packages in the renv environment.

@@
 source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
+source(paste(rootdir,"cases_and_policies/R/utils.R", sep="/"))
 source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
@@
 
-```{r testsmooth, cache=FALSE}
-source("smoothtests.R")
-```
-
 ```{r statesim, cache=FALSE}
@@
 L <- 14
-sdf <- subset(df, df$date>=datelims[1]) #+L)
+sdf <- subset(df, as.vector(df$date)>=datelims[1]) #+L)
 regs <- mainregressions(sdf,
@@
   cfnoplist <- lapply(unique(sdf$state), function(s)
-    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf, nresid=1))
+    nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
   allmeans <- rbindlist(lapply(cfnoplist, getmeans))

### 2020-11-28 cases_and_policies/rmd/regressionsWithPlots.Rmd
Minor changes in code to make sure it works with versions of packages in the renv environment.

@@
 source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
+source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
 source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
@@
 ```{r testsmooth, cache=FALSE}
-source("smoothtests.R")
+source(paste(rootdir,"cases_and_policies/rmd/smoothtests.R", sep="/"))
 ```
@@
 L <- L.c
-sdf <- subset(df, df$date>=as.Date("2020-03-07"))
+sdf <- subset(df, as.vector(df$date)>=as.Date("2020-03-07"))
 regs <- mainregressions(sdf,

### 2020-11-28 cases_and_policies/R/dataprep.R

@@ -4,7 +4,7 @@ rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
 datafile <- paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/")
 cat("Loading data last updated on ", as.character(file.mtime(datafile)), "\n")
 cat("Run ", paste(rootdir,"cases_and_policies/R/cases_and_policies.R", sep="/"), " to update data.\n")
-load(paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/"))
+base::load(paste(rootdir,"cases_and_policies/data/covidstates.Rda", sep="/"))
 source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
 
 cs <- covidstates

### 2020-11-28 cases_and_policies/R/utils.R

@@ -10,6 +10,9 @@
 #' values are desired.
 #' @return Lagged copy of x.
 panellag <- function(x, i, t, lag=1) {
+  x <- as.vector(x)
+  i <- as.vector(i)
+  t <- as.vector(t)
   if (!identical(order(i,t),1:length(i))) {
     stop("inputs not sorted.")
   }
@@ -22,7 +25,7 @@ panellag <- function(x, i, t, lag=1) {
   if (lag>0) {
     x.lag <- x[1:(length(x)-lag)]
     x.lag[i[1:(length(i)-lag)]!=i[(1+lag):length(i)] ] <- NA
-    x.lag[t[1:(length(i)-lag)]+lag!=t[(1+lag):length(i)] ] <- NA
+    x.lag[(t[1:(length(i)-lag)]+lag)!=t[(1+lag):length(i)] ] <- NA
     val <- (c(rep(NA,lag),x.lag))
   } else if (lag<0) {
     lag <- abs(lag)

### 2020-11-28 cases_and_policies/rmd/counterfactuals.R

@@ -58,6 +58,7 @@ figdatelims <- c(as.Date("2020-03-07"), max(df$date))
 createforwardsim <- function(m, df, id="state",
                        dvar="dlogdc", var="logdc",
                        difflag=7, lhsdiff=TRUE) {
+  df <- df[order(df[,id], df$t),]
   stopifnot(m$keepX)
   m$coef[is.nan(m$coef)] <- 0
 
@@ -127,7 +128,7 @@ createforwardsim <- function(m, df, id="state",
 
     date <- df$date[idx[df[idx,id]==idvar]]
     names(date) <- NULL
-    it0 <- which(df$date==date[1] & df[,id]==idvar)
+    it0 <- which(as.vector(df$date)==date[1] & as.vector(df[,id])==idvar)
     #t0 <- which(df$date==min(df$date))
     Y0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,var], k)[it0]))
     dY0 <- as.vector(sapply(1:(maxlags), function(k) lag(df[,dvar], k)[it0]))

### 2020-11-28 cases_and_policies/rmd/generatetables.R

@@ -78,8 +78,8 @@ policyreg <- function(df, # data
   w <- rep(1, length(p))
   peff <- c(sum(coef(m)[p]*w), sqrt(t(w) %*% (vcov(m)[p,p]) %*% w ))
   if (!is.null(bvars)) {
-    w <- colMeans(subset(df, df$date>=as.Date("2020-04-01") &
-                             df$date<=as.Date("2020-04-10"))[,bvars])
+    w <- colMeans(subset(df, as.vector(df$date)>=as.Date("2020-04-01") &
+                             as.vector(df$date)<=as.Date("2020-04-10"))[,bvars])
     beff <- c(sum(coef(m)[b]*w), sqrt(t(w) %*% (vcov(m)[b,b]) %*% w ))
   } else {
     beff <- NULL

### 2020-11-28 cases_and_policies/rmd/regprep.R

@@ -1,10 +1,9 @@
 cs$t <- as.numeric(cs$date)
 
 warning("Using data only up to 2020-06-03. Modify lines 4-5 of regprep.R to change")
-df <- pdata.frame(subset(cs, cs$date>=as.Date("2020-01-01") &
-                                   cs$date<=as.Date("2020-06-03")),
-                  # cs$date<=as.Date("2020-06-30")),
-                  index=c("state","t"), stringsAsFactors=FALSE)
+df <- subset(cs, cs$date>=as.Date("2020-01-01") &
+                                   cs$date<=as.Date("2020-06-03"))
+df <- df[order(df$state, df$date),]
 
 #df$tests[df$testmissing] <- 0
 df$testrate  <- df$tests/df$Population.2018*1000
@@ -113,7 +112,7 @@ nc$dlogdd.national <- dlogd(nc$deaths.national, nc$state, nc$date, lag=L,
 nc$state <- NULL
 
 ndf <- merge(ndf, nc, by="date", all=TRUE)
-ndf <- pdata.frame(ndf, index=c("state","t"))
+#ndf <- pdata.frame(ndf, index=c("state","t"))
 df <- ndf[order(ndf$state, ndf$date),]
 
 df$z.mask <- df$z.mask/100
@@ -196,3 +195,4 @@ xvars.national.d <- c(bvars.d,"dlogdd.L21","logdd.L21","dlogdd.national.L21","lo
 pols.L14 <- c("pmaskbus.L14","pk12.L14","pshelter.L14","pindex.L14")
 pols.L21 <- c("pmaskbus.L21","pk12.L21","pshelter.L21","pindex.L21")
  
+df <- pdata.frame(df, index=c("state","t"), stringsAsFactors=FALSE)

### 2021-02-23 cases_and_policies/rmd/generatetables.R
Added note to README about difference betwen paper tables and results from this code.
Improved renv usage to ensure all dependencies included and all code
works.

@@ -132,7 +132,7 @@ mainregressions <- function(df, # data
   })
 
   sdf <- df
-  sdf$month <- panellag(sdf$month,sdf$state,sdf$date,-L)
+  sdf$month <- as.factor(panellag(sdf$month,sdf$state,sdf$date,-L))
   ijs <- expand.grid(1:length(bvars), 1:length(plist))
   pib <- list()
   for (k in 1:length(xlist)) {
@@ -340,7 +340,7 @@ mainregressions2 <- function(df, # data
 
   sdf <- df
   sdf$pmaskbus <- 0*sdf$pmaskbus
-  sdf$month <- panellag(sdf$month,sdf$state,sdf$date,-L)
+  sdf$month <- as.factor(panellag(sdf$month,sdf$state,sdf$date,-L))
   ijs <- expand.grid(1:length(bvars), 1:length(plist))
   pib <- list()
   for (k in 1:length(xlist)) {

### 2021-12-21 cases_and_policies/rmd/cf-pindex.Rmd

@@
 if (!file.exists("cfmask_idx.Rda") || rerunsim) {
-  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
+  cfmasklist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
@@
 if (!file.exists("cfshelter_idx.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
+  cfshelterlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
@@
 if (!file.exists("cfindex_idx.Rda") || rerunsim) {
-  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
+  cfindexlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
   indexall <- rbindlist(lapply(cfindexlist, getmeans))
@@
 if (!file.exists("cfmask_deaths_idx.Rda") || rerunsim) {
-  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
+  cfmasklist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
@@
 if (!file.exists("cfshelter_deaths_idx.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
+  cfshelterlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
@@
 if (!file.exists("cfindex_deaths_idx.Rda") || rerunsim) {
-  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
+  cfindexlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
   indexall <- rbindlist(lapply(cfindexlist, getmeans))

### 2021-12-21 cases_and_policies/rmd/counterfactuals.Rmd

@@
 if (!file.exists("cfnop.Rda") || rerunsim) {
-  cfnoplist <- lapply(unique(sdf$state), function(s)
+  cfnoplist <- lapply(unique(as.character(sdf$state)), function(s)
     nopstatesim(s,simobs=simmodel,simcf=nopmodel,data=sdf))
@@
 if (!file.exists("cfnb.Rda") || rerunsim) {
-  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, nresid=1))
+  cfnblist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, nresid=1))
   nball <- rbindlist(lapply(cfnblist, getmeans))
@@
 if (!file.exists("cfmask.Rda") || rerunsim) {
-  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, nresid=1))
+  cfmasklist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, nresid=1))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
@@
 if (!file.exists("cfshelter.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
+  cfshelterlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf, nresid=1))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
@@
 if (!file.exists("cfindex.Rda") || rerunsim) {
-  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf, nresid=1))
+  cfindexlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf, nresid=1))
   indexall <- rbindlist(lapply(cfindexlist, getmeans))
@@
 if (!file.exists("cfnop_deaths.Rda") || rerunsim) {
-  cfnoplist <- lapply(unique(df$state), function(s)
+  cfnoplist <- lapply(unique(as.character(df$state)), function(s)
     nopstatesim(s,simobs=simmodel,simcf=nopmodel, data=sdf, var="logdd", dvar="dlogdd"))
@@
 if (!file.exists("cfnb_deaths.Rda") || rerunsim) {
-  cfnblist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
+  cfnblist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=nobmodel, data=sdf, var="logdd", dvar="dlogdd"))
   nball <- rbindlist(lapply(cfnblist, getmeans))
@@
 if (!file.exists("cfmask_deaths.Rda") || rerunsim) {
-  cfmasklist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
+  cfmasklist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=maskmodel, data=sdf, var="logdd", dvar="dlogdd"))
   maskall <- rbindlist(lapply(cfmasklist, getmeans))
@@
 if (!file.exists("cfshelter_deaths.Rda") || rerunsim) {
-  cfshelterlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
+  cfshelterlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noshelter, data=sdf))
   shelterall <- rbindlist(lapply(cfshelterlist, getmeans))
@@
 if (!file.exists("cfindex_deaths.Rda") || rerunsim) {
-  cfindexlist <- lapply(unique(df$state), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
+  cfindexlist <- lapply(unique(as.character(df$state)), function(s) nopstatesim(s,simobs=simmodel,simcf=noindex, data=sdf))
   indexall <- rbindlist(lapply(cfindexlist, getmeans))

### 2021-12-21 cases_and_policies/rmd/datasummary.Rmd

@@
 ---
-title       : ""
+title       : "Data"
 subtitle    : "Data description and summary statistics"

### 2021-12-21 cases_and_policies/rmd/regressionsWithPlots.Rmd

@@
 ```{r setup, cache=FALSE, warning=FALSE, include=FALSE}
-
+library(conflicted)
+unloadNamespace("dplyr") 
+unloadNamespace("dotwhisker")
 library(lfe)
@@
 
-# WARNINGS!  dplyr and dotwhisker seem to interfare with our regression codes --- need to call them only after calling regression codes; otherwise, they will improperly change our regression results
+# WARNINGS!  dplyr and dotwhisker seem to interfare with our
+# regression codes --- need to call them only after calling regression
+# codes; otherwise, they will improperly change our regression results
 library(dplyr)
@@
   df$y <- df[,outcome]
-  df$week <- week(df$date)
+  df$week <- data.table::week(df$date)
   m <- lm(y ~ as.factor(date) + as.factor(date):p, data=df)

### 2021-12-21 cases_and_policies/rmd/replytolemoine.Rmd

---
title       : "Causal impact of masks, policies, behavior on early COVID-19 pandemic in the U.S."
subtitle    : "Response to Lemoine"
author      :  Victor Chernozhukov, Paul Schrimpf, Hiro Kasahara
date        : "`r format(Sys.time(), '%d %B, %Y')`"
bibliography: "covid.bib"
link-citations: true
always_allow_html: true
output      :
    html_document :
        toc : true
        toc_depth : 3
        toc_float : true
        number_sections : true
        #theme : journal
        #css : 628notes.css
        code_folding: hide
        #lib_dir : deps
        self_contained: yes
        mode: selfcontained
        fig_width: 8
        fig_height: 6
---

 <!-- To create html files from this, in R enter  -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd") -->
 <!-- To create tex files from this, in R enter ` -->
 <!-- library(rmarkdown) -->
 <!-- rmarkdown::render("regressionsWithPlots.Rmd", output_format=latex_fragment()) -->

 <!-- to create the gh-pages hosted website in R enter -->
 <!-- rmarkdown::render_site() -->

 <!-- Then in a shell  -->
 <!-- $ ghp-import -p -n -m "$(date)" _site -->

```{r, include=FALSE}
options(knitr.table.format = function() {
  if (knitr::is_latex_output()) 'latex' else 'pandoc'
})
if (knitr::is_latex_output()) {
  knitr::opts_chunk$set(warning=FALSE, message=FALSE, echo=FALSE)
}
```

Recently, Phillippe Lemoine posted a critique of our
paper  \citep{chernozhukov2021}   at his post titled \href{https://cspicenter.org/blog/waronscience/lockdowns-econometrics-and-the-art-of-putting-lipstick-on-a-pig/}{``Lockdowns, econometrics and the art of putting lipstick on a pig''}.
Although Lemoine's critique appears ideologically driven and overly emotional, some of his points are worth addressing. In particular,  the sensitivity of our estimation results for (i)  including ``masks in public spaces”  and (ii) updating the data seems important critiques.

# Mask Mandates for employees versus for all

In our paper we analyze the effect of mask mandates for employees in
public facing businesses. Our dataset also contains information on
mask mandates for all people in public spaces. Our decision to focus
on the former and not the later is largely a historical
coindcidence. We used the excellent [“COVID-19 US state policy
database.” from  Raifman J, Nocka K, Jones D, Bor J, Lipson S, Jay J,
and Chan P. (2020).](www.tinyurl.com/statepolicies). This database has
been updated throughout the pandemic. When we began our paper it
didn't even include information on mask mandates. When information on
mask mandates were added and we began analyzing them, mask mandates
for all people were quite rare, so we chose to focus on the more
common policy of mask mandates for businesses. However, by the time we
were finalizing the published version of our paper, public mask
mandates were common enough that we could have analyzed them
separately. Let's do that now.

When analyzing the impact of state-level policies it is very important
to be aware of the range of policy variation observed in the data. For
example, all states closed schools at nearly the same time, so it is
impossible to reliably estimate the effect of school closures from
variation across states. In the case of public versus business mask
mandates, it is essential to realize that (except in Delaware for 3
days), states with a public mask mandate always also had a business
mask mandate. The set of observed mask policies across states are (i)
no mask mandate (ii) mask mandate only for businesses, and (iii) both
business and public mask mandate.

You can estimate a regression with separate indicators for business
and public mask mandates, but the interpretation of the coefficients
will not be straightforward. The coefficient on public mask mandate is
not the effect of a public mask mandate by itself. Estimating such an
effect is impossible because we never observe only a public mask
mandate. Instead, the coefficient on public mask mandate is the added
effect of a public mask mandate on top of a business mask mandate. If
you want to compare states with public mask mandates (which always
also have a business mask mandate) to states with no mask mandates,
then you should look at the sum of the coefficients on business mask
mandates and public mask mandates. Alternatively, you can estimate the
model with the mask policies encoded to represent the the mask mandate
combinations that we observe. This is what is shown below.

```{r setup, cache=FALSE, warning=FALSE, include=FALSE}
library(conflicted)
unloadNamespace("dotwhisker")
unloadNamespace("dplyr") 
library(lfe)
library(stargazer)
library(knitr)
library(plm)
library(latex2exp)
library(ggplot2)
library(ggthemes)
library(estimatr)
library(gridExtra)
library(grid)
library(mvtnorm)
library(kableExtra)
library(data.table)
library(reshape2)
library(AER)
library(hdm)
library(randomForest)
library(glmnet)

sgtype <- opts_knit$get("rmarkdown.pandoc.to")
sgstyle <- 'default'
colors <-  scale_color_solarized
colors_fill <- scale_fill_solarized
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]

# Set time lag:
timing <- 0 # 0 or 1
if (timing==0){
  L.c <- 14
  L.d <- 21
} else if (timing==1) {
  L.c <- 7
  L.d <- 24
}

source(paste(rootdir,"cases_and_policies/R/exploratory.R", sep="/"))
source(paste(rootdir,"cases_and_policies/R/utils.R",sep="/"))
source(paste(rootdir,"cases_and_policies/R/dataprep.R",sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/varlabels.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/regprep.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/generatetables.R", sep="/"))
source(paste(rootdir,"cases_and_policies/rmd/bootstrap_felm.R", sep="/"))
```

```{r caseregs, cache=TRUE, warning=FALSE, results='hide'}

infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
tvars <- "dlogtests" 
pols  <-  c("pmaskbusonly","pmaskall","pk12","pshelter","pindex") 
L <- L.c
sdf <- subset(df, as.vector(df$date)>=as.Date("2020-03-07"))
regs <- mainregressions(sdf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
summary(regs$piy[[3]]$reg)

```

```{r casetables, cache=FALSE, warning=FALSE, results='asis'}
showhtmltables(NULL, NULL, regs$piy)
```

```{r deathregs, cache=TRUE, warning=FALSE, results='hide'}
df$zero <- 0
infovarsd <- list(c("dlogdd", "logdd"),
                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
iv <- list("0","0")
L <- L.d
sdf <- subset(df, df$date>=as.Date("2020-03-07")) #+L)
dregs <- mainregressions(sdf,
                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)
```

```{r deathtables, cache=FALSE, warning=FALSE, results='asis'}
showhtmltables(NULL, NULL, dregs$piy)
```

# Revised Data

```{r dataupdate, cache=FALSE, warning=FALSE, results='hide'}
rootdir <- system("git rev-parse --show-toplevel", intern=TRUE)[1]
datafile <- paste(rootdir,"cases_and_policies/data/covidstates-updated.Rda", sep="/")
cat("Loading data last updated on ", as.character(file.mtime(datafile)), "\n")
cat("Run ", paste(rootdir,"cases_and_policies/R/cases_and_policies_updated.R", sep="/"), " to update data.\n")
base::load(datafile)

source(paste(rootdir,"cases_and_policies/rmd/regprep-updated.R",sep="/"))
csu <- regprep(covidstates)

udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=max(sdf$date)) )
```

Another important criticism of Lemoine is that our results change
substantially when using revised data. Here, we compare the original
with updated data. The changes in cases and deaths are relatively
small. However, the change in test counts are substantial. There is
also some change in policies.

```{r datarevision, cache=TRUE}
mdf <- merge(udf, sdf, by=c("state","fips","date"))

fig <- ggplot(mdf, aes(x=date, y=cases.x - cases.y, colour=state, label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in cumulative cases") + ggtitle("Change in cumulative cases") + figtheme
fig

fig <- ggplot(mdf, aes(x=date, y=dcases.x - dcases.y, colour=state, label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in weekly cases") + ggtitle("Change in weekly cases") + figtheme
fig

fig <- ggplot(mdf, aes(x=date, y=dlogdc.x-dlogdc.y, colour=state, , label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in ΔlogΔc") + ggtitle("Change in case growth") +
  figtheme
fig

fig <- ggplot(mdf, aes(x=date, y=dlogdd.x-dlogdd.y, colour=state, , label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in ΔlogΔc") + ggtitle("Change in death growth")
  figtheme
fig

fig <- ggplot(mdf, aes(x=date, y=dlogtests.x-dlogtests.y, colour=state, label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in dlogtests") + ggtitle("Change in dlogtests") + figtheme
fig

fig <- ggplot(mdf, aes(x=date, y=pmaskbusonly.x-pmaskbusonly.y, colour=state, label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in pmaskbusonly") + ggtitle("Change in pmaskbusonly") + figtheme
fig

fig <- ggplot(mdf, aes(x=date, y=pmaskall.x-pmaskall.y, colour=state, label=state)) +
  geom_text(show.legend=FALSE) +
  ylab("Change in pmaskall") + ggtitle("Change in pmaskall") + figtheme
fig
```

If we use the updated data with the same time span and regression
specification as the original data, some results do change. The
coefficients on masks are often smaller.

```{r caseregsu,cache=FALSE, warning=FALSE, results='hide'}
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants

infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
udf$testmiss <- is.na(udf$dlogtests)
udf$dlogtests_nomiss <- udf$dlogtests
udf$dlogtests_nomiss[is.na(udf$dlogtests_nomiss)] <- 0
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall","pk12","pshelter","pindex")

L <- L.c
regs <- mainregressions(udf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
summary(regs$piy[[3]]$reg)
```

```{r casetablesu, cache=FALSE, warning=FALSE, results='asis'}
showhtmltables(NULL, NULL, regs$piy)
```

```{r deathregsu, cache=FALSE, warning=FALSE, results='hide'}
udf$zero <- 0
infovarsd <- list(c("dlogdd", "logdd"),
                  c("dlogdd", "logdd","dlogdd.national", "logdd.national"))
iv <- list("0","0")
L <- L.d
dregs <- mainregressions(udf,
                        "dlogdd", pols, bvars, infovarsd, "zero", xlist, ilist, iv, L=L)
```

```{r deathtablesu, cache=FALSE, warning=FALSE, results='asis'}
showhtmltables(NULL, NULL, dregs$piy)
```

## Influence

What caused the results to change? One diagnostic is to look at the
influence of each observation on the coefficients on the mask
policies. Here we plot $$ \frac{-1}{1 - x_{it}' ({X}'{X})^{-1} x_{it}}
({X}'{X})^{-1} {pmask}_{it} \hat{\epsilon}_{it} $$ which is how much
the coefficient would change if that observation were deleted. A state
with positive influence values means that state contributes to the
estimate being negative, and vice versa.

```{r influence, warning=FALSE, cache=TRUE}
calculateinfluence  <- function(mod, data) {
  r <- felm(mod$formula, data=data, keepX=TRUE)

  cols <- !is.nan(r$coefficients)
  X  <- r$X[,cols]
  Q <- solve(t(X) %*% X)

  influ <- 0*r$X
  for (i in 1:nrow(r$X)) {
    influ[i,cols]  <- -1/(1- t(X[i,]) %*% Q %*% X[i,])[1] * Q %*% X[i,] * r$residuals[i]
  }

  data$idx <- as.numeric(1:nrow(data))
  ifmla <- as.character(mod$formula)
  ifmla[2] <- "idx"
  ifmla <- formula(paste(ifmla[2], ifmla[1], ifmla[3], collapse=" "))
  inc <- felm(ifmla, data=data)$response

  rownames(influ) <- inc
  return(influ)
}

plotinfluence  <- function(mod, data, variables) {
  influ <- calculateinfluence(mod, data)
  tmp <- sdf
  tmp$infl <- NA
  tmp[as.integer(rownames(influ)),'infl'] <- influ[,3]
  return(lapply(variables, function(v) {
    tmp[as.integer(rownames(influ)),'infl'] <- influ[,v]
    fig <- ggplot(data=tmp, aes(x=date, y=infl, colour=state,
                                label=state)) +
      geom_text(show.legend=FALSE) + ylab(v)  + figtheme 
    return(fig)
  }))
}

figs <- plotinfluence(regs$piy[[1]]$reg, sdf,
                      c("lag(pmaskbusonly, 14)","lag(pmaskall, 14)"))
newfigs <- plotinfluence(regs$piy[[1]]$reg, udf,
                         c("lag(pmaskbusonly, 14)","lag(pmaskall, 14)"))
i  <- 1
yl <- c(-0.005, 0.005)
grid.arrange(figs[[i]] + ggtitle("Original Data") + ylim(yl),
            newfigs[[i]]+ ggtitle("Updated Data") + ylim(yl), ncol=1)

yl <- c(-0.005, 0.015)
i  <- 2
grid.arrange(figs[[i]] +  ggtitle("Original Data") + ylim(yl),
             newfigs[[i]]+ ggtitle("Updated Data") + ylim(yl), ncol=1)

for (i in 1:2) {
  pdf(sprintf("%s/tex/tables_and_figures/influence_original_%s.pdf",rootdir, i), width=4, height=3.5)
  print(figs[[i]] + ggtitle("Original Data") + ylim(yl))
  dev.off()
  pdf(sprintf("%s/tex/tables_and_figures/influence_new_%s.pdf",rootdir, i), width=4, height=3.5) 
  print(newfigs[[i]]+ ggtitle("Updated Data") + ylim(yl))
  dev.off()
}

```

Hawaii is a notable high influence state. Comparing the old and new
data, case counts are identical in Hawaii. Test numbers change, and
most importantly, mask mandate dates change. The old data has Hawaii
beginning both business and public mask mandates on April
16, 2020. The new data lists business mask mandates beginning on June
11, 2020 and public mandates on November 16, 2020. In reality, the
Hawaii [governor proclamation on April 16,
2020](https://governor.hawaii.gov/wp-content/uploads/2020/04/2004088-ATG_Fifth-Supplementary-Proclamation-for-COVID-19-distribution-signed.pdf)
*recommended* face masks in public and mandated face masks for
employess and customers in essential businesses. The same requirements
are repeated in [a proclamation on June 10,
2020.](https://governor.hawaii.gov/wp-content/uploads/2020/06/2006097A-ATG_Ninth-Supplementary-Proclamation-COVID-19-distribution-signed.pdf). A
[proclamation on November 16](https://governor.hawaii.gov/wp-content/uploads/2020/11/2011051-ATG_Fifteenth-Proclamation-Related-to-the-COVID-19-Emergency-distribution-signed.pdf) extends the mask mandate to all persons in public. Thus, it appears neither the old or new data is entirely correct. The original coding of the data is understandable --- the order on April 16, 2020 is stronger than requiring masks for employees in businesses, since it includes customers too. The updated data appears to just be
a mistake.

We also check mask policies for the following states that have high influence on mask coefficients: Vermont, Connecticut, Montana, North Dakota, and West Virginia. Then, we find that  business mask mandates for North Dakota were announced on April 29 and then became effective on May 1 while the revised data has business mask mandate in North Dakota beginning on April 28 (https://www.governor.nd.gov/executive-orders)(https://www.governor.nd.gov/sites/www/files/documents/executive-orders/Executive%20Order%202020-06.4.pdf).   In the revised data,  West Virginia starts its universal public mask mandates on July 7 of 2021 but this public mask mandates only covers when social distancing is not possible. We keep West Virginia's starting date for public mask mandates as it is.

For Vermont, Connecticut, and Montana,   we don't find any descrepancy between the starting dates for business or public mask mandates in the revised data and those we find in state government's website.

Let's correct the updated data to have Hawaii with a
business mask mandate beginning on April 16, 2020 and to have that of North Dakota beginning on May 1. Since Hawaii's
mandates includes customers, one could argue that it should be
considered to be pmaskall as in the old data, but we will not make
this change.

```{r correcthawaii, warning=FALSE, results='asis'}
covidstates[covidstates$state=="Hawaii","Business.face.mask.mandate.start"] <-
  as.Date("2020-04-16")
covidstates[covidstates$state=="North Dakota","Business.face.mask.mandate.start"] <-
  as.Date("2020-05-01")
csu <- regprep(covidstates)

udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=max(sdf$date)) )
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants

infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
udf$testmiss <- is.na(udf$dlogtests)
udf$dlogtests_nomiss <- udf$dlogtests
udf$dlogtests_nomiss[is.na(udf$dlogtests_nomiss)] <- 0
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall",
            "pk12","pshelter","pindex") 
L <- L.c
regs <- mainregressions(udf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
```

Once again, the effect of mask mandates are substantial and
statistically significant. Unlike with the original data, a public
mask mandate does not have a larger point estimate than a business mask
mandate. However, in either case, the confidence intervals overlap
substantially.

## Dropping Hawaii

Given the concerns about the large influence of Hawaii, one could
consider dropping it from the sample. Choosing to drop observations
based on their influence in a regression is a suspect procedure. One
could argue that Hawaii is unusual due to being an island state far
from all others, so there is some reason to treat it
differently.

### Original Data
```{r noHIold, warning=FALSE, results='asis'}
regs <- mainregressions(subset(sdf,state!="Hawaii"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
```

### New Data
```{r noHInew, warning=FALSE, results='asis'}
regs <- mainregressions(subset(udf,state!="Hawaii"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
```

## Dropping Vermont, Connecticut, Montana, North Dakota, and West Virginia, one by one

We may also check how dropping other influential states (Vermont, Connecticut, Montana, North Dakota, and West Virginia) one by one changes the result using the updated data.

### No VT
```{r noVT, warning=FALSE, results='asis'}
regs <- mainregressions(subset(udf,state!="Vermont"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-revised-noVT")
```

### No CT
```{r noCT, warning=FALSE, results='asis'}
regs <- mainregressions(subset(udf,state!="Connecticut"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-revised-noCT")
```

### No MT
```{r noMT, warning=FALSE, results='asis'}
regs <- mainregressions(subset(udf,state!="Montana"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-revised-noMT")
```

### No ND
```{r noND, warning=FALSE, results='asis'}
regs <- mainregressions(subset(udf,state!="North Dakota"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-revised-noND")
```

### No WV
```{r noWV, warning=FALSE, results='asis'}

regs <- mainregressions(subset(udf,state!="West Virginia"),
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-revised-noWV")
```

Overall, the results do not drastically change as we drop these influential states one by one.

# Expand Sample Period

We fix the sampel period of the data set to the beginning of June, 2020, when we work on the revision of our paper for publication because constantly updating the data makes the revision process very difficult. Now we have access to a longer sample period and we can examine how changing the sample periods results in different estimates. Extending sample periods has an advantage of identifying the coefficient of universal mask mandates variable because many states implemented universal mask mandates after June.

Here, we consider four different end dates for the sample: July 1, August 1, September 1, and October 1 of 2020. Note that the data set does not contain the data on school re-opening dates probably because the school opening dates vary across school districts within states and their modes of school openings (e.g., in-person vs. online) differ. Also, there is almost no change in recorded policies after September 1.
For this reason, we believe that the data set up to August 1 before none of school districts has opened their schools may be a good data set to look at.

## Sample Period up to July 1, 2020
```{r longsample_July, warning=FALSE, results='asis'}
udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=as.Date("2020-07-01")
                          ))
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall","pk12","pshelter","pindex")

L <- L.c
regs <- mainregressions(udf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy) 
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-long-july")

```

## Sample Period up to August 1, 2020
```{r longsample_Aug, warning=FALSE, results='asis'}
udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=as.Date("2020-08-01")
                          ))
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0") 
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall","pk12","pshelter","pindex")

L <- L.c
regs <- mainregressions(udf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-long-aug")

```

## Sample Period up to Sept 1, 2020
```{r longsample_Sept, warning=FALSE, results='asis'}
udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=as.Date("2020-09-01")
                          ))
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall","pk12","pshelter","pindex")

L <- L.c
regs <- mainregressions(udf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-long-sept")

```

## Sample Period up to October 1, 2020

```{r longsample_Oct, warning=FALSE, results='asis'}
udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=as.Date("2020-10-01")
                          ))
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants
infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall","pk12","pshelter","pindex")

L <- L.c
regs <- mainregressions(udf,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
showhtmltables(NULL, NULL, regs$piy)
#savetextables(NULL, NULL,  regs$piy, prefix="cases-maskall-long-oct")

```

Across different sample periods, the estimates consistently indicate the strong association of both business and universal mask mandates with the subsequent case growth.

# Updated Plots

```{r npolicies, warning=FALSE, cache=TRUE}
udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=as.Date("2020-12-11")
                          ))
udf <- udf[order(udf$state,udf$date),]
# pdates <- c("Date.closed.K.12.schools",
#             "Stay.at.home..shelter.in.place",
#             "Closed.movie.theaters",
#             "Closed.restaurants.except.take.out",
#             "Closed.non.essential.businesses",
#             "Mandate.face.mask.use.by.all.individuals.in.public.spaces",
#             "Mandate.face.mask.use.by.employees.in.public.facing.businesses",
#             "State.of.emergency")
# enddates <- c(NA,
#               "End.relax.stay.at.home.shelter.in.place",
#               "Reopened.movie.theaters",
#               "Reopen.restaurants",
#               "Began.to.reopen.businesses.statewide",
#               NA,
#               NA,
#               NA)
pdates <- c("Closed.K.12.public.schools",
            "Stay.at.home.shelter.in.place",
            "Closed.movie.theaters",
            "Closed.restaurants",
            "Closed.other.non.essential.businesses",
            "Public.face.mask.mandate.start",
            "Business.face.mask.mandate.start",
                "State.of.emergency.issued")
enddates <- c(NA,
              "End.stay.at.home.shelter.in.place",
              "Reopened.movie.theaters",
                  "Reopened.restaurants", 
              "Began.to.reopen.businesses.statewide",
              "Face.mask.mandate.end",
              "Face.mask.mandate.end",
              "State.of.emergency.lifted")

pvars <- c("pk12", "pshelter", "pmovie", "prestaurant", "pnonessential",
           "pmaskall","pmaskbus", "psoe")

for(i in 1:length(pdates)) udf[,pvars[i]] <-
                             smoothedpolicy(udf,pdates[i],enddatevar=enddates[i],type="ma",bw=0)

pols <- c("pmaskbus","pmaskall","pshelter", "pmovie",
          "prestaurant", "pnonessential","pk12", "psoe")

startday <- as.Date("2020-03-01")
endday <- as.Date("2020-10-01")

ppplot <- function(pol, udf) {
  udf$y <- udf[,pol]
  adf <- aggregate(y ~ date, data=udf, function(x) mean(x, na.rm=TRUE))
  fig <- ggplot(adf, aes(x=date, y=y)) + geom_line(size=1.2) +
    plottheme + ylab("Portion of States") +
    colors() + ggtitle(varlabels[varlabels$Variable==pol,"Label"]) +
    theme(legend.position="none") + 
    xlim(startday, endday) + ylim(0,1)
  return(fig)
}
figs <- lapply(pols,function(p) ppplot(p, udf))

for (i in 1:length(pols)) {
  pdf(sprintf("%s/tex/tables_and_figures/%s_p_update.pdf",rootdir, pols[i]), width=4, height=2)
  print(figs[[i]])
  dev.off()
}
grid.arrange(grobs=figs[1:7], nrow=4, ncol=2)
```

# Permutation Test

We also implemente our own placebo test to examine if we can find spurious effects. Specifically, we generated 500 samples, where each sample is identical to the original sample except that state-level mask variables (i.e., ``masks in public spaces'' and ``masks for employees only") are replaced with those of other states that are randoly selected through permutation. Because each state's mask variables are replaced with randomly selected other state's mask variables, we expect that the estimated coefficients of mask variables would be centered around zero if there is no spurious effects arising from, for example, aggregate time effect.

## Using the sample from March 7 to June 3, 2020

```{r correcthawaii-test, warning=FALSE, results='asis'}
covidstates[covidstates$state=="Hawaii","Business.face.mask.mandate.start"] <-
  as.Date("2020-04-16")
covidstates[covidstates$state=="North Dakota","Business.face.mask.mandate.start"] <-
  as.Date("2020-05-01")
csu <- regprep(covidstates)

udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=max(sdf$date)) )
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants

infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
udf$testmiss <- is.na(udf$dlogtests)
udf$dlogtests_nomiss <- udf$dlogtests
udf$dlogtests_nomiss[is.na(udf$dlogtests_nomiss)] <- 0
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall",
            "pk12","pshelter","pindex")

iter = 500
maskall_vals <- vector(mode = "list", length = iter)
busonly_vals <- vector(mode = "list", length = iter)
pols  <-  c("pmaskbusonly","pmaskall",
            "pk12","pshelter","pindex")
L <- L.c

system.time(
for(i in 1:iter){
  # permutation of states and policy dates
  fipz <- unique(udf$fips)
  vperm <- sample(fipz)
  sudf <- udf
  sudf$fips_perm <- vperm[match(sudf$fips, fipz)]
  temp = subset(sudf, select = c('fips_perm','pmaskbusonly','pmaskall','date'))
  colnames(temp) <- c('fips', 'pmaskbusonly_perm', 'pmaskall_perm', 'date')
  sudf <- merge(sudf,temp,by = c('fips','date'), all = TRUE, sort = FALSE)
  udf_b <- udf
  udf_b$pmaskbusonly <- sudf$pmaskbusonly_perm
  udf_b$pmaskall <- sudf$pmaskall_perm
  regs_perm <- mainregressions(udf_b,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
  busonly_vals[i] <- summary(regs_perm$piy[[3]]$reg)$coefficients[2]
  maskall_vals[i] <- summary(regs_perm$piy[[3]]$reg)$coefficients[3]
})

meltedbus = reshape2::melt(busonly_vals)
meltedall = reshape2::melt(maskall_vals)
meltedbus$Coefficients <- "pmaskbusonly"
meltedall$Coefficients <- "pmaskall"
sim_coef<-rbind(meltedbus[,c(1,3)],meltedall[,c(1,3)])
sim_coef <- data.frame(sim_coef)

fig <- ggplot(sim_coef,aes(x=Coefficients, y=value)) +
  geom_boxplot() + scale_y_continuous(breaks=seq(-0.16,0.16,0.04))
fig

pdf(sprintf("%s/tex/tables_and_figures/boxplot_sim.pdf",rootdir, i), width=4, height=3)
print(fig)
dev.off()

```

## Using the extended sample from March 7 to August 1, 2020

```{r correcthawaii-test2, warning=FALSE, results='asis'}
covidstates[covidstates$state=="Hawaii","Business.face.mask.mandate.start"] <-
  as.Date("2020-04-16")
covidstates[covidstates$state=="North Dakota","Business.face.mask.mandate.start"] <-
  as.Date("2020-05-01")
csu <- regprep(covidstates)
udf <- subset(csu, as.vector(as.vector(csu$date)>=min(sdf$date))
              & as.vector(as.vector(csu$date) <=as.Date("2020-08-01")
                          ))
udf$pmovie <- udf$pmovie.theaters
udf$prestaurant <- udf$prestaurants

infovars <- list(c("dlogdc", "logdc"),
                 c("dlogdc", "logdc","dlogdc.national", "logdc.national"))
xlist <- list("","")
interactions <- list("month", statevars)
ilist <- list(interactions, interactions)
iv <- list("0","0")
udf$testmiss <- is.na(udf$dlogtests)
udf$dlogtests_nomiss <- udf$dlogtests
udf$dlogtests_nomiss[is.na(udf$dlogtests_nomiss)] <- 0
tvars <- "dlogtests"

pols  <-  c("pmaskbusonly","pmaskall",
            "pk12","pshelter","pindex")

iter = 500
maskall_vals <- vector(mode = "list", length = iter)
busonly_vals <- vector(mode = "list", length = iter)
pols  <-  c("pmaskbusonly","pmaskall",
            "pk12","pshelter","pindex")
L <- L.c

system.time(
for(i in 1:iter){
  # permutation of states and policy dates
  fipz <- unique(udf$fips)
  vperm <- sample(fipz)
  sudf <- udf
  sudf$fips_perm <- vperm[match(sudf$fips, fipz)]
  temp = subset(sudf, select = c('fips_perm','pmaskbusonly','pmaskall','date'))
  colnames(temp) <- c('fips', 'pmaskbusonly_perm', 'pmaskall_perm', 'date')
  sudf <- merge(sudf,temp,by = c('fips','date'), all = TRUE, sort = FALSE)
  udf_b <- udf
  udf_b$pmaskbusonly <- sudf$pmaskbusonly_perm
  udf_b$pmaskall <- sudf$pmaskall_perm
  regs_perm <- mainregressions(udf_b,
                        "dlogdc", pols, bvars, infovars, tvars, xlist, ilist, iv, L=L)
  busonly_vals[i] <- summary(regs_perm$piy[[3]]$reg)$coefficients[2]
  maskall_vals[i] <- summary(regs_perm$piy[[3]]$reg)$coefficients[3]
})

meltedbus = reshape2::melt(busonly_vals)
meltedall = reshape2::melt(maskall_vals)
meltedbus$Coefficients <- "pmaskbusonly"
meltedall$Coefficients <- "pmaskall"
sim_coef<-rbind(meltedbus[,c(1,3)],meltedall[,c(1,3)])
sim_coef <- data.frame(sim_coef)

fig <- ggplot(sim_coef,aes(x=Coefficients, y=value)) +
  geom_boxplot() + scale_y_continuous(breaks=seq(-0.18,0.18,0.04))
fig

pdf(sprintf("%s/tex/tables_and_figures/boxplot_sim2.pdf",rootdir, i), width=4, height=3)
print(fig)
dev.off()

```

When we use the sample from March 7 to June 3, 2020, the estimates are roughly centered around -0.02 for both mask variables while the inter-quartile ranges for  ``masks in public spaces'' and ``masks for employees only" are around $[-0.05,0.02]$. When we use the extended sample from March 7 to August 1, 2020, the estimates are  centered around $-0.01$ while the inter-quartile ranges are narrower with  $[-0.03,0.02]$. Overall, the result of our placebo test indicates that there won't be any systematic spurious effects we may find from our regression analysis, especially if we use the data set with the extended sample period to August 1, 2020.
